### 2021-06-04 survey/code/deconv_corr.Rmd
Deconv correlation experiments

---
title: Correlation analysis to inferred infection rates (deconvolved case rates)
---

```{r, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, autodep = TRUE, cache.comments = TRUE, 
                      message = FALSE, warning = FALSE)
```

```{r, messages = FALSE}
library(covidcast)
library(dplyr)
library(ggplot2)

cases <- covidcast_signal(data_source = "usa-facts",
                           signal = "confirmed_7dav_incidence_prop",
                           start_day = "2020-04-15",
                           end_day = "2021-04-15", 
                           geo_type = "state")
survey <- covidcast_signal(data_source = "fb-survey",
                           signal = "smoothed_whh_cmnty_cli",
                           start_day = "2020-04-15",
                           end_day = "2021-04-15", 
                           geo_type = "state")

deconv <- read.csv("deconv_data.csv")
deconv <- as.covidcast_signal(deconv, signal = deconv$signal[1])
deconv$time_value <- as.Date(deconv$time_value)
```

## Geo-wise correlations (sliced by time)

```{r}
# Compute correlation per time, over all counties
cor_geowise1 <- covidcast_cor(deconv, survey, by = "time_value")
cor_geowise2 <- covidcast_cor(cases, survey, by = "time_value")
cor_geowise <- rbind(cor_geowise1, cor_geowise2)
cor_geowise$signal = as.factor(
  c(rep("deconv-survey", nrow(cor_geowise1)),
    rep("cases-survey", nrow(cor_geowise2))))

ggplot(cor_geowise, aes(x = time_value, y = value)) +
  geom_line(aes(color = signal)) +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
               date_minor_breaks = "1 month") + 
  labs(x = "Date", y = "Correlation") + theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank())
#ggsave("../fig/deconv_geowise_corr.pdf", width = 5, height = 3.5)
```

## Time-wise correlations (sliced by location)

```{r}
# Compute correlation per location, over all time
cor_timewise1 <- covidcast_cor(deconv, survey, by = "geo_value")
cor_timewise2 <- covidcast_cor(cases, survey, by = "geo_value")
cor_timewise <- rbind(cor_timewise1, cor_timewise2)
cor_timewise$signal = as.factor(
  c(rep("deconv-survey", nrow(cor_timewise1)),
    rep("cases-survey", nrow(cor_timewise2))))

ggplot(cor_timewise, aes(value)) +
  geom_density(aes(color = signal, fill = signal), alpha = 0.4) +
  labs(x = "Correlation", y = "Density") + theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank())
#ggsave("../fig/deconv_timewise_corr.pdf", width = 5, height = 3.5)
```

## More systematic lag analysis

```{r}
# Loop over values for dt, and compute correlations per state
dt_vec <- (-20):10
cor_timewise_list1 <- vector("list", length(dt_vec))
for (i in 1:length(dt_vec)) {
  cor_timewise_list1[[i]] <- covidcast_cor(deconv, survey, dt_x = dt_vec[i],
                               by = "geo_value")
  cor_timewise_list1[[i]]$dt <- dt_vec[i]
}

# Stack into one big data frame, and then plot the median correlation by dt
cor_timewise_dt1 <- do.call(rbind, cor_timewise_list1)
cor_timewise_dt1 %>%
  group_by(dt) %>%
  summarize(value = mean(value, na.rm = TRUE), .groups = "drop_last") %>%
  ggplot(aes(x = dt, y = value)) + geom_line() + geom_point() +
  labs(x = "dt", y = "Correlation") + theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank())
```

```{r}
# Loop over values for dt, and compute correlations per state
cor_timewise_list2 <- vector("list", length(dt_vec))
for (i in 1:length(dt_vec)) {
  cor_timewise_list2[[i]] <- covidcast_cor(cases, survey, dt_x = dt_vec[i],
                               by = "geo_value")
  cor_timewise_list2[[i]]$dt <- dt_vec[i]
}

# Stack into one big data frame, and then plot the median correlation by dt
cor_timewise_dt2 <- do.call(rbind, cor_timewise_list2)
cor_timewise_dt2 %>%
  group_by(dt) %>%
  summarize(value = mean(value, na.rm = TRUE), .groups = "drop_last") %>%
  ggplot(aes(x = dt, y = value)) + geom_line() + geom_point() +
  labs(x = "dt", y = "Correlation") + theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank())
```

### 2021-06-05 survey/code/deconv_corr.Rmd
Updated deconv experiments

@@
 
+start_day <- "2020-04-15"
+end_day <- "2021-04-15"
+geo_type <- "state"
+cli <- covidcast_signal(data_source = "fb-survey",
+                        signal = "smoothed_wcli",
+                        start_day, end_day, geo_type)
+cmt <- covidcast_signal(data_source = "fb-survey",
+                        signal = "smoothed_whh_cmnty_cli",
+                        start_day, end_day, geo_type)
 cases <- covidcast_signal(data_source = "usa-facts",
                           signal = "confirmed_7dav_incidence_prop",
-                           start_day = "2020-04-15",
-                           end_day = "2021-04-15", 
-                           geo_type = "state")
-survey <- covidcast_signal(data_source = "fb-survey",
-                           signal = "smoothed_whh_cmnty_cli",
-                           start_day = "2020-04-15",
-                           end_day = "2021-04-15", 
-                           geo_type = "state")
+                        start_day, end_day, geo_type)
 
@@
 # Compute correlation per time, over all counties
-cor_geowise1 <- covidcast_cor(deconv, survey, by = "time_value")
-cor_geowise2 <- covidcast_cor(cases, survey, by = "time_value")
-cor_geowise <- rbind(cor_geowise1, cor_geowise2)
-cor_geowise$signal = as.factor(
-  c(rep("deconv-survey", nrow(cor_geowise1)),
-    rep("cases-survey", nrow(cor_geowise2))))
+by = "time_value"
+cor_geowise <- rbind(
+  data.frame(covidcast_cor(cli, deconv, by = by), signal = "cli-deconv"),
+  data.frame(covidcast_cor(cmt, deconv, by = by), signal = "cmt-deconv"),
+  data.frame(covidcast_cor(cli, cases, by = by), signal = "cli-cases"),
+  data.frame(covidcast_cor(cmt, cases, by = by), signal = "cmt-cases"))
 
@@
 # Compute correlation per location, over all time
-cor_timewise1 <- covidcast_cor(deconv, survey, by = "geo_value")
-cor_timewise2 <- covidcast_cor(cases, survey, by = "geo_value")
-cor_timewise <- rbind(cor_timewise1, cor_timewise2)
-cor_timewise$signal = as.factor(
-  c(rep("deconv-survey", nrow(cor_timewise1)),
-    rep("cases-survey", nrow(cor_timewise2))))
+by <- "geo_value"
+cor_timewise <- rbind(
+  data.frame(covidcast_cor(cli, deconv, by = by), signal = "cli-deconv"),
+  data.frame(covidcast_cor(cmt, deconv, by = by), signal = "cmt-deconv"),
+  data.frame(covidcast_cor(cli, cases, by = by), signal = "cli-cases"),
+  data.frame(covidcast_cor(cmt, cases, by = by), signal = "cmt-cases"))
 
@@
 # Loop over values for dt, and compute correlations per state
-dt_vec <- (-20):10
-cor_timewise_list1 <- vector("list", length(dt_vec))
+dt_vec <- (-20):20
+by <- "geo_value"
+cor_timewise_list <- vector("list", length(dt_vec))
 for (i in 1:length(dt_vec)) {
-  cor_timewise_list1[[i]] <- covidcast_cor(deconv, survey, dt_x = dt_vec[i],
-                               by = "geo_value")
-  cor_timewise_list1[[i]]$dt <- dt_vec[i]
+  cor_timewise_list[[i]] <- rbind(
+    data.frame(covidcast_cor(cli, deconv, dt_x = dt_vec[i], by = by), 
+               dt = dt_vec[i], signal = "cli-deconv"),
+    data.frame(covidcast_cor(cmt, deconv, dt_x = dt_vec[i], by = by), 
+               dt = dt_vec[i], signal = "cmt-deconv"),
+    data.frame(covidcast_cor(cli, cases, dt_x = dt_vec[i], by = by), 
+               dt = dt_vec[i], signal = "cli-cases"),
+    data.frame(covidcast_cor(cmt, cases, dt_x = dt_vec[i], by = by), 
+               dt = dt_vec[i], signal = "cmt-cases"),
+    data.frame(covidcast_cor(cli, cmt, dt_x = dt_vec[i], by = by), 
+               dt = dt_vec[i], signal = "cli-cmt"))
 }
 
-# Stack into one big data frame, and then plot the median correlation by dt
-cor_timewise_dt1 <- do.call(rbind, cor_timewise_list1)
-cor_timewise_dt1 %>%
-  group_by(dt) %>%
+# Stack into one big data frame, and then plot the mean correlation by dt
+cor_timewise_dt <- do.call(rbind, cor_timewise_list)
+cor_timewise_dt %>%
+  group_by(dt, signal) %>%
   summarize(value = mean(value, na.rm = TRUE), .groups = "drop_last") %>%
-  ggplot(aes(x = dt, y = value)) + geom_line() + geom_point() +
+  ggplot(aes(x = dt, y = value)) + 
+  geom_line(aes(color = signal)) + geom_point(aes(color = signal)) +
+  labs(x = "dt", y = "Correlation") + theme_bw() +
+  theme(legend.position = "bottom", legend.title = element_blank())
+
+cor_timewise_dt %>%
+  group_by(dt, signal) %>%
+  summarize(value = median(value, na.rm = TRUE), .groups = "drop_last") %>%
+  ggplot(aes(x = dt, y = value)) + 
+  geom_line(aes(color = signal)) + geom_point(aes(color = signal)) +
   labs(x = "dt", y = "Correlation") + theme_bw() +
@@
 ```
-
-```{r}
-# Loop over values for dt, and compute correlations per state
-cor_timewise_list2 <- vector("list", length(dt_vec))
-for (i in 1:length(dt_vec)) {
-  cor_timewise_list2[[i]] <- covidcast_cor(cases, survey, dt_x = dt_vec[i],
-                               by = "geo_value")
-  cor_timewise_list2[[i]]$dt <- dt_vec[i]
-}
-
-# Stack into one big data frame, and then plot the median correlation by dt
-cor_timewise_dt2 <- do.call(rbind, cor_timewise_list2)
-cor_timewise_dt2 %>%
-  group_by(dt) %>%
-  summarize(value = mean(value, na.rm = TRUE), .groups = "drop_last") %>%
-  ggplot(aes(x = dt, y = value)) + geom_line() + geom_point() +
-  labs(x = "dt", y = "Correlation") + theme_bw() +
-  theme(legend.position = "bottom", legend.title = element_blank())
-```

### 2021-06-11 forecast/paper/paper.tex
Transfer stuff over
- Mainly transferring stuff over from old GitHub repos
- Light updates to indicators paper, forecast paper intro

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

\input{../../common/prelim.tex}

\title{Can COVID-19 Indicators Improve Forecasts and Hotspot Prediction?}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,c,1]{Author One}
\author[b,1,2]{Author Two} 
\author[a]{Author Three}

\affil[a]{Affiliation One}
\affil[b]{Affiliation Two}
\affil[c]{Affiliation Three}

% Please give the surname of the lead author for the running footer
\leadauthor{Lead author last name}

% Please add a significance statement to explain the relevance of your work
\significancestatement{Authors must submit a 120-word maximum statement about
  the significance of their research paper written at a level understandable to
  an undergraduate educated scientist outside their field of speciality. The
  primary goal of the significance statement is to explain the relevance of the
  work in broad context to a broad readership. The significance statement
  appears in the paper itself and is required for all research papers.}

% Please include corresponding author, author contribution and author
% declaration information 
\authorcontributions{Please provide details of author contributions here.}
\authordeclaration{Please declare any competing interests here.}
\equalauthors{\textsuperscript{1}A.O.(Author One) contributed equally to this
  work with A.T. (Author Two) (remove if not applicable).} 
\correspondingauthor{\textsuperscript{2}To whom correspondence should be
  addressed. E-mail: author.two\@email.com}

% At least three keywords are required at submission. Please provide three to
% five keywords, separated by the pipe symbol. 
\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}

\begin{abstract}
Please provide an abstract of no more than 250 words in a single paragraph.
Abstracts should explain to the general reader the major contributions of the
article. References in the abstract must be cited in full within the abstract
itself and cited in the text. 
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

\dropcap{T}racking and forecasting indicators from public health reporting1
streams---such as confirmed cases and deaths in the COVID-19 pandemic---is
crucial for understanding disease spread, formulating public policy responses,
and anticipating future public health resource needs.  In a companion paper, we
describe our research group's (Delphi's) efforts in curating and maintaining a 
database of real-time indicators that track COVID-19 activity and other relevant
phenomena. The signals (a term we use synonomously with ``indicators'') in this
database are accessible through the COVIDcast API \cite{CovidcastAPI}, with
associated R \cite{CovidcastR} and Python \cite{CovidcastPy} packages for
convenient data fetching and processing tools. In the current paper, we aim to
quantify the additional utility provided by a core set of these indicators for
two fundamental prediction tasks: probabilistic forecasting of COVID-19 case
rates and prediction of future COVID-19 hotspots (where a hotspot is defined by
the event that the relative increase in COVID-19 cases in a given location
exceeds a certain threshold).

At the outset, we should be clear that our intent in this paper is \emph{not} to
provide an authoritative take on cutting-edge COVID-19 forecasting methods.
Instead, our purpose here is to provide a rigorous, quantitative assessment of
the utility that several auxiliary indicators---such as those derived from
internet surveys or medical insurance claims---provide in tasks that involve
predicting future trends in confirmed COVID-19 cases. To assess such utility in
as simple terms as possible, we center our study in the framework of a basic
autoregressive model (in which COVID cases in the near future are predicted from  
a linear combination of COVID cases in the near past), and ask whether the 
inclusion of an auxiliary indicator as an exogenous feature in such a model
improves predictive power. While forecasting carries a rich literature that
offers a wide range of techniques, see e.g., \cite{Hyndman:2018}, we purposely
constrain ourselves to very simple models, avoiding common enhancements such as
order selection, correction of outliers/anomalies in the data, and inclusion of
regularization or nonlinearities. That said, analyses of forecasts submitted to
the COVID-19 Forecast Hub \cite{ForecastHub} by a large community of modelers
have shown that simple, robust models have consistently been among the
best-performing over the pandemic \cite{Cramer:2021}, including time series
models similar to those we consider in what follows.

In our companion paper, we analyze correlations between various indicators and   
COVID case rates. These correlations are natural summaries of the
contemporaneous association between an indicator and COVID cases, but they fall
short of delivering a satisfactory answer to the question that motivates the
current article: is the information contained in an indicator demonstrably
useful for the prediction tasks we care about? For such a question, lagged
correlations (e.g., measuring the correlation between an indicator and COVID
case rates several days in the future) move us in the right direction, but still
do not move us all the way there. The question about \emph{utility for
  prediction} is focused on a much higher standard than simply asking 
about correlations; to be useful in forecasting or hotspot prediction, an 
indicator must provide relevant information that is not otherwise contained in 
past values of the case rate itself. We address this in the most direct way
possible, by inpsecting the difference in predictive performance of simple 
autoregressive models trained with and without access to past values of a 
particular indicator.

There is a second and more subtle issue in evaluating predictive utility that
deserves explicit mention, as it will play a key role in our
analysis. Indicators computed from surveillance streams will often be subject to
latency and/or revision. For example, a signal based on aggregated medical
insurance claims may be available after just a few days, but it can then be
substantially revised over the next several weeks as additional claims are
submitted and processed late. Correlations between such a signal and case rates 
calculated ``after the fact'' (meaning, we look retrospectively and use the
finalized values of this indicator) will not deliver an honest answer to the
question of whether this indicator would have been useful in real time. Instead, 
we build predictive models using only the data that would have been available 
\emph{as of} the prediction date, and compare the ensuing predictions in terms
of accuracy. This leverages our group's {\tt evalcast} R package, which plugs
into the COVIDcast API's data versioning system, and facilitates honest
backtesting of forecasters.

Furthermore, adopting this practically-oriented forecasting
perspective to indicator evaluation requires one to confront
fundamental limitations inherent to public health surveillance
systems.  In particular, many reporting systems are plagued by
latency, with delays in reporting that can sometimes span
multiple weeks.  Latency manifests itself in different ways in different data sources.  Some data sources may only make a signal available one week after the date it describes; other
data sources may provide immediate imperfect initial estimates, which are
gradually ``backfilled'' over a period of days, weeks, or months (in which previous estimates are updated as new
data trickle in).  And few of
these latency patterns are stable or predictable, with holidays,
extreme weather events, and changes in the reporting process all
affecting the dynamics of the delay process.  To fully and accurately
understand an indicator's usefulness for forecasting requires
accounting for the unfortunate reality of data latency.

Examining the importance of additional features for prediction is a core question
in inferential statistics, and there are many ways to examine this question
under a variety of assumptions. For time series, work in
this area can be traced at least to Granger causality~\cite{GrangerCausality} and
related multivariate extensions~\cite{GewekeTimeSeries} under the condition 
that the series are linearly related autoregressions among other assumptions.
These techniques are especially popular in econometrics and neuroscience~\cite{mccracken2007asymptotics,NeuroscienceGranger}.

We choose to take a
predictive angle, which is simple but also fairly robust, compared to various
other classical inferential approaches. In fact, drawing rigorous inference based on
predictions, without (or with lean) assumptions, is an active field of
research~\cite{dai2021significance,fryer2021shapley,zhang2020floodgate,williamson2020unified}.
Nonetheless, the predictive approach directly examines the target of interest: do
these signals help to produce forecasts? Furthermore, directly comparing
forecast errors without strong assumptions through tests of predictive accuracy
are common in epidemiology, econometrics, and other
fields~\cite{DieboldMarianoTest,DieboldRetrospective}.

% - then make the broad point that "feature importance" is a core question in
% statistics, and there are many ways to look at it and lots of literature (cite
% a bunch fo stuff).  We choose to take a predictive angle, which is simple but
% also fairly robust, compared to various other classical inferential approaches
% that make lots of assumptions.  In fact, drawing rigorous inference based on
% predictions, without (or with lean) assumptions, is an active field of
% research.  Cite several papers (Jing's conformal paper that describes LOCO,  
% commentary by Ale, Ryan, Larry in Stat Sci, a bunch of model X stuff including
% Lukas Janson's recent paper, Eugene + Aaditya's recent paper?).  Larry has put
% some references in the bib.  Put more if we can find them, especially on time
% series.

% RJT: 
% What is a leading indicator?
% Conceptual vs empirical.

\section{Methods}

\subsection{Signals and Locations}

The indicators we focus on provide information not generally available from
standard public health reporting. The COVID Tracking and Impact Survey (CTIS),
run by the Delphi group in partnership with Facebook, reaches roughly 50,000
respondents across the United States daily, and provides information on
symptoms, social distancing, and other attitudes and behaviors related to
COVID. The particular indicator we examine, which we refer to as \fb, is the
estimated percentage of people who know someone with COVID-like illness in their
the local community.

In some ways, this can be seen as a
proxy for the strength of social contacts between infected and susceptible
populations, a key feature for disease modeling~\cite{Zhang1481} The Google
indicator, which we refer to as \gs, tracks the local intensity of Google searches related to lack-of-smell
(anosmia) or lack-of-taste (ageusia), both of which have been connected to
COVID19~\cite{KlopfenAnosmia,vaira2020anosmia}. Two
indicators are derived from de-identified insurance claims data from Change Healthcare, a large
healthcare technology company that aggregates data from many providers. The
first signal, which we refer to as \chngcov, is the estimated percentage of outpatient doctor visits with
confirmed COVID-19 diagnosis (potentially collected earlier than the related
report to the local health authority), while the second, which we
refer to as \chngcli, is the estimated percentage of outpatient
visits with COVID-like illness. Finally, \attn{describe \dv~indicator, perhaps best described in conjunction with \chngcov}.  The hope is that each
of these signals serves as a ``leading indicator'' of future COVID cases,
improving the forecastibility of COVID cases and hotspots. For more details on
these and other indicators, see the companion paper.

% RJT: We should refer to the companion (indicators) paper for more details.
% Also, I'm wondering what the point of the table (below) is?  Does it offer new
% info in addition to the paragraph?  If it's just a nice visual cue, then we
% can turn the above paragraph into a bullet point list, one bullet per signal,
% and cut the table. @DM/JB: Let me know what you think.
% DJM: agreed, table gone for now. I'll leave the bullet decision for later.

% \begin{table}
% \begin{tabular}{@{}p{1in}p{1in}p{1in}p{3in}@{}}
%  \toprule
%     Data source &   Signal name &   Abbreviation    &  
%         Description\cite{DelphiAPI}\\
%  \midrule
%    \attn{COVIDcast Survey via Facebook} &   Smoothed CLI-in-community & FBCLI &
%         Estimated percentage of people reporting illness in their local community,
%         including their household   \\
%     Change Healthcare   &   Smoothed adjusted oupatient CLI  &   CHCLI   &
%         Estimated percentage of outpatient doctor visits primarily about
%         COVID-related symptoms, with day-of-week effects removed  \\
%     Change Healthcare   &   Smoothed adjusted oupatient COVID  &   CHCOV &
%         Estimated percentage of outpatient doctor visits with confirmed
%         COVID-19, with day-of-week effects removed  \\
%     Google Symptoms     & Smoothed sum of anosmia and ageusia   &   GSSAA   &
%         The sum of Google search volume for anosmia and ageusia related searches,
%         in an arbitrary units that are normalized for overall search users\\
%  \bottomrule
% \end{tabular}
% \caption{Syndromic signals available in COVIDcast.\label{tab:signals-used}}
% \end{table}

% RJT: We need to explain WHY HRRs.  This is not familiar to basically anybody!
% It's an invention of academics who work in healthcare policy, from Dartmouth.
% We should cite them: https://www.dartmouthatlas.org/faq/.  Here's the
% rationale for us using them: 
% - They are groups of counties driven by the same hospital referral system.
% - They are somewhat homogenous population-wise.  Don't have a long tail of 
% many HRRs with small populations (like counties).  We should check this!  And
% could even the distribution vs the distribution with respect to counties (or
% some other comparison point). 
% - They would be more relevant to **hospitalization forecasting** given their
% definition, but our indicators should provide bigger boosts to case forecasts.
% (Hospitalizations | cases shouldn't have much dependence with indicators.
% Here I'm using this as in X | Y, i.e., X conditional on Y).
% - But to the extent that case projections inform hospitalization projects,
% this is still quite relevant & sensible to use HRRs. 
% @DM/JB: Please take a stab at this.
% DJM: stabbed.

As for geographic resolution, we consider prediction of COVID case rates (and
hotspots) aggregated at the level of an individual hospital referral region
(HRR). HRRs correspond to groups of counties within the same hospital referral
system. The Dartmouth Atlas of Healthcare Policy~\cite{DartmouthHRR}, defines
these 306 regions based on a number of characteristics. They are contiguous
regions containing at least one city where major cardiovascular or neurological
precedures are performed. Each HRR has a minimum population of about 120,000, and
the majority of hospital services for that population are performed by hospitals
within the region. Some HRRs are quite large (Los Angeles has
about 10 million), but generally HRRs are much more homogenous than the
$\sim$3100 counties 
would be but at much higher resolution than states. They would be more relevant for
hospitalization forecasting, given their definition. We have chosen to focus on
forecasting cases at this resolution because the indicators should provide a
bigger boost. Once we observe case information, there should not be much
dependence of future
hospitalizations on the indicators, but predicting cases should provide more timely
warnings of future hospitalization needs as opposed to waiting for the case
numbers.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig/hrr_pop_histogram.pdf}
  \caption{Histogram of population size for HRR. Note that the $x$-axis is on
    the log scale.}
  \label{fig:hrr-histogram}
\end{figure}

\subsection{Vintage Training Data}

% RJT: We need to fill this out.  I'm convinced we need to promote "as of"
% training as our main point of study.  We need to explain why this is so
% important.  And we can and should present results with finalized data in the
% appendix.  I guess we can wait until we get the experiments in to fill out
% this section, because it might influence the prose we use here, or we could
% also start writing this right away. And we can refer to the companion
% (indicators) paper, which also describes the revision tracking feature in the
% API.  @DM/JB: Your call, whether to start writing this or wait.
% DJM: drafted

In this manuscript, all models are estimated using only ``vintage training
data''. This means that for a given forecast date, say November 1st 2020, we
estimate our models using only the data that would have been available to us on
November 1st. Essentially, the training data
is rewound appropriately to dates in the past. It is likely, for example, that
no data would actually exist 
for the forecast date on the forecast date because there is some latency between
the time signals are reported and the dates for which they are reported.
Similarly, any future revisions made to the indicators or case numbers would
not have been present on the forecast date, and so would not be available to the
forecaster. Figure~\ref{fig:vintage} shows three signals for the Los Angeles HRR
with a hypothetical forecast date of November 1. On November 1, no signals are
available for October 31. The cases data is one day latent while the Facebook
survey is 2 days latent. Doctor's visits are 3 days latent while also exhibiting
revisions.  
By training our models in this way, the evaluations are more realistic in the sense
that we are accurately mimicking the difficulty faced by real-time epidemic
forecasters. The COVIDcast API maintains all versions of the data, thereby allowing
these sorts of synthetic forecasting exercises. For more details about data
versioning, see the companion paper.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig/vintage.pdf}
  \caption{Changes in availability and revisions for three signals in the Los
    Angeles HRR. The vertical black line indicates a hypothetical forecast date
    of November 1 while the colored lines get progressively darker as we collect
  later and later vintages of data.}
  \label{fig:vintage}
\end{figure}

\subsection{Analysis Tasks}

%\subsubsection{Forecasting Commonalities} % PNAS wants Title Caps for everything

% RJT: Just a quick few quick comments for now on the text below.  
% Also, shouldn't we
% point out it's relationship to quantile score, and give it's equivalent form
% there?
% DJM: done.

On every day between \attn{July 1 and December 31}, we make two types of
forecasts for each of the 306 hospital referral regions in the United States:
(1) we forecast the seven-day trailing average of incident cases, or "smoothed
incident cases"; and (2) we predict whether the HRR will be a hotspot, a
location with a large increase in cases. For both tasks we try to predict the
outcome 7, 8, 9, up to \attn{21} days into the future. For both tasks, we
estimate a baseline global model across all HRRs using the most recent
observation of the target, as well the observations from 1 and 2 weeks earlier.
We train a separate model for each forecast date and for each forecast horizon,
but we do not assume any HRR-specific effects; that is, a single global model is
used to 
describe all HRRs for each forecast date-horizon pair. This model is
re-estimated for every forecast date and each future target resulting in
\attn{$180\times 15 = 2700$} estimated baseline models for each task.  
We then add current and lagged values (analogous to the baseline) for each of
the four signals separately. We now provide further details on each of these
tasks.

\subsubsection{Task 1: Case Forecasting}

To fix notation, let $Y_{\ell,t}$ denote the smoothed COVID-19 case
incidence rate per 100,000 people for location (HRR) $\ell$ and time (day) $t$. 
Let $X_{\ell,t}$ denote
the one of the indicators described above, for example the Facebook \%
CLI-in-community signal, 
for location $\ell$ and time $t$.
We evaluate five models of the form:
\begin{align*}
Y_{\ell,t+d}
&\approx \alpha + \sum_{j=0}^2 \beta_j Y_{\ell,t-7j} \\
Y_{\ell,t+d}
&\approx \alpha + \sum_{j=0}^2 \beta_j Y_{\ell,t-7j} +
\sum_{j=0}^2 \gamma_j X_{\ell,t-7j}, 
\end{align*}
where we also exchange $X_{\ell, t}$ for the signals from \attn{CHNG and
  Google.} 
Here $d=7,\ldots, 21$ depending on the target value
(number of days ahead we predict).

Informally, the first model, which we will call the ``Cases'' model, 
bases its predictions of future case rates on the following three features:
current COVID-19 case rates and those 1 and 2 weeks back.
The second model, ``Cases + Facebook'', additionally incorporates the 
current Facebook signal and the Facebook signal from 1 and 2 weeks in the past.
The third model, ``Cases + Google'', is exactly the same but substitutes the 
Google signal instead of the Facebook one and similarly for ``Cases + Change''.

For each model, in order to make a forecast at time $t_0$
(to predict case rates at time $t_0+d$),
we estimate the unknown parameters using quantile
regression~\cite{KoenkerXiao2006}, 
training over all locations $\ell$,
and all times $t$ that are within the most recent 21 days of data
available up to and including time $t_0$.
Rather than predicting only the median (or mean) value for each forecast date
and each horizon, we estimate seven quantiles, coinciding 
with the county-level forecast submission guidelines for the
COVID-19 Forecast Hub~\cite{Cramer:2021}.  These are \{0.025, 0.1,
0.25, 0.5, 
0.75, 0.9, 0.975\}. This provides a notion of uncertainty in the forecasts. An
example of this type of forecast is shown in \autoref{fig:trajectory}.

\begin{figure}[tb!]
    \centering
    \includegraphics[width=.9\linewidth]{fig/trajectory.pdf}
    \caption{Forecasts and observed cases for HRR x. \attn{Create a figure here
        for 1 HRR. Show the fan for the quantiles.}} 
    \label{fig:trajectory}
\end{figure}

% Noncrossing is enforced via post hoc ordering.

\subsubsection{Task 2: Hotspot Prediction}

We also examine whether alternative indicators help in predicting hotspots. On a
given date, we say an HRR is a hotspot if the average number of observed cases
over the past 7 days has increased by at least $25$\% compared to the preceding
week. We remove all data where, on average, fewer than $30$ cases were observed
over the preceding $7$ days. Thus, hotspot status can be represented by a
$\{0,1\}$-valued variable, and hotspot prediction is a binary classification
problem.

As in the case forecasting task, in hotspot prediction we train models that use
previously observed values of cases and alternative indicators as features to
predict a response (in our case, hotspot status). In addition to the differently
defined response, there are two other differences between the hotspot prediction
and case forecasting tasks. First, in hotspot prediction we always use relative
changes in previous observed cases/indicators as features. That is, rather than
$Y_{t,\ell}$ we use 
$$
Y^\Delta_{\ell,t} = \frac{Y_{\ell,t} - Y_{\ell,
    t-7}}{Y_{\ell,t-7}}\quad\text{and}\quad X^\Delta_{\ell,t} =
\frac{X_{\ell,t} - X_{\ell, t-7}}{X_{\ell,t-7}} 
$$ 
Second, because hotspot prediction is a binary classification problem, we
estimate the parameters $\alpha$, $\beta$ and $\gamma$ by logistic regression
rather than quantile regression.

\subsection{Evaluation Metrics}

We evaluate forecasts for each forecast date and horizon using the weighted
interval score. Weighted interval score (WIS), a well-known quantile-based
approximation of the commonly-used continuous ranked probability
score~\cite{gneiting-proper-score}. WIS is a proper score, meaning that lower
scores imply better forecasters, and can be thought of as a distributional
generalization of absolute error. The CDC has embraced this metric in the
context of epidemic forecasting and especially for COVID-19~\cite{wis-paper}.
Essentially, pairs of quantile forecasts are treated as a central  interval. For
example, the 0.1 and 0.9 quantiles create an 80\% interval that should, if the
forecaster is well-calibrated, contain the truth about 80\% of the time. A
forecaster whose 80\% interval contains the truth less frequently is overly
confident, while one whose interval contains the truth more frequently (say, by
using an infinitely-wide interval) is failing to take a stand on the future
outcome.

Here, denoting $K$ predicted intervals along with the median  as
$\{\alpha_K/2,\ldots,\alpha_2/2,\  
\alpha_1/2,\ 0.5, 1-\alpha_1/2,\ 1-\alpha_2/2,\ldots, 1-\alpha_K/2\}$, the associated
forecasts as $\{l_K,\ldots, l_2,\ l_1,\ m,\ u_1,\ u_2,\ldots, u_K \}$, and the observed
value (after the forecast was made) as $y$, this metric can be expressed as
\attn{someone please check this} 
\begin{align*}
 \operatorname{WIS}_K &= \frac{1}{K}|m -y | + \frac{2}{K+1}\sum_{k=1}^K 
\frac{1}{\alpha_k}(u_k - l_k)\\ &\quad + \sum_{k=1}^K  
(l_k - y)\indicator{y < l_k} +
(y - u_k)\indicator{y> u_k},
\end{align*}
where $\indicator{a} = 1$ if $a$ is true and $0$ otherwise. Essentially, this
metric decomposes into 4 parts: (1) a penalty for the median, the point forecast
or best guess, being far from the truth; (2) a penalty for the width of each
interval ($u_k - l_k$); (3)  a penalty for overprediction, when the observed
value falls below $l_k$ and hence underneath the interval; and (4) a
corresponding penalty for underprediction. Weighted interval score can also be
written in terms of an average of the quantile or ``pinball'' losses. Setting $\tau_1 =
\alpha_K/2,\ \ldots,\ \tau_{K+1} = 0.5,\ \ldots,\ \tau_{2K+1} = 1-\alpha_K/2$ and
similarly rewriting $q_1 = l_K,\ \ldots,\ q_{K+1}=m,\ \ldots,\ q_{2K+1}=u_K$, then
\begin{align*}
  \operatorname{WIS}_K &= \frac{2}{2K+1} \sum_{k=1}^{2K+1}
                         \left(\indicator{y \leq q_k} -\tau_k\right)
                         (q_k - y).
\end{align*}
While this second expression does not have the convenient decomposition
discussed above, the pinball loss is precisely the objective function that
quantile regression minimizes. So, mathematically, quantile regression should
give parameter estimates that minimize the in-sample weighted interval score.
% DJM: In trying to figure out where 'pinball' comes from, it looks like the
% preprint of the massively cited Koenker and Hallock (2001) JEP article. See
% http://www.econ.uiuc.edu/~roger/research/intro/rq.pdf. They say that tilting
% the absolute value is `pinball logic' which is not a connection I would have
% ever made. The language doesn't seem to have made it to the final version.
% Maybe there's an earlier reference?

When comparing the aggregate performance of the baseline quantile autoregression
without additional signals with 
the indicator-assisted quantile autoregression forecasters, we first
normalize the individual weighted interval scores by the weighted interval
score of a boneheaded model: predicting all future values to be the same as the
most recent observation. The dummy model uses the errors such a forecast has
made in-sample to form the other quantiles. We normalize by 
the dummy model to account for the fact that different locations or
time periods have different magnitudes of COVID-19 incidence (even after
scaling by population). Performing well by this metric means that a particular
forecaster is doing well at all locations and time periods. Without the scaling,
a forecaster that is good at locations and time periods with relatively high
case numbers but poor everywhere else could appear preferable.

For hotspots, we use the area under the curve (AUC) computed by examining the
true-positive and false-negative rates across HRRs at each horizon and forecast
date. That is, for a particular forecast date and a particular horizon, we ask,
how many true hotspots were called hotspots compared to the total number of
hotspots (specificity). The false-negative rate (sensitivity) is the ratio of
correctly classified non-hotspots to the total number of non-hotspots. A perfect
classifier would have both sensitivity and specificity equal to 1. Plotting the
sensitivity on the $y$-axis and $1-$specificity on the $x$-axis where each point
is a forecast date produces an ROC curve for each classifier. We compute the
area under this curve for each forecast horizon to compare the cases-only model
to those adding indicators.

\paragraph{\attn{Methods section graveyard, most for Appendix}}

\begin{itemize}
\item Discuss the "how many days does the indicator buy" interpretation here or below?

\item Do we explain why regularization/CV not really helpful for both
forecasting and hotspots

\item Implementation was via quantgen, cite here?

\item some people don't like relative WIS beacuse it's no longer a
``proper score''.  directly address that here?

\item Do we talk about log loss?

\item Do we explain that we are using infinite-backfilled data?

\end{itemize}

\section{Results}

\subsection{Primary Results}

Figure \ref{fig:topline-overall-results} displays the primary results
of this paper, showing the effect of using indicators on the two
forecasting tasks as a function of the horizon of the forecast (i.e.,
the number of days ahead one is forecasting). For the case rate forecasting
task, the left panel shows the ratio of the mean
WIS 
of each forecaster relative to the mean WIS of the
baseline. \attn{Further description and justification of this metric
  should be added to the ``Evaluation Metrics'' section rather than
  here it seems?}  Both means are taken over all HRRs and forecast
dates ranging from June 9, 2020 to December 31, 2020 \attn{verify}.
All curves are well below 1, which means that all methods, including \ar, outperform the baseline across all
considered forecast horizons. The claims-based
indicators \dv~and \chngcov~show the most substantial gains, whereas
the \chngcli~does not show any improvement over \ar.  The \fb~and
\gs~indicators also show a non-negligible improvement.  Notably,
\fb~is the only indicator to offer a clear improvement at far aheads.
%\begin{itemize}
%\item \dv~and \chngcov~perform very well
%\item \fb~and \gs~perform well
%\item \chngcli~performs poorly
%\item Interestingly, fb is the only one offering clear improvement at
 % far aheads
%\item Note the y-axis: consistent, clear improvement over baseline
%\end{itemize}

For the hotspot prediction task, the right panel of Figure
\ref{fig:topline-overall-results} shows the AUC of each method.
The monotonic decrease of all curves reflects the fact that hotspot
prediction gets harder as the forecasting horizon increases. At the
highest considered ahead (nearly one month ahead), the AUCs are close
  to 0.5, which is no better than random guessing.
In the 7-14 day ahead range, the indicators show a noticeable
  improvement over \ar.  For larger horizons there does not
  appear to be much difference in methods.
The approximate ordering of methods is similar to what is seen
  in the forecasting task: \chngcov~and \dv~are the strongest
  performing and \ar~the weakest.
Even for 7-day ahead predictions, the AUC is below 0.7,
  suggesting that all methods struggle with this classification
  task. \attn{Do people agree that an AUC of 0.7 is not all that great?}
  That said, \dv~and \chngcov~attain roughly a 10\% increase in AUC
  over the \ar~classifier.

%\begin{itemize}
%\item Monotonic decrease reflects the fact that hotspot
%  prediction gets harder as ahead increases.
%\item At the highest ahead (almost one month ahead) the AUCs are close
%  to 0.5, which is no better than random guessing.
%\item In the 7-14 day ahead range, the indicators show a noticeable
%  improvement over the \ar.  For larger horizons there does not
%  appear to be much difference in methods
%\item The approximate ordering of methods is similar to what is seen
%  in the forecasting task: chng\_cov and dv are the strongest
%  performing and \ar the weakest.
%\item Even for 7-day ahead predictions, the AUC is below 0.7,
%  suggesting that all methods struggle with this classification task.
%  That said, dv and chng\_cov attain roughly a 10\% increase in AUC
%  over the \ar classifier.
%\end{itemize}

% We see that all three \attn{will need to
% include imputed-zero google here, hopefully four} indicator-assisted
% forecasters improve upon the baseline QAR3 forecaster.  The QAR3 model
% assisted by the Change Healthcare confirmed COVID-19 signal
% (QAR3+CHCOV3) sees the strongest improvement.
% C`

\begin{figure}[tb!]
  \centering
  \includegraphics[width=\linewidth]{fig/figure5.pdf}
  \caption{\label{fig:topline-overall-results}  Forecasting performance as a
    function of number of days 
    ahead, aggregated across all HRRs and forecast dates. \attn{How's this?}}
\end{figure}

\subsection{Avoiding misleading retrospective evaluations}

Figure \ref{fig:finalized-vs-honest} quantifies the effect of
not properly accounting for the question of ``what was known when'' in
performing retrospective evaluations of forecasters.  When methods are
given the finalized version of the data rather than the version available at the time that the forecast would have been made, all methods
appear (falsely) to have better performance.  For example, for forecasting case rates
7-days ahead, the WIS of all methods is at least 8\% larger than what would have been
recorded using the finalized values of the data.  This effect
diminishes as the forecasting horizon increases, reflecting the fact
that these forecasters rely less heavily on recent data than very
short-term forecasters.  Crucially, some methods are
``helped'' more than others by the less scrupulous retrospective
evaluation, underscoring the difficulty of avoiding misleading
conclusions when performing retrospective evaluations of forecasters.

The \chngcli~indicator (along with the other claims-based signals) is the most
affected by this distinction, reflecting the latency in claims-based
reporting. \attn{Actually, \dv~is the least affected for forecasting
  whereas it is one of the most affected for hotspots.}  This supports
the importance of efforts to provide ``nowcasts'' for claims signals
(which corresponds to a 0-ahead ``forecast'' of what the claims
signal's value will be once all data has been collected).

Even the \ar~model is affected by this distinction, reflecting the
fact that the case rates themselves (i.e., the response values) are also
subject to revision.  The forecasters based on indicators are thus
affected both by revisions to the indicators and by revisions to the
case rates.

\begin{itemize}
%\item Motivates the importance of nowcasting for claims signals.
  \item For forecasting task: Also, \chngcli~and \dv~perform very similarly here, which is
    reassuring because they're measuring the same thing (in
    principle).  Yet they must have very different backfill profiles…
  \item For hotspots task: dv seems to get a bigger boost than
    \chngcov - interesting dichotomy to forecasting
\item But in both forecasting/hotspots \chngcli~is affected a lot by backfill
\item \attn{We should be careful about which indicators we were not
      able to track latency for.  In fact, we should probably remove
      these from the plot since they would be more strongly affected
      than is shown in the plot.}

\item \attn{We can also discuss the distinction
      between indicators that could in principle have low latency
      (e.g. GS) versus those where latency is inherent
      (e.g. claims-based indicators). If the goal is assessing the
      performance of GS ``for the next pandemic'' then it would make
      sense to give GS a pass since in theory they could have had it
      with very low latency (as opposed to claims based signals where
      backfill may be unavoidable).  Of course, in
    ``the next pandemic'' there might not be such highly specific
    symptoms such as A+A as there was in the COVID-19 case.} 
\end{itemize}

\begin{figure}[tb!]
     \centering
     \includegraphics[width=\linewidth]{fig/figure6.pdf}
     \caption{How misleading is a retrospective analysis that uses the
       finalized version of the data rather than the version that was
       in fact available at the time the forecast was made?  Plots
       show the ratio in performance (using mean WIS fore the
       forecasting task and AUC for the hotspot prediction
       task). \label{fig:finalized-vs-honest} Cheating vs. honest plots.} 
\end{figure}

\subsection{Additional sensitivity analyses}

In addition to whether one properly accounts for signal latency, there
are several other choices in evaluating forecasters that can also have
a non-trivial impact on results.

For example, in Figure
\ref{fig:topline-overall-results}, we show the ratio
of means.  Other choices of aggregation and scaling are also
possible.  For example, taking the mean of ratios of errors is
possible and would give more importance to locations where
errors are small (either because the location-time pairs had low case
rates or because the forecasting task was itself easier).  By
contrast, the ratio of the mean of errors may be dominated by
location-time pairs where the error is large.
Figure \ref{fig:
  scale-aggregate-order} of the Supplementary Information shows...
\begin{itemize}
\item Note the y-axis: barely an improvement over baseline, and this goes away for AR after 11 days ahead, and for all other models around 13-14 days ahead.
\item Using trimmed means here because otherwise the results are extremely unstable (as the denominators can be small, when WIS of baseline is small).
\item Similar message to before (non-adjusted), but \dv~and \chngcov~the best, \fb~and \gs~still good, but now \chngcli~seems to do much better!  
\item What's our interpretation here?  That somehow \chngcli~just
  does particularly bad in tasks where the baseline is also bad?
  \item We might include a histogram of WIS with a log-transformed
    x-axis to suggest that the geometric mean may be reasonable here.
\end{itemize}
Using instead the geometric mean makes the order of aggregation and
scaling immaterial since the ratio of geometric means is the same as
the geometric mean of ratios.  Figure \ref{fig:geometric-mean} of the
Supplementary Information shows how Figure
\ref{fig:topline-overall-results} panel A would change if
one replaces the arithmetic mean by the geometric mean.

In Supplementary Information \ref{sec:2021}, we investigate the effect of lengthening
the time window to include 2021.
\begin{itemize}
\item For forecasting: results get better, whether we look at
  non-adjusted (first plot) or adjusted scores (second plot).  Other
  than that, interpretation is qualitatively similar, except fb now
  seems to be worse than gs.  And \chngcli~is strong, even in the
  non-adjusted view.
  \item For hotspots: Alden will write something explaining why we
    don't do hotspot detection in 2021.
\end{itemize}

In Supplementary Information Section \ref{sec:gs-locations}, we look at various approaches to
accounting for the fact that the \gs~indicator is only available
for XX HRRs.

\begin{itemize}
  \item For forecasting task: clearly gs is not missing at random, and when it's present, it tends to be predictive.  Hence high values have a low false positivity rate.  Most visible in the adjusted view below, where gs triumphs.  Also \chngcli~gets a lot worse.

\item For hotspots: All methods look better, suggesting that the “GS locations” are easier for hotspot prediction. E.g. at 7-days ahead, the AUCs computed based on all locations range from 0.61-0.66; restricted to GS locations, the AUCs range from about 0.65-0.69. (These are all eyeballed, should get actual numbers if we want to say something like this in paper.)
\item gs\_subset appears to be particularly helped by this subsetting
  (which makes sense since on those other locations it was 0-imputed)
  \item Daniel will also try gs\_impute as in forecasting
\end{itemize}

% Figure \ref{fig:forecast-median-relative-wis-faceted} splits the topline
% median relative WIS results over four time periods:  June and July, which
% corresponds to the small, initial surge experienced in the US;
% August and September, when incident cases nationally were somewhat stable;
% October, when a second surge was witnessed; and November and December,
% which saw the largest surges to date.  In essentially all the time periods,
% the indicator-assisted models perform in aggregate as well or better than
% the baseline QAR3 model.  The results in October do not show much
% differentiation between the baseline QAR3 and the indicator-assisted models.
% However, we note that during that time period, even the baseline QAR3
% model is doing very well in terms of relative WIS (especially compared to
% other time periods), so perhaps there is not much ``room for improvement'',
% so to speak.

% \begin{figure*}[tb!]
%      \centering
%      \begin{subfigure}[b]{0.49\textwidth}
%          \centering
%          \includegraphics[width=\textwidth]{fig/reduction_mean_scaled_wis_facet_honest.pdf}
%          \caption{Forecasting task: mean WIS relative to mean WIS of baseline}
%          \label{fig:forecast-reduction-mean-scaled-wis-facet}
%      \end{subfigure}
%      \hfill
%      \begin{subfigure}[b]{0.49\textwidth}
%          \centering
%          \includegraphics[width=\textwidth]{fig/auc_by_forecast_period.pdf}
%          \caption{Hotspot prediction task: AUC}
%          \label{fig:hotspot-auc-faceted}
%      \end{subfigure}
%      \label{fig:topline-faceted-results}
%      \caption{Topline results from June - December 2020, split into
%      four representative time periods.}
% \end{figure*}

\paragraph{\attn{Additional results section material, some for Discussion}}

\begin{figure}
    \centering
    \includegraphics[width=.95\linewidth]{fig/fake-map.pdf}
    \caption{Weighted interval score for forecasts made on April 1, 2021. Scores
      are averaged over horizons from 1 to 14 days ahead and are relative to
      population size.} 
    \label{fig:fake-map}
\end{figure}

\begin{enumerate}
%\item How does this vary over time? Table showing 7-day ahead and 14-day ahead
%  (just picking out two of the d=5,...,20). Include a national curve.
\item Spatial map of errors across HRRs for both tasks. See
  \autoref{fig:fake-map} for an example. 
\item Unlikely we can find it, but we should try anyway. Possibly put in the
  intro or discussion instead. Find an example of a forecast altered by the
  presence of an indicator.  Plot the two trajectories.
\item Examine a specific HRR over time. What does it mean to predict certain
  hotspots? How can we understand what is happening?
 \item What can we say about statistical significance? NRI, paired t-test?
 \item Figures showing hotspots make sense
 \item Horizontal line at 0.5 on AUC curve, discribe that this is random
   guessing. 
\end{enumerate}

\section{Discussion}

We have assessed the performance of [NUMBER] indicators in terms of their
usefulness for performing two forecasting tasks: probabilistic
forecasting of case
rates and hotspot prediction at the HRR-level, 1--[NUMBER] days in advance.
In measuring the practical usefulness of an indicator to forecasting,
we distinguish between honest versus aspirational assessments.  Data
latency is an important factor when considering the usefulness of an
indicator.  Retrospective evaluations that ignore these messy data
issues fail to provide the full picture.  While one can hope that reporting
systems will operate faster and more consistently in ``the next
pandemic,'' some reasons for this latency appear unavoidable.  The
chaos and strain inherent to pandemic times make indicators that can
be reported and collected with quickly with minimal effort or manual
involvement desirable.

We conclude by observing that the
approach we have taken in this paper is somewhat opposite to that of much of the COVID-19
forecasting literature.  We have chosen to consider only very simple forecasting models while devoting most of our effort to accounting for as much of the complexity of
the data and evaluation as possible.  By contrast, many papers focus
on very complicated forecasting approaches but then evaluate them
under unrealistic, retrospective conditions.

\begin{enumerate}

\item Discussion should contain more detailed lit review ("context"), rather
  than in the intro, per the instructions.

\item Use of syndromic surveillance for other diseases?

\item Cite any papers that use our signals in forecasting.
\end{enumerate}

\paragraph{\attn{Other potential discussion topics}}

\begin{itemize}

\item When was this data actually available? (As of issue. The argument is: sure
  we didn't have it, but it would have helped. We should have this next time.)
\item Side issues, are there places / times where one signal is especially good?
  Especially bad?
\item What do we do about holidays? How do we understand the performance when
  the data is trash? If the data is trash, all the more reason to have other
  data that is perhaps less susceptible to these criticisms.
\item Other signals we might want? Mobility? Discuss how SEIR models need
  contact matrices, relationship with ``in community'' signals.

\end{itemize}

\showmatmethods{} % Display the Materials and Methods section

\acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}

\showacknow{} % Display the acknowledgments section

% Bibliography
\bibliography{../../common/covidcast.bib}

\end{document}

### 2021-06-11 forecast/paper/supplement.tex
Transfer stuff over
- Mainly transferring stuff over from old GitHub repos
- Light updates to indicators paper, forecast paper intro

\documentclass[9pt,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnassupportinginfo}

\input{forecast-prelim.tex}

\title{Your main manuscript title}
\author{Author1, Author2 and Author3 (complete author list)}
\correspondingauthor{Corresponding Author name.\\E-mail: author.two@email.com}

\begin{document}

%% Comment out or remove this line before generating final copy for submission; this will also remove the warning re: "Consecutive odd pages found".
\instructionspage

\maketitle

%% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
\SItext

\subsection*{Subhead}
Type or paste text here. This should be additional explanatory text such as an extended technical description of results, full details of mathematical models, etc.

\section*{Heading}
\subsection*{Subhead}
Type or paste text here. You may break this section up into subheads as needed (e.g., one section on ``Materials'' and one on ``Methods'').

\subsection*{Materials}
Add a materials subsection if you need to.

\subsection*{Methods}
Add a methods subsection if you need to.

%%% Each figure should be on its own page
\begin{figure}
\centering
\includegraphics[width=\textwidth]{example-image}
\caption{First figure}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{frog}
\caption{Second figure}
\end{figure}

\begin{table}\centering
\caption{This is a table}

\begin{tabular}{lrrr}
Species & CBS & CV & G3 \\
\midrule
1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
\bottomrule
\end{tabular}
\end{table}

%%% Add this line AFTER all your figures and tables
\FloatBarrier

\movie{Type legend for the movie here.}

\movie{Type legend for the other movie here. Adding longer text to show what happens, to decide on alignment and/or indentations.}

\movie{A third movie, just for kicks.}

\dataset{dataset_one.txt}{Type or paste legend here.}

\dataset{dataset_two.txt}{Type or paste legend here. Adding longer text to show what happens, to decide on alignment and/or indentations for multi-line or paragraph captions.}

\bibliography{pnas-sample}

\end{document}

### 2021-06-11 indicators/paper/paper.tex
Transfer stuff over
- Mainly transferring stuff over from old GitHub repos
- Light updates to indicators paper, forecast paper intro

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasresearcharticle} % Choose template
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

\input{../../common/prelim.tex}

\title{An Open-Source Repository of Real-Time COVID-19 Indicators}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,c,1]{Author One}
\author[b,1,2]{Author Two}
\author[a]{Author Three}

\affil[a]{Affiliation One}
\affil[b]{Affiliation Two}
\affil[c]{Affiliation Three}

% Please give the surname of the lead author for the running footer
\leadauthor{Lead author last name}

% Please add a significance statement to explain the relevance of your work
\significancestatement{To properly study the COVID-19 pandemic, its effects on
  society, and measures for reducing its spread, researchers need detailed data
  on the course of the pandemic. Standard public health data streams suffer
  inconsistent reporting and frequent, unexpected revisions. They also miss some
  important aspects worthy of study, such as certain aspects of public
  behavior. We present an open database of COVID signals in the United States,
  measured primarily at the county level and updated daily. These include
  traditionally reported COVID cases and deaths, but also many others: signals
  based on measures of mobility, social distancing behaviors, internet search
  trends, self-reported symptoms, and patterns COVID-related activity in
  de-identified medical insurance claims. The database provides all signals in a
  common, easy-to-use format, empowering both public health research and
  operational decision-making.
}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{Please provide details of author contributions here.}
\authordeclaration{Please declare any competing interests here.}
\equalauthors{\textsuperscript{1}A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable).}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: author.two\@email.com}

% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
\keywords{open data $|$ COVID-19 $|$ Keyword 3 $|$ ...}

\begin{abstract}
  Please provide an abstract of no more than 250 words in a single paragraph.
  Abstracts should explain to the general reader the major contributions of the
  article. References in the abstract must be cited in full within the abstract
  itself and cited in the text.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

\dropcap{P}ublic health decision-makers, healthcare providers, epidemiological
researchers, employers, institutions, and the general public benefit from
promptly and readily accessible data regarding COVID-19 activity levels,
countermeasures, and pandemic impact.  Real-time indicators of COVID-19 activity
levels, such as statistics on cases, deaths, test positivity, and
hospitalizations, enable reports and interactive dashboard applications for
situational awareness \cite{Dong:2020, NYTimes, USAFacts}, and are essential for
most analyses of the pandemic.  These data are available for locations across
the United States from a number of official sources and independent aggregators
in varied and inconsistent formats.  Different data types and sources vary in
timeliness, based on when measured events occur in the progression of the
disease, testing capabilities, the reporting pipeline, and their publication
schedules.

Additional, auxiliary data sources can improve on the timeliness, scope, and
utility of the ``topline'' indicators (cases, test positivity, hospitalizations,
deaths) coming from the public health reporting system. For example, in the
context of other infectious diseases: syndromic surveillance in ambulatory
clinics and emergency rooms improves the accuracy of outbreak detection for
emerging pathogens such as H1N1 \cite{Kass-Hout:2012}; and digital surveillance
(based on, e.g., search and social media trends) enable more accurate
``nowcasts'' and forecasts of traditional disease surveillance streams such as
the CDC's ILINet \cite{Santillana:2015, Farrow:2016}, as do publication formats
providing access to historical versions of a data set \cite{Brooks:2018,
  Brooks:2020}. Numerous other examples exist spanning a wide variety of data
platforms and diseases \cite{Browstein:2009, Kass-Hout:2011, Salathe:2012,
  Kass-Hout:2013}. During the COVID-19 pandemic, digital data streams have
permitted faster prediction of case increases \cite{Ahmad:2020, Kogan:2021},
while enabling analyses of the impact of public health policies on public
behavior, the economy, and disease spread \cite{Bonaccorsi:2020, Nouvellet:2021,
  Adjodah:2021, Jewell:2021}.

The Delphi Group works with partner organizations and public data sets to build
a massive database of indicators tracking COVID-19 activity and other relevant
phenomena, which has been publicly available and continuously updated since
April 2020. Alongside public data on reported cases and deaths, this database
includes several unique data streams, including indicators extracted from
de-identified medical claims data, antigen test results from a major testing
manufacturer, large-scale public surveys that measure symptoms and public
behavior, and symptom indicators based on Google search query volume. (We use
the terms ``indicator'' and ``signal'' interchangeably.) We make aggregate
signals publicly available, generally at the county level, via the COVIDcast API  
\cite{CovidcastAPI}. We store and provide access to all previous (historical)
versions of the signals, a key feature that exposes the effects of data
revisions.  Lastly, we provide R \cite{CovidcastR} and Python \cite{CovidcastPy}  
packages to facilitate interaction with the API, and an online dashboard to
visualize the data \cite{CovidcastViz}.

% RJT: We should fiddle with the bib so that the URLs actually appear in the
% references at the end.  Otherwise it's kind of pointless to cite manuals or
% websites like the COVIDcast API site or R/Python manuals.

In a companion paper, we analyze the utility provided by a core set of the
indicators in COVID-19 forecasting and hotspot prediction models.  In another
companion paper, we elaborate on our research group's (Delphi's) large-scale
public surveys, run in partnership with Facebook and available in aggregate form
in the COVIDcast API.

\section{Methods}

\subsection{Data Collection}

We receive data daily from healthcare partners, technology companies, and from
surveys conducted daily by Delphi in partnership with Facebook. These data
sources provide information not available from standard public health reporting
or other common sources, such as:
\begin{description}
\item[Health Insurance Claims] Based on de-identified medical insurance claims
  from Change Healthcare and other health system partners, we release indicators
  on the estimated percentage of covered outpatient visits and hospitalizations
  that involved COVID diagnoses or symptoms.
\item[Internet-Based Surveys] Conducted in partnership with Facebook, Delphi's
  COVID Tracking and Impact Survey receives about 50,000 responses daily, and
  has received over 20 million responses since April 2020 \cite{DelphiSurvey,
    Kreuter:2020}. From the surveys, we construct indicators on symptoms,
  social distancing, vaccination, and other attitudes and behaviors related to
  COVID.
\item[COVID Antigen Tests] Based on data from Quidel, a major manufacturer of
  COVID antigen tests in the United States, we calculate and release
  (Quidel-specific) test volumes and positivity rates.
\item[Search Trends] Based on Google's COVID-19 Search Trends symptoms data set
  \cite{GoogleSymptoms}, we provide indicators reflecting COVID-related search
  activity.
\item[Mobility Data] Cell phone mobility data is collected by SafeGraph
  \cite{SafeGraphSD, SafeGraphPatterns}, and is available to researchers under a
  data use agreement; we aggregate it to the county level and make indicators
  publicly available.
\end{description}

We also scrape data accessible from other public sources, such as cases and
deaths data aggregated from public reporting by JHU CSSE \cite{Dong:2020} and by
USAFacts \cite{USAFacts}, so that we can track revisions and updates to this data
(see Section~\ref{subsec:revision_tracking}).

Together, we collect TODO signals from TODO distinct sources, aggregate them,
and provide them in a single common format for public access. This hence
provides a common format both for unique data streams not available elsewhere
and for public data streams, enabling efficient comparison and modeling use.

A summary of the data sources we currently receive and make available is
presented in Table~\ref{data-sources}, and detailed documentation of the signals
available from each source is available at
\url{https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html}.

\begin{table*}
\centering
\caption{Data sources available in Delphi's COVIDcast API \cite{CovidcastAPI},
  as of date of publication. The first group of data sources are produced by
  Delphi from data not otherwise available publicly (or only available in
  limited form); the second group is mirrored from public sources.}
\begin{tabular}{>{\raggedright}p{1.2in} p{4.0in} r >{\raggedright\arraybackslash}p{0.5in}}\toprule
\textbf{Data source} & \textbf{Signals available} & \textbf{First date} &
\textbf{Resolution} \\\midrule
  % Exclusive data
  Change Healthcare & Percentage of outpatient visits with COVID diagnostic codes
or codes indicating COVID-like symptoms. Based on de-identified claims data
processed by Change Healthcare. & 2020-02-01 & County$^*$ \\
  Doctor Visits & Percentage of outpatient visits primarily about COVID-like
symptoms, based on de-identified claims data provided by health system
partners. & 2020-02-01 & County$^*$ \\
  Hospital Admissions & Percentage of new hospital admissions with COVID
diagnostic codes, based on de-identified claims data provided by health system
partners. & 2020-02-01 & County$^{**}$\\
  Quidel & Test positivity rates for COVID-19 antigen tests produced by
Quidel. & 2020-05-26 & County$^{**}$ \\
  SafeGraph & Mobility metrics, such as time away from home or visits to bars
and restaurants, based on cell phone mobility data collected by SafeGraph
\cite{SafeGraphSD, SafeGraphPatterns}. & 2019-01-01 & County \\
 COVID Tracking and Impact Survey & COVID symptoms, social distancing behaviors,
mental health, economic impact, behavior (e.g., mask-wearing, vaccination
attitudes), and COVID testing signals based on daily surveys conducted
nationally by Delphi through Facebook \cite{DelphiSurvey, Kreuter:2020}. &
2020-04-06 & County$^{**}$ \\
  \midrule
    % Mirrored data from other sources
  Health \& Human Services & Counts of hospital admissions due to confirmed or
suspected COVID-19, as reported by the Department of Health \& Human Services. &
2019-12-31 & State \\
    COVID Act Now & COVID-19 testing results, such as positivity and number of
tests, compiled by COVID Act Now from CDC reporting. & 2020-03-02 &
County$^*$ \\
  Google Symptoms & Trends in Google search volume for terms related to anosmia
and ageusia (loss of smell or taste), which correlate with COVID activity, based
on data shared by Google \cite{GoogleSymptoms}. & 2020-02-13 & County$^{***}$ \\
  Cases and Deaths & Confirmed COVID-19 cases and deaths, compiled by JHU CSSE
\cite{Dong:2020} and by USAFacts \cite{USAFacts}. & 2020-01-22 & County \\
  NCHS Mortality & Weekly totals of deaths broken down by cause, such as COVID,
flu, or pneumonia, compiled by the National Center for Health Statistics [TODO
cite]. & 2020-01-26 & State \\
    \bottomrule
\end{tabular}
% TODO PNAS wants a short title above and caption below, but they can't possibly
% mean for the caption to be formatted so weirdly...
\addtabletext{Location availability often drops lower for some times and
  signals, e.g., due to variable reporting volume. $^*$Available at $>60$\% of
  counties; $^{**}$available at 20--60\% of counties; $^{***}$available at
  $<20$\% of counties.}
\label{data-sources}
\end{table*}

\subsection{Signal Processing}

Because each data source reports data in different formats, we must convert each
source to a common format. In this format, each record represents an observation
of one quantity at one time point in one location. Locations are coded
consistently using standard identifiers such as FIPS codes; the sample size and
standard error for each observation is also reported when applicable. Each
signal is reported at the finest geographic resolution its source supports (such
as county or state) and also aggregated to metropolitan statistical areas,
Health and Human Services regions, and Hospital Referral Regions. National
averages are also provided. Crucially, each record includes an \textit{issue
  date} recording when the value was first issued, as described below. This
allows tracking of multiple revisions to individual observations, each revision
with a separate issue date.

When appropriate, additional post-processing (often nontrivial) is applied to
the data. For example, data on visits to doctor's offices is subject to strong
day-of-week effects, and so regression is used to adjust for these
effects. Other indicators are available in raw versions and versions smoothed
with a seven-day moving average. All processing is done using open-source code
written primarily in Python and R, and available publicly at
\url{https://github.com/cmu-delphi/covidcast-indicators/}.

\subsection{Revision Tracking}
\label{subsec:revision_tracking}

Many data sources that are useful for epidemic tracking are subject to revision
after their initial publication. For example, aggregated medical claims data may
be initially published after several days, but additional claims and corrections
may take days to weeks to be discovered, processed, and aggregated. Medical
testing data are also often subject to backlogs and reporting delays, and
estimates for any particular date are revised over time as errors are found or
additional data becomes available. This revision process is generally referred
to as \textit{backfill}.

For this reason, the COVIDcast API annotates every observation with two dates:
the \textit{time value}, the date the underlying events (such as tests or
doctor's visits) occurred, and the \textit{issue date} when Delphi aggregated
and reported the data for that time value. Importantly, there can be multiple
observations for a single time value with different issue dates, for example if
data is revised or claims records arrive late. Delphi tracks revisions to all
data sources we ingest, including external data sources (such as sources
tracking cases and deaths). Many external sources do not keep a public or
conveniently accessible record of revisions of their data.

For many purposes it is sufficient to know the most recently issued observation
for any time value, and the COVIDcast API reports the most recent issue by
default. However, for some purposes it is crucial to know what was known
\textit{as of} a specific date. For example, an epidemic forecasting model will
be called upon to make its forecasts based on preliminary data about recent
trends, and so when it is trained using historical data, it should be trained
using the initial versions of that data, not updates that would have been
received later. Furthermore, these revision records allow models to be modified
to account for noise and bias in early data versions, or to exclude data that is
too new to be considered stable, and to ``rewind'' time and simulate how these
revised models would have performed using only the versions of data available
\textit{as of} those times.

Research on data revisions in the context of influenza-like illness has shown
that backfill can significantly alter forecast performance \cite{Brooks:2018,
  Reich:2019}, and that training to use preliminary data can reduce this
influence \cite{Brooks:2018, Brooks:2020}. This is also investigated in our
companion paper on forecasting, where we observe that training and validating
models on finalized data yields overly optimistic estimates of true test-time
performance.

\subsection{Public API}

The data described above is publicly available through the Delphi COVIDcast API
\cite{CovidcastAPI}.  By making HTTP requests specifying the data source,
signal, geographic level, and time period desired, users can receive data in
JSON or CSV form. For added convenience, we have written \texttt{covidcast} R
\cite{CovidcastR} and Python \cite{CovidcastPy} packages with functions to
request data, format it as a data frame, plot and map it, and combine it with
data from other sources.

The R and Python package software is public and open-source, at
\url{https://github.com/cmu-delphi/covidcast/}.  The underlying server software
is itself also public and open-source, at
\url{https://github.com/cmu-delphi/delphi-epidata/}.  Lastly, most data sources
are provided under the Creative Commons Attribution license, and a small number
have additional restrictions imposed by the data source; see
\url{https://cmu-delphi.github.io/delphi-epidata/api/covidcast_licensing.html}.

\subsection{Interactive Visualization}

% RJT: @AR I'm now thinking we should add a very short paragraph or two on
% this. We can say that we provide various interactive maps highlighting both
% temporal (time series) and spatial aspects (choropleth or beehive maps), but
% also exploratory tools for correlations and anomalies.  (Very much like the
% correlations computed in the next section.)

% Another reason why I'm in favor of this is that we can now add everybody as a
% middle author who now contributed anything at all to COVIDcast, from database
% to API to indicator to visualization.  I'm sure the HCII students & faculty
% who worked with us would very much appreciate being on this paper.

\section{Results}

The indicators that are available in the COVIDcast API have been used in
dashboards produced by COVID Act Now \cite{CovidActNow}, COVID Exit Strategy
\cite{CovidExitStrategy}, and others; to inform the Delphi, DeepCOVID
\cite{Rodriguez:2021}, and the Institute for Health Metrics and Evaluation
(IHME) \cite{IHMEProj} COVID forecasting models; in various federal and state
government reports and analyses; and in a range of news stories. Aside from
operational use in decision-making and forecasting, they have also facilitated
numerous analyses studying the impacts of COVID-19 on the public, the
effectiveness of policy interventions, and factors that influenced the spread of
the pandemic \cite{Adjodah:2021, Pierri:2021, Jewell:2021, Chakrabarti:2020,
  Doerr:2021}. The API currently serves hundreds of thousands of requests each
day, to thousands of users worldwide.

In what follows, we present illustrative examples of the usefulness of some of
the novel signals available in the API. These examples demonstrate that such
indicators are meaningfully related to COVID activity, that they provide
alternate views on pandemic activity that are not subject to the same reporting
glitches and delays as traditional public health surveillance streams, and that
they provide information about public behavior and attitudes that are not
available from any other source. Code to reproduce all examples (which uses the
\texttt{covidcast} R package and fetches data directly from the API) can be
found at \url{TODO}.

\subsection{Tracking Trends}

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{fig/time_trends_national.pdf}
  \caption{National trends, from April 2020 to April 2021, of four signals in
    the COVIDcast API. The auxiliary signals, based on medical claims data and
    massive surveys, track changes in officially reported cases quite
    well. (They have all been placed on the same scale as reported cases per
    100,000 people.)}
  \label{fig:time_trends_national}
\end{figure}

Many of the indicators in the COVIDcast API are intended to track COVID
activity. Five indicators in particular have the closest connections to
confirmed COVID cases:
\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Tracking and Impact Survey CLI (CTIS-CLI): The estimated percentage
  of the population with COVID-like illness based on Delphi's surveys of
  Facebook users.
\item COVID Tracking and Impact Survey CLI-in-community (CTIS-CLI-in-community):
  The estimated percentage of the population who knows someone who is sick,
  based on Delphi's surveys of Facebook users.
\item Quidel COVID test positivity rate (Quidel-TPR): The percentage of positive
  test results among Quidel COVID antigen tests.
\end{itemize}
Figure~\ref{fig:time_trends_national} compares the first three of these signals
to reported COVID cases in the United States (from JHU CSSE, and smoothed with a
7-day trailing average) over the first year of the pandemic (April 15, 2020 to
April 15, 2021), illustrating how they  appear to track national trends in cases
quite well. Importantly, this same relationship persists broadly across multiple
resolutions, down to smaller geographic regions such as states and counties, as
shown in the supplement. This will also be illustrated in a more detailed
correlation analysis in the next subsection.

% RJT: @AR Note what I claimed we would do above!  I'm thinking we can just
% produce a big batch of case trends over the same time period and for the same
% signals, but for each US state (and territory?), and for the 50 most populous
% counties. AND maybe for this (below national) we should use CLI-in-community
% rather than CLI, from the survey

Besides tracking contemporaneous COVID activity, these and other indicators can
be used to improve forecasts of future COVID case trends, as investigated in a
companion paper.

% RJT: @AR Consider for the supplement.
% - Hospitalization trends?
% - Hospitalization correlations?
%  (I see you already have a plot for the first ... but I'm a bit bothered by
%  the fact that the HHS data looks to have such strong weekday effects that we
%  don't attempt to correct for ... perhaps we use do a 7-day rolling avg. Made
%  easy with modeltools)
% If we do so, we should refer to it in the text, this subsection (for time
% trends) and next (for correlations)

\subsection{Correlation Analyses}

To quantify the ability of the signals described above to track trends in COVID
cases, we use the Spearman (rank) correlation and analyze two key correlation
patterns, between each signal and confirmed COVID case rates (cases per 100,000
people):
\begin{enumerate}
\item \textit{Geo-wise correlations} (i.e., on a specific date, do values of the
  signal correlate with case rates across locations?): Formally, let $X_t$ and
  $Y_t$ be vectors of values of a particular signal and case rates, over all
  locations, on date $t$. The geo-wise correlation at time $t$ is defined as
  $\cor(X_t, Y_t)$. This examines whether a signal has the capability to help
  spot locations with high case rates at any given time.

\item \textit{Time-wise correlations} (i.e., at a specific location, do values
  of the signal correlate with case rates across time?): Formally, let $X_\ell$
  and $Y_\ell$ be vectors of a values of a particular signal and case rates,
  over all times, at location $\ell$. The time-wise correlation at time $t$ is
  defined as $\cor(X_\ell, Y_\ell)$. This examines whether changes in a signal
  over time correspond to changes in reported cases at the same location.
\end{enumerate}

\begin{figure}[t]

  \includegraphics[width=\columnwidth]{fig/geo_wise_corr.pdf}
  \caption{Geo-wise correlations with case rates, from April 15, 2020 to April
    15, 2021, calculated over all counties for which all signals were available
    and which had at least 500 cumulative cases by the end of this period.}
 \label{fig:geo_wise_correlation}
\end{figure}

Figure~\ref{fig:geo_wise_correlation} shows the geo-wise correlations achieved
by the five signals and COVID case rates (from USAFacts, and again smoothed
using a 7-day trailing average), from April 15, 2020 to April 15, 2021. The
locations we consider here are counties with at least 500 cumulative cases by
the end of this period, and at which all signals are available (956 counties in
total). The  significantly positive correlations suggest that these signals
could be useful in hotspot detection  (identifying counties with relatively high
COVID activity, at a given time). Somewhat surprisingly, the CLI-in-community
signal derived from the survey shows the strongest correlations for much of the
time period. This clearly demonstrates the value of a large-scale survey such as
CTIS for tracking symptoms and case trends, especially when other data is
unavailable.

Figure~\ref{fig:time_wise_correlation} summarizes time-wise correlations of
these five signals over the same time period, and for the same set of
counties. For each signal, we display the set correlations that it achieves in
histogram form (more precisely, using a kernel density estimate). All signals
produce positive correlations in the majority of counties considered (with very
little mass in each estimated density being to the left of zero). The largest
correlations, in bulk, are achieved by the CHNG-COVID signal; the
CTIS-CLI-in-community is a close second, and the CHNG-CLI signal is third.
There are two remarks deserving attention:

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{fig/time_wise_correlation.pdf}
  \caption{Time-wise correlations with case rates, from April 15, 2020 to April
    15, 2021, calculated over all counties for which all signals were available
    and which had at least 500 cumulative cases by the end of this period.}
  \label{fig:time_wise_correlation}
 \end{figure}

\begin{itemize}
\item This is different from what is observed in
  Figure~\ref{fig:geo_wise_correlation}, where the CTIS-CLI-in-community signal
  achieves clearly the highest correlations for most of the time
  period. However, it is worth emphasizing that time-wise and geo-wise
  correlations are truly measuring different properties of a signal; and the
  claims signals (CHNG-COVID and CHNG-CLI) seem more appropriate for
  temporal---rather than spatial---comparisons.  We revisit this point in the
  discussion.
\item It is still quite impressive and surprising that the CTIS-CLI-in-community
  signal, based on people reporting on the symptoms of others around them, can
  achieve  nearly as strong time-wise correlations to confirmed cases as can a
  signal that is based on picking up the occurrence of a confirmed case passing
  through the outpatient system.
\end{itemize}

\subsection{Helping Robustness}

Public health reporting of COVID tests, cases, deaths, and hospitalizations is
subject to a number of possible delays and problems. For example, COVID testing
data is reported inconsistently by different states using different definitions
and inclusion criteria, and differences in reporting processes mean state data
often does not match data reported to the federal government
\cite{Schechtman:2021}. Case and death data is frequently backlogged and
corrected, resulting in artificial spikes and drops \cite{Simon:2021,
  ArvisaisAnhalt:2021}.

As an example, looking back at Figure \ref{fig:time_trends_national}, we can see
clear dips in the confirmed COVID case curve that occur around the Thanksgiving
and New Year's holidays. This is artificial, and due to the fact that public
health departments usually close over holiday periods, which delays case and
death reporting (for this reason, the artificial dips persist at the state- and
county-level as well). The CLI signal from the survey, on the other hand,
displays no such dips. The claims signals actually display holiday effects going
in the other direction: they exhibit \textit{spikes} around Thanksgiving and New
Year's. This is because they measure the fraction of all outpatient visits with
a certain condition, and the denominator here (total outpatient visits) drops
disproportionately during holiday periods, as people are likely less willing to
go to the doctor for more routine issues. Fortunately, in principle, the holiday
effects in claims signals should be correctable: they are mainly due to
\textit{overall} changes in medical seeking behavior during holiday, periods,
and we can estimate such effects using historical claims data.

As a further example, Figure~\ref{fig:bexar-compare} displays data from Bexar
County, Texas (which contains San Antonio) during July 2020. On July 16, 2020,
4,810 backlogged cases were reported after reporting problems prevented them
from being reported over the past two weeks \cite{Palacios:2021}. This results
in a clearly visible spike in the left-hand panel of the figure (case data from
JHU CSSE, smoothed with a 7-day trailing average).  However, Delphi's COVID
Tracking and Impact Survey averaged around 350 responses per day in Bexar
County over the same time period, and was able to estimate the fraction
of the population who know someone in their community with COVID-Like Illness
(CLI). As we can see in the right-hand panel of the figure, this signal was not
affected by Bexar County's reporting problems and, as shown in the last
subsection, it is (in general) highly correlated with case rates, providing an
alternate stream of data about COVID activity unaffected by backlogs. Similar
reporting problems have occurred in many jurisdictions across the United States,
making it valuable to cross-check against external sources not part of the same
reporting systems.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{fig/bexar_compare.pdf}
  \caption{Reported cases per day in Bexar County, Texas during the summer of
    2020. Only July 16, 4,810 backlogged cases were reported, though they
    actually occurred over the preceding two weeks (this shows up as a prolonged
    spike in the left panel due to the 7-day trailing averaging applied to the
    case counts). Daily CTIS estimates of CLI-in-community showed more stable
    underlying trends.}
  \label{fig:bexar-compare}
\end{figure}

\subsection{Revisions Matter}

The revision tracking feature in the API assists in model-building and
evaluation. Figure~\ref{fig:dv_as_of} illustrates how one COVIDcast medical
claims signal evolved as it was revised across multiple issue dates, in four
different states, between June 1 and August 1, 2020. In each plot, the rightmost
ends of the lines correspond to estimates for the last day that data are
available for each issue date, which are generally the most tentative estimates,
and appear to be significantly biased upward in Arizona in June 2020, and
significantly biased downward in New York throughout June and July 2020.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{fig/dv_as_of.pdf}
\caption{Estimated percentage of doctor's visits due to COVID-like illness
  displayed across multiple issue dates, with later issue dates adding
  additional data and revising past data from prior issue dates.}
\label{fig:dv_as_of}
\end{figure}

Claims-based signals typically undergo heavy backfill as additional claims are
processed and errors are corrected; the median relative error between initial
reports and final values is over 20\% for such data, and only after roughly 35
days do estimates typically match finalized values within 5\%. However, the
systematic nature of this backfill, as illustrated in Figure~\ref{fig:dv_as_of},
suggests that statistical models could be fit (potentially separately for each
location) to estimate the final values from preliminary reports. On the other
hand, official public health reporting of COVID cases and deaths can be subject
to revision as death certificates are audited and backlogs cleared, resulting in
thousands of cases and deaths being added or removed. This process is much more
difficult to predict, and thus claims data and other sources may be a useful
stand-in while public health reports are aggregated and corrected.

To reiterate a previous point, when training and validating forecast models (on
historical data), users will want to use data that was known \textit{as of} the
forecast date, not revised versions that only became available much later. The
COVIDcast API makes all historical versions available and easily accessible for
this purpose; and this feature plays a prominent role in our own analysis of
forecasting and hotspot prediction models appearing in a companion paper.

\subsection{New Perspectives}

Auxiliary signals (outside of the standard public health reporting streams) can
serve as indicators of COVID activity, but they can also illuminate other
aspects of the pandemic. To understand the effect of the pandemic on society and
predict its future spread requires a comprehensive view of not just cases and
deaths, but public behavior and beliefs. Measures of mobility reflect social
distancing policies and the potential rate of COVID spread; medical claims data
reflects healthcare-seeking behavior and how it may be affected by local case
rates; measures of COVID vaccine acceptance can guide outreach efforts
\cite{TODO}.

% AR: Potentially interesting things to show or comment on:
% - Medical claims data can give info about medical-seeking behavior and how it
% varies with caseload -- e.g. are people deterred from seeking medical care
% because of the high case load?
% - Mobility changes
% - Map of one of the unique signals during an outbreak, e.g. Michigan in April
% 2021

% RJT: @AR Since we already stuff on claims, can we do: mobility, masks, vaccine
% uptake or hesitancy, distancing behaviors, how worried are you?

% How about this (in each case, each point in said plot is a county's signal
% values averaged over some period)
% - Bivariate plot of mask use versus how worried are you
% - Bivariate plot of of how worried are you vs case rates
%- Bivariate plot of of how worried are you vs hospitalizations
% I would be curious to see all three.  And perhaps color code the points by
% region to highlight regional trends

\section{Discussion}

% AR: I think our discussion should also make an argument: that providing many
% data sources in a common format, with revision tracking, is a valuable goal.
% Each of our individual data sources has various limitations -- SafeGraph has
% various biases, the claims data is limited by geographically varying market
% share, the surveys have demographic biases, etc. So if the argument is solely
% "We have better data than everyone else," it's a tough sell; if the argument
% is that our data being convenient to use and accessible in a single unified
% way makes it more valuable than the individual parts, and that other
% epidemiological data should be available the same way, then I think we have a
% strong case.
% RJT: I agree

% RJT Proposed format for this discussion.
% 1. COVID indicators can help supplement public health reports by filling out
% the severity pyramid, improving timeliness, helping with robustness, etc.
% 2. They also provide info well beyond public health reports (behaviors,
% attitudes)
% 3. There is an argument to be made for having data all in one unified
% convenient access format (with all revisions)

% There are a number of open questions, and challenges to be recognized.  We
% should be clear in identifying survey bias; geo bias in claims data; and
% backfill and holiday corrections in claims data.

% Finish with list of iniatives our API can power (like we have now, but
% compress it quite a lot.)

% Signals in the COVIDcast API indicating amounts of COVID-19 and
% COVID-like illness supplement publicly available case and death data by
% capturing and focusing on different levels of severity and parts of the
% timelines of COVID-19 infections. For example, the survey-based COVID-like
% illness indicators may reflect mild or recently-developed COVID-19 infections
% before or without a visit to a healthcare provider or COVID-19
% test. Additionally, the reporting pipeline for various digital signals can be
% quite short. For example, many survey signals are released the morning after the
% surveys were taken, and describe symptoms experienced just one to two days in
% the past; initial estimates of outpatient visits and inpatient admissions are
% released after a few days, generally within a week. The timeliness of these
% signals may enable users to more quickly detect and predict changes in the
% direction of COVID-19 activity.

% TODO move backfill and correlation discussion here?
%
% Discuss other applications of the data:
% - Early warning (informal: just call a human to look at things when the
%   indicators blow up)
% - Hotspot prediction (formal: build a model to predict a surge)
% - Forecasting (refer out to the paper we're writing in parallel)
% - Nowcasting (i.e., estimating latent infection curve)
% - Reopening criteria (refer to COVID Exit Strategy?)
%
% COVIDcast facilitates several types of real-time applications and retrospective
% analyses.
% \begin{itemize}
% \item
% \emph{Situational awareness dashboards and reports}
% summarize activity levels and outcomes, public behavior, vaccination rates, and
% other relevant statistics.
% %
% Features of the COVIDcast API and \texttt{covidcast} packages that are
% particularly relevant here are the common data format across all signals and the
% ability to easily request data for individual locations and for less-commonly
% used geographical resolutions such as Hospital Referral Regions and metropolitan
% statistical areas, and for newer types of data such as vaccination status.
% %
% \item
% \emph{Early warning} systems flag fast rises in COVID-19 activity indicators;
% formal variants first define what counts as elevated activity or surges and
% perform statistical \emph{hotspot predictions}.
% %
% For individuals and local-level response, signals with sufficient availability
% and high time-wise correlations with an outcome of interest (e.g., official case
% or hospitalization data) are natural choices to monitor.
% %
% For resource allocation across locations or drug trial site selection, natural
% choices would be indicators directly describing the stage of the disease being
% targeted or signals with high geo-wise correlations to these indicators.
% %
% \item
% \emph{Forecasts} directly predict COVID-19 activity levels using point estimates
% or probability distributions.
% %
% These models may operate at fine geographical and temporal resolutions and/or
% draw on a wide variety of data, making a common data format and signals offered
% by county and by date particularly useful.
% %
% Revision tracking is especially useful here, helping to explain anomalies in
% past forecasts and improve future ones.
% %
% \item
% \emph{Nowcasts} estimate current or near-future COVID-19 activity levels or the
% true number of infections.
% %
% For these applications, it is important to have signals with time
% values that relate to when symptoms started or were present, or when a healthcare
% visit occurred, rather than to the time that these values were eventually
% reported, unless there is a fixed reporting delay or auxiliary data regarding the
% distribution of reporting delays.
% %
% For example, time values for survey data in COVIDcast relates to when symptoms
% were occurring, doctor visit data indicates the date of the visit, and inpatient
% data indicates the date of admission.
% %
% This contrasts with nationwide aggregated case and death data, which is
% usually aligned by the date that a case or death was reported and subject to
% variable delays based on testing availability and location-specific reporting pipelines.
% %
% % TODO antibodies, revisions, positivity
% \item
% \emph{Reopening criteria} are informed by COVID-19 activity levels, public
% behavior, and vaccination levels.  Methodologies to develop and evaluate these
% criteria may require such information, as well as additional data on timing of
% public health interventions and on confounding variables, in order to draw valid
% conclusions or to improve estimated effects of potential policies.
% \end{itemize}

% TODO refer out to the paper we're writing in parallel

\matmethods{Please describe your materials and methods here. This can be more than one paragraph, and may contain subsections and equations as required.

\subsection*{Subsection for Method}
Example text for subsection.
}

\showmatmethods{} % Display the Materials and Methods section

\acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}

\showacknow{} % Display the acknowledgments section

% Bibliography
\bibliography{../../common/covidcast.bib}

\end{document}

### 2021-06-12 indicators/paper/paper.tex
Small language tweak

@@
 
-For many purposes it is sufficient to know the most recently issued observation
-for any time value, and the COVIDcast API reports the most recent issue by
-default. However, for some purposes it is crucial to know what was known
+For many purposes it is sufficient to use the most recently issued observation
+for any time value, and thus the COVIDcast API reports the most recent issue by
+default. However, for some applications it is crucial to know what was known
 \textit{as of} a specific date. For example, an epidemic forecasting model will

### 2021-06-12 indicators/paper/paper.tex
Sig statement is exactly 120 words

@@
 % Please add a significance statement to explain the relevance of your work
-\significancestatement{To properly study the COVID-19 pandemic, its effects on
-  society, and measures for reducing its spread, researchers need detailed data
-  on the course of the pandemic. Standard public health data streams suffer
-  inconsistent reporting and frequent, unexpected revisions. They also miss some
-  important aspects worthy of study, such as certain aspects of public
-  behavior. We present an open database of COVID signals in the United States,
-  measured primarily at the county level and updated daily. These include
-  traditionally reported COVID cases and deaths, but also many others: signals
-  based on measures of mobility, social distancing behaviors, internet search
-  trends, self-reported symptoms, and patterns COVID-related activity in
-  de-identified medical insurance claims. The database provides all signals in a
-  common, easy-to-use format, empowering both public health research and
-  operational decision-making.
+\significancestatement{To study the COVID-19 pandemic, its effects on society,
+  and measures for reducing its spread, researchers need detailed data on the
+  course of the pandemic. Standard public health data streams suffer
+  inconsistent reporting and frequent, unexpected revisions. They also miss
+  other aspects of a population's behavior, worthy of consideration. We
+  present an open database of COVID signals in the United States, measured at
+  the county level and updated daily. These include traditionally reported COVID
+  cases and deaths, and many others: signals based on measures of mobility,
+  social distancing, internet search trends, self-reported symptoms, and
+  patterns COVID-related activity in de-identified medical insurance claims. The
+  database provides all signals in a common, easy-to-use format, empowering both
+  public health research and operational decision-making.
 }

### 2021-06-12 forecast/paper/paper.tex
Edits up through intro

@@
 % Please add a significance statement to explain the relevance of your work
-\significancestatement{Authors must submit a 120-word maximum statement about
-  the significance of their research paper written at a level understandable to
-  an undergraduate educated scientist outside their field of speciality. The
-  primary goal of the significance statement is to explain the relevance of the
-  work in broad context to a broad readership. The significance statement
-  appears in the paper itself and is required for all research papers.}
+% TODO write one. This is the indicator's one.
+\significancestatement{To study the COVID-19 pandemic, its effects on society,
+  and measures for reducing its spread, researchers need detailed data on the
+  course of the pandemic. Standard public health data streams suffer
+  inconsistent reporting and frequent, unexpected revisions. They also miss
+  other aspects of a population's behavior, worthy of consideration. We
+  present an open database of COVID signals in the United States, measured at
+  the county level and updated daily. These include traditionally reported COVID
+  cases and deaths, and many others: signals based on measures of mobility,
+  social distancing, internet search trends, self-reported symptoms, and
+  patterns COVID-related activity in de-identified medical insurance claims. The
+  database provides all signals in a common, easy-to-use format, empowering both
+  public health research and operational decision-making.
+}
 
@@
 convenient data fetching and processing tools. In the current paper, we aim to
-quantify the additional utility provided by a core set of these indicators for
-two fundamental prediction tasks: probabilistic forecasting of COVID-19 case
-rates and prediction of future COVID-19 hotspots (where a hotspot is defined by
-the event that the relative increase in COVID-19 cases in a given location
-exceeds a certain threshold).
+quantify the utility provided by a core set of these indicators for two
+fundamental prediction tasks: probabilistic forecasting of COVID-19 case 
+rates and prediction of future COVID-19 hotspots (where we define a hotspot by a
+relative increase in COVID-19 cases, to be spelled out precisely later).
 
@@
   prediction} is focused on a much higher standard than simply asking 
-about correlations; to be useful in forecasting or hotspot prediction, an 
-indicator must provide relevant information that is not otherwise contained in 
-past values of the case rate itself. We address this in the most direct way
-possible, by inpsecting the difference in predictive performance of simple 
+about correlations; to be useful in forecast or hotspot models, an indicator
+must provide relevant information that is not otherwise contained in past values
+of the case rate series itself. We will assess this in the most direct way
+possible: by inpsecting the difference in predictive performance of simple
 autoregressive models trained with and without access to past values of a
@@
 
-There is a second and more subtle issue in evaluating predictive utility that
-deserves explicit mention, as it will play a key role in our
-analysis. Indicators computed from surveillance streams will often be subject to
+There is another, more subtle issue in evaluating predictive utility that 
+deserves explicit mention, as it will play a key role in our analysis.
+Signals computed from surveillance streams will often be subject to  
 latency and/or revision. For example, a signal based on aggregated medical
@@
 substantially revised over the next several weeks as additional claims are
-submitted and processed late. Correlations between such a signal and case rates 
-calculated ``after the fact'' (meaning, we look retrospectively and use the
-finalized values of this indicator) will not deliver an honest answer to the
-question of whether this indicator would have been useful in real time. Instead, 
+submitted and/or processed late. Correlations between such a signal and case
+rates calculated ``after the fact'' (computed retrospectively, using the
+finalized values of this signal) will not deliver an honest answer to the    
+question of whether this signal would have been useful in real time. Instead,
 we build predictive models using only the data that would have been available
@@
 
-Furthermore, adopting this practically-oriented forecasting
-perspective to indicator evaluation requires one to confront
-fundamental limitations inherent to public health surveillance
-systems.  In particular, many reporting systems are plagued by
-latency, with delays in reporting that can sometimes span
-multiple weeks.  Latency manifests itself in different ways in different data sources.  Some data sources may only make a signal available one week after the date it describes; other
-data sources may provide immediate imperfect initial estimates, which are
-gradually ``backfilled'' over a period of days, weeks, or months (in which previous estimates are updated as new
-data trickle in).  And few of
-these latency patterns are stable or predictable, with holidays,
-extreme weather events, and changes in the reporting process all
-affecting the dynamics of the delay process.  To fully and accurately
-understand an indicator's usefulness for forecasting requires
-accounting for the unfortunate reality of data latency.
+% Furthermore, adopting this practically-oriented forecasting
+% perspective to indicator evaluation requires one to confront
+% fundamental limitations inherent to public health surveillance
+% systems.  In particular, many reporting systems are plagued by
+% latency, with delays in reporting that can sometimes span
+% multiple weeks.  Latency manifests itself in different ways in different data sources.  Some data sources may only make a signal available one week after the date it describes; other
+% data sources may provide immediate imperfect initial estimates, which are
+% gradually ``backfilled'' over a period of days, weeks, or months (in which previous estimates are updated as new
+% data trickle in).  And few of
+% these latency patterns are stable or predictable, with holidays,
+% extreme weather events, and changes in the reporting process all
+% affecting the dynamics of the delay process.  To fully and accurately
+% understand an indicator's usefulness for forecasting requires
+% accounting for the unfortunate reality of data latency.
 
-Examining the importance of additional features for prediction is a core question
-in inferential statistics, and there are many ways to examine this question
-under a variety of assumptions. For time series, work in
-this area can be traced at least to Granger causality~\cite{GrangerCausality} and
-related multivariate extensions~\cite{GewekeTimeSeries} under the condition 
-that the series are linearly related autoregressions among other assumptions.
-These techniques are especially popular in econometrics and neuroscience~\cite{mccracken2007asymptotics,NeuroscienceGranger}.
+% Examining the importance of additional features for prediction is a core question
+% in inferential statistics, and there are many ways to examine this question
+% under a variety of assumptions. For time series, work in
+% this area can be traced at least to Granger causality~\cite{GrangerCausality} and
+% related multivariate extensions~\cite{GewekeTimeSeries} under the condition 
+% that the series are linearly related autoregressions among other assumptions.
+% These techniques are especially popular in econometrics and neuroscience~\cite{mccracken2007asymptotics,NeuroscienceGranger}.
 
-We choose to take a
-predictive angle, which is simple but also fairly robust, compared to various
-other classical inferential approaches. In fact, drawing rigorous inference based on
-predictions, without (or with lean) assumptions, is an active field of
-research~\cite{dai2021significance,fryer2021shapley,zhang2020floodgate,williamson2020unified}.
-Nonetheless, the predictive approach directly examines the target of interest: do
-these signals help to produce forecasts? Furthermore, directly comparing
-forecast errors without strong assumptions through tests of predictive accuracy
-are common in epidemiology, econometrics, and other
-fields~\cite{DieboldMarianoTest,DieboldRetrospective}.
+% We choose to take a
+% predictive angle, which is simple but also fairly robust, compared to various
+% other classical inferential approaches. In fact, drawing rigorous inference based on
+% predictions, without (or with lean) assumptions, is an active field of
+% research~\cite{dai2021significance,fryer2021shapley,zhang2020floodgate,williamson2020unified}.
+% Nonetheless, the predictive approach directly examines the target of interest: do
+% these signals help to produce forecasts? Furthermore, directly comparing
+% forecast errors without strong assumptions through tests of predictive accuracy
+% are common in epidemiology, econometrics, and other
+% fields~\cite{DieboldMarianoTest,DieboldRetrospective}.
 
@@
 % series.
-
-% RJT: 
-% What is a leading indicator?
-% Conceptual vs empirical.
 

### 2021-06-12 forecast/paper/paper.tex
Edits up through methods A

@@
 
-\title{Can COVID-19 Indicators Improve Forecasts and Hotspot Prediction?}
+\title{Can Auxiliary Indicators Improve COVID-19 Forecasts and Hotspot
+  Prediction?}
 
@@
 The indicators we focus on provide information not generally available from
-standard public health reporting. The COVID Tracking and Impact Survey (CTIS),
-run by the Delphi group in partnership with Facebook, reaches roughly 50,000
-respondents across the United States daily, and provides information on
-symptoms, social distancing, and other attitudes and behaviors related to
-COVID. The particular indicator we examine, which we refer to as \fb, is the
-estimated percentage of people who know someone with COVID-like illness in their
-the local community.
+standard public health reporting. Among the many available in Delphi's COVIDcast
+API, we consider the following five auxiliary indicators: 
+\begin{itemize}
+\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
+  outpatient visits that are primarily about COVID-related symptoms, based on
+  de-identified Change Healthcare claims data.
+\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
+  with confirmed COVID-19, based on the same claims data.
+\item COVID Tracking and Impact Survey CLI-in-community (CTIS-CLI-in-community): 
+  The estimated percentage of the population who knows someone who is sick,
+  based on Delphi's surveys of Facebook users.
+\item Doctor Visits COVID-like illness (DV-CLI): The same as CHNG-CLI, but
+  computed based on de-identified medical insurance claims from other health
+  systems partners.  
+\item Google Search Trends for anosmia and ageusia (Google-AA): A measure of 
+  COVID-19 related Google search volume for queries related to anosmia (loss of
+  smell), ageusia (loss of taste), or both.  
+\end{itemize}
+In short, we choose these indicators because, conceptually speaking, they
+measure aspects of an individual's disease progression that would plausibly
+precede the occurence of (at worst, co-occur  with) the report of a positive
+COVID-19 test, through standard public health reporting streams.
 
-In some ways, this can be seen as a
-proxy for the strength of social contacts between infected and susceptible
-populations, a key feature for disease modeling~\cite{Zhang1481} The Google
-indicator, which we refer to as \gs, tracks the local intensity of Google searches related to lack-of-smell
-(anosmia) or lack-of-taste (ageusia), both of which have been connected to
-COVID19~\cite{KlopfenAnosmia,vaira2020anosmia}. Two
-indicators are derived from de-identified insurance claims data from Change Healthcare, a large
-healthcare technology company that aggregates data from many providers. The
-first signal, which we refer to as \chngcov, is the estimated percentage of outpatient doctor visits with
-confirmed COVID-19 diagnosis (potentially collected earlier than the related
-report to the local health authority), while the second, which we
-refer to as \chngcli, is the estimated percentage of outpatient
-visits with COVID-like illness. Finally, \attn{describe \dv~indicator, perhaps best described in conjunction with \chngcov}.  The hope is that each
-of these signals serves as a ``leading indicator'' of future COVID cases,
-improving the forecastibility of COVID cases and hotspots. For more details on
-these and other indicators, see the companion paper.
+For more details on the five indicators (including how these are precisely
+computed from the underlying data streams) we refer to
+\url{https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html},
+which documents all of the COVIDcast signals, as well our companion paper on 
+COVIDcast API and database. For CTIS in particular, we refer to our companion 
+paper on this survey. For the Google Search Trends data set, see
+\cite{GoogleSymptoms}, and \cite{Klopfen:2020, Vaira:2020} for articles on the
+relevance of anosmia and ageusia.
 
-% RJT: We should refer to the companion (indicators) paper for more details.
-% Also, I'm wondering what the point of the table (below) is?  Does it offer new
-% info in addition to the paragraph?  If it's just a nice visual cue, then we
-% can turn the above paragraph into a bullet point list, one bullet per signal,
-% and cut the table. @DM/JB: Let me know what you think.
-% DJM: agreed, table gone for now. I'll leave the bullet decision for later.
+As for geographic resolution, we consider the prediction of COVID-19 case rates
+and hotspots (to be precisely defined shortly) aggregated at the level of an
+individual hospital referral region (HRR). HRRs correspond to groups of counties
+in the United States within the same hospital referral system. The Dartmouth
+Atlas of Healthcare Policy~\cite{DartmouthHRR}, defines these 306 regions based
+on a number of characteristics. They are contiguous regions containing at least
+one city where major procedures (cardiovascular or neurological) are
+performed. Each HRR has a minimum population of about 120,000, and the majority
+of hospital services for that population are performed by hospitals within the
+region. Some HRRs are quite large (such as the one containing Los Angeles, which
+has about 10 million people), but generally HRRs are much more homogenous in
+size than the (approximately) 3200 counties are, and offer a nice middle ground
+in between counties and states.
 
-% \begin{table}
-% \begin{tabular}{@{}p{1in}p{1in}p{1in}p{3in}@{}}
-%  \toprule
-%     Data source &   Signal name &   Abbreviation    &  
-%         Description\cite{DelphiAPI}\\
-%  \midrule
-%    \attn{COVIDcast Survey via Facebook} &   Smoothed CLI-in-community & FBCLI &
-%         Estimated percentage of people reporting illness in their local community,
-%         including their household   \\
-%     Change Healthcare   &   Smoothed adjusted oupatient CLI  &   CHCLI   &
-%         Estimated percentage of outpatient doctor visits primarily about
-%         COVID-related symptoms, with day-of-week effects removed  \\
-%     Change Healthcare   &   Smoothed adjusted oupatient COVID  &   CHCOV &
-%         Estimated percentage of outpatient doctor visits with confirmed
-%         COVID-19, with day-of-week effects removed  \\
-%     Google Symptoms     & Smoothed sum of anosmia and ageusia   &   GSSAA   &
-%         The sum of Google search volume for anosmia and ageusia related searches,
-%         in an arbitrary units that are normalized for overall search users\\
-%  \bottomrule
-% \end{tabular}
-% \caption{Syndromic signals available in COVIDcast.\label{tab:signals-used}}
-% \end{table}
-
-% RJT: We need to explain WHY HRRs.  This is not familiar to basically anybody!
-% It's an invention of academics who work in healthcare policy, from Dartmouth.
-% We should cite them: https://www.dartmouthatlas.org/faq/.  Here's the
-% rationale for us using them: 
-% - They are groups of counties driven by the same hospital referral system.
-% - They are somewhat homogenous population-wise.  Don't have a long tail of 
-% many HRRs with small populations (like counties).  We should check this!  And
-% could even the distribution vs the distribution with respect to counties (or
-% some other comparison point). 
-% - They would be more relevant to **hospitalization forecasting** given their
-% definition, but our indicators should provide bigger boosts to case forecasts.
-% (Hospitalizations | cases shouldn't have much dependence with indicators.
-% Here I'm using this as in X | Y, i.e., X conditional on Y).
-% - But to the extent that case projections inform hospitalization projects,
-% this is still quite relevant & sensible to use HRRs. 
-% @DM/JB: Please take a stab at this.
-% DJM: stabbed.
-
-As for geographic resolution, we consider prediction of COVID case rates (and
-hotspots) aggregated at the level of an individual hospital referral region
-(HRR). HRRs correspond to groups of counties within the same hospital referral
-system. The Dartmouth Atlas of Healthcare Policy~\cite{DartmouthHRR}, defines
-these 306 regions based on a number of characteristics. They are contiguous
-regions containing at least one city where major cardiovascular or neurological
-precedures are performed. Each HRR has a minimum population of about 120,000, and
-the majority of hospital services for that population are performed by hospitals
-within the region. Some HRRs are quite large (Los Angeles has
-about 10 million), but generally HRRs are much more homogenous than the
-$\sim$3100 counties 
-would be but at much higher resolution than states. They would be more relevant for
-hospitalization forecasting, given their definition. We have chosen to focus on
-forecasting cases at this resolution because the indicators should provide a
-bigger boost. Once we observe case information, there should not be much
-dependence of future
-hospitalizations on the indicators, but predicting cases should provide more timely
-warnings of future hospitalization needs as opposed to waiting for the case
-numbers.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{fig/hrr_pop_histogram.pdf}
-  \caption{Histogram of population size for HRR. Note that the $x$-axis is on
-    the log scale.}
-  \label{fig:hrr-histogram}
-\end{figure}
+HRRs, by their definition, would be most relevant for hospitalization
+forecasting.  We have chosen to focus on forecasting cases at this resolution 
+because the indicators that we consider should intuitively be more useful in
+latter task.  Predicting case rates and case hotspots at the HRR level is still
+a reasonable goal, especially predicted case information could then be fed
+downstream into hospitalization forecasting models.
 

### 2021-06-12 forecast/paper/paper.tex
Edits almost thru methods B

@@
 
-\dropcap{T}racking and forecasting indicators from public health reporting1
+\dropcap{T}racking and forecasting indicators from public health reporting
 streams---such as confirmed cases and deaths in the COVID-19 pandemic---is
@@
 fundamental prediction tasks: probabilistic forecasting of COVID-19 case 
-rates and prediction of future COVID-19 hotspots (where we define a hotspot by a
-relative increase in COVID-19 cases, to be spelled out precisely later).
+rates and prediction of future COVID-19 case hotspots (defined by the event that
+a relative increase in COVID-19 cases exceeds a certain threshold).
 
-At the outset, we should be clear that our intent in this paper is \emph{not} to
-provide an authoritative take on cutting-edge COVID-19 forecasting methods.
+At the outset, we should be clear that our intent in this paper is \textit{not}
+to provide an authoritative take on cutting-edge COVID-19 forecasting methods.
 Instead, our purpose here is to provide a rigorous, quantitative assessment of
@@
 case rates several days in the future) move us in the right direction, but still
-do not move us all the way there. The question about \emph{utility for
+do not move us all the way there. The question about \textit{utility for 
   prediction} is focused on a much higher standard than simply asking 
@@
 we build predictive models using only the data that would have been available
-\emph{as of} the prediction date, and compare the ensuing predictions in terms
-of accuracy. This leverages our group's {\tt evalcast} R package, which plugs
-into the COVIDcast API's data versioning system, and facilitates honest
-backtesting of forecasters.
+\textit{as of} the prediction date, and compare the ensuing predictions in terms 
+of accuracy. To do so, we leverage Delphi's \texttt{evalcast} R package 
+\cite{EvalcastR}, which plugs into the COVIDcast API's data versioning system,  
+and facilitates honest backtesting.
 
@@
 
+We consider prediction of future COVID-19 case rates or case hotspots (to be
+defined precisely shortly).  By case rate, we mean the case count per 100,000
+people (the standard in epidemiology).  We use reported case data aggregated by 
+JHU CSSE \cite{Dong:2020}, which is accessible through the COVIDcast API, like
+the auxiliary indicators that we use to supplement the basic purely
+autoregressive models.
+
 The indicators we focus on provide information not generally available from
-standard public health reporting. Among the many available in Delphi's COVIDcast
-API, we consider the following five auxiliary indicators: 
+standard public health reporting. Among the many indicators in the API, we
+consider the following five:
 \begin{itemize}
@@
 As for geographic resolution, we consider the prediction of COVID-19 case rates
-and hotspots (to be precisely defined shortly) aggregated at the level of an
-individual hospital referral region (HRR). HRRs correspond to groups of counties
-in the United States within the same hospital referral system. The Dartmouth
-Atlas of Healthcare Policy~\cite{DartmouthHRR}, defines these 306 regions based
-on a number of characteristics. They are contiguous regions containing at least
-one city where major procedures (cardiovascular or neurological) are
-performed. Each HRR has a minimum population of about 120,000, and the majority
-of hospital services for that population are performed by hospitals within the
-region. Some HRRs are quite large (such as the one containing Los Angeles, which
-has about 10 million people), but generally HRRs are much more homogenous in
-size than the (approximately) 3200 counties are, and offer a nice middle ground
-in between counties and states.
+and hotspots aggregated at the level of an individual \textit{hospital referral
+  region} (HRR). HRRs correspond to groups of counties in the United States
+within the same hospital referral system. The Dartmouth Atlas of Healthcare 
+Policy~\cite{DartmouthHRR}, defines these 306 regions based on a number of
+characteristics. They are contiguous regions containing at least one city where
+major procedures (cardiovascular or neurological) are performed.  Most hospital
+services for the population in an HRR are performed by hospitals within the
+region.  Each HRR has a minimum population of about 120,000; some HRRs are quite
+large (such as the one containing Los Angeles, which has about 10 million
+people), but generally HRRs are much more homogenous in size than the
+(approximately) 3200 counties are, and they serve as a nice middle ground in 
+between counties and states.
 
-HRRs, by their definition, would be most relevant for hospitalization
-forecasting.  We have chosen to focus on forecasting cases at this resolution 
-because the indicators that we consider should intuitively be more useful in
-latter task.  Predicting case rates and case hotspots at the HRR level is still
-a reasonable goal, especially predicted case information could then be fed
-downstream into hospitalization forecasting models.
+HRRs, by their definition, would be most relevant for forecasting hospital
+admissions.  We have chosen to focus on cases (forecasting and predicting
+hotspots) at this resolution since the indicators considered should be more
+useful in predicting case versus hospitalization activity, the latter being
+intuitively more closely connected (in time) to the events that are measured by
+the given five indicators. Predicting case rates (and hotspots) at the HRR level
+is still a reasonable goal in its own right; and moreover, it could be used to
+feed predicted case information into downstream hospitalization models.
 
@@
 
-% RJT: We need to fill this out.  I'm convinced we need to promote "as of"
-% training as our main point of study.  We need to explain why this is so
-% important.  And we can and should present results with finalized data in the
-% appendix.  I guess we can wait until we get the experiments in to fill out
-% this section, because it might influence the prose we use here, or we could
-% also start writing this right away. And we can refer to the companion
-% (indicators) paper, which also describes the revision tracking feature in the
-% API.  @DM/JB: Your call, whether to start writing this or wait.
-% DJM: drafted
+In this paper, all models are fit using only ``vintage'' training data. This
+means that for a given prediction date, say September 1, 2020, we train models
+using only the data that would have been available to us \textit{as of}
+September 1.  Essentially, we ``rewind'' the clock to September 1, and query the
+COVIDcast API to get the latest data that it would have had available at
+that point in time.  This is made possible by the COVIDcast API's comprehensive
+data versioning system (described in more detail in our companion paper).  We
+also use the \texttt{evalcast} R package \cite{EvalcastR}, which streamlines the
+process of training arbitrary prediction models over a sequence of prediction
+dates, by automatically constructing the (properly-versioned) sequence of
+vintage training data sets.
 
-In this manuscript, all models are estimated using only ``vintage training
-data''. This means that for a given forecast date, say November 1st 2020, we
-estimate our models using only the data that would have been available to us on
-November 1st. Essentially, the training data
-is rewound appropriately to dates in the past. It is likely, for example, that
-no data would actually exist 
-for the forecast date on the forecast date because there is some latency between
-the time signals are reported and the dates for which they are reported.
-Similarly, any future revisions made to the indicators or case numbers would
-not have been present on the forecast date, and so would not be available to the
-forecaster. Figure~\ref{fig:vintage} shows three signals for the Los Angeles HRR
-with a hypothetical forecast date of November 1. On November 1, no signals are
-available for October 31. The cases data is one day latent while the Facebook
-survey is 2 days latent. Doctor's visits are 3 days latent while also exhibiting
-revisions.  
-By training our models in this way, the evaluations are more realistic in the sense
-that we are accurately mimicking the difficulty faced by real-time epidemic
-forecasters. The COVIDcast API maintains all versions of the data, thereby allowing
-these sorts of synthetic forecasting exercises. For more details about data
-versioning, see the companion paper.
+Vintage training data means different things, in practice, for different
+signals. The three signals based on medical claims, CHNG-CLI, CHNG-COVID, and
+DV-CLI, are most often 3-5 days latent, and subject to a considerable degree of
+revision or ``backfill'' after their initial publication date.  The survey-based
+signal, CTIS-CLI-in-community, is 2 days latent, and rarely undergoes any 
+revision at all.  The target itself, reported case rates (or a binary hotspot
+variable, defined based on reported case rates), is 1 day latent, and exhibits
+frequent, unpredictable revisions (sometimes clear anomalies) after initial
+publication.  Compared to the revision pattern in the medical claims signals,
+which are appears to be much more systematic, the revision pattern in the case
+rates signal can be highly erratic, resulting in big spikes as reporting
+backlogs are cleared, changes in case definitions are made, other corrections
+are applied, etc.
 
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{fig/vintage.pdf}
-  \caption{Changes in availability and revisions for three signals in the Los
-    Angeles HRR. The vertical black line indicates a hypothetical forecast date
-    of November 1 while the colored lines get progressively darker as we collect
-  later and later vintages of data.}
-  \label{fig:vintage}
-\end{figure}
+TODO What is ground truth??
+TODO Better example of backfill in DV-CLI?
 
@@
 %\subsubsection{Forecasting Commonalities} % PNAS wants Title Caps for everything
-
-% RJT: Just a quick few quick comments for now on the text below.  
-% Also, shouldn't we
-% point out it's relationship to quantile score, and give it's equivalent form
-% there?
-% DJM: done.
 
@@
 \section{Results}
+
+Code to reproduce all examples (which uses the
+\texttt{evalcast} R package and fetches data directly from the API) can be
+found at \url{TODO}.
 

### 2021-06-13 indicators/paper/paper.tex
Title tweak?
Bunch of keywords
Add repo URL, small tweaks

@@
 
-\title{An Open-Source Repository of Real-Time COVID-19 Indicators}
+\title{An Open Repository of Real-Time COVID-19 Indicators}
 
@@
 % At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
-\keywords{open data $|$ COVID-19 $|$ Keyword 3 $|$ ...}
+\keywords{open data $|$ COVID-19 $|$ digital surveillance $|$ medical insurance
+  claims $|$ internet surveys $|$ search trends $|$ mobility patterns}
 
@@
 For many purposes it is sufficient to use the most recently issued observation
-for any time value, and thus the COVIDcast API reports the most recent issue by
-default. However, for some applications it is crucial to know what was known
-\textit{as of} a specific date. For example, an epidemic forecasting model will
-be called upon to make its forecasts based on preliminary data about recent
-trends, and so when it is trained using historical data, it should be trained
-using the initial versions of that data, not updates that would have been
-received later. Furthermore, these revision records allow models to be modified
-to account for noise and bias in early data versions, or to exclude data that is
-too new to be considered stable, and to ``rewind'' time and simulate how these
-revised models would have performed using only the versions of data available
-\textit{as of} those times.
+at any a given time value, and the COVIDcast API returns the most recent issue
+as its default. However, for some applications it is crucial to know what was
+known \textit{as of} a specific date. For example, an epidemic forecasting model
+will be called upon to make its forecasts based on preliminary data about recent
+trends, so when it is trained using historical data, it should be trained using
+the initial versions of that data, not updates that would have been received
+later. Moreover, these revision records allow models to be modified to account
+for noise and bias in early data versions, or to exclude data that is too new to
+be considered stable, and to ``rewind'' time and simulate how these revised
+models would have performed using only the versions of data available \textit{as
+of} those times.
 
@@
 that backfill can significantly alter forecast performance \cite{Brooks:2018,
-  Reich:2019}, and that training to use preliminary data can reduce this
-influence \cite{Brooks:2018, Brooks:2020}. This is also investigated in our
-companion paper on forecasting, where we observe that training and validating
-models on finalized data yields overly optimistic estimates of true test-time
-performance.
+  Reich:2019}, and that careful training on preliminary data can reduce this
+influence \cite{Brooks:2020}. We also investigate this in our companion paper on
+forecasting, where we observe that training and validating models on finalized 
+data yields overly optimistic estimates of true test-time performance.
 
@@
 \texttt{covidcast} R package and fetches data directly from the API) can be
-found at \url{TODO}.
+found at
+\url{https://www.github.com/cmu-delphi/covidcast-pnas/indicators/code/}.
 

### 2021-06-13 forecast/paper/paper.tex
Edits up thru methods D (formerly C)

@@
 % Please add a significance statement to explain the relevance of your work
-% TODO write one. This is the indicator's one.
+% TODO write a real one. This is the indicator's one.
 \significancestatement{To study the COVID-19 pandemic, its effects on society,
@@
 a linear combination of COVID cases in the near past), and ask whether the 
-inclusion of an auxiliary indicator as an exogenous feature in such a model
+inclusion of an auxiliary indicator as an additional feature in such a model
 improves predictive power. While forecasting carries a rich literature that
@@
 \url{https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html},
-which documents all of the COVIDcast signals, as well our companion paper on 
-COVIDcast API and database. For CTIS in particular, we refer to our companion 
-paper on this survey. For the Google Search Trends data set, see
+which documents all of the signals in the COVIDcast API, and our companion paper
+on the API and database. For CTIS in particular, we refer to our companion paper
+on this survey. For the Google Search Trends data set, see
 \cite{GoogleSymptoms}, and \cite{Klopfen:2020, Vaira:2020} for articles on the
-relevance of anosmia and ageusia.
+relevance of anosmia or ageusia (loss of smell or taste) to COVID.
 
@@
 Policy~\cite{DartmouthHRR}, defines these 306 regions based on a number of
-characteristics. They are contiguous regions containing at least one city where
-major procedures (cardiovascular or neurological) are performed.  Most hospital
-services for the population in an HRR are performed by hospitals within the
-region.  Each HRR has a minimum population of about 120,000; some HRRs are quite
-large (such as the one containing Los Angeles, which has about 10 million
-people), but generally HRRs are much more homogenous in size than the
-(approximately) 3200 counties are, and they serve as a nice middle ground in 
-between counties and states.
+characteristics. They are contiguous regions such that most of the hospital
+services for the underlying population are performed by hospitals within the
+region. Each HRR also contains at least one city where major procedures
+(cardiovascular or neurological) are performed. The smallest HRR has a
+population of about 120,000. While some are quite large (such as the one
+containing Los Angeles, which has about 10 million people), generally HRRs 
+are much more homogenous in size than the (approximately) 3200 counties,  
+and they serve as a nice middle ground in between counties and states.
 
 HRRs, by their definition, would be most relevant for forecasting hospital
-admissions.  We have chosen to focus on cases (forecasting and predicting
-hotspots) at this resolution since the indicators considered should be more
-useful in predicting case versus hospitalization activity, the latter being
+demand.  We have chosen to focus on cases (forecasting and predicting
+hotspots) at the HRR level since the indicators considered should be more 
+useful in predicting case activity versus hospital demand, since the latter is
 intuitively more closely connected (in time) to the events that are measured by
@@
 
-In this paper, all models are fit using only ``vintage'' training data. This
-means that for a given prediction date, say September 1, 2020, we train models
+In this paper, all models are fit with ``vintage'' training data. This means
+that for a given prediction date, say, September 1, 2020, we train models 
 using only the data that would have been available to us \textit{as of}
-September 1.  Essentially, we ``rewind'' the clock to September 1, and query the
-COVIDcast API to get the latest data that it would have had available at
-that point in time.  This is made possible by the COVIDcast API's comprehensive
-data versioning system (described in more detail in our companion paper).  We
-also use the \texttt{evalcast} R package \cite{EvalcastR}, which streamlines the
+September 1 (imagine that we ``rewind'' the clock to September 1, and query the
+COVIDcast API to get the latest data it would have had available at that point
+in time.)  This is made possible by the COVIDcast API's comprehensive data  
+versioning system (described in more detail in our companion paper).  We also
+use the \texttt{evalcast} R package \cite{EvalcastR}, which streamlines the
 process of training arbitrary prediction models over a sequence of prediction
-dates, by automatically constructing the (properly-versioned) sequence of
-vintage training data sets.
+dates, by constructing the proper sequence of vintage training data sets.
 
@@
 signals. The three signals based on medical claims, CHNG-CLI, CHNG-COVID, and
-DV-CLI, are most often 3-5 days latent, and subject to a considerable degree of
-revision or ``backfill'' after their initial publication date.  The survey-based
-signal, CTIS-CLI-in-community, is 2 days latent, and rarely undergoes any 
-revision at all.  The target itself, reported case rates (or a binary hotspot
-variable, defined based on reported case rates), is 1 day latent, and exhibits
-frequent, unpredictable revisions (sometimes clear anomalies) after initial
-publication.  Compared to the revision pattern in the medical claims signals,
-which are appears to be much more systematic, the revision pattern in the case
-rates signal can be highly erratic, resulting in big spikes as reporting
-backlogs are cleared, changes in case definitions are made, other corrections
-are applied, etc.
+DV-CLI, are most often 3-5 days latent, and subject to a considerable but
+regular degree of revision or ``backfill'' after their initial publication date.
+The survey-based signal, CTIS-CLI-in-community, is 2 days latent, and rarely
+undergoes any revision at all.  The target variable itself, reported COVID-19
+case rates, is 1 day latent, and exhibits frequent, unpredictable revisions
+(sometimes clear anomalies) after initial publication.  Compared to the revision
+pattern in the medical claims signals, which are appears to be much more
+systematic in nature, the revision pattern in the case rates signal can be
+highly erratic, resulting in big spikes as reporting backlogs are cleared,
+changes in case definitions are made, other corrections are applied, etc.
 
-TODO What is ground truth??
-TODO Better example of backfill in DV-CLI?
+TODO Are the above problems mentioned about case rates really to do with vintage 
+data, or would they still exist in finalized data?
+
+TODO Should we discuss the use of ``bad'' case rates as ground truth??
+
+TODO Discuss Google-AA.  We're using all finalized data, but explain why we're
+OK with this ... (they never backfill).
+
+TODO For a figure, use Jacob's ``famous'' example of backfill in DV-CLI?   
+\url{https://cmu-delphi.github.io/covidcast/talks/ml-summit/talk.html#(20)}
+
+\subsection{Missing Data Imputation}
+
+TODO Briefly describe that we zero-fill Google-AA.
 
@@
 
-%\subsubsection{Forecasting Commonalities} % PNAS wants Title Caps for everything
+The first task we consider---\textit{forecasting}---is to predict the HRR-level
+COVID-19 case incidence rates, averaged over some 7-day time period in the
+future (the average number of new daily reported cases per 100,000 people, over
+the 7-day period).  The second task---\textit{hotspot prediction}---is to
+predict a binary variable defined in terms of relative changes of the
+aforementioned target variable. Why define the target variable here via 7-day 
+averaging? The short answer is robustness: averaging stabilizes the case time
+series, and accounts for uninteresting artifacts like weekend-weekday
+differences in the series.  Note that we can of course equivalently view this
+(equivalent up to a constant factor) as predicting HRR-level COVID-19 case
+incidence rates \textit{summed} over some 7-day period in the future (the total
+number of new reported cases per 100,000 people, over the 7-day period), or
+predicting a binary variable defined in terms of its relative changes.
 
-On every day between \attn{July 1 and December 31}, we make two types of
-forecasts for each of the 306 hospital referral regions in the United States:
-(1) we forecast the seven-day trailing average of incident cases, or "smoothed
-incident cases"; and (2) we predict whether the HRR will be a hotspot, a
-location with a large increase in cases. For both tasks we try to predict the
-outcome 7, 8, 9, up to \attn{21} days into the future. For both tasks, we
-estimate a baseline global model across all HRRs using the most recent
-observation of the target, as well the observations from 1 and 2 weeks earlier.
-We train a separate model for each forecast date and for each forecast horizon,
-but we do not assume any HRR-specific effects; that is, a single global model is
-used to 
-describe all HRRs for each forecast date-horizon pair. This model is
-re-estimated for every forecast date and each future target resulting in
-\attn{$180\times 15 = 2700$} estimated baseline models for each task.  
-We then add current and lagged values (analogous to the baseline) for each of
-the four signals separately. We now provide further details on each of these
-tasks.
+To fix notation, let $Y_{\ell,t}$ denote the 7-day trailing average of COVID-19
+case incidence rates in location (HRR) $\ell$ and at time (day) $t$.  In 
+forecasting, a time $t$, we seek to predict $Y_{\ell,t+a}$ for each ``ahead''
+value $a=7,\ldots,21$. Meanwhile, in hotspot prediction, we seek to predict
+$Z_{\ell,t+a}$, a binary indicator defined in terms of the relative difference 
+$(Y_{\ell,t+a}-Y_{\ell,t+a-7}) / Y_{\ell,t+a-7}$, again for each
+$a=7,\ldots,21$.
 
-\subsubsection{Task 1: Case Forecasting}
+For each prediction date $t$, we use a 21-day trailing window of data to train
+our forecast or hotspot prediction models (so, e.g., the trained models will
+differ from those at prediction date $t-1$).  This is done to account for
+(potential) nonstationarity.  For simplicity, the forecasting and hotspot
+prediction models are always fit to data across all HRRs (i.e., the coefficients
+in the models do not account for location-specific effects).
 
-To fix notation, let $Y_{\ell,t}$ denote the smoothed COVID-19 case
-incidence rate per 100,000 people for location (HRR) $\ell$ and time (day) $t$. 
-Let $X_{\ell,t}$ denote
-the one of the indicators described above, for example the Facebook \%
-CLI-in-community signal, 
-for location $\ell$ and time $t$.
-We evaluate five models of the form:
-\begin{align*}
-Y_{\ell,t+d}
-&\approx \alpha + \sum_{j=0}^2 \beta_j Y_{\ell,t-7j} \\
-Y_{\ell,t+d}
-&\approx \alpha + \sum_{j=0}^2 \beta_j Y_{\ell,t-7j} +
-\sum_{j=0}^2 \gamma_j X_{\ell,t-7j}, 
-\end{align*}
-where we also exchange $X_{\ell, t}$ for the signals from \attn{CHNG and
-  Google.} 
-Here $d=7,\ldots, 21$ depending on the target value
-(number of days ahead we predict).
+In our analysis, we let the prediction date $t$ run over each day in between
+early June and December 31, 2020.  The precise start date differs for
+forecasting and hotspots prediction; for each task it was chosen to be the
+earliest date at which the data needed to train all models was available, which
+ends up being (per our setup, with 21 days of training data, and lagged values
+of signals for features as we will detail shortly) June 6, 2020 for forecasting,
+and June 13, 2020 for hotspot prediction. (The bottleneck here is the
+CTIS-CLI-in-community signal, which does not exist before early April 2020, when
+the survey was first launched). The end date was chosen with a consideration
+again to align both tasks as best as possible, and because few hotspots exist
+post December 31, 2020, due to the general and gradual decline of the pandemic
+in 2021.
 
-Informally, the first model, which we will call the ``Cases'' model, 
-bases its predictions of future case rates on the following three features:
-current COVID-19 case rates and those 1 and 2 weeks back.
-The second model, ``Cases + Facebook'', additionally incorporates the 
-current Facebook signal and the Facebook signal from 1 and 2 weeks in the past.
-The third model, ``Cases + Google'', is exactly the same but substitutes the 
-Google signal instead of the Facebook one and similarly for ``Cases + Change''.
+\subsubsection{Task 1: Forecasting}
 
-For each model, in order to make a forecast at time $t_0$
-(to predict case rates at time $t_0+d$),
-we estimate the unknown parameters using quantile
-regression~\cite{KoenkerXiao2006}, 
-training over all locations $\ell$,
-and all times $t$ that are within the most recent 21 days of data
-available up to and including time $t_0$.
-Rather than predicting only the median (or mean) value for each forecast date
-and each horizon, we estimate seven quantiles, coinciding 
-with the county-level forecast submission guidelines for the
-COVID-19 Forecast Hub~\cite{Cramer:2021}.  These are \{0.025, 0.1,
-0.25, 0.5, 
-0.75, 0.9, 0.975\}. This provides a notion of uncertainty in the forecasts. An
-example of this type of forecast is shown in \autoref{fig:trajectory}.
+Recall $Y_{t,\ell}$ denotes the 7-day trailing average of COVID-19 case
+incidence rates in location $\ell$ and at time $t$.  Separately for each
+$a=7,\ldots,28$, to predict $Y_{\ell,t+a}$ for ahead value $a$, we consider a
+simple probabilistic forecasting model of the form:     
+\begin{equation}
+\label{eq:forecast_ar}
+\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s}, s \leq t)  
+= \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j}.  
+\end{equation}
+This model uses current case rates, and the case rates 7 and 14 days ago, in
+order to predict (the quantiles of) case rates in the future.  We consider a
+total of 7 quantile levels (chosen in accordance with the county-level quantile
+levels suggested by the COVID-19 Forecast Hub),      
+\begin{equation}
+\label{eq:quantile_levels}
+\tau \in \{0.025, 0.1, 0.25, 0.5, 0.75, 0.9, 0.975 \}.
+\end{equation}
+To estimate the coefficients, we use \textit{quantile regression}
+\cite{Koenker:1978, Koenker:2005, Koenker:2006} fit separately for each $\tau$,
+using data from all 306 HRRs, and within each HRR, using the most recent 21 days
+of training data.  To be clear, this gives 6,426 training samples for each
+quantile regression problem.
+
+In addition to the above ``pure'' autoregression model, we consider five
+probabilistic forecasting models of the form: 
+\begin{multline}
+\label{eq:forecast_ar_x}
+\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s}, X_{\ell,s}, s \leq t)  
+= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j} + 
+\sum_{j=0}^2 \gamma^{a,\tau}_j X_{\ell,t-7j},
+\end{multline}
+where $X_{\ell,t}$ denotes any one of the five auxiliary indicators---CHNG-CLI, 
+CHNG-COVID, CTIS-CLI-in-community, DV-CLI, or Google-AA---at location $\ell$ and
+time $t$. Note that we apply the same lags (current value, along with the values
+7 and 14 days ago) for the auxiliary signals as we do for case rates. Estimation
+proceeds just as before: we use the same 7 quantile levels as in
+\eqref{eq:quantile_levels}, and again use quantile regression separately for
+each $\tau$, fit using data from all 306 HRRs and a trailing window of 21 days
+of training data.
+
+At prediction time, in order to deal with possible crossing violations (a
+predicted quantile at level $\tau$ exceeds a predicted quantile at level $\tau'
+> \tau$), we apply a simple post-hoc sorting.  See Figure \ref{fig:trajectory}
+for example forecast.
 
@@
 \centering
-    \includegraphics[width=.9\linewidth]{fig/trajectory.pdf}
-    \caption{Forecasts and observed cases for HRR x. \attn{Create a figure here
-        for 1 HRR. Show the fan for the quantiles.}} 
+\includegraphics[width=\columnwidth]{fig/trajectory.pdf}
+\caption{Forecasts and observed cases for HRR x. TODO 
+  \attn{Create a figure here for 1 HRR. Show the fan for the quantiles.}}  
 \label{fig:trajectory}
@@
 
-% Noncrossing is enforced via post hoc ordering.
-
 \subsubsection{Task 2: Hotspot Prediction}
 
-We also examine whether alternative indicators help in predicting hotspots. On a
-given date, we say an HRR is a hotspot if the average number of observed cases
-over the past 7 days has increased by at least $25$\% compared to the preceding
-week. We remove all data where, on average, fewer than $30$ cases were observed
-over the preceding $7$ days. Thus, hotspot status can be represented by a
-$\{0,1\}$-valued variable, and hotspot prediction is a binary classification
-problem.
+Define the binary indicator:
+$$
+Z_{t,\ell} = \indicator{Y^\Delta_{\ell,t} \geq 0.25},
+$$
+where we use the notation \smash{$Y^\Delta_{\ell,t} = (Y_{\ell,t} - Y_{\ell,
+    t-7})/(Y_{\ell,t-7})$}. In other words, $Z_{t,\ell}=1$ if the number of
+newly reported cases over the past 7 days has increased by at least 25\%
+compared to the preceding week.  When this occurs, we call location $\ell$ a
+\textit{hotspot} at time $t$.
 
-As in the case forecasting task, in hotspot prediction we train models that use
-previously observed values of cases and alternative indicators as features to
-predict a response (in our case, hotspot status). In addition to the differently
-defined response, there are two other differences between the hotspot prediction
-and case forecasting tasks. First, in hotspot prediction we always use relative
-changes in previous observed cases/indicators as features. That is, rather than
-$Y_{t,\ell}$ we use 
-$$
-Y^\Delta_{\ell,t} = \frac{Y_{\ell,t} - Y_{\ell,
-    t-7}}{Y_{\ell,t-7}}\quad\text{and}\quad X^\Delta_{\ell,t} =
-\frac{X_{\ell,t} - X_{\ell, t-7}}{X_{\ell,t-7}} 
-$$ 
-Second, because hotspot prediction is a binary classification problem, we
-estimate the parameters $\alpha$, $\beta$ and $\gamma$ by logistic regression
-rather than quantile regression.
+We treat hotspot prediction as a binary classification problem and use a setup 
+altogether quite similar to the forecasting setup described previously.
+Separately for each $a=7,\ldots,28$, to predict $Z_{\ell,t+a}$, we consider a
+simple logistic model:
+\begin{equation}
+\label{eq:hotspot_ar}
+\logit(Z_{\ell,t+a} \,|\, Y_{\ell,s}, s \leq t)  
+= \alpha^{a,\tau} + \sum_{j=0}^2 \^{a,\tau}_j Y^\Delta_{\ell,t-7j}.  
+\end{equation}
+In addition to this ``pure'' autoregression model, we also consider five
+logistic models of the form: 
+\begin{multline}
+\label{eq:hotspot_ar_x}
+\logit(Z_{\ell,t+a} \,|\, Y_{\ell,s}, X_{\ell,s}, s \leq t)  
+= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j} + 
+\sum_{j=0}^2 \gamma^{a,\tau}_j X_{\ell,t-7j},
+\end{multline}
+where we use \smash{$X^\Delta_{\ell,t} = (X_{\ell,t} - X_{\ell,
+    t-7})/(X_{\ell,t-7})$}, and again $X_{\ell,t}$ stands for any of the five  
+auxiliary indicators at location $\ell$ and time $t$.  We fit all models
+\eqref{eq:hotspot_ar}, \eqref{eq:hotspot_ar_x} using logistic regression, 
+pooling all 306 HRRs and using a 21-day trailing window for the training data.
+
+An important detail is that in hotspot prediction we remove all data from
+training and evaluation where, on average, fewer than 30 cases (this refers to a
+count, not a rate) are observed over the preceding 7 days. This avoids having to
+make arbitrary calls for a hotspot (or lack thereof) based on small counts.
 
@@
 
-Code to reproduce all examples (which uses the
-\texttt{evalcast} R package and fetches data directly from the API) can be
-found at \url{TODO}.
+Code to reproduce all results (which uses the \texttt{evalcast} R package) can 
+be found at
+\url{https://www.github.com/cmu-delphi/covidcast-pnas/forecast/code/}.
 

### 2021-06-13 indicators/paper/paper.tex
Update discussion of signals

@@
 or other common sources, such as:
-\begin{description}
-\item[Health Insurance Claims] Based on de-identified medical insurance claims
-  from Change Healthcare and other health system partners, we release indicators
-  on the estimated percentage of covered outpatient visits and hospitalizations
-  that involved COVID diagnoses or symptoms.
-\item[Internet-Based Surveys] Conducted in partnership with Facebook, Delphi's
-  COVID Tracking and Impact Survey receives about 50,000 responses daily, and
-  has received over 20 million responses since April 2020 \cite{DelphiSurvey,
-    Kreuter:2020}. From the surveys, we construct indicators on symptoms,
-  social distancing, vaccination, and other attitudes and behaviors related to
-  COVID.
-\item[COVID Antigen Tests] Based on data from Quidel, a major manufacturer of
+
+\paragraph{Health Insurance Claims} Based on de-identified medical insurance 
+claims from Change Healthcare and other health system partners, we release
+indicators on the estimated percentage of covered outpatient visits and
+hospitalizations that involved COVID diagnoses or symptoms.
+
+\paragraph{Internet-Based Surveys} Conducted in partnership with Facebook,
+Delphi's COVID Tracking and Impact Survey receives about 50,000 responses daily,  
+and has received over 20 million responses since April 2020 \cite{DelphiSurvey,
+  Kreuter:2020}. From the surveys, we construct indicators on symptoms, social 
+distancing, vaccination, and other attitudes and behaviors related to COVID.
+
+\paragraph{COVID Antigen Tests} Based on data from Quidel, a manufacturer of
 COVID antigen tests in the United States, we calculate and release
 (Quidel-specific) test volumes and positivity rates.
-\item[Search Trends] Based on Google's COVID-19 Search Trends symptoms data set
+
+\paragraph{Search Trends} Based on Google's COVID-19 Search Trends data set
 \cite{GoogleSymptoms}, we provide indicators reflecting COVID-related search  
 activity.
-\item[Mobility Data] Cell phone mobility data is collected by SafeGraph
-  \cite{SafeGraphSD, SafeGraphPatterns}, and is available to researchers under a
-  data use agreement; we aggregate it to the county level and make indicators
-  publicly available.
-\end{description}
 
+\paragraph{Mobility Data} SafeGraph, a company that collects geospatial data
+from smartphone apps, calculates COVID-related mobility signals
+\cite{SafeGraphSD, SafeGraphPatterns} and makes it available to researchers
+under a data use agreement; we aggregate (some of) these signals to the county
+level and make them publicly available.
+
+\medskip
 We also scrape data accessible from other public sources, such as cases and
@@
 
-Together, we collect TODO signals from TODO distinct sources, aggregate them,
-and provide them in a single common format for public access. This hence
-provides a common format both for unique data streams not available elsewhere
-and for public data streams, enabling efficient comparison and modeling use.
-
-A summary of the data sources we currently receive and make available is
-presented in Table~\ref{data-sources}, and detailed documentation of the signals
-available from each source is available at
+Altogether, we collect 110 signals from 12 distinct sources, aggregate and
+process them, and provide them in a common format for public access. This
+unifies both unique (not available anywhere else) and standard COVID data
+streams into a single common format, enabling efficient comparison and modeling.
+A summary of the data sources and signals in the API is in
+Table~\ref{data-sources}, and detailed documentation is available at
 \url{https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html}.
@@
   limited form); the second group is mirrored from public sources.}
-\begin{tabular}{>{\raggedright}p{1.2in} p{4.0in} r >{\raggedright\arraybackslash}p{0.5in}}\toprule
+\begin{tabular}{>{\raggedright}p{1.2in} p{4.0in} l >{\raggedright\arraybackslash}p{0.5in}}
+\toprule
 \textbf{Data source} & \textbf{Signals available} & \textbf{First date} &
@@
 2019-12-31 & State \\
-    COVID Act Now & COVID-19 testing results, such as positivity and number of
+COVID Act Now & COVID-19 testing results, such as positivity rate and number of
  tests, compiled by COVID Act Now from CDC reporting. & 2020-03-02 & 
@@
 \end{tabular}
-% TODO PNAS wants a short title above and caption below, but they can't possibly
-% mean for the caption to be formatted so weirdly...
-\addtabletext{Location availability often drops lower for some times and
-  signals, e.g., due to variable reporting volume. $^*$Available at $>60$\% of
-  counties; $^{**}$available at 20--60\% of counties; $^{***}$available at
-  $<20$\% of counties.}
+\addtabletext{$^*$Available at $>60$\% of counties; $^{**}$available at 20--60\%
+  of counties; $^{***}$available at $<20$\% of counties.  For some signals,
+  location availability varies over time, e.g., due to variable reporting
+  volume.}
 \label{data-sources}

### 2021-06-13 indicators/code/count_signals.R

@@ -1,0 +1,16 @@
+library(covidcast)
+library(dplyr)
+
+covidcast_meta() %>%
+  group_by(data_source, signal) %>%
+  summarize(county = ifelse("county" %in% geo_type, "*", ""),
+            msa = ifelse("msa" %in% geo_type, "*", ""),
+            hrr = ifelse("hrr" %in% geo_type, "*", ""),
+            state = ifelse("state" %in% geo_type, "*", ""),
+            max_time = min(max_time, na.rm = TRUE)) %>%
+  mutate(signal = paste(data_source, signal, sep=":")) %>%
+  ungroup() %>%
+  slice(grep("(raw|7dav|\\_w)", signal, invert = TRUE)) %>%
+  filter(max_time >= "2021-06-01") %>%
+  summarize(num_sources = length(unique(data_source)),
+            num_signals = length(unique(signal)))

### 2021-06-13 indicators/paper/paper.tex
Improve URLs, other formatting

@@
 from smartphone apps, calculates COVID-related mobility signals
-\cite{SafeGraphSD, SafeGraphPatterns} and makes it available to researchers
+\cite{SafeGraphSocial, SafeGraphPatterns} and makes it available to researchers 
 under a data use agreement; we aggregate (some of) these signals to the county
@@
 A summary of the data sources and signals in the API is in
-Table~\ref{data-sources}, and detailed documentation is available at
+Table~\ref{tab:sources_signals}, and detailed documentation is available at 
 \url{https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html}.
 
-\begin{table*}
+\begin{table*}[t]
   \centering
@@
     % Exclusive data
-  Change Healthcare & Percentage of outpatient visits with COVID diagnostic codes
-or codes indicating COVID-like symptoms. Based on de-identified claims data
-processed by Change Healthcare. & 2020-02-01 & County$^*$ \\
+    Change Healthcare & Percentage of outpatient visits with COVID diagnostic
+codes or codes indicating COVID-like symptoms. Based on de-identified claims
+data processed by Change Healthcare. & 2020-02-01 & County$^*$ \\
     Doctor Visits & Percentage of outpatient visits primarily about COVID-like
@@
 and restaurants, based on cell phone mobility data collected by SafeGraph
-\cite{SafeGraphSD, SafeGraphPatterns}. & 2019-01-01 & County \\
- COVID Tracking and Impact Survey & COVID symptoms, social distancing behaviors,
-mental health, economic impact, behavior (e.g., mask-wearing, vaccination
-attitudes), and COVID testing signals based on daily surveys conducted
-nationally by Delphi through Facebook \cite{DelphiSurvey, Kreuter:2020}. &
-2020-04-06 & County$^{**}$ \\
+\cite{SafeGraphSocial, SafeGraphPatterns}. & 2019-01-01 & County \\
+    COVID Tracking and Impact Survey & COVID symptoms, social distancing
+behaviors, mental health, economic impact, behavior (e.g., mask-wearing,
+vaccination attitudes), and COVID testing signals based on daily surveys
+conducted nationally by Delphi through Facebook \cite{DelphiSurvey,
+Kreuter:2020}. & 2020-04-06 & County$^{**}$ \\
     \midrule
@@
 2019-12-31 & State \\
-COVID Act Now & COVID-19 testing results, such as positivity rate and number of
- tests, compiled by COVID Act Now from CDC reporting. & 2020-03-02 & 
+    COVID Act Now & COVID-19 testing results, such as positivity rate and number
+of tests, compiled by COVID Act Now from CDC reporting. & 2020-03-02 &
 County$^*$ \\
-  Google Symptoms & Trends in Google search volume for terms related to anosmia
-and ageusia (loss of smell or taste), which correlate with COVID activity, based
-on data shared by Google \cite{GoogleSymptoms}. & 2020-02-13 & County$^{***}$ \\
+    Google Symptoms & Trends in Google search volume for terms related to
+anosmia and ageusia (loss of smell or taste), which correlate with COVID
+activity, based on data shared by Google \cite{GoogleSymptoms}. & 2020-02-13 &
+County$^{***}$ \\
     Cases and Deaths & Confirmed COVID-19 cases and deaths, compiled by JHU CSSE 
 \cite{Dong:2020} and by USAFacts \cite{USAFacts}. & 2020-01-22 & County \\
-  NCHS Mortality & Weekly totals of deaths broken down by cause, such as COVID,
-flu, or pneumonia, compiled by the National Center for Health Statistics [TODO
-cite]. & 2020-01-26 & State \\
+    NCHS Mortality & Weekly totals of deaths broken down by cause, such as
+COVID, flu, or pneumonia, compiled by the National Center for Health Statistics
+[TODO cite]. & 2020-01-26 & State \\
     \bottomrule
   \end{tabular}
-\addtabletext{$^*$Available at $>60$\% of counties; $^{**}$available at 20--60\%
-  of counties; $^{***}$available at $<20$\% of counties.  For some signals,
-  location availability varies over time, e.g., due to variable reporting
-  volume.}
-\label{data-sources}
+  \addtabletext{$^*$Available at $>60$\% of counties; $^{**}$available at
+    20--60\% of counties; $^{***}$available at $<20$\% of counties.  For some
+    signals, location availability varies over time, e.g., due to variable
+    reporting volume.}
+  \label{tab:sources_signals}
 \end{table*}
@@
 The R and Python package software is public and open-source, at
-\url{https://github.com/cmu-delphi/covidcast/}.  The underlying server software
-is itself also public and open-source, at
+\url{https://github.com/cmu-delphi/covidcast/}.  The API server software is
+itself also public and open-source, at
 \url{https://github.com/cmu-delphi/delphi-epidata/}.  Lastly, most data sources 
@@
 the pandemic \cite{Adjodah:2021, Pierri:2021, Jewell:2021, Chakrabarti:2020,
-  Doerr:2021}. The API currently serves hundreds of thousands of requests each
-day, to thousands of users worldwide.
+  Doerr:2021}. The API currently serves hundreds of thousands of requests to
+thousands of users, each day.
 
@@
 
-\begin{figure}
+\begin{figure}[t]
   \centering
@@
 \begin{figure}[t]
-
   \includegraphics[width=\columnwidth]{fig/geo_wise_corr.pdf}
@@
 
-\begin{figure}
+\begin{figure}[t]
 \centering

### 2021-06-13 forecast/paper/paper.tex
Updates thru methods C (formerly D)

@@
 inclusion of an auxiliary indicator as an additional feature in such a model
-improves predictive power. While forecasting carries a rich literature that
-offers a wide range of techniques, see e.g., \cite{Hyndman:2018}, we purposely
-constrain ourselves to very simple models, avoiding common enhancements such as
-order selection, correction of outliers/anomalies in the data, and inclusion of
-regularization or nonlinearities. That said, analyses of forecasts submitted to
-the COVID-19 Forecast Hub \cite{ForecastHub} by a large community of modelers
-have shown that simple, robust models have consistently been among the
-best-performing over the pandemic \cite{Cramer:2021}, including time series
-models similar to those we consider in what follows.
+improves its predictions.
+
+While forecasting carries a rich literature that offers a wide range of
+techniques, see e.g., \cite{Hyndman:2018}, we purposely constrain ourselves to
+very simple models, avoiding common enhancements such as order selection,
+correction of outliers/anomalies in the data, and inclusion of regularization or
+nonlinearities. That said, analyses of forecasts submitted to the COVID-19
+Forecast Hub \cite{ForecastHub} by a large community of modelers have shown that
+simple, robust models have consistently been among the best-performing over the
+pandemic \cite{Cramer:2021}, including time series models similar to those we
+consider in what follows.
 
@@
 people (the standard in epidemiology).  We use reported case data aggregated by 
-JHU CSSE \cite{Dong:2020}, which is accessible through the COVIDcast API, like
-the auxiliary indicators that we use to supplement the basic purely
-autoregressive models.
+JHU CSSE \cite{Dong:2020}, which is accessible through the COVIDcast API
+\cite{CovidcastAPI}, like the auxiliary indicators that we use to supplement the
+basic purely autoregressive models.
 
 The indicators we focus on provide information not generally available from
-standard public health reporting. Among the many indicators in the API, we
-consider the following five:
+standard public health reporting. Among the many auxiliary indicators collected
+in the API, we study the following five:  
 \begin{itemize}
@@
 systematic in nature, the revision pattern in the case rates signal can be
-highly erratic, resulting in big spikes as reporting backlogs are cleared,
-changes in case definitions are made, other corrections are applied, etc.
-
-TODO Are the above problems mentioned about case rates really to do with vintage 
-data, or would they still exist in finalized data?
+highly erratic. Big spikes or other anomalies can occur in the data as reporting
+backlogs are cleared, changes in case definitions are made, etc. Groups like JHU 
+CSSE then work tirelessly to correct such anomalies (e.g., back-distribute a
+huge spike when a reporting backlog is cleared), resulting in highly nontrivial
+and unpredictable revisions.
 
@@
 
-\subsection{Missing Data Imputation}
-
-TODO Briefly describe that we zero-fill Google-AA.
-
 \subsection{Analysis Tasks}
 
-The first task we consider---\textit{forecasting}---is to predict the HRR-level
-COVID-19 case incidence rates, averaged over some 7-day time period in the
-future (the average number of new daily reported cases per 100,000 people, over
-the 7-day period).  The second task---\textit{hotspot prediction}---is to
-predict a binary variable defined in terms of relative changes of the
-aforementioned target variable. Why define the target variable here via 7-day 
-averaging? The short answer is robustness: averaging stabilizes the case time
-series, and accounts for uninteresting artifacts like weekend-weekday
-differences in the series.  Note that we can of course equivalently view this
-(equivalent up to a constant factor) as predicting HRR-level COVID-19 case
-incidence rates \textit{summed} over some 7-day period in the future (the total
-number of new reported cases per 100,000 people, over the 7-day period), or
-predicting a binary variable defined in terms of its relative changes.
+\begin{table*}[t]
+  \centering
+  \caption{Summary of forecasting and hotspot prediction tasks considered in
+    this paper.}
+  \begin{tabular}{l p{2.75in} p{2.75in}}
+    \toprule
+    & \textbf{Forecasting} & \textbf{Hotspot prediction} \\
+    \midrule
+    Response variable & $Y_{\ell,t}$ (7-day trailing average of COVID-19 case
+ incidence rates, per location $\ell$ and time $t$) & $Z_{\ell,t} =
+\indicator{Y_{\ell,t} \geq 1.25 \cdot Y_{\ell, t-7}}$ (indicator that $Y_{\ell,t}$ 
+grows by more than 25\% relative to the preceding week) \\
+    Geographic resolution & Hospital referral region (HRR) & Hospital referral
+region (HRR) \\ 
+    Forecast period & June 6--December 31, 2020 & June 13--December 31, 2020 \\  
+    Model type & Quantile regression & Logistic regression \\
+    Evaluation metric & Weighted interval score (WIS) & Area under curve (AUC) \\
+    \bottomrule
+  \end{tabular}
+\label{tab:analysis_tasks}
+\end{table*}
 
 To fix notation, let $Y_{\ell,t}$ denote the 7-day trailing average of COVID-19
-case incidence rates in location (HRR) $\ell$ and at time (day) $t$.  In 
-forecasting, a time $t$, we seek to predict $Y_{\ell,t+a}$ for each ``ahead''
-value $a=7,\ldots,21$. Meanwhile, in hotspot prediction, we seek to predict
-$Z_{\ell,t+a}$, a binary indicator defined in terms of the relative difference 
-$(Y_{\ell,t+a}-Y_{\ell,t+a-7}) / Y_{\ell,t+a-7}$, again for each
-$a=7,\ldots,21$.
+case incidence rates in location (HRR) $\ell$ and at time (day) $t$.  To be
+clear, this is the number of new daily reported cases per 100,000
+people, averaged over the 7-day period $t-6, \ldots t$. The first task we
+consider---\textit{forecasting}---is to predict $Y_{\ell,t+a}$ for each
+``ahead'' value $a=7,\ldots,21$.  The second task---\textit{hotspot
+  prediction}---is to predict a binary variable defined in terms of the relative
+change of $Y_{\ell,t+a}$ (relative to its value one week prior,
+$Y_{\ell,t+a-7}$), again for each $a=7,\ldots,21$.
+
+Why do we define the response variables via 7-day averaging? The short answer 
+is robustness: averaging stabilizes the case time series, and accounts for 
+uninteresting artifacts like weekend-weekday differences in the series.  Note
+that we can of course equivalently view this (equivalent up to a constant
+factor) as predicting the HRR-level COVID-19 case incidence rate \textit{summed} 
+over some 7-day period in the future, and predicting a binary variable derived 
+from this.
+
+In what follows, we cover more details on our two analysis
+tasks. Table~\ref{tab:analysis_tasks} presents a summary.
+
+\paragraph{Dynamic Re-Training}
 
@@
 (potential) nonstationarity.  For simplicity, the forecasting and hotspot
-prediction models are always fit to data across all HRRs (i.e., the coefficients
-in the models do not account for location-specific effects).
+prediction models are always trained on data across all HRRs (i.e., the
+coefficients in the models do not account for location-specific effects).
+
+\paragraph{Prediction Period}
 
 In our analysis, we let the prediction date $t$ run over each day in between
-early June and December 31, 2020.  The precise start date differs for
+early/mid June and December 31, 2020.  The precise start date differs for 
 forecasting and hotspots prediction; for each task it was chosen to be the
@@
 CTIS-CLI-in-community signal, which does not exist before early April 2020, when
-the survey was first launched). The end date was chosen with a consideration
-again to align both tasks as best as possible, and because few hotspots exist
-post December 31, 2020, due to the general and gradual decline of the pandemic
-in 2021.
+the survey was first launched). The end date was chosen again with a
+consideration to align both tasks as best as possible, and because few hotspots
+exist post December 31, 2020, due to the general and gradual decline of the
+pandemic in 2021.
 
-\subsubsection{Task 1: Forecasting}
+\paragraph{Forecasting Models}
 
@@
 \end{equation}
-To estimate the coefficients, we use \textit{quantile regression}
-\cite{Koenker:1978, Koenker:2005, Koenker:2006} fit separately for each $\tau$,
+We fit \eqref{eq:forecast_ar} using \textit{quantile regression}
+\cite{Koenker:1978, Koenker:2005, Koenker:2006} separately for each $\tau$, 
 using data from all 306 HRRs, and within each HRR, using the most recent 21 days
-of training data.  To be clear, this gives 6,426 training samples for each
-quantile regression problem.
+of training data.  This gives us 6,426 training samples for each quantile 
+regression problem.
 
@@
 time $t$. Note that we apply the same lags (current value, along with the values
-7 and 14 days ago) for the auxiliary signals as we do for case rates. Estimation
-proceeds just as before: we use the same 7 quantile levels as in
-\eqref{eq:quantile_levels}, and again use quantile regression separately for
-each $\tau$, fit using data from all 306 HRRs and a trailing window of 21 days
-of training data.
+7 and 14 days ago) for the auxiliary indicators as we do for the case
+rates. Training then proceeds just as before: we use the same 7 quantile levels
+in \eqref{eq:quantile_levels}, and fit quantile regression separately for each 
+level $\tau$, using data from all 306 HRRs and a trailing window of 21 days of
+training data.
 
@@
 > \tau$), we apply a simple post-hoc sorting.  See Figure \ref{fig:trajectory}
-for example forecast.
+for an example forecast.
 
@@
 
-\subsubsection{Task 2: Hotspot Prediction}
+\paragraph{Hotspot Prediction Models}
 
@@
 newly reported cases over the past 7 days has increased by at least 25\%
-compared to the preceding week.  When this occurs, we call location $\ell$ a
+compared to the preceding week.  When this occurs, we say location $\ell$ is a
 \textit{hotspot} at time $t$.
@@
 \logit(Z_{\ell,t+a} \,|\, Y_{\ell,s}, X_{\ell,s}, s \leq t)  
-= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j} + 
-\sum_{j=0}^2 \gamma^{a,\tau}_j X_{\ell,t-7j},
+= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j} +  
+\sum_{j=0}^2 \gamma^{a,\tau}_j X^\Delta_{\ell,t-7j},
 \end{multline}
@@
 to those adding indicators.
+
+\subsection{Other Considerations}
+
+\paragraph{Missing Data Imputation}
+
+TODO Briefly describe that we zero-fill Google-AA.
+
+\paragraph{Spatial Heterogeneity}
+
+\paragraph{Backfill, Nowcasting}
 

### 2021-06-13 forecast/paper/paper.tex
Update discussion of case revisions

@@
 (sometimes clear anomalies) after initial publication.  Compared to the revision
-pattern in the medical claims signals, which are appears to be much more
-systematic in nature, the revision pattern in the case rates signal can be
-highly erratic. Big spikes or other anomalies can occur in the data as reporting
-backlogs are cleared, changes in case definitions are made, etc. Groups like JHU 
-CSSE then work tirelessly to correct such anomalies (e.g., back-distribute a
-huge spike when a reporting backlog is cleared), resulting in highly nontrivial
-and unpredictable revisions.
-
-TODO Should we discuss the use of ``bad'' case rates as ground truth??
-
-TODO Discuss Google-AA.  We're using all finalized data, but explain why we're
-OK with this ... (they never backfill).
+pattern in the medical claims signals, which are much more systematic in nature,
+revisions in case reports can be highly erratic. Big spikes or other anomalies
+can occur in the data as reporting backlogs are cleared, changes in case
+definitions are made, etc. Groups like JHU CSSE then work tirelessly to correct
+such anomalies after first publication (e.g., they will attempt to
+back-distribute a spike when a reporting backlog is cleared, by working with a
+local authority to figure out how this should best be done), which can result
+in very nontrivial and often unpredictable revisions.
 
@@
 \url{https://cmu-delphi.github.io/covidcast/talks/ml-summit/talk.html#(20)}
+
+Lastly, our treatment of the Google-AA signal is different from the rest.
+Because Google's team did not start publishing this signal until early
+September, 2020, we do not have true vintage data before then.  However, this
+signal is never revised after initial publication (confirmed via personal
+communication with the Google team that produces this signal), so we simply use
+finalized values for it in our analysis.
 
@@
 \paragraph{Backfill, Nowcasting}
+
+TODO Should we discuss the use of ``bad'' case rates as ground truth??
 

### 2021-06-13 forecast/paper/paper.tex
Edits through methods D

@@
 compared to the preceding week.  When this occurs, we say location $\ell$ is a
-\textit{hotspot} at time $t$.
+\textit{hotspot} at time $t$.  Empirically, this rule labels about 27\% of
+location-time pairs as hotspots, during the prediction period (June 13--December
+31, 2020).
 
@@
 
-We evaluate forecasts for each forecast date and horizon using the weighted
-interval score. Weighted interval score (WIS), a well-known quantile-based
-approximation of the commonly-used continuous ranked probability
-score~\cite{gneiting-proper-score}. WIS is a proper score, meaning that lower
-scores imply better forecasters, and can be thought of as a distributional
-generalization of absolute error. The CDC has embraced this metric in the
-context of epidemic forecasting and especially for COVID-19~\cite{wis-paper}.
-Essentially, pairs of quantile forecasts are treated as a central  interval. For
-example, the 0.1 and 0.9 quantiles create an 80\% interval that should, if the
-forecaster is well-calibrated, contain the truth about 80\% of the time. A
-forecaster whose 80\% interval contains the truth less frequently is overly
-confident, while one whose interval contains the truth more frequently (say, by
-using an infinitely-wide interval) is failing to take a stand on the future
-outcome.
+For forecasting, we evaluate the probabilistic forecasts produced by the
+quantile models \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x} using
+\textit{weighted interval score} (WIS), which is a well-known quantile-based
+scoring rule; see, e.g., \cite{Gneiting:2007}.  WIS is a proper score, which
+means that its expectation is minimized by the true quantiles of the target
+variable.  The use of WIS in scoring COVID-19 forecasts is discussed by
+\cite{Bracher:2021}; and WIS serves as the primary metric for forecast
+evaluation in the COVID-19 Forecast Hub.
 
-Here, denoting $K$ predicted intervals along with the median  as
-$\{\alpha_K/2,\ldots,\alpha_2/2,\  
-\alpha_1/2,\ 0.5, 1-\alpha_1/2,\ 1-\alpha_2/2,\ldots, 1-\alpha_K/2\}$, the associated
-forecasts as $\{l_K,\ldots, l_2,\ l_1,\ m,\ u_1,\ u_2,\ldots, u_K \}$, and the observed
-value (after the forecast was made) as $y$, this metric can be expressed as
-\attn{someone please check this} 
-\begin{align*}
- \operatorname{WIS}_K &= \frac{1}{K}|m -y | + \frac{2}{K+1}\sum_{k=1}^K 
-\frac{1}{\alpha_k}(u_k - l_k)\\ &\quad + \sum_{k=1}^K  
-(l_k - y)\indicator{y < l_k} +
-(y - u_k)\indicator{y> u_k},
-\end{align*}
-where $\indicator{a} = 1$ if $a$ is true and $0$ otherwise. Essentially, this
-metric decomposes into 4 parts: (1) a penalty for the median, the point forecast
-or best guess, being far from the truth; (2) a penalty for the width of each
-interval ($u_k - l_k$); (3)  a penalty for overprediction, when the observed
-value falls below $l_k$ and hence underneath the interval; and (4) a
-corresponding penalty for underprediction. Weighted interval score can also be
-written in terms of an average of the quantile or ``pinball'' losses. Setting $\tau_1 =
-\alpha_K/2,\ \ldots,\ \tau_{K+1} = 0.5,\ \ldots,\ \tau_{2K+1} = 1-\alpha_K/2$ and
-similarly rewriting $q_1 = l_K,\ \ldots,\ q_{K+1}=m,\ \ldots,\ q_{2K+1}=u_K$, then
-\begin{align*}
-  \operatorname{WIS}_K &= \frac{2}{2K+1} \sum_{k=1}^{2K+1}
-                         \left(\indicator{y \leq q_k} -\tau_k\right)
-                         (q_k - y).
-\end{align*}
-While this second expression does not have the convenient decomposition
-discussed above, the pinball loss is precisely the objective function that
-quantile regression minimizes. So, mathematically, quantile regression should
-give parameter estimates that minimize the in-sample weighted interval score.
-% DJM: In trying to figure out where 'pinball' comes from, it looks like the
-% preprint of the massively cited Koenker and Hallock (2001) JEP article. See
-% http://www.econ.uiuc.edu/~roger/research/intro/rq.pdf. They say that tilting
-% the absolute value is `pinball logic' which is not a connection I would have
-% ever made. The language doesn't seem to have made it to the final version.
-% Maybe there's an earlier reference?
+WIS is typically defined for quantile-based forecasts where the quantile levels
+are symmetric around 0.5.  This is the case for our choice in
+\eqref{eq:quantile_levels}.  Let $F$ be a forecaster comprised of predicted
+quantiles $q_\tau$ parametrized by a quantile level $\tau$.  In the symmetric
+case, this is equivalent to a collection of central prediction intervals
+$(\ell_\alpha, u_\alpha)$, parametrized by an exclusion probability $\alpha$. 
+The WIS of the forecaster $F$, evaluated at the target variable $Y$, is defined
+by  
+\begin{equation}
+\label{eq:wis_intervals}
+\WIS(F,Y) = \sum_\alpha \Big\{\alpha(u_\alpha - \ell_\alpha) + 2 \cdot
+\dist(Y, [\ell_\alpha, u_\alpha])\Big\},  
+\end{equation}
+where $\dist(a, S)$ is the distance between a point $a$ and set $S$ (the
+smallest distance between $a$ and an element of $S$).  Note that, corresponding 
+to \eqref{eq:quantile_levels}, the exclusion probabilities are $\alpha \in
+\{0.05, 0.2, 0.5, 1\}$, resulting in 4 terms in the above sum.  By 
+straightforward algebra, it is not hard to see WIS has an alternative 
+representation in terms of the predicted quantiles themselves,
+\begin{equation}
+\label{eq:wis_quantiles}
+\WIS(F,Y) = 2 \sum_\tau \phi_\tau(Y - q_\tau), 
+\end{equation}
+where $\phi_\tau(x) = \tau |x|$ for $x \geq 0$ and $\phi_\tau(x) = (1-\tau)
+|x|$ for $x<0$, which is often called the ``tilted absolute'' loss.  While
+\eqref{eq:wis_quantiles} is more general (it can accomodate asymmetric quantile
+levels), the first definition \eqref{eq:wis_intervals} is usually preferred in
+presentation, because the score nicely decouples into a ``sharpness'' component
+(first term in each summand) and ``under/overprediction'' component (second term
+in each summand).  The second form given in \eqref{eq:wis_quantiles} is
+especially noteworthy in our current study as it reveals WIS is the same as the
+quantile regression loss that we use to train our forecast models (in other
+words, our models are fit to optimize WIS average over the training set).
 
-When comparing the aggregate performance of the baseline quantile autoregression
-without additional signals with 
-the indicator-assisted quantile autoregression forecasters, we first
-normalize the individual weighted interval scores by the weighted interval
-score of a boneheaded model: predicting all future values to be the same as the
-most recent observation. The dummy model uses the errors such a forecast has
-made in-sample to form the other quantiles. We normalize by 
-the dummy model to account for the fact that different locations or
-time periods have different magnitudes of COVID-19 incidence (even after
-scaling by population). Performing well by this metric means that a particular
-forecaster is doing well at all locations and time periods. Without the scaling,
-a forecaster that is good at locations and time periods with relatively high
-case numbers but poor everywhere else could appear preferable.
-
-For hotspots, we use the area under the curve (AUC) computed by examining the
-true-positive and false-negative rates across HRRs at each horizon and forecast
-date. That is, for a particular forecast date and a particular horizon, we ask,
-how many true hotspots were called hotspots compared to the total number of
-hotspots (specificity). The false-negative rate (sensitivity) is the ratio of
-correctly classified non-hotspots to the total number of non-hotspots. A perfect
-classifier would have both sensitivity and specificity equal to 1. Plotting the
-sensitivity on the $y$-axis and $1-$specificity on the $x$-axis where each point
-is a forecast date produces an ROC curve for each classifier. We compute the
-area under this curve for each forecast horizon to compare the cases-only model
-to those adding indicators.
+For hotspot prediction, we evaluate the probabilistic classifiers produced by
+the logistic models \eqref{eq:hotspot_ar}, \eqref{eq:hotspot_ar_x} using the
+area under the curve (AUC) of their true positive versus false positive rate
+curve (which is traced out by varying the discrimination threshold).
 

### 2021-06-13 forecast/paper/paper.tex
FINISH edits thru methods

@@
 \includegraphics[width=\columnwidth]{fig/trajectory.pdf}
-\caption{Forecasts and observed cases for HRR x. TODO 
-  \attn{Create a figure here for 1 HRR. Show the fan for the quantiles.}}  
+\caption{TODO}
 \label{fig:trajectory}
@@
 For forecasting, we evaluate the probabilistic forecasts produced by the
-quantile models \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x} using
-\textit{weighted interval score} (WIS), which is a well-known quantile-based
-scoring rule; see, e.g., \cite{Gneiting:2007}.  WIS is a proper score, which
-means that its expectation is minimized by the true quantiles of the target
-variable.  The use of WIS in scoring COVID-19 forecasts is discussed by
-\cite{Bracher:2021}; and WIS serves as the primary metric for forecast
-evaluation in the COVID-19 Forecast Hub.
+quantile models in \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x} using 
+\textit{weighted interval score} (WIS), a quantile-based scoring rule; see e.g.,
+\cite{Gneiting:2007}.  WIS is a proper score, meaning that its expectation
+is minimized by the population quantiles of the target variable.  The use of WIS
+in COVID-19 forecast scoring is discussed in \cite{Bracher:2021}; WIS is also
+the main evaluation metric used in the COVID-19 Forecast Hub.
 
@@
 \eqref{eq:quantile_levels}.  Let $F$ be a forecaster comprised of predicted
-quantiles $q_\tau$ parametrized by a quantile level $\tau$.  In the symmetric
-case, this is equivalent to a collection of central prediction intervals
-$(\ell_\alpha, u_\alpha)$, parametrized by an exclusion probability $\alpha$. 
-The WIS of the forecaster $F$, evaluated at the target variable $Y$, is defined
-by  
+quantiles $q_\tau$ parametrized by a quantile level $\tau$.  In the case of
+symmetric quantile levels, this is equivalent to a collection of central
+prediction intervals $(\ell_\alpha, u_\alpha)$, parametrized by an exclusion
+probability $\alpha$. The WIS of the forecaster $F$, evaluated at the target
+variable $Y$, is defined by:
 \begin{equation}
@@
 straightforward algebra, it is not hard to see WIS has an alternative 
-representation in terms of the predicted quantiles themselves,
+representation in terms of the predicted quantiles themselves:
 \begin{equation}
@@
 especially noteworthy in our current study as it reveals WIS is the same as the
-quantile regression loss that we use to train our forecast models (in other
-words, our models are fit to optimize WIS average over the training set).
+quantile regression loss that we use to train our forecasting models (i.e., our
+models are fit to optimize WIS averaged over the training set).
 
 For hotspot prediction, we evaluate the probabilistic classifiers produced by
-the logistic models \eqref{eq:hotspot_ar}, \eqref{eq:hotspot_ar_x} using the
+the logistic models in \eqref{eq:hotspot_ar}, \eqref{eq:hotspot_ar_x} using the 
 area under the curve (AUC) of their true positive versus false positive rate
 curve (which is traced out by varying the discrimination threshold).
+
+The primary aggregation scheme that we will use in model evaluation and
+comparisons will be to average WIS per forecaster per ahead value $a$, over all
+forecast dates $t$ and locations $\ell$; and similarly, to compute AUC per
+classifier and per ahead value $a$, over all forecast dates $t$ and locations
+$\ell$.
 
@@
 
-TODO Briefly describe that we zero-fill Google-AA.
+Over the prediction period, all auxiliary indicators are available (in the
+proper vintage sense) for all locations and prediction times, except for the
+Google-AA signal, which is only observed for an average of 105 (of 306) HRRs.
+This is due to the fact that that the COVID-19 search trends data is released
+after applying differential privacy techniques \cite{Bavadekar:2020}, and a
+missing signal value means that the level of noise added in the differential
+privacy mechanism is high compared to the underyling search count.  In other
+words, values of the Google-AA signal are clearly \textit{not} missing at
+random.  It seems most appropriate to impute missing values by zero, and this 
+is what we do in our analysis.
+
+\paragraph{Backfill and Nowcasting}
+
+As described previously, the auxiliary indicators defined in terms of medical
+claims, CHNG-CLI, CHNG-COVID, and DV-CLI, undergo a significant and systematic
+pattern of revision, or ``backfill'', after their initial publication.  Given
+their somewhat statistically-regular backfill profiles, it would be reasonable
+to attempt to estimate their finalized values based on vintage data---a problem
+we refer to as \textit{nowcasting}---as a pre-processing step before using them
+as features in the models in \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar_x}. 
+Nowcasting is itself a highly nontrivial modeling problem, and we do not attempt
+it in this paper (though it is a topic of ongoing work in our research group).
+We do note, however, that this would likely improve the performance of the
+models involving claims-based indicators in particular.
 
@@
 
-\paragraph{Backfill, Nowcasting}
-
-TODO Should we discuss the use of ``bad'' case rates as ground truth??
-
-\paragraph{\attn{Methods section graveyard, most for Appendix}}
-
-\begin{itemize}
-\item Discuss the "how many days does the indicator buy" interpretation here or below?
-
-\item Do we explain why regularization/CV not really helpful for both
-forecasting and hotspots
-
-\item Implementation was via quantgen, cite here?
-
-\item some people don't like relative WIS beacuse it's no longer a
-``proper score''.  directly address that here?
-
-\item Do we talk about log loss?
-
-\item Do we explain that we are using infinite-backfilled data?
-
-\end{itemize}
+Some signals display a significant degree of spatial heterogeneity, by which we
+mean the values that they take at different geographic locations are not 
+comparable.  This is true of the Google-AA signal (due to the way in which the 
+underlying search trends time series is self-normalized, see
+\cite{GoogleSymptoms}) and of the claims-based signals (see the discussion in
+our companion paper on the API and database).  Such spatial heterogeneity likely
+hurts the performance of the predictive models that rely on these signals,
+since the models are trained by pooling data from all locations.  In this paper,
+we do not attempt to address this issue (it is again a topic of ongoing work in
+our research group), and we simply note that location-specific effects (or 
+pre-processing to remove spatial bias) would likely improve the performance of
+the models involving Google-AA and the claims-based indicators.
 

### 2021-06-14 indicators/paper/paper.tex
Subsections on viz and forecasting

@@
 
-% RJT: @AR I'm now thinking we should add a very short paragraph or two on
-% this. We can say that we provide various interactive maps highlighting both
-% temporal (time series) and spatial aspects (choropleth or beehive maps), but
-% also exploratory tools for correlations and anomalies.  (Very much like the
-% correlations computed in the next section.)
+Since April 2020, Delphi has been maintaining and continually improving various
+online visualization tools for the COVIDcast indicators \cite{CovidcastViz}.
+These tools fetch data directly from the API, and allow for exploration of both 
+temporal (e.g., time series graphs) and spatial (e.g., choropleth maps) trends
+in the signals, as well as many other aspects, such as correlations, anomalies,
+and backfill.  There is also a dedicated dashboard for exploring results from
+the COVID Tracking and Impact Survey.
 
-% Another reason why I'm in favor of this is that we can now add everybody as a
-% middle author who now contributed anything at all to COVIDcast, from database
-% to API to indicator to visualization.  I'm sure the HCII students & faculty
-% who worked with us would very much appreciate being on this paper.
+\subsection{COVID-19 Forecasting Systems}
+
+Since July 2020, Delphi has been regularly submitting short-term forecasts of
+COVID-19 case and death incidence, at the state and county levels, to the
+COVID-19 Forecast Hub \cite{ForecastHub}, under the name ``CMU-TimeSeries''.
+Building, training, and deploying our forecasting models leverages much of the
+infrastructure described in this paper (such as the COVIDcast API's \textit{as
+  of} feature), and some of our forecasting models rely on auxiliary indicators
+(such as survey-based and claims-based COVID-like illness indicators, described
+below, which we have used in case forecasting).
 
@@
   Doerr:2021}. The API currently serves hundreds of thousands of requests to
-thousands of users, each day.
+thousands of users every day.
 
-In what follows, we present illustrative examples of the usefulness of some of
-the novel signals available in the API. These examples demonstrate that such
-indicators are meaningfully related to COVID activity, that they provide
-alternate views on pandemic activity that are not subject to the same reporting
-glitches and delays as traditional public health surveillance streams, and that
-they provide information about public behavior and attitudes that are not
-available from any other source. Code to reproduce all examples (which uses the
-\texttt{covidcast} R package and fetches data directly from the API) can be
-found at
+In what follows, we present examples of the usefulness of some of the novel
+signals available in the API. These examples demonstrate that such indicators
+are meaningfully related to COVID activity, that they provide alternate views on
+pandemic activity that are not subject to the same reporting glitches and delays
+as traditional public health surveillance streams, and that they provide
+information about public behavior and attitudes that are not available from any
+other source. Code to reproduce all examples (which uses the \texttt{covidcast}
+R package and fetches data directly from the API) can be found at
 \url{https://www.github.com/cmu-delphi/covidcast-pnas/indicators/code/}.
@@
 activity. Five indicators in particular have the closest connections to
-confirmed COVID cases:
+confirmed cases:
 \begin{itemize}
@@
 
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{fig/geo_wise_corr.pdf}
+  \caption{Geo-wise correlations with case rates, from April 15, 2020 to April
+    15, 2021, calculated over all counties for which all signals were available
+    and which had at least 500 cumulative cases by the end of this period.}
+ \label{fig:geo_wise_correlation}
+\end{figure}
+
 To quantify the ability of the signals described above to track trends in COVID
@@
 
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/geo_wise_corr.pdf}
-  \caption{Geo-wise correlations with case rates, from April 15, 2020 to April
-    15, 2021, calculated over all counties for which all signals were available
-    and which had at least 500 cumulative cases by the end of this period.}
- \label{fig:geo_wise_correlation}
-\end{figure}
-
 Figure~\ref{fig:geo_wise_correlation} shows the geo-wise correlations achieved
@@
 
-Figure~\ref{fig:time_wise_correlation} summarizes time-wise correlations of
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{fig/time_wise_correlation.pdf}
+  \caption{Time-wise correlations with case rates, from April 15, 2020 to April
+    15, 2021, calculated over all counties for which all signals were available
+    and which had at least 500 cumulative cases by the end of this period.}
+  \label{fig:time_wise_correlation}
+ \end{figure}
+
+Figure~\ref{fig:time_wise_correlation} summarizes time-wise correlations from 
 these five signals over the same time period, and for the same set of
@@
 There are two remarks deserving attention:
-
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/time_wise_correlation.pdf}
-  \caption{Time-wise correlations with case rates, from April 15, 2020 to April
-    15, 2021, calculated over all counties for which all signals were available
-    and which had at least 500 cumulative cases by the end of this period.}
-  \label{fig:time_wise_correlation}
- \end{figure}
 
@@
   discussion.
-\item It is still quite impressive and surprising that the CTIS-CLI-in-community
-  signal, based on people reporting on the symptoms of others around them, can
-  achieve  nearly as strong time-wise correlations to confirmed cases as can a
-  signal that is based on picking up the occurrence of a confirmed case passing
-  through the outpatient system.
+\item It is still quite impressive (and surprising) that the
+  CTIS-CLI-in-community signal, based on people reporting on the symptoms of
+  others around them, can achieve  nearly as strong time-wise correlations to
+  confirmed cases as can a signal that is based on picking up the occurrence of
+  a confirmed case passing through the outpatient system.
 \end{itemize}
@@
 \subsection{Helping Robustness}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\columnwidth]{fig/bexar_compare.pdf}
+  \caption{Reported cases per day in Bexar County, Texas during the summer of
+    2020. Only July 16, 4,810 backlogged cases were reported, though they
+    actually occurred over the preceding two weeks (this shows up as a prolonged
+    spike in the left panel due to the 7-day trailing averaging applied to the
+    case counts). Daily CTIS estimates of CLI-in-community showed more stable
+    underlying trends.}
+  \label{fig:bexar-compare}
+\end{figure}
 
@@
 
+\subsection{Revisions Matter}
+
 \begin{figure}[t]
 \centering
-  \includegraphics[width=\columnwidth]{fig/bexar_compare.pdf}
-  \caption{Reported cases per day in Bexar County, Texas during the summer of
-    2020. Only July 16, 4,810 backlogged cases were reported, though they
-    actually occurred over the preceding two weeks (this shows up as a prolonged
-    spike in the left panel due to the 7-day trailing averaging applied to the
-    case counts). Daily CTIS estimates of CLI-in-community showed more stable
-    underlying trends.}
-  \label{fig:bexar-compare}
+\includegraphics[width=\columnwidth]{fig/dv_as_of.pdf}
+\caption{Estimated percentage of doctor's visits due to COVID-like illness
+  displayed across multiple issue dates, with later issue dates adding
+  additional data and revising past data from prior issue dates.}
+\label{fig:dv_as_of}
 \end{figure}
-
-\subsection{Revisions Matter}
 
@@
 significantly biased downward in New York throughout June and July 2020.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=\columnwidth]{fig/dv_as_of.pdf}
-\caption{Estimated percentage of doctor's visits due to COVID-like illness
-  displayed across multiple issue dates, with later issue dates adding
-  additional data and revising past data from prior issue dates.}
-\label{fig:dv_as_of}
-\end{figure}
 

### 2021-06-14 forecast/paper/paper.tex
Edits through results summary

@@
 submitted and/or processed late. Correlations between such a signal and case
-rates calculated ``after the fact'' (computed retrospectively, using the
+rates calculated ``after the fact'' (i.e., computed retrospectively, using the
 finalized values of this signal) will not deliver an honest answer to the    
@@
 
-% Furthermore, adopting this practically-oriented forecasting
-% perspective to indicator evaluation requires one to confront
-% fundamental limitations inherent to public health surveillance
-% systems.  In particular, many reporting systems are plagued by
-% latency, with delays in reporting that can sometimes span
-% multiple weeks.  Latency manifests itself in different ways in different data sources.  Some data sources may only make a signal available one week after the date it describes; other
-% data sources may provide immediate imperfect initial estimates, which are
-% gradually ``backfilled'' over a period of days, weeks, or months (in which previous estimates are updated as new
-% data trickle in).  And few of
-% these latency patterns are stable or predictable, with holidays,
-% extreme weather events, and changes in the reporting process all
-% affecting the dynamics of the delay process.  To fully and accurately
-% understand an indicator's usefulness for forecasting requires
-% accounting for the unfortunate reality of data latency.
-
-% Examining the importance of additional features for prediction is a core question
-% in inferential statistics, and there are many ways to examine this question
-% under a variety of assumptions. For time series, work in
-% this area can be traced at least to Granger causality~\cite{GrangerCausality} and
-% related multivariate extensions~\cite{GewekeTimeSeries} under the condition 
-% that the series are linearly related autoregressions among other assumptions.
-% These techniques are especially popular in econometrics and neuroscience~\cite{mccracken2007asymptotics,NeuroscienceGranger}.
-
-% We choose to take a
-% predictive angle, which is simple but also fairly robust, compared to various
-% other classical inferential approaches. In fact, drawing rigorous inference based on
-% predictions, without (or with lean) assumptions, is an active field of
-% research~\cite{dai2021significance,fryer2021shapley,zhang2020floodgate,williamson2020unified}.
-% Nonetheless, the predictive approach directly examines the target of interest: do
-% these signals help to produce forecasts? Furthermore, directly comparing
-% forecast errors without strong assumptions through tests of predictive accuracy
-% are common in epidemiology, econometrics, and other
-% fields~\cite{DieboldMarianoTest,DieboldRetrospective}.
-
-% - then make the broad point that "feature importance" is a core question in
-% statistics, and there are many ways to look at it and lots of literature (cite
-% a bunch fo stuff).  We choose to take a predictive angle, which is simple but
-% also fairly robust, compared to various other classical inferential approaches
-% that make lots of assumptions.  In fact, drawing rigorous inference based on
-% predictions, without (or with lean) assumptions, is an active field of
-% research.  Cite several papers (Jing's conformal paper that describes LOCO,  
-% commentary by Ale, Ryan, Larry in Stat Sci, a bunch of model X stuff including
-% Lukas Janson's recent paper, Eugene + Aaditya's recent paper?).  Larry has put
-% some references in the bib.  Put more if we can find them, especially on time
-% series.
+Finally, it is worth noting that examining the importance of additional features
+for prediction is a core question in inferential statistics and econometrics,
+with work dating back to at least \cite{Granger:1969}. Still today, drawing
+rigorous inference based on predictions, without (or with lean) assumptions, is
+an active field of research from both the applied and theoretical angles;
+see, e.g., \cite{Diebold:2002, McCraken:2007, Diebold:2015, Stokes:2017,
+  Lei:2018, Rinaldo:2019, Williamsom:2020, Zhang:2020, Dai:2021, Fryer:2021}.
+Our take on this problem is in line with much of this literature; however, in
+order to avoid making any explicit assumptions, we do attempt to make explicit
+significance statements, and instead, broadly examine the stability of our
+conclusions with respect to numerous modes of analysis.
 
@@
 
+\begin{figure}[tb!]
+\centering
+\includegraphics[width=\columnwidth]{fig/vintage.pdf}
+\caption{TODO}
+\label{fig:vintage}
+\end{figure}
+
 Vintage training data means different things, in practice, for different
@@
 local authority to figure out how this should best be done), which can result
-in very nontrivial and often unpredictable revisions.
-
-TODO For a figure, use Jacob's ``famous'' example of backfill in DV-CLI?   
-\url{https://cmu-delphi.github.io/covidcast/talks/ml-summit/talk.html#(20)}
+in very nontrivial and often unpredictable revisions.  See Figure
+\ref{fig:vintage} for an example.
 
@@
 
-In addition to the above ``pure'' autoregression model, we consider five
+In addition to the above ``pure'' autoregressive model, we consider five
 probabilistic forecasting models of the form: 
@@
 \end{equation}
-In addition to this ``pure'' autoregression model, we also consider five
+In addition to this ``pure'' autoregressive model, we also consider five
 logistic models of the form: 
@@
 Nowcasting is itself a highly nontrivial modeling problem, and we do not attempt
-it in this paper (though it is a topic of ongoing work in our research group).
-We do note, however, that this would likely improve the performance of the
-models involving claims-based indicators in particular.
+it in this paper (it is a topic of ongoing work in our research group), but we
+note that nowcasting would likely improve the performance of the models
+involving claims-based signals in particular.
 
@@
 hurts the performance of the predictive models that rely on these signals,
-since the models are trained by pooling data from all locations.  In this paper,
-we do not attempt to address this issue (it is again a topic of ongoing work in
-our research group), and we simply note that location-specific effects (or 
-pre-processing to remove spatial bias) would likely improve the performance of
-the models involving Google-AA and the claims-based indicators.
+as we train the models on data pooled over all locations.  In the current 
+paper, we do not attempt to address this issue (it is again a topic of ongoing
+work in our group), and we simply note that location-specific effects
+(or pre-processing to remove spatial bias) would likely improve the performance 
+of the models involving Google-AA and the claims-based indicators.
 
 \section{Results}
+
+We begin with a summary of the high-level conclusions.  Here, and in the
+following subsections, we will use ``AR'' to refer to the pure autoregressive
+model both in forecasting \eqref{eq:forecast_ar} and in hotspot prediction
+\eqref{eq:hotspot_ar} (where the reference to the task should be clear from the
+context). We will also use the name of an auxiliary indicator---``CHNG-CLI'', 
+``CHNG-CLI'', ``CHNG-COVID'', ``CTIS-CLI-in-community'', ``DV-CLI'', or
+``Google-AA''---interchangeably with the model in forecasting
+\eqref{eq:forecast_ar_x} or hotspot prediction \eqref{eq:hotspot_ar_x} that uses
+this particular indicator as a feature (the meaning should be clear from the
+context).  So, e.g., the CHNG-CLI model in forecasting is the one in
+\eqref{eq:forecast_ar_x} that sets $X_{\ell,t}$ to be the value of the CHNG-CLI
+indicator at location $\ell$ and time $t$, fits the coefficients based on the
+training data, and makes its predictions by setting $X_{\ell,t}$ to the
+appropriate CHNG-CLI values.  Finally, we use the term ``indicator model'' to 
+broadly refer to any one of the ten models of the form \eqref{eq:forecast_ar_x},
+\eqref{eq:hotspot_ar_x} (five in each of the forecasting and hotspot prediction
+tasks).
+
+\begin{itemize}
+\item Stratifying predictions by the ahead value ($a=7,\ldots,21$), and
+  aggregating results over the prediction period (early June through end of
+  December 2020), each of the indicator models generally gives a boost in
+  predictive accuracy over the AR model, in both the forecasting and hotspot 
+  prediction tasks.  The gains in accuracy generally attenuate as the ahead 
+  value grows.
+
+\item In the same aggregate view, CHNG-COVID and DV-CLI offer the biggest 
+  gains in each of forecasting and hotspot prediction.  CHNG-CLI is
+  inconsistent: it provides a big gain in hotspot prediction, but little gain 
+  in forecasting (it seems to be hurt by a notable lack of robustness, due
+  to backfill).  CTIS-CLI-in-community and Google-AA each provide decent 
+  gains in forecasting and hotspot prediction.  The former's performance in 
+  forecasting is notable in that it clearly improves on AR even at the largest
+  ahead values.
+
+\item In a more detailed analysis of forecasting performance, the indicator
+  models tend to be worse than AR when case rates are increasing (this is most
+  notable in CHNG-CLI and DV-CLI), and better than AR when case rates are flat
+  or decreasing (most notable in CHNG-COV and CTIS-CLI-in-community).
+  More rarely does an indicator model tend to beat AR when case rates are
+  increasing, but there appears to be some evidence of this for Google-AA. 
+  % In an increasing period, when an indicator model is worse than AR, it tends
+  % to underpredict relative to AR (its median forecast is lower).  In a flat
+  % or decreasing  period, when an indicator model is better than AR, AR tends
+  % to overpredict relative to it (the median AR forecast is higher).
+
+\item In this same analysis, when an indicator model performs better than AR  
+  in a decreasing period, this tends to co-occur with instances in which the
+  indicator ``leads'' case rates (meaning, roughly, on a short-time scale in a
+  given location, its behavior mimics that of case rates some number of days
+   ahead).  On the other hand, neither leading nor lagging behavior tend to
+  co-occur as consistently with an indicator model doing better in periods of
+  increase, or worse in periods of increase or decrease.  
+\end{itemize}
+
+Finally, to quantify the importance of training and making predictions using
+proper vintage data, we ran a parallel set of forecasting and hotspot prediction 
+experiments using finalized data. The results, given in the supplement, show
+that training and making predictions on finalized data can result in grossly
+optimistic estimates of true test-time performance (up to 10\% better in terms
+of average WIS or AUC).
 

### 2021-06-14 indicators/paper/paper.tex
Tweak forecasting text

@@
 
-\subsection{COVID-19 Forecasting Systems}
+\subsection{COVID Forecasting}
 
@@
 COVID-19 Forecast Hub \cite{ForecastHub}, under the name ``CMU-TimeSeries''.  
-Building, training, and deploying our forecasting models leverages much of the
-infrastructure described in this paper (such as the COVIDcast API's \textit{as
-  of} feature), and some of our forecasting models rely on auxiliary indicators
-(such as survey-based and claims-based COVID-like illness indicators, described
-below, which we have used in case forecasting).
+The process of building, training, and deploying our forecasting models
+leverages much of the infrastructure described in this paper (such as the
+COVIDcast API's \textit{as of} feature), and some of our forecasting systems
+rely on auxiliary indicators (such as survey-based and claims-based COVID-like 
+illness signals, described below, which we have used in case forecasting).
 

### 2021-06-15 forecast/paper/paper.tex
Fix logistic typo

@@
 \label{eq:hotspot_ar}
-\logit(Z_{\ell,t+a} \,|\, Y_{\ell,s}, s \leq t)  
-= \alpha^{a,\tau} + \sum_{j=0}^2 \^{a,\tau}_j Y^\Delta_{\ell,t-7j}.  
+\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s}, s \leq t) \big) 
+= \alpha^{a,\tau} + \sum_{j=0}^2 \^{a,\tau}_j Y^\Delta_{\ell,t-7j},
 \end{equation}
+where $\logit(p) = \log(p/(1-p))$, the log-odds of $p$.
+
 In addition to this ``pure'' autoregressive model, we also consider five
@@
 \label{eq:hotspot_ar_x}
-\logit(Z_{\ell,t+a} \,|\, Y_{\ell,s}, X_{\ell,s}, s \leq t)  
+\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s}, X_{\ell,s}, s \leq t) \big)
 = \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j} +  
@@
 appropriate CHNG-CLI values.  Finally, we use the term ``indicator model'' to 
-broadly refer to any one of the ten models of the form \eqref{eq:forecast_ar_x},
-\eqref{eq:hotspot_ar_x} (five in each of the forecasting and hotspot prediction
-tasks).
+refer to any one of the ten models of the form \eqref{eq:forecast_ar_x},
+\eqref{eq:hotspot_ar_x} (five from each of the forecasting and hotspot
+prediction tasks).
 
@@
   aggregating results over the prediction period (early June through end of
-  December 2020), each of the indicator models generally gives a boost in
-  predictive accuracy over the AR model, in both the forecasting and hotspot 
-  prediction tasks.  The gains in accuracy generally attenuate as the ahead 
-  value grows.
+  December 2020), we find that each of the indicator models generally gives a
+  boost in predictive accuracy over the AR model, in both the forecasting and
+  hotspot prediction tasks.  The gains in accuracy generally attenuate as the
+  ahead value grows.
 
@@
 
-\item In a more detailed analysis of forecasting performance, the indicator
-  models tend to be worse than AR when case rates are increasing (this is most
-  notable in CHNG-CLI and DV-CLI), and better than AR when case rates are flat
-  or decreasing (most notable in CHNG-COV and CTIS-CLI-in-community).
+\item In a more detailed analysis of forecasting performance, we find that the
+  indicator models tend to be worse than AR when case rates are increasing (this
+  is most notable in CHNG-CLI and DV-CLI), and better than AR when case rates
+  are flat or decreasing (most notable in CHNG-COVID and CTIS-CLI-in-community).
   More rarely does an indicator model tend to beat AR when case rates are 
-  increasing, but there appears to be some evidence of this for Google-AA. 
+  increasing, but there appears to be some evidence of this for the Google-AA
+  model. 
   % In an increasing period, when an indicator model is worse than AR, it tends
@@
 
-\item In this same analysis, when an indicator model performs better than AR  
-  in a decreasing period, this tends to co-occur with instances in which the
+\item In this same analysis, when an indicator model performs better than AR in
+  a decreasing period, this tends to co-occur with instances in which the
   indicator ``leads'' case rates (meaning, roughly, on a short-time scale in a
@@
 \url{https://www.github.com/cmu-delphi/covidcast-pnas/forecast/code/}.
+
+\subsection{Aggregate Results by Ahead Value}
+
+\subsection{Implicit Regularization Hypothesis}
+
+---which is that 
+  defining features from lagged indicator values in the  models in
+  \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar_x} simply provides a kind of
+  implicit regularization on top of the basic AR model (so the benefits have
+  nothing to do with the actual indicator values)---
+
+To test the ``implicit regularization'' hypothesis (which we explain
+  precisely later), we reran all of the prediction experiments but with
+  $X_{\ell,t}$ in each indicator model replaced by suitable random noise.  This 
+  resulted in vastly different (worse) performance compared to the indicator
+  models fit with actual indicator values, and gave no improvement over the 
+  pure AR model, in both forecasting and hotspot prediction, providing a strong
+  rejection (informally speaking) of the implicit regularization hypothesis.
+
+\subsection{Performance in Up, Down, Flat Periods}
+
+\subsection{Leading-ness or Lagging-ness}
+
+TODO
+
+% TODO: make th
+
+A Primary 
+B Bootstrap
+C up/down/flat (Dan's first histogram) --> table in appendix
+D leading/lagging
 

### 2021-06-15 indicators/paper/paper.tex
Updates throughout, new discussion

@@
 
-\dropcap{P}ublic health decision-makers, healthcare providers, epidemiological
+\dropcap{P}ublic health decision makers, healthcare providers, epidemiological 
 researchers, employers, institutions, and the general public benefit from
 promptly and readily accessible data regarding COVID-19 activity levels,
-countermeasures, and pandemic impact.  Real-time indicators of COVID-19 activity
-levels, such as statistics on cases, deaths, test positivity, and
+countermeasures, and pandemic impact.  Real-time indicators of COVID-19 
+activity levels, such as statistics on cases, deaths, test positivity, and
 hospitalizations, enable reports and interactive dashboard applications for
@@
 a massive database of indicators tracking COVID-19 activity and other relevant
-phenomena, which has been publicly available and continuously updated since
-April 2020. Alongside public data on reported cases and deaths, this database
-includes several unique data streams, including indicators extracted from
-de-identified medical claims data, antigen test results from a major testing
-manufacturer, large-scale public surveys that measure symptoms and public
-behavior, and symptom indicators based on Google search query volume. (We use
-the terms ``indicator'' and ``signal'' interchangeably.) We make aggregate
-signals publicly available, generally at the county level, via the COVIDcast API  
-\cite{CovidcastAPI}. We store and provide access to all previous (historical)
-versions of the signals, a key feature that exposes the effects of data
-revisions.  Lastly, we provide R \cite{CovidcastR} and Python \cite{CovidcastPy}  
-packages to facilitate interaction with the API, and an online dashboard to
-visualize the data \cite{CovidcastViz}.
-
-% RJT: We should fiddle with the bib so that the URLs actually appear in the
-% references at the end.  Otherwise it's kind of pointless to cite manuals or
-% websites like the COVIDcast API site or R/Python manuals.
+phenomena in the United States, which has been publicly available and
+continuously updated since April 2020. Alongside public data on reported cases
+and deaths, this database includes several unique data streams, including
+indicators extracted from de-identified medical claims data, antigen test
+results from a major testing manufacturer, large-scale public surveys that
+measure symptoms and public behavior, and indicators based on particular Google
+search queries. (We use the terms ``indicator'' and ``signal'' interchangeably.)
+We make aggregate signals publicly available, generally at the county level, via
+the COVIDcast API \cite{CovidcastAPI}. We store and provide access to all
+previous (historical) versions of the signals, a key feature that exposes the
+effects of data revisions.  Lastly, we provide R \cite{CovidcastR} and Python
+\cite{CovidcastPy} packages to facilitate interaction with the API, and an
+online dashboard to visualize the data \cite{CovidcastViz}.
 
@@
 \paragraph{Internet-Based Surveys} Conducted in partnership with Facebook,
-Delphi's COVID Tracking and Impact Survey receives about 50,000 responses daily,  
-and has received over 20 million responses since April 2020 \cite{DelphiSurvey,
-  Kreuter:2020}. From the surveys, we construct indicators on symptoms, social 
-distancing, vaccination, and other attitudes and behaviors related to COVID.
+Delphi's COVID Trends and Impact Survey receives an average of 50,000 responses
+daily, and has received over 20 million responses since April 2020
+\cite{DelphiSurvey, Kreuter:2020}. From the surveys, we construct indicators on
+symptoms, social distancing, vaccination, and other attitudes and behaviors
+related to COVID.
 
@@
 \cite{SafeGraphSocial, SafeGraphPatterns}. & 2019-01-01 & County \\
-    COVID Tracking and Impact Survey & COVID symptoms, social distancing
+    COVID Trends and Impact Survey & COVID symptoms, social distancing 
 behaviors, mental health, economic impact, behavior (e.g., mask-wearing,
@@
 COVID, flu, or pneumonia, compiled by the National Center for Health Statistics
-[TODO cite]. & 2020-01-26 & State \\
+\cite{TODO}. & 2020-01-26 & State \\ 
     \bottomrule
@@
 as county or state) and also aggregated to metropolitan statistical areas,
-Health and Human Services regions, and Hospital Referral Regions. National
-averages are also provided. Crucially, each record includes an \textit{issue
-  date} recording when the value was first issued, as described below. This
-allows tracking of multiple revisions to individual observations, each revision
-with a separate issue date.
+Health and Human Services regions, and hospital referral regions. National
+averages are also provided. Crucially, each record is tagged with an
+\textit{issue date} referring to when the value was first issued, as described 
+below. This allows tracking of revisions made to individual observations, as
+each revision is tagged with its own issue date.
 
@@
 effects. Other indicators are available in raw versions and versions smoothed 
-with a seven-day moving average. All processing is done using open-source code
+with a 7-day trailing average. All processing is done using open-source code 
 written primarily in Python and R, and available publicly at
@@
 and backfill.  There is also a dedicated dashboard for exploring results from
-the COVID Tracking and Impact Survey.
+the COVID Trends and Impact Survey.
 
@@
 confirmed cases:
+
 \begin{itemize}
@@
   with confirmed COVID-19, based on the same claims data.
-\item COVID Tracking and Impact Survey CLI (CTIS-CLI): The estimated percentage
+\item COVID Trends and Impact Survey CLI (CTIS-CLI): The estimated percentage
   of the population with COVID-like illness based on Delphi's surveys of
   Facebook users.
-\item COVID Tracking and Impact Survey CLI-in-community (CTIS-CLI-in-community):
-  The estimated percentage of the population who knows someone who is sick,
-  based on Delphi's surveys of Facebook users.
-\item Quidel COVID test positivity rate (Quidel-TPR): The percentage of positive
-  test results among Quidel COVID antigen tests.
+\item COVID Trends and Impact Survey CLI in the community
+  (CTIS-CLI-in-community): The estimated percentage of the population who knows
+  someone in their local community that is sick, based the same surveys.
+\item Quidel test positivity rate (Quidel-TPR): The percentage of positive
+  results among Quidel COVID antigen tests.  
 \end{itemize}
+
 Figure~\ref{fig:time_trends_national} compares the first three of these signals
-to reported COVID cases in the United States (from JHU CSSE, and smoothed with a
-7-day trailing average) over the first year of the pandemic (April 15, 2020 to
-April 15, 2021), illustrating how they  appear to track national trends in cases
-quite well. Importantly, this same relationship persists broadly across multiple
-resolutions, down to smaller geographic regions such as states and counties, as
-shown in the supplement. This will also be illustrated in a more detailed
-correlation analysis in the next subsection.
+to COVID cases in the United States (from JHU CSSE, smoothed with a 7-day
+trailing average) over a year of the pandemic (April 15, 2020 to April 15,
+2021), illustrating how they track national trends quite well.  Importantly, 
+this same relationship persists across multiple resolutions of the data, down to
+smaller geographic regions such as states and counties, as shown in the
+supplement.  This will also be illustrated in a more detailed correlation 
+analysis in the next subsection.
 
+% TODO
 % RJT: @AR Note what I claimed we would do above!  I'm thinking we can just
@@
 
-% RJT: @AR Consider for the supplement.
+% TODO 
+% RJT: @AR Consider for the supplement.  [Or skip it?  Not all that important 
+% but I think it would be nice to do.  Maybe you can delegate this to Logan,
+% Aaron, Jingjing, etc.] 
 % - Hospitalization trends?
@@
 people):
+
 \begin{enumerate}
@@
 Figure~\ref{fig:geo_wise_correlation} shows the geo-wise correlations achieved
-by the five signals and COVID case rates (from USAFacts, and again smoothed
-using a 7-day trailing average), from April 15, 2020 to April 15, 2021. The
-locations we consider here are counties with at least 500 cumulative cases by
-the end of this period, and at which all signals are available (956 counties in
-total). The  significantly positive correlations suggest that these signals 
-could be useful in hotspot detection  (identifying counties with relatively high
-COVID activity, at a given time). Somewhat surprisingly, the CLI-in-community
-signal derived from the survey shows the strongest correlations for much of the
-time period. This clearly demonstrates the value of a large-scale survey such as
-CTIS for tracking symptoms and case trends, especially when other data is
-unavailable.
+by the five signals and COVID case rates (from USAFacts, smoothed using a 7-day
+trailing average), from April 15, 2020 to April 15, 2021. This calculation is
+performed over counties with at least 500 cumulative cases by the end of this 
+period, and at which all indicators are available (956 counties in total). The
+significantly positive correlations suggest that these signals could be useful
+in hotspot detection (identifying counties with relatively high COVID activity,
+at a given time). Somewhat surprisingly, the survey-based CLI-in-community
+signal shows the strongest correlations for much of the time period. This
+clearly demonstrates the value of a large-scale survey such as CTIS for tracking
+symptoms and case trends, especially when other data is unavailable.
 
@@
 CTIS-CLI-in-community is a close second, and the CHNG-CLI signal is third.
-There are two remarks deserving attention:
+There are two noteworthy points:
 
@@
   discussion.
+
 \item It is still quite impressive (and surprising) that the
@@
     underlying trends.}
-  \label{fig:bexar-compare}
+  \label{fig:bexar_compare}
 \end{figure}
@@
 
-As a further example, Figure~\ref{fig:bexar-compare} displays data from Bexar
+As a further example, Figure~\ref{fig:bexar_compare} displays data from Bexar
 County, Texas (which contains San Antonio) during July 2020. On July 16, 2020,
 4,810 backlogged cases were reported after reporting problems prevented them
-from being reported over the past two weeks \cite{Palacios:2021}. This results
-in a clearly visible spike in the left-hand panel of the figure (case data from
-JHU CSSE, smoothed with a 7-day trailing average).  However, Delphi's COVID
-Tracking and Impact Survey averaged around 350 responses per day in Bexar
-County over the same time period, and was able to estimate the fraction
-of the population who know someone in their community with COVID-Like Illness
-(CLI). As we can see in the right-hand panel of the figure, this signal was not
-affected by Bexar County's reporting problems and, as shown in the last
-subsection, it is (in general) highly correlated with case rates, providing an
-alternate stream of data about COVID activity unaffected by backlogs. Similar
-reporting problems have occurred in many jurisdictions across the United States,
-making it valuable to cross-check against external sources not part of the same
-reporting systems.
+from being reported over the past two weeks \cite{Palacios:2021}, resulting in a
+clearly visible spike in the left-hand panel of Figure~\ref{fig:bexar_compare}
+(case data from JHU CSSE, smoothed using a 7-day trailing average).  Meanwhile,
+Delphi's COVID Trends and Impact Survey averaged around 350 responses per day in
+Bexar County over the same time period, and was able to estimate the fraction of
+the population who know someone in their community with COVID-Like Illness
+(CLI). As we can see in the right-hand panel of Figure~\ref{fig:bexar_compare},
+this signal was not affected by Bexar County's reporting problems and, as shown
+in the last subsection, it is (in general) highly correlated with case rates,
+providing an alternate stream of data about COVID activity unaffected by
+backlogs. Similar reporting problems have occurred in many jurisdictions across
+the United States, making it valuable to cross-check against external sources
+not part of the same reporting systems.
 
@@
 
+% TODO
+
 % AR: Potentially interesting things to show or comment on:
@@
 % uptake or hesitancy, distancing behaviors, how worried are you?
-
 % How about this (in each case, each point in said plot is a county's signal
@@
 
-% AR: I think our discussion should also make an argument: that providing many
-% data sources in a common format, with revision tracking, is a valuable goal.
-% Each of our individual data sources has various limitations -- SafeGraph has
-% various biases, the claims data is limited by geographically varying market
-% share, the surveys have demographic biases, etc. So if the argument is solely
-% "We have better data than everyone else," it's a tough sell; if the argument
-% is that our data being convenient to use and accessible in a single unified
-% way makes it more valuable than the individual parts, and that other
-% epidemiological data should be available the same way, then I think we have a
-% strong case.
-% RJT: I agree
+The COVIDcast API provides open access to real-time and geographically-detailed
+indicators that support and enhance standard public health reporting streams in 
+several ways.
 
-% RJT Proposed format for this discussion.
-% 1. COVID indicators can help supplement public health reports by filling out
-% the severity pyramid, improving timeliness, helping with robustness, etc.
-% 2. They also provide info well beyond public health reports (behaviors,
-% attitudes)
-% 3. There is an argument to be made for having data all in one unified
-% convenient access format (with all revisions)
+First, several signals in the API closely track COVID activity (over both time
+and space); yet they are derived from different data streams (e.g., surveys,
+medical insurance claims, medical devices), and are therefore not subject to the
+same sources of error as public health reporting data.  This can be important
+both for downstream modeling purposes (enhancing the accuracy in nowcasting and
+forecasting models; see our companion paper on forecasting) and also for
+robustness (providing sanity checks when faced with potential anomalies in
+standard surveillance streams).
 
-% There are a number of open questions, and challenges to be recognized.  We
-% should be clear in identifying survey bias; geo bias in claims data; and
-% backfill and holiday corrections in claims data.
+Second, the API features many other signals that are relevant to understanding
+aspects of the pandemic and its effects on the United States population that are
+not found in traditional public health streams, e.g., data on mobility patterns,
+internet search trends, mask wearing, and vaccine hesitancy, to name just a few
+(the last two are derived from the COVID Trends and Impact Survey; see our
+companion paper on this survey for a detailed view of its features and
+capabilities).
 
-% Finish with list of iniatives our API can power (like we have now, but
-% compress it quite a lot.)
+Third, the underlying database tracks all revisions made to the data, allowing
+us to query the API to learn ``what was known when'', critical for understanding
+the behavior (and potential pitfalls) of real-time surveillance signals.
 
-% Signals in the COVIDcast API indicating amounts of COVID-19 and
-% COVID-like illness supplement publicly available case and death data by
-% capturing and focusing on different levels of severity and parts of the
-% timelines of COVID-19 infections. For example, the survey-based COVID-like
-% illness indicators may reflect mild or recently-developed COVID-19 infections
-% before or without a visit to a healthcare provider or COVID-19
-% test. Additionally, the reporting pipeline for various digital signals can be
-% quite short. For example, many survey signals are released the morning after the
-% surveys were taken, and describe symptoms experienced just one to two days in
-% the past; initial estimates of outpatient visits and inpatient admissions are
-% released after a few days, generally within a week. The timeliness of these
-% signals may enable users to more quickly detect and predict changes in the
-% direction of COVID-19 activity.
+Finally, we emphasize that unifying many relevant signals into a single common
+format, with comprehensive revision tracking, is a valuable goal in it of
+itself.  No one data source is perfect; each is subject to error.  Bringing
+numerous signals from different data streams all together, and making access as
+convenient as possible for public health researchers, policy makers, modelers,
+etc., is a strong testament to our belief in ``the whole being greater than the
+sum of its parts''.
 
-% TODO move backfill and correlation discussion here?
-%
-% Discuss other applications of the data:
-% - Early warning (informal: just call a human to look at things when the
-%   indicators blow up)
-% - Hotspot prediction (formal: build a model to predict a surge)
-% - Forecasting (refer out to the paper we're writing in parallel)
-% - Nowcasting (i.e., estimating latent infection curve)
-% - Reopening criteria (refer to COVID Exit Strategy?)
-%
-% COVIDcast facilitates several types of real-time applications and retrospective
-% analyses.
-% \begin{itemize}
-% \item
-% \emph{Situational awareness dashboards and reports}
-% summarize activity levels and outcomes, public behavior, vaccination rates, and
-% other relevant statistics.
-% %
-% Features of the COVIDcast API and \texttt{covidcast} packages that are
-% particularly relevant here are the common data format across all signals and the
-% ability to easily request data for individual locations and for less-commonly
-% used geographical resolutions such as Hospital Referral Regions and metropolitan
-% statistical areas, and for newer types of data such as vaccination status.
-% %
-% \item
-% \emph{Early warning} systems flag fast rises in COVID-19 activity indicators;
-% formal variants first define what counts as elevated activity or surges and
-% perform statistical \emph{hotspot predictions}.
-% %
-% For individuals and local-level response, signals with sufficient availability
-% and high time-wise correlations with an outcome of interest (e.g., official case
-% or hospitalization data) are natural choices to monitor.
-% %
-% For resource allocation across locations or drug trial site selection, natural
-% choices would be indicators directly describing the stage of the disease being
-% targeted or signals with high geo-wise correlations to these indicators.
-% %
-% \item
-% \emph{Forecasts} directly predict COVID-19 activity levels using point estimates
-% or probability distributions.
-% %
-% These models may operate at fine geographical and temporal resolutions and/or
-% draw on a wide variety of data, making a common data format and signals offered
-% by county and by date particularly useful.
-% %
-% Revision tracking is especially useful here, helping to explain anomalies in
-% past forecasts and improve future ones.
-% %
-% \item
-% \emph{Nowcasts} estimate current or near-future COVID-19 activity levels or the
-% true number of infections.
-% %
-% For these applications, it is important to have signals with time
-% values that relate to when symptoms started or were present, or when a healthcare
-% visit occurred, rather than to the time that these values were eventually
-% reported, unless there is a fixed reporting delay or auxiliary data regarding the
-% distribution of reporting delays.
-% %
-% For example, time values for survey data in COVIDcast relates to when symptoms
-% were occurring, doctor visit data indicates the date of the visit, and inpatient
-% data indicates the date of admission.
-% %
-% This contrasts with nationwide aggregated case and death data, which is
-% usually aligned by the date that a case or death was reported and subject to
-% variable delays based on testing availability and location-specific reporting pipelines.
-% %
-% % TODO antibodies, revisions, positivity
-% \item
-% \emph{Reopening criteria} are informed by COVID-19 activity levels, public
-% behavior, and vaccination levels.  Methodologies to develop and evaluate these
-% criteria may require such information, as well as additional data on timing of
-% public health interventions and on confounding variables, in order to draw valid
-% conclusions or to improve estimated effects of potential policies.
-% \end{itemize}
-
-% TODO refer out to the paper we're writing in parallel
+There are a number of open questions, and challenges that remain.  
+% TODO:
+% RJT: we should write quick bullet points on 1. survey bias, 2. geo bias in
+% claims data, 3. backfill and holiday corrections in claims data. 
+% @AR: can you give this a stab?  You can also delegate (Aaron e.g. should be
+% able to write something on the claims stuff.)  Let's keep it succint!
 

### 2021-06-15 indicators/paper/paper.tex
Fix typo

@@
 \item COVID Trends and Impact Survey CLI in the community
-  (CTIS-CLI-in-community): The estimated percentage of the population who knows
+  (CTIS-CLI-in-community): The estimated percentage of the population who know 
   someone in their local community that is sick, based the same surveys.

### 2021-06-15 forecast/paper/supplement.Rmd
add code, begin supplement generation

---
bibliography: [../../common/covidcast.bib]
output: 
  pdf_document:
    keep_tex: true
    template: pnas-suppl-template.tex
params:
  fd_casefloor: 30
  flag_jumps: 15
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 5, fig.height = 3,
                      fig.align = "center",
                      fig.path = "fig/",
                      cache = TRUE, 
                      autodep = TRUE,
                      message = FALSE, 
                      warning = FALSE)
source("eval_funs.R")
source("hotspot_funs.R")
source("trajectory_plots.R")
library(scales)
path_to_data <- "~/Documents/COVID-Delphi/fcast-indicators-paper/temp_data/"
```

```{r data-loading}
actuals <- readRDS(file.path(path_to_data, "actuals.RDS"))
fcasts_honest <- readRDS(file.path(path_to_data, "results_honest.RDS")) %>%
  process_res_cases(actuals) %>%
  filter(forecaster != "gs_inherit") 
fcasts_finalized <- readRDS(file.path(path_to_data, "results_dishonest.RDS")) %>%
  process_res_cases(actuals) %>%
  filter(forecaster != "gs_inherit")
hotspots_honest <- readRDS(file.path(path_to_data, "hotspots_honest.RDS")) %>%
  process_res_hotspots() %>%
  filter(forecaster != "gs_inherit")
hotspots_finalized <- readRDS(
  file.path(path_to_data, "hotspots_dishonest.RDS")) %>%
  process_res_hotspots() %>%
  filter(forecaster != "gs_inherit")
cases7dav <- readRDS(file.path(path_to_data, "cases7dav.RDS")) %>%
  rename(case_num = value)
if (params$fd_casefloor >= 0) {
  fcasts_honest <- filter_case_floor(fcasts_honest, cases7dav, params)
  fcasts_finalized <- filter_case_floor(fcasts_finalized, cases7dav, params)
}
if (params$flag_jumps > 0) {
  fcasts_honest <- filter_jumps(fcasts_honest, actuals, params)
  fcasts_finalized <- filter_jumps(fcasts_finalized, actuals, params)
  hotspots_honest <- filter_jumps(hotspots_honest, actuals, params)
  hotspots_finalized <- filter_jumps(hotspots_finalized, actuals, params)
}
```

```{r gs-processing}
# only honest
fcasts_gs <- intersect_averagers(
  fcasts_honest, "forecaster", 
  c("geo_value", "ahead", "forecast_date", "target_end_date")) %>%
  filter(forecaster != "gs_subset") %>%
  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
fcasts_honest <- filter(fcasts_honest, forecaster != "gs_subset") %>%
  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
fcasts_finalized <- filter(fcasts_finalized, forecaster != "gs_subset")%>%
  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
hotspots_gs <- intersect_averagers(
  hotspots_honest %>% filter(!is.na(value), !is.na(actual)), 
  "forecaster", 
  c("geo_value", "ahead", "forecast_date", "target_end_date")) %>%
  filter(forecaster != "gs_subset") %>%
  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
hotspots_honest <- intersect_averagers( # Deal with DV NaN's
  hotspots_honest %>% 
    filter(! is.na(value), !is.na(actual), forecaster != "gs_subset"), 
  "forecaster", 
  c("geo_value", "ahead", "forecast_date", "target_end_date")) %>%
  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
hotspots_finalized <- intersect_averagers( 
  hotspots_finalized %>% 
    filter(! is.na(value), !is.na(actual), forecaster != "gs_subset"), 
  "forecaster", 
  c("geo_value", "ahead", "forecast_date", "target_end_date")) %>%
  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
```

## Forecasts

```{r fcast}
fig5a <- plotter(fcasts_honest %>% filter(period != "jm"), 
        "wis", Mean, scaler = "strawman_wis", 
        order_of_operations = c("aggr","scale")) +
  ylab("Mean WIS (relative to baseline)")
fig5a
```

```{r fcast-finalized}
plotter(fcasts_finalized %>% filter(period != "jm"), 
        "wis", Mean, scaler = "strawman_wis", 
        order_of_operations = c("aggr","scale")) +
  ylab("Mean WIS (relative to baseline)")
```

```{r fcast-honest-v-finalized}
fig6a <- plotter(
  left_join(
    fcasts_honest %>% filter(period != "jm") %>% select(forecaster:wis),
    fcasts_finalized %>% filter(period != "jm") %>% select(forecaster:wis) %>%
      rename(finalized_wis = wis)
    ), 
  "wis", Mean, scaler = "finalized_wis", 
  order_of_operations = c("aggr","scale")) +
  geom_hline(yintercept = 1) +
  ylab("Mean WIS (vintage / finalized)")
fig6a
```

```{r fcast-adjusted}
plotter(fcasts_honest %>% filter(period != "jm"), 
        "wis", GeoMean, scaler = "strawman_wis", 
        order_of_operations = c("scale","aggr")) +
  ylab("Geometric mean of WIS (relative to baseline)")
```

```{r wis-densities}
ggplot(fcasts_honest %>% filter(period != "jm"), 
       aes(wis, fill = forecaster)) +
  geom_density() +
  scale_x_log10() +
  theme_bw(base_size = 12) +
  xlab("WIS") +
  scale_fill_brewer(palette = "Set1") +
  facet_wrap(~forecaster) +
  theme(legend.position = "none")
```

### All periods

```{r fcast-alldates}
plotter(fcasts_honest,
        "wis", Mean, scaler = "strawman_wis", 
        order_of_operations = c("aggr","scale")) +
  ylab("Mean WIS (relative to baseline)")
```

```{r fcast-alldates-adjusted}
plotter(fcasts_honest, 
        "wis", GeoMean, scaler = "strawman_wis", 
        order_of_operations = c("scale","aggr")) +
  ylab("Geometric mean of WIS (relative to baseline)")
```

### Only GS locations

```{r fcast-gs-locations}
plotter(fcasts_gs %>% filter(period != "jm"),
        "wis", Mean, scaler = "strawman_wis", 
        order_of_operations = c("aggr","scale")) +
  ylab("Mean WIS (relative to baseline)")
```

```{r fcast-gs-locations-adjusted}
plotter(fcasts_gs %>% filter(period != "jm"), 
        "wis", GeoMean, scaler = "strawman_wis", 
        order_of_operations = c("scale","aggr")) +
  ylab("Geometric mean of WIS (relative to baseline)")
```

## Trajectories

```{r hrr-names}
hrr_names <- read_csv(file.path(path_to_data, "Hospital_Referral_Regions.csv"), 
                      col_types = "_ccc__") %>%
  mutate(state = substr(hrrcity, 1, 2),
         hrrname = paste(HRR_lbl, state)) %>%
  select(hrrnum, hrrname)
hrr_tab <- pull(hrr_names, hrrname)
names(hrr_tab) <- pull(hrr_names, hrrnum)
```

```{r traj-data, eval = FALSE}
preds <- readRDS(file.path(path_to_data, "predictions_honest.RDS"))
traj_best <- get_trajectory_plots(fcasts_honest, preds, actuals, hrr_tab, 
                                  "only2020", "best")
traj_worst <- get_trajectory_plots(fcasts_honest, preds, actuals, hrr_tab,
                                   "only2020", "worst")
rm(preds)
```

```{r trajectory-plots, eval = FALSE}
for (nm in names(traj_best)) print(trajectory_panel(nm, traj_best, traj_worst))
```

## Hotspots

```{r hot}
fig5b <- plotter_hotspots(hotspots_honest %>% filter(period != "jm")) + 
  geom_hline(yintercept = 0.5)
fig5b
```

```{r hot-finalized}
plotter_hotspots(hotspots_finalized %>% filter(period != "jm")) +
  geom_hline(yintercept = 0.5)
```

```{r hot-honest-v-finalized}
fig6b <- left_join(
  hotspots_honest %>%
    group_by(forecaster, ahead) %>%
    summarise(auc_honest = auc(value, actual)),
  hotspots_finalized %>%
    group_by(forecaster, ahead) %>%
    summarise(auc_finalized = auc(value, actual))) %>%
  ggplot(aes(ahead, auc_honest / auc_finalized, color = forecaster)) + 
  geom_line() + geom_point() +
  theme_bw(base_size = 12) + 
  scale_color_brewer(palette = "Set1", guide = guide_legend(nrow=1)) +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  geom_hline(yintercept = 1) +
  ylab("AUC (vintage / finalized)")
fig6b
```

### All periods

```{r hot-alldates}
plotter_hotspots(hotspots_honest)
```

### Only GS locations

```{r hot-gs-locations}
plotter_hotspots(hotspots_gs %>% filter(period != "jm")) +
  geom_hline(yintercept = 0.5)
```

### Figure 5

* Currently, this is the "main" figure. WIS and hotspots, 2 panels, only 2020, honest training.

```{r figure5, eval = FALSE, fig.width=9, fig.height=4}
library(cowplot)
leg <- get_legend(fig5a)
fig5a <- fig5a + theme(legend.position = "none")
fig5b <- fig5b + theme(legend.position = "none")
fig5 <- plot_grid(fig5a, fig5b, align = "hv", nrow = 1)
fig5 <- plot_grid(fig5, leg, rel_heights = c(1,.1), nrow = 2)
fig5
```

### Figure 6 (misleading retrospective analysis)

* WIS and hotspots, 2 panels, only 2020, honest training.

```{r figure6, fig.width=9, fig.height=4}
fig6a <- fig6a + theme(legend.position = "none")
fig6b <- fig6b + theme(legend.position = "none")
fig6 <- plot_grid(fig6a, fig6b, align = "hv", nrow = 1)
fig6 <- plot_grid(fig6, leg, rel_heights = c(1,.1), nrow = 2)
fig6
```

### Correlations with lagged actuals

```{r cor-with-actuals}
pct_change <- function(x, n = 14, col_name = "pct_change"){
  # courtesy of Alden
  dt <- c(0,-n)
  y = covidcast:::apply_shifts(x,dt) %>%
    rename(value = tidyselect::matches("value\\+"),
           lag_value = tidyselect::matches("value\\-")) %>%
    mutate(`:=`(!!col_name,(value - lag_value)/lag_value))
  return(y)
}
comparison_df <- left_join(
  fcasts_honest %>% filter(forecaster != "QAR3", period != "jm"),
  fcasts_honest %>% filter(forecaster == "QAR3", period != "jm") %>%
    select(-c(forecaster,target_end_date)) %>%
    rename(ar3_wis = wis),
  by = c("geo_value", "ahead", "forecast_date")) %>%
  group_by(forecaster, forecast_date, geo_value) %>%
  summarize(wis_ratio = mean(wis) / mean(ar3_wis), .groups = "drop")

pct_chng_df <- readRDS(file.path(
  path_to_data, 
  "jhu-csse_confirmed_7dav_incidence_num_2021-05-18.RDS")) %>%
  pct_change(n = 7) %>%
  select(geo_value, time_value, pct_change) %>%
  filter(time_value %in% unique(comparison_df$forecast_date))

comparison_df <- comparison_df %>% 
  left_join(pct_chng_df, 
            by = c("forecast_date" = "time_value", "geo_value"))

pct_chng_df <- comparison_df %>% 
  group_by(forecaster, geo_value) %>%
  summarize(
    cor_wisratio_pctchange = cor(wis_ratio, pct_change, method = "spearman"),
    cor_abs_wisratio_minus_1_pctchange = cor(
      abs(wis_ratio - 1), pct_change,method = "spearman"), .groups = "drop")
```

```{r cor-wis-ratio}
ggplot(pct_chng_df %>% 
         group_by(forecaster) %>% 
         mutate(median = median(cor_wisratio_pctchange)), 
       aes(x = cor_wisratio_pctchange)) +
  geom_histogram(aes(y = ..count.. / sum(..count..), fill = forecaster)) +
  scale_fill_brewer(palette = "Set1") +
  theme_bw(base_size = 14) +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  geom_vline(aes(xintercept = median)) + 
  geom_vline(xintercept = 0, linetype = "dotted") +
  facet_wrap(~forecaster) + 
  xlab("Spearman correlation") + ylab("relative frequency") +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1))
```

```{r cor-wis-ratio-m1}
ggplot(pct_chng_df %>% 
         group_by(forecaster) %>% 
         mutate(median = median(cor_abs_wisratio_minus_1_pctchange)), 
       aes(x = cor_abs_wisratio_minus_1_pctchange)) +
  geom_histogram(aes(y = ..count.. / sum(..count..), fill = forecaster)) +
  scale_fill_brewer(palette = "Set1") +
  theme_bw(base_size = 14) +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  geom_vline(aes(xintercept = median)) + 
  geom_vline(xintercept = 0, linetype = "dotted") +
  facet_wrap(~forecaster) + 
  xlab("Spearman correlation") + ylab("Relative frequency") +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1))
```

## Upswings vs downswings

```{r up-down-processing}
preds <- readRDS(file.path(path_to_data, "predictions_honest.RDS"))
preds <- preds %>% 
  filter(abs(quantile - 0.5) < 1e-6) %>% 
  select(-quantile) %>%
  mutate(forecaster = recode(forecaster,
                             AR3 = "AR",
                             AR3CHCLI3 = "CHNG-CLI",
                             AR3CHCOV3 = "CHNG-COV",
                             AR3DVCLI3 = "DV-CLI",
                             AR3FBCLI3 = "CTIS-CLI-in-community",
                             AR3GSSAA3_Subset = "gs_subset",
                             AR3GSSAA3_Zero = "Google-AA",
                             Baseline = "strawman"
  )) %>%
  filter(! forecaster %in% c("gs_subset", "strawman"))

up_down <- actuals %>%
  arrange(geo_value, target_end_date) %>%
  group_by(geo_value) %>%
  left_join(cases7dav, c("target_end_date" = "time_value", "geo_value")) %>%
  mutate(pct_change = (actual - lag(actual,7)) / lag(actual,7),
         udf = case_when(
           pct_change >= .25 ~ "up",
           pct_change <= -.25 ~ "down",
           TRUE ~ "flat")) %>%
  filter(case_num > 30, !is.na(pct_change)) #%>%
  #rename(forecast_date = target_end_date)

corr_df <- left_join(fcasts_honest, preds, 
                     by = c("geo_value", "ahead", "forecaster", 
                            "forecast_date")) %>%
  select(-actual_fd, -starts_with("strawman"), -ae) %>%
  rename(med = value)
rm(preds)

df <- left_join(corr_df %>% 
                  ungroup() %>%
                  filter(forecaster != "AR"),
                corr_df %>% 
                  ungroup() %>%
                  filter(forecaster == "AR") %>%
                  select(-forecaster, -actual_td) %>%
                  rename(AR_wis = wis, AR_med = med)) %>%
  inner_join(up_down, by = c("geo_value","target_end_date"))
```

```{r upswing-histogram}
df %>% 
  filter(period != "jm") %>%
  group_by(forecaster, geo_value, udf) %>%
  ggplot(aes(AR_wis - wis, fill = forecaster)) +
  geom_histogram(bins = 100) +
  facet_grid(udf ~ forecaster) +
  theme_bw(base_size = 14) +
  theme(legend.position = "none") +
  scale_fill_brewer(palette = "Set1") +
  scale_y_log10(breaks = c(10,1000,100000), 
                labels = trans_format("log10", math_format(10^.x))) +
  xlab("AR WIS - forecaster WIS") +
  geom_vline(xintercept = 0)
```

```{r upswing-histogram-logged}
df_summary <- df %>%
  filter(period != "jm") %>%
  group_by(forecaster, udf) %>%
  summarise(m = GeoMean((AR_wis + 1) / (wis + 1)))
df %>% 
  filter(period != "jm") %>%
  group_by(forecaster, udf) %>%
  ggplot(aes((AR_wis + 1) / (wis + 1), fill = forecaster)) +
  geom_histogram(bins = 100) +
  facet_grid(udf ~ forecaster) +
  theme_bw(base_size = 14) +
  theme(legend.position = "none") +
  scale_fill_brewer(palette = "Set1") +
  scale_x_log10() +
  scale_y_log10(breaks = c(10,1000,100000), 
                labels = trans_format("log10", math_format(10^.x))) +
  xlab("AR WIS - forecaster WIS") +
  geom_vline(xintercept = 1) +
  geom_vline(data = df_summary, aes(xintercept = m), linetype = "dotted")
```

```{r upswing-corr-table}
df %>% 
  filter(period != "jm") %>% 
  group_by(forecaster, udf) %>%
  summarise(cor = cor(AR_wis - wis, AR_med - med)) %>%
  pivot_wider(names_from = forecaster, values_from = cor) %>%
  kableExtra::kbl(booktabs = TRUE, digits = 2)
```

## Leadingness and laggingness

```{r lead-lag-processing}
# see lead-lag-analysis.R for signal downloading and processing
lead_lag <- function(x, y, lag.max = NULL, na.len = 14) {
  if (sum(is.na(x)) >= na.len || sum(is.na(y)) >= na.len) {
    return(c(leading = NA, lagging = NA))
  }
  out <- ccf(x, y, lag.max = lag.max, plot = FALSE, na.action = na.omit)
  clim <- qnorm((1 + .95) / 2) / sqrt(out$n.used)
  lag_vec <- drop(out$lag)
  cc <- drop(out$acf)
  cc <- cc * (cc > clim)
  return(c(leading = mean(cc[lag_vec < 0]),
           lagging = mean(cc[lag_vec > 0])))
}

slide_lead_lag <- function(x, y, min_len = 56, lag.max = NULL, na.len = 14) {
  n <- length(x)
  outlist <- list()
  idx <- 1:min_len
  for (i in min_len:n) {
    outlist[[i]] <- lead_lag(x[idx], y[idx], lag.max, na.len)
    idx <- idx + 1
  }
  outlist <- bind_rows(outlist)
  bind_rows(
    tibble(leading = rep(NA, min_len-1), lagging = rep(NA, min_len -1)),
    outlist)
}

all_sigs <- readRDS(file.path(
  path_to_data, "all_signals_wide_as_of_2020-05-15.RDS"))

all_sigs_imputed <- all_sigs %>% 
  group_by(geo_value) %>% 
  arrange(time_value) %>%
  mutate(across(contains("value+0"), 
                ~zoo::na.locf(.x, na.rm = FALSE, maxgap = 13))) %>%
  pivot_longer(contains("value+0"))

sigs_cor <- left_join(
  all_sigs_imputed %>% filter(! str_detect(name, "jhu-csse")),
  all_sigs_imputed %>% filter(str_detect(name, "jhu-csse")) %>%
    select(-name) %>%
    rename(cases = value))

lead_lag_metric <- sigs_cor %>% 
  group_by(geo_value, name) %>%
  arrange(time_value) %>%
  group_modify(~ {
    bind_cols(time_value = .x$time_value, slide_lead_lag(.x$value, .x$cases))
  })

lead_lag_metric <- lead_lag_metric %>%
  mutate(
    forecaster = recode(
      name,
      `value+0:chng_smoothed_adj_outpatient_cli` = "CHNG-CLI",
      `value+0:chng_smoothed_adj_outpatient_covid` = "CHNG-COV",
      `value+0:doctor-visits_smoothed_adj_cli` = "DV-CLI",
      `value+0:fb-survey_smoothed_hh_cmnty_cli` = "CTIS-CLI-in-community",
      `value+0:google-symptoms_sum_anosmia_ageusia_smoothed_search` = "Google-AA")) %>%
  ungroup() %>%
  select(-name)

df2 <- left_join(
  df, 
  lead_lag_metric,
  by = c("geo_value","forecaster","target_end_date" = "time_value"))
```

```{r leading-and-lagging, fig.width=8, fig.height=4}
df2 %>%
  filter(period != "jm") %>%
  group_by(forecaster, udf) %>%
  summarise(
    leadingness = cor(AR_wis - wis, leading, use = "pairwise.complete.obs"),
    laggingness = cor(AR_wis - wis, lagging, use = "pairwise.complete.obs")) %>%
  pivot_longer(leadingness:laggingness) %>%
  ggplot(aes(udf, value, fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  geom_hline(yintercept = 0) +
  scale_fill_brewer(palette = "Set1") +
  facet_wrap(~name) +
  theme_bw(base_size = 14) +
  ylab("correlation") +
  xlab("") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

```{r diff-in-lead-lag, fig.width=8, fig.height=4}
df2 %>%
  filter(period != "jm") %>%
  group_by(forecaster, udf) %>%
  summarise(cor = cor(AR_wis - wis, leading - lagging, 
                      use = "pairwise.complete.obs")) %>%
  ggplot(aes(udf, cor, fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  geom_hline(yintercept = 0) +
  scale_fill_brewer(palette = "Set1") +
  theme_bw(base_size = 14) +
  ylab("correlation") +
  xlab("") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

### 2021-06-15 forecast/code/ccf-dv-finalized.R

@@ -1,0 +1,42 @@
+library(covidcast)
+library(tidyverse)
+
+sigs <- covidcast_signals(
+  data_source = c("jhu-csse", "doctor-visits"),
+  signal = c("confirmed_7dav_incidence_prop", "smoothed_adj_cli"),
+  start_day = "2020-10-01",
+  end_day = "2020-12-15",
+  as_of = "2021-05-15", #2020-12-21",
+  geo_type = "hrr",
+  geo_values = "303") %>% # NYC
+  aggregate_signals(format = "wide")
+
+nas <- complete.cases(sigs)
+sigs <- sigs[nas,]
+
+names(sigs)[3:4] <- c("cases", "doctor-visits")
+
+std_sigs <- sigs %>%
+  mutate(across(c("cases","doctor-visits"), ~ (.x - mean(.x)) / sd(.x))) %>%
+  select(-geo_value) %>%
+  pivot_longer(-time_value)
+  
+g1 <- std_sigs %>%
+  ggplot(aes(time_value, value, color = name)) +
+  geom_line() +
+  theme_bw(base_size = 14) +
+  xlab("date") + ylab("standardized signal") +
+  scale_color_viridis_d(begin=.25, end=.75, name = "") +
+  theme(legend.position = "right")
+
+cc <- ccf(sigs$`doctor-visits`, sigs$cases, plot = FALSE)
+g2 <- ggplot(tibble(lag = drop(cc$lag), ccf = drop(cc$acf)), aes(lag)) +
+  geom_segment(aes(xend = lag, yend = ccf), y = 0, color = "cornflowerblue") +
+  geom_hline(yintercept = c(1,-1) * qnorm((1 + .95) / 2) / sqrt(cc$n.used),
+             color = "darkorange", linetype = "dashed") +
+  geom_hline(yintercept = 0) +
+  geom_vline(xintercept = 0) +
+  ylab("cross correlation") +
+  theme_bw(base_size = 14) 
+cowplot::plot_grid(g2, g1, rel_widths = c(4,6))
+ggsave("../gfx/ccf-dv-finalized.pdf", width = 8, height = 4)

### 2021-06-15 forecast/code/eval_funs.R

@@ -1,0 +1,192 @@
+library(tidyverse)
+library(lubridate)
+library(evalcast)
+
+Mean <- function(x) mean(x, na.rm = TRUE)
+Median <- function(x) median(x, na.rm = TRUE)
+TrimMean01 <- function(x) mean(x, trim = .01, na.rm = TRUE)
+TrimMean025 <- function(x) mean(x, trim = .025, na.rm = TRUE)
+TrimMean05 <- function(x) mean(x, trim = .05, na.rm = TRUE)
+TrimMean1 <- function(x) mean(x, trim = .1, na.rm = TRUE)
+GeoMean <- function(x) exp(mean(log(x), na.rm = TRUE))
+
+summarizer <- function(df, y, centerer, scaler, aggr, order_of_operations) {
+  for (i in seq_along(order_of_operations)) {
+    df <- switch(
+      order_of_operations[i],
+      center = {
+        if (is.null(centerer)) df
+        else mutate(df, !!y := !!sym(y) - !!sym(centerer))},
+      scale = {
+        if (is.null(scaler)) df
+        else mutate(df, across(all_of(c(y, centerer)), ~.x / !!sym(scaler)))},
+      aggr = summarise(df, across(all_of(c(y, centerer, scaler)), ~aggr(.x))))
+  }
+  return(df)
+}
+
+
+plotter <- function(df, y, aggr,
+                    centerer = NULL, scaler = NULL, 
+                    facet_by_period = FALSE,
+                    order_of_operations = c("center", "scale", "aggr")) {
+  df <- df %>% group_by(ahead, forecaster)
+  facet_layer <- NULL
+  if (facet_by_period) {
+    period_labels <- c(jj = "June - July", as = "Aug - Sep", o = "Oct",
+                       nd = "Nov - Dec", jm = "Jan - Mar")
+    df <- df %>% group_by(period, .add = TRUE)
+    facet_layer <- facet_wrap(~period, scales = "free_y", 
+                              labeller = labeller(period = period_labels))
+  }
+  
+  df <- summarizer(df, y, centerer, scaler, aggr, order_of_operations)
+  
+  
+  ggplot(df, aes(ahead, wis, color = forecaster)) + 
+    geom_line() + geom_point() +
+    facet_layer +
+    theme_bw(base_size = 14) + 
+    scale_color_brewer(palette = "Set1", guide = guide_legend(nrow = 1)) +
+    theme(legend.position = "bottom", legend.title = element_blank())
+}
+
[…]

### 2021-06-15 forecast/code/hotspot-contingency-table.R

@@ -1,0 +1,35 @@
+la <- hotspots_honest %>%
+  filter(geo_value == "56", period != "jm") %>%
+  left_join(actuals %>% filter(geo_value == "56") %>% rename(observed = actual))
+
+la %>%
+  ggplot(aes(target_end_date, value, color = ahead)) +
+  geom_vline(data = la %>% filter(actual == 1), aes(xintercept = target_end_date),
+             color = "orange", size = .2) +
+  geom_point() +
+  scale_fill_viridis_c() +
+  facet_wrap(~forecaster, ncol = 1) +
+  theme_bw()
+
+hot_udf <- inner_join(hotspots_honest %>% filter(period != "jm"),
+                      up_down %>% select(geo_value, target_end_date, udf))
+n_expected <- hot_udf %>% group_by(udf,forecaster) %>% summarise(n_forecasts = n())
+
+
+tab <- hot_udf %>% 
+  mutate(deciles = ntile(value, 10)) %>%
+  group_by(udf, deciles, forecaster) %>%
+  summarise(m = n()) %>%
+  left_join(n_expected) %>%
+  mutate(expected = n_forecasts * .1, deviation = m / expected)
+
+ggplot(tab, aes(deciles, deviation, color = forecaster)) +
+  geom_line() +
+  geom_point() +
+  geom_hline(yintercept = 1) +
+  facet_wrap(~udf, ncol = 1) +
+  scale_color_brewer(palette = "Set1", guide = guide_legend(nrow = 1)) +
+  scale_x_continuous(breaks = 1:10, labels = 1:10) +
+  theme_bw() +
+  ylab("predictions / expected") +
+  theme(legend.position = "bottom", legend.title = element_blank())

### 2021-06-15 forecast/code/hrr_population.R

@@ -1,0 +1,28 @@
+library(tidyverse)
+fips_pop <- read_csv("https://raw.githubusercontent.com/cmu-delphi/covidcast-indicators/main/_delphi_utils_python/delphi_utils/data/fips_pop.csv",
+                     col_types = "ci")
+hrrtab <- read_csv("https://raw.githubusercontent.com/cmu-delphi/covidcast-indicators/main/_delphi_utils_python/delphi_utils/data/fips_hrr_table.csv",
+                   col_types = "ccd")
+hrrs <- left_join(hrrtab, fips_pop, by = "fips")
+hrrs <- mutate(hrrs, wpop = pop * weight)
+hrr_pop <- hrrs %>%
+  group_by(hrr) %>%
+  summarise(pop = sum(wpop, na.rm=TRUE))
+# note, 02270, 46113, 51515 have pop NA. These seem no longer to exist (renumbered or removed)
+
+hrr_names <- read_csv("Hospital_Referral_Regions.csv", col_types = "_ccc__") %>%
+  mutate(state = substr(hrrcity, 1, 2))
+hrr_pop <- left_join(hrr_pop, hrr_names, by = c("hrr" = "hrrnum")) %>%
+  mutate(name = paste(HRR_lbl,State)) %>%
+  select(hrr, name, pop)
+
+ggplot(hrr_pop, aes(pop)) +
+  geom_histogram(aes(y=(..count..)/sum(..count..)), bins = 20, fill = "cornflowerblue") +
+  xlab("population") + ylab("frequency") +
+  scale_x_log10(labels = scales::label_comma()) +
+  scale_y_continuous(expand = expansion(c(0,.05)),
+                     labels = scales::label_percent(accuracy = 1),
+                     breaks = scales::breaks_pretty(4)) +
+  theme_bw(base_family = "Times")
+
+ggsave("../gfx/hrr_pop_histogram.pdf", width=6, height=3)

### 2021-06-15 forecast/code/ny-trajectory.R

@@ -1,0 +1,38 @@
+library(tidyverse)
+library(lubridate)
+
+ny_preds <- readRDS("../temp_data/predictions_honest.RDS") %>%
+  filter(geo_value == "303", 
+         forecast_date == "2020-10-15",
+         forecaster == "AR3") %>%
+  mutate(target_end_date = forecast_date + days(ahead),
+         incidence_period = "day")
+
+actuals <- readRDS("../temp_data/actuals.RDS") %>%
+  filter(geo_value == "303", 
+         target_end_date > "2020-09-01", 
+         target_end_date < "2020-11-15") %>%
+  rename(value = actual, time_value = target_end_date)
+
+pd <- evalcast:::setup_plot_trajectory(ny_preds, side_truth = actuals)
+pd$truth_df <- pd$truth_df %>% rename(target_end_date = time_value)
+
+ggplot(pd$truth_df, mapping = aes(x = target_end_date)) +
+  geom_ribbon(
+    data = pd$quantiles_df,
+    mapping = aes(ymin = lower, ymax = upper, fill = interval)) +
+  scale_fill_brewer(palette = "Blues") +
+  geom_line(aes(y = value)) + # reported
+  geom_point(aes(y = value)) + # reported gets dots
+  geom_line(data = pd$points_df, 
+            mapping = aes(y = value),
+            color = "orange", size = .5) +
+  geom_point(data = pd$points_df, 
+            mapping = aes(y = value),
+            color = "orange", size = 1) +
+  geom_vline(xintercept = as.Date("2020-10-15")) +
+  theme_bw(base_size = 14) + 
+  theme(legend.position = "none") + 
+  ylab("QAR3 forecast for case rate in NYC") + 
+  xlab("date")
+ggsave("../gfx/ny-trajectory-plot.pdf", width = 8, height = 5)

### 2021-06-15 forecast/code/revisions-dv-jhu.R

@@ -1,0 +1,32 @@
+library(covidcast)
+library(tidyverse)
+
+# Known major case revision on Dec 29
+# HRR 303 is NYC
+
+as_ofs <- seq(as.Date("2020-12-3"), as.Date("2020-12-31"), by = "week")
+cases_as_of <- map_dfr(as_ofs, function(as_of) {
+  covidcast_signal(data_source = "jhu-csse", 
+                   signal = "confirmed_incidence_prop",
+                   start_day = "2020-10-01", end_day = "2020-12-01", 
+                   geo_type = "hrr", geo_values = "303", as_of = as_of) %>%
+    mutate(as_of = as_of)
+})
+dv_as_of <-  map_dfr(as_ofs, function(as_of) {
+  covidcast_signal(data_source = "doctor-visits", 
+                   signal = "smoothed_adj_cli",
+                   start_day = "2020-10-01", end_day = "2020-12-01", 
+                   geo_type = "hrr", geo_values = "303", as_of = as_of) %>%
+    mutate(as_of = as_of)
+})
+
+bind_rows(cases_as_of, dv_as_of) %>% 
+  mutate(as_of = fct_relabel(factor(as_of), function(x) strftime(x, "%b %d"))) %>%
+  ggplot(aes(x = time_value, y = value)) + 
+  geom_line(aes(color = factor(as_of))) + 
+  facet_wrap(~data_source, scales = "free_y") +
+  theme_bw(base_size = 14) +
+  labs(x = "", y = "", color = "") +
+  theme(legend.position = "bottom") +
+  scale_color_viridis_d(end = .9, begin = .1)
+ggsave("../gfx/revisions-dv-jhu.pdf", width = 8, height = 4)

### 2021-06-15 forecast/code/vintage.R

@@ -1,0 +1,36 @@
+library(covidcast)
+library(tidyverse)
+
+as_ofs <- lubridate::ymd("2020-11-01") + 0:6
+sigs <- list()
+for (i in seq_along(as_ofs)) {
+  sigs[[i]] <- covidcast_signals(
+    data_source = c("jhu-csse","fb-survey","doctor-visits"),
+    signal = c("confirmed_incidence_num","smoothed_wcli","smoothed_cli"),
+    geo_type = "hrr",
+    geo_values = "56", # LA HRR
+    start_day = "2020-10-01",
+    end_day = "2020-11-01",
+    as_of = as_ofs[i]) %>%
+    aggregate_signals(format = "long") %>%
+    mutate(as_of = as_ofs[i])
+}
+sigs <- bind_rows(sigs) %>% select(-c(issue:dt))
+nam <- c("jhu-csse" = "cases", 
+         "fb-survey" = "fb-survey",
+         "doctor-visits" = "doctor-visits")
+
+sigs %>% 
+  mutate(vintage = fct_relevel(as.factor(as_of), rev)) %>%
+  ggplot(aes(time_value, value, color = vintage)) +
+  geom_line() + 
+  facet_wrap(~data_source, scales = "free_y", ncol = 1, 
+             labeller = labeller(data_source = nam)) +
+  theme_bw() +
+  xlab("date") + ylab("") +
+  scale_color_viridis_d() +
+  geom_vline(xintercept = lubridate::ymd("2020-11-01")) +
+  scale_x_date(limits = lubridate::ymd(c("2020-10-26","2020-11-01")),
+               expand = expansion(c(0,.05)))
+
+ggsave("../gfx/vintage.pdf", width = 6, height = 4)

### 2021-06-16 forecast/paper/paper.tex
Edits thru discussion
- Left parts C and D of results blank, until Dan gets figures
- Discussion still needs to be fleshed out
- Figs 1 and 2 still need to be replaced by their real versions

@@
 
-\title{Can Auxiliary Indicators Improve COVID-19 Forecasts and Hotspot
+\title{Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot  
   Prediction?}
@@
 % Please add a significance statement to explain the relevance of your work
-% TODO write a real one. This is the indicator's one.
-\significancestatement{To study the COVID-19 pandemic, its effects on society,
-  and measures for reducing its spread, researchers need detailed data on the
-  course of the pandemic. Standard public health data streams suffer
-  inconsistent reporting and frequent, unexpected revisions. They also miss
-  other aspects of a population's behavior, worthy of consideration. We
-  present an open database of COVID signals in the United States, measured at
-  the county level and updated daily. These include traditionally reported COVID
-  cases and deaths, and many others: signals based on measures of mobility,
-  social distancing, internet search trends, self-reported symptoms, and
-  patterns COVID-related activity in de-identified medical insurance claims. The
-  database provides all signals in a common, easy-to-use format, empowering both
-  public health research and operational decision-making.
+\significancestatement{
+  TODO % For reference, below is the indicators paper one.
+
+  % To study the COVID-19 pandemic, its effects on society,
+  % and measures for reducing its spread, researchers need detailed data on the
+  % course of the pandemic. Standard public health data streams suffer
+  % inconsistent reporting and frequent, unexpected revisions. They also miss
+  % other aspects of a population's behavior, worthy of consideration. We
+  % present an open database of COVID signals in the United States, measured at
+  % the county level and updated daily. These include traditionally reported
+  % COVID cases and deaths, and many others: signals based on measures of
+  % mobility, social distancing, internet search trends, self-reported symptoms,
+  % and patterns COVID-related activity in de-identified medical insurance 
+  % claims. The database provides all signals in a common, easy-to-use format,
+  % empowering both  public health research and operational decision-making. 
 }
@@
   with confirmed COVID-19, based on the same claims data.
-\item COVID Tracking and Impact Survey CLI-in-community (CTIS-CLI-in-community): 
-  The estimated percentage of the population who knows someone who is sick,
-  based on Delphi's surveys of Facebook users.
+\item COVID Trends and Impact Survey CLI in the community
+  (CTIS-CLI-in-community): The estimated percentage of the population who know
+  someone in their local community that is sick, based on Delphi's surveys of
+  Facebook users, called the COVID Trends and Impact Survey. 
 \item Doctor Visits COVID-like illness (DV-CLI): The same as CHNG-CLI, but
@@
 \item Google Search Trends for anosmia and ageusia (Google-AA): A measure of 
-  COVID-19 related Google search volume for queries related to anosmia (loss of
-  smell), ageusia (loss of taste), or both.  
+  COVID-19 related Google search volume for queries relating to anosmia or
+  ageusia (loss of smell or taste), based on Google's Search Trends data set. 
 \end{itemize}
@@
 \begin{figure}[tb!]
-\centering
 \includegraphics[width=\columnwidth]{fig/vintage.pdf}
@@
 local authority to figure out how this should best be done), which can result
-in very nontrivial and often unpredictable revisions.  See Figure
-\ref{fig:vintage} for an example.
+in very nontrivial and often unpredictable revisions.  See
+Figure~\ref{fig:vintage} for an example.
 
@@
 incidence rates, per location $\ell$ and time $t$) & $Z_{\ell,t} =
-\indicator{Y_{\ell,t} \geq 1.25 \cdot Y_{\ell, t-7}}$ (indicator that $Y_{\ell,t}$ 
-grows by more than 25\% relative to the preceding week) \\
+\indicator{Y_{\ell,t} \geq 1.25 \cdot Y_{\ell,t-7}}$ (indicator that 
+$Y_{\ell,t}$ grows by more than 25\% relative to the preceding week) \\ 
   Geographic resolution & Hospital referral region (HRR) & Hospital referral
 region (HRR) \\ 
-    Forecast period & June 6--December 31, 2020 & June 13--December 31, 2020 \\  
+  Forecast period & June 9--December 31, 2020 & June 16--December 31, 2020 \\  
   Model type & Quantile regression & Logistic regression \\
@@
 ends up being (per our setup, with 21 days of training data, and lagged values
-of signals for features as we will detail shortly) June 6, 2020 for forecasting,
-and June 13, 2020 for hotspot prediction. (The bottleneck here is the
+of signals for features as we will detail shortly) June 9, 2020 for forecasting,
+and June 16, 2020 for hotspot prediction. (The bottleneck here is the
 CTIS-CLI-in-community signal, which does not exist before early April 2020, when
@@
 
-Recall $Y_{t,\ell}$ denotes the 7-day trailing average of COVID-19 case
+Recall $Y_{\ell,t}$ denotes the 7-day trailing average of COVID-19 case
 incidence rates in location $\ell$ and at time $t$.  Separately for each
@@
 predicted quantile at level $\tau$ exceeds a predicted quantile at level $\tau'
-> \tau$), we apply a simple post-hoc sorting.  See Figure \ref{fig:trajectory}
+> \tau$), we apply a simple post-hoc sorting.  See Figure~\ref{fig:trajectory}
 for an example forecast.
 
-\begin{figure}[tb!]
-\centering
+\begin{figure}[t]
 \includegraphics[width=\columnwidth]{fig/trajectory.pdf}
@@
 $$
-Z_{t,\ell} = \indicator{Y^\Delta_{\ell,t} \geq 0.25},
+Z_{\ell,t} = \indicator{Y^\Delta_{\ell,t} \geq 0.25},
 $$
 where we use the notation \smash{$Y^\Delta_{\ell,t} = (Y_{\ell,t} - Y_{\ell,
-    t-7})/(Y_{\ell,t-7})$}. In other words, $Z_{t,\ell}=1$ if the number of
+    t-7})/(Y_{\ell,t-7})$}. In other words, $Z_{\ell,t}=1$ if the number of
 newly reported cases over the past 7 days has increased by at least 25\%
@@
 \textit{hotspot} at time $t$.  Empirically, this rule labels about 27\% of
-location-time pairs as hotspots, during the prediction period (June 13--December
+location-time pairs as hotspots, during the prediction period (June 16--December
 31, 2020).
@@
 
-Some signals display a significant degree of spatial heterogeneity, by which we
-mean the values that they take at different geographic locations are not 
-comparable.  This is true of the Google-AA signal (due to the way in which the 
+Some signals have a significant amount of spatial heterogeneity, by which we
+mean their values across different geographic locations are not comparable.
+This is the case for the Google-AA signal (due to the way in which the
 underlying search trends time series is self-normalized, see
-\cite{GoogleSymptoms}) and of the claims-based signals (see the discussion in
-our companion paper on the API and database).  Such spatial heterogeneity likely
+\cite{GoogleSymptoms}) and the claims-based signals (see the discussion in our
+companion paper on the API and database).  Such spatial heterogeneity likely
 hurts the performance of the predictive models that rely on these signals,
-as we train the models on data pooled over all locations.  In the current 
+because we train the models on data pooled over all locations.  In the current
 paper, we do not attempt to address this issue (it is again a topic of ongoing
-work in our group), and we simply note that location-specific effects
-(or pre-processing to remove spatial bias) would likely improve the performance 
-of the models involving Google-AA and the claims-based indicators.
+work in our group), and we simply note that location-specific effects (or
+pre-processing to remove spatial bias) would likely improve the performance of
+the models involving Google-AA and the claims-based indicators.
 
@@
 experiments using finalized data. The results, given in the supplement, show
-that training and making predictions on finalized data can result in grossly
+that training and making predictions on finalized data can result in overly
 optimistic estimates of true test-time performance (up to 10\% better in terms
 of average WIS or AUC).
+% TODO note what we said here about the supplement
 
@@
 
+Figure~\ref{fig:forecast} displays the results for forecasting, stratified by
+ahead value and averaged over all HRRs and forecast dates (June 9--December 31,
+2020).  Shown is the average WIS for each forecast model divided by the averaged
+WIS from a ``baseline'' model, which is basically a flat-line forecaster (its
+median forecast for $Y_{\ell,t+a}$ is always $Y_{\ell,t}$, with predicted
+quantiles defined around this flat line based on historical variation). This is
+the same baseline model as that used in the COVID-19 Forecast Hub.  Here, we  
+divide by the average WIS of the baseline model in order to put the y-axis on an 
+interpretable scale.  We can see that all curves are well below 1, which means
+that all of the models, including AR, outperform the baseline on average over
+the forecast period (though their advantage over the baseline clearly narrows as
+the ahead value grows).
+
+We can also see from the figure that CHNG-COVID and DV-CLI offer the biggest
+gains over AR at small ahead values, followed by CTIS-CLI-in-community and
+Google-AA, with the former providing the biggest gains at large ahead values.
+The CHNG-CLI model performs basically the same as AR.  This is likely due to the
+fact that CHNG-CLI suffers from some sort of volatility in forecasting due to
+backfill.  The evidence for this is twofold: the CHNG-CLI model benefits from a
+more robust method of aggregating WIS (geometric mean; shown in the supplement),
+and when we train and make predictions on all-finalized data, it provides a big
+gain over AR, on par with the best-performing models (also shown in the
+supplement).
+% TODO note what we said here about the supplement
+
+\begin{figure}[t]
+\includegraphics[width=\columnwidth]{fig/fcast-1.pdf}
+\caption{TODO}
+\label{fig:forecast}
+\end{figure}
+
+Figure~\ref{fig:hotspot} displays the results for hotspot prediction, again
+stratified by ahead value and averaged over all HRRs and prediction dates (June 
+16--December 31, 2020).  We can see many similarities to the forecasting
+results: CHNG-COVID and DV-CLI offer the biggest improvement over AR, all models
+including AR degrade in performance towards the baseline (this is now just a
+classifier based on random guessing, which achieves an AUC of 0.5) as the ahead
+values grow.  However, one clear difference is that the CHNG-CLI model performs
+quite well in hotspot prediction, close to the best-performing indicator models
+for many of the ahead values. [TODO: should we insert one line explaining
+this---you're somewhat protected form volatiliy in a logistic model thanks to
+the sigmoid function used to define the probabilities?]
+
 \subsection{Implicit Regularization Hypothesis}
 
----which is that 
-  defining features from lagged indicator values in the  models in
-  \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar_x} simply provides a kind of
-  implicit regularization on top of the basic AR model (so the benefits have
-  nothing to do with the actual indicator values)---
+One might ask if the benefits we observe in forecasting and hotspot have
+anything to do with the actual indicator themselves; could it instead be that
+the indicators are simply providing ``implicit regularization'' on top of the
+basic AR model, when we include them as lagged features in
+\eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar_x}?
 
-To test the ``implicit regularization'' hypothesis (which we explain
-  precisely later), we reran all of the prediction experiments but with
-  $X_{\ell,t}$ in each indicator model replaced by suitable random noise.  This 
-  resulted in vastly different (worse) performance compared to the indicator
-  models fit with actual indicator values, and gave no improvement over the 
-  pure AR model, in both forecasting and hotspot prediction, providing a strong
+To test this implicit regularization hypothesis, we reran all of the prediction
+experiments but with $X_{\ell,t}$ in each indicator model \eqref{eq:forecast_ar_x},
+\eqref{eq:hotspot_ar_x} replaced with suitable random noise (bootstrap
+samples form an indicator's historical distribution).  The results, shown and
+explained more precisely in the supplement, are vastly different (worse) than
+the original set of results.  In both forecasting and hotspot prediction, the
+``fake'' indicator models (with random noise as auxiliary features) offered
+essentially no improvement over the pure AR model, which provides a strong
 rejection (informally speaking) of the implicit regularization hypothesis.
+
+\begin{figure}[t]
+\includegraphics[width=\columnwidth]{fig/hot-1.pdf}
+\caption{TODO}
+\label{fig:hotspot}
+\end{figure}
 
@@
 
-\subsection{Leading-ness or Lagging-ness}
+TODO
+
+\subsection{Effects of Leading or Lagging Behavior}
 
@@
 
-% TODO: make th
-
-A Primary 
-B Bootstrap
-C up/down/flat (Dan's first histogram) --> table in appendix
-D leading/lagging
-
-\subsection{Primary Results}
-
-Figure \ref{fig:topline-overall-results} displays the primary results
-of this paper, showing the effect of using indicators on the two
-forecasting tasks as a function of the horizon of the forecast (i.e.,
-the number of days ahead one is forecasting). For the case rate forecasting
-task, the left panel shows the ratio of the mean
-WIS 
-of each forecaster relative to the mean WIS of the
-baseline. \attn{Further description and justification of this metric
-  should be added to the ``Evaluation Metrics'' section rather than
-  here it seems?}  Both means are taken over all HRRs and forecast
-dates ranging from June 9, 2020 to December 31, 2020 \attn{verify}.
-All curves are well below 1, which means that all methods, including \ar, outperform the baseline across all
-considered forecast horizons. The claims-based
-indicators \dv~and \chngcov~show the most substantial gains, whereas
-the \chngcli~does not show any improvement over \ar.  The \fb~and
-\gs~indicators also show a non-negligible improvement.  Notably,
-\fb~is the only indicator to offer a clear improvement at far aheads.
-%\begin{itemize}
-%\item \dv~and \chngcov~perform very well
-%\item \fb~and \gs~perform well
-%\item \chngcli~performs poorly
-%\item Interestingly, fb is the only one offering clear improvement at
- % far aheads
-%\item Note the y-axis: consistent, clear improvement over baseline
-%\end{itemize}
-
-For the hotspot prediction task, the right panel of Figure
-\ref{fig:topline-overall-results} shows the AUC of each method.
-The monotonic decrease of all curves reflects the fact that hotspot
-prediction gets harder as the forecasting horizon increases. At the
-highest considered ahead (nearly one month ahead), the AUCs are close
-  to 0.5, which is no better than random guessing.
-In the 7-14 day ahead range, the indicators show a noticeable
-  improvement over \ar.  For larger horizons there does not
-  appear to be much difference in methods.
-The approximate ordering of methods is similar to what is seen
-  in the forecasting task: \chngcov~and \dv~are the strongest
-  performing and \ar~the weakest.
-Even for 7-day ahead predictions, the AUC is below 0.7,
-  suggesting that all methods struggle with this classification
-  task. \attn{Do people agree that an AUC of 0.7 is not all that great?}
-  That said, \dv~and \chngcov~attain roughly a 10\% increase in AUC
-  over the \ar~classifier.
-
-%\begin{itemize}
-%\item Monotonic decrease reflects the fact that hotspot
-%  prediction gets harder as ahead increases.
-%\item At the highest ahead (almost one month ahead) the AUCs are close
-%  to 0.5, which is no better than random guessing.
-%\item In the 7-14 day ahead range, the indicators show a noticeable
-%  improvement over the \ar.  For larger horizons there does not
-%  appear to be much difference in methods
-%\item The approximate ordering of methods is similar to what is seen
-%  in the forecasting task: chng\_cov and dv are the strongest
-%  performing and \ar the weakest.
-%\item Even for 7-day ahead predictions, the AUC is below 0.7,
-%  suggesting that all methods struggle with this classification task.
-%  That said, dv and chng\_cov attain roughly a 10\% increase in AUC
-%  over the \ar classifier.
-%\end{itemize}
-
-% We see that all three \attn{will need to
-% include imputed-zero google here, hopefully four} indicator-assisted
-% forecasters improve upon the baseline QAR3 forecaster.  The QAR3 model
-% assisted by the Change Healthcare confirmed COVID-19 signal
-% (QAR3+CHCOV3) sees the strongest improvement.
-% C`
-
-\begin{figure}[tb!]
-  \centering
-  \includegraphics[width=\linewidth]{fig/figure5.pdf}
-  \caption{\label{fig:topline-overall-results}  Forecasting performance as a
-    function of number of days 
-    ahead, aggregated across all HRRs and forecast dates. \attn{How's this?}}
-\end{figure}
-
-\subsection{Avoiding misleading retrospective evaluations}
-
-Figure \ref{fig:finalized-vs-honest} quantifies the effect of
-not properly accounting for the question of ``what was known when'' in
-performing retrospective evaluations of forecasters.  When methods are
-given the finalized version of the data rather than the version available at the time that the forecast would have been made, all methods
-appear (falsely) to have better performance.  For example, for forecasting case rates
-7-days ahead, the WIS of all methods is at least 8\% larger than what would have been
-recorded using the finalized values of the data.  This effect
-diminishes as the forecasting horizon increases, reflecting the fact
-that these forecasters rely less heavily on recent data than very
-short-term forecasters.  Crucially, some methods are
-``helped'' more than others by the less scrupulous retrospective
-evaluation, underscoring the difficulty of avoiding misleading
-conclusions when performing retrospective evaluations of forecasters.
-
-The \chngcli~indicator (along with the other claims-based signals) is the most
-affected by this distinction, reflecting the latency in claims-based
-reporting. \attn{Actually, \dv~is the least affected for forecasting
-  whereas it is one of the most affected for hotspots.}  This supports
-the importance of efforts to provide ``nowcasts'' for claims signals
-(which corresponds to a 0-ahead ``forecast'' of what the claims
-signal's value will be once all data has been collected).
-
-Even the \ar~model is affected by this distinction, reflecting the
-fact that the case rates themselves (i.e., the response values) are also
-subject to revision.  The forecasters based on indicators are thus
-affected both by revisions to the indicators and by revisions to the
-case rates.
-
-\begin{itemize}
-%\item Motivates the importance of nowcasting for claims signals.
-  \item For forecasting task: Also, \chngcli~and \dv~perform very similarly here, which is
-    reassuring because they're measuring the same thing (in
-    principle).  Yet they must have very different backfill profiles…
-  \item For hotspots task: dv seems to get a bigger boost than
-    \chngcov - interesting dichotomy to forecasting
-\item But in both forecasting/hotspots \chngcli~is affected a lot by backfill
-\item \attn{We should be careful about which indicators we were not
-      able to track latency for.  In fact, we should probably remove
-      these from the plot since they would be more strongly affected
-      than is shown in the plot.}
-
-\item \attn{We can also discuss the distinction
-      between indicators that could in principle have low latency
-      (e.g. GS) versus those where latency is inherent
-      (e.g. claims-based indicators). If the goal is assessing the
-      performance of GS ``for the next pandemic'' then it would make
-      sense to give GS a pass since in theory they could have had it
-      with very low latency (as opposed to claims based signals where
-      backfill may be unavoidable).  Of course, in
-    ``the next pandemic'' there might not be such highly specific
-    symptoms such as A+A as there was in the COVID-19 case.} 
-\end{itemize}
-
-\begin{figure}[tb!]
-     \centering
-     \includegraphics[width=\linewidth]{fig/figure6.pdf}
-     \caption{How misleading is a retrospective analysis that uses the
-       finalized version of the data rather than the version that was
-       in fact available at the time the forecast was made?  Plots
-       show the ratio in performance (using mean WIS fore the
-       forecasting task and AUC for the hotspot prediction
-       task). \label{fig:finalized-vs-honest} Cheating vs. honest plots.} 
-\end{figure}
-
-\subsection{Additional sensitivity analyses}
-
-In addition to whether one properly accounts for signal latency, there
-are several other choices in evaluating forecasters that can also have
-a non-trivial impact on results.
-
-For example, in Figure
-\ref{fig:topline-overall-results}, we show the ratio
-of means.  Other choices of aggregation and scaling are also
-possible.  For example, taking the mean of ratios of errors is
-possible and would give more importance to locations where
-errors are small (either because the location-time pairs had low case
-rates or because the forecasting task was itself easier).  By
-contrast, the ratio of the mean of errors may be dominated by
-location-time pairs where the error is large.
-Figure \ref{fig:
-  scale-aggregate-order} of the Supplementary Information shows...
-\begin{itemize}
-\item Note the y-axis: barely an improvement over baseline, and this goes away for AR after 11 days ahead, and for all other models around 13-14 days ahead.
-\item Using trimmed means here because otherwise the results are extremely unstable (as the denominators can be small, when WIS of baseline is small).
-\item Similar message to before (non-adjusted), but \dv~and \chngcov~the best, \fb~and \gs~still good, but now \chngcli~seems to do much better!  
-\item What's our interpretation here?  That somehow \chngcli~just
-  does particularly bad in tasks where the baseline is also bad?
-  \item We might include a histogram of WIS with a log-transformed
-    x-axis to suggest that the geometric mean may be reasonable here.
-\end{itemize}
-Using instead the geometric mean makes the order of aggregation and
-scaling immaterial since the ratio of geometric means is the same as
-the geometric mean of ratios.  Figure \ref{fig:geometric-mean} of the
-Supplementary Information shows how Figure
-\ref{fig:topline-overall-results} panel A would change if
-one replaces the arithmetic mean by the geometric mean.
-
-In Supplementary Information \ref{sec:2021}, we investigate the effect of lengthening
-the time window to include 2021.
-\begin{itemize}
-\item For forecasting: results get better, whether we look at
-  non-adjusted (first plot) or adjusted scores (second plot).  Other
-  than that, interpretation is qualitatively similar, except fb now
-  seems to be worse than gs.  And \chngcli~is strong, even in the
-  non-adjusted view.
-  \item For hotspots: Alden will write something explaining why we
-    don't do hotspot detection in 2021.
-\end{itemize}
-
-In Supplementary Information Section \ref{sec:gs-locations}, we look at various approaches to
-accounting for the fact that the \gs~indicator is only available
-for XX HRRs.
-
-\begin{itemize}
-  \item For forecasting task: clearly gs is not missing at random, and when it's present, it tends to be predictive.
-Hence high values have a low false positivity rate.
-Most visible in the adjusted view below, where gs triumphs.
-Also \chngcli~gets a lot worse.
-
-\item For hotspots: All methods look better, suggesting that the “GS locations” are easier for hotspot prediction.
-E.g.
-at 7-days ahead, the AUCs computed based on all locations range from 0.61-0.66; restricted to GS locations, the AUCs range from about 0.65-0.69.
-(These are all eyeballed, should get actual numbers if we want to say something like this in paper.)
-\item gs\_subset appears to be particularly helped by this subsetting
-  (which makes sense since on those other locations it was 0-imputed)
-  \item Daniel will also try gs\_impute as in forecasting
-\end{itemize}
-
-% Figure \ref{fig:forecast-median-relative-wis-faceted} splits the topline
-% median relative WIS results over four time periods:  June and July, which
-% corresponds to the small, initial surge experienced in the US;
-% August and September, when incident cases nationally were somewhat stable;
-% October, when a second surge was witnessed; and November and December,
-% which saw the largest surges to date.  In essentially all the time periods,
-% the indicator-assisted models perform in aggregate as well or better than
-% the baseline QAR3 model.  The results in October do not show much
-% differentiation between the baseline QAR3 and the indicator-assisted models.
-% However, we note that during that time period, even the baseline QAR3
-% model is doing very well in terms of relative WIS (especially compared to
-% other time periods), so perhaps there is not much ``room for improvement'',
-% so to speak.
-
-% \begin{figure*}[tb!]
-%      \centering
-%      \begin{subfigure}[b]{0.49\textwidth}
-%          \centering
-%          \includegraphics[width=\textwidth]{fig/reduction_mean_scaled_wis_facet_honest.pdf}
-%          \caption{Forecasting task: mean WIS relative to mean WIS of baseline}
-%          \label{fig:forecast-reduction-mean-scaled-wis-facet}
-%      \end{subfigure}
-%      \hfill
-%      \begin{subfigure}[b]{0.49\textwidth}
-%          \centering
-%          \includegraphics[width=\textwidth]{fig/auc_by_forecast_period.pdf}
-%          \caption{Hotspot prediction task: AUC}
-%          \label{fig:hotspot-auc-faceted}
-%      \end{subfigure}
-%      \label{fig:topline-faceted-results}
-%      \caption{Topline results from June - December 2020, split into
-%      four representative time periods.}
-% \end{figure*}
-
-\paragraph{\attn{Additional results section material, some for Discussion}}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=.95\linewidth]{fig/fake-map.pdf}
-    \caption{Weighted interval score for forecasts made on April 1, 2021. Scores
-      are averaged over horizons from 1 to 14 days ahead and are relative to
-      population size.} 
-    \label{fig:fake-map}
-\end{figure}
-
-\begin{enumerate}
-%\item How does this vary over time? Table showing 7-day ahead and 14-day ahead
-%  (just picking out two of the d=5,...,20). Include a national curve.
-\item Spatial map of errors across HRRs for both tasks. See
-  \autoref{fig:fake-map} for an example. 
-\item Unlikely we can find it, but we should try anyway. Possibly put in the
-  intro or discussion instead. Find an example of a forecast altered by the
-  presence of an indicator.  Plot the two trajectories.
-\item Examine a specific HRR over time. What does it mean to predict certain
-  hotspots? How can we understand what is happening?
- \item What can we say about statistical significance? NRI, paired t-test?
- \item Figures showing hotspots make sense
- \item Horizontal line at 0.5 on AUC curve, discribe that this is random
-   guessing. 
-\end{enumerate}
-
 \section{Discussion}
 
-We have assessed the performance of [NUMBER] indicators in terms of their
-usefulness for performing two forecasting tasks: probabilistic
-forecasting of case
-rates and hotspot prediction at the HRR-level, 1--[NUMBER] days in advance.
-In measuring the practical usefulness of an indicator to forecasting,
-we distinguish between honest versus aspirational assessments.  Data
-latency is an important factor when considering the usefulness of an
-indicator.  Retrospective evaluations that ignore these messy data
-issues fail to provide the full picture.  While one can hope that reporting
-systems will operate faster and more consistently in ``the next
-pandemic,'' some reasons for this latency appear unavoidable.  The
-chaos and strain inherent to pandemic times make indicators that can
-be reported and collected with quickly with minimal effort or manual
-involvement desirable.
+Can axuiliary indicators improve COVID-19 forecasting and hotspot prediction
+models?  Our answer, based on analyzing five auxiliary indicators from the  
+COVIDcast API (defined using from medical insurance claims, internet-based
+surveys, and Google search trends) is undoubtedbly ``yes''. However, there are 
+levels of nuance to such an answer that must be explained.  None of the
+indicators that we have investigated appear to be the ``silver bullet'' that 
+one  might have hoped for, revolutionizing the tractability of the prediction 
+problem, rendering it ``easy'' when it was once ``hard'', in the absence of
+auxiliary information.  Rather, the gains in accuracy offered by the indicator
+models appear to be nontrivial, and consistent across modes of analysis, but
+modest---and more about tipping the bias-variance tradeoff in our favor than
+providing a consistent leading indication of future events.
 
-We conclude by observing that the
-approach we have taken in this paper is somewhat opposite to that of much of the COVID-19
-forecasting literature.  We have chosen to consider only very simple forecasting models while devoting most of our effort to accounting for as much of the complexity of
-the data and evaluation as possible.  By contrast, many papers focus
-on very complicated forecasting approaches but then evaluate them
-under unrealistic, retrospective conditions.
+[TODO:  Explain how insights from up/down/flat, lagging/leading, etc., are
+relevant to this overall arc.  Reiterate the importance of vintage data.  Finish
+with something about what's left for future work?  ]
 
-\begin{enumerate}
+% RJT: some text commented out below.  May be useful for finishing off the
+% discussion
 
-\item Discussion should contain more detailed lit review ("context"), rather
-  than in the intro, per the instructions.
+% This paper is as much about demonstrating how one might go about answering
+% such questions about feature importance in forecasting and hotspot prediction 
+% general, as providing an empirically rigorous answer to this precise question
+% in the context of the COVID-19 pandemic.
 
-\item Use of syndromic surveillance for other diseases?
-
-\item Cite any papers that use our signals in forecasting.
-\end{enumerate}
-
-\paragraph{\attn{Other potential discussion topics}}
-
-\begin{itemize}
-
-\item When was this data actually available? (As of issue. The argument is: sure
-  we didn't have it, but it would have helped. We should have this next time.)
-\item Side issues, are there places / times where one signal is especially good?
-  Especially bad?
-\item What do we do about holidays? How do we understand the performance when
-  the data is trash? If the data is trash, all the more reason to have other
-  data that is perhaps less susceptible to these criticisms.
-\item Other signals we might want? Mobility? Discuss how SEIR models need
-  contact matrices, relationship with ``in community'' signals.
-
-\end{itemize}
+% We conclude by observing that the approach we have taken in this paper is
+% somewhat opposite to that of much of the COVID-19 forecasting literature.  We
+% have chosen to consider only very simple forecasting models while devoting
+% most of our effort to accounting for as much of the complexity of the data and 
+% evaluation as possible.  By contrast, many papers focus on very complicated
+% forecasting approaches but then evaluate them under unrealistic, retrospective 
+% conditions.
 

### 2021-06-16 forecast/paper/unused.tex
Edits thru discussion
- Left parts C and D of results blank, until Dan gets figures
- Discussion still needs to be fleshed out
- Figs 1 and 2 still need to be replaced by their real versions

%%%%%

Using instead the geometric mean makes the order of aggregation and
scaling immaterial since the ratio of geometric means is the same as
the geometric mean of ratios.  Figure \ref{fig:geometric-mean} shows how the
results in forecasting would change if one replaces the arithmetic mean by the
geometric mean.

%%%%%

We also investigate the effect of lengthening the time window to include 2021. 
\begin{itemize}
\item For forecasting: results get better, whether we look at
  non-adjusted (first plot) or adjusted scores (second plot).  Other
  than that, interpretation is qualitatively similar, except fb now
  seems to be worse than gs.  And \chngcli~is strong, even in the
  non-adjusted view.
  \item For hotspots: Alden will write something explaining why we
    don't do hotspot detection in 2021.
\end{itemize}

%%%%%

We look at various approaches to
accounting for the fact that the \gs~indicator is only available
for XX HRRs.

\begin{itemize}
  \item For forecasting task: clearly gs is not missing at random, and when it's present, it tends to be predictive.  Hence high values have a low false positivity rate.  Most visible in the adjusted view below, where gs triumphs.  Also \chngcli~gets a lot worse.

\item For hotspots: All methods look better, suggesting that the “GS locations” are easier for hotspot prediction. E.g. at 7-days ahead, the AUCs computed based on all locations range from 0.61-0.66; restricted to GS locations, the AUCs range from about 0.65-0.69. (These are all eyeballed, should get actual numbers if we want to say something like this in paper.)
\item gs\_subset appears to be particularly helped by this subsetting
  (which makes sense since on those other locations it was 0-imputed)
  \item Daniel will also try gs\_impute as in forecasting
\end{itemize}

%%%%%

Figure \ref{fig:finalized-vs-honest} quantifies the effect of
not properly accounting for the question of ``what was known when'' in
performing retrospective evaluations of forecasters.  When methods are
given the finalized version of the data rather than the version available at the time that the forecast would have been made, all methods
appear (falsely) to have better performance.  For example, for forecasting case rates
7-days ahead, the WIS of all methods is at least 8\% larger than what would have been
recorded using the finalized values of the data.  This effect
diminishes as the forecasting horizon increases, reflecting the fact
that these forecasters rely less heavily on recent data than very
short-term forecasters.  Crucially, some methods are
``helped'' more than others by the less scrupulous retrospective
evaluation, underscoring the difficulty of avoiding misleading
conclusions when performing retrospective evaluations of forecasters.

The \chngcli~indicator (along with the other claims-based signals) is the most
affected by this distinction, reflecting the latency in claims-based
reporting. \attn{Actually, \dv~is the least affected for forecasting
  whereas it is one of the most affected for hotspots.}  This supports
the importance of efforts to provide ``nowcasts'' for claims signals
(which corresponds to a 0-ahead ``forecast'' of what the claims
signal's value will be once all data has been collected).

Even the \ar~model is affected by this distinction, reflecting the
fact that the case rates themselves (i.e., the response values) are also
subject to revision.  The forecasters based on indicators are thus
affected both by revisions to the indicators and by revisions to the
case rates.

\begin{itemize}
%\item Motivates the importance of nowcasting for claims signals.
  \item For forecasting task: Also, \chngcli~and \dv~perform very similarly here, which is
    reassuring because they're measuring the same thing (in
    principle).  Yet they must have very different backfill profiles…
  \item For hotspots task: dv seems to get a bigger boost than
    \chngcov - interesting dichotomy to forecasting
\item But in both forecasting/hotspots \chngcli~is affected a lot by backfill
\item \attn{We should be careful about which indicators we were not
      able to track latency for.  In fact, we should probably remove
      these from the plot since they would be more strongly affected
      than is shown in the plot.}

\item \attn{We can also discuss the distinction
      between indicators that could in principle have low latency
      (e.g. GS) versus those where latency is inherent
      (e.g. claims-based indicators). If the goal is assessing the
      performance of GS ``for the next pandemic'' then it would make
      sense to give GS a pass since in theory they could have had it
      with very low latency (as opposed to claims based signals where
      backfill may be unavoidable).  Of course, in
    ``the next pandemic'' there might not be such highly specific
    symptoms such as A+A as there was in the COVID-19 case.} 
\end{itemize}

\begin{figure}[t]
     \centering
     \includegraphics[width=\textwidth]{fig/fcast-honest-v-finalized-1.pdf}
     \includegraphics[width=\textwidth]{fig/hot-honest-v-finalized-1.pdf}
     \caption{How misleading is a retrospective analysis that uses the
       finalized version of the data rather than the version that was
       in fact available at the time the forecast was made?  Plots
       show the ratio in performance (using mean WIS fore the
       forecasting task and AUC for the hotspot prediction
       task). \label{fig:finalized-vs-honest} Cheating vs. honest plots.} 
\end{figure}

### 2021-06-16 forecast/paper/supplement.Rmd
add supplement and rebuilt figs

@@
 ---
-bibliography: [../../common/covidcast.bib]
+bibliography: [../../common/covidcast.bib, pnas-materials/pnas-sample.bib]
 output: 
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, fig.width = 5, fig.height = 3,
+knitr::opts_chunk$set(echo = FALSE, fig.width = 6.5, fig.height = 4.5,
                       fig.align = "center",
@@
                       cache = TRUE, 
+                      out.width = "\\linewidth",
                       autodep = TRUE,
@@
                       warning = FALSE)
-source("eval_funs.R")
-source("hotspot_funs.R")
-source("trajectory_plots.R")
+```
+
+```{r source-funs, cache=FALSE}
+source("../code/eval_funs.R")
+source("../code/hotspot_funs.R")
+source("../code/trajectory_plot_funs.R")
 library(scales)
+library(cowplot)
 path_to_data <- "~/Documents/COVID-Delphi/fcast-indicators-paper/temp_data/"
@@
 
+# Figures from results
+
 ## Forecasts
@@
   scale_x_log10() +
-  theme_bw(base_size = 12) +
+  theme_bw() +
   xlab("WIS") +
-  scale_fill_brewer(palette = "Set1") +
+  scale_fill_manual(values = fcast_colors) +
   facet_wrap(~forecaster) +
@@
 
-## Trajectories
+## Trajectories (deprecated)
 
@@
   ggplot(aes(ahead, auc_honest / auc_finalized, color = forecaster)) + 
-  geom_line() + geom_point() +
-  theme_bw(base_size = 12) + 
-  scale_color_brewer(palette = "Set1", guide = guide_legend(nrow=1)) +
+  geom_line() + 
+  geom_point() +
+  theme_bw() + 
+  scale_color_manual(values = fcast_colors, guide = guide_legend(nrow=1)) +
   theme(legend.position = "bottom", legend.title = element_blank()) +
@@
 
-```{r figure5, eval = FALSE, fig.width=9, fig.height=4}
-library(cowplot)
+```{r figure5}
 leg <- get_legend(fig5a)
@@
 
-```{r figure6, fig.width=9, fig.height=4}
+```{r figure6}
 fig6a <- fig6a + theme(legend.position = "none")
@@
 
-### Correlations with lagged actuals
+## Correlations with lagged actuals
 
@@
 comparison_df <- left_join(
-  fcasts_honest %>% filter(forecaster != "QAR3", period != "jm"),
-  fcasts_honest %>% filter(forecaster == "QAR3", period != "jm") %>%
+  fcasts_honest %>% filter(forecaster != "AR", period != "jm"),
+  fcasts_honest %>% filter(forecaster == "AR", period != "jm") %>%
     select(-c(forecaster,target_end_date)) %>%
@@
   geom_histogram(aes(y = ..count.. / sum(..count..), fill = forecaster)) +
-  scale_fill_brewer(palette = "Set1") +
-  theme_bw(base_size = 14) +
-  theme(legend.position = "bottom", legend.title = element_blank()) +
+  scale_fill_manual(values = fcast_colors) +
+  theme_bw() +
+  theme(legend.position = "none") +
   geom_vline(aes(xintercept = median)) + 
@@
   facet_wrap(~forecaster) + 
-  xlab("Spearman correlation") + ylab("relative frequency") +
+  xlab("Spearman correlation") + ylab("Relative frequency") +
   scale_y_continuous(labels = scales::label_percent(accuracy = 1))
@@
                              AR3DVCLI3 = "DV-CLI",
-                             AR3FBCLI3 = "CTIS-CLI-in-community",
+                             AR3FBCLI3 = "CTIS-CLIIC",
                              AR3GSSAA3_Subset = "gs_subset",
@@
   facet_grid(udf ~ forecaster) +
-  theme_bw(base_size = 14) +
+  theme_bw() +
+  ylab("Count") +
   theme(legend.position = "none") +
-  scale_fill_brewer(palette = "Set1") +
+  scale_fill_manual(values = fcast_colors) +
   scale_y_log10(breaks = c(10,1000,100000), 
@@
   facet_grid(udf ~ forecaster) +
-  theme_bw(base_size = 14) +
+  theme_bw() +
+  ylab("Count") +
   theme(legend.position = "none") +
-  scale_fill_brewer(palette = "Set1") +
+  scale_fill_manual(values = fcast_colors) +
   scale_x_log10() +
@@
       `value+0:doctor-visits_smoothed_adj_cli` = "DV-CLI",
-      `value+0:fb-survey_smoothed_hh_cmnty_cli` = "CTIS-CLI-in-community",
+      `value+0:fb-survey_smoothed_hh_cmnty_cli` = "CTIS-CLIIC",
       `value+0:google-symptoms_sum_anosmia_ageusia_smoothed_search` = "Google-AA")) %>%
@@
 
-```{r leading-and-lagging, fig.width=8, fig.height=4}
+```{r leading-and-lagging}
 df2 %>%
@@
   geom_hline(yintercept = 0) +
-  scale_fill_brewer(palette = "Set1") +
+  scale_fill_manual(values = fcast_colors) +
   facet_wrap(~name) +
-  theme_bw(base_size = 14) +
-  ylab("correlation") +
+  theme_bw() +
+  ylab("Correlation") +
   xlab("") +
@@
 
-```{r diff-in-lead-lag, fig.width=8, fig.height=4}
+```{r diff-in-lead-lag}
 df2 %>%
@@
   geom_hline(yintercept = 0) +
-  scale_fill_brewer(palette = "Set1") +
-  theme_bw(base_size = 14) +
-  ylab("correlation") +
+  scale_fill_manual(values = fcast_colors) +
+  theme_bw() +
+  ylab("Correlation") +
   xlab("") +
@@
 ```
+
+# Side figures (in their own scripts)
+
+```{r ccf-dv-finalized}
+library(covidcast)
+source("../code/ccf-dv-finalized.R")
+gg
+```
+
+```{r ny-trajectory}
+source("../code/ny-trajectory.R")
+gg
+```
+
+```{r revisions-dv-jhu}
+source("../code/revisions-dv-jhu.R")
+pp
+```

### 2021-06-16 forecast/code/ccf-dv-finalized.R

@@ -1,6 +1,3 @@
-library(covidcast)
-library(tidyverse)
-
 sigs <- covidcast_signals(
   data_source = c("jhu-csse", "doctor-visits"),
   signal = c("confirmed_7dav_incidence_prop", "smoothed_adj_cli"),
@@ -38,5 +35,7 @@ g2 <- ggplot(tibble(lag = drop(cc$lag), ccf = drop(cc$acf)), aes(lag)) +
   geom_vline(xintercept = 0) +
   ylab("cross correlation") +
   theme_bw(base_size = 14) 
-cowplot::plot_grid(g2, g1, rel_widths = c(4,6))
-ggsave("../gfx/ccf-dv-finalized.pdf", width = 8, height = 4)
+
+
+gg <- cowplot::plot_grid(g2, g1, rel_widths = c(4,6))
+

### 2021-06-16 forecast/code/eval_funs.R

@@ -10,6 +10,10 @@ TrimMean05 <- function(x) mean(x, trim = .05, na.rm = TRUE)
 TrimMean1 <- function(x) mean(x, trim = .1, na.rm = TRUE)
 GeoMean <- function(x) exp(mean(log(x), na.rm = TRUE))
 
+fcast_colors <- c(RColorBrewer::brewer.pal(5, "Set1"), "#000000")
+names(fcast_colors) <- c("CHNG-CLI", "CHNG-COV", "CTIS-CLIIC", "DV-CLI",
+                         "Google-AA", "AR")
+
 summarizer <- function(df, y, centerer, scaler, aggr, order_of_operations) {
   for (i in seq_along(order_of_operations)) {
     df <- switch(
@@ -46,8 +50,9 @@ plotter <- function(df, y, aggr,
   ggplot(df, aes(ahead, wis, color = forecaster)) + 
     geom_line() + geom_point() +
     facet_layer +
-    theme_bw(base_size = 14) + 
-    scale_color_brewer(palette = "Set1", guide = guide_legend(nrow = 1)) +
+    theme_bw() + 
+    xlab("Days ahead") +
+    scale_color_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
     theme(legend.position = "bottom", legend.title = element_blank())
 }
 
@@ -65,10 +70,13 @@ plotter_hotspots <- function(df, facet_by_period = FALSE) {
   df <- summarize(df, auc = auc(value, actual))
   
   ggplot(df, aes(ahead, auc, color = forecaster)) + 
-    geom_line() + geom_point() +
+    geom_line() + 
+    geom_point() +
     facet_layer +
-    theme_bw(base_size = 14) + 
-    scale_color_brewer(palette = "Set1", guide = guide_legend(nrow=1)) +
+    theme_bw() + 
+    ylab("AUC") +
+    xlab("Days ahead") +
+    scale_color_manual(values = fcast_colors, guide = guide_legend(nrow=1)) +
     theme(legend.position = "bottom", legend.title = element_blank())
 }
 
@@ -90,9 +98,9 @@ process_res_hotspots <- function(res) {
     mutate(forecaster = recode(forecaster,
                                AR3 = "AR",
                                AR3CHCLI3 = "CHNG-CLI",
-                               AR3CHCOV3 = "CHNG-COVID",
+                               AR3CHCOV3 = "CHNG-COV",
[…]

### 2021-06-16 forecast/code/hotspot-contingency-table.R

@@ -1,35 +1,0 @@
-la <- hotspots_honest %>%
-  filter(geo_value == "56", period != "jm") %>%
-  left_join(actuals %>% filter(geo_value == "56") %>% rename(observed = actual))
-
-la %>%
-  ggplot(aes(target_end_date, value, color = ahead)) +
-  geom_vline(data = la %>% filter(actual == 1), aes(xintercept = target_end_date),
-             color = "orange", size = .2) +
-  geom_point() +
-  scale_fill_viridis_c() +
-  facet_wrap(~forecaster, ncol = 1) +
-  theme_bw()
-
-hot_udf <- inner_join(hotspots_honest %>% filter(period != "jm"),
-                      up_down %>% select(geo_value, target_end_date, udf))
-n_expected <- hot_udf %>% group_by(udf,forecaster) %>% summarise(n_forecasts = n())
-
-
-tab <- hot_udf %>% 
-  mutate(deciles = ntile(value, 10)) %>%
-  group_by(udf, deciles, forecaster) %>%
-  summarise(m = n()) %>%
-  left_join(n_expected) %>%
-  mutate(expected = n_forecasts * .1, deviation = m / expected)
-
-ggplot(tab, aes(deciles, deviation, color = forecaster)) +
-  geom_line() +
-  geom_point() +
-  geom_hline(yintercept = 1) +
-  facet_wrap(~udf, ncol = 1) +
-  scale_color_brewer(palette = "Set1", guide = guide_legend(nrow = 1)) +
-  scale_x_continuous(breaks = 1:10, labels = 1:10) +
-  theme_bw() +
-  ylab("predictions / expected") +
-  theme(legend.position = "bottom", legend.title = element_blank())

### 2021-06-16 forecast/code/hrr_population.R

@@ -1,28 +1,0 @@
-library(tidyverse)
-fips_pop <- read_csv("https://raw.githubusercontent.com/cmu-delphi/covidcast-indicators/main/_delphi_utils_python/delphi_utils/data/fips_pop.csv",
-                     col_types = "ci")
-hrrtab <- read_csv("https://raw.githubusercontent.com/cmu-delphi/covidcast-indicators/main/_delphi_utils_python/delphi_utils/data/fips_hrr_table.csv",
-                   col_types = "ccd")
-hrrs <- left_join(hrrtab, fips_pop, by = "fips")
-hrrs <- mutate(hrrs, wpop = pop * weight)
-hrr_pop <- hrrs %>%
-  group_by(hrr) %>%
-  summarise(pop = sum(wpop, na.rm=TRUE))
-# note, 02270, 46113, 51515 have pop NA. These seem no longer to exist (renumbered or removed)
-
-hrr_names <- read_csv("Hospital_Referral_Regions.csv", col_types = "_ccc__") %>%
-  mutate(state = substr(hrrcity, 1, 2))
-hrr_pop <- left_join(hrr_pop, hrr_names, by = c("hrr" = "hrrnum")) %>%
-  mutate(name = paste(HRR_lbl,State)) %>%
-  select(hrr, name, pop)
-
-ggplot(hrr_pop, aes(pop)) +
-  geom_histogram(aes(y=(..count..)/sum(..count..)), bins = 20, fill = "cornflowerblue") +
-  xlab("population") + ylab("frequency") +
-  scale_x_log10(labels = scales::label_comma()) +
-  scale_y_continuous(expand = expansion(c(0,.05)),
-                     labels = scales::label_percent(accuracy = 1),
-                     breaks = scales::breaks_pretty(4)) +
-  theme_bw(base_family = "Times")
-
-ggsave("../gfx/hrr_pop_histogram.pdf", width=6, height=3)

### 2021-06-16 forecast/code/ny-trajectory.R

@@ -1,14 +1,12 @@
-library(tidyverse)
-library(lubridate)
 
-ny_preds <- readRDS("../temp_data/predictions_honest.RDS") %>%
+ny_preds <- readRDS(file.path(path_to_data, "predictions_honest.RDS")) %>%
   filter(geo_value == "303", 
          forecast_date == "2020-10-15",
          forecaster == "AR3") %>%
   mutate(target_end_date = forecast_date + days(ahead),
          incidence_period = "day")
 
-actuals <- readRDS("../temp_data/actuals.RDS") %>%
+actuals <- readRDS(file.path(path_to_data, "actuals.RDS")) %>%
   filter(geo_value == "303", 
          target_end_date > "2020-09-01", 
          target_end_date < "2020-11-15") %>%
@@ -17,7 +15,7 @@ actuals <- readRDS("../temp_data/actuals.RDS") %>%
 pd <- evalcast:::setup_plot_trajectory(ny_preds, side_truth = actuals)
 pd$truth_df <- pd$truth_df %>% rename(target_end_date = time_value)
 
-ggplot(pd$truth_df, mapping = aes(x = target_end_date)) +
+gg <- ggplot(pd$truth_df, mapping = aes(x = target_end_date)) +
   geom_ribbon(
     data = pd$quantiles_df,
     mapping = aes(ymin = lower, ymax = upper, fill = interval)) +
@@ -33,6 +31,5 @@ ggplot(pd$truth_df, mapping = aes(x = target_end_date)) +
   geom_vline(xintercept = as.Date("2020-10-15")) +
   theme_bw(base_size = 14) + 
   theme(legend.position = "none") + 
-  ylab("QAR3 forecast for case rate in NYC") + 
+  ylab("AR forecast for case rate in NYC") + 
   xlab("date")
-ggsave("../gfx/ny-trajectory-plot.pdf", width = 8, height = 5)

### 2021-06-16 forecast/code/revisions-dv-jhu.R

@@ -1,6 +1,3 @@
-library(covidcast)
-library(tidyverse)
-
 # Known major case revision on Dec 29
 # HRR 303 is NYC
 
@@ -20,13 +17,15 @@ dv_as_of <-  map_dfr(as_ofs, function(as_of) {
     mutate(as_of = as_of)
 })
 
-bind_rows(cases_as_of, dv_as_of) %>% 
+
+pp <- bind_rows(cases_as_of, dv_as_of) %>% 
   mutate(as_of = fct_relabel(factor(as_of), function(x) strftime(x, "%b %d"))) %>%
   ggplot(aes(x = time_value, y = value)) + 
   geom_line(aes(color = factor(as_of))) + 
   facet_wrap(~data_source, scales = "free_y") +
-  theme_bw(base_size = 14) +
+  theme_bw() +
   labs(x = "", y = "", color = "") +
   theme(legend.position = "bottom") +
   scale_color_viridis_d(end = .9, begin = .1)
-ggsave("../gfx/revisions-dv-jhu.pdf", width = 8, height = 4)
+
+

### 2021-06-16 forecast/code/vintage.R

@@ -1,36 +1,0 @@
-library(covidcast)
-library(tidyverse)
-
-as_ofs <- lubridate::ymd("2020-11-01") + 0:6
-sigs <- list()
-for (i in seq_along(as_ofs)) {
-  sigs[[i]] <- covidcast_signals(
-    data_source = c("jhu-csse","fb-survey","doctor-visits"),
-    signal = c("confirmed_incidence_num","smoothed_wcli","smoothed_cli"),
-    geo_type = "hrr",
-    geo_values = "56", # LA HRR
-    start_day = "2020-10-01",
-    end_day = "2020-11-01",
-    as_of = as_ofs[i]) %>%
-    aggregate_signals(format = "long") %>%
-    mutate(as_of = as_ofs[i])
-}
-sigs <- bind_rows(sigs) %>% select(-c(issue:dt))
-nam <- c("jhu-csse" = "cases", 
-         "fb-survey" = "fb-survey",
-         "doctor-visits" = "doctor-visits")
-
-sigs %>% 
-  mutate(vintage = fct_relevel(as.factor(as_of), rev)) %>%
-  ggplot(aes(time_value, value, color = vintage)) +
-  geom_line() + 
-  facet_wrap(~data_source, scales = "free_y", ncol = 1, 
-             labeller = labeller(data_source = nam)) +
-  theme_bw() +
-  xlab("date") + ylab("") +
-  scale_color_viridis_d() +
-  geom_vline(xintercept = lubridate::ymd("2020-11-01")) +
-  scale_x_date(limits = lubridate::ymd(c("2020-10-26","2020-11-01")),
-               expand = expansion(c(0,.05)))
-
-ggsave("../gfx/vintage.pdf", width = 6, height = 4)

### 2021-06-16 forecast/paper/supplement.tex
Edits thru discussion
- Left parts C and D of results blank, until Dan gets figures
- Discussion still needs to be fleshed out
- Figs 1 and 2 still need to be replaced by their real versions

@@
 
-\input{forecast-prelim.tex}
+\input{../../common/prelim.tex}
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 \SItext
+
+\section*{PNAS Template material, to be removed}
 
@@
 \centering
-\includegraphics[width=\textwidth]{frog}
+\includegraphics[width=\textwidth]{pnas-materials/frog}
 \caption{Second figure}
@@
 
+\clearpage
+
+\hypertarget{figures-from-results}{%
+\section{Figures from results}\label{figures-from-results}}
+
+\hypertarget{forecasts}{%
+\subsection{Forecasts}\label{forecasts}}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/fcast-1} \end{center}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/fcast-finalized-1} \end{center}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/fcast-honest-v-finalized-1} \end{center}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/fcast-adjusted-1} \end{center}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/wis-densities-1} \end{center}
+
+\hypertarget{all-periods}{%
+\subsubsection{All periods}\label{all-periods}}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/fcast-alldates-1} \end{center}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/fcast-alldates-adjusted-1} \end{center}
+
+\hypertarget{only-gs-locations}{%
+\subsubsection{Only GS locations}\label{only-gs-locations}}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/fcast-gs-locations-1} \end{center}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/fcast-gs-locations-adjusted-1} \end{center}
+
+\hypertarget{trajectories-deprecated}{%
+\subsection{Trajectories (deprecated)}\label{trajectories-deprecated}}
+
+\hypertarget{hotspots}{%
+\subsection{Hotspots}\label{hotspots}}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/hot-1} \end{center}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/hot-finalized-1} \end{center}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/hot-honest-v-finalized-1} \end{center}
+
+\hypertarget{all-periods-1}{%
+\subsubsection{All periods}\label{all-periods-1}}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/hot-alldates-1} \end{center}
+
+\hypertarget{only-gs-locations-1}{%
+\subsubsection{Only GS locations}\label{only-gs-locations-1}}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/hot-gs-locations-1} \end{center}
+
+\hypertarget{figure-5}{%
+\subsubsection{Figure 5}\label{figure-5}}
+
+\begin{itemize}
+\tightlist
+\item
+  Currently, this is the ``main'' figure. WIS and hotspots, 2 panels,
+  only 2020, honest training.
+\end{itemize}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/figure5-1} \end{center}
+
+\hypertarget{figure-6-misleading-retrospective-analysis}{%
+\subsubsection{Figure 6 (misleading retrospective
+analysis)}\label{figure-6-misleading-retrospective-analysis}}
+
+\begin{itemize}
+\tightlist
+\item
+  WIS and hotspots, 2 panels, only 2020, honest training.
+\end{itemize}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/figure6-1} \end{center}
+
+\hypertarget{correlations-with-lagged-actuals}{%
+\subsection{Correlations with lagged
+actuals}\label{correlations-with-lagged-actuals}}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/cor-wis-ratio-1} \end{center}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/cor-wis-ratio-m1-1} \end{center}
+
+\hypertarget{upswings-vs-downswings}{%
+\subsection{Upswings vs downswings}\label{upswings-vs-downswings}}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/upswing-histogram-1} \end{center}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/upswing-histogram-logged-1} \end{center}
+
+\begin{tabular}[t]{lrrrrr}
+\toprule
+udf & CHNG-CLI & CHNG-COV & CTIS-CLIIC & DV-CLI & Google-AA\\
+\midrule
+down & 0.79 & 0.80 & 0.83 & 0.81 & 0.81\\
+flat & 0.12 & 0.17 & 0.28 & 0.18 & 0.17\\
+up & -0.57 & -0.56 & -0.53 & -0.53 & -0.47\\
+\bottomrule
+\end{tabular}
+
+\hypertarget{leadingness-and-laggingness}{%
+\subsection{Leadingness and
+laggingness}\label{leadingness-and-laggingness}}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/leading-and-lagging-1} \end{center}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/diff-in-lead-lag-1} \end{center}
+
+\hypertarget{side-figures-in-their-own-scripts}{%
+\section{Side figures (in their own
+scripts)}\label{side-figures-in-their-own-scripts}}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/ccf-dv-finalized-1} \end{center}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/ny-trajectory-1} \end{center}
+
+\begin{center}\includegraphics[width=\linewidth]{fig/revisions-dv-jhu-1} \end{center}
+
 %%% Add this line AFTER all your figures and tables
@@
 
-\bibliography{pnas-sample}
+\bibliography{../../common/covidcast.bib,pnas-materials/pnas-sample.bib}
 

### 2021-06-16 forecast/paper/paper.tex
Revise dated and add figs marked TODO to manuscript.

@@
 \begin{figure}[tb!]
-\includegraphics[width=\columnwidth]{fig/vintage.pdf}
-\caption{TODO}
+\includegraphics[width=\columnwidth]{fig/revisions-dv-jhu-1.pdf}
+\caption{Revision behavior for signals in the New York City HRR from two
+  sources. Each colored line 
+  corresponds to the data as reported on that date. DV-CLI is revised upward
+  throughout the month of December, though the amount fades as we look further
+  back in time. In contrast, JHU-CSSE data remains ``as reported'' on
+  December 3 until a major correction is made on December 31, affecting all data
+  over the period. }
 \label{fig:vintage}
@@
 \begin{figure}[t]
-\includegraphics[width=\columnwidth]{fig/trajectory.pdf}
-\caption{TODO}
+\includegraphics[width=\columnwidth]{fig/ny-trajectory-1.pdf}
+\caption{Forecast from the autoregressive model for the New York City HRR made
+  on October 15 (vertical line). 
+  The fan displays 50\%, 80\% and 95\% intervals while the orange curve shows
+  the median forecast. The connected black points represent ``finalized'' data,
+  as reported in May 2021.}
 \label{fig:trajectory}
@@
   \includegraphics[width=\columnwidth]{fig/fcast-1.pdf}
-\caption{TODO}
+  \caption{Average WIS for each forecast model divided by the averaged WIS of
+    the flat-line forecaster.}
   \label{fig:forecast}
@@
 \includegraphics[width=\columnwidth]{fig/hot-1.pdf}
-\caption{TODO}
+\caption{Area under the curve for each forecast model averaged over all
+  prediction dates and HRRS.}
 \label{fig:hotspot}
@@
 
-TODO
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{fig/upswing-histogram-1.pdf}
+  \caption{Histogram of the difference in WIS for the AR model relative to each
+    forecaster. Larger values are better for the indicator assisted model.}
+  \label{fig:up-down-flat}
+\end{figure}
 
@@
 
-TODO
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{fig/leading-and-lagging-1.pdf}
+  \caption{Correlation of the difference in WIS with the ``leadingness'' and
+    ``laggingness'' of the indicator at the target date.}
+  \label{fig:leading-lagging}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{fig/diff-in-lead-lag-1.pdf}
+  \caption{Correlation of the difference in WIS with difference between the
+    ``leadingness'' and ``laggingness'' of the indicator at the target date.}
+  \label{fig:diff-btw-leading-lagging}
+\end{figure}
 

### 2021-06-16 forecast/code/revisions-dv-jhu.R

@@ -19,7 +19,9 @@ dv_as_of <-  map_dfr(as_ofs, function(as_of) {
 
 
 pp <- bind_rows(cases_as_of, dv_as_of) %>% 
-  mutate(as_of = fct_relabel(factor(as_of), function(x) strftime(x, "%b %d"))) %>%
+  mutate(as_of = fct_relabel(factor(as_of), function(x) strftime(x, "%b %d")),
+         data_source = recode(data_source, `doctor-visits` = "DV-CLI",
+                              `jhu-csse` = "JHU-CSSE")) %>%
   ggplot(aes(x = time_value, y = value)) + 
   geom_line(aes(color = factor(as_of))) + 
   facet_wrap(~data_source, scales = "free_y") +

### 2021-06-16 forecast/paper/paper.tex
Update discussion (unfinished)

--- 2021-06-16 forecast/paper/paper.tex
@@
 one  might have hoped for, revolutionizing the tractability of the prediction 
-problem, rendering it ``easy'' when it was once ``hard'', in the absence of
-auxiliary information.  Rather, the gains in accuracy offered by the indicator
-models appear to be nontrivial, and consistent across modes of analysis, but
-modest---and more about tipping the bias-variance tradeoff in our favor than
-providing a consistent leading indication of future events.
+problem, rendering it easy when it was once hard (in the absence of auxiliary 
+information).  Rather, the gains in accuracy offered by the indicator models 
+(over an autoregressive model based on past case rates along) appear to be
+nontrivial, and consistent across modes of analysis, but modest.  In
+forecasting, the indicator models appear to be most useful in periods in which 
+case rates are flat or trending down, rather than periods in which case rates
+are trending up (as one might hope to see, as the hypothetical benefit provided
+by an idealized ``leading indicator'').
 
-[TODO:  Explain how insights from up/down/flat, lagging/leading, etc., are
-relevant to this overall arc.  Reiterate the importance of vintage data.  Finish
-with something about what's left for future work?  ]
+As mentioned previously, the indicator models could certainly benefit from a 
+careful handling of location-specific effects (to account for spatial
+heterogeneity in the search trends and claims-based indicators in particular),
+as well as nowcasting techniques to estimate finalized indicator values before
+using them as features (to account for backfill in the claims-based indicators
+in particular).  Beyond this, it is certainly possible that more sophisticated 
+models for forecasting or hotspot prediction would lead to different results,
+and potentially even different insights.  Natural directions to explore include  
+using interactions, using multiple indicators in tandem in a single model, and 
+leveraging HRR demographics or mobility patterns.  That said, we are doubtful
+that the ``topline'' conclusion would change--that auxiliary indicators can
+provide a nontrivial, consistent, but modest improvement in forecasting and
+hotspot prediction.
+
+Furthermore, from our experience working on forecasting  TODO
+
+[  Finish by reiterating the importance of vintage data.  Emphasizing the rigor
+or our approach, how we have provided a template for others to be able to do the
+same thing with new models ]
 

### 2021-06-16 indicators/paper/paper.tex
No need for centering
Minor copyedits

@@
 
+% TODO: Cite this? https://arxiv.org/abs/2106.04420
+
 \subsection{Public API}
@@
 \begin{figure}[t]
-\centering
 \includegraphics[width=\columnwidth]{fig/time_trends_national.pdf}
@@
 Figure~\ref{fig:geo_wise_correlation} shows the geo-wise correlations achieved
-by the five signals and COVID case rates (from USAFacts, smoothed using a 7-day
+by the five signals and COVID case rates (from JHU CSSE, smoothed using a 7-day
 trailing average), from April 15, 2020 to April 15, 2021. This calculation is
@@
 \begin{figure}[t]
-\centering
 \includegraphics[width=\columnwidth]{fig/bexar_compare.pdf}
@@
 \begin{figure}[t]
-\centering
 \includegraphics[width=\columnwidth]{fig/dv_as_of.pdf}

### 2021-06-16 indicators/paper/paper.tex
Add in NCHS citation
Try to finish off the discussion

@@
 COVID, flu, or pneumonia, compiled by the National Center for Health Statistics
-\cite{TODO}. & 2020-01-26 & State \\ 
+\cite{NCHS}. & 2020-01-26 & State \\
   \bottomrule
@@
 The COVIDcast API provides open access to real-time and geographically-detailed
-indicators that support and enhance standard public health reporting streams in 
-several ways.
+indicators of COVID activity in the United States that support and enhance
+standard public health reporting streams in several ways.
 
 First, several signals in the API closely track COVID activity (over both time
-and space); yet they are derived from different data streams (e.g., surveys,
-medical insurance claims, medical devices), and are therefore not subject to the
-same sources of error as public health reporting data.  This can be important
-both for downstream modeling purposes (enhancing the accuracy in nowcasting and
-forecasting models; see our companion paper on forecasting) and also for
-robustness (providing sanity checks when faced with potential anomalies in
-standard surveillance streams).
+and space); yet they are derived from different data streams (such as surveys,
+medical insurance claims, and medical devices), and are therefore not subject to
+the same sources of error as public health reporting data. This can be important
+both for robustness and situational awareness, allowing decision-makers to
+diagnose potential anomalies in standard surveillance streams, and for modeling
+tasks such as forecasting and nowcasting. Our companion paper on forecasting
+discusses this in more detail.
 
@@
 aspects of the pandemic and its effects on the United States population that are
-not found in traditional public health streams, e.g., data on mobility patterns,
-internet search trends, mask wearing, and vaccine hesitancy, to name just a few
-(the last two are derived from the COVID Trends and Impact Survey; see our
-companion paper on this survey for a detailed view of its features and
-capabilities).
+not found in traditional public health streams, such as data on mobility
+patterns, Internet search trends, mask wearing, and vaccine hesitancy, to name
+just a few. (The latter two are derived from the COVID Trends and Impact Survey;
+see our companion paper on this survey for a detailed view of its features and
+capabilities.) These signals have already supported pandemic research and
+policy-making.
 
@@
 us to query the API to learn ``what was known when'', critical for understanding
-the behavior (and potential pitfalls) of real-time surveillance signals.
+the behavior (and potential pitfalls) of real-time surveillance signals. Such
+revision data is rarely available in standardized form from other sources.
 
 Finally, we emphasize that unifying many relevant signals into a single common
-format, with comprehensive revision tracking, is a valuable goal in it of
+format, with comprehensive revision tracking, is a valuable goal in and of
 itself. No one data source is perfect; each is subject to error. Bringing
@@
 convenient as possible for public health researchers, policy makers, modelers,
-etc., is a strong testament to our belief in ``the whole being greater than the
+etc., is a strong testament to our belief that ``the whole is greater than the
 sum of its parts''.
 
-There are a number of open questions, and challenges that remain.  
-% TODO:
-% RJT: we should write quick bullet points on 1. survey bias, 2. geo bias in
-% claims data, 3. backfill and holiday corrections in claims data. 
-% @AR: can you give this a stab?  You can also delegate (Aaron e.g. should be
-% able to write something on the claims stuff.)  Let's keep it succint!
+There are a number of open questions, and challenges that remain. Each signal is
+subject to different biases, such as survey sampling and nonresponse biases or
+geographic differences in market share for medical claims data. Claims data can
+also be subject to biases during major national holidays and other events that
+change health-seeking behavior. Characterizing these biases will be important
+for future research using these signals. Several data sources are also subject
+to extensive revision and backfill, which must be studied and modeled to enable
+real-time use of these sources in forecasting and nowcasting systems. The
+breadth and unique features of the COVIDcast API will help facilitate this work,
+which will be vital to advancing pandemic modeling and preparedness.
 
-\matmethods{Please describe your materials and methods here. This can be more than one paragraph, and may contain subsections and equations as required.
+% \matmethods{Please describe your materials and methods here. This can be more than one paragraph, and may contain subsections and equations as required.
 
-\subsection*{Subsection for Method}
-Example text for subsection.
-}
+% \subsection*{Subsection for Method}
+% Example text for subsection.
+% }
 
-\showmatmethods{} % Display the Materials and Methods section
+% \showmatmethods{} % Display the Materials and Methods section
 

### 2021-06-16 forecast/code/ccf-dv-finalized.R
Fix some figure titles.

@@ -15,16 +15,17 @@ names(sigs)[3:4] <- c("cases", "doctor-visits")
 
 std_sigs <- sigs %>%
   mutate(across(c("cases","doctor-visits"), ~ (.x - mean(.x)) / sd(.x))) %>%
+  rename(`JHU-CSSE` = cases, `DV-CLI` = `doctor-visits`) %>%
   select(-geo_value) %>%
   pivot_longer(-time_value)
   
 g1 <- std_sigs %>%
   ggplot(aes(time_value, value, color = name)) +
   geom_line() +
-  theme_bw(base_size = 14) +
-  xlab("date") + ylab("standardized signal") +
+  theme_bw() +
+  xlab("Date") + ylab("Standardized signal") +
   scale_color_viridis_d(begin=.25, end=.75, name = "") +
-  theme(legend.position = "right")
+  theme(legend.position = "bottom")
 
 cc <- ccf(sigs$`doctor-visits`, sigs$cases, plot = FALSE)
 g2 <- ggplot(tibble(lag = drop(cc$lag), ccf = drop(cc$acf)), aes(lag)) +
@@ -33,9 +34,13 @@ g2 <- ggplot(tibble(lag = drop(cc$lag), ccf = drop(cc$acf)), aes(lag)) +
              color = "darkorange", linetype = "dashed") +
   geom_hline(yintercept = 0) +
   geom_vline(xintercept = 0) +
-  ylab("cross correlation") +
-  theme_bw(base_size = 14) 
+  ylab("Cross correlation") +
+  xlab("Lag") +
+  theme_bw() +
+  theme(legend.position = "none")
 
+leg <- cowplot::get_legend(g1)
 
-gg <- cowplot::plot_grid(g2, g1, rel_widths = c(4,6))
+gg <- cowplot::plot_grid(g2, g1 + theme(legend.position = "none"))
+gg <- cowplot::plot_grid(gg, leg, nrow = 2, rel_heights = c(1,.1))
 

### 2021-06-16 indicators/paper/paper.tex
Cite recent preprint on backfill in COVID
Add a new perspective

@@
   Reich:2019}, and that careful training on preliminary data can reduce this
-influence \cite{Brooks:2020}. We also investigate this in our companion paper on
-forecasting, where we observe that training and validating models on finalized 
-data yields overly optimistic estimates of true test-time performance.
-
-% TODO: Cite this? https://arxiv.org/abs/2106.04420
+influence \cite{Brooks:2020}. Recent research has shown similar results for
+COVID-19 forecasts \cite{Kamarthi:2021}. We also investigate this in our
+companion paper on forecasting, where we observe that training and validating
+models on finalized data yields overly optimistic estimates of true test-time
+performance.
 
@@
 reflects healthcare-seeking behavior and how it may be affected by local case
-rates; measures of COVID vaccine acceptance can guide outreach efforts
-\cite{TODO}.
+rates; measures of COVID vaccine acceptance can guide outreach efforts.
 
-% TODO
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{fig/mobility-drop.pdf}
+  \caption{Percentage point increase in SafeGraph estimates of the percentage of
+  people staying completely at home, based on aggregated mobile device data,
+  between March 1, 2020 and April 15, 2020.}
+  \label{fig:mobility-drop}
+\end{figure}
 
-% AR: Potentially interesting things to show or comment on:
-% - Medical claims data can give info about medical-seeking behavior and how it
-% varies with caseload -- e.g. are people deterred from seeking medical care
-% because of the high case load?
-% - Mobility changes
-% - Map of one of the unique signals during an outbreak, e.g. Michigan in April
-% 2021
-
-% RJT: @AR Since we already stuff on claims, can we do: mobility, masks, vaccine
-% uptake or hesitancy, distancing behaviors, how worried are you?
-% How about this (in each case, each point in said plot is a county's signal
-% values averaged over some period)
-% - Bivariate plot of mask use versus how worried are you
-% - Bivariate plot of of how worried are you vs case rates
-%- Bivariate plot of of how worried are you vs hospitalizations
-% I would be curious to see all three.  And perhaps color code the points by
-% region to highlight regional trends
+As an illustration, Figure~\ref{fig:mobility-drop} maps the sharp increase in
+rates of people staying at home between March 1 and April 15, 2020. Using
+SafeGraph data on the fraction of mobile devices included in SafeGraph's panel
+that did not leave the immediate area of their home, it illustrates the sharp
+drop in travel and work outside the home that occurred in the early stages of
+the pandemic. It also shows that this drop was much more pronounced in some
+states than others, enabling analysis of policy impacts and disease spread.
+Similar maps can be quickly constructed for any signal in the COVIDcast API.
 

### 2021-06-16 indicators/paper/paper.tex
Copyedits to discussion

@@
 influence \cite{Brooks:2020}. Recent research has shown similar results for
-COVID-19 forecasts \cite{Kamarthi:2021}. We also investigate this in our
-companion paper on forecasting, where we observe that training and validating
-models on finalized data yields overly optimistic estimates of true test-time
+COVID-19 forecasts \cite{Kamarthi:2021}. We also examine this in our companion  
+paper on forecasting, where we observe that training and validating models on
+finalized data yields overly optimistic estimates of true test-time
 performance.
@@
 The COVIDcast API provides open access to real-time and geographically-detailed
-indicators of COVID activity in the United States that support and enhance
+indicators of COVID activity in the United States, which supports and enhances
 standard public health reporting streams in several ways.
@@
 and space); yet they are derived from different data streams (such as surveys,
-medical insurance claims, and medical devices), and are therefore not subject to
-the same sources of error as public health reporting data. This can be important
+medical insurance claims, and medical devices), and are thus not subject to the
+same sources of error as public health reporting streams. This can be important  
 both for robustness and situational awareness, allowing decision-makers to
@@
 patterns, Internet search trends, mask wearing, and vaccine hesitancy, to name
-just a few. (The latter two are derived from the COVID Trends and Impact Survey;
-see our companion paper on this survey for a detailed view of its features and
-capabilities.) These signals have already supported pandemic research and
-policy-making.
+just a few. (The latter two signals are derived from the COVID Trends and Impact
+Survey; our companion paper on this survey gives more a detailed view of its
+features and capabilities.) These signals have already supported pandemic
+research and  policy-making.
 
@@
 the behavior (and potential pitfalls) of real-time surveillance signals. Such
-revision data is rarely available in standardized form from other sources.
+revision data is rarely available in standardized format from other sources.
 
@@
 
-There are a number of open questions, and challenges that remain. Each signal is
-subject to different biases, such as survey sampling and nonresponse biases or
-geographic differences in market share for medical claims data. Claims data can
-also be subject to biases during major national holidays and other events that
-change health-seeking behavior. Characterizing these biases will be important
-for future research using these signals. Several data sources are also subject
-to extensive revision and backfill, which must be studied and modeled to enable
-real-time use of these sources in forecasting and nowcasting systems. The
-breadth and unique features of the COVIDcast API will help facilitate this work,
-which will be vital to advancing pandemic modeling and preparedness.
+There are a number of open questions, and challenges that remain. Several
+signals are subject to biases, such as survey sampling and nonresponse biases,
+or geographic differences in market share for medical claims data. Claims data
+can also be subject to biases during major national holidays and other events
+that change health-seeking behavior. Characterizing these biases will be
+important for future research and operational systems that use these
+signals. Several data sources are also subject to extensive revision and
+backfill, which must be studied and modeled to enable effective real-time use of
+these sources in forecasting and nowcasting systems. The breadth and unique
+features of the COVIDcast API will help facilitate this work, which will be
+vital to advancing pandemic modeling and preparedness.
 

### 2021-06-16 forecast/paper/pnas-suppl-template.tex
add code, begin supplement generation
add supplement and rebuilt figs
reorganize supplement; clean up all figures

--- 2021-06-11 forecast/paper/supplement.tex
@@
 
-\input{forecast-prelim.tex}
+\input{../../common/prelim.tex}
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+$for(header-includes)$
+$header-includes$
+$endfor$
 
@@
 
-%% Comment out or remove this line before generating final copy for submission; this will also remove the warning re: "Consecutive odd pages found".
-\instructionspage
+$body$
 
-\maketitle
-
-%% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
-\SItext
-
-\subsection*{Subhead}
-Type or paste text here. This should be additional explanatory text such as an extended technical description of results, full details of mathematical models, etc.
-
-\section*{Heading}
-\subsection*{Subhead}
-Type or paste text here. You may break this section up into subheads as needed (e.g., one section on ``Materials'' and one on ``Methods'').
-
-\subsection*{Materials}
-Add a materials subsection if you need to.
-
-\subsection*{Methods}
-Add a methods subsection if you need to.
-
-%%% Each figure should be on its own page
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{example-image}
-\caption{First figure}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{frog}
-\caption{Second figure}
-\end{figure}
-
-\begin{table}\centering
-\caption{This is a table}
-
-\begin{tabular}{lrrr}
-Species & CBS & CV & G3 \\
-\midrule
-1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
-2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
-3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-%%% Add this line AFTER all your figures and tables
-\FloatBarrier
-
-\movie{Type legend for the movie here.}
-
-\movie{Type legend for the other movie here. Adding longer text to show what happens, to decide on alignment and/or indentations.}
-
-\movie{A third movie, just for kicks.}
-
-\dataset{dataset_one.txt}{Type or paste legend here.}
-
-\dataset{dataset_two.txt}{Type or paste legend here. Adding longer text to show what happens, to decide on alignment and/or indentations for multi-line or paragraph captions.}
-
-\bibliography{pnas-sample}
+\bibliography{$for(bibliography)$$bibliography$$sep$,$endfor$}
 

### 2021-06-16 forecast/paper/supplement.Rmd
reorganize supplement; clean up all figures

@@
 
+<!-- PNAS materials -->
+
+<!-- Comment out or remove this line before generating final copy for submission; this will also remove the warning re: "Consecutive odd pages found". -->
+\instructionspage
+
+<!-- we need this -->
+\maketitle
+
+<!-- Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text. -->
+
+\SItext
+
+All our text goes here. Reference figures with the R chuck label as Figure~\ref{fig:fcast-finalized}.
+
+All figures go on their own page after all of the text...
+
+Edit `pnas-suppl-template.tex` for the correct Author list and Title.
+
+# Examining the relative advantage of using finalized rather than vintage data
+
+This section referes to Figures~\ref{fig:fcast-finalized}--\ref{fig:hot-honest-v-finalized}.
+
+# Aggregating with geometric mean
+
+* The weighted interval score is bounded below by zero and can be very large. This behavior is typical of right-skewed distributions.
+* Figure~\ref{fig:wis-densities} illustrates that the densities appear log-Gaussian.
+* This suggests aggregating by the geometric mean rather than the mean for comparisons. See Figure~\ref{fig:fcast-adjusted}.
+
+# Bootstrap results
+
+Here we discuss how implicit regularization is not the reason for improved performance
+
+# Correlations with lagged actuals
+
+Alden's histograms are in Figure~\ref{fig:cor-wis-ratio} and Figure~\ref{fig:cor-wis-ratio-m1}.
+
+# Upswings and Downswings
+
+Logged version of Figure 5 in the manuscript is in Figure~\ref{fig:upswing-histogram-logged}.
+
+See also Table~\ref{tab:upswing-corr-table}.
+
+# Leadingness and laggingness
+
+Currently, both figures are in the manuscript. Probably just need text here.
+
+# Examining data in 2021
+
+See Figures \ref{fig:fcast-alldates} -- \ref{fig:hot-alldates}.
+
+# Deprecated
+
+There are a few blocks at the bottom (figures with Google symptoms only and the
+old trajectory plots) that we can remove once we decide.
+
 ```{r setup, include=FALSE}
@@
                       cache = TRUE, 
-                      out.width = "\\linewidth",
+                      out.width = "\\textwidth",
                       autodep = TRUE,
@@
 ```{r source-funs, cache=FALSE}
+library(scales)
+library(cowplot)
+library(covidcast)
+library(tidyverse)
+library(lubridate)
+library(evalcast)
+
 source("../code/eval_funs.R")
 source("../code/hotspot_funs.R")
-source("../code/trajectory_plot_funs.R")
-library(scales)
-library(cowplot)
+
 path_to_data <- "~/Documents/COVID-Delphi/fcast-indicators-paper/temp_data/"
@@
 
-# Figures from results
-
-## Forecasts
-
-```{r fcast}
-fig5a <- plotter(fcasts_honest %>% filter(period != "jm"), 
-        "wis", Mean, scaler = "strawman_wis", 
-        order_of_operations = c("aggr","scale")) +
-  ylab("Mean WIS (relative to baseline)")
-fig5a
-```
-
-```{r fcast-finalized}
-plotter(fcasts_finalized %>% filter(period != "jm"), 
-        "wis", Mean, scaler = "strawman_wis", 
-        order_of_operations = c("aggr","scale")) +
-  ylab("Mean WIS (relative to baseline)")
-```
-
-```{r fcast-honest-v-finalized}
-fig6a <- plotter(
-  left_join(
-    fcasts_honest %>% filter(period != "jm") %>% select(forecaster:wis),
-    fcasts_finalized %>% filter(period != "jm") %>% select(forecaster:wis) %>%
-      rename(finalized_wis = wis)
-    ), 
-  "wis", Mean, scaler = "finalized_wis", 
-  order_of_operations = c("aggr","scale")) +
-  geom_hline(yintercept = 1) +
-  ylab("Mean WIS (vintage / finalized)")
-fig6a
-```
-
-```{r fcast-adjusted}
-plotter(fcasts_honest %>% filter(period != "jm"), 
-        "wis", GeoMean, scaler = "strawman_wis", 
-        order_of_operations = c("scale","aggr")) +
-  ylab("Geometric mean of WIS (relative to baseline)")
-```
-
-```{r wis-densities}
-ggplot(fcasts_honest %>% filter(period != "jm"), 
-       aes(wis, fill = forecaster)) +
-  geom_density() +
-  scale_x_log10() +
-  theme_bw() +
-  xlab("WIS") +
-  scale_fill_manual(values = fcast_colors) +
-  facet_wrap(~forecaster) +
-  theme(legend.position = "none")
-```
-
-### All periods
-
-```{r fcast-alldates}
-plotter(fcasts_honest,
-        "wis", Mean, scaler = "strawman_wis", 
-        order_of_operations = c("aggr","scale")) +
-  ylab("Mean WIS (relative to baseline)")
-```
-
-```{r fcast-alldates-adjusted}
-plotter(fcasts_honest, 
-        "wis", GeoMean, scaler = "strawman_wis", 
-        order_of_operations = c("scale","aggr")) +
-  ylab("Geometric mean of WIS (relative to baseline)")
-```
-
-### Only GS locations
-
-```{r fcast-gs-locations}
-plotter(fcasts_gs %>% filter(period != "jm"),
-        "wis", Mean, scaler = "strawman_wis", 
-        order_of_operations = c("aggr","scale")) +
-  ylab("Mean WIS (relative to baseline)")
-```
-
-```{r fcast-gs-locations-adjusted}
-plotter(fcasts_gs %>% filter(period != "jm"), 
-        "wis", GeoMean, scaler = "strawman_wis", 
-        order_of_operations = c("scale","aggr")) +
-  ylab("Geometric mean of WIS (relative to baseline)")
-```
-
-## Trajectories (deprecated)
-
 ```{r hrr-names}
@@
 
-```{r traj-data, eval = FALSE}
-preds <- readRDS(file.path(path_to_data, "predictions_honest.RDS"))
-traj_best <- get_trajectory_plots(fcasts_honest, preds, actuals, hrr_tab, 
-                                  "only2020", "best")
-traj_worst <- get_trajectory_plots(fcasts_honest, preds, actuals, hrr_tab,
-                                   "only2020", "worst")
-rm(preds)
+<!-- All figures and tables below this line -->
+
+<!-- Finalized v. vintage -->
+
+```{r fcast, include=FALSE}
+# Figure 3 in the paper
+plotter(fcasts_honest %>% filter(period != "jm"), 
+        "wis", Mean, scaler = "strawman_wis", 
+        order_of_operations = c("aggr","scale")) +
+  ylab("Mean WIS (relative to baseline)")
 ```
 
-```{r trajectory-plots, eval = FALSE}
-for (nm in names(traj_best)) print(trajectory_panel(nm, traj_best, traj_worst))
+```{r hot, include=FALSE}
+# Figure 4 in the manuscript
+plotter_hotspots(hotspots_honest %>% filter(period != "jm")) + 
+  geom_hline(yintercept = 0.5)
 ```
 
-## Hotspots
-
-```{r hot}
-fig5b <- plotter_hotspots(hotspots_honest %>% filter(period != "jm")) + 
-  geom_hline(yintercept = 0.5)
-fig5b
+```{r fcast-finalized, fig.cap="Forecasting performance using finalized data. Compare to Figure 3 in the manuscript."}
+plotter(fcasts_finalized %>% filter(period != "jm"), 
+        "wis", Mean, scaler = "strawman_wis", 
+        order_of_operations = c("aggr","scale")) +
+  ylab("Mean WIS (relative to baseline)")
 ```
 
-```{r hot-finalized}
+\clearpage
+
+```{r hot-finalized, fig.cap="Hotspot prediction performance using finalized data. Compare to Figure 4 in the manuscript."}
 plotter_hotspots(hotspots_finalized %>% filter(period != "jm")) +
@@
 
-```{r hot-honest-v-finalized}
-fig6b <- left_join(
+\clearpage
+
+```{r fcast-honest-v-finalized, fig.cap="Relative forecast WIS with vintage compared to finalized data. Using finalized data leads to overly optimistic performance."}
+plotter(
+  left_join(
+    fcasts_honest %>% filter(period != "jm") %>% select(forecaster:wis),
+    fcasts_finalized %>% filter(period != "jm") %>% select(forecaster:wis) %>%
+      rename(finalized_wis = wis)
+    ), 
+  "wis", Mean, scaler = "finalized_wis", 
+  order_of_operations = c("aggr","scale")) +
+  geom_hline(yintercept = 1) +
+  ylab("Mean WIS (vintage / finalized)")
+
+```
+
+\clearpage
+
+```{r hot-honest-v-finalized, fig.cap="Relative AUC with vintage compared to finalized data. Using finalized data leads to overly optimistic hotspot performance."}
+left_join(
   hotspots_honest %>%
@@
   ylab("AUC (vintage / finalized)")
-fig6b
 ```
 
-### All periods
+\clearpage
 
-```{r hot-alldates}
-plotter_hotspots(hotspots_honest)
+<!-- Geometric mean justifications -->
+
+```{r wis-densities, fig.cap="Weighted interval score appears to more closely resemble a log-Gaussian distribution."}
+ggplot(fcasts_honest %>% filter(period != "jm"), 
+       aes(wis, fill = forecaster)) +
+  geom_density() +
+  scale_x_log10() +
+  theme_bw() +
+  xlab("WIS") +
+  scale_fill_manual(values = fcast_colors) +
+  facet_wrap(~forecaster) +
+  theme(legend.position = "none")
 ```
 
-### Only GS locations
+\clearpage
 
-```{r hot-gs-locations}
-plotter_hotspots(hotspots_gs %>% filter(period != "jm")) +
+```{r fcast-adjusted, fig.cap="Relative forecast performance using vintage data and summarizing with the more robust geometric mean."}
+plotter(fcasts_honest %>% filter(period != "jm"), 
+        "wis", GeoMean, scaler = "strawman_wis", 
+        order_of_operations = c("scale","aggr")) +
+  ylab("Geometric mean of WIS (relative to baseline)")
+```
+
+\clearpage
+
+<!-- Bootstrap stuff -->
+
+```{r bootstrap-loading}
+fcasts_booted <- readRDS(
+  file.path(path_to_data,"results_honest_bootstrapped.RDS")) %>%
+  process_res_cases(actuals) %>%
+  filter(forecaster != "gs_inherit")
+if (params$fd_casefloor >= 0) {
+  fcasts_booted <- filter_case_floor(fcasts_booted, cases7dav, params)
+}
+hotspots_booted <- readRDS(
+  file.path(path_to_data, "hotspots_honest_bootstrapped.RDS")) %>%
+  process_res_hotspots() %>%
+  filter(forecaster != "gs_inherit")
+if (params$flag_jumps > 0) {
+  hotspots_booted <- filter_jumps(hotspots_booted, actuals, params)
+  fcasts_booted <- filter_jumps(fcasts_booted, actuals, params)
+}
+fcasts_booted <- filter(fcasts_booted, forecaster != "gs_subset") %>%
+  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
+hotspots_booted <- intersect_averagers( # Deal with DV NaN's
+  hotspots_booted %>% 
+    filter(! is.na(value), !is.na(actual), forecaster != "gs_subset"), 
+  "forecaster", 
+  c("geo_value", "ahead", "forecast_date", "target_end_date")) %>%
+  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
+```
+
+```{r fcast-booted, fig.cap = "Forecast performance when indicators are replaced with samples from their empirical distribution. Performance is largely similar to the AR model."}
+plotter(fcasts_booted %>% filter(period != "jm"), 
+        "wis", Mean, scaler = "strawman_wis", 
+        order_of_operations = c("aggr","scale")) +
+  ylab("Mean WIS (relative to baseline)")
+```
+
+\clearpage
+
+```{r fcast-booted-adjusted, fig.cap = "Forecast performance as measured with the geometric mean when indicators are replaced with samples from their empirical distribution.
+Performance is largely similar to the AR model."}
+plotter(fcasts_booted %>% filter(period != "jm"), 
+        "wis", GeoMean, scaler = "strawman_wis", 
+        order_of_operations = c("scale","aggr")) +
+  ylab("Geometric mean of WIS (relative to baseline)")
+```
+
+\clearpage
+
+```{r hot-booted, fig.cap = "Hotspot prediction performance when indicators are replaced with samples from their empirical distribution. Performance is largely similar to the AR model."}
+plotter_hotspots(hotspots_booted %>% filter(period != "jm")) + 
   geom_hline(yintercept = 0.5)
@@
 
-### Figure 5
+\clearpage
 
-* Currently, this is the "main" figure. WIS and hotspots, 2 panels, only 2020, honest training.
-
-```{r figure5}
-leg <- get_legend(fig5a)
-fig5a <- fig5a + theme(legend.position = "none")
-fig5b <- fig5b + theme(legend.position = "none")
-fig5 <- plot_grid(fig5a, fig5b, align = "hv", nrow = 1)
-fig5 <- plot_grid(fig5, leg, rel_heights = c(1,.1), nrow = 2)
-fig5
-```
-
-### Figure 6 (misleading retrospective analysis)
-
-* WIS and hotspots, 2 panels, only 2020, honest training.
-
-```{r figure6}
-fig6a <- fig6a + theme(legend.position = "none")
-fig6b <- fig6b + theme(legend.position = "none")
-fig6 <- plot_grid(fig6a, fig6b, align = "hv", nrow = 1)
-fig6 <- plot_grid(fig6, leg, rel_heights = c(1,.1), nrow = 2)
-fig6
-```
-
-## Correlations with lagged actuals
+<!-- Correlations with lagged actuals -->
 
 ```{r cor-with-actuals}
-pct_change <- function(x, n = 14, col_name = "pct_change"){
-  # courtesy of Alden
-  dt <- c(0,-n)
-  y = covidcast:::apply_shifts(x,dt) %>%
-    rename(value = tidyselect::matches("value\\+"),
-           lag_value = tidyselect::matches("value\\-")) %>%
-    mutate(`:=`(!!col_name,(value - lag_value)/lag_value))
-  return(y)
-}
 comparison_df <- left_join(
@@
 
-```{r cor-wis-ratio}
+```{r cor-wis-ratio, fig.cap="This is one of the correlation plots Alden made.
+It shows histograms of the Spearman correlation between the ratio of AR to AR WIS with the percent change in 7dav cases relative to 7 days earlier."}
 ggplot(pct_chng_df %>% 
@@
 
-```{r cor-wis-ratio-m1}
+\clearpage
+
+```{r cor-wis-ratio-m1, fig.cap="This is Alden's second set of histograms.
+Here we have the correlation of the absolute value of WIS ratio - 1 with the percent change in 7dav cases relative to 7 days earlier"}
 ggplot(pct_chng_df %>% 
@@
 
-## Upswings vs downswings
+\clearpage
+
+<!-- Upswings vs. downswings -->
 
@@
                              AR3CHCLI3 = "CHNG-CLI",
-                             AR3CHCOV3 = "CHNG-COV",
+                             AR3CHCOV3 = "CHNG-COVID",
                              AR3DVCLI3 = "DV-CLI",
@@
 
-```{r upswing-histogram}
+```{r upswing-histogram, include=FALSE}
+# Figure 5 in the manuscript
 df %>% 
@@
 
-```{r upswing-histogram-logged}
+```{r upswing-histogram-logged, fig.cap="Not sure if we want this here. Similar to Figure 5 in the manuscript but taking logs. "}
 df_summary <- df %>%
@@
 
+\clearpage
+
 ```{r upswing-corr-table}
@@
   pivot_wider(names_from = forecaster, values_from = cor) %>%
-  kableExtra::kbl(booktabs = TRUE, digits = 2)
+  kableExtra::kbl(
+    booktabs = TRUE, digits = 2, centering = TRUE,
+    caption = paste("Correlation of the difference in WIS between the AR",
+                    "model with the difference in median predictions. In down",
+                    "periods, improvements in forecast risk are highly",
+                    "correlated with lower median predictions. The opposite",
+                    "is true in up periods. This suggests, as one might expect",
+                    "that improved performance of the indicator-assisted",
+                    "model is attributable to being closer to the truth then",
+                    "the AR model. This conclusion is stronger in down",
+                    "periods then in up periods."))
 ```
 
-## Leadingness and laggingness
+\clearpage
+
+```{r hotspots-upswing-downswing, fig.cap="Classification and loglikelihood separated into periods of upswing, downswing, and flat cases.
+Like the analysis of the forecasting task in the main paper (see Figure 7), performance is better during down and flat periods."}
+hot_udf <- inner_join(
+  hotspots_honest %>% filter(period != "jm"),
+  up_down %>% select(geo_value, target_end_date, udf))
+
+cutoff <- mean(hot_udf %>% filter(forecaster == "AR") %>% pull(actual))
+
+con_tab <- hot_udf %>% 
+  filter(!is.na(actual)) %>%
+  mutate(pred = value > cutoff) %>%
+  group_by(forecaster, udf) %>%
+  summarise(m = mean(pred != actual)) %>%
+  ungroup()
+
+con_tab <- left_join(
+  con_tab %>% filter(forecaster != "AR"),
+  con_tab %>% filter(forecaster == "AR") %>%
+    select(-forecaster) %>%
+    rename(mar = m))
+
+llike_tab <- hot_udf %>%
+  filter(!is.na(actual)) %>%
+  mutate( # kill prob 0-1 predictions
+    value = case_when(
+      value < 1e-8 ~ 1e-8,
+      value > 1-1e-8 ~ 1-1e-8,
+      TRUE ~ value
+    ),
+    llike = (actual == 1) * log(value) + (actual == 0) * log(1 - value)) %>%
+  group_by(forecaster, udf) %>%
+  summarise(m = mean(llike)) %>%
+  ungroup()
+
+llike_tab <- left_join(
+  llike_tab %>% filter(forecaster != "AR"),
+  llike_tab %>% filter(forecaster == "AR") %>%
+    select(-forecaster) %>%
+    rename(mar = m))
+
+bind_rows(con_tab %>% mutate(err = "Classification error"), 
+          llike_tab %>% mutate(err = "Log likelihood")) %>%
+  ggplot(aes(udf, (m - mar) / abs(mar) , fill = forecaster)) +
+  geom_bar(width = 0.6, position = position_dodge(width=0.6),
+           stat = "identity") +
+  scale_fill_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
+  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
+  theme_bw() +
+  geom_hline(yintercept = 0) +
+  facet_wrap(~err) +
+  ylab("Change relative to AR") +
+  xlab("") +
+  theme(legend.position = "bottom", legend.title = element_blank())
+```
+
+\clearpage
+
+<!-- Leadingness and laggingness -->
 
@@
       `value+0:chng_smoothed_adj_outpatient_cli` = "CHNG-CLI",
-      `value+0:chng_smoothed_adj_outpatient_covid` = "CHNG-COV",
+      `value+0:chng_smoothed_adj_outpatient_covid` = "CHNG-COVID",
       `value+0:doctor-visits_smoothed_adj_cli` = "DV-CLI",
@@
 
-```{r leading-and-lagging}
+```{r leading-and-lagging, include=FALSE}
+# Figure 6 in the manuscript
 df2 %>%
@@
 
-```{r diff-in-lead-lag}
+```{r diff-in-lead-lag, include=FALSE}
+# Figure 7 in the manuscript
 df2 %>%
@@
 
-# Side figures (in their own scripts)
+<!-- Examining data in 2021 -->
 
-```{r ccf-dv-finalized}
-library(covidcast)
+```{r fcast-alldates, fig.cap="Forecast performance over all periods. Performance largely improves for all forecasters with the inclusion of data in 2021."}
+plotter(fcasts_honest,
+        "wis", Mean, scaler = "strawman_wis", 
+        order_of_operations = c("aggr","scale")) +
+  ylab("Mean WIS (relative to baseline)")
+```
+
+\clearpage
+
+```{r fcast-alldates-adjusted, fig.cap="Forcast performance over all periods aggregaged with the geometric mean. Again, the inclusion of data in 2021 leads to improved performance."}
+plotter(fcasts_honest, 
+        "wis", GeoMean, scaler = "strawman_wis", 
+        order_of_operations = c("scale","aggr")) +
+  ylab("Geometric mean of WIS (relative to baseline)")
+```
+
+\clearpage
+
+```{r hot-alldates, fig.cap="Area under the curve for hotspot predictions including data in 2021.
+Performance degrades relative to the period in 2020.
+However, there are far fewer hotspots during this period as case rates declined in much of the country."}
+plotter_hotspots(hotspots_honest)
+```
+
+\clearpage
+
+<!-- Figures in the manuscript -->
+
+```{r revisions-dv-jhu, include=FALSE}
+# Figure 1 in the manuscript
+source("../code/revisions-dv-jhu.R")
+pp
+```
+
+```{r ny-trajectory, include=FALSE}
+# Figure 2 in the manuscript
+source("../code/ny-trajectory.R")
+gg
+```
+
+```{r ccf-dv-finalized, include=FALSE}
+# Figure ?? in the manuscript
 source("../code/ccf-dv-finalized.R")
@@
 
-```{r ny-trajectory}
-source("../code/ny-trajectory.R")
-gg
+<!-- All code below is deprecated -->
+
+```{r fcast-gs-locations, eval=FALSE}
+plotter(fcasts_gs %>% filter(period != "jm"),
+        "wis", Mean, scaler = "strawman_wis", 
+        order_of_operations = c("aggr","scale")) +
+  ylab("Mean WIS (relative to baseline)")
 ```
 
-```{r revisions-dv-jhu}
-source("../code/revisions-dv-jhu.R")
-pp
+```{r fcast-gs-locations-adjusted, eval=FALSE}
+plotter(fcasts_gs %>% filter(period != "jm"), 
+        "wis", GeoMean, scaler = "strawman_wis", 
+        order_of_operations = c("scale","aggr")) +
+  ylab("Geometric mean of WIS (relative to baseline)")
 ```
+
+```{r hot-gs-locations, eval=FALSE}
+plotter_hotspots(hotspots_gs %>% filter(period != "jm")) +
+  geom_hline(yintercept = 0.5)
+```
+
+```{r traj-data, eval = FALSE}
+source("../code/trajectory_plot_funs.R")
+preds <- readRDS(file.path(path_to_data, "predictions_honest.RDS"))
+traj_best <- get_trajectory_plots(fcasts_honest, preds, actuals, hrr_tab, 
+                                  "only2020", "best")
+traj_worst <- get_trajectory_plots(fcasts_honest, preds, actuals, hrr_tab,
+                                   "only2020", "worst")
+rm(preds)
+```
+
+```{r trajectory-plots, eval = FALSE}
+for (nm in names(traj_best)) print(trajectory_panel(nm, traj_best, traj_worst))
+```
+
+\clearpage
+
+<!-- PNAS says: Add this line AFTER all your figures and tables -->
+\FloatBarrier
+
+<!-- More PNAS example material -->
+<!-- Probably remove all this -->
+
+\movie{Type legend for the movie here.}
+
+\dataset{dataset_one.txt}{Type or paste legend here.}
+
+\dataset{dataset_two.txt}{Type or paste legend here. Adding longer text to show what happens, to decide on alignment and/or indentations for multi-line or paragraph captions.}

### 2021-06-16 forecast/paper/supplement.tex
reorganize supplement; clean up all figures

--- 2021-06-11 forecast/paper/supplement.tex
@@
 
-%% Comment out or remove this line before generating final copy for submission; this will also remove the warning re: "Consecutive odd pages found".
 \instructionspage
@@
 
-%% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
 \SItext
 
-\subsection*{Subhead}
-Type or paste text here. This should be additional explanatory text such as an extended technical description of results, full details of mathematical models, etc.
+All our text goes here. Reference figures with the R chuck label as
+Figure\textasciitilde{}\ref{fig:fcast-finalized}.
 
-\section*{Heading}
-\subsection*{Subhead}
-Type or paste text here. You may break this section up into subheads as needed (e.g., one section on ``Materials'' and one on ``Methods'').
+All figures go on their own page after all of the text\ldots{}
 
-\subsection*{Materials}
-Add a materials subsection if you need to.
+Edit \texttt{pnas-suppl-template.tex} for the correct Author list and
+Title.
 
-\subsection*{Methods}
-Add a methods subsection if you need to.
+\hypertarget{examining-the-relative-advantage-of-using-finalized-rather-than-vintage-data}{%
+\section{Examining the relative advantage of using finalized rather than
+vintage
+data}\label{examining-the-relative-advantage-of-using-finalized-rather-than-vintage-data}}
 
-%%% Each figure should be on its own page
+This section referes to
+Figures\textasciitilde{}\ref{fig:fcast-finalized}--\ref{fig:hot-honest-v-finalized}.
+
+\hypertarget{aggregating-with-geometric-mean}{%
+\section{Aggregating with geometric
+mean}\label{aggregating-with-geometric-mean}}
+
+\begin{itemize}
+\tightlist
+\item
+  The weighted interval score is bounded below by zero and can be very
+  large. This behavior is typical of right-skewed distributions.
+\item
+  Figure\textasciitilde{}\ref{fig:wis-densities} illustrates that the
+  densities appear log-Gaussian.
+\item
+  This suggests aggregating by the geometric mean rather than the mean
+  for comparisons. See Figure\textasciitilde{}\ref{fig:fcast-adjusted}.
+\end{itemize}
+
+\hypertarget{bootstrap-results}{%
+\section{Bootstrap results}\label{bootstrap-results}}
+
+Here we discuss how implicit regularization is not the reason for
+improved performance
+
+\hypertarget{correlations-with-lagged-actuals}{%
+\section{Correlations with lagged
+actuals}\label{correlations-with-lagged-actuals}}
+
+Alden's histograms are in Figure\textasciitilde{}\ref{fig:cor-wis-ratio}
+and Figure\textasciitilde{}\ref{fig:cor-wis-ratio-m1}.
+
+\hypertarget{upswings-and-downswings}{%
+\section{Upswings and Downswings}\label{upswings-and-downswings}}
+
+Logged version of Figure 5 in the manuscript is in
+Figure\textasciitilde{}\ref{fig:upswing-histogram-logged}.
+
+See also Table\textasciitilde{}\ref{tab:upswing-corr-table}.
+
+\hypertarget{leadingness-and-laggingness}{%
+\section{Leadingness and
+laggingness}\label{leadingness-and-laggingness}}
+
+Currently, both figures are in the manuscript. Probably just need text
+here.
+
+\hypertarget{examining-data-in-2021}{%
+\section{Examining data in 2021}\label{examining-data-in-2021}}
+
+See Figures \ref{fig:fcast-alldates} -- \ref{fig:hot-alldates}.
+
+\hypertarget{deprecated}{%
+\section{Deprecated}\label{deprecated}}
+
+There are a few blocks at the bottom (figures with Google symptoms only
+and the old trajectory plots) that we can remove once we decide.
+
 \begin{figure}
-\centering
-\includegraphics[width=\textwidth]{example-image}
-\caption{First figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/fcast-finalized-1}
+
+}
+
+\caption{Forecasting performance using finalized data. Compare to Figure 3 in the manuscript.}\label{fig:fcast-finalized}
 \end{figure}
 
+\clearpage
+
 \begin{figure}
-\centering
-\includegraphics[width=\textwidth]{frog}
-\caption{Second figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/hot-finalized-1}
+
+}
+
+\caption{Hotspot prediction performance using finalized data. Compare to Figure 4 in the manuscript.}\label{fig:hot-finalized}
 \end{figure}
 
-\begin{table}\centering
-\caption{This is a table}
+\clearpage
 
-\begin{tabular}{lrrr}
-Species & CBS & CV & G3 \\
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/fcast-honest-v-finalized-1}
+
+}
+
+\caption{Relative forecast WIS with vintage compared to finalized data. Using finalized data leads to overly optimistic performance.}\label{fig:fcast-honest-v-finalized}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/hot-honest-v-finalized-1}
+
+}
+
+\caption{Relative AUC with vintage compared to finalized data. Using finalized data leads to overly optimistic hotspot performance.}\label{fig:hot-honest-v-finalized}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/wis-densities-1}
+
+}
+
+\caption{Weighted interval score appears to more closely resemble a log-Gaussian distribution.}\label{fig:wis-densities}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/fcast-adjusted-1}
+
+}
+
+\caption{Relative forecast performance using vintage data and summarizing with the more robust geometric mean.}\label{fig:fcast-adjusted}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/fcast-booted-1}
+
+}
+
+\caption{Forecast performance when indicators are replaced with samples from their empirical distribution. Performance is largely similar to the AR model.}\label{fig:fcast-booted}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/fcast-booted-adjusted-1}
+
+}
+
+\caption{Forecast performance as measured with the geometric mean when indicators are replaced with samples from their empirical distribution.
+Performance is largely similar to the AR model.}\label{fig:fcast-booted-adjusted}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/hot-booted-1}
+
+}
+
+\caption{Hotspot prediction performance when indicators are replaced with samples from their empirical distribution. Performance is largely similar to the AR model.}\label{fig:hot-booted}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/cor-wis-ratio-1}
+
+}
+
+\caption{This is one of the correlation plots Alden made.
+It shows histograms of the Spearman correlation between the ratio of AR to AR WIS with the percent change in 7dav cases relative to 7 days earlier.}\label{fig:cor-wis-ratio}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/cor-wis-ratio-m1-1}
+
+}
+
+\caption{This is Alden's second set of histograms.
+Here we have the correlation of the absolute value of WIS ratio - 1 with the percent change in 7dav cases relative to 7 days earlier}\label{fig:cor-wis-ratio-m1}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/upswing-histogram-logged-1}
+
+}
+
+\caption{Not sure if we want this here. Similar to Figure 5 in the manuscript but taking logs. }\label{fig:upswing-histogram-logged}
+\end{figure}
+
+\clearpage
+
+\begin{table}
+
+\caption{\label{tab:upswing-corr-table}Correlation of the difference in WIS between the AR model with the difference in median predictions.
+In down periods, improvements in forecast risk are highly correlated with lower median predictions.
+The opposite is true in up periods.
+This suggests, as one might expect that improved performance of the indicator-assisted model is attributable to being closer to the truth then the AR model.
+This conclusion is stronger in down periods then in up periods.}
+\centering
+\begin{tabular}[t]{lrrrrr}
+\toprule
+udf & CHNG-CLI & CHNG-COVID & CTIS-CLIIC & DV-CLI & Google-AA\\
 \midrule
-1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
-2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
-3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
+down & 0.79 & 0.80 & 0.83 & 0.81 & 0.81\\
+flat & 0.12 & 0.17 & 0.28 & 0.18 & 0.17\\
+up & -0.57 & -0.56 & -0.53 & -0.53 & -0.47\\
 \bottomrule
@@
 
-%%% Add this line AFTER all your figures and tables
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/hotspots-upswing-downswing-1}
+
+}
+
+\caption{Classification and loglikelihood separated into periods of upswing, downswing, and flat cases.
+Like the analysis of the forecasting task in the main paper (see Figure 7), performance is better during down and flat periods.}\label{fig:hotspots-upswing-downswing}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/fcast-alldates-1}
+
+}
+
+\caption{Forecast performance over all periods. Performance largely improves for all forecasters with the inclusion of data in 2021.}\label{fig:fcast-alldates}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/fcast-alldates-adjusted-1}
+
+}
+
+\caption{Forcast performance over all periods aggregaged with the geometric mean. Again, the inclusion of data in 2021 leads to improved performance.}\label{fig:fcast-alldates-adjusted}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/hot-alldates-1}
+
+}
+
+\caption{Area under the curve for hotspot predictions including data in 2021.
+Performance degrades relative to the period in 2020.
+However, there are far fewer hotspots during this period as case rates declined in much of the country.}\label{fig:hot-alldates}
+\end{figure}
+
+\clearpage
+
+\clearpage
+
 \FloatBarrier
@@
 \movie{Type legend for the movie here.}
-
-\movie{Type legend for the other movie here. Adding longer text to show what happens, to decide on alignment and/or indentations.}
-
-\movie{A third movie, just for kicks.}
 

### 2021-06-16 forecast/code/eval_funs.R

@@ -1,7 +1,3 @@
-library(tidyverse)
-library(lubridate)
-library(evalcast)
-
 Mean <- function(x) mean(x, na.rm = TRUE)
 Median <- function(x) median(x, na.rm = TRUE)
 TrimMean01 <- function(x) mean(x, trim = .01, na.rm = TRUE)
@@ -11,7 +7,7 @@ TrimMean1 <- function(x) mean(x, trim = .1, na.rm = TRUE)
 GeoMean <- function(x) exp(mean(log(x), na.rm = TRUE))
 
 fcast_colors <- c(RColorBrewer::brewer.pal(5, "Set1"), "#000000")
-names(fcast_colors) <- c("CHNG-CLI", "CHNG-COV", "CTIS-CLIIC", "DV-CLI",
+names(fcast_colors) <- c("CHNG-CLI", "CHNG-COVID", "CTIS-CLIIC", "DV-CLI",
                          "Google-AA", "AR")
 
 summarizer <- function(df, y, centerer, scaler, aggr, order_of_operations) {
@@ -98,7 +94,7 @@ process_res_hotspots <- function(res) {
     mutate(forecaster = recode(forecaster,
                                AR3 = "AR",
                                AR3CHCLI3 = "CHNG-CLI",
-                               AR3CHCOV3 = "CHNG-COV",
+                               AR3CHCOV3 = "CHNG-COVID",
                                AR3DV3 = "DV-CLI",
                                AR3FB3 = "CTIS-CLIIC",
                                AR3GG3 = "gs_subset",
@@ -127,7 +123,7 @@ process_res_cases <- function(res, actuals) {
     mutate(forecaster = recode(forecaster,
                                AR3 = "AR",
                                AR3CHCLI3 = "CHNG-CLI",
-                               AR3CHCOV3 = "CHNG-COV",
+                               AR3CHCOV3 = "CHNG-COVID",
                                AR3DVCLI3 = "DV-CLI",
                                AR3FBCLI3 = "CTIS-CLIIC",
                                AR3GSSAA3_Subset = "gs_subset",
@@ -197,4 +193,15 @@ filter_holidays <- function(res) {
   holiday_craziness = c(thanksgiving, christmas)
   res <- filter(res, !(target_end_date %in% holiday_craziness))
   res
+}
+
+
+pct_change <- function(x, n = 14, col_name = "pct_change"){
+  # courtesy of Alden
+  dt <- c(0,-n)
[…]

### 2021-06-16 forecast/code/revisions-dv-jhu.R
revisions figure to Charlotte

@@ -1,19 +1,22 @@
-# Known major case revision on Dec 29
-# HRR 303 is NYC
+# Large spike in cases on 9/25 due to inclusion of positive antigen tests as 
+# probable cases. Backfilled by JHU on 10/2 or there abouts.
+# See https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data#retrospective-reporting-of-probable-cases-and-deaths
+# 
+# HRR 311 is Charlotte NC
 
-as_ofs <- seq(as.Date("2020-12-3"), as.Date("2020-12-31"), by = "week")
+as_ofs <- seq(as.Date("2020-09-28"), as.Date("2020-10-19"), by = "week")
 cases_as_of <- map_dfr(as_ofs, function(as_of) {
   covidcast_signal(data_source = "jhu-csse", 
                    signal = "confirmed_incidence_prop",
-                   start_day = "2020-10-01", end_day = "2020-12-01", 
-                   geo_type = "hrr", geo_values = "303", as_of = as_of) %>%
+                   start_day = "2020-08-15", end_day = "2020-09-26", 
+                   geo_type = "hrr", geo_values = "311", as_of = as_of) %>%
     mutate(as_of = as_of)
 })
 dv_as_of <-  map_dfr(as_ofs, function(as_of) {
   covidcast_signal(data_source = "doctor-visits", 
                    signal = "smoothed_adj_cli",
-                   start_day = "2020-10-01", end_day = "2020-12-01", 
-                   geo_type = "hrr", geo_values = "303", as_of = as_of) %>%
+                   start_day = "2020-08-15", end_day = "2020-09-26", 
+                   geo_type = "hrr", geo_values = "311", as_of = as_of) %>%
     mutate(as_of = as_of)
 })
 

### 2021-06-17 forecast/paper/paper.tex
- Talk about Google latency - Add interesting vintage vs. finalized insights - Fix typos

@@
 of the case rate series itself. We will assess this in the most direct way
-possible: by inpsecting the difference in predictive performance of simple
+possible: by inspecting the difference in predictive performance of simple
 autoregressive models trained with and without access to past values of a
@@
 Because Google's team did not start publishing this signal until early
-September, 2020, we do not have true vintage data before then.  However, this
+September, 2020, we do not have true vintage data before then.
+Furthermore, the latency of the signal was always at least one week
+through 2020.  However, this
 signal is never revised after initial publication (confirmed via personal
-communication with the Google team that produces this signal), so we simply use
-finalized values for it in our analysis.
+communication with the Google team that produces this signal) and
+furthermore the latency of the signal is not an unavoidable property
+of the data type, so we simply use finalized values for it in our analysis.
 
@@
 Google-AA signal, which is only observed for an average of 105 (of 306) HRRs.
-This is due to the fact that that the COVID-19 search trends data is released
+This is due to the fact that the COVID-19 search trends data is released
 after applying differential privacy techniques \cite{Bavadekar:2020}, and a
@@
 optimistic estimates of true test-time performance (up to 10\% better in terms
-of average WIS or AUC).
+of average WIS or AUC). Furthermore, since indicators can have
+greatly different backfill profiles, the use of finalized data in
+retrospective evaluations can even change the relative ordering of
+methods.  For example, CHNG-CLI and DV-CLI
+when trained on finalized data perform very similarly for
+the forecasting task.  This makes sense since they are both claims-based indicators measuring supposedly the
+same thing.  However, DV-CLI outperforms CHNG-CLI on vintage data,
+reflecting its less severe backfill profile.
+
 % TODO note what we said here about the supplement
@@
 be found at
-\url{https://www.github.com/cmu-delphi/covidcast-pnas/forecast/code/}.
+\url{https://github.com/cmu-delphi/covidcast-pnas/tree/main/forecast/code}.
 
@@
 
-One might ask if the benefits we observe in forecasting and hotspot have
+One might ask if the benefits we observe in forecasting and hotspot
+prediction have
 anything to do with the actual indicator themselves; could it instead be that
@@
 
-Can axuiliary indicators improve COVID-19 forecasting and hotspot prediction
+Can auxiliary indicators improve COVID-19 forecasting and hotspot prediction
 models?  Our answer, based on analyzing five auxiliary indicators from the  
 COVIDcast API (defined using from medical insurance claims, internet-based
-surveys, and Google search trends) is undoubtedbly ``yes''. However, there are 
+surveys, and Google search trends) is undoubtedly ``yes''. However, there are 
 levels of nuance to such an answer that must be explained.  None of the

### 2021-06-17 forecast/paper/supplement.tex
Writing supplement sections
- Finalized vs. vintage
- Geometric mean
- Bootstrap
- 2021

@@
 
-This section referes to
-Figures\textasciitilde{}\ref{fig:fcast-finalized}--\ref{fig:hot-honest-v-finalized}.
+The goal of this section is to quantify the effect of
+not properly accounting for the question of ``what was known when'' in
+performing retrospective evaluations of forecasters.   Figures \ref{fig:fcast-finalized} and \ref{fig:hot-finalized} show
+what Figures 3 and 4 in the main paper would have looked like if we
+had simply trained all models using the finalized data rather than
+using vintage data.  This comparison can be seen most
+straightforwardly in Figures \ref{fig:fcast-honest-v-finalized} and \ref{fig:hot-honest-v-finalized}, which show the ratio in performance between the vintage and finalized
+versions.  When methods are
+given the finalized version of the data rather than the version available at the time that the forecast would have been made, all methods
+appear (misleadingly) to have better performance than they would
+actually have had if run prospectively.  For example, for forecasting case rates
+7-days ahead, the WIS of all methods is at least 8\% larger than what would have been
+recorded using the finalized values of the data.  This effect
+diminishes as the forecasting horizon increases, reflecting the fact
+that these forecasters rely less heavily on recent data than very
+short-term forecasters.  Crucially, some methods are
+``helped'' more than others by the less scrupulous retrospective
+evaluation, underscoring the difficulty of avoiding misleading
+conclusions when performing retrospective evaluations of forecasters.
+
+The \chngcli~indicator (along with the other claims-based signals) is the most
+affected by this distinction, reflecting the latency in claims-based
+reporting.  This supports the importance of efforts to provide ``nowcasts'' for claims signals
+(which corresponds to a 0-ahead ``forecast'' of what the claims
+signal's value will be once all data has been collected). Looking at
+the \chngcli~and \dv~curves in Figure \ref{fig:fcast-finalized}, we
+can see that they perform very similarly when trained on the finalized data.  This is reassuring
+because they are in principle measuring the same thing (namely, the percentage of
+  outpatient visits that are primarily about COVID-related symptoms).
+  The substantial difference in their curves in Figure 3 of the
+  main paper must therefore reflect their having very different
+  backfill profiles.
+
+While \dv~is one of the least affected methods by using finalized
+rather than vintage values, it is one of the most affected methods for the hotspot problem.
+This is a reminder that the forecasting and hotspot problems are
+distinct problems.  For example, the hotspot problem does not measure
+the ability to distinguish between flat and downward trends.
+
+Even the \ar~model is affected by this distinction, reflecting the
+fact that the case rates themselves (i.e., the response values) are also
+subject to revision.  The forecasters based on indicators are thus
+affected both by revisions to the indicators and by revisions to the
+case rates.  In the case of the \gs~model, in which we only used finalized
+values for the \gs~indicator, the difference in performance can be
+wholly attributed to revisions of case rates.
 
@@
 
-\begin{itemize}
-\tightlist
-\item
-  The weighted interval score is bounded below by zero and can be very
-  large. This behavior is typical of right-skewed distributions.
-\item
-  Figure\textasciitilde{}\ref{fig:wis-densities} illustrates that the
-  densities appear log-Gaussian.
-\item
-  This suggests aggregating by the geometric mean rather than the mean
-  for comparisons. See Figure\textasciitilde{}\ref{fig:fcast-adjusted}.
-\end{itemize}
+In this section, we consider using the geometric mean instead of the
+arithmetic mean when aggregating the weighted interval score (WIS) across location-time pairs.
+There are three reasons why the geometric mean may be desirable.
+\begin{enumerate}
+\item  WIS is right-skewed, being bounded below by zero and having
+  occasional very large values.   Figure\textasciitilde{}\ref{fig:wis-densities} illustrates that the
+  densities appear roughly log-Gaussian.  The geometric mean is a
+  natural choice in such a context since the relative ordering of forecasters is
+  determined by the arithmetic mean of the {\em logarithm} of their WIS
+  values.
+\item In the main paper, we report the ratio of the mean WIS of a
+  forecaster to the mean WIS of the baseline forecaster. Another
+  choice could be to take the mean of the ratio of WIS values for the
+  two methods. (This latter choice would penalize a method less for
+  doing poorly where the baseline forecaster also does poorly.)
+Using instead the geometric mean makes the order of aggregation and
+scaling immaterial since the ratio of geometric means is the same as
+the geometric mean of ratios.
+\item If one imagines that a forecaster's WIS is composed of
+  multiplicative space-time effects $S_{\ell,t}$ shared across all forecasters,
+  i.e. $\WIS(F_{\ell,t,f},Y_{\ell,t})=S_{\ell,t}E_{f}$, then
+  taking the ratio of two forecasters' geometric mean WIS values will
+  effectively cancel these space-time effects.
+\end{enumerate}
+
+Figure\textasciitilde{}\ref{fig:fcast-adjusted} uses the geometric
+mean for aggregation.  Comparing this with Figure 3 of the main paper,
+we see that the main conclusions are largely unchanged; however,
+\chngcli~now does appear better than \ar.  This behavior would be
+expected if \chngcli's poor performance is attributable to a
+relatively small number of large errors (as opposed to a large number
+of moderate errors).  Indeed, Figure 5 of the main paper further
+corroborates this, in which we see the heaviest left tails occuring
+for \chngcli.
 
 \hypertarget{bootstrap-results}{%
-\section{Bootstrap results}\label{bootstrap-results}}
+\section{Evaluating the implicit regularization hypothesis}\label{bootstrap-results}}
 
-Here we discuss how implicit regularization is not the reason for
-improved performance
+As explained in Section 2.B. of the main paper, a (somewhat cynical)
+hypothesis for why we see benefits in forecasting and hotspot
+prediction is that the indicators are not actually providing useful
+information but they are instead simply acting as a sort of ``implicit
+regularization,''  leading to shrinkage on the autoregressive
+coefficients, leading to less volatile predictions.  To investigate
+this hypothesis, we consider fitting  ``noise features'' that in truth
+should have zero coefficients.  Recall (from the main paper) that at each forecast date, we
+train a model on 6,426 location-time pairs.  Indicator models are
+based on six features, corresponding to the three autoregressive terms
+and the three lagged indicator values.  To form ``noise features''
+we resample from other rows.  In particular, at each location $\ell$ and
+time $t$, we replace the triplet $(X_{\ell,t}, X_{\ell,t-7},
+X_{\ell,t-14})$ by the triplet $(X_{\ell^*,t^*}, X_{\ell^*,t^*-7},
+X_{\ell^*,t^*-14})$, where $(\ell^*,t^*)$ is a location-time pair
+sampled with replacement from the 6,426 location-time pairs.  Figures
+\ref{fig:fcast-booted}--\ref{fig:hot-booted} show the results.  No
+method exhibits a noticeable performance gain over the \ar~method,
+leading us to dismiss the implicit regularization hypothesis.
+
+\attn{Would be good for someone to double check that the bootstrap
+  procedure is correctly described.}
 
@@
 
-See Figures \ref{fig:fcast-alldates} -- \ref{fig:hot-alldates}.
+In this section, we investigate the sensitivity of the results to the
+period over which we train and evaluate the models.  In the main
+paper, we end all evaluation on December 31, 2020.  Figures
+\ref{fig:fcast-alldates} -- \ref{fig:hot-alldates} show how the
+results would differ if we extended this analysis through \attn{XX,
+  XX}, 2021. Comparing Figure \ref{fig:fcast-alldates} to Figure 3 of
+the main paper, one sees that as ahead increases most methods now improve
+relative to the baseline forecaster. When compared to other methods, \chngcli~appears much better than
+it had previously; however, all forecasters other than \chngcov~and
+\dv~are performing less well relative to the baseline than before.
+These changes are likely due to the differing nature of the pandemic
+in 2021, with flat and downward trends much more common than upward
+trajectories.  Indeed, the nature of the hotspot prediction problem is
+quite different in this period.  With a 21-day training window, it is
+common for there to be very few hotspots in training.
 

### 2021-06-17 forecast/paper/paper.tex
Temp commit so I can pull

--- 2021-06-16 forecast/paper/paper.tex
@@
   \caption{Revision behavior for signals in the New York City HRR from two
-  sources. Each colored line 
-  corresponds to the data as reported on that date. DV-CLI is revised upward
-  throughout the month of December, though the amount fades as we look further
-  back in time. In contrast, JHU-CSSE data remains ``as reported'' on
-  December 3 until a major correction is made on December 31, affecting all data
-  over the period. }
+    sources. Each colored line corresponds to the data as reported on that
+    date. DV-CLI is revised upward throughout the month of December, though the
+    amount fades as we look further back in time. In contrast, JHU-CSSE data
+    remains ``as reported'' on December 3 until a major correction is made on
+    December 31, affecting all data over the period. }
   \label{fig:vintage}
@@
 problem, rendering it easy when it was once hard (in the absence of auxiliary 
-information).  Rather, the gains in accuracy offered by the indicator models 
-(over an autoregressive model based on past case rates along) appear to be
-nontrivial, and consistent across modes of analysis, but modest.  In
-forecasting, the indicator models appear to be most useful in periods in which 
-case rates are flat or trending down, rather than periods in which case rates
-are trending up (as one might hope to see, as the hypothetical benefit provided
-by an idealized ``leading indicator'').
+information).  Rather, the gains in accuracy from the indicator models (over an
+autoregressive model based only on past case rates) appear to be nontrivial, and
+consistent across modes of analysis, but modest.  In forecasting, the indicator 
+models appear to be most useful in periods in which case rates are flat or
+trending down, rather than periods in which case rates are trending up (as one 
+might hope to see is the benefit provided by a hypothetical ``leading
+indicator'').
 
-As mentioned previously, the indicator models could certainly benefit from a 
-careful handling of location-specific effects (to account for spatial
-heterogeneity in the search trends and claims-based indicators in particular),
-as well as nowcasting techniques to estimate finalized indicator values before
-using them as features (to account for backfill in the claims-based indicators
-in particular).  Beyond this, it is certainly possible that more sophisticated 
-models for forecasting or hotspot prediction would lead to different results,
-and potentially even different insights.  Natural directions to explore include  
-using interactions, using multiple indicators in tandem in a single model, and 
-leveraging HRR demographics or mobility patterns.  That said, we are doubtful
-that the ``topline'' conclusion would change--that auxiliary indicators can
-provide a nontrivial, consistent, but modest improvement in forecasting and
-hotspot prediction.
-
-Furthermore, from our experience working on forecasting  TODO
+As mentioned previously, the indicator models could likely be improved by using 
+location-specific effects, as well as using nowcasting techniques to estimate
+finalized indicator values before using them as features (in order to account
+for backfill in the claims-based signals in particular).  Beyond this, it is
+certainly possible that more sophisticated models for forecasting or hotspot
+prediction would lead to different results, and even different
+insights. Natural directions to explore include using interactions, using
+multiple indicators in a single model, and leveraging HRR demographics or
+mobility patterns.  That said, we are doubtful that more sophisticated modeling
+techniques would change the ``topline'' conclusion---that auxiliary indicators 
+can provide nontrivial, consistent, but modest gains in forecasting and
+hotspot prediction. Whether a more sophisticated model would be able to leverage
+the indicators in such a way as to change some of the finer conclusions---e.g.,
+by offering clear improvements in periods in which cases are trending up---is 
+less clear to us.
 

### 2021-06-17 forecast/paper/unused.tex

@@
-%%%%%
-
-Using instead the geometric mean makes the order of aggregation and
-scaling immaterial since the ratio of geometric means is the same as
-the geometric mean of ratios.  Figure \ref{fig:geometric-mean} shows how the
-results in forecasting would change if one replaces the arithmetic mean by the
-geometric mean.
-
 %%%%%
@@
 %%%%%
-
-Figure \ref{fig:finalized-vs-honest} quantifies the effect of
-not properly accounting for the question of ``what was known when'' in
-performing retrospective evaluations of forecasters.  When methods are
-given the finalized version of the data rather than the version available at the time that the forecast would have been made, all methods
-appear (falsely) to have better performance.  For example, for forecasting case rates
-7-days ahead, the WIS of all methods is at least 8\% larger than what would have been
-recorded using the finalized values of the data.  This effect
-diminishes as the forecasting horizon increases, reflecting the fact
-that these forecasters rely less heavily on recent data than very
-short-term forecasters.  Crucially, some methods are
-``helped'' more than others by the less scrupulous retrospective
-evaluation, underscoring the difficulty of avoiding misleading
-conclusions when performing retrospective evaluations of forecasters.
-
-The \chngcli~indicator (along with the other claims-based signals) is the most
-affected by this distinction, reflecting the latency in claims-based
-reporting. \attn{Actually, \dv~is the least affected for forecasting
-  whereas it is one of the most affected for hotspots.}  This supports
-the importance of efforts to provide ``nowcasts'' for claims signals
-(which corresponds to a 0-ahead ``forecast'' of what the claims
-signal's value will be once all data has been collected).
-
-Even the \ar~model is affected by this distinction, reflecting the
-fact that the case rates themselves (i.e., the response values) are also
-subject to revision.  The forecasters based on indicators are thus
-affected both by revisions to the indicators and by revisions to the
-case rates.
-
-\begin{itemize}
-%\item Motivates the importance of nowcasting for claims signals.
-  \item For forecasting task: Also, \chngcli~and \dv~perform very similarly here, which is
-    reassuring because they're measuring the same thing (in
-    principle).  Yet they must have very different backfill profiles…
-  \item For hotspots task: dv seems to get a bigger boost than
-    \chngcov - interesting dichotomy to forecasting
-\item But in both forecasting/hotspots \chngcli~is affected a lot by backfill
-\item \attn{We should be careful about which indicators we were not
-      able to track latency for.  In fact, we should probably remove
-      these from the plot since they would be more strongly affected
-      than is shown in the plot.}
-
-\item \attn{We can also discuss the distinction
-      between indicators that could in principle have low latency
-      (e.g. GS) versus those where latency is inherent
-      (e.g. claims-based indicators). If the goal is assessing the
-      performance of GS ``for the next pandemic'' then it would make
-      sense to give GS a pass since in theory they could have had it
-      with very low latency (as opposed to claims based signals where
-      backfill may be unavoidable).  Of course, in
-    ``the next pandemic'' there might not be such highly specific
-    symptoms such as A+A as there was in the COVID-19 case.} 
-\end{itemize}
-
-\begin{figure}[t]
-     \centering
-     \includegraphics[width=\textwidth]{fig/fcast-honest-v-finalized-1.pdf}
-     \includegraphics[width=\textwidth]{fig/hot-honest-v-finalized-1.pdf}
-     \caption{How misleading is a retrospective analysis that uses the
-       finalized version of the data rather than the version that was
-       in fact available at the time the forecast was made?  Plots
-       show the ratio in performance (using mean WIS fore the
-       forecasting task and AUC for the hotspot prediction
-       task). \label{fig:finalized-vs-honest} Cheating vs. honest plots.} 
-\end{figure}

### 2021-06-17 indicators/paper/paper.tex
Attemt to draft a complete abstract

@@
   inconsistent reporting and frequent, unexpected revisions. They also miss
-  other aspects of a population's behavior, worthy of consideration. We
+  other aspects of a population's behavior that are worthy of consideration. We
   present an open database of COVID signals in the United States, measured at
@@
   cases and deaths, and many others: signals based on measures of mobility,
-  social distancing, internet search trends, self-reported symptoms, and
+  social distancing, Internet search trends, self-reported symptoms, and
   patterns COVID-related activity in de-identified medical insurance claims. The
@@
 
-\begin{abstract}
-  Please provide an abstract of no more than 250 words in a single paragraph.
-  Abstracts should explain to the general reader the major contributions of the
-  article. References in the abstract must be cited in full within the abstract
-  itself and cited in the text.
+\begin{abstract} % 250 words max, one paragraph
+  The COVID-19 pandemic in the United States presented enormous data challenges.
+  Policymakers, epidemiological modelers, and health researchers all require
+  up-to-date data on the pandemic and relevant public behavior, ideally at fine
+  spatial and temporal resolution. We present the COVIDcast API, which provides
+  an open database of both traditional public health surveillance signals (such
+  as cases, deaths, and hospitalizations) and unique indicators of COVID-19
+  activity, such as signals extracted from de-identified medical claims data,
+  massive online surveys, cell phone mobility data, and Internet search trends.
+  These signals are available at a fine geographic resolution (mostly at the
+  county level) and are updated daily. Unlike most public health databases, the
+  COVIDcast API also tracks all revisions to historical data, allowing modelers
+  to account for the frequent revisions and backfill that are common for many
+  public health data sources. All data is available in a common format through
+  an API and software packages. We describe the data and formats, and present
+  examples demonstrating that the non-traditional data sources present in the
+  COVIDcast API present information relevant to tracking COVID activity,
+  augmenting standard public health reporting and enabling research.
 \end{abstract}

### 2021-06-17 indicators/paper/paper.tex
Temp commit so I can pull

--- 2021-06-16 indicators/paper/paper.tex
@@
 COVID-19 case and death incidence, at the state and county levels, to the
-COVID-19 Forecast Hub \cite{ForecastHub}, under the name ``CMU-TimeSeries''.  
-The process of building, training, and deploying our forecasting models
-leverages much of the infrastructure described in this paper (such as the
-COVIDcast API's \textit{as of} feature), and some of our forecasting systems
-rely on auxiliary indicators (such as survey-based and claims-based COVID-like 
-illness signals, described below, which we have used in case forecasting).
+COVID-19 Forecast Hub \cite{ForecastHub}, with ``CMU-TimeSeries'' as the
+team-model name.  The process of building, training, and deploying our
+forecasting models leverages much of the infrastructure described in this paper
+(such as the COVIDcast API's \textit{as of} feature), and some of our
+forecasting systems rely on auxiliary indicators (such as survey-based and
+claims-based COVID-like illness signals, which are described below).
 

### 2021-06-17 forecast/paper/paper.tex
fix typo: max ahead is 28

--- 2021-06-17 forecast/paper/paper.tex
@@
 incidence rates in location $\ell$ and at time $t$.  Separately for each
-$a=7,\ldots,28$, to predict $Y_{\ell,t+a}$ for ahead value $a$, we consider a
+$a=7,\ldots,21$, to predict $Y_{\ell,t+a}$ for ahead value $a$, we consider a
 simple probabilistic forecasting model of the form:     
@@
 altogether quite similar to the forecasting setup described previously.
-Separately for each $a=7,\ldots,28$, to predict $Z_{\ell,t+a}$, we consider a
+Separately for each $a=7,\ldots,21$, to predict $Z_{\ell,t+a}$, we consider a
 simple logistic model:

### 2021-06-17 indicators/paper/paper.tex
Copyedits and plot tweaks

--- 2021-06-17 indicators/paper/paper.tex
@@
   inconsistent reporting and frequent, unexpected revisions. They also miss
-  other aspects of a population's behavior, worthy of consideration. We
+  other aspects of a population's behavior that are worthy of consideration. We
   present an open database of COVID signals in the United States, measured at
-  the county level and updated daily. These include traditionally reported COVID
-  cases and deaths, and many others: signals based on measures of mobility,
-  social distancing, internet search trends, self-reported symptoms, and
-  patterns COVID-related activity in de-identified medical insurance claims. The
-  database provides all signals in a common, easy-to-use format, empowering both
-  public health research and operational decision-making.
+  the county level and updated daily. This includes traditionally reported COVID
+  cases and deaths, and many others: measures of mobility, social distancing,
+  Internet search trends, self-reported symptoms, and patterns COVID-related
+  activity in de-identified medical insurance claims. The database provides all
+  signals in a common, easy-to-use format, empowering both public health
+  research and operational decision-making.  
 }
@@
 \begin{abstract} 
-  Please provide an abstract of no more than 250 words in a single paragraph.
-  Abstracts should explain to the general reader the major contributions of the
-  article. References in the abstract must be cited in full within the abstract
-  itself and cited in the text.
+  The COVID-19 pandemic presented enormous data challenges in the United States.
+  Policy makers, epidemiological modelers, and health researchers all require
+  up-to-date data on the pandemic and relevant public behavior, ideally at fine
+  spatial and temporal resolution. The COVIDcast API is our attempt to fill this
+  need: operational since April 2020, it provides open access to both
+  traditional public health surveillance signals (cases, deaths, and
+  hospitalizations) and many auxiliary indicators of COVID-19 activity, such as
+  signals extracted from de-identified medical claims data, massive online
+  surveys, cell phone mobility data, and Internet search trends. These 
+  are available at a fine geographic resolution (mostly at the county level) and
+  are updated daily. The COVIDcast API also tracks all revisions to historical
+  data, allowing modelers to account for the frequent revisions and backfill
+  that are common for many public health data sources. All of the data is
+  available in a common format through the API and accompanying R and Python
+  software packages. This paper describes the data sources and signals, and
+  provides examples demonstrating that the auxiliary signals in the COVIDcast
+  API present information relevant to tracking COVID activity, augmenting
+  traditional public health reporting and empowering research and
+  decision-making.         
 \end{abstract}
@@
 the CDC's ILINet \cite{Santillana:2015, Farrow:2016}, as do publication formats
-providing access to historical versions of a data set \cite{Brooks:2018,
-  Brooks:2020}. Numerous other examples exist spanning a wide variety of data
+providing access to historical versions of a given data set \cite{Brooks:2018, 
+  Brooks:2020}. Several other examples exist that span a wide variety of data  
 platforms and diseases \cite{Browstein:2009, Kass-Hout:2011, Salathe:2012,

### 2021-06-17 indicators/paper/paper.tex
Discussion copyedits

@@
 Third, the underlying database tracks all revisions made to the data, allowing
-us to query the API to learn ``what was known when'', critical for understanding
-the behavior (and potential pitfalls) of real-time surveillance signals. Such
-revision data is rarely available in standardized format from other sources.
+us to query the API to learn ``what was known when'', which is critical for
+understanding the behavior (and potential pitfalls) of real-time surveillance
+signals. Such revision data is rarely available in standardized format from
+other sources.
 

### 2021-06-17 forecast/paper/paper.tex
fix typo: beta missing in equation
Discussion copyedits

--- 2021-06-17 forecast/paper/paper.tex
@@
 \logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s}, s \leq t) \big) 
-= \alpha^{a,\tau} + \sum_{j=0}^2 \^{a,\tau}_j Y^\Delta_{\ell,t-7j},
+= \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j},
 \end{equation}

### 2021-06-17 forecast/paper/paper.tex
Fix link, move figs

--- 2021-06-17 forecast/paper/paper.tex
@@
 Because Google's team did not start publishing this signal until early
-September, 2020, we do not have true vintage data before then.
-Furthermore, the latency of the signal was always at least one week
-through 2020.  However, this
+September, 2020, we do not have true vintage data before then.  Furthermore, the
+latency of the signal was always at least one week through 2020.  However, this
 signal is never revised after initial publication (confirmed via personal
-communication with the Google team that produces this signal) and
-furthermore the latency of the signal is not an unavoidable property
-of the data type, so we simply use finalized values for it in our analysis.
+communication with the Google team that produces this signal) and furthermore
+the latency of the signal is not an unavoidable property of the data type, so we
+simply use finalized signal values it, with zero latency, in our analysis.
 
@@
 Google-AA signal, which is only observed for an average of 105 (of 306) HRRs.
-This is due to the fact that the COVID-19 search trends data is released
-after applying differential privacy techniques \cite{Bavadekar:2020}, and a
+Such missingness occurs because the COVID-19 search trends data is
+constructed using differential privacy methods \cite{Bavadekar:2020}, and a 
 missing signal value means that the level of noise added in the differential
@@
 optimistic estimates of true test-time performance (up to 10\% better in terms
-of average WIS or AUC). Furthermore, since indicators can have
-greatly different backfill profiles, the use of finalized data in
-retrospective evaluations can even change the relative ordering of
-methods.  For example, CHNG-CLI and DV-CLI
-when trained on finalized data perform very similarly for
-the forecasting task.  This makes sense since they are both claims-based indicators measuring supposedly the
-same thing.  However, DV-CLI outperforms CHNG-CLI on vintage data,
-reflecting its less severe backfill profile.
-
+of average WIS or AUC). Furthermore, since indicators can have greatly different
+backfill profiles, the use of finalized data in retrospective evaluations can
+even change the relative ordering of models.  For example, CHNG-CLI and DV-CLI
+when trained on finalized data perform very similarly for the forecasting task.
+This makes sense since they are both claims-based indicators measuring
+supposedly the same thing.  However, DV-CLI outperforms CHNG-CLI on vintage
+data, reflecting its has a less severe backfill profile.
 % TODO note what we said here about the supplement
@@
 
+Reiterate importance of AS OF
+
+Not the end of the story.  Should we even be forecasting case rates?  More
+careful models---combine hotspots and forecasting?
+
 [  Finish by reiterating the importance of vintage data.  Emphasizing the rigor

### 2021-06-17 indicators/paper/paper.tex
Fix link, move figs

@@
 
-In what follows, we present examples of the usefulness of some of the novel
-signals available in the API. These examples demonstrate that such indicators
-are meaningfully related to COVID activity, that they provide alternate views on
-pandemic activity that are not subject to the same reporting glitches and delays
-as traditional public health surveillance streams, and that they provide
-information about public behavior and attitudes that are not available from any
-other source. Code to reproduce all examples (which uses the \texttt{covidcast}
-R package and fetches data directly from the API) can be found at
-\url{https://www.github.com/cmu-delphi/covidcast-pnas/indicators/code/}.
-
-\subsection{Tracking Trends}
-
 \begin{figure}[t]
@@
 \end{figure}
+
+In what follows, we present examples of the usefulness of some of the novel
+signals available in the API. These examples demonstrate that such indicators
+are meaningfully related to COVID activity, that they provide alternate views on
+pandemic activity that are not subject to the same reporting glitches and delays
+as traditional public health surveillance streams, and that they provide
+information about public behavior and attitudes that are not available from any
+other source. Code to reproduce all examples (which uses the \texttt{covidcast}
+R package and fetches data from the API) can be found at
+\url{https://github.com/cmu-delphi/covidcast-pnas/tree/main/indicators/code/}.
+
+\subsection{Tracking Trends}
 
@@
 
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{fig/geo_wise_corr.pdf}
+  \caption{Geo-wise correlations with case rates, from April 15, 2020 to April 
+    15, 2021, calculated over all counties for which all signals were available
+    and which had at least 500 cumulative cases by the end of this period.}
+  \label{fig:geo_wise_correlation} 
+\end{figure}
+
 Figure~\ref{fig:time_trends_national} compares the first three of these signals
@@
 
-% TODO
-% RJT: @AR Note what I claimed we would do above!  I'm thinking we can just
-% produce a big batch of case trends over the same time period and for the same
-% signals, but for each US state (and territory?), and for the 50 most populous
-% counties. AND maybe for this (below national) we should use CLI-in-community
-% rather than CLI, from the survey
-
 Besides tracking contemporaneous COVID activity, these and other indicators can
@@
 
-% TODO 
-% RJT: @AR Consider for the supplement.  [Or skip it?  Not all that important 
-% but I think it would be nice to do.  Maybe you can delegate this to Logan,
-% Aaron, Jingjing, etc.] 
-% - Hospitalization trends?
-% - Hospitalization correlations?
-%  (I see you already have a plot for the first ... but I'm a bit bothered by
-%  the fact that the HHS data looks to have such strong weekday effects that we
-%  don't attempt to correct for ... perhaps we use do a 7-day rolling avg. Made
-%  easy with modeltools)
-% If we do so, we should refer to it in the text, this subsection (for time
-% trends) and next (for correlations)
-
 \subsection{Correlation Analyses}
-
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/geo_wise_corr.pdf}
-  \caption{Geo-wise correlations with case rates, from April 15, 2020 to April 
-    15, 2021, calculated over all counties for which all signals were available
-    and which had at least 500 cumulative cases by the end of this period.}
-  \label{fig:geo_wise_correlation} 
-\end{figure}
 
@@
   Figure~\ref{fig:geo_wise_correlation}, where the CTIS-CLI-in-community signal
-  achieves clearly the highest correlations for most of the time
-  period. However, it is worth emphasizing that time-wise and geo-wise
+  achieves clearly the highest correlations for most of the time period. 
+  However, it is worth emphasizing that time-wise and geo-wise 
   correlations are truly measuring different properties of a signal; and the

### 2021-06-17 indicators/paper/paper.tex
Add first group of authors

@@
 
-% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[a,c,1]{Author One}
-\author[b,1,2]{Author Two}
-\author[a]{Author Three}
+% Use letters for affiliations, numbers to show equal authorship (if applicable)
+% and to indicate the corresponding author
 
-\affil[a]{Affiliation One}
-\affil[b]{Affiliation Two}
-\affil[c]{Affiliation Three}
+% Writing group
+\author[a,1]{Alex Reinhart}
+\author[b,2]{Logan Brooks}
+\author[a,2]{Maria Jahja}
+\author[b,2]{Aaron Rumack}
+\author[c,2]{Jingjing Tang}
+
+% Middle authors
+\author[d]{Middle Authors to be Added}
+
+% Senior authors
+\author[b]{Roni Rosenfeld}
+\author[a]{Ryan Tibshirani}
+
+\affil[a]{Department of Statistics \& Data Science, Carnegie Mellon University}
+\affil[b]{Machine Learning Department, Carnegie Mellon University}
+\affil[c]{Department of Computational Biology, Carnegie Mellon University}
+\affil[d]{Remaining affiliations to be added}
 
 % Please give the surname of the lead author for the running footer
-\leadauthor{Lead author last name}
+\leadauthor{Reinhart}
 
@@
 \authordeclaration{Please declare any competing interests here.}
-\equalauthors{\textsuperscript{1}A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable).}
-\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: author.two\@email.com}
+\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: areinhar@stat.cmu.edu}
+\equalauthors{\textsuperscript{2}L.B., M.J., A.R., and J.T. contributed equally
+  to this work.}
 

### 2021-06-17 forecast/paper/paper.tex
drafted udf and lead/lag sections of the manuscript

--- 2021-06-17 forecast/paper/paper.tex
@@
 \subsection{Signals and Locations}
+\label{sec:signals-locations}
 
@@
 \caption{Revision behavior for signals in the New York City HRR from two
-    sources. Each colored line corresponds to the data as reported on that
-    date. DV-CLI is revised upward throughout the month of December, though the
-    amount fades as we look further back in time. In contrast, JHU-CSSE data
-    remains ``as reported'' on December 3 until a major correction is made on
-    December 31, affecting all data over the period. }
+  sources. Each colored line 
+  corresponds to the data as reported on that date. DV-CLI is revised upward
+  throughout the month of December, though the amount fades as we look further
+  back in time. In contrast, JHU-CSSE data remains ``as reported'' on
+  December 3 until a major correction is made on December 31, affecting all data
+  over the period. }
 \label{fig:vintage}
@@
 incidence rates in location $\ell$ and at time $t$.  Separately for each
-$a=7,\ldots,21$, to predict $Y_{\ell,t+a}$ for ahead value $a$, we consider a
+$a=7,\ldots,28$, to predict $Y_{\ell,t+a}$ for ahead value $a$, we consider a
 simple probabilistic forecasting model of the form:     
@@
 altogether quite similar to the forecasting setup described previously.
-Separately for each $a=7,\ldots,21$, to predict $Z_{\ell,t+a}$, we consider a
+Separately for each $a=7,\ldots,28$, to predict $Z_{\ell,t+a}$, we consider a
 simple logistic model:
@@
 \logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s}, s \leq t) \big) 
-= \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j},
+= \alpha^{a,\tau} + \sum_{j=0}^2 \^{a,\tau}_j Y^\Delta_{\ell,t-7j},
 \end{equation}
@@
 
-\subsection{Performance in Up, Down, Flat Periods}
+\subsection{Performance in Up, Down, and Flat Periods}
+
+The course of the pandemic has played out very differently across space and
+time. Aggregating case rates nationally shows three pronounced waves, but the
+behavior is more nuanced at the HRR level. 
+Figure \ref{fig:trajectory} is a single example of what a forecast looks like
+during a period of relatively flat cases as New York City begins what will
+become its second wave. The AR forecast's 50\% confidence interval contains this
+upswing, but the median forecast is clearly below the finalized data.
+Unfortunately, this behavior is typical of all forecasters: during upswings, the
+median forecast tends to fall below the target while the reverse is true during
+downswings.
+
+Figure \ref{fig:up-down-flat} shows histograms of the difference between the WIS for the
+forecasters with the benefit of the indicators and the WIS of the AR model, but
+it stratifies these differences by
+whether the target occurs during a period of increasing cases rates (up),
+decreasing case rates (down), or flat. To define the increasing period, we use
+the same definition we used for the hotspot task in Table
+\ref{tab:analysis_tasks}. Thus all hotspots are ``up'' while all non-hotspots
+are either flat or decreasing. For the ``down'' scenario, we simply use the opposite of
+the hotspot definition: $Y_{\ell,t}$ decreases by more than 25\% relative to the
+preceding week.
+
+While the performance of all forecast degrades during up periods (WIS is on the
+scale of cases, so higher case counts means higher WIS) there are pronounced
+patterns for the different models. 
+Google-AA is the only model that
+significantly outperforms the AR model during upswings. Both Google-AA and
+CTIS-CLI-in-community have large right tails during the flat periods while the
+claims signals have large left tails (indicating some large, relatively poor
+forecasts) during flat and up periods. CHNG-CLI, CHNG-COVID, Google-AA and
+especially CTIS-CLI-in-community have large positive outliers during the down
+period. Overall, the indicators seem to help more during periods of flat or down
+cases than during up periods, with the exception of Google-AA. In the
+supplement, we examine classification accuracy and log-likelihood for the
+hotspot task. We see a similar phenomenon: the indicators dramatically improve
+hotspot prediction accuracy during flat and down periods with more mixed
+behavior during periods of increasing cases, with CHNG-CLI, CHNG-COVID, and DV-CLI
+leading to decreased performance.
 
@@
   \caption{Histogram of the difference in WIS for the AR model relative to each
-    forecaster. Larger values are better for the indicator assisted model.}
+    forecaster. Larger values are better for the indicator assisted model. The
+    $y$-axis is on the logarithmic scale to 
+    emphasize outliers.}
   \label{fig:up-down-flat}
@@
 \subsection{Effects of Leading or Lagging Behavior}
+
+As described in Section \ref{sec:signals-locations}, each of the indicators we
+examine could be said to measure aspects on disease progression that would
+precede a positive test. That is, we imagine that these signals should ``lead''
+cases. It is entirely reasonable to imagine that, prior to an increase of
+confirmed COVID-19 tests reported by public health in a particular location, we
+would see an increase in insurance claims for COVID-19-related medical visits.
+However, it may well be the case that this leading behavior is different during
+different periods. In fact, we find that the ``leadingness'' of the indicators
+is more pronounced during downswings and flat periods than during upswings, a
+plausible explanation for the decreased performance noted above.
+
+In order to measure the leadingness on an indicator, we use the cross
+correlation function. Essentially this examines the Pearson correlation between a
+shifted version of the indicator and the case rate. So if $X_{\ell,t}$ is
+shifted forward by 7 days, and the magnitude of the correlation of that shifted series with
+$Y_{\ell,t}$ (unshifted) is high, we would say that $X_{\ell}$ leads $Y_{\ell}$
+by 7 days. This process is repeated for a range of shifts, forward and backward,
+resulting in leadingness and laggingness scores. Further details are given in
+the supplement.
+
+Figure \ref{fig:leading-lagging} shows the correlations between the difference
+in the WIS of the AR model relative to $F$ with the leading and laggingness
+scores stratified by whether the target is classified as up, down, or flat. One
+would expect that increases in $\WIS(AR) - \WIS(AR)$ would be positively
+correlated with leadingness. This relationship is strongest during down periods
+and weakest during up periods with flat in the middle. In fact, for each
+indicator, the strength of correlations between improved performance and
+leadingness can be ordered down $>$ flat $>$ up. The relationship with
+laggingness is the opposite, though less pronounced. CHNG-CLI's forecast
+performance doesn't correlate very strongly with laggingness while any
+laggingness in Google-AA during down and flat periods results in realtively
+worse forecasting performance.
+
+Because a time series can, to a degree, both simultaneously leading and lagging,
+we also examine the correlation between performance and the difference between
+leadingness and laggingness. Figure \ref{fig:diff-btw-leading-lagging} shows
+this behavior. Again, indicators that are more leading then lagging tend to
+are positively correlated with improved relative performance. This relationship
+is most pronounced during periods of decreasing case rates and attenuates in
+other scenarios.
 
@@
 problem, rendering it easy when it was once hard (in the absence of auxiliary 
-information).  Rather, the gains in accuracy from the indicator models (over an
-autoregressive model based only on past case rates) appear to be nontrivial, and
-consistent across modes of analysis, but modest.  In forecasting, the indicator 
-models appear to be most useful in periods in which case rates are flat or
-trending down, rather than periods in which case rates are trending up (as one 
-might hope to see is the benefit provided by a hypothetical ``leading
-indicator'').
+information).  Rather, the gains in accuracy offered by the indicator models 
+(over an autoregressive model based on past case rates along) appear to be
+nontrivial, and consistent across modes of analysis, but modest.  In
+forecasting, the indicator models appear to be most useful in periods in which 
+case rates are flat or trending down, rather than periods in which case rates
+are trending up (as one might hope to see, as the hypothetical benefit provided
+by an idealized ``leading indicator'').
 
-As mentioned previously, the indicator models could likely be improved by using 
-location-specific effects, as well as using nowcasting techniques to estimate
-finalized indicator values before using them as features (in order to account
-for backfill in the claims-based signals in particular).  Beyond this, it is
-certainly possible that more sophisticated models for forecasting or hotspot
-prediction would lead to different results, and even different
-insights. Natural directions to explore include using interactions, using
-multiple indicators in a single model, and leveraging HRR demographics or
-mobility patterns.  That said, we are doubtful that more sophisticated modeling
-techniques would change the ``topline'' conclusion---that auxiliary indicators 
-can provide nontrivial, consistent, but modest gains in forecasting and
-hotspot prediction. Whether a more sophisticated model would be able to leverage
-the indicators in such a way as to change some of the finer conclusions---e.g.,
-by offering clear improvements in periods in which cases are trending up---is 
-less clear to us.
+As mentioned previously, the indicator models could certainly benefit from a 
+careful handling of location-specific effects (to account for spatial
+heterogeneity in the search trends and claims-based indicators in particular),
+as well as nowcasting techniques to estimate finalized indicator values before
+using them as features (to account for backfill in the claims-based indicators
+in particular).  Beyond this, it is certainly possible that more sophisticated 
+models for forecasting or hotspot prediction would lead to different results,
+and potentially even different insights.  Natural directions to explore include  
+using interactions, using multiple indicators in tandem in a single model, and 
+leveraging HRR demographics or mobility patterns.  That said, we are doubtful
+that the ``topline'' conclusion would change--that auxiliary indicators can
+provide a nontrivial, consistent, but modest improvement in forecasting and
+hotspot prediction.
+
+Furthermore, from our experience working on forecasting  TODO
 

### 2021-06-17 indicators/paper/supplement.Rmd

---
bibliography: [../../common/covidcast.bib, pnas-materials/pnas-sample.bib]
output:
  pdf_document:
    keep_tex: true
    template: pnas-suppl-template.tex
params:
  fd_casefloor: 30
  flag_jumps: 15
---

<!-- PNAS materials -->

<!-- Comment out or remove this line before generating final copy for submission; this will also remove the warning re: "Consecutive odd pages found". -->
\instructionspage

<!-- we need this -->
\maketitle

<!-- Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text. -->

\SItext

All our text goes here. Reference figures with the R chuck label as Figure~\ref{fig:fcast-finalized}.

All figures go on their own page after all of the text...

Edit `pnas-suppl-template.tex` for the correct Author list and Title.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 6.5, fig.height = 4.5,
                      fig.align = "center",
                      fig.path = "fig/",
                      cache = TRUE,
                      out.width = "\\textwidth",
                      autodep = TRUE,
                      message = FALSE,
                      warning = FALSE)

library(covidcast)
library(gridExtra)
library(ggplot2)
library(ggforce) # for splitting across pages
library(scales)
library(dplyr)
```

\begin{figure}
\includegraphics[width=\textwidth]{fig/severity-pyramid.pdf}
\caption{The epidemiological "severity pyramid" represents the progression of
cases from the public, through infection, through increasingly severe stages of
disease. The annotations here represent the data sources collected by Delphi's
COVIDcast Epidata API.}
\end{figure}

\clearpage

```{r state-trends}
SOURCES <- c("jhu-csse", "chng", "chng", "fb-survey")
SIGNALS <- c("confirmed_7dav_incidence_prop",
             "smoothed_adj_outpatient_cli",
             "smoothed_adj_outpatient_covid",
             "smoothed_whh_cmnty_cli")
colors <- c("black", scales::hue_pal()(5))
names <- c("Cases", "CHNG-CLI", "CHNG-COVID", "CTIS-CLI-in-community")

# geo_labeler is a function from geo_value to human-readable name
get_trend_grid <- function(trends, geo_labeler) {
  case_max <- max(trends[[1]]$value)

  trends <- aggregate_signals(trends, format = "long") %>%
    group_by(signal) %>%
    mutate(value = value / max(value, na.rm = TRUE) * case_max) %>%
    ungroup() %>%
    mutate(signal = factor(signal, levels = SIGNALS, labels = names),
           geo_value = geo_labeler(geo_value))

  g <- ggplot(trends, aes(x = time_value, y = value, color = signal)) +
    geom_line() +
    scale_color_manual(breaks = names, values = colors) +
    labs(x = "Date", y = "Signal value (scaled)") +
    scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
                 date_minor_breaks = "1 month", guide = guide_axis(angle = 45)) +
    theme_bw() +
    theme(legend.position = "bottom", legend.title = element_blank())

  return(g)
}

state_trends <- covidcast_signals(
  SOURCES,
  SIGNALS,
  "2020-04-15", "2021-04-15",
  geo_type = "state"
)

g <- get_trend_grid(state_trends, function(val) { abbr_to_name(toupper(val)) })
```

```{r state-trend-grids-1, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. Cases are reported cases per 100,000 population. Other signals are scaled to have the same global maximum, so they can be presented in the same range. (part 1 of 4)", fig.height=8}
g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 1)
```

```{r state-trend-grids-2, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. (part 2 of 4)", fig.height=8}
g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 2)
```

```{r state-trend-grids-3, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. (part 3 of 4)", fig.height=8}
g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 3)
```

```{r state-trend-grids-4, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. (part 4 of 4)", fig.height=8}
g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 4)
```

\clearpage

```{r county-trends}
# Same thing again, but the 50 largest counties
## top_counties <- county_census %>%
##   filter(COUNTY != 0) %>% # these are states
##   arrange(desc(POPESTIMATE2019)) %>%
##   head(n = 50) %>%
##   pull(FIPS)

## county_trends <- covidcast_signals(
##   SOURCES,
##   SIGNALS,
##   "2020-04-15", "2021-04-15",
##   geo_type = "county",
##   geo_values = top_counties
## )

## g <- get_trend_grid(state_trends, county_fips_to_name)
```

```{r county-trend-grids-1, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 1 of 4)"}
## g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 1)
```

<!-- PNAS says: Add this line AFTER all your figures and tables -->
\FloatBarrier

### 2021-06-17 forecast/paper/paper.tex
Typo fixes, update revision fig

--- 2021-06-17 forecast/paper/paper.tex
@@
 Our take on this problem is in line with much of this literature; however, in
-order to avoid making any explicit assumptions, we do attempt to make explicit
+order to avoid making any explicit assumptions, we do not attempt to make formal 
 significance statements, and instead, broadly examine the stability of our
@@
 using only the data that would have been available to us \textit{as of}
-September 1 (imagine that we ``rewind'' the clock to September 1, and query the
-COVIDcast API to get the latest data it would have had available at that point
-in time.)  This is made possible by the COVIDcast API's comprehensive data  
+September 1 (imagine that we can ``rewind'' the clock to September 1, and query
+the COVIDcast API to get the latest data it would have had available at that
+point in time.)  This is made possible by the COVIDcast API's comprehensive data   
 versioning system (described in more detail in our companion paper).  We also
@@
 \begin{figure}[tb!]
-  \includegraphics[width=\columnwidth]{fig/revisions-dv-jhu-1.pdf}
-  \caption{Revision behavior for signals in the New York City HRR from two
-    sources. Each colored line corresponds to the data as reported on that
-    date. DV-CLI is revised upward throughout the month of December, though the
-    amount fades as we look further back in time. In contrast, JHU-CSSE data
-    remains ``as reported'' on December 3 until a major correction is made on
-    December 31, affecting all data over the period. }
+  \includegraphics[width=\columnwidth]{fig/revisions.pdf}
+  \caption{Revision behavior for two signals in the HRR containing New York City
+    HRR.  Each colored line corresponds to the data as reported on a particular
+    date (\textit{as of} dates varying from September 28 through October
+    19). The left panel shows the DV-CLI signal, which was regularly revised
+    throughout the period, though the effects fade as we look further back in   
+    time. In contrast, the right panel shows case rates reported by JHU CSSE,
+    which remain ``as reported'' on September 28, with a huge spike towards the 
+    end of this period, until a major correction is made on October 19, which
+    brings down this spike and affects all prior data as well.}
   \label{fig:vintage}
@@
 
-In addition to the above ``pure'' autoregressive model, we consider five
+In addition to this pure autoregressive model, we also consider five
 probabilistic forecasting models of the form:  
@@
 \caption{Forecast from the autoregressive model for the New York City HRR made
-  on October 15 (vertical line). 
-  The fan displays 50\%, 80\% and 95\% intervals while the orange curve shows
-  the median forecast. The connected black points represent ``finalized'' data,
-  as reported in May 2021.}
+  on October 15 (vertical line).  The fan displays 50\%, 80\% and 95\% intervals
+  while the orange curve shows the median forecast. The connected black points
+  represent ``finalized'' data, as reported in May 2021.}
 \label{fig:trajectory}
@@
 
-In addition to this ``pure'' autoregressive model, we also consider five
+In addition to this pure autoregressive model, we also consider five
 logistic models of the form:  
@@
 2020).  Shown is the average WIS for each forecast model divided by the averaged
-WIS from a ``baseline'' model, which is basically a flat-line forecaster (its
+WIS from a baseline model, which is basically a flat-line forecaster (its
 median forecast for $Y_{\ell,t+a}$ is always $Y_{\ell,t}$, with predicted
@@
 interpretable scale.  We can see that all curves are well below 1, which means
-that all of the models, including AR, outperform the baseline on average over
-the forecast period (though their advantage over the baseline clearly narrows as
+(smaller WIS is better) that all of the models, including AR, outperform the
+baseline on average over the forecast period (though this advantage narrows as
 the ahead value grows).
@@
 16--December 31, 2020).  We can see many similarities to the forecasting
-results: CHNG-COVID and DV-CLI offer the biggest improvement over AR, all models
-including AR degrade in performance towards the baseline (this is now just a
-classifier based on random guessing, which achieves an AUC of 0.5) as the ahead
-values grow.  However, one clear difference is that the CHNG-CLI model performs
-quite well in hotspot prediction, close to the best-performing indicator models
-for many of the ahead values. [TODO: should we insert one line explaining
-this---you're somewhat protected form volatiliy in a logistic model thanks to
+results (recall larger AUC is better): CHNG-COVID and DV-CLI offer the biggest
+improvement over AR, all models including AR degrade in performance towards the
+baseline (this is now just a classifier based on random guessing, which achieves
+an AUC of 0.5) as the ahead values grow.  However, one clear difference is that
+the CHNG-CLI model performs quite well in hotspot prediction, close to the
+best-performing indicator models for many of the ahead values.
+
+[TODO: should we
+insert one line explaining this---you're somewhat protected form volatiliy in a
+logistic model thanks to 
 the sigmoid function used to define the probabilities?]
@@
 anything to do with the actual indicator themselves; could it instead be that
-the indicators are simply providing ``implicit regularization'' on top of the
-basic AR model, when we include them as lagged features in
+the indicators are simply providing \textit{implicit regularization} on top of 
+the basic AR model, when we include them as lagged features in
 \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar_x}?

### 2021-06-17 forecast/code/ny-trajectory.R

@@ -1,3 +1,5 @@
+library(covidcast)
+library(tidyverse)
 
 ny_preds <- readRDS(file.path(path_to_data, "predictions_honest.RDS")) %>%
   filter(geo_value == "303", 

### 2021-06-17 forecast/code/revisions.R

@@ -1,8 +1,9 @@
 # Large spike in cases on 9/25 due to inclusion of positive antigen tests as 
 # probable cases. Backfilled by JHU on 10/2 or there abouts.
 # See https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data#retrospective-reporting-of-probable-cases-and-deaths
-# 
-# HRR 311 is Charlotte NC
+
+library(covidcast)
+library(tidyverse)
 
 as_ofs <- seq(as.Date("2020-09-28"), as.Date("2020-10-19"), by = "week")
 cases_as_of <- map_dfr(as_ofs, function(as_of) {
@@ -20,17 +21,21 @@ dv_as_of <-  map_dfr(as_ofs, function(as_of) {
     mutate(as_of = as_of)
 })
 
-
-pp <- bind_rows(cases_as_of, dv_as_of) %>% 
+  
+bind_rows(cases_as_of, dv_as_of) %>% 
   mutate(as_of = fct_relabel(factor(as_of), function(x) strftime(x, "%b %d")),
          data_source = recode(data_source, `doctor-visits` = "DV-CLI",
                               `jhu-csse` = "JHU-CSSE")) %>%
   ggplot(aes(x = time_value, y = value)) + 
   geom_line(aes(color = factor(as_of))) + 
-  facet_wrap(~data_source, scales = "free_y") +
+  facet_wrap(~data_source, scales = "free_y",
+             strip.position = "left", 
+             labeller = as_labeller(c(`DV-CLI` = "% doctor's visits due to CLI",
+                                      `JHU-CSSE` = "Cases per 100,000 people"))) +
   theme_bw() +
-  labs(x = "", y = "", color = "") +
+  labs(x = "", y = "", color = "As of:") +
   theme(legend.position = "bottom") +
   scale_color_viridis_d(end = .9, begin = .1)
 
+ggsave("../paper/fig/revisions.pdf", width = 5.5, height = 4)
 

### 2021-06-17 forecast/paper/supplement.Rmd

@@
        aes(x = cor_wisratio_pctchange)) +
-  geom_histogram(aes(y = ..count.. / sum(..count..), fill = forecaster)) +
+  geom_histogram(aes(y = ..count.. / sum(..count..), fill = forecaster), nbins = 100) +
   scale_fill_manual(values = fcast_colors) +
@@
   theme(legend.position = "none") +
-  geom_vline(aes(xintercept = median)) + 
-  geom_vline(xintercept = 0, linetype = "dotted") +
+  geom_vline(aes(xintercept = median), linetype = "dotted") + 
+  geom_vline(xintercept = 0) +
   facet_wrap(~forecaster) + 
@@
        aes(x = cor_abs_wisratio_minus_1_pctchange)) +
-  geom_histogram(aes(y = ..count.. / sum(..count..), fill = forecaster)) +
+  geom_histogram(aes(y = ..count.. / sum(..count..), fill = forecaster), nbins = 100) +
   scale_fill_manual(values = fcast_colors) +
@@
   filter(period != "jm") %>%
-  group_by(forecaster, geo_value, udf) %>%
+  group_by(forecaster, udf) %>%
+  mutate(m = Mean(AR_wis - wis)) %>%
   ggplot(aes(AR_wis - wis, fill = forecaster)) +
@@
   xlab("AR WIS - forecaster WIS") +
-  geom_vline(xintercept = 0)
+  geom_vline(xintercept = 0) +
+  geom_vline(aes(xintercept = m), linetype = "dashed")
 ```

### 2021-06-17 forecast/code/ny-trajectory.R
add trajectory plot data

@@ -1,18 +1,21 @@
 
-ny_preds <- readRDS(file.path(path_to_data, "predictions_honest.RDS")) %>%
-  filter(geo_value == "303", 
-         forecast_date == "2020-10-15",
-         forecaster == "AR3") %>%
-  mutate(target_end_date = forecast_date + days(ahead),
-         incidence_period = "day")
+# ny_preds <- readRDS(file.path(path_to_data, "predictions_honest.RDS")) %>%
+#   filter(geo_value == "303", 
+#          forecast_date == "2020-10-15",
+#          forecaster == "AR3") %>%
+#   mutate(target_end_date = forecast_date + days(ahead),
+#          incidence_period = "day")
+# 
+# ny_actuals <- readRDS(file.path(path_to_data, "actuals.RDS")) %>%
+#   filter(geo_value == "303", 
+#          target_end_date > "2020-09-01", 
+#          target_end_date < "2020-11-15") %>%
+#   rename(value = actual, time_value = target_end_date)
+#   
+ny_preds <- readRDS("../data/ny_predictions.RDS")
+ny_actuals <- readRDS("../data/ny_actuals.RDS")
 
-actuals <- readRDS(file.path(path_to_data, "actuals.RDS")) %>%
-  filter(geo_value == "303", 
-         target_end_date > "2020-09-01", 
-         target_end_date < "2020-11-15") %>%
-  rename(value = actual, time_value = target_end_date)
-
-pd <- evalcast:::setup_plot_trajectory(ny_preds, side_truth = actuals)
+pd <- evalcast:::setup_plot_trajectory(ny_preds, side_truth = ny_actuals)
 pd$truth_df <- pd$truth_df %>% rename(target_end_date = time_value)
 
 gg <- ggplot(pd$truth_df, mapping = aes(x = target_end_date)) +

### 2021-06-17 forecast/paper/paper.tex
added up/down supplement section

@@
 logistic model thanks to 
-the sigmoid function used to define the probabilities?]
+the sigmoid function used to define the probabilities? DJM: I think not, seems a
+bit far afield.]
 
@@
 
-While the performance of all forecast degrades during up periods (WIS is on the
+While the performance of all forecasters degrades during up periods (WIS is on the
 scale of cases, so higher case counts means higher WIS) there are pronounced
@@
 Google-AA is the only model that
-significantly outperforms the AR model during upswings. Both Google-AA and
-CTIS-CLI-in-community have large right tails during the flat periods while the
-claims signals have large left tails (indicating some large, relatively poor
-forecasts) during flat and up periods. CHNG-CLI, CHNG-COVID, Google-AA and
+outperforms the AR model, on average, during upswings (see the supplement for
+these averaged conclusions). Both Google-AA and
+CTIS-CLI-in-community have large right tails during the flat periods. CHNG-CLI
+and DV-CLI have large left tails (indicating some large, relatively poor
+forecasts) during flat and up periods. CHNG-CLI, CHNG-COVID, Google-AA, and
 especially CTIS-CLI-in-community have large positive outliers during the down
 period. Overall, the indicators seem to help more during periods of flat or down
-cases than during up periods, with the exception of Google-AA. In the
-supplement, we examine classification accuracy and log-likelihood for the
+cases than during periods of increasing case rates, with the exception of
+Google-AA.
+In the supplement, we examine classification accuracy and log-likelihood for the
 hotspot task. We see a similar phenomenon: the indicators dramatically improve
@@
 \end{figure}
+
+The supplement pursues this analysis further. Essentially, during downswings,
+improved performance is strongly correlated with predicing larger declines than
+the AR model. The AR model tends to overpredict relative to the decline, but
+adding indicators tends to diminish this effect. During upswings, the AR model
+tends to underpredict, but when the addition of an indicator improves
+performance (a more rare occurrence) it is rarely due to higher median
+predictions.
 

### 2021-06-17 forecast/paper/supplement.tex
added up/down supplement section

@@
 
-The goal of this section is to quantify the effect of
-not properly accounting for the question of ``what was known when'' in
-performing retrospective evaluations of forecasters.   Figures \ref{fig:fcast-finalized} and \ref{fig:hot-finalized} show
-what Figures 3 and 4 in the main paper would have looked like if we
-had simply trained all models using the finalized data rather than
-using vintage data.  This comparison can be seen most
-straightforwardly in Figures \ref{fig:fcast-honest-v-finalized} and \ref{fig:hot-honest-v-finalized}, which show the ratio in performance between the vintage and finalized
-versions.  When methods are
-given the finalized version of the data rather than the version available at the time that the forecast would have been made, all methods
-appear (misleadingly) to have better performance than they would
-actually have had if run prospectively.  For example, for forecasting case rates
-7-days ahead, the WIS of all methods is at least 8\% larger than what would have been
-recorded using the finalized values of the data.  This effect
-diminishes as the forecasting horizon increases, reflecting the fact
-that these forecasters rely less heavily on recent data than very
-short-term forecasters.  Crucially, some methods are
-``helped'' more than others by the less scrupulous retrospective
-evaluation, underscoring the difficulty of avoiding misleading
-conclusions when performing retrospective evaluations of forecasters.
+The goal of this section is to quantify the effect of not properly
+accounting for the question of ``what was known when'' in performing
+retrospective evaluations of forecasters. Figures
+\ref{fig:fcast-finalized} and \ref{fig:hot-finalized} show what Figures
+3 and 4 in the main paper would have looked like if we had simply
+trained all models using the finalized data rather than using vintage
+data. This comparison can be seen most straightforwardly in Figures
+\ref{fig:fcast-honest-v-finalized} and \ref{fig:hot-honest-v-finalized},
+which show the ratio in performance between the vintage and finalized
+versions. When methods are given the finalized version of the data
+rather than the version available at the time that the forecast would
+have been made, all methods appear (misleadingly) to have better
+performance than they would actually have had if run prospectively. For
+example, for forecasting case rates 7-days ahead, the WIS of all methods
+is at least 8\% larger than what would have been recorded using the
+finalized values of the data. This effect diminishes as the forecasting
+horizon increases, reflecting the fact that these forecasters rely less
+heavily on recent data than very short-term forecasters. Crucially, some
+methods are ``helped'' more than others by the less scrupulous
+retrospective evaluation, underscoring the difficulty of avoiding
+misleading conclusions when performing retrospective evaluations of
+forecasters.
 
-The \chngcli~indicator (along with the other claims-based signals) is the most
-affected by this distinction, reflecting the latency in claims-based
-reporting.  This supports the importance of efforts to provide ``nowcasts'' for claims signals
-(which corresponds to a 0-ahead ``forecast'' of what the claims
-signal's value will be once all data has been collected). Looking at
-the \chngcli~and \dv~curves in Figure \ref{fig:fcast-finalized}, we
-can see that they perform very similarly when trained on the finalized data.  This is reassuring
-because they are in principle measuring the same thing (namely, the percentage of
-  outpatient visits that are primarily about COVID-related symptoms).
-  The substantial difference in their curves in Figure 3 of the
-  main paper must therefore reflect their having very different
-  backfill profiles.
+The \chngcli\textasciitilde indicator (along with the other claims-based
+signals) is the most affected by this distinction, reflecting the
+latency in claims-based reporting. This supports the importance of
+efforts to provide ``nowcasts'' for claims signals (which corresponds to
+a 0-ahead ``forecast'' of what the claims signal's value will be once
+all data has been collected). Looking at the \chngcli\textasciitilde and
+\dv\textasciitilde curves in Figure \ref{fig:fcast-finalized}, we can
+see that they perform very similarly when trained on the finalized data.
+This is reassuring because they are in principle measuring the same
+thing (namely, the percentage of outpatient visits that are primarily
+about COVID-related symptoms). The substantial difference in their
+curves in Figure 3 of the main paper must therefore reflect their having
+very different backfill profiles.
 
-While \dv~is one of the least affected methods by using finalized
-rather than vintage values, it is one of the most affected methods for the hotspot problem.
-This is a reminder that the forecasting and hotspot problems are
-distinct problems.  For example, the hotspot problem does not measure
-the ability to distinguish between flat and downward trends.
+While \dv\textasciitilde is one of the least affected methods by using
+finalized rather than vintage values, it is one of the most affected
+methods for the hotspot problem. This is a reminder that the forecasting
+and hotspot problems are distinct problems. For example, the hotspot
+problem does not measure the ability to distinguish between flat and
+downward trends.
 
-Even the \ar~model is affected by this distinction, reflecting the
-fact that the case rates themselves (i.e., the response values) are also
-subject to revision.  The forecasters based on indicators are thus
-affected both by revisions to the indicators and by revisions to the
-case rates.  In the case of the \gs~model, in which we only used finalized
-values for the \gs~indicator, the difference in performance can be
+Even the \ar\textasciitilde model is affected by this distinction,
+reflecting the fact that the case rates themselves (i.e., the response
+values) are also subject to revision. The forecasters based on
+indicators are thus affected both by revisions to the indicators and by
+revisions to the case rates. In the case of the
+\gs\textasciitilde model, in which we only used finalized values for the
+\gs\textasciitilde indicator, the difference in performance can be
 wholly attributed to revisions of case rates.
@@
 In this section, we consider using the geometric mean instead of the
-arithmetic mean when aggregating the weighted interval score (WIS) across location-time pairs.
-There are three reasons why the geometric mean may be desirable.
+arithmetic mean when aggregating the weighted interval score (WIS)
+across location-time pairs. There are three reasons why the geometric
+mean may be desirable.
+
 \begin{enumerate}
 \item  WIS is right-skewed, being bounded below by zero and having
-  occasional very large values.   Figure\textasciitilde{}\ref{fig:wis-densities} illustrates that the
+  occasional very large values.   Figure \ref{fig:wis-densities} illustrates that the
   densities appear roughly log-Gaussian.  The geometric mean is a
@@
 
-Figure\textasciitilde{}\ref{fig:fcast-adjusted} uses the geometric
-mean for aggregation.  Comparing this with Figure 3 of the main paper,
-we see that the main conclusions are largely unchanged; however,
-\chngcli~now does appear better than \ar.  This behavior would be
-expected if \chngcli's poor performance is attributable to a
-relatively small number of large errors (as opposed to a large number
-of moderate errors).  Indeed, Figure 5 of the main paper further
-corroborates this, in which we see the heaviest left tails occuring
-for \chngcli.
+Figure \ref{fig:fcast-adjusted} uses the geometric mean for aggregation.
+Comparing this with Figure 3 of the main paper, we see that the main
+conclusions are largely unchanged; however, \chngcli\textasciitilde now
+does appear better than \ar. This behavior would be expected if
+\chngcli's poor performance is attributable to a relatively small number
+of large errors (as opposed to a large number of moderate errors).
+Indeed, Figure 5 of the main paper further corroborates this, in which
+we see the heaviest left tails occuring for \chngcli.
 
 \hypertarget{bootstrap-results}{%
-\section{Evaluating the implicit regularization hypothesis}\label{bootstrap-results}}
+\section{Bootstrap results}\label{bootstrap-results}}
 
 As explained in Section 2.B. of the main paper, a (somewhat cynical)
-hypothesis for why we see benefits in forecasting and hotspot
-prediction is that the indicators are not actually providing useful
-information but they are instead simply acting as a sort of ``implicit
-regularization,''  leading to shrinkage on the autoregressive
-coefficients, leading to less volatile predictions.  To investigate
-this hypothesis, we consider fitting  ``noise features'' that in truth
-should have zero coefficients.  Recall (from the main paper) that at each forecast date, we
-train a model on 6,426 location-time pairs.  Indicator models are
-based on six features, corresponding to the three autoregressive terms
-and the three lagged indicator values.  To form ``noise features''
-we resample from other rows.  In particular, at each location $\ell$ and
-time $t$, we replace the triplet $(X_{\ell,t}, X_{\ell,t-7},
-X_{\ell,t-14})$ by the triplet $(X_{\ell^*,t^*}, X_{\ell^*,t^*-7},
-X_{\ell^*,t^*-14})$, where $(\ell^*,t^*)$ is a location-time pair
-sampled with replacement from the 6,426 location-time pairs.  Figures
-\ref{fig:fcast-booted}--\ref{fig:hot-booted} show the results.  No
-method exhibits a noticeable performance gain over the \ar~method,
-leading us to dismiss the implicit regularization hypothesis.
+hypothesis for why we see benefits in forecasting and hotspot prediction
+is that the indicators are not actually providing useful information but
+they are instead simply acting as a sort of implicit regularization,"
+leading to shrinkage on the autoregressive coefficients, leading to less
+volatile predictions. To investigate this hypothesis, we consider
+fitting ``noise features'' that in truth should have zero coefficients.
+Recall (from the main paper) that at each forecast date, we train a
+model on 6,426 location-time pairs. Indicator models are based on six
+features, corresponding to the three autoregressive terms and the three
+lagged indicator values. To form noise features we resample from other
+rows. In particular, at each location \(\ell\) and time \(t\), we
+replace the triplet \((X_{\ell,t}, X_{\ell,t-7}, X_{\ell,t-14})\) by the
+triplet \((X_{\ell^*,t^*}, X_{\ell^*,t^*-7}, X_{\ell^*,t^*-14})\), where
+\((\ell^*,t^*)\) is a location-time pair sampled with replacement from
+the 6,426 location-time pairs. Figures
+\ref{fig:fcast-booted}--\ref{fig:hot-booted} show the results. No method
+exhibits a noticeable performance gain over the
+\ar\textasciitilde method, leading us to dismiss the implicit
+regularization hypothesis.
 
@@
 
-\hypertarget{correlations-with-lagged-actuals}{%
-\section{Correlations with lagged
-actuals}\label{correlations-with-lagged-actuals}}
-
-Alden's histograms are in Figure\textasciitilde{}\ref{fig:cor-wis-ratio}
-and Figure\textasciitilde{}\ref{fig:cor-wis-ratio-m1}.
-
 \hypertarget{upswings-and-downswings}{%
@@
 
-Logged version of Figure 5 in the manuscript is in
-Figure\textasciitilde{}\ref{fig:upswing-histogram-logged}.
+In this section we provide extra details about the upswing / flat /
+downswing analysis described in the main text. Figure
+\ref{fig:upswing-summary} shows the overall results, examining the
+average difference \(\WIS(AR) - \WIS(F)\) in period. Figure
+\ref{fig:hotspots-upswing-downswing} shows the same information for the
+hotspot task. On average, during downswings and flat periods, the
+indicator-assisted models have lower classification error and higher log
+likelihood than the AR model. For hotspots, both Google-AA and
+CTIS-CLIIC perform better than the AR model during upswings, in contrast
+to the forecasting task, where only Google-AA improves. For a related
+analysis, Figure \ref{fig:cor-wis-ratio} shows histograms of the
+Spearman correlation (Spearman's \(\rho\), a rank-based measure of
+association) between the \(\WIS(F)/\WIS(AR)\) and the magnitude of the
+swing. Again we see that case rate increases are positively related to
+diminished performance of the indicator models.
 
-See also Table\textasciitilde{}\ref{tab:upswing-corr-table}.
+One hypothesis for diminished relative performance during upswings is
+that the AR model tends to overpredict downswings and underpredict
+upswings. Adding indicators seems to help avoid this behavior on the
+downswing but not as much on upswings. Figure
+\ref{fig:upswing-corr-table} shows the correlation between between
+\(\WIS(AR) - \WIS(F)\) and the difference of their median forecasts.
+During downswings, this correlation is large, implying that improved
+relative performance of \(F\) is related to making lower forecasts than
+the AR model. The opposite is true during upswings. This is largely to
+be expected. However, the relationship attenuates in flat periods and
+during upswings. That is, when performance is better in those cases, it
+may be due to other factors than simply making predictions in the
+correct direction, for example, narrower confidence intervals.
 
@@
 In this section, we investigate the sensitivity of the results to the
-period over which we train and evaluate the models.  In the main
-paper, we end all evaluation on December 31, 2020.  Figures
-\ref{fig:fcast-alldates} -- \ref{fig:hot-alldates} show how the
-results would differ if we extended this analysis through \attn{XX,
-  XX}, 2021. Comparing Figure \ref{fig:fcast-alldates} to Figure 3 of
-the main paper, one sees that as ahead increases most methods now improve
-relative to the baseline forecaster. When compared to other methods, \chngcli~appears much better than
-it had previously; however, all forecasters other than \chngcov~and
-\dv~are performing less well relative to the baseline than before.
-These changes are likely due to the differing nature of the pandemic
-in 2021, with flat and downward trends much more common than upward
-trajectories.  Indeed, the nature of the hotspot prediction problem is
-quite different in this period.  With a 21-day training window, it is
-common for there to be very few hotspots in training.
+period over which we train and evaluate the models. In the main paper,
+we end all evaluation on December 31, 2020. Figures
+\ref{fig:fcast-alldates} -- \ref{fig:hot-alldates} show how the results
+would differ if we extended this analysis through March 31, 2021.
+Comparing Figure \ref{fig:fcast-alldates} to Figure 3 of the main paper,
+one sees that as ahead increases most methods now improve relative to
+the baseline forecaster. When compared to other methods,
+\chngcli\textasciitilde appears much better than it had previously;
+however, all forecasters other than \chngcov\textasciitilde and
+\dv\textasciitilde are performing less well relative to the baseline
+than before. These changes are likely due to the differing nature of the
+pandemic in 2021, with flat and downward trends much more common than
+upward trajectories. Indeed, the nature of the hotspot prediction
+problem is quite different in this period. With a 21-day training
+window, it is common for there to be very few hotspots in training.
 
@@
 
-{\centering \includegraphics[width=\textwidth]{fig/cor-wis-ratio-1}
+{\centering \includegraphics[width=\textwidth]{fig/upswing-summary-1}
 
@@
 
-\caption{This is one of the correlation plots Alden made.
-It shows histograms of the Spearman correlation between the ratio of AR to AR WIS with the percent change in 7dav cases relative to 7 days earlier.}\label{fig:cor-wis-ratio}
+\caption{Average difference between the WIS of the AR model and the WIS of the other forecasters. The indicator-assisted forecasters do best during down and flat periods.}\label{fig:upswing-summary}
 \end{figure}
-
-\clearpage
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{fig/cor-wis-ratio-m1-1}
-
-}
-
-\caption{This is Alden's second set of histograms.
-Here we have the correlation of the absolute value of WIS ratio - 1 with the percent change in 7dav cases relative to 7 days earlier}\label{fig:cor-wis-ratio-m1}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{fig/upswing-histogram-logged-1}
-
-}
-
-\caption{Not sure if we want this here. Similar to Figure 5 in the manuscript but taking logs. }\label{fig:upswing-histogram-logged}
-\end{figure}
-
-\clearpage
-
-\begin{table}
-
-\caption{\label{tab:upswing-corr-table}Correlation of the difference in WIS between the AR model with the difference in median predictions.
-In down periods, improvements in forecast risk are highly correlated with lower median predictions.
-The opposite is true in up periods.
-This suggests, as one might expect that improved performance of the indicator-assisted model is attributable to being closer to the truth then the AR model.
-This conclusion is stronger in down periods then in up periods.}
-\centering
-\begin{tabular}[t]{lrrrrr}
-\toprule
-udf & CHNG-CLI & CHNG-COVID & CTIS-CLIIC & DV-CLI & Google-AA\\
-\midrule
-down & 0.79 & 0.80 & 0.83 & 0.81 & 0.81\\
-flat & 0.12 & 0.17 & 0.28 & 0.18 & 0.17\\
-up & -0.57 & -0.56 & -0.53 & -0.53 & -0.47\\
-\bottomrule
-\end{tabular}
-\end{table}
 
@@
 Like the analysis of the forecasting task in the main paper (see Figure 7), performance is better during down and flat periods.}\label{fig:hotspots-upswing-downswing}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/cor-wis-ratio-1}
+
+}
+
+\caption{Histograms of the Spearman correlation between the ratio of AR to AR WIS with the percent change in smoothed case rates relative to 7 days earlier.}\label{fig:cor-wis-ratio}
+\end{figure}
+
+\clearpage
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/upswing-corr-table-1}
+
+}
+
+\caption{Correlation of the difference in WIS with the difference in median predictions for the AR model relative to the indicator-assisted forecaster.
+In down periods, improvements in forecast risk are highlycorrelated with lower median predictions.
+The opposite is true in up periods.
+This suggests, as one might expect that improved performance of the indicator-assisted model is attributable to being closer to the truth then the AR model.
+This conclusion is stronger in down periods then in up periods.}\label{fig:upswing-corr-table}
 \end{figure}

### 2021-06-18 forecast/paper/supplement.Rmd

--- 2021-06-16 forecast/paper/supplement.Rmd
@@
 
-This section referes to Figures~\ref{fig:fcast-finalized}--\ref{fig:hot-honest-v-finalized}.
+The goal of this section is to quantify the effect of
+not properly accounting for the question of "what was known when" in
+performing retrospective evaluations of forecasters.   Figures \ref{fig:fcast-finalized} and \ref{fig:hot-finalized} show
+what Figures 3 and 4 in the main paper would have looked like if we
+had simply trained all models using the finalized data rather than
+using vintage data.  This comparison can be seen most
+straightforwardly in Figures \ref{fig:fcast-honest-v-finalized} and \ref{fig:hot-honest-v-finalized}, which show the ratio in performance between the vintage and finalized
+versions.  When methods are
+given the finalized version of the data rather than the version available at the time that the forecast would have been made, all methods
+appear (misleadingly) to have better performance than they would
+actually have had if run prospectively.  For example, for forecasting case rates
+7-days ahead, the WIS of all methods is at least 8\% larger than what would have been
+recorded using the finalized values of the data.  This effect
+diminishes as the forecasting horizon increases, reflecting the fact
+that these forecasters rely less heavily on recent data than very
+short-term forecasters.  Crucially, some methods are
+"helped" more than others by the less scrupulous retrospective
+evaluation, underscoring the difficulty of avoiding misleading
+conclusions when performing retrospective evaluations of forecasters.
+
+The \chngcli~indicator (along with the other claims-based signals) is the most
+affected by this distinction, reflecting the latency in claims-based
+reporting.  This supports the importance of efforts to provide "nowcasts" for claims signals
+(which corresponds to a 0-ahead "forecast" of what the claims
+signal's value will be once all data has been collected). Looking at
+the \chngcli~and \dv~curves in Figure \ref{fig:fcast-finalized}, we
+can see that they perform very similarly when trained on the finalized data.  This is reassuring
+because they are in principle measuring the same thing (namely, the percentage of
+  outpatient visits that are primarily about COVID-related symptoms).
+  The substantial difference in their curves in Figure 3 of the
+  main paper must therefore reflect their having very different
+  backfill profiles.
+
+While \dv~is one of the least affected methods by using finalized
+rather than vintage values, it is one of the most affected methods for the hotspot problem.
+This is a reminder that the forecasting and hotspot problems are
+distinct problems.  For example, the hotspot problem does not measure
+the ability to distinguish between flat and downward trends.
+
+Even the \ar~model is affected by this distinction, reflecting the
+fact that the case rates themselves (i.e., the response values) are also
+subject to revision.  The forecasters based on indicators are thus
+affected both by revisions to the indicators and by revisions to the
+case rates.  In the case of the \gs~model, in which we only used finalized
+values for the \gs~indicator, the difference in performance can be
+wholly attributed to revisions of case rates.
 
@@
 
-* The weighted interval score is bounded below by zero and can be very large. This behavior is typical of right-skewed distributions.
-* Figure~\ref{fig:wis-densities} illustrates that the densities appear log-Gaussian.
-* This suggests aggregating by the geometric mean rather than the mean for comparisons. See Figure~\ref{fig:fcast-adjusted}.
+In this section, we consider using the geometric mean instead of the
+arithmetic mean when aggregating the weighted interval score (WIS) across location-time pairs.
+There are three reasons why the geometric mean may be desirable.
+\begin{enumerate}
+\item  WIS is right-skewed, being bounded below by zero and having
+  occasional very large values.   Figure \ref{fig:wis-densities} illustrates that the
+  densities appear roughly log-Gaussian.  The geometric mean is a
+  natural choice in such a context since the relative ordering of forecasters is
+  determined by the arithmetic mean of the {\em logarithm} of their WIS
+  values.
+\item In the main paper, we report the ratio of the mean WIS of a
+  forecaster to the mean WIS of the baseline forecaster. Another
+  choice could be to take the mean of the ratio of WIS values for the
+  two methods. (This latter choice would penalize a method less for
+  doing poorly where the baseline forecaster also does poorly.)
+Using instead the geometric mean makes the order of aggregation and
+scaling immaterial since the ratio of geometric means is the same as
+the geometric mean of ratios.
+\item If one imagines that a forecaster's WIS is composed of
+  multiplicative space-time effects $S_{\ell,t}$ shared across all forecasters,
+  i.e. $\WIS(F_{\ell,t,f},Y_{\ell,t})=S_{\ell,t}E_{f}$, then
+  taking the ratio of two forecasters' geometric mean WIS values will
+  effectively cancel these space-time effects.
+\end{enumerate}
+
+Figure \ref{fig:fcast-adjusted} uses the geometric
+mean for aggregation.  Comparing this with Figure 3 of the main paper,
+we see that the main conclusions are largely unchanged; however,
+\chngcli~now does appear better than \ar.  This behavior would be
+expected if \chngcli's poor performance is attributable to a
+relatively small number of large errors (as opposed to a large number
+of moderate errors).  Indeed, Figure 5 of the main paper further
+corroborates this, in which we see the heaviest left tails occuring
+for \chngcli.
 
@@
 
-Here we discuss how implicit regularization is not the reason for improved performance
+As explained in Section 2.B. of the main paper, a (somewhat cynical)
+hypothesis for why we see benefits in forecasting and hotspot
+prediction is that the indicators are not actually providing useful
+information but they are instead simply acting as a sort of implicit
+regularization,"  leading to shrinkage on the autoregressive
+coefficients, leading to less volatile predictions.  To investigate
+this hypothesis, we consider fitting  "noise features" that in truth
+should have zero coefficients.  Recall (from the main paper) that at each forecast date, we
+train a model on 6,426 location-time pairs.  Indicator models are
+based on six features, corresponding to the three autoregressive terms
+and the three lagged indicator values.  To form noise features
+we resample from other rows.  In particular, at each location $\ell$ and
+time $t$, we replace the triplet $(X_{\ell,t}, X_{\ell,t-7},
+X_{\ell,t-14})$ by the triplet $(X_{\ell^*,t^*}, X_{\ell^*,t^*-7},
+X_{\ell^*,t^*-14})$, where $(\ell^*,t^*)$ is a location-time pair
+sampled with replacement from the 6,426 location-time pairs.  Figures
+\ref{fig:fcast-booted}--\ref{fig:hot-booted} show the results.  No
+method exhibits a noticeable performance gain over the \ar~method,
+leading us to dismiss the implicit regularization hypothesis.
+
+\attn{Would be good for someone to double check that the bootstrap
+  procedure is correctly described.}
+
+<!--
 
@@
 
+-->
+
 # Upswings and Downswings
 
-Logged version of Figure 5 in the manuscript is in Figure~\ref{fig:upswing-histogram-logged}.
+In this section we provide extra details about the upswing / flat / downswing 
+analysis described in the main text. Figure \ref{fig:upswing-summary} shows
+the overall results, examining the average difference $\WIS(AR) - \WIS(F)$ in 
+period. Figure \ref{fig:hotspots-upswing-downswing} shows the same information
+for the hotspot task. On average, during downswings and flat periods, the 
+indicator-assisted models have lower classification error and higher 
+log likelihood than the AR model. For hotspots, both Google-AA and CTIS-CLIIC
+perform better than the AR model during upswings, in contrast to the forecasting
+task, where only Google-AA improves. For a related analysis, Figure 
+\ref{fig:cor-wis-ratio} shows histograms of the Spearman
+correlation (Spearman's $\rho$, a rank-based measure of association) between
+the $\WIS(F)/\WIS(AR)$ and the magnitude of the swing. Again we see that case
+rate increases are positively related to diminished performance of the 
+indicator models.
 
-See also Table~\ref{tab:upswing-corr-table}.
+One hypothesis for diminished relative performance during upswings is that 
+the AR model tends to overpredict downswings and underpredict upswings. Adding
+indicators seems to help avoid this behavior on the downswing but not as much 
+on upswings. Figure \ref{fig:upswing-corr-table} shows the correlation between
+between $\WIS(AR) - \WIS(F)$ and the difference of their median forecasts. During
+downswings, this correlation is large, implying that improved relative performance
+of $F$ is related to making lower forecasts than the AR model. The opposite is
+true during upswings. This is largely to be expected. However, the relationship
+attenuates in flat periods and during upswings. That is, when performance is better
+in those cases, it may be due to other factors than simply making predictions
+in the correct direction, for example, narrower confidence intervals.
 
@@
 
-See Figures \ref{fig:fcast-alldates} -- \ref{fig:hot-alldates}.
+In this section, we investigate the sensitivity of the results to the
+period over which we train and evaluate the models.  In the main
+paper, we end all evaluation on December 31, 2020.  Figures
+\ref{fig:fcast-alldates} -- \ref{fig:hot-alldates} show how the
+results would differ if we extended this analysis through March
+  31, 2021. Comparing Figure \ref{fig:fcast-alldates} to Figure 3 of
+the main paper, one sees that as ahead increases most methods now improve
+relative to the baseline forecaster. When compared to other methods, 
+\chngcli~appears much better than
+it had previously; however, all forecasters other than \chngcov~and
+\dv~are performing less well relative to the baseline than before.
+These changes are likely due to the differing nature of the pandemic
+in 2021, with flat and downward trends much more common than upward
+trajectories.  Indeed, the nature of the hotspot prediction problem is
+quite different in this period.  With a 21-day training window, it is
+common for there to be very few hotspots in training.
 
@@
 
-<!-- Correlations with lagged actuals -->
-
-```{r cor-with-actuals}
-comparison_df <- left_join(
-  fcasts_honest %>% filter(forecaster != "AR", period != "jm"),
-  fcasts_honest %>% filter(forecaster == "AR", period != "jm") %>%
-    select(-c(forecaster,target_end_date)) %>%
-    rename(ar3_wis = wis),
-  by = c("geo_value", "ahead", "forecast_date")) %>%
-  group_by(forecaster, forecast_date, geo_value) %>%
-  summarize(wis_ratio = mean(wis) / mean(ar3_wis), .groups = "drop")
-
-pct_chng_df <- readRDS(file.path(
-  path_to_data, 
-  "jhu-csse_confirmed_7dav_incidence_num_2021-05-18.RDS")) %>%
-  pct_change(n = 7) %>%
-  select(geo_value, time_value, pct_change) %>%
-  filter(time_value %in% unique(comparison_df$forecast_date))
-
-comparison_df <- comparison_df %>% 
-  left_join(pct_chng_df, 
-            by = c("forecast_date" = "time_value", "geo_value"))
-
-pct_chng_df <- comparison_df %>% 
-  group_by(forecaster, geo_value) %>%
-  summarize(
-    cor_wisratio_pctchange = cor(wis_ratio, pct_change, method = "spearman"),
-    cor_abs_wisratio_minus_1_pctchange = cor(
-      abs(wis_ratio - 1), pct_change,method = "spearman"), .groups = "drop")
-```
-
-```{r cor-wis-ratio, fig.cap="This is one of the correlation plots Alden made.
-It shows histograms of the Spearman correlation between the ratio of AR to AR WIS with the percent change in 7dav cases relative to 7 days earlier."}
-ggplot(pct_chng_df %>% 
-         group_by(forecaster) %>% 
-         mutate(median = median(cor_wisratio_pctchange)), 
-       aes(x = cor_wisratio_pctchange)) +
-  geom_histogram(aes(y = ..count.. / sum(..count..), fill = forecaster)) +
-  scale_fill_manual(values = fcast_colors) +
-  theme_bw() +
-  theme(legend.position = "none") +
-  geom_vline(aes(xintercept = median)) + 
-  geom_vline(xintercept = 0, linetype = "dotted") +
-  facet_wrap(~forecaster) + 
-  xlab("Spearman correlation") + ylab("Relative frequency") +
-  scale_y_continuous(labels = scales::label_percent(accuracy = 1))
-```
-
-\clearpage
-
-```{r cor-wis-ratio-m1, fig.cap="This is Alden's second set of histograms.
-Here we have the correlation of the absolute value of WIS ratio - 1 with the percent change in 7dav cases relative to 7 days earlier"}
-ggplot(pct_chng_df %>% 
-         group_by(forecaster) %>% 
-         mutate(median = median(cor_abs_wisratio_minus_1_pctchange)), 
-       aes(x = cor_abs_wisratio_minus_1_pctchange)) +
-  geom_histogram(aes(y = ..count.. / sum(..count..), fill = forecaster)) +
-  scale_fill_manual(values = fcast_colors) +
-  theme_bw() +
-  theme(legend.position = "none") +
-  geom_vline(aes(xintercept = median)) + 
-  geom_vline(xintercept = 0, linetype = "dotted") +
-  facet_wrap(~forecaster) + 
-  xlab("Spearman correlation") + ylab("Relative frequency") +
-  scale_y_continuous(labels = scales::label_percent(accuracy = 1))
-```
-
-\clearpage
-
 <!-- Upswings vs. downswings -->
@@
 
-df <- left_join(corr_df %>% 
+up_down_df <- left_join(corr_df %>% 
                   ungroup() %>%
@@
 
+```{r upswing-summary, fig.cap="Average difference between the WIS of the AR model and the WIS of the other forecasters. The indicator-assisted forecasters do best during down and flat periods."}
+# Figure 5 in the manuscript
+up_down_df_summary <- up_down_df %>%
+  filter(period != "jm") %>%
+  group_by(forecaster, udf) %>%
+  summarise(mean = Mean(AR_wis - wis))
+up_down_df_summary %>%
+  ggplot(aes(udf, mean , fill = forecaster)) +
+  geom_bar(width = 0.6, position = position_dodge(width=0.6),
+           stat = "identity") +
+  scale_fill_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
+  scale_y_continuous() +
+  theme_bw() +
+  geom_hline(yintercept = 0) +
+  ylab("Mean of WIS(AR) - WIS(F)") +
+  xlab("") +
+  theme(legend.position = "bottom", legend.title = element_blank())
+```
+
 ```{r upswing-histogram, include=FALSE}
-# Figure 5 in the manuscript
-df %>% 
+up_down_df %>% 
   filter(period != "jm") %>%
-  group_by(forecaster, geo_value, udf) %>%
+  group_by(forecaster, udf) %>%
   ggplot(aes(AR_wis - wis, fill = forecaster)) +
@@
 
-```{r upswing-histogram-logged, fig.cap="Not sure if we want this here. Similar to Figure 5 in the manuscript but taking logs. "}
-df_summary <- df %>%
-  filter(period != "jm") %>%
-  group_by(forecaster, udf) %>%
-  summarise(m = GeoMean((AR_wis + 1) / (wis + 1)))
-df %>% 
+```{r upswing-histogram-logged, eval = FALSE, fig.cap="Not sure if we want this here. Similar to Figure 5 in the manuscript but taking logs. "}
+up_down_df %>% 
   filter(period != "jm") %>%
@@
   xlab("AR WIS - forecaster WIS") +
-  geom_vline(xintercept = 1) +
-  geom_vline(data = df_summary, aes(xintercept = m), linetype = "dotted")
-```
-
-\clearpage
-
-```{r upswing-corr-table}
-df %>% 
-  filter(period != "jm") %>% 
-  group_by(forecaster, udf) %>%
-  summarise(cor = cor(AR_wis - wis, AR_med - med)) %>%
-  pivot_wider(names_from = forecaster, values_from = cor) %>%
-  kableExtra::kbl(
-    booktabs = TRUE, digits = 2, centering = TRUE,
-    caption = paste("Correlation of the difference in WIS between the AR",
-                    "model with the difference in median predictions. In down",
-                    "periods, improvements in forecast risk are highly",
-                    "correlated with lower median predictions. The opposite",
-                    "is true in up periods. This suggests, as one might expect",
-                    "that improved performance of the indicator-assisted",
-                    "model is attributable to being closer to the truth then",
-                    "the AR model. This conclusion is stronger in down",
-                    "periods then in up periods."))
+  geom_vline(xintercept = 1)
 ```
@@
   theme(legend.position = "bottom", legend.title = element_blank())
+```
+
+\clearpage
+
+<!-- Correlations with lagged actuals -->
+
+```{r cor-with-actuals}
+comparison_df <- left_join(
+  fcasts_honest %>% filter(forecaster != "AR", period != "jm"),
+  fcasts_honest %>% filter(forecaster == "AR", period != "jm") %>%
+    select(-c(forecaster,target_end_date)) %>%
+    rename(ar3_wis = wis),
+  by = c("geo_value", "ahead", "forecast_date")) %>%
+  group_by(forecaster, forecast_date, geo_value) %>%
+  summarize(wis_ratio = mean(wis) / mean(ar3_wis), .groups = "drop")
+
+pct_chng_df <- readRDS(file.path(
+  path_to_data, 
+  "jhu-csse_confirmed_7dav_incidence_num_2021-05-18.RDS")) %>%
+  pct_change(n = 7) %>%
+  select(geo_value, time_value, pct_change) %>%
+  filter(time_value %in% unique(comparison_df$forecast_date))
+
+comparison_df <- comparison_df %>% 
+  left_join(pct_chng_df, 
+            by = c("forecast_date" = "time_value", "geo_value"))
+
+pct_chng_df <- comparison_df %>% 
+  group_by(forecaster, geo_value) %>%
+  summarize(
+    cor_wisratio_pctchange = cor(wis_ratio, pct_change, method = "spearman"),
+    cor_abs_wisratio_minus_1_pctchange = cor(
+      abs(wis_ratio - 1), pct_change,method = "spearman"), .groups = "drop")
+```
+
+```{r cor-wis-ratio, fig.cap="Histograms of the Spearman correlation between the ratio of AR to AR WIS with the percent change in smoothed case rates relative to 7 days earlier."}
+ggplot(pct_chng_df %>% 
+         group_by(forecaster) %>% 
+         mutate(median = median(cor_wisratio_pctchange)), 
+       aes(x = cor_wisratio_pctchange)) +
+  geom_histogram(aes(y = ..count.. / sum(..count..), fill = forecaster), bins = 50) +
+  scale_fill_manual(values = fcast_colors) +
+  theme_bw() +
+  theme(legend.position = "none") +
+  geom_vline(aes(xintercept = median), linetype = "dotted") + 
+  geom_vline(xintercept = 0) +
+  facet_wrap(~forecaster) + 
+  xlab("Spearman correlation") + ylab("Relative frequency") +
+  scale_y_continuous(labels = scales::label_percent(accuracy = .1))
+```
+
+\clearpage
+
+```{r cor-wis-ratio-m1, eval = FALSE, fig.cap="This is Alden's second set of histograms.
+Here we have the correlation of the absolute value of WIS ratio - 1 with the percent change in 7dav cases relative to 7 days earlier"}
+ggplot(pct_chng_df %>% 
+         group_by(forecaster) %>% 
+         mutate(median = median(cor_abs_wisratio_minus_1_pctchange)), 
+       aes(x = cor_abs_wisratio_minus_1_pctchange)) +
+  geom_histogram(aes(y = ..count.. / sum(..count..), fill = forecaster), bins = 50) +
+  scale_fill_manual(values = fcast_colors) +
+  theme_bw() +
+  theme(legend.position = "none") +
+  geom_vline(aes(xintercept = median), linetype = "dotted") + 
+  geom_vline(xintercept = 0) +
+  facet_wrap(~forecaster) + 
+  xlab("Spearman correlation") + ylab("Relative frequency") +
+  scale_y_continuous(labels = scales::label_percent(accuracy = 1))
+```
+
+\clearpage
+
+```{r upswing-corr-table, fig.cap="Correlation of the difference in WIS with the difference in median predictions for the AR model relative to the indicator-assisted forecaster.
+In down periods, improvements in forecast risk are highlycorrelated with lower median predictions.
+The opposite is true in up periods.
+This suggests, as one might expect that improved performance of the indicator-assisted model is attributable to being closer to the truth then the AR model.
+This conclusion is stronger in down periods then in up periods."}
+up_down_df %>% 
+  filter(period != "jm") %>% 
+  group_by(forecaster, udf) %>%
+  summarise(cor = cor(AR_wis - wis, AR_med - med)) %>%
+  ggplot(aes(udf, cor , fill = forecaster)) +
+  geom_bar(width = 0.6, position = position_dodge(width=0.6),
+           stat = "identity") +
+  scale_fill_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
+  theme_bw() +
+  geom_hline(yintercept = 0) +
+  theme(legend.position = "bottom", legend.title = element_blank())
+
+# up_down_df %>% 
+#   filter(period != "jm") %>% 
+#   group_by(forecaster, udf) %>%
+#   summarise(cor = cor(AR_wis - wis, AR_med - med)) %>%
+#   pivot_wider(names_from = forecaster, values_from = cor) %>%
+#   kableExtra::kbl(
+#     booktabs = TRUE, digits = 2, centering = TRUE,
+#     caption = paste("Correlation of the difference in WIS between the AR",
+#                    "model with the difference in median predictions. In down",
+ #                   "periods, improvements in forecast risk are highly",
+#                    "correlated with lower median predictions. The opposite",
+#                    "is true in up periods. This suggests, as one might expect",
+#                    "that improved performance of the indicator-assisted",
+#                    "model is attributable to being closer to the truth then",
+#                    "the AR model. This conclusion is stronger in down",
+#                    "periods then in up periods."))
 ```
@@
 df2 <- left_join(
-  df, 
+  up_down_df, 
   lead_lag_metric,
@@
 
-<!-- Figures in the manuscript -->
+<!-- Figures in the manuscript in their own scripts -->
 
-```{r revisions-dv-jhu, include=FALSE}
+```{r revisions-dv-jhu, eval=FALSE}
 # Figure 1 in the manuscript
-source("../code/revisions-dv-jhu.R")
-pp
+source("../code/revisions.R")
 ```
 
-```{r ny-trajectory, include=FALSE}
+```{r ny-trajectory, eval=FALSE}
 # Figure 2 in the manuscript
@@
 
-```{r ccf-dv-finalized, include=FALSE}
+```{r ccf-dv-finalized, eval=FALSE}
 # Figure ?? in the manuscript

### 2021-06-18 indicators/paper/supplement.Rmd

@@
 disease. The annotations here represent the data sources collected by Delphi's
-COVIDcast Epidata API.}
+COVIDcast Epidata API. TODO Edit to use consistent language to paper}
 \end{figure}
@@
 Cases are reported cases per 100,000 population.
-Other signals are scaled to have the same global maximum, so they can be presented in the same range.
+Other signals are scaled to have the same global maximum across all counties and times, so they can be presented in the same range.
 (part 1 of 4)", fig.height=8}
@@
 ```
+
+\clearpage
 
@@
 
+\clearpage
+
 ```{r state-trend-grids-3, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. (part 3 of 4)", fig.height=8}
@@
 # Same thing again, but the 50 largest counties
-## top_counties <- county_census %>%
-##   filter(COUNTY != 0) %>% # these are states
-##   arrange(desc(POPESTIMATE2019)) %>%
-##   head(n = 50) %>%
-##   pull(FIPS)
+top_counties <- county_census %>%
+  filter(COUNTY != 0) %>% # these are states
+  arrange(desc(POPESTIMATE2019)) %>%
+  head(n = 50) %>%
+  pull(FIPS)
 
-## county_trends <- covidcast_signals(
-##   SOURCES,
-##   SIGNALS,
-##   "2020-04-15", "2021-04-15",
-##   geo_type = "county",
-##   geo_values = top_counties
-## )
+# Work around an API problem: specifying geo_values in the query causes it to
+# fail on the CHNG data because of bad database indexing. Filter after the
+# fact instead. DANGER: Very slow.
+county_trends_nochng <- covidcast_signals(
+  SOURCES[c(1,4)], SIGNALS[c(1,4)],
+  "2020-04-15", "2021-04-15",
+  geo_type = "county",
+  geo_values = top_counties
+)
 
-## g <- get_trend_grid(state_trends, county_fips_to_name)
+county_trends_chng <- covidcast_signals(
+  SOURCES[2:3], SIGNALS[2:3],
+  "2020-04-15", "2021-04-15",
+  geo_type = "county"
+)
+
+county_trends_chng <- lapply(
+  county_trends_chng,
+  function(df) {
+    filter(df, geo_value %in% top_counties)
+  }
+)
+
+county_trends <- append(county_trends_nochng, county_trends_chng[1])
+county_trends <- append(county_trends, county_trends_chng[2])
 ```
 
-```{r county-trend-grids-1, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 1 of 4)"}
-## g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 1)
+```{r county-trend-grids-1, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States.
+Cases are reported cases per 100,000 population.
+Other signals are scaled to have the same global maximum across all counties and times, so they can be presented in the same range.
+(part 1 of 4)", fig.height=8}
+g <- get_trend_grid(
+  county_trends,
+  function(fipsen) {
+    paste0(county_fips_to_name(fipsen), ", ", name_to_abbr(state_fips_to_name(fipsen)))
+  }
+)
+
+g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 1)
+```
+
+\clearpage
+
+```{r county-trend-grids-2, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 2 of 4)", fig.height=8}
+g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 2)
+```
+
+\clearpage
+
+```{r county-trend-grids-3, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 3 of 4)", fig.height=8}
+g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 3)
+```
+
+\clearpage
+
+```{r county-trend-grids-4, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 4 of 4)", fig.height=8}
+g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 4)
 ```

### 2021-06-18 forecast/code/revisions.R
Plot updates

@@ -37,5 +37,5 @@ bind_rows(cases_as_of, dv_as_of) %>%
   theme(legend.position = "bottom") +
   scale_color_viridis_d(end = .9, begin = .1)
 
-ggsave("../paper/fig/revisions.pdf", width = 5.5, height = 4)
+ggsave("../paper/fig/revisions.pdf", width = 5, height = 4)
 

### 2021-06-18 forecast/code/trajectory.R

@@ -1,26 +1,14 @@
 library(covidcast)
+library(evalcast)
 library(tidyverse)
 
-# ny_preds <- readRDS(file.path(path_to_data, "predictions_honest.RDS")) %>%
-#   filter(geo_value == "303", 
-#          forecast_date == "2020-10-15",
-#          forecaster == "AR3") %>%
-#   mutate(target_end_date = forecast_date + days(ahead),
-#          incidence_period = "day")
-# 
-# ny_actuals <- readRDS(file.path(path_to_data, "actuals.RDS")) %>%
-#   filter(geo_value == "303", 
-#          target_end_date > "2020-09-01", 
-#          target_end_date < "2020-11-15") %>%
-#   rename(value = actual, time_value = target_end_date)
-#   
 ny_preds <- readRDS("../data/ny_predictions.RDS")
 ny_actuals <- readRDS("../data/ny_actuals.RDS")
 
 pd <- evalcast:::setup_plot_trajectory(ny_preds, side_truth = ny_actuals)
 pd$truth_df <- pd$truth_df %>% rename(target_end_date = time_value)
 
-gg <- ggplot(pd$truth_df, mapping = aes(x = target_end_date)) +
+ggplot(pd$truth_df, mapping = aes(x = target_end_date)) +
   geom_ribbon(
     data = pd$quantiles_df,
     mapping = aes(ymin = lower, ymax = upper, fill = interval)) +
@@ -30,11 +18,14 @@ gg <- ggplot(pd$truth_df, mapping = aes(x = target_end_date)) +
   geom_line(data = pd$points_df, 
             mapping = aes(y = value),
             color = "orange", size = .5) +
-  geom_point(data = pd$points_df, 
+  geom_point(data = pd$points_df,
             mapping = aes(y = value),
             color = "orange", size = 1) +
   geom_vline(xintercept = as.Date("2020-10-15")) +
+  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
   theme_bw(base_size = 14) + 
   theme(legend.position = "none") + 
-  ylab("AR forecast for case rate in NYC") + 
-  xlab("date")
+  ylab("Cases per 100,000 people") + 
+  xlab("Date")
+
+ggsave("../paper/fig/trajectory.pdf", width = 6, height = 4)

### 2021-06-18 forecast/paper/paper.tex
Edits to up, down, flat

--- 2021-06-17 forecast/paper/paper.tex
@@
 \begin{figure}[t]
-\includegraphics[width=\columnwidth]{fig/ny-trajectory-1.pdf}
-\caption{Forecast from the autoregressive model for the New York City HRR made
+\includegraphics[width=\columnwidth]{fig/trajectory.pdf}
+\caption{Forecast from an autoregressive model for the New York City HRR made 
   on October 15 (vertical line).  The fan displays 50\%, 80\% and 95\% intervals
-  while the orange curve shows the median forecast. The connected black points
-  represent ``finalized'' data, as reported in May 2021.}
+  while the orange curve shows the median forecast. The black curve represents
+  ``finalized'' data, as reported in May 2021.} 
 \label{fig:trajectory}
@@
     t-7})/(X_{\ell,t-7})$}, and again $X_{\ell,t}$ stands for any of the five  
-auxiliary indicators at location $\ell$ and time $t$.  We fit all models
-\eqref{eq:hotspot_ar}, \eqref{eq:hotspot_ar_x} using logistic regression, 
+auxiliary indicators at location $\ell$ and time $t$.  We fit the above models,
+\eqref{eq:hotspot_ar}, \eqref{eq:hotspot_ar_x}, using logistic regression,   
 pooling all 306 HRRs and using a 21-day trailing window for the training data.
@@
 \eqref{eq:wis_quantiles} is more general (it can accomodate asymmetric quantile
-levels), the first definition \eqref{eq:wis_intervals} is usually preferred in
+levels), the first form in \eqref{eq:wis_intervals} is typically preferred in
 presentation, because the score nicely decouples into a ``sharpness'' component
@@
 words, values of the Google-AA signal are clearly \textit{not} missing at
-random.  It seems most appropriate to impute missing values by zero, and this 
-is what we do in our analysis.
+random.  It seems most appropriate to impute missing values by zero, and this is
+what we do in our analysis.
 
@@
 following subsections, we will use ``AR'' to refer to the pure autoregressive
-model both in forecasting \eqref{eq:forecast_ar} and in hotspot prediction
-\eqref{eq:hotspot_ar} (where the reference to the task should be clear from the
-context). We will also use the name of an auxiliary indicator---``CHNG-CLI'', 
-``CHNG-CLI'', ``CHNG-COVID'', ``CTIS-CLI-in-community'', ``DV-CLI'', or
-``Google-AA''---interchangeably with the model in forecasting
-\eqref{eq:forecast_ar_x} or hotspot prediction \eqref{eq:hotspot_ar_x} that uses
-this particular indicator as a feature (the meaning should be clear from the
-context).  So, e.g., the CHNG-CLI model in forecasting is the one in
+model both in forecasting, \eqref{eq:forecast_ar}, and in hotspot prediction,
+\eqref{eq:hotspot_ar} (the reference to the prediction task should always be
+clear from the context). We will also use the name of an auxiliary
+indicator---namely ``CHNG-CLI'',  ``CHNG-COVID'', ``CTIS-CLI-in-community'',
+``DV-CLI'', or ``Google-AA''---interchangeably with the model in forecasting, 
+\eqref{eq:forecast_ar_x}, or hotspot prediction, \eqref{eq:hotspot_ar_x}, that
+uses this particular indicator as a feature (the meaning should be clear from
+the context).  So, for example, the CHNG-CLI model in forecasting is the one in  
 \eqref{eq:forecast_ar_x} that sets $X_{\ell,t}$ to be the value of the CHNG-CLI 
-indicator at location $\ell$ and time $t$, fits the coefficients based on the
-training data, and makes its predictions by setting $X_{\ell,t}$ to the
-appropriate CHNG-CLI values.  Finally, we use the term ``indicator model'' to 
-refer to any one of the ten models of the form \eqref{eq:forecast_ar_x},
-\eqref{eq:hotspot_ar_x} (five from each of the forecasting and hotspot
-prediction tasks).
+indicator at location $\ell$ and time $t$.
+% fits the coefficients based on the training data, and makes its predictions by
+% setting $X_{\ell,t}$ to the appropriate CHNG-CLI values.  
+Finally, we use the term ``indicator model'' to refer to any one of the ten
+models of the form \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar_x} (five from
+each of the forecasting and hotspot prediction tasks).
 
@@
 divide by the average WIS of the baseline model in order to put the y-axis on an 
-interpretable scale.  We can see that all curves are well below 1, which means
+interpretable scale.  We can see that all curves are below 1, which means
 (smaller WIS is better) that all of the models, including AR, outperform the
-baseline on average over the forecast period (though this advantage narrows as
-the ahead value grows).
+baseline on average over the forecasting period.  On the other hand, the models 
+deliver at best an improvement of about 20\% in mean WIS over the baseline
+model, with this gap narrowing to only 10\% at the largest ahead values,
+illustrating the difficulty of the forecasting problem.
 
@@
   \includegraphics[width=\columnwidth]{fig/fcast-1.pdf}
-  \caption{Average WIS for each forecast model divided by the averaged WIS of
-    the flat-line forecaster.}
+  \caption{Average WIS for each forecast model, over all forecast dates and all
+    HRRs, divided by the average WIS achieved by a baseline model (a
+    probabilistic version of the flat-line forecaster).  Here and in all figures
+  we abbreviate CTIS-CLI-in-community by CTIS-CLIIC.}  
   \label{fig:forecast}
@@
 16--December 31, 2020).  We can see many similarities to the forecasting
-results (recall larger AUC is better): CHNG-COVID and DV-CLI offer the biggest
-improvement over AR, all models including AR degrade in performance towards the
-baseline (this is now just a classifier based on random guessing, which achieves
-an AUC of 0.5) as the ahead values grow.  However, one clear difference is that
-the CHNG-CLI model performs quite well in hotspot prediction, close to the
-best-performing indicator models for many of the ahead values.
-
-[TODO: should we
-insert one line explaining this---you're somewhat protected form volatiliy in a
-logistic model thanks to 
-the sigmoid function used to define the probabilities?]
+results (now larger AUC is better); for example, CHNG-COVID and DV-CLI offer the
+biggest improvement over AR, and all models including AR degrade in performance 
+towards the baseline (this is now just a classifier based on random guessing,
+which achieves an AUC of 0.5) as the ahead values grow.  A clear difference,
+however, is that the CHNG-CLI model performs quite well in hotspot prediction,
+close to the best-performing indicator models for many of the ahead values. This
+may be because volatility in the CHNG-CLI indicator plays less of a role in the
+associated logistic model's predicted probabilities (in general, a sigmoid
+function can absorb a lot of the variability in its input).
 
@@
 One might ask if the benefits we observe in forecasting and hotspot
-prediction have
-anything to do with the actual indicator themselves; could it instead be that
-the indicators are simply providing \textit{implicit regularization} on top of 
-the basic AR model, when we include them as lagged features in
-\eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar_x}?
+prediction have anything to do with the actual indicator themselves; could it
+instead be that the indicators are simply providing \textit{implicit
+  regularization} on top of  the basic AR model, when we include them as lagged
+features in \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar_x}?
 
-To test this implicit regularization hypothesis, we reran all of the prediction
-experiments but with $X_{\ell,t}$ in each indicator model \eqref{eq:forecast_ar_x},
-\eqref{eq:hotspot_ar_x} replaced with suitable random noise (bootstrap
-samples form an indicator's historical distribution).  The results, shown and
-explained more precisely in the supplement, are vastly different (worse) than
-the original set of results.  In both forecasting and hotspot prediction, the
-``fake'' indicator models (with random noise as auxiliary features) offered
-essentially no improvement over the pure AR model, which provides a strong
-rejection (informally speaking) of the implicit regularization hypothesis.
+To test this hypothesis, we reran all of the prediction experiments but with
+$X_{\ell,t}$ in the each indicator models in \eqref{eq:forecast_ar_x},
+\eqref{eq:hotspot_ar_x} replaced with suitable random noise (bootstrap samples
+from a signal's historical distribution).  The results, shown and explained more
+precisely in the supplement, are vastly different (worse) than the original set
+of results.  In both forecasting and hotspot prediction, the ``fake'' indicator
+models (with random noise as auxiliary features) offered essentially no
+improvement over the pure AR model, which provides a strong rejection
+(informally speaking) of the implicit regularization hypothesis.
+
+Along the lines of regularization, it is also worth noting that the use of
+$\ell_1$ regularization (tuned by cross-validation) in fitting any of the models
+in \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar}, 
+\eqref{eq:hotspot_ar_x} did not improve their generally performance at all
+(experiments not shown).  This is likely due to the fact that the number of
+training samples is quite large compared to the number of features (6,426
+training samples and only 3--6 features).
 
@@
 \includegraphics[width=\columnwidth]{fig/hot-1.pdf}
-\caption{Area under the curve for each forecast model averaged over all
-  prediction dates and HRRS.}
+\caption{Area under the curve for each hotspot prediction model, calculated over
+  all prediction dates and all HRRs.} 
 \label{fig:hotspot}
@@
 
-\subsection{Performance in Up, Down, Flat Periods}
+\subsection{Performance in Up, Down, and Flat Periods}
+
+The course of the pandemic has played out quite differently across space and
+time. Aggregating case rates nationally shows three pronounced waves, but the
+behavior is more nuanced at the HRR level.  Figure \ref{fig:trajectory} is a
+single example of a forecast in a period of relatively flat case trends, as New
+York City enters what would become its second wave. The AR forecaster's 50\%
+prediction interval contains this upswing, but its forecasted median is clearly
+below the finalized case data.  Unfortunately, this behavior is fairly typical
+of all forecasters: during upswings, the forecasted median tends to fall below 
+the target, while the reverse is true during downswings.
+
+Figure~\ref{fig:up_down_flat} shows histograms of the differences in WIS of the 
+AR model and each indicator model, where we stratify these differences by
+whether the target occurs during a period of increasing cases rates (up),
+decreasing case rates (down), or flat case rates (flat). To define the
+increasing period, we use the same definition we used for the hotspot task in
+Table \ref{tab:analysis_tasks}.  Therefore all hotspots are ``up'', while all
+non-hotspots are either ``flat'' or ``down''.  For the ``down'' scenario, we 
+simply use the opposite of the hotspot definition: $Y_{\ell,t}$ decreases by
+more than 25\% relative to the preceding week.
+
+While the performance of all forecasters, including AR, will generally degrade
+in up periods, different models exhibit different and interesting patterns.
+CHNG-CLI, CHNG-COVID, Google-AA, and especially CTIS-CLI-in-community have
+large right tails (displaying improvements over AR) over the down periods.
+Google-AA and CTIS-CLI-in-community have large right tails during the flat
+periods. CHNG-CLI and DV-CLI have large left tails (poor forecasts relative to
+AR) in flat and up periods.  Google-AA is the only model that outperforms
+the AR model, on average, over up periods.  Overall, the indicators seem to help
+more during flat or down periods than up periods, with the exception of
+Google-AA.
 
@@
   \includegraphics[width=\columnwidth]{fig/upswing-histogram-1.pdf}
-  \caption{Histogram of the difference in WIS for the AR model relative to each
-    forecaster. Larger values are better for the indicator assisted model.}
-  \label{fig:up-down-flat}
+  \caption{Histogram of the difference in WIS for the AR model and that for each
+    indicator model, stratified by up, down, or flat periods, measured in terms
+    of case activity. (Larger values are better for the indicator model.) The
+    y-axis is on the log scale to emphasize tail behavior.} 
+  \label{fig:up_down_flat}
 \end{figure}
 
+The supplement pursues this analysis further.  For example, we examine
+classification accuracy and log-likelihood for the hotspot task, and find a
+similar phenomenon: the indicators considerably improve accuracy or
+log-likelihood during flat or down periods, with more mixed behavior during up
+periods, with the CHNG-CLI, CHNG-COVID, and DV-CLI leading to decreased
+performance in particular.
+
+% Essentially, during downswings, improved performance is strongly correlated
+% with predicing larger declines than the AR model. The AR model tends to
+% overpredict relative to the decline, but adding indicators tends to diminish
+% this effect. During upswings, the AR model tends to underpredict, but when the
+% addition of an indicator improves performance (a more rare occurrence) it is 
+% rarely due to higher median predictions.
+
+% RJT: Commented out the last paragraph for now, because this is kind of a tough
+% thing to explain succintly. Can revisit if we really want to include it here,
+% or just let them find it in the supplement?
+
 \subsection{Effects of Leading or Lagging Behavior}
+
+As described in the methods section, each of the indicators we examine could be
+said to measure aspects on disease progression that would precede a positive
+test. That is, we imagine that these signals should ``lead'' cases. It is
+entirely reasonable to imagine that, prior to an increase of confirmed COVID-19
+tests reported by public health in a particular location, we would see an
+increase in insurance claims for COVID-19-related medical visits.  However, it
+may well be the case that this leading behavior is different during different
+periods. In fact, we find that the ``leadingness'' of an indicator (degree to
+which it leads) tends to be more pronounced during down and flat periods than up
+periods, a plausible explanation for the decreased performance noted above.
+
+In order to measure the leadingness on af indicator, we examine its cross
+correlations to case rates over a short time window
+
+ the cross
+correlation function. Essentially this examines the Pearson correlation between
+
+shifted version of the indicator and the case rate. So if $X_{\ell,t}$ is
+shifted forward by 7 days, and the magnitude of the correlation of that shifted series with
+$Y_{\ell,t}$ (unshifted) is high, we would say that $X_{\ell}$ leads $Y_{\ell}$
+by 7 days. This process is repeated for a range of shifts, forward and backward,
+resulting in leadingness and laggingness scores. Further details are given in
+the supplement.
+
+Figure \ref{fig:leading-lagging} shows the correlations between the difference
+in the WIS of the AR model relative to $F$ with the leading and laggingness
+scores stratified by whether the target is classified as up, down, or flat. One
+would expect that increases in $\WIS(AR) - \WIS(AR)$ would be positively
+correlated with leadingness. This relationship is strongest during down periods
+and weakest during up periods with flat in the middle. In fact, for each
+indicator, the strength of correlations between improved performance and
+leadingness can be ordered down $>$ flat $>$ up. The relationship with
+laggingness is the opposite, though less pronounced. CHNG-CLI's forecast
+performance doesn't correlate very strongly with laggingness while any
+laggingness in Google-AA during down and flat periods results in realtively
+worse forecasting performance.
+
+Because a time series can, to a degree, both simultaneously leading and lagging,
+we also examine the correlation between performance and the difference between
+leadingness and laggingness. Figure \ref{fig:diff-btw-leading-lagging} shows
+this behavior. Again, indicators that are more leading then lagging tend to
+are positively correlated with improved relative performance. This relationship
+is most pronounced during periods of decreasing case rates and attenuates in
+other scenarios.
 
@@
 COVIDcast API (defined using from medical insurance claims, internet-based
-surveys, and Google search trends) is undoubtedly ``yes''. However, there are 
+surveys, and Internet search trends) is undoubtedly ``yes''. However, there are  
 levels of nuance to such an answer that must be explained.  None of the

### 2021-06-18 indicators/paper/paper.tex
More figure jostling
Copyedits to discussion

@@
 in the signals, as well as many other aspects, such as correlations, anomalies,
-and backfill.  There is also a dedicated dashboard for exploring results from
-the COVID Trends and Impact Survey.
+and backfill. There is also a dedicated dashboard for exploring results from the
+COVID Trends and Impact Survey.
 
@@
 
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{fig/time_wise_correlation.pdf}
+  \caption{Time-wise correlations with case rates, from April 15, 2020 to April
+    15, 2021, calculated over all counties for which all signals were available
+    and which had at least 500 cumulative cases by the end of this period.}
+  \label{fig:time_wise_correlation}
+\end{figure}
+
 Figure~\ref{fig:geo_wise_correlation} shows the geo-wise correlations achieved
@@
 symptoms and case trends, especially when other data is unavailable.
-
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/time_wise_correlation.pdf}
-  \caption{Time-wise correlations with case rates, from April 15, 2020 to April
-    15, 2021, calculated over all counties for which all signals were available
-    and which had at least 500 cumulative cases by the end of this period.}
-  \label{fig:time_wise_correlation}
-\end{figure}
 
@@
 or geographic differences in market share for medical claims data. Claims data
-can also be subject to biases during major national holidays and other events
-that change health-seeking behavior. Characterizing these biases will be
-important for future research and operational systems that use these
-signals. Several data sources are also subject to extensive revision and
-backfill, which must be studied and modeled to enable effective real-time use of
-these sources in forecasting and nowcasting systems. The breadth and unique
-features of the COVIDcast API will help facilitate this work, which will be
+tends also to be subject to biases during major holidays and other events that
+change health-seeking behavior. Characterizing these biases will be important
+for future research and operational systems that use these signals. Several data
+sources are also subject to extensive revision and backfill, which must be
+studied and modeled to enable effective real-time use of these sources in
+forecasting and nowcasting systems. The breadth and unique features of the
+COVIDcast API will help facilitate this and other related work, which will be
 vital to advancing pandemic modeling and preparedness.

### 2021-06-18 forecast/paper/paper.tex
FINISH discussion

@@
 
-Along the lines of regularization, it is also worth noting that the use of
+On the topic of regularization, it is also worth noting that the use of 
 $\ell_1$ regularization (tuned by cross-validation) in fitting any of the models
@@
 (experiments not shown).  This is likely due to the fact that the number of
-training samples is quite large compared to the number of features (6,426
-training samples and only 3--6 features).
+training samples is large compared to the number of features (6,426 training
+samples and only 3--6 features).
 
@@
 
-\subsection{Performance in Up, Down, and Flat Periods}
+\subsection{Evaluation in Up, Down, and Flat Periods}
 
@@
   \caption{Histogram of the difference in WIS for the AR model and that for each
-    indicator model, stratified by up, down, or flat periods, measured in terms
-    of case activity. (Larger values are better for the indicator model.) The
-    y-axis is on the log scale to emphasize tail behavior.} 
+    indicator model, stratified by up, down, or flat period, measured in terms
+    of case trends. Note that \textit{larger} differences here are better for
+    each indicator model.  The y-axis is on the log scale to emphasize tail 
+    behavior.}   
   \label{fig:up_down_flat}
@@
 entirely reasonable to imagine that, prior to an increase of confirmed COVID-19
-tests reported by public health in a particular location, we would see an
-increase in insurance claims for COVID-19-related medical visits.  However, it
-may well be the case that this leading behavior is different during different
-periods. In fact, we find that the ``leadingness'' of an indicator (degree to
-which it leads) tends to be more pronounced during down and flat periods than up
-periods, a plausible explanation for the decreased performance noted above.
+tests reported by a public health authority in a particular location, we would
+see an increase in medical insurance claims for COVID-related outpatient
+visits. However, it may well be the case that such behavior is different
+during different periods. In fact, we find empirically that the ``leadingness''
+of an indicator (degree to which it leads case activity) tends to be more
+pronounced in down or flat periods than up periods, a plausible explanation
+for the decreased performance in up periods noted above.
 
-In order to measure the leadingness on af indicator, we examine its cross
-correlations to case rates over a short time window
-
- the cross
-correlation function. Essentially this examines the Pearson correlation between
-
-shifted version of the indicator and the case rate. So if $X_{\ell,t}$ is
-shifted forward by 7 days, and the magnitude of the correlation of that shifted series with
-$Y_{\ell,t}$ (unshifted) is high, we would say that $X_{\ell}$ leads $Y_{\ell}$
-by 7 days. This process is repeated for a range of shifts, forward and backward,
-resulting in leadingness and laggingness scores. Further details are given in
-the supplement.
-
-Figure \ref{fig:leading-lagging} shows the correlations between the difference
-in the WIS of the AR model relative to $F$ with the leading and laggingness
-scores stratified by whether the target is classified as up, down, or flat. One
-would expect that increases in $\WIS(AR) - \WIS(AR)$ would be positively
-correlated with leadingness. This relationship is strongest during down periods
-and weakest during up periods with flat in the middle. In fact, for each
-indicator, the strength of correlations between improved performance and
-leadingness can be ordered down $>$ flat $>$ up. The relationship with
-laggingness is the opposite, though less pronounced. CHNG-CLI's forecast
-performance doesn't correlate very strongly with laggingness while any
-laggingness in Google-AA during down and flat periods results in realtively
-worse forecasting performance.
-
-Because a time series can, to a degree, both simultaneously leading and lagging,
-we also examine the correlation between performance and the difference between
-leadingness and laggingness. Figure \ref{fig:diff-btw-leading-lagging} shows
-this behavior. Again, indicators that are more leading then lagging tend to
-are positively correlated with improved relative performance. This relationship
-is most pronounced during periods of decreasing case rates and attenuates in
-other scenarios.
+In the supplement, we show how to define a quantitative score to measure the
+leadingness of an indicator, at any time $t$ and any location $\ell$, based on
+cross correlations to case rates over a short time window around $t$.
+The higher this score, the greater it ``leads'' case activity.  Figure 
+\ref{fig:leading} displays correlations between the leadingness score of an 
+indicator and the WIS difference (AR model minus an indicator model), stratified
+by whether the target is classified as up, down, or flat.  One would expect that  
+the WIS difference would be positively correlated with leadingness.  Somewhat
+surprisingly, this relationship turns out to be the strongest in down periods,
+and weakest in up periods.  In fact, it is very nearly the case that for each
+indicator, the strength of correlations only decreases as we move from down to
+flat to up  periods.  In the supplement, we extend this analysis by examining
+analogous ``laggingness'' scores, but we do not find as clear patterns.
 
@@
   \includegraphics[width=\columnwidth]{fig/leading-and-lagging-1.pdf}
-  \caption{Correlation of the difference in WIS with the ``leadingness'' and
-    ``laggingness'' of the indicator at the target date.}
-  \label{fig:leading-lagging}
-\end{figure}
-
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/diff-in-lead-lag-1.pdf}
-  \caption{Correlation of the difference in WIS with difference between the
-    ``leadingness'' and ``laggingness'' of the indicator at the target date.}
-  \label{fig:diff-btw-leading-lagging}
+  \caption{Correlation of the difference in WIS with the ``leadingness'' of the
+    indicator at the target date, stratified by up, down, or flat period.} 
+  \label{fig:leading}
 \end{figure}
@@
 levels of nuance to such an answer that must be explained.  None of the
-indicators that we have investigated appear to be the ``silver bullet'' that 
-one might have hoped for, revolutionizing the tractability of the prediction 
+indicators that we have investigated appear to be the ``silver bullet'' that one
+might have hoped for, revolutionizing the tractability of the prediction
 problem, rendering it easy when it was once hard (in the absence of auxiliary
@@
 consistent across modes of analysis, but modest.  In forecasting, the indicator
-models appear to be most useful in periods in which case rates are flat or
+models are found to be most useful in periods in which case rates are flat or
 trending down, rather than periods in which case rates are trending up (as one
@@
 
-As mentioned previously, the indicator models could likely be improved by using 
-location-specific effects, as well as using nowcasting techniques to estimate
-finalized indicator values before using them as features (in order to account
+As described previously, it is likely that we could improve the indicator models
+by using location-specific effects, as well as using nowcasting techniques to
+estimate finalized indicator values before we use them as features (to account
 for backfill in the claims-based signals in particular).  Beyond this, it is
 certainly possible that more sophisticated models for forecasting or hotspot
-prediction would lead to different results, and even different
-insights. Natural directions to explore include using interactions, using
-multiple indicators in a single model, and leveraging HRR demographics or
+prediction would lead to different results, and possibly even different
+insights. Natural directions to explore include using multiple indicators in a
+single model, allowing for interaction terms, and leveraging HRR demographics or
 mobility patterns.  That said, we are doubtful that more sophisticated modeling
 techniques would change the ``topline'' conclusion---that auxiliary indicators
-can provide nontrivial, consistent, but modest gains in forecasting and
-hotspot prediction. Whether a more sophisticated model would be able to leverage
-the indicators in such a way as to change some of the finer conclusions---e.g.,
-by offering clear improvements in periods in which cases are trending up---is 
-less clear to us.
+can provide nontrivial, consistent, but modest gains in forecasting and hotspot
+prediction. Whether a more sophisticated model would be able to leverage the
+indicators in such a way as to change some of the finer conclusions---e.g., by
+offering clear improvements in periods in which cases are trending up---is less
+clear to us.
 
-Reiterate importance of AS OF
+We reiterate the importance of using vintage data for rigorous backtesting. Data
+sources that are relevant to public health surveillance are often subject to
+revision, sometimes regularly (such as medical claims data) and sometimes
+unpredictably (such as COVID-19 case reports).  When analyzing models that are
+designed to predict future events, if we train these models and make predictions
+using finalized data, then we are missing a big part of the story.  Not only
+will our sense of accuracy be unrealistic, but certain models may degrade by a
+greater or lesser extent when they are forced to work with vintage data, so 
+backtesting them with finalized data may lead us to make modeling decisions that
+may not be relevant for true test-time performance.
 
-Not the end of the story.  Should we even be forecasting case rates?  More
-careful models---combine hotspots and forecasting?
-
-[  Finish by reiterating the importance of vintage data.  Emphasizing the rigor
-or our approach, how we have provided a template for others to be able to do the
-same thing with new models ]
-
-% RJT: some text commented out below.  May be useful for finishing off the
-% discussion
-
-% This paper is as much about demonstrating how one might go about answering
-% such questions about feature importance in forecasting and hotspot prediction 
-% general, as providing an empirically rigorous answer to this precise question
-% in the context of the COVID-19 pandemic.
-
-% We conclude by observing that the approach we have taken in this paper is
-% somewhat opposite to that of much of the COVID-19 forecasting literature.  We
-% have chosen to consider only very simple forecasting models while devoting
-% most of our effort to accounting for as much of the complexity of the data and 
-% evaluation as possible.  By contrast, many papers focus on very complicated
-% forecasting approaches but then evaluate them under unrealistic, retrospective 
-% conditions.
+In this paper, we have chosen to consider only very simple forecasting models,
+while devoting most of our effort to accounting for as much of the complexity of
+the data and evaluation as possible.  In fact, our paper is as much about
+demonstrating how one might address questions about model comparisons and
+evaluation in forecasting and hotspot prediction in general, as it is about
+providing rigorous answers to such questions in the context of COVID-19 case
+rates in particular.  We hope that others will leverage our framework, and build
+on it, so that it can be used to guide work that advances the frontier of
+predictive modeling for epidemics and pandemics.
 

### 2021-06-18 forecast/paper/unused.tex
FINISH discussion

@@
+Essentially this examines the Pearson correlation between 
+shifted version of the indicator and the case rate. So if $X_{\ell,t}$ is
+shifted forward by 7 days, and the magnitude of the correlation of that shifted series with
+$Y_{\ell,t}$ (unshifted) is high, we would say that $X_{\ell}$ leads $Y_{\ell}$
+by 7 days. This process is repeated for a range of shifts, forward and backward,
+resulting in leadingness and laggingness scores. Further details are given in
+the supplement.
+
+%%%%%
+
+Because a time series can, to a degree, both simultaneously leading and lagging,
+we also examine the correlation between performance and the difference between
+leadingness and laggingness. Figure \ref{fig:diff-btw-leading-lagging} shows
+this behavior. Again, indicators that are more leading then lagging tend to
+are positively correlated with improved relative performance. This relationship
+is most pronounced during periods of decreasing case rates and attenuates in
+other scenarios.
+
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{fig/diff-in-lead-lag-1.pdf}
+  \caption{Correlation of the difference in WIS with difference between the
+    ``leadingness'' and ``laggingness'' of the indicator at the target date.}
+  \label{fig:diff-btw-leading-lagging}
+\end{figure}
+
 %%%%%

### 2021-06-18 indicators/paper/paper.tex
Typo fixes and suggestions from Jacob

@@
   cases and deaths, and many others: measures of mobility, social distancing,
-  Internet search trends, self-reported symptoms, and patterns COVID-related
+  Internet search trends, self-reported symptoms, and patterns of COVID-related
   activity in de-identified medical insurance claims. The database provides all
@@
 emerging pathogens such as H1N1 \cite{Kass-Hout:2012}; and digital surveillance
-(based on, e.g., search and social media trends) enable more accurate
+(based on, e.g., search and social media trends) enables more accurate
 ``nowcasts'' and forecasts of traditional disease surveillance streams such as
@@
 When appropriate, additional post-processing (often nontrivial) is applied to
-the data. For example, data on visits to doctor's offices is subject to strong
+the data. For example, data on visits to doctors' offices is subject to strong
 day-of-week effects, and so regression is used to adjust for these
@@
 For many purposes it is sufficient to use the most recently issued observation
-at any a given time value, and the COVIDcast API returns the most recent issue
-as its default. However, for some applications it is crucial to know what was
-known \textit{as of} a specific date. For example, an epidemic forecasting model
-will be called upon to make its forecasts based on preliminary data about recent
+at a given time value, and the COVIDcast API returns the most recent issue as
+its default. However, for some applications it is crucial to know what was known
+\textit{as of} a specific date. For example, an epidemic forecasting model will
+be called upon to make its forecasts based on preliminary data about recent
 trends, so when it is trained using historical data, it should be trained using
@@
   (CTIS-CLI-in-community): The estimated percentage of the population who know 
-  someone in their local community that is sick, based the same surveys.
+  someone in their local community that is sick, based on the same surveys.
 \item Quidel test positivity rate (Quidel-TPR): The percentage of positive
@@
   and $Y_\ell$ be vectors of a values of a particular signal and case rates,
-  over all times, at location $\ell$. The time-wise correlation at time $t$ is
-  defined as $\cor(X_\ell, Y_\ell)$. This examines whether changes in a signal
-  over time correspond to changes in reported cases at the same location.
+  over all times, at location $\ell$. The time-wise correlation at location
+  $\ell$ is defined as $\cor(X_\ell, Y_\ell)$. This examines whether changes in
+  a signal over time correspond to changes in reported cases at the same
+  location.
 \end{enumerate}
@@
 period, and at which all indicators are available (956 counties in total). The
-significantly positive correlations suggest that these signals could be useful
-in hotspot detection (identifying counties with relatively high COVID activity,
-at a given time). Somewhat surprisingly, the survey-based CLI-in-community
-signal shows the strongest correlations for much of the time period. This
-clearly demonstrates the value of a large-scale survey such as CTIS for tracking
+large positive correlations suggest that these signals could be useful in
+hotspot detection (identifying counties with relatively high COVID activity, at
+a given time). Somewhat surprisingly, the survey-based CLI-in-community signal
+shows the strongest correlations for much of the time period. This clearly
+demonstrates the value of a large-scale survey such as CTIS for tracking
 symptoms and case trends, especially when other data is unavailable.
@@
 Figure~\ref{fig:time_wise_correlation} summarizes time-wise correlations from
-these five signals over the same time period, and for the same set of
-counties. For each signal, we display the set correlations that it achieves in
+these five signals over the same time period, and for the same set of counties.
+For each signal, we display the set of correlations that it achieves in
 histogram form (more precisely, using a kernel density estimate). All signals
@@
 correlations, in bulk, are achieved by the CHNG-COVID signal; the
-CTIS-CLI-in-community is a close second, and the CHNG-CLI signal is third.
-There are two noteworthy points:
+CTIS-CLI-in-community is a close second, and the CHNG-CLI signal is third. There
+are two noteworthy points:
 
@@
   \caption{Reported cases per day in Bexar County, Texas during the summer of
-    2020. Only July 16, 4,810 backlogged cases were reported, though they 
-    actually occurred over the preceding two weeks (this shows up as a prolonged 
-    spike in the left panel due to the 7-day trailing averaging applied to the 
-    case counts). Daily CTIS estimates of CLI-in-community showed more stable 
+    2020. On July 16, 4,810 backlogged cases were reported, though they actually
+    occurred over the preceding two weeks (this shows up as a prolonged spike in
+    the left panel due to the 7-day trailing averaging applied to the case
+    counts). Daily CTIS estimates of CLI-in-community showed more stable
     underlying trends.}
@@
 just a few. (The latter two signals are derived from the COVID Trends and Impact
-Survey; our companion paper on this survey gives more a detailed view of its
+Survey; our companion paper on this survey gives a more detailed view of its
 features and capabilities.) These signals have already supported pandemic
@@
 signals are subject to biases, such as survey sampling and nonresponse biases,
-or geographic differences in market share for medical claims data. Claims data
-tends also to be subject to biases during major holidays and other events that
-change health-seeking behavior. Characterizing these biases will be important
-for future research and operational systems that use these signals. Several data
-sources are also subject to extensive revision and backfill, which must be
-studied and modeled to enable effective real-time use of these sources in
-forecasting and nowcasting systems. The breadth and unique features of the
+geographic differences in market share for medical claims data, or biases in the
+population represented in mobile phone mobility data \cite{Coston:2021}. Claims
+data tends also to be subject to biases during major holidays and other events
+that change health-seeking behavior. Characterizing these biases will be
+important for future research and operational systems that use these signals.
+Several data sources are also subject to extensive revision and backfill, which
+must be studied and modeled to enable effective real-time use of these sources
+in forecasting and nowcasting systems. The breadth and unique features of the
 COVIDcast API will help facilitate this and other related work, which will be

### 2021-06-18 forecast/paper/paper.tex
Update authors and acks

@@
 
-% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[a,c,1]{Author One}
-\author[b,1,2]{Author Two} 
-\author[a]{Author Three}
+% Use letters for affiliations, numbers to show equal authorship (if applicable)
+% and to indicate the corresponding author
 
-\affil[a]{Affiliation One}
-\affil[b]{Affiliation Two}
-\affil[c]{Affiliation Three}
+% Writing group
+\author[a,1]{Daniel McDonald}
+\author[b,2]{Jacob Bien}
+\author[c,2]{Alden Green}
+\author[c,d,2]{Addison Hu}
+
+% Middle authors
+\author[d]{Nat DeFries}
+\author[b]{Sangwon Hyun}
+\author[c,d]{Natalia L.\ Oliveira}
+\author[e]{James Sharpnack}
+\author[f]{Jingjing Tan}
+\author[g]{Robert Tibshirani}
+\author[c]{Valerie Ventura}
+\author[c,d]{Larry Wasserman}
+
+% Senior authors
+\author[c,d]{Ryan J.\ Tibshirani}
+
+\affil[a]{Department of Statistics, University of British Columbia}
+\affil[b]{Data Sciences and Operations, University of Southern California}
+\affil[c]{Department of Statistics \& Data Science, Carnegie Mellon University}
+\affil[d]{Machine Learning Department, Carnegie Mellon University}
+\affil[e]{Department of Statistics, University of California, Davis}
+\affil[f]{Computational Biology Department, Carnegie Mellon University}
+\affil[g]{Department of Statistics, Stanford University}
 
@@
 \authordeclaration{Please declare any competing interests here.}
-\equalauthors{\textsuperscript{1}A.O.(Author One) contributed equally to this
-  work with A.T. (Author Two) (remove if not applicable).} 
-\correspondingauthor{\textsuperscript{2}To whom correspondence should be
-  addressed. E-mail: author.two\@email.com}
+\correspondingauthor{\textsuperscript{1}To whom correspondence should be
+  addressed. E-mail: daniel@stat.ubc.ca} 
+\equalauthors{\textsuperscript{2}J.B., A.G., and A.H.\ contributed equally to
+  this work.}
 
@@
 prediction. Whether a more sophisticated model would be able to leverage the
-indicators in such a way as to change some of the finer conclusions---e.g., by
-offering clear improvements in periods in which cases are trending up---is less
+indicators in such a way as to change some of the finer conclusions (e.g., by 
+offering clear improvements in periods in which cases are trending up) is less 
 clear to us.
@@
 
-\showmatmethods{} % Display the Materials and Methods section
-
-\acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}
+\acknow{We thank Matthew Biggerstaff, Logan Brooks, Johannes Bracher, Michael 
+  Johansson, Evan Ray, Nicholas Reich, and Roni Rosenfeld for several
+  enlightening conversations about forecast evaluation. The research was
+  supported by Centers for Disease Control and Prevention (CDC) grant
+  U01IP001121, and gifts from Facebook and Google.}
 

### 2021-06-18 indicators/paper/paper.tex
Update authors and acks

--- 2021-06-18 indicators/paper/paper.tex
@@
 \author[b,2]{Logan Brooks}
-\author[a,2]{Maria Jahja}
+\author[a,b,2]{Maria Jahja}
 \author[b,2]{Aaron Rumack}
@@
 \author[b]{Roni Rosenfeld}
-\author[a]{Ryan Tibshirani}
+\author[a,b]{Ryan J.\ Tibshirani}
 
@@
 \affil[b]{Machine Learning Department, Carnegie Mellon University}
-\affil[c]{Department of Computational Biology, Carnegie Mellon University}
+\affil[c]{Computational Biology Department, Carnegie Mellon University}
 \affil[d]{Remaining affiliations to be added}
@@
 \authordeclaration{Please declare any competing interests here.}
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: areinhar@stat.cmu.edu}
+\correspondingauthor{\textsuperscript{1}To whom correspondence should be
+  addressed. E-mail: areinhar@stat.cmu.edu} 
 \equalauthors{\textsuperscript{2}L.B., M.J., A.R., and J.T. contributed equally
@@
 
-% \matmethods{Please describe your materials and methods here. This can be more than one paragraph, and may contain subsections and equations as required.
-
-% \subsection*{Subsection for Method}
-% Example text for subsection.
-% }
-
-% \showmatmethods{} % Display the Materials and Methods section
-
-\acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}
+\acknow{We thank TODO.
+  This research was supported by Centers for Disease Control and Prevention
+  (CDC) grant U01IP001121, and gifts from Facebook, Google, McCune, and
+  Optum.}
 

### 2021-06-18 forecast/paper/paper.tex
Update acks

@@
   supported by Centers for Disease Control and Prevention (CDC) grant
-  U01IP001121, and gifts from Facebook and Google.}
+  U01IP001121, gifts from Facebook, Google, McCune, and Optum, and National
+  Science Foundation Graduate Research Fellowship Program (NSF GRFP) award
+  DGE175016.}
 

### 2021-06-18 forecast/paper/pnas-suppl-template.tex
rebuild leadingness; reflow supplement for ease of edits

@@
 
-\title{Your main manuscript title}
-\author{Author1, Author2 and Author3 (complete author list)}
-\correspondingauthor{Corresponding Author name.\\E-mail: author.two@email.com}
+\title{Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot  
+  Prediction?}
+
+\author{Daniel J.\ McDonald, Jacob Bien, Alden Green, Addison Hu, Nat DeFries, Sangwon Hyun, Natalia L.\ Oliveira, James Sharpnack, Jingjing Tan, Robert Tibshirani, Valerie Ventura, Larry Wasserman, and Ryan J.\ Tibshirani}
+\correspondingauthor{Daniel J. McDonald.\\E-mail: daniel@stat.ubc.ca}
 
 \begin{document}
+
+\maketitle
+
+% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
+\SItext
 

### 2021-06-18 forecast/paper/supplement-text.tex
rebuild leadingness; reflow supplement for ease of edits

--- 2021-06-18 forecast/paper/supplement.Rmd
@@
----
-bibliography: [../../common/covidcast.bib, pnas-materials/pnas-sample.bib]
-output: 
-  pdf_document:
-    keep_tex: true
-    template: pnas-suppl-template.tex
-params:
-  fd_casefloor: 30
-  flag_jumps: 15
----
-
-<!-- PNAS materials -->
-
-<!-- Comment out or remove this line before generating final copy for submission; this will also remove the warning re: "Consecutive odd pages found". -->
-\instructionspage
-
-<!-- we need this -->
-\maketitle
-
-<!-- Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text. -->
-
-\SItext
-
-All our text goes here. Reference figures with the R chuck label as Figure~\ref{fig:fcast-finalized}.
-
-All figures go on their own page after all of the text...
-
-Edit `pnas-suppl-template.tex` for the correct Author list and Title.
-
-# Examining the relative advantage of using finalized rather than vintage data
+\section{Examining the relative advantage of using finalized rather than vintage data}
 
@@
 not properly accounting for the question of "what was known when" in
-performing retrospective evaluations of forecasters.   Figures \ref{fig:fcast-finalized} and \ref{fig:hot-finalized} show
+performing retrospective evaluations of forecasters.   Figures
+\ref{fig:fcast-finalized} and \ref{fig:hot-finalized} show 
 what Figures 3 and 4 in the main paper would have looked like if we
@@
 using vintage data.  This comparison can be seen most
-straightforwardly in Figures \ref{fig:fcast-honest-v-finalized} and \ref{fig:hot-honest-v-finalized}, which show the ratio in performance between the vintage and finalized
+straightforwardly in Figures \ref{fig:fcast-honest-v-finalized} and
+\ref{fig:hot-honest-v-finalized}, which show the ratio in performance between
+the vintage and finalized 
 versions.  When methods are 
-given the finalized version of the data rather than the version available at the time that the forecast would have been made, all methods
+given the finalized version of the data rather than the version available at the
+time that the forecast would have been made, all methods 
 appear (misleadingly) to have better performance than they would
@@
 
-# Aggregating with geometric mean
+\section{Aggregating with geometric mean}
 
@@
 
-# Bootstrap results
+\section{Bootstrap results}
 
@@
 
-<!--
+% # Correlations with lagged actuals
 
-# Correlations with lagged actuals
+% Alden's histograms are in Figure~\ref{fig:cor-wis-ratio} and Figure~\ref{fig:cor-wis-ratio-m1}.
 
-Alden's histograms are in Figure~\ref{fig:cor-wis-ratio} and Figure~\ref{fig:cor-wis-ratio-m1}.
-
--->
-
-# Upswings and Downswings
+\section{Upswings and Downswings}
 
@@
 on upswings. Figure \ref{fig:upswing-corr-table} shows the correlation between
-between $\WIS(AR) - \WIS(F)$ and the difference of their median forecasts. During
-downswings, this correlation is large, implying that improved relative performance
+between $\WIS(AR) - \WIS(F)$ and the difference of their median forecasts.
+During 
+downswings, this correlation is large, implying that improved relative
+performance 
 of $F$ is related to making lower forecasts than the AR model. The opposite is
 true during upswings. This is largely to be expected. However, the relationship
-attenuates in flat periods and during upswings. That is, when performance is better
+attenuates in flat periods and during upswings. That is, when performance is
+better 
 in those cases, it may be due to other factors than simply making predictions
@@
 
-# Leadingness and laggingness
+\section{Leadingness and laggingness}
 
@@
 
-# Examining data in 2021
+\section{Examining data in 2021}
 
@@
 
-# Deprecated
+\section{Deprecated}
 
@@
 old trajectory plots) that we can remove once we decide.
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, fig.width = 6.5, fig.height = 4.5,
-                      fig.align = "center",
-                      fig.path = "fig/",
-                      cache = TRUE, 
-                      out.width = "\\textwidth",
-                      autodep = TRUE,
-                      message = FALSE, 
-                      warning = FALSE)
-```
-
-```{r source-funs, cache=FALSE}
-library(scales)
-library(cowplot)
-library(covidcast)
-library(tidyverse)
-library(lubridate)
-library(evalcast)
-
-source("../code/eval_funs.R")
-source("../code/hotspot_funs.R")
-
-path_to_data <- "~/Documents/COVID-Delphi/fcast-indicators-paper/temp_data/"
-```
-
-```{r data-loading}
-actuals <- readRDS(file.path(path_to_data, "actuals.RDS"))
-fcasts_honest <- readRDS(file.path(path_to_data, "results_honest.RDS")) %>%
-  process_res_cases(actuals) %>%
-  filter(forecaster != "gs_inherit") 
-fcasts_finalized <- readRDS(file.path(path_to_data, "results_dishonest.RDS")) %>%
-  process_res_cases(actuals) %>%
-  filter(forecaster != "gs_inherit")
-hotspots_honest <- readRDS(file.path(path_to_data, "hotspots_honest.RDS")) %>%
-  process_res_hotspots() %>%
-  filter(forecaster != "gs_inherit")
-hotspots_finalized <- readRDS(
-  file.path(path_to_data, "hotspots_dishonest.RDS")) %>%
-  process_res_hotspots() %>%
-  filter(forecaster != "gs_inherit")
-cases7dav <- readRDS(file.path(path_to_data, "cases7dav.RDS")) %>%
-  rename(case_num = value)
-if (params$fd_casefloor >= 0) {
-  fcasts_honest <- filter_case_floor(fcasts_honest, cases7dav, params)
-  fcasts_finalized <- filter_case_floor(fcasts_finalized, cases7dav, params)
-}
-if (params$flag_jumps > 0) {
-  fcasts_honest <- filter_jumps(fcasts_honest, actuals, params)
-  fcasts_finalized <- filter_jumps(fcasts_finalized, actuals, params)
-  hotspots_honest <- filter_jumps(hotspots_honest, actuals, params)
-  hotspots_finalized <- filter_jumps(hotspots_finalized, actuals, params)
-}
-```
-
-```{r gs-processing}
-# only honest
-fcasts_gs <- intersect_averagers(
-  fcasts_honest, "forecaster", 
-  c("geo_value", "ahead", "forecast_date", "target_end_date")) %>%
-  filter(forecaster != "gs_subset") %>%
-  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
-fcasts_honest <- filter(fcasts_honest, forecaster != "gs_subset") %>%
-  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
-fcasts_finalized <- filter(fcasts_finalized, forecaster != "gs_subset")%>%
-  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
-hotspots_gs <- intersect_averagers(
-  hotspots_honest %>% filter(!is.na(value), !is.na(actual)), 
-  "forecaster", 
-  c("geo_value", "ahead", "forecast_date", "target_end_date")) %>%
-  filter(forecaster != "gs_subset") %>%
-  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
-hotspots_honest <- intersect_averagers( # Deal with DV NaN's
-  hotspots_honest %>% 
-    filter(! is.na(value), !is.na(actual), forecaster != "gs_subset"), 
-  "forecaster", 
-  c("geo_value", "ahead", "forecast_date", "target_end_date")) %>%
-  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
-hotspots_finalized <- intersect_averagers( 
-  hotspots_finalized %>% 
-    filter(! is.na(value), !is.na(actual), forecaster != "gs_subset"), 
-  "forecaster", 
-  c("geo_value", "ahead", "forecast_date", "target_end_date")) %>%
-  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
-```
-
-```{r hrr-names}
-hrr_names <- read_csv(file.path(path_to_data, "Hospital_Referral_Regions.csv"), 
-                      col_types = "_ccc__") %>%
-  mutate(state = substr(hrrcity, 1, 2),
-         hrrname = paste(HRR_lbl, state)) %>%
-  select(hrrnum, hrrname)
-hrr_tab <- pull(hrr_names, hrrname)
-names(hrr_tab) <- pull(hrr_names, hrrnum)
-```
-
-<!-- All figures and tables below this line -->
-
-<!-- Finalized v. vintage -->
-
-```{r fcast, include=FALSE}
-# Figure 3 in the paper
-plotter(fcasts_honest %>% filter(period != "jm"), 
-        "wis", Mean, scaler = "strawman_wis", 
-        order_of_operations = c("aggr","scale")) +
-  ylab("Mean WIS (relative to baseline)")
-```
-
-```{r hot, include=FALSE}
-# Figure 4 in the manuscript
-plotter_hotspots(hotspots_honest %>% filter(period != "jm")) + 
-  geom_hline(yintercept = 0.5)
-```
-
-```{r fcast-finalized, fig.cap="Forecasting performance using finalized data. Compare to Figure 3 in the manuscript."}
-plotter(fcasts_finalized %>% filter(period != "jm"), 
-        "wis", Mean, scaler = "strawman_wis", 
-        order_of_operations = c("aggr","scale")) +
-  ylab("Mean WIS (relative to baseline)")
-```
-
-\clearpage
-
-```{r hot-finalized, fig.cap="Hotspot prediction performance using finalized data. Compare to Figure 4 in the manuscript."}
-plotter_hotspots(hotspots_finalized %>% filter(period != "jm")) +
-  geom_hline(yintercept = 0.5)
-```
-
-\clearpage
-
-```{r fcast-honest-v-finalized, fig.cap="Relative forecast WIS with vintage compared to finalized data. Using finalized data leads to overly optimistic performance."}
-plotter(
-  left_join(
-    fcasts_honest %>% filter(period != "jm") %>% select(forecaster:wis),
-    fcasts_finalized %>% filter(period != "jm") %>% select(forecaster:wis) %>%
-      rename(finalized_wis = wis)
-    ), 
-  "wis", Mean, scaler = "finalized_wis", 
-  order_of_operations = c("aggr","scale")) +
-  geom_hline(yintercept = 1) +
-  ylab("Mean WIS (vintage / finalized)")
-
-```
-
-\clearpage
-
-```{r hot-honest-v-finalized, fig.cap="Relative AUC with vintage compared to finalized data. Using finalized data leads to overly optimistic hotspot performance."}
-left_join(
-  hotspots_honest %>%
-    group_by(forecaster, ahead) %>%
-    summarise(auc_honest = auc(value, actual)),
-  hotspots_finalized %>%
-    group_by(forecaster, ahead) %>%
-    summarise(auc_finalized = auc(value, actual))) %>%
-  ggplot(aes(ahead, auc_honest / auc_finalized, color = forecaster)) + 
-  geom_line() + 
-  geom_point() +
-  theme_bw() + 
-  scale_color_manual(values = fcast_colors, guide = guide_legend(nrow=1)) +
-  theme(legend.position = "bottom", legend.title = element_blank()) +
-  geom_hline(yintercept = 1) +
-  ylab("AUC (vintage / finalized)")
-```
-
-\clearpage
-
-<!-- Geometric mean justifications -->
-
-```{r wis-densities, fig.cap="Weighted interval score appears to more closely resemble a log-Gaussian distribution."}
-ggplot(fcasts_honest %>% filter(period != "jm"), 
-       aes(wis, fill = forecaster)) +
-  geom_density() +
-  scale_x_log10() +
-  theme_bw() +
-  xlab("WIS") +
-  scale_fill_manual(values = fcast_colors) +
-  facet_wrap(~forecaster) +
-  theme(legend.position = "none")
-```
-
-\clearpage
-
-```{r fcast-adjusted, fig.cap="Relative forecast performance using vintage data and summarizing with the more robust geometric mean."}
-plotter(fcasts_honest %>% filter(period != "jm"), 
-        "wis", GeoMean, scaler = "strawman_wis", 
-        order_of_operations = c("scale","aggr")) +
-  ylab("Geometric mean of WIS (relative to baseline)")
-```
-
-\clearpage
-
-<!-- Bootstrap stuff -->
-
-```{r bootstrap-loading}
-fcasts_booted <- readRDS(
-  file.path(path_to_data,"results_honest_bootstrapped.RDS")) %>%
-  process_res_cases(actuals) %>%
-  filter(forecaster != "gs_inherit")
-if (params$fd_casefloor >= 0) {
-  fcasts_booted <- filter_case_floor(fcasts_booted, cases7dav, params)
-}
-hotspots_booted <- readRDS(
-  file.path(path_to_data, "hotspots_honest_bootstrapped.RDS")) %>%
-  process_res_hotspots() %>%
-  filter(forecaster != "gs_inherit")
-if (params$flag_jumps > 0) {
-  hotspots_booted <- filter_jumps(hotspots_booted, actuals, params)
-  fcasts_booted <- filter_jumps(fcasts_booted, actuals, params)
-}
-fcasts_booted <- filter(fcasts_booted, forecaster != "gs_subset") %>%
-  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
-hotspots_booted <- intersect_averagers( # Deal with DV NaN's
-  hotspots_booted %>% 
-    filter(! is.na(value), !is.na(actual), forecaster != "gs_subset"), 
-  "forecaster", 
-  c("geo_value", "ahead", "forecast_date", "target_end_date")) %>%
-  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
-```
-
-```{r fcast-booted, fig.cap = "Forecast performance when indicators are replaced with samples from their empirical distribution. Performance is largely similar to the AR model."}
-plotter(fcasts_booted %>% filter(period != "jm"), 
-        "wis", Mean, scaler = "strawman_wis", 
-        order_of_operations = c("aggr","scale")) +
-  ylab("Mean WIS (relative to baseline)")
-```
-
-\clearpage
-
-```{r fcast-booted-adjusted, fig.cap = "Forecast performance as measured with the geometric mean when indicators are replaced with samples from their empirical distribution.
-Performance is largely similar to the AR model."}
-plotter(fcasts_booted %>% filter(period != "jm"), 
-        "wis", GeoMean, scaler = "strawman_wis", 
-        order_of_operations = c("scale","aggr")) +
-  ylab("Geometric mean of WIS (relative to baseline)")
-```
-
-\clearpage
-
-```{r hot-booted, fig.cap = "Hotspot prediction performance when indicators are replaced with samples from their empirical distribution. Performance is largely similar to the AR model."}
-plotter_hotspots(hotspots_booted %>% filter(period != "jm")) + 
-  geom_hline(yintercept = 0.5)
-```
-
-\clearpage
-
-<!-- Upswings vs. downswings -->
-
-```{r up-down-processing}
-preds <- readRDS(file.path(path_to_data, "predictions_honest.RDS"))
-preds <- preds %>% 
-  filter(abs(quantile - 0.5) < 1e-6) %>% 
-  select(-quantile) %>%
-  mutate(forecaster = recode(forecaster,
-                             AR3 = "AR",
-                             AR3CHCLI3 = "CHNG-CLI",
-                             AR3CHCOV3 = "CHNG-COVID",
-                             AR3DVCLI3 = "DV-CLI",
-                             AR3FBCLI3 = "CTIS-CLIIC",
-                             AR3GSSAA3_Subset = "gs_subset",
-                             AR3GSSAA3_Zero = "Google-AA",
-                             Baseline = "strawman"
-  )) %>%
-  filter(! forecaster %in% c("gs_subset", "strawman"))
-
-up_down <- actuals %>%
-  arrange(geo_value, target_end_date) %>%
-  group_by(geo_value) %>%
-  left_join(cases7dav, c("target_end_date" = "time_value", "geo_value")) %>%
-  mutate(pct_change = (actual - lag(actual,7)) / lag(actual,7),
-         udf = case_when(
-           pct_change >= .25 ~ "up",
-           pct_change <= -.25 ~ "down",
-           TRUE ~ "flat")) %>%
-  filter(case_num > 30, !is.na(pct_change)) #%>%
-  #rename(forecast_date = target_end_date)
-
-corr_df <- left_join(fcasts_honest, preds, 
-                     by = c("geo_value", "ahead", "forecaster", 
-                            "forecast_date")) %>%
-  select(-actual_fd, -starts_with("strawman"), -ae) %>%
-  rename(med = value)
-rm(preds)
-
-up_down_df <- left_join(corr_df %>% 
-                  ungroup() %>%
-                  filter(forecaster != "AR"),
-                corr_df %>% 
-                  ungroup() %>%
-                  filter(forecaster == "AR") %>%
-                  select(-forecaster, -actual_td) %>%
-                  rename(AR_wis = wis, AR_med = med)) %>%
-  inner_join(up_down, by = c("geo_value","target_end_date"))
-```
-
-```{r upswing-summary, fig.cap="Average difference between the WIS of the AR model and the WIS of the other forecasters. The indicator-assisted forecasters do best during down and flat periods."}
-# Figure 5 in the manuscript
-up_down_df_summary <- up_down_df %>%
-  filter(period != "jm") %>%
-  group_by(forecaster, udf) %>%
-  summarise(mean = Mean(AR_wis - wis))
-up_down_df_summary %>%
-  ggplot(aes(udf, mean , fill = forecaster)) +
-  geom_bar(width = 0.6, position = position_dodge(width=0.6),
-           stat = "identity") +
-  scale_fill_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
-  scale_y_continuous() +
-  theme_bw() +
-  geom_hline(yintercept = 0) +
-  ylab("Mean of WIS(AR) - WIS(F)") +
-  xlab("") +
-  theme(legend.position = "bottom", legend.title = element_blank())
-```
-
-```{r upswing-histogram, include=FALSE}
-up_down_df %>% 
-  filter(period != "jm") %>%
-  group_by(forecaster, udf) %>%
-  ggplot(aes(AR_wis - wis, fill = forecaster)) +
-  geom_histogram(bins = 100) +
-  facet_grid(udf ~ forecaster) +
-  theme_bw() +
-  ylab("Count") +
-  theme(legend.position = "none") +
-  scale_fill_manual(values = fcast_colors) +
-  scale_y_log10(breaks = c(10,1000,100000), 
-                labels = trans_format("log10", math_format(10^.x))) +
-  xlab("AR WIS - forecaster WIS") +
-  geom_vline(xintercept = 0)
-```
-
-```{r upswing-histogram-logged, eval = FALSE, fig.cap="Not sure if we want this here. Similar to Figure 5 in the manuscript but taking logs. "}
-up_down_df %>% 
-  filter(period != "jm") %>%
-  group_by(forecaster, udf) %>%
-  ggplot(aes((AR_wis + 1) / (wis + 1), fill = forecaster)) +
-  geom_histogram(bins = 100) +
-  facet_grid(udf ~ forecaster) +
-  theme_bw() +
-  ylab("Count") +
-  theme(legend.position = "none") +
-  scale_fill_manual(values = fcast_colors) +
-  scale_x_log10() +
-  scale_y_log10(breaks = c(10,1000,100000), 
-                labels = trans_format("log10", math_format(10^.x))) +
-  xlab("AR WIS - forecaster WIS") +
-  geom_vline(xintercept = 1)
-```
-
-\clearpage
-
-```{r hotspots-upswing-downswing, fig.cap="Classification and loglikelihood separated into periods of upswing, downswing, and flat cases.
-Like the analysis of the forecasting task in the main paper (see Figure 7), performance is better during down and flat periods."}
-hot_udf <- inner_join(
-  hotspots_honest %>% filter(period != "jm"),
-  up_down %>% select(geo_value, target_end_date, udf))
-
-cutoff <- mean(hot_udf %>% filter(forecaster == "AR") %>% pull(actual))
-
-con_tab <- hot_udf %>% 
-  filter(!is.na(actual)) %>%
-  mutate(pred = value > cutoff) %>%
-  group_by(forecaster, udf) %>%
-  summarise(m = mean(pred != actual)) %>%
-  ungroup()
-
-con_tab <- left_join(
-  con_tab %>% filter(forecaster != "AR"),
-  con_tab %>% filter(forecaster == "AR") %>%
-    select(-forecaster) %>%
-    rename(mar = m))
-
-llike_tab <- hot_udf %>%
-  filter(!is.na(actual)) %>%
-  mutate( # kill prob 0-1 predictions
-    value = case_when(
-      value < 1e-8 ~ 1e-8,
-      value > 1-1e-8 ~ 1-1e-8,
-      TRUE ~ value
-    ),
-    llike = (actual == 1) * log(value) + (actual == 0) * log(1 - value)) %>%
-  group_by(forecaster, udf) %>%
-  summarise(m = mean(llike)) %>%
-  ungroup()
-
-llike_tab <- left_join(
-  llike_tab %>% filter(forecaster != "AR"),
-  llike_tab %>% filter(forecaster == "AR") %>%
-    select(-forecaster) %>%
-    rename(mar = m))
-
-bind_rows(con_tab %>% mutate(err = "Classification error"), 
-          llike_tab %>% mutate(err = "Log likelihood")) %>%
-  ggplot(aes(udf, (m - mar) / abs(mar) , fill = forecaster)) +
-  geom_bar(width = 0.6, position = position_dodge(width=0.6),
-           stat = "identity") +
-  scale_fill_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
-  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
-  theme_bw() +
-  geom_hline(yintercept = 0) +
-  facet_wrap(~err) +
-  ylab("Change relative to AR") +
-  xlab("") +
-  theme(legend.position = "bottom", legend.title = element_blank())
-```
-
-\clearpage
-
-<!-- Correlations with lagged actuals -->
-
-```{r cor-with-actuals}
-comparison_df <- left_join(
-  fcasts_honest %>% filter(forecaster != "AR", period != "jm"),
-  fcasts_honest %>% filter(forecaster == "AR", period != "jm") %>%
-    select(-c(forecaster,target_end_date)) %>%
-    rename(ar3_wis = wis),
-  by = c("geo_value", "ahead", "forecast_date")) %>%
-  group_by(forecaster, forecast_date, geo_value) %>%
-  summarize(wis_ratio = mean(wis) / mean(ar3_wis), .groups = "drop")
-
-pct_chng_df <- readRDS(file.path(
-  path_to_data, 
-  "jhu-csse_confirmed_7dav_incidence_num_2021-05-18.RDS")) %>%
-  pct_change(n = 7) %>%
-  select(geo_value, time_value, pct_change) %>%
-  filter(time_value %in% unique(comparison_df$forecast_date))
-
-comparison_df <- comparison_df %>% 
-  left_join(pct_chng_df, 
-            by = c("forecast_date" = "time_value", "geo_value"))
-
-pct_chng_df <- comparison_df %>% 
-  group_by(forecaster, geo_value) %>%
-  summarize(
-    cor_wisratio_pctchange = cor(wis_ratio, pct_change, method = "spearman"),
-    cor_abs_wisratio_minus_1_pctchange = cor(
-      abs(wis_ratio - 1), pct_change,method = "spearman"), .groups = "drop")
-```
-
-```{r cor-wis-ratio, fig.cap="Histograms of the Spearman correlation between the ratio of AR to AR WIS with the percent change in smoothed case rates relative to 7 days earlier."}
-ggplot(pct_chng_df %>% 
-         group_by(forecaster) %>% 
-         mutate(median = median(cor_wisratio_pctchange)), 
-       aes(x = cor_wisratio_pctchange)) +
-  geom_histogram(aes(y = ..count.. / sum(..count..), fill = forecaster), bins = 50) +
-  scale_fill_manual(values = fcast_colors) +
-  theme_bw() +
-  theme(legend.position = "none") +
-  geom_vline(aes(xintercept = median), linetype = "dotted") + 
-  geom_vline(xintercept = 0) +
-  facet_wrap(~forecaster) + 
-  xlab("Spearman correlation") + ylab("Relative frequency") +
-  scale_y_continuous(labels = scales::label_percent(accuracy = .1))
-```
-
-\clearpage
-
-```{r cor-wis-ratio-m1, eval = FALSE, fig.cap="This is Alden's second set of histograms.
-Here we have the correlation of the absolute value of WIS ratio - 1 with the percent change in 7dav cases relative to 7 days earlier"}
-ggplot(pct_chng_df %>% 
-         group_by(forecaster) %>% 
-         mutate(median = median(cor_abs_wisratio_minus_1_pctchange)), 
-       aes(x = cor_abs_wisratio_minus_1_pctchange)) +
-  geom_histogram(aes(y = ..count.. / sum(..count..), fill = forecaster), bins = 50) +
-  scale_fill_manual(values = fcast_colors) +
-  theme_bw() +
-  theme(legend.position = "none") +
-  geom_vline(aes(xintercept = median), linetype = "dotted") + 
-  geom_vline(xintercept = 0) +
-  facet_wrap(~forecaster) + 
-  xlab("Spearman correlation") + ylab("Relative frequency") +
-  scale_y_continuous(labels = scales::label_percent(accuracy = 1))
-```
-
-\clearpage
-
-```{r upswing-corr-table, fig.cap="Correlation of the difference in WIS with the difference in median predictions for the AR model relative to the indicator-assisted forecaster.
-In down periods, improvements in forecast risk are highlycorrelated with lower median predictions.
-The opposite is true in up periods.
-This suggests, as one might expect that improved performance of the indicator-assisted model is attributable to being closer to the truth then the AR model.
-This conclusion is stronger in down periods then in up periods."}
-up_down_df %>% 
-  filter(period != "jm") %>% 
-  group_by(forecaster, udf) %>%
-  summarise(cor = cor(AR_wis - wis, AR_med - med)) %>%
-  ggplot(aes(udf, cor , fill = forecaster)) +
-  geom_bar(width = 0.6, position = position_dodge(width=0.6),
-           stat = "identity") +
-  scale_fill_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
-  theme_bw() +
-  geom_hline(yintercept = 0) +
-  theme(legend.position = "bottom", legend.title = element_blank())
-
-# up_down_df %>% 
-#   filter(period != "jm") %>% 
-#   group_by(forecaster, udf) %>%
-#   summarise(cor = cor(AR_wis - wis, AR_med - med)) %>%
-#   pivot_wider(names_from = forecaster, values_from = cor) %>%
-#   kableExtra::kbl(
-#     booktabs = TRUE, digits = 2, centering = TRUE,
-#     caption = paste("Correlation of the difference in WIS between the AR",
-#                    "model with the difference in median predictions. In down",
- #                   "periods, improvements in forecast risk are highly",
-#                    "correlated with lower median predictions. The opposite",
-#                    "is true in up periods. This suggests, as one might expect",
-#                    "that improved performance of the indicator-assisted",
-#                    "model is attributable to being closer to the truth then",
-#                    "the AR model. This conclusion is stronger in down",
-#                    "periods then in up periods."))
-```
-
-\clearpage
-
-<!-- Leadingness and laggingness -->
-
-```{r lead-lag-processing}
-# see lead-lag-analysis.R for signal downloading and processing
-lead_lag <- function(x, y, lag.max = NULL, na.len = 14) {
-  if (sum(is.na(x)) >= na.len || sum(is.na(y)) >= na.len) {
-    return(c(leading = NA, lagging = NA))
-  }
-  out <- ccf(x, y, lag.max = lag.max, plot = FALSE, na.action = na.omit)
-  clim <- qnorm((1 + .95) / 2) / sqrt(out$n.used)
-  lag_vec <- drop(out$lag)
-  cc <- drop(out$acf)
-  cc <- cc * (cc > clim)
-  return(c(leading = mean(cc[lag_vec < 0]),
-           lagging = mean(cc[lag_vec > 0])))
-}
-
-slide_lead_lag <- function(x, y, min_len = 56, lag.max = NULL, na.len = 14) {
-  n <- length(x)
-  outlist <- list()
-  idx <- 1:min_len
-  for (i in min_len:n) {
-    outlist[[i]] <- lead_lag(x[idx], y[idx], lag.max, na.len)
-    idx <- idx + 1
-  }
-  outlist <- bind_rows(outlist)
-  bind_rows(
-    tibble(leading = rep(NA, min_len-1), lagging = rep(NA, min_len -1)),
-    outlist)
-}
-
-all_sigs <- readRDS(file.path(
-  path_to_data, "all_signals_wide_as_of_2020-05-15.RDS"))
-
-all_sigs_imputed <- all_sigs %>% 
-  group_by(geo_value) %>% 
-  arrange(time_value) %>%
-  mutate(across(contains("value+0"), 
-                ~zoo::na.locf(.x, na.rm = FALSE, maxgap = 13))) %>%
-  pivot_longer(contains("value+0"))
-
-sigs_cor <- left_join(
-  all_sigs_imputed %>% filter(! str_detect(name, "jhu-csse")),
-  all_sigs_imputed %>% filter(str_detect(name, "jhu-csse")) %>%
-    select(-name) %>%
-    rename(cases = value))
-
-lead_lag_metric <- sigs_cor %>% 
-  group_by(geo_value, name) %>%
-  arrange(time_value) %>%
-  group_modify(~ {
-    bind_cols(time_value = .x$time_value, slide_lead_lag(.x$value, .x$cases))
-  })
-
-lead_lag_metric <- lead_lag_metric %>%
-  mutate(
-    forecaster = recode(
-      name,
-      `value+0:chng_smoothed_adj_outpatient_cli` = "CHNG-CLI",
-      `value+0:chng_smoothed_adj_outpatient_covid` = "CHNG-COVID",
-      `value+0:doctor-visits_smoothed_adj_cli` = "DV-CLI",
-      `value+0:fb-survey_smoothed_hh_cmnty_cli` = "CTIS-CLIIC",
-      `value+0:google-symptoms_sum_anosmia_ageusia_smoothed_search` = "Google-AA")) %>%
-  ungroup() %>%
-  select(-name)
-
-df2 <- left_join(
-  up_down_df, 
-  lead_lag_metric,
-  by = c("geo_value","forecaster","target_end_date" = "time_value"))
-```
-
-```{r leading-and-lagging, include=FALSE}
-# Figure 6 in the manuscript
-df2 %>%
-  filter(period != "jm") %>%
-  group_by(forecaster, udf) %>%
-  summarise(
-    leadingness = cor(AR_wis - wis, leading, use = "pairwise.complete.obs"),
-    laggingness = cor(AR_wis - wis, lagging, use = "pairwise.complete.obs")) %>%
-  pivot_longer(leadingness:laggingness) %>%
-  ggplot(aes(udf, value, fill = forecaster)) +
-  geom_bar(width = 0.6, position = position_dodge(width=0.6),
-           stat = "identity") +
-  geom_hline(yintercept = 0) +
-  scale_fill_manual(values = fcast_colors) +
-  facet_wrap(~name) +
-  theme_bw() +
-  ylab("Correlation") +
-  xlab("") +
-  theme(legend.position = "bottom", legend.title = element_blank())
-```
-
-```{r diff-in-lead-lag, include=FALSE}
-# Figure 7 in the manuscript
-df2 %>%
-  filter(period != "jm") %>%
-  group_by(forecaster, udf) %>%
-  summarise(cor = cor(AR_wis - wis, leading - lagging, 
-                      use = "pairwise.complete.obs")) %>%
-  ggplot(aes(udf, cor, fill = forecaster)) +
-  geom_bar(width = 0.6, position = position_dodge(width=0.6),
-           stat = "identity") +
-  geom_hline(yintercept = 0) +
-  scale_fill_manual(values = fcast_colors) +
-  theme_bw() +
-  ylab("Correlation") +
-  xlab("") +
-  theme(legend.position = "bottom", legend.title = element_blank())
-```
-
-<!-- Examining data in 2021 -->
-
-```{r fcast-alldates, fig.cap="Forecast performance over all periods. Performance largely improves for all forecasters with the inclusion of data in 2021."}
-plotter(fcasts_honest,
-        "wis", Mean, scaler = "strawman_wis", 
-        order_of_operations = c("aggr","scale")) +
-  ylab("Mean WIS (relative to baseline)")
-```
-
-\clearpage
-
-```{r fcast-alldates-adjusted, fig.cap="Forcast performance over all periods aggregaged with the geometric mean. Again, the inclusion of data in 2021 leads to improved performance."}
-plotter(fcasts_honest, 
-        "wis", GeoMean, scaler = "strawman_wis", 
-        order_of_operations = c("scale","aggr")) +
-  ylab("Geometric mean of WIS (relative to baseline)")
-```
-
-\clearpage
-
-```{r hot-alldates, fig.cap="Area under the curve for hotspot predictions including data in 2021.
-Performance degrades relative to the period in 2020.
-However, there are far fewer hotspots during this period as case rates declined in much of the country."}
-plotter_hotspots(hotspots_honest)
-```
-
-\clearpage
-
-<!-- Figures in the manuscript in their own scripts -->
-
-```{r revisions-dv-jhu, eval=FALSE}
-# Figure 1 in the manuscript
-source("../code/revisions.R")
-```
-
-```{r ny-trajectory, eval=FALSE}
-# Figure 2 in the manuscript
-source("../code/ny-trajectory.R")
-gg
-```
-
-```{r ccf-dv-finalized, eval=FALSE}
-# Figure ?? in the manuscript
-source("../code/ccf-dv-finalized.R")
-gg
-```
-
-<!-- All code below is deprecated -->
-
-```{r fcast-gs-locations, eval=FALSE}
-plotter(fcasts_gs %>% filter(period != "jm"),
-        "wis", Mean, scaler = "strawman_wis", 
-        order_of_operations = c("aggr","scale")) +
-  ylab("Mean WIS (relative to baseline)")
-```
-
-```{r fcast-gs-locations-adjusted, eval=FALSE}
-plotter(fcasts_gs %>% filter(period != "jm"), 
-        "wis", GeoMean, scaler = "strawman_wis", 
-        order_of_operations = c("scale","aggr")) +
-  ylab("Geometric mean of WIS (relative to baseline)")
-```
-
-```{r hot-gs-locations, eval=FALSE}
-plotter_hotspots(hotspots_gs %>% filter(period != "jm")) +
-  geom_hline(yintercept = 0.5)
-```
-
-```{r traj-data, eval = FALSE}
-source("../code/trajectory_plot_funs.R")
-preds <- readRDS(file.path(path_to_data, "predictions_honest.RDS"))
-traj_best <- get_trajectory_plots(fcasts_honest, preds, actuals, hrr_tab, 
-                                  "only2020", "best")
-traj_worst <- get_trajectory_plots(fcasts_honest, preds, actuals, hrr_tab,
-                                   "only2020", "worst")
-rm(preds)
-```
-
-```{r trajectory-plots, eval = FALSE}
-for (nm in names(traj_best)) print(trajectory_panel(nm, traj_best, traj_worst))
-```
-
-\clearpage
-
-<!-- PNAS says: Add this line AFTER all your figures and tables -->
-\FloatBarrier
-
-<!-- More PNAS example material -->
-<!-- Probably remove all this -->
-
-\movie{Type legend for the movie here.}
-
-\dataset{dataset_one.txt}{Type or paste legend here.}
-
-\dataset{dataset_two.txt}{Type or paste legend here. Adding longer text to show what happens, to decide on alignment and/or indentations for multi-line or paragraph captions.}

### 2021-06-18 forecast/paper/supplement.Rmd
rebuild leadingness; reflow supplement for ease of edits

@@
 
-<!-- PNAS materials -->
-
-<!-- Comment out or remove this line before generating final copy for submission; this will also remove the warning re: "Consecutive odd pages found". -->
-\instructionspage
-
-<!-- we need this -->
-\maketitle
-
-<!-- Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text. -->
-
-\SItext
-
-All our text goes here. Reference figures with the R chuck label as Figure~\ref{fig:fcast-finalized}.
-
-All figures go on their own page after all of the text...
-
-Edit `pnas-suppl-template.tex` for the correct Author list and Title.
-
-# Examining the relative advantage of using finalized rather than vintage data
-
-The goal of this section is to quantify the effect of
-not properly accounting for the question of "what was known when" in
-performing retrospective evaluations of forecasters.   Figures \ref{fig:fcast-finalized} and \ref{fig:hot-finalized} show
-what Figures 3 and 4 in the main paper would have looked like if we
-had simply trained all models using the finalized data rather than
-using vintage data.  This comparison can be seen most
-straightforwardly in Figures \ref{fig:fcast-honest-v-finalized} and \ref{fig:hot-honest-v-finalized}, which show the ratio in performance between the vintage and finalized
-versions.  When methods are
-given the finalized version of the data rather than the version available at the time that the forecast would have been made, all methods
-appear (misleadingly) to have better performance than they would
-actually have had if run prospectively.  For example, for forecasting case rates
-7-days ahead, the WIS of all methods is at least 8\% larger than what would have been
-recorded using the finalized values of the data.  This effect
-diminishes as the forecasting horizon increases, reflecting the fact
-that these forecasters rely less heavily on recent data than very
-short-term forecasters.  Crucially, some methods are
-"helped" more than others by the less scrupulous retrospective
-evaluation, underscoring the difficulty of avoiding misleading
-conclusions when performing retrospective evaluations of forecasters.
-
-The \chngcli~indicator (along with the other claims-based signals) is the most
-affected by this distinction, reflecting the latency in claims-based
-reporting.  This supports the importance of efforts to provide "nowcasts" for claims signals
-(which corresponds to a 0-ahead "forecast" of what the claims
-signal's value will be once all data has been collected). Looking at
-the \chngcli~and \dv~curves in Figure \ref{fig:fcast-finalized}, we
-can see that they perform very similarly when trained on the finalized data.  This is reassuring
-because they are in principle measuring the same thing (namely, the percentage of
-  outpatient visits that are primarily about COVID-related symptoms).
-  The substantial difference in their curves in Figure 3 of the
-  main paper must therefore reflect their having very different
-  backfill profiles.
-
-While \dv~is one of the least affected methods by using finalized
-rather than vintage values, it is one of the most affected methods for the hotspot problem.
-This is a reminder that the forecasting and hotspot problems are
-distinct problems.  For example, the hotspot problem does not measure
-the ability to distinguish between flat and downward trends.
-
-Even the \ar~model is affected by this distinction, reflecting the
-fact that the case rates themselves (i.e., the response values) are also
-subject to revision.  The forecasters based on indicators are thus
-affected both by revisions to the indicators and by revisions to the
-case rates.  In the case of the \gs~model, in which we only used finalized
-values for the \gs~indicator, the difference in performance can be
-wholly attributed to revisions of case rates.
-
-# Aggregating with geometric mean
-
-In this section, we consider using the geometric mean instead of the
-arithmetic mean when aggregating the weighted interval score (WIS) across location-time pairs.
-There are three reasons why the geometric mean may be desirable.
-\begin{enumerate}
-\item  WIS is right-skewed, being bounded below by zero and having
-  occasional very large values.   Figure \ref{fig:wis-densities} illustrates that the
-  densities appear roughly log-Gaussian.  The geometric mean is a
-  natural choice in such a context since the relative ordering of forecasters is
-  determined by the arithmetic mean of the {\em logarithm} of their WIS
-  values.
-\item In the main paper, we report the ratio of the mean WIS of a
-  forecaster to the mean WIS of the baseline forecaster. Another
-  choice could be to take the mean of the ratio of WIS values for the
-  two methods. (This latter choice would penalize a method less for
-  doing poorly where the baseline forecaster also does poorly.)
-Using instead the geometric mean makes the order of aggregation and
-scaling immaterial since the ratio of geometric means is the same as
-the geometric mean of ratios.
-\item If one imagines that a forecaster's WIS is composed of
-  multiplicative space-time effects $S_{\ell,t}$ shared across all forecasters,
-  i.e. $\WIS(F_{\ell,t,f},Y_{\ell,t})=S_{\ell,t}E_{f}$, then
-  taking the ratio of two forecasters' geometric mean WIS values will
-  effectively cancel these space-time effects.
-\end{enumerate}
-
-Figure \ref{fig:fcast-adjusted} uses the geometric
-mean for aggregation.  Comparing this with Figure 3 of the main paper,
-we see that the main conclusions are largely unchanged; however,
-\chngcli~now does appear better than \ar.  This behavior would be
-expected if \chngcli's poor performance is attributable to a
-relatively small number of large errors (as opposed to a large number
-of moderate errors).  Indeed, Figure 5 of the main paper further
-corroborates this, in which we see the heaviest left tails occuring
-for \chngcli.
-
-# Bootstrap results
-
-As explained in Section 2.B. of the main paper, a (somewhat cynical)
-hypothesis for why we see benefits in forecasting and hotspot
-prediction is that the indicators are not actually providing useful
-information but they are instead simply acting as a sort of implicit
-regularization,"  leading to shrinkage on the autoregressive
-coefficients, leading to less volatile predictions.  To investigate
-this hypothesis, we consider fitting  "noise features" that in truth
-should have zero coefficients.  Recall (from the main paper) that at each forecast date, we
-train a model on 6,426 location-time pairs.  Indicator models are
-based on six features, corresponding to the three autoregressive terms
-and the three lagged indicator values.  To form noise features
-we resample from other rows.  In particular, at each location $\ell$ and
-time $t$, we replace the triplet $(X_{\ell,t}, X_{\ell,t-7},
-X_{\ell,t-14})$ by the triplet $(X_{\ell^*,t^*}, X_{\ell^*,t^*-7},
-X_{\ell^*,t^*-14})$, where $(\ell^*,t^*)$ is a location-time pair
-sampled with replacement from the 6,426 location-time pairs.  Figures
-\ref{fig:fcast-booted}--\ref{fig:hot-booted} show the results.  No
-method exhibits a noticeable performance gain over the \ar~method,
-leading us to dismiss the implicit regularization hypothesis.
-
-\attn{Would be good for someone to double check that the bootstrap
-  procedure is correctly described.}
-
-<!--
-
-# Correlations with lagged actuals
-
-Alden's histograms are in Figure~\ref{fig:cor-wis-ratio} and Figure~\ref{fig:cor-wis-ratio-m1}.
-
--->
-
-# Upswings and Downswings
-
-In this section we provide extra details about the upswing / flat / downswing 
-analysis described in the main text. Figure \ref{fig:upswing-summary} shows
-the overall results, examining the average difference $\WIS(AR) - \WIS(F)$ in 
-period. Figure \ref{fig:hotspots-upswing-downswing} shows the same information
-for the hotspot task. On average, during downswings and flat periods, the 
-indicator-assisted models have lower classification error and higher 
-log likelihood than the AR model. For hotspots, both Google-AA and CTIS-CLIIC
-perform better than the AR model during upswings, in contrast to the forecasting
-task, where only Google-AA improves. For a related analysis, Figure 
-\ref{fig:cor-wis-ratio} shows histograms of the Spearman
-correlation (Spearman's $\rho$, a rank-based measure of association) between
-the $\WIS(F)/\WIS(AR)$ and the magnitude of the swing. Again we see that case
-rate increases are positively related to diminished performance of the 
-indicator models.
-
-One hypothesis for diminished relative performance during upswings is that 
-the AR model tends to overpredict downswings and underpredict upswings. Adding
-indicators seems to help avoid this behavior on the downswing but not as much 
-on upswings. Figure \ref{fig:upswing-corr-table} shows the correlation between
-between $\WIS(AR) - \WIS(F)$ and the difference of their median forecasts. During
-downswings, this correlation is large, implying that improved relative performance
-of $F$ is related to making lower forecasts than the AR model. The opposite is
-true during upswings. This is largely to be expected. However, the relationship
-attenuates in flat periods and during upswings. That is, when performance is better
-in those cases, it may be due to other factors than simply making predictions
-in the correct direction, for example, narrower confidence intervals.
-
-# Leadingness and laggingness
-
-Currently, both figures are in the manuscript. Probably just need text here.
-
-# Examining data in 2021
-
-In this section, we investigate the sensitivity of the results to the
-period over which we train and evaluate the models.  In the main
-paper, we end all evaluation on December 31, 2020.  Figures
-\ref{fig:fcast-alldates} -- \ref{fig:hot-alldates} show how the
-results would differ if we extended this analysis through March
-  31, 2021. Comparing Figure \ref{fig:fcast-alldates} to Figure 3 of
-the main paper, one sees that as ahead increases most methods now improve
-relative to the baseline forecaster. When compared to other methods, 
-\chngcli~appears much better than
-it had previously; however, all forecasters other than \chngcov~and
-\dv~are performing less well relative to the baseline than before.
-These changes are likely due to the differing nature of the pandemic
-in 2021, with flat and downward trends much more common than upward
-trajectories.  Indeed, the nature of the hotspot prediction problem is
-quite different in this period.  With a 21-day training window, it is
-common for there to be very few hotspots in training.
-
-# Deprecated
-
-There are a few blocks at the bottom (figures with Google symptoms only and the
-old trajectory plots) that we can remove once we decide.
+\input{supplement-text.tex}
 
@@
 
-```{r leading-and-lagging, include=FALSE}
+```{r leading-only, include=FALSE}
 # Figure 6 in the manuscript
+df2 %>%
+  filter(period != "jm") %>%
+  group_by(forecaster, udf) %>%
+  summarise(
+    leadingness = cor(AR_wis - wis, leading, use = "pairwise.complete.obs"),
+    laggingness = cor(AR_wis - wis, lagging, use = "pairwise.complete.obs")) %>%
+  ggplot(aes(udf, leadingness, fill = forecaster)) +
+  geom_bar(width = 0.6, position = position_dodge(width=0.6),
+           stat = "identity") +
+  geom_hline(yintercept = 0) +
+  scale_fill_manual(values = fcast_colors) +
+  theme_bw() +
+  ylab("Correlation") +
+  xlab("") +
+  theme(legend.position = "bottom", legend.title = element_blank())
+```
+
+```{r leading-and-lagging, eval=FALSE}
+# Deprecated
 df2 %>%
@@
   facet_wrap(~name) +
+  theme_bw() +
+  ylab("Correlation") +
+  xlab("") +
+  theme(legend.position = "bottom", legend.title = element_blank())
+```
+
+```{r lagging-only, eval=FALSE}
+# Deprecated
+df2 %>%
+  filter(period != "jm") %>%
+  group_by(forecaster, udf) %>%
+  summarise(
+    leadingness = cor(AR_wis - wis, leading, use = "pairwise.complete.obs"),
+    laggingness = cor(AR_wis - wis, lagging, use = "pairwise.complete.obs")) %>%
+  ggplot(aes(udf, laggingness, fill = forecaster)) +
+  geom_bar(width = 0.6, position = position_dodge(width=0.6),
+           stat = "identity") +
+  geom_hline(yintercept = 0) +
+  scale_fill_manual(values = fcast_colors) +
   theme_bw() +

### 2021-06-18 forecast/paper/supplement.tex
rebuild leadingness; reflow supplement for ease of edits

--- 2021-06-16 forecast/paper/supplement.tex
@@
 
-\title{Your main manuscript title}
-\author{Author1, Author2 and Author3 (complete author list)}
-\correspondingauthor{Corresponding Author name.\\E-mail: author.two@email.com}
+\title{Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot  
+  Prediction?}
+
+\author{Daniel J.\ McDonald, Jacob Bien, Alden Green, Addison Hu, Nat DeFries, Sangwon Hyun, Natalia L.\ Oliveira, James Sharpnack, Jingjing Tan, Robert Tibshirani, Valerie Ventura, Larry Wasserman, and Ryan J.\ Tibshirani}
+\correspondingauthor{Daniel J. McDonald.\\E-mail: daniel@stat.ubc.ca}
 
@@
 
-\instructionspage
-
 \maketitle
 
+% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
 \SItext
 
-All our text goes here. Reference figures with the R chuck label as
-Figure\textasciitilde{}\ref{fig:fcast-finalized}.
-
-All figures go on their own page after all of the text\ldots{}
-
-Edit \texttt{pnas-suppl-template.tex} for the correct Author list and
-Title.
-
-\hypertarget{examining-the-relative-advantage-of-using-finalized-rather-than-vintage-data}{%
-\section{Examining the relative advantage of using finalized rather than
-vintage
-data}\label{examining-the-relative-advantage-of-using-finalized-rather-than-vintage-data}}
-
-This section referes to
-Figures\textasciitilde{}\ref{fig:fcast-finalized}--\ref{fig:hot-honest-v-finalized}.
-
-\hypertarget{aggregating-with-geometric-mean}{%
-\section{Aggregating with geometric
-mean}\label{aggregating-with-geometric-mean}}
-
-\begin{itemize}
-\tightlist
-\item
-  The weighted interval score is bounded below by zero and can be very
-  large. This behavior is typical of right-skewed distributions.
-\item
-  Figure\textasciitilde{}\ref{fig:wis-densities} illustrates that the
-  densities appear log-Gaussian.
-\item
-  This suggests aggregating by the geometric mean rather than the mean
-  for comparisons. See Figure\textasciitilde{}\ref{fig:fcast-adjusted}.
-\end{itemize}
-
-\hypertarget{bootstrap-results}{%
-\section{Bootstrap results}\label{bootstrap-results}}
-
-Here we discuss how implicit regularization is not the reason for
-improved performance
-
-\hypertarget{correlations-with-lagged-actuals}{%
-\section{Correlations with lagged
-actuals}\label{correlations-with-lagged-actuals}}
-
-Alden's histograms are in Figure\textasciitilde{}\ref{fig:cor-wis-ratio}
-and Figure\textasciitilde{}\ref{fig:cor-wis-ratio-m1}.
-
-\hypertarget{upswings-and-downswings}{%
-\section{Upswings and Downswings}\label{upswings-and-downswings}}
-
-Logged version of Figure 5 in the manuscript is in
-Figure\textasciitilde{}\ref{fig:upswing-histogram-logged}.
-
-See also Table\textasciitilde{}\ref{tab:upswing-corr-table}.
-
-\hypertarget{leadingness-and-laggingness}{%
-\section{Leadingness and
-laggingness}\label{leadingness-and-laggingness}}
-
-Currently, both figures are in the manuscript. Probably just need text
-here.
-
-\hypertarget{examining-data-in-2021}{%
-\section{Examining data in 2021}\label{examining-data-in-2021}}
-
-See Figures \ref{fig:fcast-alldates} -- \ref{fig:hot-alldates}.
-
-\hypertarget{deprecated}{%
-\section{Deprecated}\label{deprecated}}
-
-There are a few blocks at the bottom (figures with Google symptoms only
-and the old trajectory plots) that we can remove once we decide.
+\input{supplement-text.tex}
 

### 2021-06-18 indicators/paper/paper.tex

@@
   signal correlate with case rates across locations?): Formally, let $X_t$ and
-  $Y_t$ be vectors of values of a particular signal and case rates, over all
-  locations, on date $t$. The geo-wise correlation at time $t$ is defined as
-  $\cor(X_t, Y_t)$. This examines whether a signal has the capability to help
-  spot locations with high case rates at any given time.
+  $Y_t$ be vectors of values of a signal and case rates, over all locations, on
+  date $t$. The geo-wise correlation at time $t$ is defined as $\cor(X_t, 
+  Y_t)$ (where here and throughout $\cor(\cdot,\cdot)$ denotes Spearman 
+  correlation). This examines whether a signal has the capability to help spot
+  locations with high case rates at any given time.
 
@@
   of the signal correlate with case rates across time?): Formally, let $X_\ell$
-  and $Y_\ell$ be vectors of a values of a particular signal and case rates,
-  over all times, at location $\ell$. The time-wise correlation at location
-  $\ell$ is defined as $\cor(X_\ell, Y_\ell)$. This examines whether changes in
-  a signal over time correspond to changes in reported cases at the same
+  and $Y_\ell$ be vectors of values of a signal and case rates, over all
+  times, at location $\ell$. The time-wise correlation at location $\ell$ is
+  defined as $\cor(X_\ell, Y_\ell)$. This examines whether changes in a signal
+  over time correspond to changes in reported cases at the same 
   location.
@@
 large positive correlations suggest that these signals could be useful in
-hotspot detection (identifying counties with relatively high COVID activity, at
-a given time). Somewhat surprisingly, the survey-based CLI-in-community signal
-shows the strongest correlations for much of the time period. This clearly
-demonstrates the value of a large-scale survey such as CTIS for tracking
-symptoms and case trends, especially when other data is unavailable.
+hotspot detection (identifying counties that have relatively high COVID
+activity, at a given time). Somewhat surprisingly, the survey-based
+CLI-in-community signal shows the strongest correlations for much of the time
+period. This clearly demonstrates the value of a large-scale survey such as CTIS
+for tracking symptoms and case trends, especially when other data is unavailable.
 
@@
 correlations, in bulk, are achieved by the CHNG-COVID signal; the
-CTIS-CLI-in-community is a close second, and the CHNG-CLI signal is third. There
-are two noteworthy points:
+CTIS-CLI-in-community signal is a close second, and the CHNG-CLI signal is
+third. There are two noteworthy points:
 
@@
 geographic differences in market share for medical claims data, or biases in the
-population represented in mobile phone mobility data \cite{Coston:2021}. Claims
-data tends also to be subject to biases during major holidays and other events
-that change health-seeking behavior. Characterizing these biases will be
-important for future research and operational systems that use these signals.
-Several data sources are also subject to extensive revision and backfill, which
-must be studied and modeled to enable effective real-time use of these sources
-in forecasting and nowcasting systems. The breadth and unique features of the
-COVIDcast API will help facilitate this and other related work, which will be
-vital to advancing pandemic modeling and preparedness.
+population represented in mobile device mobility data.
+% \cite{Coston:2021}. RJT: commenting out because if we cite only this, then I 
+% feel like we will have to provided references for the other sources of bias.
+% Since otherwise it might look like we don't think there's credible work on the
+% others. 
+Claims data tends also to be subject to biases during major national holidays
+and other events that change health-seeking behavior. Characterizing these
+biases will be important for future research and operational systems that use
+these signals.  Several data sources are also subject to extensive revision and
+backfill, which must be studied and modeled to enable effective real-time use of
+these sources in forecasting and nowcasting systems. The breadth and unique
+features of the COVIDcast API will help facilitate this and other related work,
+which will be vital to advancing pandemic modeling and preparedness.
 

### 2021-06-18 forecast/paper/supplement.Rmd
wide histogram

@@
 
-```{r upswing-histogram, include=FALSE}
+```{r upswing-histogram, include=FALSE, fig.width=12}
 up_down_df %>% 
@@
   scale_fill_manual(values = fcast_colors) +
-  scale_y_log10(breaks = c(10,1000,100000), 
+  scale_y_log10(breaks = c(1,100,10000), 
                 labels = trans_format("log10", math_format(10^.x))) +
+  coord_cartesian(ylim=c(1,2e4)) +
   xlab("AR WIS - forecaster WIS") +

### 2021-06-18 forecast/paper/supplement.Rmd
Add wide combined fig 3/4

@@
 # Figure 3 in the paper
-plotter(fcasts_honest %>% filter(period != "jm"), 
+fig3 <- plotter(fcasts_honest %>% filter(period != "jm"), 
         "wis", Mean, scaler = "strawman_wis", 
@@
   ylab("Mean WIS (relative to baseline)")
+fig3
 ```
@@
 # Figure 4 in the manuscript
-plotter_hotspots(hotspots_honest %>% filter(period != "jm")) + 
+fig4 <- plotter_hotspots(hotspots_honest %>% filter(period != "jm")) + 
   geom_hline(yintercept = 0.5)
+fig4
+```
+
+```{r fcast-hot-combo, include=FALSE, fig.width=8, fig.height=4}
+leg <- get_legend(fig3)
+comb1 <- plot_grid(fig3 + theme(legend.position = "none"),
+                   fig4 + theme(legend.position = "none"), 
+                   nrow = 1)
+plot_grid(comb1, leg, ncol = 1, rel_heights = c(1,.1))
 ```

### 2021-06-18 forecast/paper/paper.tex
rebuild leadingness; reflow supplement for ease of edits
Two column for fig 5
Attempt sig statement

@@
 % Writing group
-\author[a,1]{Daniel McDonald}
+\author[a,1]{Daniel J.\ McDonald}
 \author[b,2]{Jacob Bien}
@@
 % Please give the surname of the lead author for the running footer
-\leadauthor{Lead author last name}
+\leadauthor{McDonald}
 
 % Please add a significance statement to explain the relevance of your work
-\significancestatement{
-  TODO % For reference, below is the indicators paper one.
+\significancestatement{Forecasting is a key component in the public health
+  response to any fast-moving epidemic or pandemic.  Forecasting the future
+  spread some disease process, based on disease activity in the recent
+  past---say, using what is known as autoregressive modeling in the time series
+  literature---is a basic and widely-used method, which can be especially 
+  effective in noisy data settings.  In this paper we ask whether including
+  auxiliary indicators as additional features in such an autoregressive model
+  can be useful in COVID-19 case forecasting and hotspot prediction.  We find 
+  that auxiliary indicators based on de-identified medical insurance claims,
+  self-reported symptoms through online surveys, and COVID-related Internet
+  search activity all provide a nontrivial boost in accuracy, though their
+  utility generally depends the phase of the pandemic that we are in at
+  prediction time (upswing, downswing, or flat period).
+}
 
+  % TODO % For reference, below is the indicators paper one.
   % To study the COVID-19 pandemic, its effects on society,
@@
   % empowering both  public health research and operational decision-making. 
-}
+%}
 
@@
 
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/upswing-histogram-1.pdf}
+\begin{figure*}[t]
+  \includegraphics[width=\textwidth]{fig/upswing-histogram-1.pdf}
   \caption{Histogram of the difference in WIS for the AR model and that for each
@@
   \label{fig:up_down_flat}
-\end{figure}
+\end{figure*}
 
@@
 \begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/leading-and-lagging-1.pdf}
+  \includegraphics[width=\columnwidth]{fig/leading-only-1.pdf}
   \caption{Correlation of the difference in WIS with the ``leadingness'' of the

### 2021-06-18 indicators/paper/paper.tex
Incorporate some of Nigam's language on public health
To make our contributions to public health clearer.

@@
 Auxiliary signals (outside of the standard public health reporting streams) can
-serve as indicators of COVID activity, but they can also illuminate other
-aspects of the pandemic. To understand the effect of the pandemic on society and
-predict its future spread requires a comprehensive view of not just cases and
-deaths, but public behavior and beliefs. Measures of mobility reflect social
-distancing policies and the potential rate of COVID spread; medical claims data
-reflects healthcare-seeking behavior and how it may be affected by local case
-rates; measures of COVID vaccine acceptance can guide outreach efforts.
+serve as indicators of COVID activity, but they can illustrate the effect of
+mitigating actions (such as shelter-in-place orders) and can guide resource
+allocation for fighting the pandemic. For example, medical claims data reflects
+healthcare-seeking behavior; measures of mobility reflect adherence to public
+health recommendations; and measures of COVID vaccine acceptance can guide
+outreach efforts.
 
@@
 format, with comprehensive revision tracking, is a valuable goal in and of
-itself. No one data source is perfect; each is subject to error. Bringing
-numerous signals from different data streams all together, and making access as
-convenient as possible for public health researchers, policy makers, modelers,
-etc., is a strong testament to our belief that ``the whole is greater than the
-sum of its parts''.
+itself. The ability to combine public health reporting data, syndromic
+surveillance data, and digital surveillance measures of mobility and behaviors
+creates a platform that goes beyond providing situational awareness. Providing
+convenient and real-time access to this data provides continuous telemetry
+summarizing how things are, how they are expected to change, which areas need
+additional resources to be allocated in response, and how effective public
+communication is.
 

### 2021-06-18 forecast/paper/paper.tex
fixes to figure sizes / margins

--- 2021-06-18 forecast/paper/paper.tex
@@
 % Please give the surname of the lead author for the running footer
-\leadauthor{Lead author last name}
+\leadauthor{McDonald}
 
@@
 
-Figure~\ref{fig:forecast} displays the results for forecasting, stratified by
+Figure~\ref{fig:forecast} (right panel) displays the results for forecasting, stratified by
 ahead value and averaged over all HRRs and forecast dates (June 9--December 31,
@@
 
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/fcast-1.pdf}
-  \caption{Average WIS for each forecast model, over all forecast dates and all
-    HRRs, divided by the average WIS achieved by a baseline model (a
-    probabilistic version of the flat-line forecaster).  Here and in all figures
-  we abbreviate CTIS-CLI-in-community by CTIS-CLIIC.}  
-  \label{fig:forecast}
-\end{figure}
+% \begin{figure}[t]
+%   \includegraphics[width=\columnwidth]{fig/fcast-1.pdf}
+%   \caption{Average WIS for each forecast model, over all forecast dates and all
+%     HRRs, divided by the average WIS achieved by a baseline model (a
+%     probabilistic version of the flat-line forecaster).  Here and in all figures
+%   we abbreviate CTIS-CLI-in-community by CTIS-CLIIC.}  
+%   \label{fig:forecast}
+% \end{figure}
 
-Figure~\ref{fig:hotspot} displays the results for hotspot prediction, again
+Figure~\ref{fig:forecast} (left panel) displays the results for hotspot prediction, again
 stratified by ahead value and averaged over all HRRs and prediction dates (June 
@@
 function can absorb a lot of the variability in its input).
+
+\begin{figure*}[t]
+  \includegraphics[width=\textwidth]{fig/fcast-hot-combo-1.pdf}
+  \caption{Left: Average WIS for each forecast model, over all forecast dates and all
+    HRRs, divided by the average WIS achieved by a baseline model (a
+    probabilistic version of the flat-line forecaster).  Here and in all figures
+    we abbreviate CTIS-CLI-in-community by CTIS-CLIIC.
+    Right: Area under the curve for each hotspot prediction model, calculated over
+    all prediction dates and all HRRs.
+  }  
+  \label{fig:forecast}
+\end{figure*}
 
@@
 
-\begin{figure}[t]
-\includegraphics[width=\columnwidth]{fig/hot-1.pdf}
-\caption{Area under the curve for each hotspot prediction model, calculated over
-  all prediction dates and all HRRs.} 
-\label{fig:hotspot}
-\end{figure}
+% \begin{figure}[t]
+% \includegraphics[width=\columnwidth]{fig/hot-1.pdf}
+% \caption{Area under the curve for each hotspot prediction model, calculated over
+%   all prediction dates and all HRRs.} 
+% \label{fig:hotspot}
+% \end{figure}
 

### 2021-06-18 forecast/paper/supplement.Rmd
fixes to figure sizes / margins

@@
 
-```{r upswing-histogram, include=FALSE, fig.width=12}
+```{r upswing-histogram, include=FALSE, fig.width=8, fig.height=4}
 up_down_df %>% 
@@
 
-```{r leading-only, include=FALSE}
+```{r leading-only, include=FALSE, fig.width=6, fig.height=4}
 # Figure 6 in the manuscript
@@
   scale_fill_manual(values = fcast_colors) +
-  theme_bw() +
+  guides(fill = guide_legend(nrow = 1)) +
+  theme_bw(base_size = 16) +
   ylab("Correlation") +
   xlab("") +
-  theme(legend.position = "bottom", legend.title = element_blank())
+  theme(legend.position = c(1,-.15), 
+        legend.justification = "right",
+        legend.title = element_blank(), 
+        legend.text = element_text(size = 10),
+        plot.margin = margin(b = 20, l = 4),
+        legend.background = element_rect(fill = "transparent"))
 ```

### 2021-06-18 forecast/code/revisions.R

@@ -23,19 +23,22 @@ dv_as_of <-  map_dfr(as_ofs, function(as_of) {
 
   
 bind_rows(cases_as_of, dv_as_of) %>% 
-  mutate(as_of = fct_relabel(factor(as_of), function(x) strftime(x, "%b %d")),
-         data_source = recode(data_source, `doctor-visits` = "DV-CLI",
-                              `jhu-csse` = "JHU-CSSE")) %>%
+  mutate(as_of = fct_relabel(factor(as_of), function(x) strftime(x, "%b %d"))) %>%
   ggplot(aes(x = time_value, y = value)) + 
   geom_line(aes(color = factor(as_of))) + 
   facet_wrap(~data_source, scales = "free_y",
              strip.position = "left", 
-             labeller = as_labeller(c(`DV-CLI` = "% doctor's visits due to CLI",
-                                      `JHU-CSSE` = "Cases per 100,000 people"))) +
+             labeller = as_labeller(
+               c(`doctor-visits` = "% doctor's visits due to CLI",
+                 `jhu-csse` = "Cases per 100,000 people"))) +
   theme_bw() +
   labs(x = "", y = "", color = "As of:") +
-  theme(legend.position = "bottom") +
-  scale_color_viridis_d(end = .9, begin = .1)
+  theme(legend.position = c(.5, -.15),
+        legend.direction = "horizontal",
+        plot.margin = margin(b = 20),
+        legend.background = element_rect(fill = "transparent")) +
+  scale_color_viridis_d(end = .9, begin = .1) +
+  guides(color = guide_legend(nrow = 1))
 
 ggsave("../paper/fig/revisions.pdf", width = 5, height = 4)
 

### 2021-06-18 forecast/paper/paper.tex

@@
 % Please add a significance statement to explain the relevance of your work
-\significancestatement{
-  TODO % For reference, below is the indicators paper one.
+\significancestatement{Forecasting is a key component in the public health
+  response to any fast-moving epidemic or pandemic.  Forecasting the future
+  spread some disease process, based on disease activity in the recent
+  past---say, using what is known as autoregressive modeling in the time series
+  literature---is a basic and widely-used method, which can be especially 
+  effective in noisy data settings.  In this paper we ask whether including
+  auxiliary indicators as additional features in such an autoregressive model
+  can be useful in COVID-19 case forecasting and hotspot prediction.  We find 
+  that auxiliary indicators based on de-identified medical insurance claims,
+  self-reported symptoms through online surveys, and COVID-related Internet
+  search activity all provide a nontrivial boost in accuracy, though their
+  utility generally depends the phase of the pandemic that we are in at
+  prediction time (upswing, downswing, or flat period).
+}
 
+  % TODO % For reference, below is the indicators paper one.
   % To study the COVID-19 pandemic, its effects on society,

### 2021-06-18 forecast/paper/supplement-text.tex
fix quotation marks

@@
 prediction is that the indicators are not actually providing useful
-information but they are instead simply acting as a sort of implicit
-regularization,"  leading to shrinkage on the autoregressive
+information but they are instead simply acting as a sort of ``implicit
+regularization,''  leading to shrinkage on the autoregressive
 coefficients, leading to less volatile predictions.  To investigate
-this hypothesis, we consider fitting  "noise features" that in truth
+this hypothesis, we consider fitting  ``noise features'' that in truth
 should have zero coefficients.  Recall (from the main paper) that at each forecast date, we

### 2021-06-18 forecast/code/revisions.R
Adjust fig sizes

@@ -40,5 +40,5 @@ bind_rows(cases_as_of, dv_as_of) %>%
   scale_color_viridis_d(end = .9, begin = .1) +
   guides(color = guide_legend(nrow = 1))
 
-ggsave("../paper/fig/revisions.pdf", width = 5, height = 4)
+ggsave("../paper/fig/revisions.pdf", width = 5, height = 3.5)
 

### 2021-06-18 forecast/code/trajectory.R

@@ -28,4 +28,4 @@ ggplot(pd$truth_df, mapping = aes(x = target_end_date)) +
   ylab("Cases per 100,000 people") + 
   xlab("Date")
 
-ggsave("../paper/fig/trajectory.pdf", width = 6.5, height = 4.5)
+ggsave("../paper/fig/trajectory.pdf", width = 6, height = 4)

### 2021-06-18 forecast/paper/paper.tex
fix author name typo
Update sig statement

--- 2021-06-18 forecast/paper/paper.tex
@@
 % Please add a significance statement to explain the relevance of your work
-\significancestatement{
-  TODO % For reference, below is the indicators paper one.
+\significancestatement{Forecasting is a critical element of the public health
+  response to any fast-moving epidemic or pandemic.  Predicting the future
+  spread some disease process, based on disease activity in the recent
+  past---say, using what is known as autoregressive modeling in the time series
+  literature---is a basic and widely-used method.  In this paper we study
+  whether including auxiliary indicators as additional features in such an
+  autoregressive model can be useful in COVID-19 case forecasting and hotspot
+  prediction.  We find that indicators based on de-identified medical insurance 
+  claims, self-reported symptoms through online surveys, and COVID-related
+  Internet search activity all provide a nontrivial boost in accuracy, though
+  their utility generally depends the pandemic's local dynamics (upswing,
+  downswing, or flat period) at prediction time.
+}
 
+  % TODO % For reference, below is the indicators paper one.
   % To study the COVID-19 pandemic, its effects on society,
@@
 
-Figure~\ref{fig:forecast} (right panel) displays the results for forecasting, stratified by
-ahead value and averaged over all HRRs and forecast dates (June 9--December 31,
-2020).  Shown is the average WIS for each forecast model divided by the averaged
-WIS from a baseline model, which is basically a flat-line forecaster (its
-median forecast for $Y_{\ell,t+a}$ is always $Y_{\ell,t}$, with predicted
-quantiles defined around this flat line based on historical variation). This is
-the same baseline model as that used in the COVID-19 Forecast Hub.  Here, we  
-divide by the average WIS of the baseline model in order to put the y-axis on an 
-interpretable scale.  We can see that all curves are below 1, which means
-(smaller WIS is better) that all of the models, including AR, outperform the
-baseline on average over the forecasting period.  On the other hand, the models 
-deliver at best an improvement of about 20\% in mean WIS over the baseline
-model, with this gap narrowing to only 10\% at the largest ahead values,
-illustrating the difficulty of the forecasting problem.
+Figure~\ref{fig:forecast} (left panel) displays evaluation results for
+forecasting, stratified by ahead value and averaged over all HRRs and forecast
+dates.  Shown is the average WIS for each forecast model divided by that from a
+baseline model, which is basically a flat-line forecaster (its median forecast
+for $Y_{\ell,t+a}$ is always $Y_{\ell,t}$, with predicted quantiles defined
+around this based on historical variation). This is the same baseline model as
+in the COVID-19 Forecast Hub. Here, we use the baseline model in order to scale
+mean WIS so that it gets put on an interpretable, unitless scale.  In the
+figure, we can see that all curves are below 1, which means (smaller WIS is
+better) that all of the models, including AR, outperform the baseline on average
+over the forecasting period.  On the other hand, the models deliver at best an
+improvement of about 20\% in average WIS over the baseline model, with this gap
+narrowing to about 10\% at the largest ahead values, illustrating the difficulty 
+of the forecasting problem.
 
@@
 
-% \begin{figure}[t]
-%   \includegraphics[width=\columnwidth]{fig/fcast-1.pdf}
-%   \caption{Average WIS for each forecast model, over all forecast dates and all
-%     HRRs, divided by the average WIS achieved by a baseline model (a
-%     probabilistic version of the flat-line forecaster).  Here and in all figures
-%   we abbreviate CTIS-CLI-in-community by CTIS-CLIIC.}  
-%   \label{fig:forecast}
-% \end{figure}
-
-Figure~\ref{fig:forecast} (left panel) displays the results for hotspot prediction, again
-stratified by ahead value and averaged over all HRRs and prediction dates (June 
-16--December 31, 2020).  We can see many similarities to the forecasting
-results (now larger AUC is better); for example, CHNG-COVID and DV-CLI offer the
-biggest improvement over AR, and all models including AR degrade in performance 
-towards the baseline (this is now just a classifier based on random guessing,
-which achieves an AUC of 0.5) as the ahead values grow.  A clear difference,
-however, is that the CHNG-CLI model performs quite well in hotspot prediction,
-close to the best-performing indicator models for many of the ahead values. This
-may be because volatility in the CHNG-CLI indicator plays less of a role in the
-associated logistic model's predicted probabilities (in general, a sigmoid
-function can absorb a lot of the variability in its input).
+Figure~\ref{fig:forecast} (right panel) displays the results for hotspot
+prediction, again stratified by ahead value and averaged over all HRRs and
+prediction dates.  We can see many similarities to the forecasting results (now
+larger AUC is better); for example, CHNG-COVID and DV-CLI offer the biggest
+improvement over AR, and all models including AR degrade in performance towards
+the baseline (this is now just a classifier based on random guessing, which
+achieves an AUC of 0.5) as the ahead values grow, illustrating the difficulty of
+the hotspot prediction problem A clear difference, however, is that the CHNG-CLI
+model performs quite well in hotspot prediction, close to the best-performing
+indicator models for many of the ahead values. This may be because volatility in
+the CHNG-CLI indicator plays less of a role in the associated logistic model's
+predicted probabilities (in general, a sigmoid function can absorb a lot of the
+variability in its input).
 
@@
   \includegraphics[width=\textwidth]{fig/fcast-hot-combo-1.pdf}
-  \caption{Left: Average WIS for each forecast model, over all forecast dates and all
-    HRRs, divided by the average WIS achieved by a baseline model (a
-    probabilistic version of the flat-line forecaster).  Here and in all figures
-    we abbreviate CTIS-CLI-in-community by CTIS-CLIIC.
-    Right: Area under the curve for each hotspot prediction model, calculated over
-    all prediction dates and all HRRs.
+  \caption{Left: average WIS for each forecast model, over all forecast dates
+    and all HRRs, divided by the average WIS achieved by a baseline model (a
+    probabilistic version of the flat-line forecaster).  Right: area under the
+    curve for each hotspot prediction model, calculated over all prediction
+    dates and all HRRs.  Here and in all figures we abbreviate
+    CTIS-CLI-in-community by CTIS-CLIIC.  
   }  

### 2021-06-18 forecast/paper/supplement.tex

@@
 
-\author{Daniel J.\ McDonald, Jacob Bien, Alden Green, Addison Hu, Nat DeFries, Sangwon Hyun, Natalia L.\ Oliveira, James Sharpnack, Jingjing Tan, Robert Tibshirani, Valerie Ventura, Larry Wasserman, and Ryan J.\ Tibshirani}
+\author{Daniel J.\ McDonald, Jacob Bien, Alden Green, Addison Hu, Nat DeFries, Sangwon Hyun, Natalia L.\ Oliveira, James Sharpnack, Jingjing Tang, Robert Tibshirani, Valerie Ventura, Larry Wasserman, and Ryan J.\ Tibshirani}
 \correspondingauthor{Daniel J. McDonald.\\E-mail: daniel@stat.ubc.ca}

### 2021-06-18 forecast/paper/supplement.Rmd
incorporate author name typo; resize figs

--- 2021-06-18 forecast/paper/supplement.Rmd
@@
 
-```{r fcast-hot-combo, include=FALSE, fig.width=8, fig.height=4}
+```{r fcast-hot-combo, include=FALSE, fig.width=9, fig.height=4.5}
 leg <- get_legend(fig3)
@@
            pct_change >= .25 ~ "up",
-           pct_change <= -.25 ~ "down",
+           pct_change <= -.2 ~ "down",
            TRUE ~ "flat")) %>%
@@
 
-```{r upswing-histogram, include=FALSE, fig.width=12}
+```{r upswing-histogram, include=FALSE, fig.width=9, fig.height=4.5}
+# Figure 4
 up_down_df %>% 
@@
   scale_fill_manual(values = fcast_colors) +
+  guides(fill = guide_legend(nrow = 1)) +
   theme_bw() +
@@
   xlab("") +
-  theme(legend.position = "bottom", legend.title = element_blank())
+  theme(legend.position = c(1,-.15), 
+        legend.justification = "right",
+        legend.title = element_blank(), 
+        legend.text = element_text(size = 10),
+        plot.margin = margin(b = 20, l = 4),
+        legend.background = element_rect(fill = "transparent"))
 ```

### 2021-06-18 forecast/paper/paper.tex
Down is now 20\%

@@
 \author[e]{James Sharpnack}
-\author[f]{Jingjing Tan}
+\author[f]{Jingjing Tang}
 \author[g]{Robert Tibshirani}
@@
 simply use the opposite of the hotspot definition: $Y_{\ell,t}$ decreases by
-more than 25\% relative to the preceding week.
+more than 20\% relative to the preceding week.
 

### 2021-06-18 indicators/paper/paper.tex
Copyedits to discussion

@@
 Auxiliary signals (outside of the standard public health reporting streams) can
-serve as indicators of COVID activity, but they can illustrate the effect of
-mitigating actions (such as shelter-in-place orders) and can guide resource
+serve as indicators of COVID activity, but they can also illustrate the effect
+of mitigating actions (such as shelter-in-place orders) and can guide resource 
 allocation for fighting the pandemic. For example, medical claims data reflects
@@
 Finally, we emphasize that unifying many relevant signals into a single common
-format, with comprehensive revision tracking, is a valuable goal in and of
+format, with comprehensive revision tracking, is an important goal in and of 
 itself. The ability to combine public health reporting data, syndromic
-surveillance data, and digital surveillance measures of mobility and behaviors
-creates a platform that goes beyond providing situational awareness. Providing
-convenient and real-time access to this data provides continuous telemetry
-summarizing how things are, how they are expected to change, which areas need
-additional resources to be allocated in response, and how effective public
-communication is.
+surveillance data, and digital measures of mobility and behaviors goes beyond
+providing traditional situational awareness.  Convenient and real-time access to
+this data enables continuous telemetry summarizing how things are, how they are
+expected to change, which areas need additional resources to be allocated in
+response, and how effective public communication is.
 

### 2021-06-18 forecast/paper/paper.tex
abstract (241 words)

@@
 \begin{abstract}
-Please provide an abstract of no more than 250 words in a single paragraph.
-Abstracts should explain to the general reader the major contributions of the
-article. References in the abstract must be cited in full within the abstract
-itself and cited in the text. 
+Reliable short-term, location-specific forecasts of a pandemic are a
+key ingredient in effective public health decision making. Through the
+COVID-19 pandemic, a plethora of real-time data signals were
+collected, curated, and made publicly available by our team and data
+partners, providing multiple views of the pandemic's progress. The
+goal of this work is to measure the practical utility of these
+auxiliary indicators from a forecasting perspective. We focus on five
+indicators, derived from medical insurance claims data, web search
+queries, and survey responses. For each indicator, we ask whether
+including it in a simple model leads to improved forecasting
+performance relative to a similar model without it. Performance is considered across two tasks
+at the state level at a forecasting horizon of up to about one month:
+Probabilistic forecasting of seven-day average case rates and binary
+prediction of ``hotspots.'' Since the values of indicators are
+commonly revised over time, we take special care to ensure that the data provided
+to a forecaster is the version that would have been available at the
+time that the forecast was made. We demonstrate how less careful
+backtesting can lead to misleading retrospective evaluations. Our analysis suggests consistent, although
+modest, gains in forecasting performance by using these indicators,
+and we investigate whether the auxiliary signals are acting as ``leading indicators'' of case rates.
+Evaluating forecasters thoroughly and fairly is an important way that
+we can learn from this pandemic to do better in the next.
+
+% Please provide an abstract of no more than 250 words in a single paragraph.
+% Abstracts should explain to the general reader the major contributions of the
+% article. References in the abstract must be cited in full within the abstract
+% itself and cited in the text.
 \end{abstract}

### 2021-06-18 forecast/paper/paper.tex
Add UBC acks

--- 2021-06-18 forecast/paper/paper.tex
@@
   supported by Centers for Disease Control and Prevention (CDC) grant
-  U01IP001121, gifts from Facebook, Google, McCune, and Optum, and National
+  U01IP001121; gifts from Facebook, Google, McCune, and Optum; the National
   Science Foundation Graduate Research Fellowship Program (NSF GRFP) award
-  DGE175016.}
+  DGE175016; the Canadian Statistical Sciences Institute; and the National
+  Sciences and Engineering Research Council of Canada (NSERC) RGPIN-2021-02618.}
 

### 2021-06-18 forecast/paper/paper.tex
Fixed a reference, proofreading up to results.

--- 2021-06-18 forecast/paper/paper.tex
@@
   response to any fast-moving epidemic or pandemic.  Predicting the future
-  spread some disease process, based on disease activity in the recent
+  spread of some disease process, based on disease activity in the recent
   past---say, using what is known as autoregressive modeling in the time series
@@
   claims, self-reported symptoms through online surveys, and COVID-related
-  Internet search activity all provide a nontrivial boost in accuracy, though
+  internet% Capitalization is apparently
+  % in question: https://en.wikipedia.org/wiki/Capitalization_of_Internet
+  search activity all provide a nontrivial boost in accuracy, though
   their utility generally depends the pandemic's local dynamics (upswing,
@@
 \begin{abstract}
-Please provide an abstract of no more than 250 words in a single paragraph.
-Abstracts should explain to the general reader the major contributions of the
-article. References in the abstract must be cited in full within the abstract
-itself and cited in the text. 
+Reliable short-term, location-specific forecasts of a pandemic are a
+key ingredient in effective public health decision making. Throughout the
+COVID-19 pandemic, a plethora of real-time data signals were
+collected, curated, and made publicly available by our team and data
+partners, providing multiple views of the pandemic's progress. The
+goal of this work is to measure the practical utility of these
+auxiliary indicators from a forecasting perspective. We focus on five
+indicators, derived from medical insurance claims data, web search
+queries, and survey responses. For each indicator, we ask whether its
+inclusion in a simple model leads to improved forecasting
+performance relative to a similar model excluding it. Performance is considered across two tasks
+at a forecasting horizon of up to about one month:
+probabilistic forecasting of seven-day average case rates and binary
+prediction of ``hotspots.'' Since the values of indicators are
+commonly revised over time, we take special care to ensure that the data provided
+to a forecaster is the version that would have been available at the
+time that the forecast was made. We demonstrate how less careful
+backtesting can lead to misleading retrospective evaluations. Our analysis
+suggests consistent, although 
+modest, gains in forecasting performance by using these indicators,
+and further investigation suggests that these gains are related to periods when the
+auxiliary signals are acting as ``leading 
+indicators'' of case rates. 
+\attn{Evaluating forecasters thoroughly and fairly is an important way that
+we can learn from this pandemic to do better in the next. 251 :(}
+
+% Please provide an abstract of no more than 250 words in a single paragraph.
+% Abstracts should explain to the general reader the major contributions of the
+% article. References in the abstract must be cited in full within the abstract
+% itself and cited in the text.
 \end{abstract}
@@
 people (the standard in epidemiology).  We use reported case data aggregated by 
-JHU CSSE \cite{Dong:2020}, which is accessible through the COVIDcast API
-\cite{CovidcastAPI}, like the auxiliary indicators that we use to supplement the
-basic purely autoregressive models.
+JHU CSSE \cite{Dong:2020}, which, like the auxiliary indicators that we use to supplement the
+basic purely autoregressive models, is accessible through the COVIDcast API
+\cite{CovidcastAPI}.
 
@@
   with confirmed COVID-19, based on the same claims data.
-\item COVID Trends and Impact Survey CLI in the community
+\item COVID Trends and Impact Survey COVID-like illness in the community
   (CTIS-CLI-in-community): The estimated percentage of the population who know
@@
 \item Google Search Trends for anosmia and ageusia (Google-AA): A measure of 
-  COVID-19 related Google search volume for queries relating to anosmia or
+  Google search volume for queries relating to anosmia or
   ageusia (loss of smell or taste), based on Google's Search Trends data set. 
@@
 on this survey. For the Google Search Trends data set, see
-\cite{GoogleSymptoms}, and \cite{Klopfen:2020, Vaira:2020} for articles on the
-relevance of anosmia or ageusia (loss of smell or taste) to COVID.
+\cite{GoogleSymptoms}, and see \cite{Klopfen:2020, Vaira:2020} for justification of the
+relevance of anosmia or ageusia to COVID-19 infection.
 
@@
 (cardiovascular or neurological) are performed. The smallest HRR has a
-population of about 120,000. While some are quite large (such as the one
-containing Los Angeles, which has about 10 million people), generally HRRs 
+population of about 125,000. While some are quite large (such as the one
+containing Los Angeles, which has more than 10 million people), generally HRRs 
 are much more homogenous in size than the (approximately) 3200 counties,  
@@
 demand.  We have chosen to focus on cases (forecasting and predicting
-hotspots) at the HRR level since the indicators considered should be more 
-useful in predicting case activity versus hospital demand, since the latter is
-intuitively more closely connected (in time) to the events that are measured by
+hotspots) at the HRR level because the indicators considered should be more 
+useful in predicting case activity rather than hospital demand. The \attn{former?} is
+intuitively more contemporaneous to the events that are measured by
 the given five indicators. Predicting case rates (and hotspots) at the HRR level
@@
 In this paper, all models are fit with ``vintage'' training data. This means
-that for a given prediction date, say, September 1, 2020, we train models 
+that for a given prediction date, say, September 28, 2020, we train models 
 using only the data that would have been available to us \textit{as of}
-September 1 (imagine that we can ``rewind'' the clock to September 1, and query
+September 28 (imagine that we can ``rewind'' the clock to September 28 and query
 the COVIDcast API to get the latest data it would have had available at that
@@
   \includegraphics[width=\columnwidth]{fig/revisions.pdf}
-  \caption{Revision behavior for two signals in the HRR containing New York City
-    HRR.  Each colored line corresponds to the data as reported on a particular
+  \caption{This figure displays revision behavior for two signals in the HRR
+    containing New York 
+    City.  Each colored line corresponds to the data as reported on a particular 
     date (\textit{as of} dates varying from September 28 through October
@@
 case rates, is 1 day latent, and exhibits frequent, unpredictable revisions
-(sometimes clear anomalies) after initial publication.  Compared to the revision
+after initial publication.  Compared to the revision
 pattern in the medical claims signals, which are much more systematic in nature,
@@
 local authority to figure out how this should best be done), which can result
-in very nontrivial and often unpredictable revisions.  See
+in very nontrivial revisions.  See
 Figure~\ref{fig:vintage} for an example.
@@
 the latency of the signal is not an unavoidable property of the data type, so we
-simply use finalized signal values it, with zero latency, in our analysis.
+simply use finalized signal values, with zero latency, in our analysis.
 
@@
 clear, this is the number of new daily reported cases per 100,000
-people, averaged over the 7-day period $t-6, \ldots t$. The first task we
+people, averaged over the 7-day period $t-6, \ldots, t$. The first task we
 consider---\textit{forecasting}---is to predict $Y_{\ell,t+a}$ for each
@@
 is robustness: averaging stabilizes the case time series, and accounts for 
-uninteresting artifacts like weekend-weekday differences in the series.  Note
-that we can of course equivalently view this (equivalent up to a constant
+uninteresting artifacts like day-of-the-week effects in the series.  Note
+that we can equivalently view this (equivalent up to a constant
 factor) as predicting the HRR-level COVID-19 case incidence rate \textit{summed}
@@
 earliest date at which the data needed to train all models was available, which
-ends up being (per our setup, with 21 days of training data, and lagged values
-of signals for features as we will detail shortly) June 9, 2020 for forecasting,
+ends up being (per our setup, with 21 days of training data and lagged values
+of signals for features, as we will detail shortly) June 9, 2020 for forecasting,
 and June 16, 2020 for hotspot prediction. (The bottleneck here is the
@@
 \label{eq:forecast_ar}
-\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s}, s \leq t)  
+\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s},\ s \leq t)  
 = \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j}.  
@@
 \label{eq:quantile_levels}
-\tau \in \{0.025, 0.1, 0.25, 0.5, 0.75, 0.9, 0.975 \}.
+\tau \in \{0.025,\ 0.1,\ 0.25,\ 0.5,\ 0.75,\ 0.9,\ 0.975 \}.
 \end{equation}
@@
 \label{eq:forecast_ar_x}
-\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s}, X_{\ell,s}, s \leq t)  
+\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s},\ X_{\ell,s},\ s \leq t)  
 = \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j} + 
@@
 
-At prediction time, in order to deal with possible crossing violations (a
-predicted quantile at level $\tau$ exceeds a predicted quantile at level $\tau'
-> \tau$), we apply a simple post-hoc sorting.  See Figure~\ref{fig:trajectory}
+At prediction time, in order to avoid crossing violations (that is,
+for two levels $\tau' > \tau$,
+the predicted quantile at level $\tau$ exceeds the predicted quantile at level
+$\tau'$), we apply a simple post-hoc sorting.  See Figure~\ref{fig:trajectory}
 for an example forecast.
@@
 \includegraphics[width=\columnwidth]{fig/trajectory.pdf}
-\caption{Forecast from an autoregressive model for the New York City HRR made 
+\caption{This figure displays a forecast from an autoregressive model for the
+  HRR containing
+  New York City made  
   on October 15 (vertical line).  The fan displays 50\%, 80\% and 95\% intervals
@@
 \label{eq:hotspot_ar}
-\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s}, s \leq t) \big) 
+\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s},\ s \leq t) \big) 
 = \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j},
@@
 \label{eq:hotspot_ar_x}
-\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s}, X_{\ell,s}, s \leq t) \big)
+\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s},\ X_{\ell,s},\ s \leq t) \big)
 = \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j} +  
@@
 For forecasting, we evaluate the probabilistic forecasts produced by the
-quantile models in \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x} using 
+quantile models in \eqref{eq:forecast_ar} and \eqref{eq:forecast_ar_x} using 
 \textit{weighted interval score} (WIS), a quantile-based scoring rule; see e.g.,
@@
 to \eqref{eq:quantile_levels}, the exclusion probabilities are $\alpha \in
-\{0.05, 0.2, 0.5, 1\}$, resulting in 4 terms in the above sum.  By 
+\{0.05,\ 0.2,\ 0.5,\ 1\}$, resulting in 4 terms in the above sum.  By 
 straightforward algebra, it is not hard to see WIS has an alternative 
@@
 presentation, because the score nicely decouples into a ``sharpness'' component
-(first term in each summand) and ``under/overprediction'' component (second term
+(first term in each summand) and an ``under/overprediction'' component (second term
 in each summand).  The second form given in \eqref{eq:wis_quantiles} is
-especially noteworthy in our current study as it reveals WIS is the same as the
+especially noteworthy in our current study because it reveals WIS to be the same as the
 quantile regression loss that we use to train our forecasting models (i.e., our
@@
 For hotspot prediction, we evaluate the probabilistic classifiers produced by
-the logistic models in \eqref{eq:hotspot_ar}, \eqref{eq:hotspot_ar_x} using the 
+the logistic models in \eqref{eq:hotspot_ar} and \eqref{eq:hotspot_ar_x} using the 
 area under the curve (AUC) of their true positive versus false positive rate
@@
 The primary aggregation scheme that we will use in model evaluation and
-comparisons will be to average WIS per forecaster per ahead value $a$, over all
+comparisons will be to average WIS per forecaster at ahead value $a$ over all
 forecast dates $t$ and locations $\ell$; and similarly, to compute AUC per
-classifier and per ahead value $a$, over all forecast dates $t$ and locations
+classifier at ahead value $a$ over all forecast dates $t$ and locations
 $\ell$.
@@
 As described previously, the auxiliary indicators defined in terms of medical
-claims, CHNG-CLI, CHNG-COVID, and DV-CLI, undergo a significant and systematic
+claims (CHNG-CLI, CHNG-COVID, and DV-CLI) undergo a significant and systematic
 pattern of revision, or ``backfill'', after their initial publication.  Given

### 2021-06-19 forecast/paper/paper.tex
Proofed up to discussion

@@
 we refer to as \textit{nowcasting}---as a pre-processing step before using them
-as features in the models in \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar_x}. 
+as features in the models in \eqref{eq:forecast_ar_x} and \eqref{eq:hotspot_ar_x}. 
 Nowcasting is itself a highly nontrivial modeling problem, and we do not attempt
@@
 
-We begin with a summary of the high-level conclusions.  Here, and in the
+Here, and in the
 following subsections, we will use ``AR'' to refer to the pure autoregressive
@@
 Finally, we use the term ``indicator model'' to refer to any one of the ten
-models of the form \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar_x} (five from
-each of the forecasting and hotspot prediction tasks).
+models of the form \eqref{eq:forecast_ar_x} or \eqref{eq:hotspot_ar_x} (five from
+each of the forecasting and hotspot prediction tasks). We begin with a summary
+of the high-level conclusions.
 
@@
 \item In the same aggregate view, CHNG-COVID and DV-CLI offer the biggest 
-  gains in each of forecasting and hotspot prediction.  CHNG-CLI is
+  gains in both forecasting and hotspot prediction.  CHNG-CLI is
   inconsistent: it provides a big gain in hotspot prediction, but little gain 
@@
 \item In a more detailed analysis of forecasting performance, we find that the
-  indicator models tend to be worse than AR when case rates are increasing (this
-  is most notable in CHNG-CLI and DV-CLI), and better than AR when case rates
-  are flat or decreasing (most notable in CHNG-COVID and CTIS-CLI-in-community).
+  indicator models tend to be
+  better than AR when case rates
+  are flat or decreasing (most notable in CHNG-COVID and CTIS-CLI-in-community),
+  but they are
+  worse than AR when case rates are increasing (this
+  is most notable in CHNG-CLI and DV-CLI).
   More rarely does an indicator model tend to beat AR when case rates are 
@@
   indicator ``leads'' case rates (meaning, roughly, on a short-time scale in a
-  given location, its behavior mimics that of case rates some number of days
-  ahead).  On the other hand, neither leading nor lagging behavior tend to
-  co-occur as consistently with an indicator model doing better in periods of
-  increase, or worse in periods of increase or decrease.   
+  given location, its behavior mimics that of future case rate behavior).  On
+  the other hand, if an an indicator model does better in periods of
+  increase, or worse in periods of increase or decrease, its performance is
+  unrelated to leadingness.   
 \end{itemize}
@@
 of average WIS or AUC). Furthermore, since indicators can have greatly different
-backfill profiles, the use of finalized data in retrospective evaluations can
-even change the relative ordering of models.  For example, CHNG-CLI and DV-CLI
-when trained on finalized data perform very similarly for the forecasting task.
-This makes sense since they are both claims-based indicators measuring
-supposedly the same thing.  However, DV-CLI outperforms CHNG-CLI on vintage
+backfill profiles, the use of finalized data in retrospective evaluations
+changes the relative ranking of models.  For example, CHNG-CLI and DV-CLI,
+when trained on finalized data, perform very similarly for the forecasting task.
+This makes sense since they are both claims-based indicators supposedly measuring
+the same thing.  However, DV-CLI outperforms CHNG-CLI on vintage
 data, reflecting its has a less severe backfill profile.
-% TODO note what we said here about the supplement
 
@@
 The CHNG-CLI model performs basically the same as AR.  This is likely due to the
-fact that CHNG-CLI suffers from some sort of volatility in forecasting due to
-backfill.  The evidence for this is twofold: the CHNG-CLI model benefits from a
-more robust method of aggregating WIS (geometric mean; shown in the supplement),
-and when we train and make predictions on all-finalized data, it provides a big
-gain over AR, on par with the best-performing models (also shown in the
+fact that CHNG-CLI suffers from spatially heterogeneous volatility due to
+backfill.  The evidence for this explanation is twofold: (1) the CHNG-CLI model benefits from a
+more robust method of aggregating WIS (geometric mean; shown in the supplement);
+and (2) when we train and make predictions on finalized data, it handily beats
+AR, on par with the best-performing models (also shown in the 
 supplement).
@@
 prediction dates.  We can see many similarities to the forecasting results (now
-larger AUC is better); for example, CHNG-COVID and DV-CLI offer the biggest
-improvement over AR, and all models including AR degrade in performance towards
-the baseline (this is now just a classifier based on random guessing, which
+larger AUC is better). For example, CHNG-COVID and DV-CLI offer the biggest
+improvement over AR, and all models, including AR, degrade in performance towards
+the baseline (in this context, a classifier based on random guessing, which
 achieves an AUC of 0.5) as the ahead values grow, illustrating the difficulty of
-the hotspot prediction problem A clear difference, however, is that the CHNG-CLI
+the hotspot prediction problem. A clear difference, however, is that the CHNG-CLI
 model performs quite well in hotspot prediction, close to the best-performing
@@
   \includegraphics[width=\textwidth]{fig/fcast-hot-combo-1.pdf}
-  \caption{Left: average WIS for each forecast model, over all forecast dates
+  \caption{Main results for both tasks. Left: average WIS for each forecast
+    model, over all forecast dates 
     and all HRRs, divided by the average WIS achieved by a baseline model (a
@@
 One might ask if the benefits we observe in forecasting and hotspot
-prediction have anything to do with the actual indicator themselves; could it
-instead be that the indicators are simply providing \textit{implicit
-  regularization} on top of  the basic AR model, when we include them as lagged
-features in \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar_x}?
+prediction have anything to do with the actual indicator themselves. A plausible
+alternative explanation is that the indicators are simply providing \textit{implicit
+  regularization} on top of  the basic AR model when we include them as lagged
+features in \eqref{eq:forecast_ar_x} and \eqref{eq:hotspot_ar_x}.
 
 To test this hypothesis, we reran all of the prediction experiments but with
-$X_{\ell,t}$ in the each indicator models in \eqref{eq:forecast_ar_x},
-\eqref{eq:hotspot_ar_x} replaced with suitable random noise (bootstrap samples
+$X_{\ell,t}$ in the each indicator models
+%in \eqref{eq:forecast_ar_x} \eqref{eq:hotspot_ar_x}
+replaced with suitable random noise (bootstrap samples
 from a signal's historical distribution).  The results, shown and explained more
@@
 of results.  In both forecasting and hotspot prediction, the ``fake'' indicator
-models (with random noise as auxiliary features) offered essentially no
+models
+% (with random noise as auxiliary features)
+offered essentially no
 improvement over the pure AR model, which provides a strong rejection
@@
 $\ell_1$ regularization (tuned by cross-validation) in fitting any of the models
-in \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar}, 
-\eqref{eq:hotspot_ar_x} did not improve their generally performance at all
+in \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar}, or 
+\eqref{eq:hotspot_ar_x} did not generally improve their performance
 (experiments not shown).  This is likely due to the fact that the number of
 training samples is large compared to the number of features (6,426 training
-samples and only 3--6 features).
+samples and only 3--6 features). [DJM: and that proper CV requires longer
+training windows, exacerbating the nonstationarity, a scenario in which its
+behavior is poorly understood.]
 
@@
 CHNG-CLI, CHNG-COVID, Google-AA, and especially CTIS-CLI-in-community have
-large right tails (displaying improvements over AR) over the down periods.
+large right tails (displaying improvements over AR) during the down periods.
 Google-AA and CTIS-CLI-in-community have large right tails during the flat
@@
 The supplement pursues this analysis further.  For example, we examine
-classification accuracy and log-likelihood for the hotspot task, and find a
+classification accuracy and log-likelihood for the hotspot task and find a
 similar phenomenon: the indicators considerably improve accuracy or
 log-likelihood during flat or down periods, with more mixed behavior during up
-periods, with the CHNG-CLI, CHNG-COVID, and DV-CLI leading to decreased
-performance in particular.
+periods when  CHNG-CLI, CHNG-COVID, and DV-CLI, in particular, lead to decreased
+performance.
 
@@
 % or just let them find it in the supplement?
+% DJM: Agreed. In the supplement.
 
@@
 As described in the methods section, each of the indicators we examine could be
-said to measure aspects on disease progression that would precede a positive
+said to measure aspects of disease progression that would precede a positive
 test. That is, we imagine that these signals should ``lead'' cases. It is
@@
 of an indicator (degree to which it leads case activity) tends to be more
-pronounced in down or flat periods than up periods, a plausible explanation
+pronounced in down or flat periods than in up periods, a plausible explanation
 for the decreased performance in up periods noted above.
@@
 the WIS difference would be positively correlated with leadingness.  Somewhat
-surprisingly, this relationship turns out to be the strongest in down periods,
+surprisingly, this relationship turns out to be strongest in down periods
 and weakest in up periods.  In fact, it is very nearly the case that for each

### 2021-06-19 forecast/paper/paper.tex
Done with proofing.

@@
 COVIDcast API (defined using from medical insurance claims, internet-based
-surveys, and Internet search trends) is undoubtedly ``yes''. However, there are
+surveys, and internet search trends) is undoubtedly ``yes''. However, there are
 levels of nuance to such an answer that must be explained.  None of the
@@
 backtesting them with finalized data may lead us to make modeling decisions that
-may not be relevant for true test-time performance.
+are suboptimal for true test-time performance.
 

### 2021-06-19 forecast/paper/paper.tex
Temp commit so I can merge

--- 2021-06-19 forecast/paper/paper.tex
@@
   claims, self-reported symptoms through online surveys, and COVID-related
-  internet% Capitalization is apparently
-  % in question: https://en.wikipedia.org/wiki/Capitalization_of_Internet
-  search activity all provide a nontrivial boost in accuracy, though
+  internet search activity all provide a nontrivial boost in accuracy, though
   their utility generally depends the pandemic's local dynamics (upswing,
@@
 }
-
-  % TODO % For reference, below is the indicators paper one.
-  % To study the COVID-19 pandemic, its effects on society,
-  % and measures for reducing its spread, researchers need detailed data on the
-  % course of the pandemic. Standard public health data streams suffer
-  % inconsistent reporting and frequent, unexpected revisions. They also miss
-  % other aspects of a population's behavior, worthy of consideration. We
-  % present an open database of COVID signals in the United States, measured at
-  % the county level and updated daily. These include traditionally reported
-  % COVID cases and deaths, and many others: signals based on measures of
-  % mobility, social distancing, internet search trends, self-reported symptoms,
-  % and patterns COVID-related activity in de-identified medical insurance 
-  % claims. The database provides all signals in a common, easy-to-use format,
-  % empowering both  public health research and operational decision-making. 
-%}
 

### 2021-06-19 forecast/paper/supplement-text.tex
Edit bootstrap supplement

@@
 based on six features, corresponding to the three autoregressive terms
-and the three lagged indicator values.  To form noise features
-we resample from other rows.  In particular, at each location $\ell$ and
-time $t$, we replace the triplet $(X_{\ell,t}, X_{\ell,t-7},
-X_{\ell,t-14})$ by the triplet $(X_{\ell^*,t^*}, X_{\ell^*,t^*-7},
+and the three lagged indicator values.  To form noise indicator features,
+we resample from other rows (while keeping the autoregressive
+features fixed).  In particular, at each location $\ell$ and
+time $t$, for the forecasting task we replace the triplet $(X_{\ell,t}, X_{\ell,t-7},
+X_{\ell,t-14})$ in Eq. (3) of the main paper with the triplet $(X_{\ell^*,t^*}, X_{\ell^*,t^*-7},
 X_{\ell^*,t^*-14})$, where $(\ell^*,t^*)$ is a location-time pair
-sampled with replacement from the 6,426 location-time pairs.  Figures
+sampled with replacement from the 6,426 location-time pairs.  
+Likewise in the hotspot prediction task, we replace the triplet $(X_{\ell,t}^\Delta, X_{\ell,t-7}^\Delta,
+X_{\ell,t-14}^\Delta)$ in Eq. (5) of the main paper with $(X_{\ell^*,t^*}^\Delta, X_{\ell^*,t^*-7}^\Delta,
+X_{\ell^*,t^*-14}^\Delta)$.
+Figures
 \ref{fig:fcast-booted}--\ref{fig:hot-booted} show the results.  No
@@
 leading us to dismiss the implicit regularization hypothesis.
-
-\attn{Would be good for someone to double check that the bootstrap
-  procedure is correctly described.}
 

### 2021-06-19 indicators/paper/paper.tex

@@
   cases and deaths, and many others: measures of mobility, social distancing,
-  Internet search trends, self-reported symptoms, and patterns of COVID-related 
+  internet search trends, self-reported symptoms, and patterns of COVID-related 
   activity in de-identified medical insurance claims. The database provides all
@@
   signals extracted from de-identified medical claims data, massive online
-  surveys, cell phone mobility data, and Internet search trends. These 
+  surveys, cell phone mobility data, and internet search trends. These 
   are available at a fine geographic resolution (mostly at the county level) and
@@
 not found in traditional public health streams, such as data on mobility
-patterns, Internet search trends, mask wearing, and vaccine hesitancy, to name
+patterns, internet search trends, mask wearing, and vaccine hesitancy, to name
 just a few. (The latter two signals are derived from the COVID Trends and Impact

### 2021-06-19 forecast/paper/paper.tex
Copyedit abstract

@@
 % five keywords, separated by the pipe symbol. 
-\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}
+\keywords{COVID-19 $|$ forecasting $|$ hotspot prediction $|$ time series | 
+  digital surveillance}
 
 \begin{abstract}
-Reliable short-term, location-specific forecasts of a pandemic are a
-key ingredient in effective public health decision making. Throughout the
-COVID-19 pandemic, a plethora of real-time data signals were
-collected, curated, and made publicly available by our team and data
-partners, providing multiple views of the pandemic's progress. The
-goal of this work is to measure the practical utility of these
-auxiliary indicators from a forecasting perspective. We focus on five
-indicators, derived from medical insurance claims data, web search
-queries, and survey responses. For each indicator, we ask whether its
-inclusion in a simple model leads to improved forecasting
-performance relative to a similar model excluding it. Performance is considered across two tasks
-at a forecasting horizon of up to about one month:
-probabilistic forecasting of seven-day average case rates and binary
-prediction of ``hotspots.'' Since the values of indicators are
-commonly revised over time, we take special care to ensure that the data provided
-to a forecaster is the version that would have been available at the
-time that the forecast was made. We demonstrate how less careful
-backtesting can lead to misleading retrospective evaluations. Our analysis
-suggests consistent, although 
-modest, gains in forecasting performance by using these indicators,
-and further investigation suggests that these gains are related to periods when the
-auxiliary signals are acting as ``leading 
+  Reliable, short-term forecasts of traditional public health reporting streams
+  (such as cases, hospitalizations, and deaths) are a key ingredient in
+  effective public health decision-making during a pandemic. Since April 2020,
+  our research group has worked with data partners to collect, curate, and make
+  publicly available numerous real-time COVID-19 indicators, providing multiple
+  views of pandemic activity.  This paper studies the utility of these
+  indicators from a forecasting perspective. We focus on five indicators,
+  derived from medical insurance claims data, web search queries, and online
+  survey responses. For each indicator, we ask whether its inclusion in a simple
+  model leads to improved predictive accuracy relative to a similar model
+  excluding it. We consider both probabilistic forecasting of confirmed COVID-19 
+  case rates and binary prediction of case ``hotspots''.  Since the values of 
+  indicators (and case rates) are commonly revised over time, we take special
+  care to ensure that the data provided to a forecaster is the version that
+  would have been available at the time the forecast was made.   
+  % We demonstrate how less careful backtesting can lead to misleading
+  % retrospective evaluations.   
+  Our  analysis shows that consistent but modest gains in predictive accuracy 
+  are obtained by using these indicators, and furthermore, these gains are
+  related to periods in which the auxiliary indicators behave as ``leading
   indicators'' of case rates.     
-\attn{Evaluating forecasters thoroughly and fairly is an important way that
-we can learn from this pandemic to do better in the next. 251 :(}
-
-% Please provide an abstract of no more than 250 words in a single paragraph.
-% Abstracts should explain to the general reader the major contributions of the
-% article. References in the abstract must be cited in full within the abstract
-% itself and cited in the text.
 \end{abstract}

### 2021-06-19 forecast/paper/paper.tex
Full pass of edits

@@
 % Please add a significance statement to explain the relevance of your work
-\significancestatement{Forecasting is a critical element of the public health
+\significancestatement{Forecasting is a critical element in the public health
   response to any fast-moving epidemic or pandemic.  Predicting the future
@@
 people (the standard in epidemiology).  We use reported case data aggregated by 
-JHU CSSE \cite{Dong:2020}, which, like the auxiliary indicators that we use to supplement the
-basic purely autoregressive models, is accessible through the COVIDcast API
-\cite{CovidcastAPI}.
+JHU CSSE \cite{Dong:2020}, which, like the auxiliary indicators that we use to
+supplement the basic autoregressive models, is accessible through the COVIDcast
+API \cite{CovidcastAPI}.
 
@@
   (CTIS-CLI-in-community): The estimated percentage of the population who know
-  someone in their local community that is sick, based on Delphi's surveys of
-  Facebook users, called the COVID Trends and Impact Survey. 
+  someone in their local community that is sick, based on Delphi's COVID Trends
+  and Impact Survey, in partnership with Facebook.
 \item Doctor Visits COVID-like illness (DV-CLI): The same as CHNG-CLI, but
@@
 \item Google Search Trends for anosmia and ageusia (Google-AA): A measure of 
-  Google search volume for queries relating to anosmia or
-  ageusia (loss of smell or taste), based on Google's Search Trends data set. 
+  Google search volume for queries that relate to anosmia or ageusia (loss of
+  smell or taste), based on Google's Search Trends data set.  
 \end{itemize}
@@
 on this survey. For the Google Search Trends data set, see
-\cite{GoogleSymptoms}, and see \cite{Klopfen:2020, Vaira:2020} for justification of the
-relevance of anosmia or ageusia to COVID-19 infection.
+\cite{GoogleSymptoms}, and see \cite{Klopfen:2020, Vaira:2020} for justification
+of the relevance of anosmia or ageusia to COVID-19 infection.
 
@@
 hotspots) at the HRR level because the indicators considered should be more 
-useful in predicting case activity rather than hospital demand. The \attn{former?} is
+useful in predicting case activity rather than hospital demand, as the former is  
 intuitively more contemporaneous to the events that are measured by
@@
 that for a given prediction date, say, September 28, 2020, we train models 
-using only the data that would have been available to us \textit{as of}
+using data that would have been available to us \textit{as of}
 September 28 (imagine that we can ``rewind'' the clock to September 28 and query
 the COVIDcast API to get the latest data it would have had available at that
-point in time.)  This is made possible by the COVIDcast API's comprehensive data   
-versioning system (described in more detail in our companion paper).  We also
-use the \texttt{evalcast} R package \cite{EvalcastR}, which streamlines the
+point in time.)  This is possible because of the COVIDcast API's comprehensive
+data versioning system (described in more detail in our companion paper).  We
+also use the \texttt{evalcast} R package \cite{EvalcastR}, which streamlines the 
 process of training arbitrary prediction models over a sequence of prediction
@@
   \includegraphics[width=\columnwidth]{fig/revisions.pdf}
-  \caption{This figure displays revision behavior for two signals in the HRR
-    containing New York 
+  % RJT: Ensuring we start with a "declarative title" :) 
+  % https://www.internationalscienceediting.com/how-to-write-a-figure-caption/
+  \caption{Revision behavior for two signals in the HRR containing New York  
     City.  Each colored line corresponds to the data as reported on a particular 
@@
 case rates, is 1 day latent, and exhibits frequent, unpredictable revisions
-after initial publication.  Compared to the revision
-pattern in the medical claims signals, which are much more systematic in nature,
-revisions in case reports can be highly erratic. Big spikes or other anomalies
-can occur in the data as reporting backlogs are cleared, changes in case
-definitions are made, etc. Groups like JHU CSSE then work tirelessly to correct
-such anomalies after first publication (e.g., they will attempt to
-back-distribute a spike when a reporting backlog is cleared, by working with a
-local authority to figure out how this should best be done), which can result
-in very nontrivial revisions.  See
-Figure~\ref{fig:vintage} for an example.
+after intial publication.  Compared to the revision pattern in the medical
+claims signals, which are much more systematic in nature, revisions in case
+reports can be highly erratic. Big spikes or other anomalies can occur in the
+data as reporting backlogs are cleared, changes in case definitions are made,
+etc. Groups like JHU CSSE then work tirelessly to correct such anomalies after
+first publication (e.g., they will attempt to back-distribute a spike when a
+reporting backlog is cleared, by working with a local authority to figure out
+how this should best be done), which can result in very nontrivial revisions.
+See Figure~\ref{fig:vintage} for an example.
 
@@
 is robustness: averaging stabilizes the case time series, and accounts for 
-uninteresting artifacts like day-of-the-week effects in the series.  Note
-that we can equivalently view this (equivalent up to a constant
-factor) as predicting the HRR-level COVID-19 case incidence rate \textit{summed}
-over some 7-day period in the future, and predicting a binary variable derived
-from this.
+uninteresting artifacts like day-of-the-week effects in the series.  Note that
+we can also equivalently view this (equivalent up to a constant factor) as
+predicting the HRR-level case incidence rate \textit{summed} over some 7-day
+period in the future, and predicting a binary variable derived from this.
 
@@
 ends up being (per our setup, with 21 days of training data and lagged values 
-of signals for features, as we will detail shortly) June 9, 2020 for forecasting,
-and June 16, 2020 for hotspot prediction. (The bottleneck here is the
-CTIS-CLI-in-community signal, which does not exist before early April 2020, when
-the survey was first launched). The end date was chosen again with a
+of signals for features, as we will detail shortly) June 9, 2020 for
+forecasting, and June 16, 2020 for hotspot prediction. (The bottleneck here is
+the CTIS-CLI-in-community signal, which does not exist before early April 2020,
+when the survey was first launched). The end date was chosen again with a 
 consideration to align both tasks as best as possible, and because few hotspots
@@
 \label{eq:forecast_ar}
-\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s},\ s \leq t)  
+\Quantile_\tau(Y_{\ell,t+a} \ | \ Y_{\ell,s},\ s \leq t)   
 = \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j}.  
@@
 \label{eq:forecast_ar_x}
-\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s},\ X_{\ell,s},\ s \leq t)  
+\Quantile_\tau(Y_{\ell,t+a} \ | \ Y_{\ell,s},\ X_{\ell,s},\ s \leq t) 
 = \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j} + 
@@
 
-At prediction time, in order to avoid crossing violations (that is,
-for two levels $\tau' > \tau$,
-the predicted quantile at level $\tau$ exceeds the predicted quantile at level
-$\tau'$), we apply a simple post-hoc sorting.  See Figure~\ref{fig:trajectory}
-for an example forecast.
+At prediction time, in order to avoid crossing violations (that is, for two
+levels $\tau' > \tau$, the predicted quantile at level $\tau$ exceeds the
+predicted quantile at level $\tau'$), we apply a simple post-hoc sorting.  See
+Figure~\ref{fig:trajectory} for an example forecast.
 
@@
 \includegraphics[width=\columnwidth]{fig/trajectory.pdf}
-\caption{This figure displays a forecast from an autoregressive model for the
-  HRR containing
-  New York City made  
-  on October 15 (vertical line).  The fan displays 50\%, 80\% and 95\% intervals
-  while the orange curve shows the median forecast. The black curve represents
-  ``finalized'' data, as reported in May 2021.} 
+\caption{Forecast for the HRR containing New York City from an autoregressive
+  model made on October 15 (vertical line).  The fan displays 50\%, 80\% and
+  95\% intervals while the orange curve shows the median forecast. The black
+  curve shows ``finalized'' data, as reported in May 2021.}   
 \label{fig:trajectory}
@@
 \label{eq:hotspot_ar}
-\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s},\ s \leq t) \big) 
+\logit \big( \P(Z_{\ell,t+a} = 1 \ | \ Y_{\ell,s},\ s \leq t) \big) 
 = \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j},
@@
 \label{eq:hotspot_ar_x}
-\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s},\ X_{\ell,s},\ s \leq t) \big)
+\logit \big( \P(Z_{\ell,t+a} = 1 \ | \ Y_{\ell,s},\ X_{\ell,s},\ s \leq t) \big)
 = \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j} +  
@@
 \textit{weighted interval score} (WIS), a quantile-based scoring rule; see e.g.,
-\cite{Gneiting:2007}.  WIS is a proper score, meaning that its expectation
+\cite{Gneiting:2007}.  WIS is a proper score, which means that its expectation 
 is minimized by the population quantiles of the target variable.  The use of WIS
@@
 levels), the first form in \eqref{eq:wis_intervals} is typically preferred in
-presentation, because the score nicely decouples into a ``sharpness'' component
-(first term in each summand) and an ``under/overprediction'' component (second term
-in each summand).  The second form given in \eqref{eq:wis_quantiles} is
-especially noteworthy in our current study because it reveals WIS to be the same as the
-quantile regression loss that we use to train our forecasting models (i.e., our
-models are fit to optimize WIS averaged over the training set).
+presentation, as the score nicely decouples into a ``sharpness'' component
+(first term in each summand) and an ``under/overprediction'' component (second
+term in each summand).  But the second form given in \eqref{eq:wis_quantiles} is 
+especially noteworthy in our current study because it reveals WIS to be the same
+as the quantile regression loss that we use to train our forecasting models
+(i.e., we fit by optimizing WIS averaged over the training data).
 
 For hotspot prediction, we evaluate the probabilistic classifiers produced by
-the logistic models in \eqref{eq:hotspot_ar} and \eqref{eq:hotspot_ar_x} using the 
-area under the curve (AUC) of their true positive versus false positive rate
+the logistic models in \eqref{eq:hotspot_ar} and \eqref{eq:hotspot_ar_x} using
+the area under the curve (AUC) of their true positive versus false positive rate 
 curve (which is traced out by varying the discrimination threshold).
@@
 
-Here, and in the
-following subsections, we will use ``AR'' to refer to the pure autoregressive
-model both in forecasting, \eqref{eq:forecast_ar}, and in hotspot prediction,
-\eqref{eq:hotspot_ar} (the reference to the prediction task should always be
-clear from the context). We will also use the name of an auxiliary
+Here, and in the following subsections, we will use ``AR'' to refer to the pure
+autoregressive model both in forecasting, \eqref{eq:forecast_ar}, and in hotspot
+prediction, \eqref{eq:hotspot_ar} (the reference to the prediction task should
+always be clear from the context). We will also use the name of an auxiliary 
 indicator---namely ``CHNG-CLI'',  ``CHNG-COVID'', ``CTIS-CLI-in-community'',
@@
 Finally, we use the term ``indicator model'' to refer to any one of the ten
-models of the form \eqref{eq:forecast_ar_x} or \eqref{eq:hotspot_ar_x} (five from
-each of the forecasting and hotspot prediction tasks). We begin with a summary
-of the high-level conclusions.
+models of the form \eqref{eq:forecast_ar_x} or \eqref{eq:hotspot_ar_x} (five
+from each of the forecasting and hotspot prediction tasks).
+
+We begin with a summary of the high-level conclusions.
 
@@
 \item In a more detailed analysis of forecasting performance, we find that the
-  indicator models tend to be
-  better than AR when case rates
-  are flat or decreasing (most notable in CHNG-COVID and CTIS-CLI-in-community),
-  but they are
-  worse than AR when case rates are increasing (this
-  is most notable in CHNG-CLI and DV-CLI).
-  More rarely does an indicator model tend to beat AR when case rates are 
-  increasing, but there appears to be some evidence of this for the Google-AA
-  model. 
+  indicator models tend to be better than AR when case rates are flat or
+  decreasing (most notable in CHNG-COVID and CTIS-CLI-in-community), but they
+  are worse than AR when case rates are increasing (this is most notable in
+  CHNG-CLI and DV-CLI). More rarely does an indicator model tend to beat AR when
+  case rates are increasing, but there appears to be some evidence of this for
+  the Google-AA model. 
   % In an increasing period, when an indicator model is worse than AR, it tends
@@
   indicator ``leads'' case rates (meaning, roughly, on a short-time scale in a
-  given location, its behavior mimics that of future case rate behavior).  On
+  given location, its behavior mimics that of future case rates).  On
   the other hand, if an an indicator model does better in periods of
   increase, or worse in periods of increase or decrease, its performance is
-  unrelated to leadingness.   
+  not as related to leadingness.   
 \end{itemize}
@@
 changes the relative ranking of models.  For example, CHNG-CLI and DV-CLI,
-when trained on finalized data, perform very similarly for the forecasting task.
-This makes sense since they are both claims-based indicators supposedly measuring
-the same thing.  However, DV-CLI outperforms CHNG-CLI on vintage
+when trained on finalized data, perform very similarly in forecasting.
+This makes sense since they are both claims-based indicators that are supposedly
+measuring the same thing.  However, DV-CLI outperforms CHNG-CLI on vintage
 data, reflecting its has a less severe backfill profile.
@@
 \subsection{Aggregate Results by Ahead Value}
+
+\begin{figure*}[t]
+  \includegraphics[width=\textwidth]{fig/fcast-hot-combo-1.pdf}
+  \caption{Main results for both tasks. Left: average WIS for each forecast
+    model, over all forecast dates and all HRRs, divided by the average WIS
+    achieved by a baseline model (a probabilistic version of the flat-line
+    forecaster).  Right: area under the curve for each hotspot prediction model,
+    calculated over all prediction dates and all HRRs.  Here and in all figures
+    we abbreviate CTIS-CLI-in-community by CTIS-CLIIC.  
+  }  
+  \label{fig:forecast}
+\end{figure*}
 
@@
 The CHNG-CLI model performs basically the same as AR.  This is likely due to the
-fact that CHNG-CLI suffers from spatially heterogeneous volatility due to
-backfill.  The evidence for this explanation is twofold: (1) the CHNG-CLI model benefits from a
-more robust method of aggregating WIS (geometric mean; shown in the supplement);
-and (2) when we train and make predictions on finalized data, it handily beats
-AR, on par with the best-performing models (also shown in the 
-supplement).
-% TODO note what we said here about the supplement
+fact that CHNG-CLI suffers from volatility due to backfill.  The evidence for
+this explanation is twofold: (1) the CHNG-CLI model benefits from a more robust
+method of aggregating WIS (geometric mean; shown in the supplement); and (2)
+when we train and make predictions on finalized data, it handily beats AR, on
+par with the best-performing models (also shown in the supplement).
 
@@
 larger AUC is better). For example, CHNG-COVID and DV-CLI offer the biggest
-improvement over AR, and all models, including AR, degrade in performance towards
-the baseline (in this context, a classifier based on random guessing, which
-achieves an AUC of 0.5) as the ahead values grow, illustrating the difficulty of
-the hotspot prediction problem. A clear difference, however, is that the CHNG-CLI
-model performs quite well in hotspot prediction, close to the best-performing
-indicator models for many of the ahead values. This may be because volatility in
-the CHNG-CLI indicator plays less of a role in the associated logistic model's
-predicted probabilities (in general, a sigmoid function can absorb a lot of the
-variability in its input).
-
-\begin{figure*}[t]
-  \includegraphics[width=\textwidth]{fig/fcast-hot-combo-1.pdf}
-  \caption{Main results for both tasks. Left: average WIS for each forecast
-    model, over all forecast dates 
-    and all HRRs, divided by the average WIS achieved by a baseline model (a
-    probabilistic version of the flat-line forecaster).  Right: area under the
-    curve for each hotspot prediction model, calculated over all prediction
-    dates and all HRRs.  Here and in all figures we abbreviate
-    CTIS-CLI-in-community by CTIS-CLIIC.  
-  }  
-  \label{fig:forecast}
-\end{figure*}
+improvement over AR, and all models, including AR, degrade in performance
+towards the baseline (in this context, a classifier based on random guessing,
+which achieves an AUC of 0.5) as the ahead values grow, illustrating the
+difficulty of the hotspot prediction problem. A clear difference, however, is
+that the CHNG-CLI model performs quite well in hotspot prediction, close to the
+best-performing indicator models for many of the ahead values. This may be
+because volatility in the CHNG-CLI indicator plays less of a role in the
+associated logistic model's predicted probabilities (in general, a sigmoid
+function can absorb a lot of the variability in its input).
 
@@
 To test this hypothesis, we reran all of the prediction experiments but with
-$X_{\ell,t}$ in the each indicator models
-%in \eqref{eq:forecast_ar_x} \eqref{eq:hotspot_ar_x}
-replaced with suitable random noise (bootstrap samples
-from a signal's historical distribution).  The results, shown and explained more
-precisely in the supplement, are vastly different (worse) than the original set
-of results.  In both forecasting and hotspot prediction, the ``fake'' indicator
-models
-% (with random noise as auxiliary features)
-offered essentially no
-improvement over the pure AR model, which provides a strong rejection
-(informally speaking) of the implicit regularization hypothesis.
+$X_{\ell,t}$ in the each indicator models replaced with suitable random noise
+(bootstrap samples from a signal's history).  The results, shown and explained
+more precisely in the supplement, are vastly different (worse) than the original
+set of results.  In both forecasting and hotspot prediction, the ``fake''
+indicator models offered essentially no improvement over the pure AR model,
+which strongly rejects (informally speaking) the implicit regularization
+hypothesis.
 
@@
 training samples is large compared to the number of features (6,426 training
-samples and only 3--6 features). [DJM: and that proper CV requires longer
-training windows, exacerbating the nonstationarity, a scenario in which its
-behavior is poorly understood.]
-
-% \begin{figure}[t]
-% \includegraphics[width=\columnwidth]{fig/hot-1.pdf}
-% \caption{Area under the curve for each hotspot prediction model, calculated over
-%   all prediction dates and all HRRs.} 
-% \label{fig:hotspot}
-% \end{figure}
+samples and only 3--6 features). 
+% DJM: and that proper CV requires longer training windows, exacerbating the
+% nonstationarity, a scenario in which its behavior is poorly understood.
 
 \subsection{Evaluation in Up, Down, and Flat Periods}
+
+\begin{figure*}[t]
+  \includegraphics[width=\textwidth]{fig/upswing-histogram-1.pdf}
+  \caption{Histogram of the difference in WIS for the AR model and that for each
+    indicator model, stratified by up, down, or flat period, measured in terms
+    of case trends. Note that \textit{larger} differences here are better for
+    each indicator model.  The y-axis is on the log scale to emphasize tail 
+    behavior.}   
+  \label{fig:up_down_flat}
+\end{figure*}
 
@@
 
-\begin{figure*}[t]
-  \includegraphics[width=\textwidth]{fig/upswing-histogram-1.pdf}
-  \caption{Histogram of the difference in WIS for the AR model and that for each
-    indicator model, stratified by up, down, or flat period, measured in terms
-    of case trends. Note that \textit{larger} differences here are better for
-    each indicator model.  The y-axis is on the log scale to emphasize tail 
-    behavior.}   
-  \label{fig:up_down_flat}
-\end{figure*}
-
 The supplement pursues this analysis further.  For example, we examine
@@
 
-% Essentially, during downswings, improved performance is strongly correlated
-% with predicing larger declines than the AR model. The AR model tends to
-% overpredict relative to the decline, but adding indicators tends to diminish
-% this effect. During upswings, the AR model tends to underpredict, but when the
-% addition of an indicator improves performance (a more rare occurrence) it is 
-% rarely due to higher median predictions.
+\subsection{Effects of Leading or Lagging Behavior}
 
-% RJT: Commented out the last paragraph for now, because this is kind of a tough
-% thing to explain succintly. Can revisit if we really want to include it here,
-% or just let them find it in the supplement?
-% DJM: Agreed. In the supplement.
-
-\subsection{Effects of Leading or Lagging Behavior}
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{fig/leading-only-1.pdf}
+  \caption{Correlation of the difference in WIS with the ``leadingness'' of the
+    indicator at the target date, stratified by up, down, or flat period.} 
+  \label{fig:leading}
+\end{figure}
 
@@
 indicator and the WIS difference (AR model minus an indicator model), stratified
-by whether the target is classified as up, down, or flat.  One would expect that  
-the WIS difference would be positively correlated with leadingness.  Somewhat
-surprisingly, this relationship turns out to be strongest in down periods
-and weakest in up periods.  In fact, it is very nearly the case that for each
-indicator, the strength of correlations only decreases as we move from down to
-flat to up  periods.  In the supplement, we extend this analysis by examining
+by whether the target is classified as up, down, or flat.  One would naturally
+expect that the WIS difference would be positively correlated with leadingness.
+Somewhat surprisingly, this relationship turns out to be strongest in down
+periods and weakest in up periods.  In fact, it is very nearly the case that for
+each indicator, the strength of correlations only decreases as we move from down
+to flat to up  periods.  In the supplement, we extend this analysis by studying 
 analogous ``laggingness'' scores, but we do not find as clear patterns.
-
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/leading-only-1.pdf}
-  \caption{Correlation of the difference in WIS with the ``leadingness'' of the
-    indicator at the target date, stratified by up, down, or flat period.} 
-  \label{fig:leading}
-\end{figure}
 
@@
   supported by Centers for Disease Control and Prevention (CDC) grant
-  U01IP001121; gifts from Facebook, Google, McCune, and Optum; the National
-  Science Foundation Graduate Research Fellowship Program (NSF GRFP) award
-  DGE175016; the Canadian Statistical Sciences Institute; and the National
-  Sciences and Engineering Research Council of Canada (NSERC) RGPIN-2021-02618.}
+  U01IP001121; gifts from Facebook, Google, McCune, and Optum; the Canadian
+  Statistical Sciences Institute; National Sciences and Engineering Research
+  Council of Canada (NSERC) RGPIN-2021-02618; and National Science Foundation
+  Graduate Research Fellowship Program (NSF GRFP) award DGE175016.}
 

### 2021-06-19 forecast/code/revisions.R
Improve revisions fig

@@ -4,11 +4,13 @@
 
 library(covidcast)
 library(tidyverse)
+library(gridExtra)
+library(grid)
 
 as_ofs <- seq(as.Date("2020-09-28"), as.Date("2020-10-19"), by = "week")
 cases_as_of <- map_dfr(as_ofs, function(as_of) {
   covidcast_signal(data_source = "jhu-csse", 
-                   signal = "confirmed_incidence_prop",
+                   signal = "confirmed_7dav_incidence_prop",
                    start_day = "2020-08-15", end_day = "2020-09-26", 
                    geo_type = "hrr", geo_values = "311", as_of = as_of) %>%
     mutate(as_of = as_of)
@@ -22,23 +24,62 @@ dv_as_of <-  map_dfr(as_ofs, function(as_of) {
 })
 
   
-bind_rows(cases_as_of, dv_as_of) %>% 
+p1 <- dv_as_of %>%
   mutate(as_of = fct_relabel(factor(as_of), function(x) strftime(x, "%b %d"))) %>%
   ggplot(aes(x = time_value, y = value)) + 
   geom_line(aes(color = factor(as_of))) + 
-  facet_wrap(~data_source, scales = "free_y",
-             strip.position = "left", 
-             labeller = as_labeller(
-               c(`doctor-visits` = "% doctor's visits due to CLI",
-                 `jhu-csse` = "Cases per 100,000 people"))) +
-  theme_bw() +
-  labs(x = "", y = "", color = "As of:") +
-  theme(legend.position = c(.5, -.15),
-        legend.direction = "horizontal",
-        plot.margin = margin(b = 20),
-        legend.background = element_rect(fill = "transparent")) +
+  labs(title = "DV-CLI", x = "", y = "% doctor's visits due to CLI",
+       color = "As of:") +
+  theme_bw() + 
+  theme(legend.position = "bottom") +
   scale_color_viridis_d(end = .9, begin = .1) +
   guides(color = guide_legend(nrow = 1))
 
-ggsave("../paper/fig/revisions.pdf", width = 5, height = 3.5)
+p2 <- cases_as_of %>%
+  mutate(as_of = fct_relabel(factor(as_of), function(x) strftime(x, "%b %d"))) %>%
+  ggplot(aes(x = time_value, y = value)) + 
+  geom_line(aes(color = factor(as_of))) + 
+  labs(title = "Cases", x = "", y = "Cases per 100,000 people",
[…]

### 2021-06-19 indicators/paper/paper.tex
Only five keywords allowed
Typos, copyedits

@@
 % At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
-\keywords{open data $|$ COVID-19 $|$ digital surveillance $|$ medical insurance
-  claims $|$ internet surveys $|$ search trends $|$ mobility patterns}
+\keywords{COVID-19 $|$ open data $|$ digital surveillance $|$ medical insurance  
+  claims $|$ internet surveys}
 
@@
 \item \textit{Time-wise correlations} (i.e., at a specific location, do values
-  of the signal correlate with case rates across time?): Formally, let $X_\ell$
+  of the signal correlate with case rates across time?): Let $X_\ell$
   and $Y_\ell$ be vectors of values of a signal and case rates, over all
@@
 trailing average), from April 15, 2020 to April 15, 2021. This calculation is
-performed over counties with at least 500 cumulative cases by the end of this
-period, and at which all indicators are available (956 counties in total). The
-large positive correlations suggest that these signals could be useful in
-hotspot detection (identifying counties that have relatively high COVID
-activity, at a given time). Somewhat surprisingly, the survey-based
+performed over all counties with at least 500 cumulative cases by the end of
+this period, and at which all indicators are available (956 counties in
+total). The large positive correlations suggest that these signals could be
+useful in hotspot detection (identifying counties that have relatively high
+COVID activity, at a given time). Somewhat surprisingly, the survey-based 
 CLI-in-community signal shows the strongest correlations for much of the time
@@
 
+[TODO: improve/revisit this example?]
 As an illustration, Figure~\ref{fig:mobility-drop} maps the sharp increase in
@@
 geographic differences in market share for medical claims data, or biases in the
-population represented in mobile device mobility data.
+population represented in app-based mobility data.
 % \cite{Coston:2021}. RJT: commenting out because if we cite only this, then I 

### 2021-06-19 forecast/paper/paper.tex
Typos, copyedits

@@
   systems partners.  
-\item Google Search Trends for anosmia and ageusia (Google-AA): A measure of 
+\item Google search trends for anosmia and ageusia (Google-AA): A measure of  
   Google search volume for queries that relate to anosmia or ageusia (loss of
-  smell or taste), based on Google's Search Trends data set.  
+  smell or taste), based on Google's COVID-19 Search Trends data set.  
 \end{itemize}
@@
 on the API and database. For CTIS in particular, we refer to our companion paper
-on this survey. For the Google Search Trends data set, see
-\cite{GoogleSymptoms}, and see \cite{Klopfen:2020, Vaira:2020} for justification
-of the relevance of anosmia or ageusia to COVID-19 infection.
+on this survey. For the Google COVID-19 Search Trends data set, see
+\cite{GoogleSymptoms}; see also \cite{Klopfen:2020, Vaira:2020} for a 
+justification of the relevance of anosmia or ageusia to COVID-19 infection.
 
@@
     throughout the period, though the effects fade as we look further back in   
-    time. In contrast, the right panel shows case rates reported by JHU CSSE,
-    which remain ``as reported'' on September 28, with a huge spike towards the 
-    end of this period, until a major correction is made on October 19, which
-    brings down this spike and affects all prior data as well.}
+    time. In contrast, the right panel shows case rates reported by JHU CSSE
+    (smoothed with a 7-day trailing average), which remain ``as reported'' on
+    September 28, with a spike towards the end of this period, until a major
+    correction is made on October 19, which brings this down and affects all
+    prior data as well.}  
   \label{fig:vintage}
@@
 signals. The three signals based on medical claims, CHNG-CLI, CHNG-COVID, and
-DV-CLI, are most often 3-5 days latent, and subject to a considerable but
+DV-CLI, are typically 3-5 days latent, and subject to a considerable but
 regular degree of revision or ``backfill'' after their initial publication date.
@@
 case rates, is 1 day latent, and exhibits frequent, unpredictable revisions
-after intial publication.  Compared to the revision pattern in the medical
+after intial publication.  Compared to the pattern of revisions in the medical 
 claims signals, which are much more systematic in nature, revisions in case
@@
 underlying search trends time series is self-normalized, see
-\cite{GoogleSymptoms}) and the claims-based signals (see the discussion in our
-companion paper on the API and database).  Such spatial heterogeneity likely
-hurts the performance of the predictive models that rely on these signals,
-because we train the models on data pooled over all locations.  In the current
-paper, we do not attempt to address this issue (it is again a topic of ongoing
-work in our group), and we simply note that location-specific effects (or
-pre-processing to remove spatial bias) would likely improve the performance of
-the models involving Google-AA and the claims-based indicators.
+\cite{GoogleSymptoms}) and the claims-based signals (due to market-share 
+differences, and/or differences in health-seeking behavior).  Such spatial
+heterogeneity likely hurts the performance of the predictive models that rely on
+these signals, because we train the models on data pooled over all locations.
+In the current paper, we do not attempt to address this issue (it is again a
+topic of ongoing work in our group), and we simply note that location-specific
+effects (or pre-processing to remove spatial bias) would likely improve the
+performance of the models involving Google-AA and the claims-based indicators.
 
@@
 
-Here, and in the following subsections, we will use ``AR'' to refer to the pure
+Here, and in what follows, we will use ``AR'' to refer to the pure 
 autoregressive model both in forecasting, \eqref{eq:forecast_ar}, and in hotspot
@@
 while devoting most of our effort to accounting for as much of the complexity of
-the data and evaluation as possible.  In fact, our paper is as much about
-demonstrating how one might address questions about model comparisons and
+the underlying data and evaluation as possible.  In fact, our paper is as much
+about demonstrating how one might address questions about model comparisons and
 evaluation in forecasting and hotspot prediction in general, as it is about
@@
   Johansson, Evan Ray, Nicholas Reich, and Roni Rosenfeld for several
-  enlightening conversations about forecast evaluation. The research was
-  supported by Centers for Disease Control and Prevention (CDC) grant
-  U01IP001121; gifts from Facebook, Google, McCune, and Optum; the Canadian
-  Statistical Sciences Institute; National Sciences and Engineering Research
-  Council of Canada (NSERC) RGPIN-2021-02618; and National Science Foundation
-  Graduate Research Fellowship Program (NSF GRFP) award DGE175016.}
+  enlightening conversations about scoring and evaluation.  The research was  
+  supported by gifts from Facebook, Google, McCune, and Optum; 
+  Centers for Disease Control and Prevention (CDC) grant U01IP001121; the 
+  Canadian Statistical Sciences Institute; National Sciences and Engineering
+  Research Council of Canada (NSERC) grant RGPIN-2021-02618; and National
+  Science Foundation Graduate Research Fellowship Program (NSF GRFP) award
+  DGE175016.}
 

### 2021-06-21 forecast/paper/paper.tex
Update acks

@@
   Science Foundation Graduate Research Fellowship Program (NSF GRFP) award
-  DGE175016.}
+  DGE1745016.}
 

### 2021-06-21 indicators/paper/paper.tex
Update acks

@@
 \acknow{We thank TODO.
-  This research was supported by Centers for Disease Control and Prevention
-  (CDC) grant U01IP001121, and gifts from Facebook, Google, McCune, and
-  Optum.}
+  This research was supported by gifts from Facebook, Google, McCune, and
+  Optum; Centers for Disease Control and Prevention (CDC) grant U01IP001121;
+  National Science Foundation Graduate Research Fellowship Program (NSF GRFP)
+  awards DGE1745016; and the Center for Machine Learning and Health (CMLH) at
+  Carnegie Mellon University.}
 

### 2021-06-21 forecast/paper/supplement-text.tex

@@
 had simply trained all models using the finalized data rather than
-using vintage data.  This comparison can be seen most
+using vintage data.  This comparison can be seen more
 straightforwardly in Figures \ref{fig:fcast-honest-v-finalized} and
@@
 appear (misleadingly) to have better performance than they would
-actually have had if run prospectively.  For example, for forecasting case rates
+have had if run prospectively.  For example, for forecasting case rates
 7-days ahead, the WIS of all methods is at least 8\% larger than what would have been
-recorded using the finalized values of the data.  This effect
+achieved using finalized data.  This effect
 diminishes as the forecasting horizon increases, reflecting the fact
-that these forecasters rely less heavily on recent data than very
-short-term forecasters.  Crucially, some methods are
-"helped" more than others by the less scrupulous retrospective
+that longer-horizon forecasters rely less heavily on recent data than very
+short-horizon forecasters.  Crucially, some methods are
+``helped'' more than others by the less scrupulous retrospective
 evaluation, underscoring the difficulty of avoiding misleading
@@
 
-The \chngcli~indicator (along with the other claims-based signals) is the most
+\chngcli~(and, to a lesser extent, the other claims-based
+signals) is the most  
 affected by this distinction, reflecting the latency in claims-based
-reporting.  This supports the importance of efforts to provide "nowcasts" for claims signals
-(which corresponds to a 0-ahead "forecast" of what the claims
+reporting.  This underscores the importance of efforts to provide ``nowcasts'' for
+claims signals 
+(which corresponds to a 0-ahead forecast of what the claims
 signal's value will be once all data has been collected). Looking at
 the \chngcli~and \dv~curves in Figure \ref{fig:fcast-finalized}, we
-can see that they perform very similarly when trained on the finalized data.  This is reassuring
-because they are in principle measuring the same thing (namely, the percentage of
+can see that they perform very similarly when trained on the finalized data.
+This is reassuring 
+because they are, in principle, measuring the same thing (namely, the percentage of
 outpatient visits that are primarily about COVID-related symptoms).
@@
 
-While \dv~is one of the least affected methods by using finalized
-rather than vintage values, it is one of the most affected methods for the hotspot problem.
-This is a reminder that the forecasting and hotspot problems are
-distinct problems.  For example, the hotspot problem does not measure
+While using finalized rather than vintage data affects \dv~the least for forecasting,
+it is one of the most affected methods for the
+hotspot problem. 
+This is a reminder that the forecasting and hotspot problems are fundamentally
+distinct.  For example, the hotspot problem does not measure
 the ability to distinguish between flat and downward trends.
@@
 In this section, we consider using the geometric mean instead of the
-arithmetic mean when aggregating the weighted interval score (WIS) across location-time pairs.
-There are three reasons why the geometric mean may be desirable.
+arithmetic mean when aggregating the weighted interval score (WIS) across
+location-time pairs. 
+There are three reasons why using the geometric mean may be desirable.
 \begin{enumerate}
@@
   choice could be to take the mean of the ratio of WIS values for the
-  two methods. (This latter choice would penalize a method less for
-  doing poorly where the baseline forecaster also does poorly.)
+  two methods. This latter choice would penalize a method less for
+  doing poorly where the baseline forecaster also does poorly.
   Using instead the geometric mean makes the order of aggregation and
@@
   multiplicative space-time effects $S_{\ell,t}$ shared across all forecasters,
-  i.e. $\WIS(F_{\ell,t,f},Y_{\ell,t})=S_{\ell,t}E_{f}$, then
+  i.e. $\WIS(F_{\ell,t,f},Y_{\ell,t})=S_{\ell,t}E_{f,t}$ with $E_{F,t}$ a
+  forecaster-specific error, then
   taking the ratio of two forecasters' geometric mean WIS values will
@@
 we see that the main conclusions are largely unchanged; however,
-\chngcli~now does appear better than \ar.  This behavior would be
+\chngcli~now appears better than \ar.  This behavior would be
 expected if \chngcli's poor performance is attributable to a
@@
 prediction is that the indicators are not actually providing useful
-information but they are instead simply acting as a sort of ``implicit
+information but they are instead acting as a sort of ``implicit
 regularization,''  leading to shrinkage on the autoregressive
-coefficients, leading to less volatile predictions.  To investigate
+coefficients and therefore to less volatile predictions.  To investigate
 this hypothesis, we consider fitting  ``noise features'' that in truth
-should have zero coefficients.  Recall (from the main paper) that at each forecast date, we
+should have zero coefficients.  Recall (from the main paper) that at each
+forecast date, we 
 train a model on 6,426 location-time pairs.  Indicator models are
@@
 and the three lagged indicator values.  To form noise indicator features,
-we resample from other rows (while keeping the autoregressive
+we replace their values with those from a randomly chosen time-space pair
+(while keeping the autoregressive 
 features fixed).  In particular, at each location $\ell$ and
 time $t$, for the forecasting task we replace the triplet $(X_{\ell,t}, X_{\ell,t-7},
-X_{\ell,t-14})$ in Eq. (3) of the main paper with the triplet $(X_{\ell^*,t^*}, X_{\ell^*,t^*-7},
+X_{\ell,t-14})$ in Eq. (3) of the main paper with the triplet $(X_{\ell^*,t^*},
+X_{\ell^*,t^*-7}, 
 X_{\ell^*,t^*-14})$, where $(\ell^*,t^*)$ is a location-time pair
 sampled with replacement from the 6,426 location-time pairs.  
-Likewise in the hotspot prediction task, we replace the triplet $(X_{\ell,t}^\Delta, X_{\ell,t-7}^\Delta,
-X_{\ell,t-14}^\Delta)$ in Eq. (5) of the main paper with $(X_{\ell^*,t^*}^\Delta, X_{\ell^*,t^*-7}^\Delta,
+Likewise in the hotspot prediction task, we replace the triplet
+$(X_{\ell,t}^\Delta, X_{\ell,t-7}^\Delta, 
+X_{\ell,t-14}^\Delta)$ in Eq. (5) of the main paper with
+$(X_{\ell^*,t^*}^\Delta, X_{\ell^*,t^*-7}^\Delta, 
 X_{\ell^*,t^*-14}^\Delta)$.
@@
 leading us to dismiss the implicit regularization hypothesis.
-
-% # Correlations with lagged actuals
-
-% Alden's histograms are in Figure~\ref{fig:cor-wis-ratio} and Figure~\ref{fig:cor-wis-ratio-m1}.
 
@@
 the AR model tends to overpredict downswings and underpredict upswings. Adding
-indicators seems to help avoid this behavior on the downswing but not as much 
+indicators appears to help avoid this behavior on the downswing but not as much 
 on upswings. Figure \ref{fig:upswing-corr-table} shows the correlation between
@@
 paper, we end all evaluation on December 31, 2020.  Figures
-\ref{fig:fcast-alldates} -- \ref{fig:hot-alldates} show how the
+\ref{fig:fcast-alldates}--\ref{fig:hot-alldates} show how the
 results would differ if we extended this analysis through March
@@
 quite different in this period.  With a 21-day training window, it is
-common for there to be very few hotspots in training.
-
-\section{Deprecated}
-
-There are a few blocks at the bottom (figures with Google symptoms only and the
-old trajectory plots) that we can remove once we decide.
+common for there to be many fewer hotspots in training.

### 2021-06-21 forecast/paper/paper.tex
More acks

--- 2021-06-19 forecast/paper/paper.tex
@@
   Johansson, Evan Ray, Nicholas Reich, and Roni Rosenfeld for several
-  enlightening conversations about scoring and evaluation.  The research was  
-  supported by gifts from Facebook, Google, McCune, and Optum; 
-  Centers for Disease Control and Prevention (CDC) grant U01IP001121; the 
-  Canadian Statistical Sciences Institute; National Sciences and Engineering
-  Research Council of Canada (NSERC) grant RGPIN-2021-02618; and National
-  Science Foundation Graduate Research Fellowship Program (NSF GRFP) award
-  DGE175016.}
+  enlightening conversations about forecast scoring and evaluation.  
+  This material is based on work supported by gifts from Facebook, Google.org,
+  the Foundation, and Optum; Centers for Disease Control and Prevention (CDC)
+  grant U01IP001121; the Canadian Statistical Sciences Institute; National
+  Sciences and Engineering Research Council of Canada (NSERC) grant
+  RGPIN-2021-02618; and National Science Foundation Graduate Research Fellowship 
+  Program (NSF GRFP) award DGE1745016.}
 

### 2021-06-21 indicators/paper/paper.tex
More acks

--- 2021-06-19 indicators/paper/paper.tex
@@
 
-\acknow{We thank TODO.
-  This research was supported by Centers for Disease Control and Prevention
-  (CDC) grant U01IP001121, and gifts from Facebook, Google, McCune, and
-  Optum.}
+\acknow{We thank Carrie Reed, Matt Biggerstaff, Michael Johansson, Rachel
+  Slayton, Velma Lopez, Jo Walker and others on the CDC COVID-19 Modeling Team;
+  Hal Varian, Brett Slatkin, and others on Google's Surveys team and
+  Google.org's CMU-Delphi Fellowship team; Curtiss Cobb and others on Facebook's   
+  Demography and Survey Science, Data for Good and Health teams; Alex Smola and
+  others at Amazon Web Services; Tim Suther and others at Change Healthcare;
+  Paul Nielsen and others at Optum; and John Tamerius and others at Quidel.
+  This material is based on work supported by gifts from Facebook, Google.org,
+  the McCune Foundation, and Optum; Centers for Disease Control and Prevention
+  (CDC) grant U01IP001121; National Science Foundation Graduate Research
+  Fellowship Program (NSF GRFP) award DGE1745016; and the Center for Machine
+  Learning and Health (CMLH) at Carnegie Mellon University.}
 

### 2021-06-21 indicators/paper/paper.tex
Add nearly all middle authors

@@
 % Middle authors
-\author[d]{Middle Authors to be Added}
+\author[i]{Sumit Agrawal}
+\author[l]{Wael Al Saeed}
+\author[g]{Taylor Arnold}
+\author[j]{Amartya Basu}
+\author[d]{Jacob Bien}
+\author[r]{Ángel Alexander Cabrera}
+\author[b]{Andrew Chin}
+\author[b]{Eu Jing Chua}
+\author[b]{Brian Clark}
+\author[i]{Sarah Colquhoun}
+\author[b]{Nat DeFries}
+\author[i]{David C. Farrow}
+\author[r]{Jodi Forlizzi}
+\author[i]{Jed Grabman}
+\author[b]{Samuel Gratzl}
+\author[a]{Alden Green}
+\author[r]{Robin Han}
+\author[i]{Kate Harwood}
+\author[a,b]{Addison J. Hu}
+\author[i]{Raphael Hyde}
+\author[d]{Sangwon Hyun}
+\author[l]{Ananya Joshi}
+\author[h]{Jimi Kim}
+\author[r]{Andrew Kuznetsov}
+\author[b]{Wichada La~Motte-Kerr}
+\author[f]{Balasubramanian Narasimhan}
+\author[r]{Jenny Yeon Jin Lee}
+\author[p]{Kenneth Lee}
+\author[b]{Zachary C. Lipton}
+\author[r]{Michael Xieyang Liu}
+\author[k]{Lester Mackey}
+\author[b]{Kathryn Mazaitis}
+\author[o]{Daniel J. McDonald}
+\author[i]{Phillip McGuinness}
+\author[i]{Michael P. O'Brien}
+\author[a,b]{Natalia L. Oliveira}
+\author[a,b]{Pratik Patil}
+\author[r]{Adam Perer}
+\author[b]{Collin A. Politsch}
+\author[l]{Dawn Rucker}
+\author[i]{Chris Scott}
+\author[s]{Nigam Shah}
+\author[n]{Vishnu Shankar}
+\author[p]{James Sharpnack}
+\author[b]{Dmitry Shemetov}
+\author[q]{Noah Simon}
+\author[i]{Benjamin Y. Smith}
+\author[b]{Vishakha Srivastava}
+\author[e,f]{Robert Tibshirani}
+\author[e]{Elena Tuzhilina}
+\author[b]{Ana Karina Van Nortwick}
+\author[a]{Valérie Ventura}
+\author[a]{Larry Wasserman}
+\author[i]{Benjamin Weaver}
+\author[m]{Jeremy C. Weiss}
+\author[i]{Spencer Whitman}
+\author[r]{Kristin Williams}
 
@@
 \affil[c]{Computational Biology Department, Carnegie Mellon University}
-\affil[d]{Remaining affiliations to be added}
+\affil[d]{Department of Data Sciences and Operations, University of Southern California}
+\affil[e]{Department of Statistics, Stanford University}
+\affil[f]{Department of Biomedical Data Science, Stanford University}
+\affil[g]{Linguistics Program, University of Richmond}
+\affil[h]{School of Natural Sciences and Mathematics, University of Texas at Dallas}
+\affil[i]{Google.org}
+\affil[j]{Information Networking Institute, Carnegie Mellon University}
+\affil[k]{Microsoft Research New England}
+\affil[l]{Computer Science Department, Carnegie Mellon University}
+\affil[m]{Heinz College of Information Systems and Public Policy, Carnegie
+  Mellon University}
+\affil[n]{Program in Immunology, Stanford University School of Medicine}
+\affil[o]{Department of Statistics, University of British Columbia}
+\affil[p]{Department of Statistics, University of California, Davis}
+\affil[q]{Department of Biostatistics, University of Washington}
+\affil[r]{Human-Computer Interaction Institute, Carnegie Mellon University}
+\affil[s]{Department of Medicine, Stanford University}
 

### 2021-06-21 indicators/paper/supplement.Rmd
Recreate/Adapt Fig 1--3 in paper to hosp data, put in supplement

@@
 
+\clearpage
+
+```{r hospitalization_time_trends_national, fig.cap="National trends, from April 2020 to April 2021, of HHS confirmed COVID-19 hospital admissions, plus three other signals in the COVIDcast API.
+Here, the HHS data has been limited to values from 2020-08-01 onward due to changes in reporting behavior, and processed into a 7-day trailing average rate per 100,000 resident population using Census Bureau estimates for 2019.
+HSP-HOSP is the percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners, smoothed in time and adjusted for systematic day-of-week effects.
+Auxiliary signals have been scaled to have the same maximum value as the HHS data."}
+start_day = "2020-04-15"
+end_day = "2021-04-15"
+as_of_day = "2021-06-01"
+
+HOSP_SOURCES <-
+  c("hhs", "hospital-admissions", "jhu-csse", "chng")
+HOSP_SIGNALS <-
+  c("confirmed_admissions_covid_1d",
+    "smoothed_adj_covid19_from_claims",
+    "confirmed_7dav_incidence_prop",
+    "smoothed_adj_outpatient_covid")
+hosp_names = c(
+  "Hospitalizations", "HSP-HOSP",
+  "Cases", "CHNG-COVID"
+)
+hosp_signal_has_national = c(
+  TRUE, TRUE,
+  TRUE, TRUE
+)
+hhs_start_date = as.Date("2020-08-01")
+
+hosp_colors = c("black", scales::hue_pal()(length(HOSP_SIGNALS)-1))
+
+hosp_raw_signal_data <- covidcast_signals(
+  HOSP_SOURCES[hosp_signal_has_national],
+  HOSP_SIGNALS[hosp_signal_has_national],
+  start_day, end_day,
+  as_of=as_of_day,
+  geo_type="nation"
+)
+
+## For HHS data, remove pre-cutoff-date period, and take a 7-day rolling average rate:
+postprocess_hhs_hosp_df <- function(hhs_hosp_df, geo_type, population_df) {
+  stopifnot(nrow(distinct(hhs_hosp_df, data_source, signal)) == 1)
+  hhs_hosp_df %>%
+    filter(time_value >= hhs_start_date) %>%
+    group_by(geo_value) %>%
+    arrange(time_value) %>%
+    mutate(value = c(rep(NA, 6), sum(value[1:7]) + c(0, cumsum(tail(value,-7) - head(value, -7))))/7) %>% # (could also use zoo::rollmeanr)
+    slice(-(1:6)) %>%
+    left_join(population_df) %>%
+    mutate(value = value/population*100000, population=NULL) %>%
+    ungroup() %>%
+    {as.covidcast_signal(., signal=unique(.$signal), geo_type=geo_type, time_type="day")}
+}
+hosp_signal_data <- hosp_raw_signal_data
+hosp_signal_data[[1]] <- postprocess_hhs_hosp_df(
+  hosp_raw_signal_data[[1]], "nation",
+  ## use national entry appearing in state_census
+  state_census %>% transmute(geo_value=tolower(ABBR), population=POPESTIMATE2019)
+)
+
+hosp_main_signal_max <- max(hosp_signal_data[[1]]$value, na.rm=TRUE)
+
+hosp_rescaled <- aggregate_signals(hosp_signal_data, format = "long") %>%
+  group_by(signal) %>%
+  mutate(value = value / max(value, na.rm=TRUE) * hosp_main_signal_max) %>%
+  ungroup()
+
+hosp_rescaled$signal <- factor(hosp_rescaled$signal, levels=HOSP_SIGNALS, labels=hosp_names)
+
+ggplot(hosp_rescaled, aes(x = time_value, y = value, color = signal)) +
+  geom_line() +
+  scale_color_manual(breaks = hosp_names, values = hosp_colors) +
+  labs(x = "Date", y = "Signal value (scaled)") +
+  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
+               date_minor_breaks = "1 month") +
+  theme_bw() +
+  theme(legend.position = "bottom", legend.title = element_blank())
+```
+
+```{r hosp-correlation-data}
+df_hosp_signals = vector("list", length(HOSP_SIGNALS))
+for (i in seq_along(HOSP_SIGNALS)) {
+  df_hosp_signals[[i]] =
+    covidcast_signal(HOSP_SOURCES[i], HOSP_SIGNALS[i],
+                     start_day, end_day, as_of=as_of_day,
+                     geo_type="state")
+}
+state_populations = state_census %>%
+  transmute(geo_value=tolower(ABBR), population=POPESTIMATE2019)
+df_hosp_signals[[1]] <- postprocess_hhs_hosp_df(df_hosp_signals[[1]], "state", state_populations)
+
+## Fetch HHS confirmed hospital admissions and smooth with 7-day
+## trailing average rate:
+df_raw_hospitalizations =
+  covidcast_signal("hhs", "confirmed_admissions_covid_1d",
+                   start_day, end_day, as_of=as_of_day,
+                   geo_type="state")
+df_hospitalizations = postprocess_hhs_hosp_df(df_raw_hospitalizations, "state", state_populations)
+
+## Consider only state-like jurisdictions for which all signals of interest have at least 50 rows:
+geo_values = unique(df_hospitalizations$geo_value)
+for (i in seq_along(HOSP_SIGNALS)) {
+  geo_values = intersect(geo_values, df_hosp_signals[[i]] %>% count(geo_value) %>% filter(n > 50) %>% pull(geo_value))
+}
+
+cat('Number of unique jurisdictions included:\n')
+length(geo_values)
+```
+
+Correlations sliced by time (geo-wise correlation):
+```{r hosp-correlations-by-time, fig.cap="Geo-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated for all times with sufficient available data within this period, over all state-like jurisdictions for which each signal was reported on at least 50 days during this period."}
+n_hosp_signals = length(HOSP_SIGNALS)
+df_list_hosp_geowise_cor = vector("list", n_hosp_signals)
+for (i in 1:n_hosp_signals) {
+  df_list_hosp_geowise_cor[[i]] = covidcast_cor(
+    df_hosp_signals[[i]] %>% filter(geo_value %in% geo_values),
+    df_hospitalizations %>% filter(geo_value %in% geo_values),
+    by = "time_value", method = "spearman"
+  )
+  df_list_hosp_geowise_cor[[i]]$signal = hosp_names[[i]]
+}
+df_long_hosp_geowise_cor = do.call(rbind, df_list_hosp_geowise_cor[-1])
+df_long_hosp_geowise_cor$signal = factor(df_long_hosp_geowise_cor$signal, hosp_names[-1])
+
+ggplot(df_long_hosp_geowise_cor, aes(x = time_value, y = value)) +
+  geom_line(aes(color = signal)) +
+  scale_color_manual(breaks = hosp_names, values = hosp_colors) +
+  expand_limits(y=0) +
+  labs(x = "Date", y = "Correlation") +
+  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
+               date_minor_breaks = "1 month") +
+  theme_bw() +
+  theme(legend.position = "bottom", legend.title = element_blank())
+```
+
+Correlations sliced by state (time-series correlation)
+```{r hosp-correlations-by-state, fig.cap="Time-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated over all state-like jurisdictions for which each signal was reported on at least 50 days during this period."}
+df_list_hosp_timewise_cor = vector("list", length(HOSP_SIGNALS))
+for (i in seq_along(HOSP_SIGNALS)) {
+  df_list_hosp_timewise_cor[[i]] = covidcast_cor(
+    df_hosp_signals[[i]] %>% filter(geo_value %in% geo_values),
+    df_hospitalizations %>% filter(geo_value %in% geo_values),
+    by = "geo_value", method = "spearman"
+  )
+  df_list_hosp_timewise_cor[[i]]$signal = hosp_names[[i]]
+}
+df_long_hosp_timewise_cor = do.call(rbind, df_list_hosp_timewise_cor[-1])
+df_long_hosp_timewise_cor$signal = factor(df_long_hosp_timewise_cor$signal, hosp_names[-1])
+
+ggplot(df_long_hosp_timewise_cor, aes(value)) +
+  geom_density(aes(color = signal, fill = signal), alpha = 0.4) +
+  labs(x = "Correlation", y = "Density") +
+  expand_limits(x=c(-1,1)) +
+  scale_color_manual(breaks = hosp_names, values = hosp_colors) +
+  scale_fill_manual(breaks = hosp_names, values = hosp_colors) +
+  theme_bw() +
+  theme(legend.position = "bottom", legend.title = element_blank())
+```
+
 <!-- PNAS says: Add this line AFTER all your figures and tables -->

### 2021-06-21 forecast/paper/paper.tex
Authors, affiliations

@@
 % Writing group
-\author[a,1]{Daniel J.\ McDonald}
+\author[a,1]{Daniel J. McDonald}
 \author[b,2]{Jacob Bien}
 \author[c,2]{Alden Green}
-\author[c,d,2]{Addison Hu}
+\author[c,d,2]{Addison J. Hu}
 
@@
 \author[b]{Sangwon Hyun}
-\author[c,d]{Natalia L.\ Oliveira}
+\author[c,d]{Natalia L. Oliveira}
 \author[e]{James Sharpnack}
 \author[f]{Jingjing Tang}
-\author[g]{Robert Tibshirani}
-\author[c]{Valerie Ventura}
+\author[g,h]{Robert Tibshirani}
+\author[c]{Val{\'e}rie Ventura}
 \author[c,d]{Larry Wasserman}
@@
 % Senior authors
-\author[c,d]{Ryan J.\ Tibshirani}
+\author[c,d]{Ryan J. Tibshirani}
 
 \affil[a]{Department of Statistics, University of British Columbia}
-\affil[b]{Data Sciences and Operations, University of Southern California}
+\affil[b]{Department of Data Sciences and Operations, University of Southern
+  California} 
 \affil[c]{Department of Statistics \& Data Science, Carnegie Mellon University}
@@
 \affil[g]{Department of Statistics, Stanford University}
+\affil[h]{Department of Biomedical Data Science, Stanford University}
 
@@
   addressed. E-mail: daniel@stat.ubc.ca} 
-\equalauthors{\textsuperscript{2}J.B., A.G., and A.H.\ contributed equally to
+\equalauthors{\textsuperscript{2}J.B., A.G., and A.J.H. contributed equally to 
   this work.}

### 2021-06-21 forecast/paper/paper.tex
Rebuilt figs; finished supplement draft; added Justin's suggestion

--- 2021-06-21 forecast/paper/paper.tex
@@
   % https://www.internationalscienceediting.com/how-to-write-a-figure-caption/
-  \caption{Revision behavior for two signals in the HRR containing New York  
-    City.  Each colored line corresponds to the data as reported on a particular 
+  \caption{Revision behavior for two signals in the HRR containing Charlotte,
+    North Carolina.  Each colored line corresponds to the data as reported on a
+    particular 
     date (\textit{as of} dates varying from September 28 through October
@@
 alternative explanation is that the indicators are simply providing \textit{implicit
-  regularization} on top of  the basic AR model when we include them as lagged
+  regularization} on top of the basic AR model, in the same way any noise
+variable might, when we include them as lagged
 features in \eqref{eq:forecast_ar_x} and \eqref{eq:hotspot_ar_x}.
@@
   Sciences and Engineering Research Council of Canada (NSERC) grant
-  RGPIN-2021-02618; and National Science Foundation Graduate Research Fellowship 
-  Program (NSF GRFP) award DGE1745016.}
+  RGPIN-2021-02618; and National Science Foundation Graduate Research
+  Fellowship Program (NSF GRFP) award DGE1745016.}
 

### 2021-06-21 forecast/paper/supplement-text.tex
Rebuilt figs; finished supplement draft; added Justin's suggestion

@@
 
-Currently, both figures are in the manuscript. Probably just need text here.
+In Section 2.D of the main text, we discuss the extent to which the indicators
+are leading or lagging case rates during different periods. To define the amount
+of leadingness or laggingness, we use the cross correlation function (CCF)
+between the two time series. The CCF is of an indicator $X_{\ell,t}$ and case
+rates $Y_{\ell,t}$ is defined as
+\[
+\CCF_{\ell,t}(a) = \frac{1}{n_a} \sum_{i=1}^{n_a} \left( ( X_{\ell,t,i+a} -
+  \overline{X}_{\ell,t}) / s^{X}_{\ell,t}\right) \left( ( Y_{\ell,t,i} -
+  \overline{Y}_{\ell,t}) / s^{Y}_{\ell,t}\right), 
+\]
+where $n_a$ is the number of available time points when $X_{\ell,t}$ has been
+shifted in time by $a$ days, and $\overline{X}_{\ell,t}$, $s^X_{\ell,t}$
+are the sample mean and
+standard deviation of $X_{\ell,t}$ (respectively $Y_{\ell,t}$). The result is a
+sequence of Pearson correlations for each selected $a$, where we align
+$X_{\ell,t}$ with the values of $Y_{\ell,t}$ that occurred $a$ days earlier.
+Thus, for $a>0$, $\CCF_{\ell,t}(a)>0$ indicates that
+$Y_{\ell,t}$ is moving together with the values of $X_{\ell,t}$ that occur $a$
+days in the future. In this case we say that $X_{\ell,t}$ is lagging
+$Y_{\ell,t}$. For $a<0$, $\CCF_{\ell,t}(a)>0$ means that $Y_{\ell,t}$ is
+positively correlated with $X_{\ell,t}$ shifted earlier, so we say that
+$X_{\ell,t}$ leads $Y_{\ell,t}$.
+
+Figure \ref{fig:ccf-dv-finalized} shows the standardized signals for the HRR
+containing Charlotte, North Carolina, from August 1, 2020 until the end of
+September. These are the same signals shown in Figure 1 in the manuscript. To
+define ``leadingness'' we compute $\CCF_{\ell,t}(a)$ for each $a\in
+\{-15,\ldots,15\}$ using the 56 days leading up to time $t$ (this is the same
+amount of data used to train the forecasters: 21 days of training data, 21 days
+to get the response at $a=21$, and 14 days for the longest lagged value). The
+orange dashed horizontal line represents the 97.5\% significance threshold for
+correlations based on 56 observations. Any correlations larger in magnitude than
+this value are considered statistically significant under the null hypothesis of
+no relationship. We define leadingness to be the sum of the significant
+correlations that are leading (those above the dashed line with $a<0$) while
+laggingness is the same but for $a>0$. In this case, there are no significant
+correlations, so both scores will be 0: on September 28, \dv~is neither
+leading nor lagging cases.
+
+Figure \ref{fig:lagging-only} shows the correlation between laggingness and the
+difference in indicator WIS and AR WIS. Unlike leadingness (Figure 5 in the
+manuscript) there is no obvious relationship. This is heartening as laggingness
+should not aid forecasting performance. On the other hand, if an indicator is
+more lagging than it is leading, this may suggest diminished performance.
+Figure~\ref{fig:diff-in-lead-lag} shows the correlation of the difference in
+leadingness and laggingness with the difference in WIS. The pattern here is
+largely similar to the pattern in leadingness described in the manuscript: the
+relationship is strongest in down periods and weakest in up periods with the
+strength diminishing as we move from down to flat to up for all indicators.
 

### 2021-06-21 forecast/paper/supplement.Rmd
typos and minor supplement corrections
Rebuilt figs; finished supplement draft; added Justin's suggestion

@@
 
-```{r lagging-only, eval=FALSE}
+```{r ccf-dv-finalized, eval=TRUE, fig.cap="Illustration of the cross-correlation function between DV-CLI and cases.
+The left panel shows the standardized signals over the period from August 1 to September 28 (as of September 28, 2020).
+The right panel shows $\\CCF(a)$ for different values of $a$ as vertical blue bars.
+The orange dashed lines indicate the 95\\% significance threshold.
+By our leadingness metric, DV-CLI is neither leading nor lagging cases over this period."}
+source("../code/ccf-dv-finalized.R")
+gg
+```
+
+\clearpage
+
+```{r lagging-only, fig.cap="Correlation of the difference in WIS with the  laggingness of the indicator at the target date, stratified by up, down, or flat period.
+Compare to Figure 5 in the manuscript."}
 # Deprecated
@@
 
-```{r diff-in-lead-lag, include=FALSE}
-# Figure 7 in the manuscript
+\clearpage
+
+```{r diff-in-lead-lag, fig.cap = "Correlation of the difference between leadingness and laggingness with the difference in WIS.
+The relationship is essentially the same as described in the manuscript and shown in Figure 5."}
+# Deprecated
 df2 %>%
@@
   theme(legend.position = "bottom", legend.title = element_blank())
+```
+
+\clearpage
+
+<!-- Figures in the manuscript in their own scripts -->
+
+```{r revisions-dv-jhu, eval=FALSE}
+# Figure 1 in the manuscript
+source("../code/revisions.R")
+```
+
+```{r ny-trajectory, eval=FALSE}
+# Figure 2 in the manuscript
+source("../code/ny-trajectory.R")
+gg
 ```
@@
 \clearpage
-
-<!-- Figures in the manuscript in their own scripts -->
-
-```{r revisions-dv-jhu, eval=FALSE}
-# Figure 1 in the manuscript
-source("../code/revisions.R")
-```
-
-```{r ny-trajectory, eval=FALSE}
-# Figure 2 in the manuscript
-source("../code/ny-trajectory.R")
-gg
-```
-
-```{r ccf-dv-finalized, eval=FALSE}
-# Figure ?? in the manuscript
-source("../code/ccf-dv-finalized.R")
-gg
-```
 
@@
 
-\clearpage
-
 <!-- PNAS says: Add this line AFTER all your figures and tables -->
@@
 
-\movie{Type legend for the movie here.}
-
 \dataset{dataset_one.txt}{Type or paste legend here.}

### 2021-06-21 forecast/paper/supplement.tex
typos and minor supplement corrections
Rebuilt figs; finished supplement draft; added Justin's suggestion

@@
 
+{\centering \includegraphics[width=\textwidth]{fig/ccf-dv-finalized-1}
+
+}
+
+\caption{Illustration of the cross-correlation function between DV-CLI and cases.
+The left panel shows the standardized signals over the period from August 1 to September 28 (as of September 28, 2020).
+The right panel shows $\CCF(a)$ for different values of $a$ as vertical blue bars.
+The orange dashed lines indicate the 95\% significance threshold.
+By our leadingness metric, DV-CLI is neither leading nor lagging cases over this period.}\label{fig:ccf-dv-finalized}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/lagging-only-1}
+
+}
+
+\caption{Correlation of the difference in WIS with the  laggingness of the indicator at the target date, stratified by up, down, or flat period.
+Compare to Figure 5 in the manuscript.}\label{fig:lagging-only}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/diff-in-lead-lag-1}
+
+}
+
+\caption{Correlation of the difference between leadingness and laggingness with the difference in WIS.
+The relationship is essentially the same as described in the manuscript and shown in Figure 5.}\label{fig:diff-in-lead-lag}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
 {\centering \includegraphics[width=\textwidth]{fig/fcast-alldates-1}
@@
 
-\clearpage
-
 \FloatBarrier
-
-\movie{Type legend for the movie here.}
 

### 2021-06-21 forecast/code/ccf-dv-finalized.R

@@ -1,11 +1,11 @@
 sigs <- covidcast_signals(
   data_source = c("jhu-csse", "doctor-visits"),
   signal = c("confirmed_7dav_incidence_prop", "smoothed_adj_cli"),
-  start_day = "2020-10-01",
-  end_day = "2020-12-15",
-  as_of = "2021-05-15", #2020-12-21",
+  start_day = "2020-08-01",
+  end_day = "2020-09-28",
+  as_of = "2020-09-28", #2020-12-21",
   geo_type = "hrr",
-  geo_values = "303") %>% # NYC
+  geo_values = "311") %>% # NYC
   aggregate_signals(format = "wide")
 
 nas <- complete.cases(sigs)
@@ -15,7 +15,7 @@ names(sigs)[3:4] <- c("cases", "doctor-visits")
 
 std_sigs <- sigs %>%
   mutate(across(c("cases","doctor-visits"), ~ (.x - mean(.x)) / sd(.x))) %>%
-  rename(`JHU-CSSE` = cases, `DV-CLI` = `doctor-visits`) %>%
+  rename(Cases = cases, `DV-CLI` = `doctor-visits`) %>%
   select(-geo_value) %>%
   pivot_longer(-time_value)
   
@@ -23,24 +23,25 @@ g1 <- std_sigs %>%
   ggplot(aes(time_value, value, color = name)) +
   geom_line() +
   theme_bw() +
-  xlab("Date") + ylab("Standardized signal") +
+  ylab("Standardized signal") +
+  xlab("") +
   scale_color_viridis_d(begin=.25, end=.75, name = "") +
   theme(legend.position = "bottom")
 
 cc <- ccf(sigs$`doctor-visits`, sigs$cases, plot = FALSE)
 g2 <- ggplot(tibble(lag = drop(cc$lag), ccf = drop(cc$acf)), aes(lag)) +
   geom_segment(aes(xend = lag, yend = ccf), y = 0, color = "cornflowerblue") +
-  geom_hline(yintercept = c(1,-1) * qnorm((1 + .95) / 2) / sqrt(cc$n.used),
+  geom_hline(yintercept = qnorm((1 + .95) / 2) / sqrt(cc$n.used),
              color = "darkorange", linetype = "dashed") +
   geom_hline(yintercept = 0) +
   geom_vline(xintercept = 0) +
   ylab("Cross correlation") +
-  xlab("Lag") +
+  xlab("Lag (a)") +
[…]

### 2021-06-21 indicators/paper/paper.tex
Small typo fixes and corrections

@@
 \author[d]{Jacob Bien}
-\author[r]{Ángel Alexander Cabrera}
+\author[r]{{\'A}ngel A. Cabrera}
 \author[b]{Andrew Chin}
@@
 \author[r]{Andrew Kuznetsov}
-\author[b]{Wichada La~Motte-Kerr}
-\author[f]{Balasubramanian Narasimhan}
-\author[r]{Jenny Yeon Jin Lee}
+\author[b]{Wichada La Motte-Kerr}
+\author[e,f]{Balasubramanian Narasimhan}
+\author[r]{Yeon Jin Lee}
 \author[p]{Kenneth Lee}
 \author[b]{Zachary C. Lipton}
-\author[r]{Michael Xieyang Liu}
+\author[r]{Michael X. Liu}
 \author[k]{Lester Mackey}
@@
 \author[b]{Ana Karina Van Nortwick}
-\author[a]{Valérie Ventura}
+\author[a]{Val{\'e}rie Ventura}
 \author[a]{Larry Wasserman}
@@
 \author[b]{Roni Rosenfeld}
-\author[a,b]{Ryan J.\ Tibshirani}
+\author[a,b]{Ryan J. Tibshirani}
 
@@
 from smartphone apps, calculates COVID-related mobility signals
-\cite{SafeGraphSocial, SafeGraphPatterns} and makes it available to researchers 
+\cite{SafeGraphSocial, SafeGraphPatterns} and makes them available to researchers
 under a data use agreement; we aggregate (some of) these signals to the county
@@
   Doerr:2021}. The API currently serves hundreds of thousands of requests to
-thousands of users every day.
-
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/time_trends_national.pdf}
-  \caption{National trends, from April 2020 to April 2021, of four signals in 
-    the COVIDcast API. The auxiliary signals, based on medical claims data and 
-    massive surveys, track changes in officially reported cases quite
-    well. (They have all been placed on the same scale as reported cases per 
-    100,000 people.)}
-  \label{fig:time_trends_national}
-\end{figure}
+hundreds of users every day.
 
@@
   (CTIS-CLI-in-community): The estimated percentage of the population who know 
-  someone in their local community that is sick, based on the same surveys.
+  someone in their local community who is sick, based on the same surveys.
 \item Quidel test positivity rate (Quidel-TPR): The percentage of positive
@@
 \end{itemize}
-
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/geo_wise_corr.pdf}
-  \caption{Geo-wise correlations with case rates, from April 15, 2020 to April 
-    15, 2021, calculated over all counties for which all signals were available
-    and which had at least 500 cumulative cases by the end of this period.}
-  \label{fig:geo_wise_correlation} 
-\end{figure}
 
@@
 analysis in the next subsection.
+
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{fig/time_trends_national.pdf}
+  \caption{National trends, from April 2020 to April 2021, of four signals in 
+    the COVIDcast API. The auxiliary signals, based on medical claims data and 
+    massive surveys, track changes in officially reported cases quite
+    well. (They have all been placed on the same scale as reported cases per 
+    100,000 people.)}
+  \label{fig:time_trends_national}
+\end{figure}
 
@@
 
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/time_wise_correlation.pdf}
-  \caption{Time-wise correlations with case rates, from April 15, 2020 to April
-    15, 2021, calculated over all counties for which all signals were available
-    and which had at least 500 cumulative cases by the end of this period.}
-  \label{fig:time_wise_correlation}
-\end{figure}
-
 Figure~\ref{fig:geo_wise_correlation} shows the geo-wise correlations achieved
@@
 
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{fig/geo_wise_corr.pdf}
+  \caption{Geo-wise correlations with case rates, from April 15, 2020 to April 
+    15, 2021, calculated over all counties for which all signals were available
+    and which had at least 500 cumulative cases by the end of this period.}
+  \label{fig:geo_wise_correlation} 
+\end{figure}
+
 \begin{itemize}
@@
 \subsection{Helping Robustness}
-
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/bexar_compare.pdf}
-  \caption{Reported cases per day in Bexar County, Texas during the summer of
-    2020. On July 16, 4,810 backlogged cases were reported, though they actually
-    occurred over the preceding two weeks (this shows up as a prolonged spike in
-    the left panel due to the 7-day trailing averaging applied to the case
-    counts). Daily CTIS estimates of CLI-in-community showed more stable 
-    underlying trends.}
-  \label{fig:bexar_compare}
-\end{figure}
 
@@
 
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{fig/time_wise_correlation.pdf}
+  \caption{Time-wise correlations with case rates, from April 15, 2020 to April
+    15, 2021, calculated over all counties for which all signals were available
+    and which had at least 500 cumulative cases by the end of this period.}
+  \label{fig:time_wise_correlation}
+\end{figure}
+
 As a further example, Figure~\ref{fig:bexar_compare} displays data from Bexar
 County, Texas (which contains San Antonio) during July 2020. On July 16, 2020,
-4,810 backlogged cases were reported after reporting problems prevented them
-from being reported over the past two weeks \cite{Palacios:2021}, resulting in a
-clearly visible spike in the left-hand panel of Figure~\ref{fig:bexar_compare}
-(case data from JHU CSSE, smoothed using a 7-day trailing average).  Meanwhile,
-Delphi's COVID Trends and Impact Survey averaged around 350 responses per day in
-Bexar County over the same time period, and was able to estimate the fraction of
-the population who know someone in their community with COVID-Like Illness
-(CLI). As we can see in the right-hand panel of Figure~\ref{fig:bexar_compare},
-this signal was not affected by Bexar County's reporting problems and, as shown
-in the last subsection, it is (in general) highly correlated with case rates,
-providing an alternate stream of data about COVID activity unaffected by
-backlogs. Similar reporting problems have occurred in many jurisdictions across
-the United States, making it valuable to cross-check against external sources
-not part of the same reporting systems.
+San Antonio reported 4,810 backlogged cases after reporting problems prevented
+them from being reported over the past two weeks \cite{Palacios:2021}, resulting
+in a clearly visible spike in the left-hand panel of
+Figure~\ref{fig:bexar_compare} (case data from JHU CSSE, smoothed using a 7-day
+trailing average). Meanwhile, Delphi's COVID Trends and Impact Survey averaged
+around 350 responses per day in Bexar County over the same time period, and was
+able to estimate the fraction of the population who know someone in their
+community with COVID-Like Illness (CLI). As we can see in the right-hand panel
+of Figure~\ref{fig:bexar_compare}, this signal was not affected by Bexar
+County's reporting problems and, as shown in the last subsection, it is (in
+general) highly correlated with case rates, providing an alternate stream of
+data about COVID activity unaffected by backlogs. Similar reporting problems
+have occurred in many jurisdictions across the United States, making it valuable
+to cross-check against external sources not part of the same reporting systems.
+
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{fig/bexar_compare.pdf}
+  \caption{Reported cases per day in Bexar County, Texas during the summer of
+    2020. On July 16, 4,810 backlogged cases were reported, though they actually
+    occurred over the preceding two weeks (this shows up as a prolonged spike in
+    the left panel due to the 7-day trailing averaging applied to the case
+    counts). Daily CTIS estimates of CLI-in-community showed more stable 
+    underlying trends.}
+  \label{fig:bexar_compare}
+\end{figure}
 
 \subsection{Revisions Matter}
-
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/dv_as_of.pdf}
-  \caption{Estimated percentage of doctor's visits due to COVID-like illness
-    displayed across multiple issue dates, with later issue dates adding
-    additional data and revising past data from prior issue dates.}
-  \label{fig:dv_as_of}
-\end{figure}
 
@@
 significantly biased downward in New York throughout June and July 2020.
+
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{fig/dv_as_of.pdf}
+  \caption{Estimated percentage of doctor's visits due to COVID-like illness
+    displayed across multiple issue dates, with later issue dates adding
+    additional data and revising past data from prior issue dates.}
+  \label{fig:dv_as_of}
+\end{figure}
 
@@
 Third, the underlying database tracks all revisions made to the data, allowing
-us to query the API to learn ``what was known when'', which is critical for
+us to query the API to learn ``what was known when,'' which is critical for
 understanding the behavior (and potential pitfalls) of real-time surveillance

### 2021-06-21 forecast/paper/paper.tex
Update acks, other small edits

--- 2021-06-21 forecast/paper/paper.tex
@@
   (CTIS-CLI-in-community): The estimated percentage of the population who know
-  someone in their local community that is sick, based on Delphi's COVID Trends
+  someone in their local community thatwho is sick, based on Delphi's COVID Trends
   and Impact Survey, in partnership with Facebook.
@@
   % https://www.internationalscienceediting.com/how-to-write-a-figure-caption/
-  \caption{Revision behavior for two signals in the HRR containing New York  
-    City.  Each colored line corresponds to the data as reported on a particular 
-    date (\textit{as of} dates varying from September 28 through October
-    19). The left panel shows the DV-CLI signal, which was regularly revised
-    throughout the period, though the effects fade as we look further back in   
-    time. In contrast, the right panel shows case rates reported by JHU CSSE
-    (smoothed with a 7-day trailing average), which remain ``as reported'' on
-    September 28, with a spike towards the end of this period, until a major
+  \caption{Revision behavior for two indicators in the HRR containing Charlotte, 
+    North Carolina.  Each colored line corresponds to the data as reported on a
+    particular date (\textit{as of} dates varying from September 28 through
+    October 19). The left panel shows the DV-CLI signal, which was regularly
+    revised throughout the period, though the effects fade as we look further
+    back in time. In contrast, the right panel shows case rates reported by JHU
+    CSSE (smoothed with a 7-day trailing average), which remain ``as reported''
+    on September 28, with a spike towards the end of this period, until a major 
     correction is made on October 19, which brings this down and affects all
@@
 alternative explanation is that the indicators are simply providing \textit{implicit
-  regularization} on top of  the basic AR model when we include them as lagged
-features in \eqref{eq:forecast_ar_x} and \eqref{eq:hotspot_ar_x}.
+  regularization} on top of the basic AR model, in the same way any noise
+variable might, when we include them as lagged features in
+\eqref{eq:forecast_ar_x} and \eqref{eq:hotspot_ar_x}.
 
@@
   This material is based on work supported by gifts from Facebook, Google.org,
-  the Foundation, and Optum; Centers for Disease Control and Prevention (CDC)
-  grant U01IP001121; the Canadian Statistical Sciences Institute; National
+  the McCune Foundation, and Optum; Centers for Disease Control and Prevention
+  (CDC) grant U01IP001121; the Canadian Statistical Sciences Institute; National 
   Sciences and Engineering Research Council of Canada (NSERC) grant
-  RGPIN-2021-02618; and National Science Foundation Graduate Research Fellowship 
-  Program (NSF GRFP) award DGE1745016.}
+  RGPIN-2021-02618; and National Science Foundation Graduate Research
+  Fellowship Program (NSF GRFP) award DGE1745016.}
 

### 2021-06-21 forecast/paper/paper.tex
Copyedit ack

@@
   Johansson, Evan Ray, Nicholas Reich, and Roni Rosenfeld for several
-  enlightening conversations about forecast scoring and evaluation.  
+  enlightening conversations about forecasting, scoring, and evaluation.  
   This material is based on work supported by gifts from Facebook, Google.org,

### 2021-06-21 indicators/paper/paper.tex

@@
 from smartphone apps, calculates COVID-related mobility signals
-\cite{SafeGraphSocial, SafeGraphPatterns} and makes them available to researchers
-under a data use agreement; we aggregate (some of) these signals to the county
-level and make them publicly available.
+\cite{SafeGraphSocial, SafeGraphPatterns} and makes them available to
+researchers under a data use agreement; we aggregate (some of) these signals to
+the county level and make them publicly available.
 
@@
   Doerr:2021}. The API currently serves hundreds of thousands of requests to
-hundreds of users every day.
+thousands of users every day.
 
@@
 them from being reported over the past two weeks \cite{Palacios:2021}, resulting
-in a clearly visible spike in the left-hand panel of
-Figure~\ref{fig:bexar_compare} (case data from JHU CSSE, smoothed using a 7-day
-trailing average). Meanwhile, Delphi's COVID Trends and Impact Survey averaged
-around 350 responses per day in Bexar County over the same time period, and was
-able to estimate the fraction of the population who know someone in their
-community with COVID-Like Illness (CLI). As we can see in the right-hand panel
-of Figure~\ref{fig:bexar_compare}, this signal was not affected by Bexar
-County's reporting problems and, as shown in the last subsection, it is (in
-general) highly correlated with case rates, providing an alternate stream of
-data about COVID activity unaffected by backlogs. Similar reporting problems
-have occurred in many jurisdictions across the United States, making it valuable
-to cross-check against external sources not part of the same reporting systems.
+in a clearly visible spike in the left-hand panel of the figure (case data from
+JHU CSSE, smoothed using a 7-day trailing average). Meanwhile, Delphi's COVID
+Trends and Impact Survey averaged around 350 responses per day in Bexar County
+over the same time period, and was able to estimate the fraction of the
+population who know someone in their local community with COVID-Like Illness
+(CLI). As we can see in the right-hand panel of the figure, this signal was not
+affected by Bexar County's reporting problems and, as shown in the last
+subsection, it is (in general) highly correlated with case rates, providing an
+alternate stream of data about COVID activity unaffected by backlogs. Similar
+reporting problems have occurred in many jurisdictions across the United States,
+making it valuable to cross-check against external sources not part of the same
+reporting systems.
 
@@
   Slayton, Velma Lopez, Jo Walker and others on the CDC COVID-19 Modeling Team; 
-  Hal Varian, Brett Slatkin, and others on Google's Surveys team and
-  Google.org's CMU-Delphi Fellowship team; Curtiss Cobb and others on Facebook's   
-  Demography and Survey Science, Data for Good and Health teams; Alex Smola and
-  others at Amazon Web Services; Tim Suther and others at Change Healthcare;
-  Paul Nielsen and others at Optum; and John Tamerius and others at Quidel.
+  Hal Varian, Brett Slatkin, and others on Google's Surveys team;
+  Erin Hattersley, Rasmi Elasmar, and others at Google.org; 
+  Evgeniy Gabrilovich and others on Google's Health team; 
+  Curtiss Cobb and others on Facebook's  Demography and Survey Science, Data for
+  Good and Health teams; 
+  Alex Smola and others at Amazon Web Services;
+  Tim Suther and others at Change Healthcare;
+  Paul Nielsen and others at Optum; 
+  John Tamerius and others at Quidel; and
+  Ross Epstein and others at SafeGraph.
   This material is based on work supported by gifts from Facebook, Google.org,
@@
   Fellowship Program (NSF GRFP) award DGE1745016; and the Center for Machine
-  Learning and Health (CMLH) at Carnegie Mellon University.}
+  Learning and Health (CMLH) at Carnegie Mellon.}
 

### 2021-06-21 forecast/paper/paper.tex
Author declaration; supplement mods
Typo

@@
   (CTIS-CLI-in-community): The estimated percentage of the population who know
-  someone in their local community thatwho is sick, based on Delphi's COVID Trends
+  someone in their local community who is sick, based on Delphi's COVID Trends
   and Impact Survey, in partnership with Facebook.

### 2021-06-21 forecast/paper/paper.tex
Added author contributions.

--- 2021-06-21 forecast/paper/paper.tex
@@
 % declaration information 
-\authorcontributions{Please provide details of author contributions here.}
-\authordeclaration{Please declare any competing interests here.}
+\authorcontributions{D.J.M., J.B., A.G., A.J.H., and R.J.T wrote the manuscript,
+produced figures, and developed software necessary to perform the analyses. All
+authors performed research and provided intellectual contributions. R.J.T
+acquired funding.}
+\authordeclaration{The authors have no competing interests to declare.}
 \correspondingauthor{\textsuperscript{1}To whom correspondence should be

### 2021-06-21 forecast/paper/supplement.Rmd

@@
 ```
-
-<!-- PNAS says: Add this line AFTER all your figures and tables -->
-\FloatBarrier
-
-<!-- More PNAS example material -->
-<!-- Probably remove all this -->
-
-\dataset{dataset_one.txt}{Type or paste legend here.}
-
-\dataset{dataset_two.txt}{Type or paste legend here. Adding longer text to show what happens, to decide on alignment and/or indentations for multi-line or paragraph captions.}

### 2021-06-21 indicators/paper/paper.tex

@@
 % Middle authors
-\author[i]{Sumit Agrawal}
-\author[l]{Wael Al Saeed}
-\author[g]{Taylor Arnold}
-\author[j]{Amartya Basu}
-\author[d]{Jacob Bien}
-\author[r]{{\'A}ngel A. Cabrera}
+\author[d]{Sumit Agrawal} 
+\author[e]{Wael Al Saeed} 
+\author[f]{Taylor Arnold} 
+\author[g]{Amartya Basu}
+\author[h]{Jacob Bien} 
+\author[i]{{\'A}ngel A. Cabrera} 
 \author[b]{Andrew Chin}
@@
 \author[b]{Brian Clark}
-\author[i]{Sarah Colquhoun}
+\author[d]{Sarah Colquhoun}
 \author[b]{Nat DeFries}
-\author[i]{David C. Farrow}
-\author[r]{Jodi Forlizzi}
-\author[i]{Jed Grabman}
+\author[d]{David C. Farrow}
+\author[i]{Jodi Forlizzi}
+\author[d]{Jed Grabman}
 \author[b]{Samuel Gratzl}
 \author[a]{Alden Green}
-\author[r]{Robin Han}
-\author[i]{Kate Harwood}
+\author[i]{Robin Han}
+\author[d]{Kate Harwood}
 \author[a,b]{Addison J. Hu}
-\author[i]{Raphael Hyde}
-\author[d]{Sangwon Hyun}
-\author[l]{Ananya Joshi}
-\author[h]{Jimi Kim}
-\author[r]{Andrew Kuznetsov}
+\author[d]{Raphael Hyde}
+\author[h]{Sangwon Hyun}
+\author[e]{Ananya Joshi}
+\author[j]{Jimi Kim}
+\author[i]{Andrew Kuznetsov}
 \author[b]{Wichada La Motte-Kerr}
-\author[e,f]{Balasubramanian Narasimhan}
-\author[r]{Yeon Jin Lee}
-\author[p]{Kenneth Lee}
+\author[i]{Yeon Jin Lee}
+\author[k]{Kenneth Lee} 
 \author[b]{Zachary C. Lipton}
-\author[r]{Michael X. Liu}
-\author[k]{Lester Mackey}
+\author[i]{Michael X. Liu}
+\author[l]{Lester Mackey} 
 \author[b]{Kathryn Mazaitis}
-\author[o]{Daniel J. McDonald}
-\author[i]{Phillip McGuinness}
-\author[i]{Michael P. O'Brien}
+\author[m]{Daniel J. McDonald} 
+\author[d]{Phillip McGuinness}
+\author[n,o]{Balasubramanian Narasimhan} 
+\author[d]{Michael P. O'Brien}
 \author[a,b]{Natalia L. Oliveira}
 \author[a,b]{Pratik Patil}
-\author[r]{Adam Perer}
+\author[i]{Adam Perer}
 \author[b]{Collin A. Politsch}
-\author[l]{Dawn Rucker}
-\author[i]{Chris Scott}
-\author[s]{Nigam Shah}
-\author[n]{Vishnu Shankar}
-\author[p]{James Sharpnack}
+\author[o]{Samyak Rajanala}
+\author[e]{Dawn Rucker}
+\author[d]{Chris Scott}
+\author[p]{Nigam Shah} %s
+\author[q]{Vishnu Shankar} %n
+\author[k]{James Sharpnack}
 \author[b]{Dmitry Shemetov}
-\author[q]{Noah Simon}
-\author[i]{Benjamin Y. Smith}
+\author[r]{Noah Simon} %q
+\author[d]{Benjamin Y. Smith}
 \author[b]{Vishakha Srivastava}
-\author[e,f]{Robert Tibshirani}
-\author[e]{Elena Tuzhilina}
+\author[k]{Shiyu Tan}
+\author[n,o]{Robert Tibshirani}
+\author[n]{Elena Tuzhilina}
 \author[b]{Ana Karina Van Nortwick}
 \author[a]{Val{\'e}rie Ventura}
-\author[a]{Larry Wasserman}
-\author[i]{Benjamin Weaver}
-\author[m]{Jeremy C. Weiss}
-\author[i]{Spencer Whitman}
-\author[r]{Kristin Williams}
+\author[a,b]{Larry Wasserman}
+\author[d]{Benjamin Weaver}
+\author[s]{Jeremy C. Weiss} %m
+\author[d]{Spencer Whitman}
+\author[i]{Kristin Williams}
 
@@
 \affil[c]{Computational Biology Department, Carnegie Mellon University}
-\affil[d]{Department of Data Sciences and Operations, University of Southern California}
-\affil[e]{Department of Statistics, Stanford University}
-\affil[f]{Department of Biomedical Data Science, Stanford University}
-\affil[g]{Linguistics Program, University of Richmond}
-\affil[h]{School of Natural Sciences and Mathematics, University of Texas at Dallas}
-\affil[i]{Google.org}
-\affil[j]{Information Networking Institute, Carnegie Mellon University}
-\affil[k]{Microsoft Research New England}
-\affil[l]{Computer Science Department, Carnegie Mellon University}
-\affil[m]{Heinz College of Information Systems and Public Policy, Carnegie
+\affil[d]{Google.org}
+\affil[e]{Computer Science Department, Carnegie Mellon University}
+\affil[f]{Linguistics Program, University of Richmond}
+\affil[g]{Information Networking Institute, Carnegie Mellon University}
+\affil[h]{Department of Data Sciences and Operations, University of Southern California}
+\affil[i]{Human-Computer Interaction Institute, Carnegie Mellon University}
+\affil[j]{School of Natural Sciences and Mathematics, University of Texas at Dallas}
+\affil[k]{Department of Statistics, University of California, Davis}
+\affil[l]{Microsoft Research New England}
+\affil[m]{Department of Statistics, University of British Columbia}
+\affil[n]{Department of Statistics, Stanford University}
+\affil[o]{Department of Biomedical Data Science, Stanford University}
+\affil[p]{Department of Medicine, Stanford University}
+\affil[q]{Program in Immunology, Stanford University School of Medicine}
+\affil[r]{Department of Biostatistics, University of Washington}
+\affil[s]{Heinz College of Information Systems and Public Policy, Carnegie
   Mellon University}
-\affil[n]{Program in Immunology, Stanford University School of Medicine}
-\affil[o]{Department of Statistics, University of British Columbia}
-\affil[p]{Department of Statistics, University of California, Davis}
-\affil[q]{Department of Biostatistics, University of Washington}
-\affil[r]{Human-Computer Interaction Institute, Carnegie Mellon University}
-\affil[s]{Department of Medicine, Stanford University}
 

### 2021-06-21 forecast/paper/supplement-text.tex
Add extra leading/lagging text to supplement, rebuild.

@@
 
+In calculating the CCF and the associated leadingness and laggingness scores, we
+have used the finalized data, and we look at the behavior when $t$ is the
+target date of the forecast. That is we are using the same data to
+evaluate predictive accuracy as to determine
+leadingness and laggingness. It should be noted that the leadingness of the
+indicator at the time the model is trained may also be important. Thus, we could
+calculate separate leadningness and laggingness scores for the trained model and
+for the evaluation data and examine their combination in some way. We do not
+pursue this combination further and leave this question for future work.
+
 \section{Examining data in 2021}

### 2021-06-21 indicators/paper/supplement.Rmd
Rename HSP-HOSP -> HSP-Hosp in indicators supplement

@@
 Here, the HHS data has been limited to values from 2020-08-01 onward due to changes in reporting behavior, and processed into a 7-day trailing average rate per 100,000 resident population using Census Bureau estimates for 2019.
-HSP-HOSP is the percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners, smoothed in time and adjusted for systematic day-of-week effects.
+HSP-Hosp is the percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners, smoothed in time and adjusted for systematic day-of-week effects.
 Auxiliary signals have been scaled to have the same maximum value as the HHS data."}
@@
 hosp_names = c(
-  "Hospitalizations", "HSP-HOSP",
+  "Hospitalizations", "HSP-Hosp",
   "Cases", "CHNG-COVID"

### 2021-06-21 indicators/paper/supplement.Rmd
Calculate hosp cor's on common geotemp set in indicators supplement

@@
 ## Consider only state-like jurisdictions for which all signals of interest have at least 50 rows:
-geo_values = unique(df_hospitalizations$geo_value)
+hosp_geo_values = unique(df_hospitalizations$geo_value)
 for (i in seq_along(HOSP_SIGNALS)) {
-  geo_values = intersect(geo_values, df_hosp_signals[[i]] %>% count(geo_value) %>% filter(n > 50) %>% pull(geo_value))
+  hosp_geo_values = intersect(hosp_geo_values, df_hosp_signals[[i]] %>% count(geo_value) %>% filter(n > 50) %>% pull(geo_value))
 }
@@
 cat('Number of unique jurisdictions included:\n')
-length(geo_values)
+length(hosp_geo_values)
+```
+
+```{r hosp-correlation-instances}
+## Find common set of (geo_value, time_value) instances across all signals:
+hosp_common_instances =
+  select(df_hospitalizations, geo_value, time_value) %>%
+  filter(geo_value %in% hosp_geo_values)
+
+for (i in seq_along(HOSP_SIGNALS)) {
+  hosp_common_instances <- inner_join(hosp_common_instances,
+                                      select(df_hosp_signals[[i]], geo_value, time_value),
+                                      by=c("geo_value","time_value"))
+}
+
+cat('Number of common instances among all signals of interest:\n')
+nrow(hosp_common_instances)
 ```
@@
 Correlations sliced by time (geo-wise correlation):
-```{r hosp-correlations-by-time, fig.cap="Geo-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated for all times with sufficient available data within this period, over all state-like jurisdictions for which each signal was reported on at least 50 days during this period."}
+```{r hosp-correlations-by-time, fig.cap="Geo-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated for all times with sufficient available data within this period, over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which all signals of interest are available."}
 n_hosp_signals = length(HOSP_SIGNALS)
@@
   df_list_hosp_geowise_cor[[i]] = covidcast_cor(
-    df_hosp_signals[[i]] %>% filter(geo_value %in% geo_values),
-    df_hospitalizations %>% filter(geo_value %in% geo_values),
+    df_hosp_signals[[i]] %>% inner_join(hosp_common_instances, by=c("geo_value","time_value")),
+    df_hospitalizations %>% inner_join(hosp_common_instances, by=c("geo_value","time_value")),
     by = "time_value", method = "spearman"
@@
 Correlations sliced by state (time-series correlation)
-```{r hosp-correlations-by-state, fig.cap="Time-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated over all state-like jurisdictions for which each signal was reported on at least 50 days during this period."}
+```{r hosp-correlations-by-state, fig.cap="Time-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which all signals of interest are available."}
 df_list_hosp_timewise_cor = vector("list", length(HOSP_SIGNALS))
@@
   df_list_hosp_timewise_cor[[i]] = covidcast_cor(
-    df_hosp_signals[[i]] %>% filter(geo_value %in% geo_values),
-    df_hospitalizations %>% filter(geo_value %in% geo_values),
+    df_hosp_signals[[i]] %>% inner_join(hosp_common_instances, by=c("geo_value","time_value")),
+    df_hospitalizations %>% inner_join(hosp_common_instances, by=c("geo_value","time_value")),
     by = "geo_value", method = "spearman"

### 2021-06-21 indicators/paper/supplement.Rmd
Tweak caption description of HHS hospitalization data in indicators supplement

@@
 
-```{r hospitalization_time_trends_national, fig.cap="National trends, from April 2020 to April 2021, of HHS confirmed COVID-19 hospital admissions, plus three other signals in the COVIDcast API.
+```{r hospitalization_time_trends_national, fig.cap="National trends, from April 2020 to April 2021, of HHS-reported confirmed COVID-19 hospital admissions, plus three other signals in the COVIDcast API.
 Here, the HHS data has been limited to values from 2020-08-01 onward due to changes in reporting behavior, and processed into a 7-day trailing average rate per 100,000 resident population using Census Bureau estimates for 2019.
@@
 
-## Fetch HHS confirmed hospital admissions and smooth with 7-day
+## Fetch HHS-reported confirmed hospital admissions and smooth with 7-day
 ## trailing average rate:

### 2021-06-21 indicators/paper/supplement.Rmd
Limit hosp trend&cor figs in indicator supplement to hosp signals

@@
 
-```{r hospitalization_time_trends_national, fig.cap="National trends, from April 2020 to April 2021, of HHS-reported confirmed COVID-19 hospital admissions, plus three other signals in the COVIDcast API.
+```{r hospitalization_time_trends_national, fig.cap="National trends, from April 2020 to April 2021, of HHS-reported confirmed COVID-19 hospital admissions, plus another hospitalization signal in the COVIDcast API.
 Here, the HHS data has been limited to values from 2020-08-01 onward due to changes in reporting behavior, and processed into a 7-day trailing average rate per 100,000 resident population using Census Bureau estimates for 2019.
-HSP-Hosp is the percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners, smoothed in time and adjusted for systematic day-of-week effects.
-Auxiliary signals have been scaled to have the same maximum value as the HHS data."}
+HSP-Hosp is the percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners, smoothed in time and adjusted for systematic day-of-week effects (HSP-Hosp is available from an earlier date when working at finer geographical resolutions).
+HSP-Hosp has been scaled to have the same maximum value as the HHS data."}
 start_day = "2020-04-15"
@@
 
-HOSP_SOURCES <-
-  c("hhs", "hospital-admissions", "jhu-csse", "chng")
-HOSP_SIGNALS <-
-  c("confirmed_admissions_covid_1d",
-    "smoothed_adj_covid19_from_claims",
-    "confirmed_7dav_incidence_prop",
-    "smoothed_adj_outpatient_covid")
-hosp_names = c(
-  "Hospitalizations", "HSP-Hosp",
-  "Cases", "CHNG-COVID"
-)
-hosp_signal_has_national = c(
-  TRUE, TRUE,
-  TRUE, TRUE
-)
+HOSP_SOURCES <- c("hhs", "hospital-admissions")
+HOSP_SIGNALS <- c("confirmed_admissions_covid_1d", "smoothed_adj_covid19_from_claims")
+hosp_names = c("Hospitalizations", "HSP-Hosp")
+hosp_signal_has_national = c(TRUE, TRUE)
 hhs_start_date = as.Date("2020-08-01")
@@
 Correlations sliced by time (geo-wise correlation):
-```{r hosp-correlations-by-time, fig.cap="Geo-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated for all times with sufficient available data within this period, over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which all signals of interest are available."}
+```{r hosp-correlations-by-time, fig.cap="Geo-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated for all times with sufficient available data within this period, over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which both signals are available."}
 n_hosp_signals = length(HOSP_SIGNALS)
@@
 Correlations sliced by state (time-series correlation)
-```{r hosp-correlations-by-state, fig.cap="Time-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which all signals of interest are available."}
+```{r hosp-correlations-by-state, fig.cap="Time-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which both signals are available."}
 df_list_hosp_timewise_cor = vector("list", length(HOSP_SIGNALS))

### 2021-06-21 indicators/paper/supplement.Rmd
Tweak non-caption text between hosp-related figs in indicators supplement

@@
 
+\clearpage
+
 ```{r hosp-correlation-data}
@@
 
-cat('Number of unique jurisdictions included:\n')
+cat('Number of unique jurisdictions included in correlation analysis below:\n')
 length(hosp_geo_values)
@@
 
-cat('Number of common instances among all signals of interest:\n')
+cat('Number of common instances among all signals of interest in correlation analysis below:\n')
 nrow(hosp_common_instances)
@@
 
-Correlations sliced by time (geo-wise correlation):
 ```{r hosp-correlations-by-time, fig.cap="Geo-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated for all times with sufficient available data within this period, over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which both signals are available."}
@@
 
-Correlations sliced by state (time-series correlation)
 ```{r hosp-correlations-by-state, fig.cap="Time-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which both signals are available."}

### 2021-06-21 forecast/paper/pnas-suppl-template.tex
Author declaration; supplement mods
Add data placeholders to supplement.

@@
 
-% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
 \SItext
@@
 
-\bibliography{$for(bibliography)$$bibliography$$sep$,$endfor$}
+\FloatBarrier
+
+\dataset{predictions.zip}{Archived {\tt .RDS} (R objects) files containing
+  all predictions for forecasting and hotspots using vintage and finalized data.
+Persistent DOI to be added at publication.}
+
+\dataset{evaluations.zip}{Archived {\tt .RDS} (R objects) files containing
+  all evaluations for forecasting and hotspots using vintage and finalized data.
+  Persistent DOI to be added at publication.}
+
+\dataset{analysis.zip}{Archived {\tt .RDS} (R objects) files containing
+  additional data used to produce graphics and conclusions in the manuscript.
+  Persistent DOI to be added at publication.}
+
+\dataset{code.zip}{R script files containing
+  all code used to reproduce the analysis described in the manuscript.
+  Persistent DOI to be added at publication.}
+
+%\bibliography{$for(bibliography)$$bibliography$$sep$,$endfor$}
 

### 2021-06-21 forecast/paper/supplement.tex
Author declaration; supplement mods
Add extra leading/lagging text to supplement, rebuild.
Add data placeholders to supplement.

@@
 
-\dataset{dataset_one.txt}{Type or paste legend here.}
+\dataset{predictions.zip}{Archived {\tt .RDS} (R objects) files containing
+  all predictions for forecasting and hotspots using vintage and finalized data.
+Persistent DOI to be added at publication.}
 
-\dataset{dataset_two.txt}{Type or paste legend here. Adding longer text to show what happens, to decide on alignment and/or indentations for multi-line or paragraph captions.}
+\dataset{evaluations.zip}{Archived {\tt .RDS} (R objects) files containing
+  all evaluations for forecasting and hotspots using vintage and finalized data.
+  Persistent DOI to be added at publication.}
 
-\bibliography{../../common/covidcast.bib,pnas-materials/pnas-sample.bib}
+\dataset{analysis.zip}{Archived {\tt .RDS} (R objects) files containing
+  additional data used to produce graphics and conclusions in the manuscript.
+  Persistent DOI to be added at publication.}
+
+\dataset{code.zip}{R script files containing
+  all code used to reproduce the analysis described in the manuscript.
+  Persistent DOI to be added at publication.}
+
+%\bibliography{../../common/covidcast.bib,pnas-materials/pnas-sample.bib}
 

### 2021-06-21 forecast/paper/pnas-suppl-template.tex
Minor typos.

@@
 
-\author{Daniel J.\ McDonald, Jacob Bien, Alden Green, Addison Hu, Nat DeFries, Sangwon Hyun, Natalia L.\ Oliveira, James Sharpnack, Jingjing Tang, Robert Tibshirani, Valerie Ventura, Larry Wasserman, and Ryan J.\ Tibshirani}
+\author{Daniel J. McDonald, Jacob Bien, Alden Green, Addison Hu, Nat DeFries,
+  Sangwon Hyun, Natalia L. Oliveira, James Sharpnack, Jingjing Tang, Robert
+  Tibshirani, Valerie Ventura, Larry Wasserman, and Ryan J. Tibshirani} 
 \correspondingauthor{Daniel J. McDonald.\\E-mail: daniel@stat.ubc.ca}

### 2021-06-21 forecast/paper/supplement-text.tex
Minor typos.

@@
   multiplicative space-time effects $S_{\ell,t}$ shared across all forecasters,
-  i.e. $\WIS(F_{\ell,t,f},Y_{\ell,t})=S_{\ell,t}E_{f,t}$ with $E_{F,t}$ a
+  i.e. $\WIS(F_{\ell,t,f},Y_{\ell,t})=S_{\ell,t}E_{f,t}$ with $E_{f,t}$ a
   forecaster-specific error, then

### 2021-06-21 forecast/paper/supplement-text.tex
minor changes

--- 2021-06-21 forecast/paper/supplement-text.tex
@@
 The goal of this section is to quantify the effect of
-not properly accounting for the question of "what was known when" in
+not properly accounting for the question of ``what was known when'' in
 performing retrospective evaluations of forecasters.   Figures
@@
 of leadingness or laggingness, we use the cross correlation function (CCF)
-between the two time series. The CCF is of an indicator $X_{\ell,t}$ and case
+between the two time series. The CCF of an indicator $X_{\ell,t}$ and case
 rates $Y_{\ell,t}$ is defined as
@@
 difference in indicator WIS and AR WIS. Unlike leadingness (Figure 5 in the
-manuscript) there is no obvious relationship. This is heartening as laggingness
+manuscript) there is no obvious relationship that holds consistently
+across indicators. This is heartening as laggingness
 should not aid forecasting performance. On the other hand, if an indicator is

### 2021-06-21 indicators/paper/paper.tex
Authorship fixes
Remove old aff letters
Author contributions

@@
 \author[a]{Alden Green}
+\author[b]{George Haff}
 \author[i]{Robin Han}
@@
 \author[b]{Collin A. Politsch}
-\author[o]{Samyak Rajanala}
+\author[n]{Samyak Rajanala}
 \author[e]{Dawn Rucker}
 \author[d]{Chris Scott}
-\author[p]{Nigam Shah} %s
-\author[q]{Vishnu Shankar} %n
+\author[p]{Nigam H. Shah} 
+\author[q]{Vishnu Shankar} 
 \author[k]{James Sharpnack}
 \author[b]{Dmitry Shemetov}
-\author[r]{Noah Simon} %q
+\author[r]{Noah Simon} 
 \author[d]{Benjamin Y. Smith}
 \author[b]{Vishakha Srivastava}
-\author[k]{Shiyu Tan}
+\author[m]{Shuyi Tan}
 \author[n,o]{Robert Tibshirani}
@@
 \author[d]{Benjamin Weaver}
-\author[s]{Jeremy C. Weiss} %m
+\author[s]{Jeremy C. Weiss} 
 \author[d]{Spencer Whitman}
@@
 % Please include corresponding author, author contribution and author declaration information
-\authorcontributions{Please provide details of author contributions here.}
-\authordeclaration{Please declare any competing interests here.}
+\authorcontributions{Author contributions: A.R., L.B., M.J., A.R., J.T., A.C,
+  E.J.C., N.D., D.C.F, A.J.H., S.H., M.J., J.K., W.L.M.K, Z.C.L., L.M., N.L.O.,
+  J.C.W, R.R., and R.J.T., obtained and managed data;
+  A.R., J.B., J.F., Z.C.L., L.M., D.J.M., A.P., S.R., D.R., C.S. N.S., R.T.,
+  V.V., L.W., B.W., J.C.W, S.W., R.R., and R.J.T. designed software and
+  research,
+  A.R., L.B., M.J., A.R., J.T., W.A.S., T.A., A.B., J.B., L.B., A.A.C., A.C.,
+  E.J.C., B.C., S.C., N.D., D.C.F., J.G., S.G., A.G., G.H., R.H., K.H., A.J.H.,
+  S.H., M.J., A.J., J.K., A.K., M.X.L., L.M., K.M., D.J.M., B.N., M.P.O.,
+  N.L.O., P.P., C.A.P., S.R., C.S., V.S, J.S, D.S., N.S., B.Y.S., V.S., S.T.,
+  J.T., R.T., J.C.W, and R.J.T. developed software;
+  S.A., W.A.S., A.C., B.C., D.C.F., S.G., G.H., K.M., D.J.M., B.N., C.S., J.S.,
+  D.S., B.Y.S., A.K.V.N. managed technical infrastructure and systems;
+  A.R., L.B., M.J., A.R., J.T, J.B., A.C., E.J.C., N.D., D.C.F., A.G., K.H.,
+  A.J.H., R.H., S.H., A.J., J.K., K.L., Y.J.L, Z.C.L., L.M., D.J.M., P.M., B.N.,
+  N.L.O., P.P., C.A.P., S.R., N.S., V.S., J.S., D.S., N.S., S.T., J.T., R.T.,
+  V.V., L.W., J.C.W., R.R., and R.J.T. performed research;
+  and A.R., L.B., M.J., A.R., J.T., and R.J.T. wrote the paper.
+}
+\authordeclaration{No competing interests to declare.}
 \correspondingauthor{\textsuperscript{1}To whom correspondence should be

### 2021-06-21 indicators/paper/paper.tex
Remove TODO

@@
 
-[TODO: improve/revisit this example?]
 As an illustration, Figure~\ref{fig:mobility-drop} maps the sharp increase in

### 2021-06-22 forecast/paper/supplement-text.tex
Go to 95% and fix up the leading/lagging

--- 2021-06-21 forecast/paper/supplement-text.tex
@@
 of leadingness or laggingness, we use the cross correlation function (CCF)
-between the two time series. The CCF is of an indicator $X_{\ell,t}$ and case
-rates $Y_{\ell,t}$ is defined as
-\[
-\CCF_{\ell,t}(a) = \frac{1}{n_a} \sum_{i=1}^{n_a} \left( ( X_{\ell,t,i+a} -
-  \overline{X}_{\ell,t}) / s^{X}_{\ell,t}\right) \left( ( Y_{\ell,t,i} -
-  \overline{Y}_{\ell,t}) / s^{Y}_{\ell,t}\right), 
-\]
-where $n_a$ is the number of available time points when $X_{\ell,t}$ has been
-shifted in time by $a$ days, and $\overline{X}_{\ell,t}$, $s^X_{\ell,t}$
-are the sample mean and
-standard deviation of $X_{\ell,t}$ (respectively $Y_{\ell,t}$). The result is a
-sequence of Pearson correlations for each selected $a$, where we align
-$X_{\ell,t}$ with the values of $Y_{\ell,t}$ that occurred $a$ days earlier.
-Thus, for $a>0$, $\CCF_{\ell,t}(a)>0$ indicates that
-$Y_{\ell,t}$ is moving together with the values of $X_{\ell,t}$ that occur $a$
-days in the future. In this case we say that $X_{\ell,t}$ is lagging
-$Y_{\ell,t}$. For $a<0$, $\CCF_{\ell,t}(a)>0$ means that $Y_{\ell,t}$ is
-positively correlated with $X_{\ell,t}$ shifted earlier, so we say that
-$X_{\ell,t}$ leads $Y_{\ell,t}$.
+between the two time series. The $\CCF_\ell(a)$ of an indicator $X_{\ell}$ and case
+rates $Y_{\ell}$ is defined as
+% \[
+% \CCF_{\ell,t}(a) = \frac{1}{n_a} \sum_{i=1}^{n_a} \left( ( X_{\ell,t,i+a} -
+%   \overline{X}_{\ell,t}) / s^{X}_{\ell,t}\right) \left( ( Y_{\ell,t,i} -
+%   \overline{Y}_{\ell,t}) / s^{Y}_{\ell,t}\right), 
+% \]
+% where $n_a$ is the number of available time points when $X_{\ell,t}$ has been
+% shifted in time by $a$ days, and $\overline{X}_{\ell,t}$, $s^X_{\ell,t}$
+% are the sample mean and
+% standard deviation of $X_{\ell,t}$ (respectively $Y_{\ell,t}$). The result is a
+their Pearson correlation where $X_\ell$ has been
+% sequence of Pearson correlations for each selected $a$, where we
+aligned
+with the values of $Y_{\ell}$ that occurred $a$ days earlier.
+Thus, for any $a>0$, $\CCF_{\ell}(a)>0$ indicates that
+$Y_{\ell,t}$ is moving together with the values of $X_{\ell,t+a}$.
+% days in the future.
+In this case we say that $X_{\ell}$ is lagging
+$Y_{\ell}$. For $a<0$, $\CCF_{\ell}(a)>0$ means that $Y_{\ell,t}$ is
+positively correlated with $X_{\ell,t-a}$, so we say that
+$X_{\ell}$ leads $Y_{\ell}$.
 
@@
 containing Charlotte, North Carolina, from August 1, 2020 until the end of
-September. These are the same signals shown in Figure 1 in the manuscript. To
-define ``leadingness'' we compute $\CCF_{\ell,t}(a)$ for each $a\in
-\{-15,\ldots,15\}$ using the 56 days leading up to time $t$ (this is the same
+September. These are the same signals shown in Figure 1 in the manuscript but
+using finalized data. To
+define ``leadingness'' we compute $\CCF_{\ell}(a)$ for each $a\in
+\{-15,\ldots,15\}$ using the 56 days leading up to the target date (this is the same
 amount of data used to train the forecasters: 21 days of training data, 21 days
 to get the response at $a=21$, and 14 days for the longest lagged value). The
-orange dashed horizontal line represents the 97.5\% significance threshold for
+orange dashed horizontal line represents the 95\% significance threshold for
 correlations based on 56 observations. Any correlations larger in magnitude than

### 2021-06-22 forecast/paper/supplement.Rmd
Go to 95% and fix up the leading/lagging

@@
   out <- ccf(x, y, lag.max = lag.max, plot = FALSE, na.action = na.omit)
-  clim <- qnorm((1 + .95) / 2) / sqrt(out$n.used)
+  clim <- qnorm((1 + .9) / 2) / sqrt(out$n.used)
   lag_vec <- drop(out$lag)

### 2021-06-22 forecast/code/ccf-dv-finalized.R

@@ -3,7 +3,7 @@ sigs <- covidcast_signals(
   signal = c("confirmed_7dav_incidence_prop", "smoothed_adj_cli"),
   start_day = "2020-08-01",
   end_day = "2020-09-28",
-  as_of = "2020-09-28", #2020-12-21",
+  as_of = "2021-05-15", #2020-12-21",
   geo_type = "hrr",
   geo_values = "311") %>% # NYC
   aggregate_signals(format = "wide")
@@ -31,7 +31,7 @@ g1 <- std_sigs %>%
 cc <- ccf(sigs$`doctor-visits`, sigs$cases, plot = FALSE)
 g2 <- ggplot(tibble(lag = drop(cc$lag), ccf = drop(cc$acf)), aes(lag)) +
   geom_segment(aes(xend = lag, yend = ccf), y = 0, color = "cornflowerblue") +
-  geom_hline(yintercept = qnorm((1 + .95) / 2) / sqrt(cc$n.used),
+  geom_hline(yintercept = qnorm((1 + .9) / 2) / sqrt(cc$n.used),
              color = "darkorange", linetype = "dashed") +
   geom_hline(yintercept = 0) +
   geom_vline(xintercept = 0) +

### 2021-06-22 forecast/paper/supplement.Rmd
Final culling of leading/lagging.

@@
 ```{r ccf-dv-finalized, eval=TRUE, fig.cap="Illustration of the cross-correlation function between DV-CLI and cases.
-The left panel shows the standardized signals over the period from August 1 to September 28 (as of September 28, 2020).
-The right panel shows $\\CCF(a)$ for different values of $a$ as vertical blue bars.
+The left panel shows the standardized signals over the period from August 1 to September 28 (as of May 15, 2021).
+The right panel shows $\\CCF_{\\ell}(a)$ for different values of $a$ as vertical blue bars.
 The orange dashed lines indicate the 95\\% significance threshold.
-By our leadingness metric, DV-CLI is neither leading nor lagging cases over this period."}
+By our leadingness/laggingness metric, DV-CLI is leading (but notlagging) cases over this period."}
 source("../code/ccf-dv-finalized.R")

### 2021-06-22 forecast/paper/supplement-text.tex

@@
 are leading or lagging case rates during different periods. To define the amount
-of leadingness or laggingness, we use the cross correlation function (CCF)
+of leadingness or laggingness at location $\ell$, we use the cross correlation
+function (CCF) 
 between the two time series. The $\CCF_\ell(a)$ of an indicator $X_{\ell}$ and case
@@
 Thus, for any $a>0$, $\CCF_{\ell}(a)>0$ indicates that
-$Y_{\ell,t}$ is moving together with the values of $X_{\ell,t+a}$.
+$Y_{\ell,t}$ is moving together with $X_{\ell,t+a}$.
 % days in the future.
@@
 using finalized data. To
-define ``leadingness'' we compute $\CCF_{\ell}(a)$ for each $a\in
-\{-15,\ldots,15\}$ using the 56 days leading up to the target date (this is the same
+define ``leadingness'' we compute $\CCF_{\ell}(a)$ (as implemented with the R
+function {\tt ccf()}) for each $a\in
+\{-15,\ldots,15\}$ using the 56 days leading up to the target date. This is the same
 amount of data used to train the forecasters: 21 days of training data, 21 days
-to get the response at $a=21$, and 14 days for the longest lagged value). The
+to get the response at $a=21$, and 14 days for the longest lagged value. The
 orange dashed horizontal line represents the 95\% significance threshold for
@@
 correlations that are leading (those above the dashed line with $a<0$) while
-laggingness is the same but for $a>0$. In this case, there are no significant
-correlations, so both scores will be 0: on September 28, \dv~is neither
-leading nor lagging cases.
+laggingness is the same but for $a>0$. In the figure, there are three
+significant correlations on the ``leading'' side (at $a = -5, -4, -3$), so
+leadingness will be the sum of those values while laggingness is 0:
+%correlations, so both scores will be 0:
+on September 28 in Charlotte, \dv~is leading cases
+leading but not lagging.
 
@@
 In calculating the CCF and the associated leadingness and laggingness scores, we
-have used the finalized data, and we look at the behavior when $t$ is the
+have used the finalized data, and we look at the behavior at the 
 target date of the forecast. That is we are using the same data to
@@
 indicator at the time the model is trained may also be important. Thus, we could
-calculate separate leadningness and laggingness scores for the trained model and
+calculate separate leadingness and laggingness scores for the trained model and
 for the evaluation data and examine their combination in some way. We do not
-pursue this combination further and leave this question for future work.
+pursue this combination further and leave this investigation for future work.
 

### 2021-06-22 forecast/paper/supplement.tex

@@
 \caption{Illustration of the cross-correlation function between DV-CLI and cases.
-The left panel shows the standardized signals over the period from August 1 to September 28 (as of September 28, 2020).
-The right panel shows $\CCF(a)$ for different values of $a$ as vertical blue bars.
+The left panel shows the standardized signals over the period from August 1 to September 28 (as of May 15, 2021).
+The right panel shows $\CCF_{\ell}(a)$ for different values of $a$ as vertical blue bars.
 The orange dashed lines indicate the 95\% significance threshold.
-By our leadingness metric, DV-CLI is neither leading nor lagging cases over this period.}\label{fig:ccf-dv-finalized}
+By our leadingness/laggingness metric, DV-CLI is leading (but notlagging) cases over this period.}\label{fig:ccf-dv-finalized}
 \end{figure}

### 2021-06-22 indicators/paper/pnas-suppl-template.tex
Fill out supplement template

--- 2021-06-16 forecast/paper/pnas-suppl-template.tex
@@
 
-\title{Your main manuscript title}
-\author{Author1, Author2 and Author3 (complete author list)}
-\correspondingauthor{Corresponding Author name.\\E-mail: author.two@email.com}
+\title{An Open Repository of Real-Time COVID-19 Indicators}
+\author{Alex Reinhart, Logan Brooks, Maria Jahja, Aaron Rumack, Jingjing Tang,
+  et al.}
+\correspondingauthor{Alex Reinhart.\\E-mail: areinhar@stat.cmu.edu}
 

### 2021-06-22 indicators/paper/paper.tex
Formatting fixes
Slightly expand viz section

@@
   E.J.C., N.D., D.C.F, A.J.H., S.H., M.J., J.K., W.L.M.K, Z.C.L., L.M., N.L.O.,
-  J.C.W, R.R., and R.J.T., obtained and managed data;
+  J.C.W, R.R., and R.J.T. obtained and managed data;
   A.R., J.B., J.F., Z.C.L., L.M., D.J.M., A.P., S.R., D.R., C.S. N.S., R.T.,
   V.V., L.W., B.W., J.C.W, S.W., R.R., and R.J.T. designed software and
-  research,
+  research;
   A.R., L.B., M.J., A.R., J.T., W.A.S., T.A., A.B., J.B., L.B., A.A.C., A.C.,
@@
   S.A., W.A.S., A.C., B.C., D.C.F., S.G., G.H., K.M., D.J.M., B.N., C.S., J.S.,
-  D.S., B.Y.S., A.K.V.N. managed technical infrastructure and systems;
+  D.S., B.Y.S., and A.K.V.N. managed technical infrastructure and systems;
   A.R., L.B., M.J., A.R., J.T, J.B., A.C., E.J.C., N.D., D.C.F., A.G., K.H.,
@@
   V.V., L.W., J.C.W., R.R., and R.J.T. performed research;
+  R.R. and R.J.T. acquired funding;
   and A.R., L.B., M.J., A.R., J.T., and R.J.T. wrote the paper.
@@
 and backfill. There is also a dedicated dashboard for exploring results from the
-COVID Trends and Impact Survey.
+COVID Trends and Impact Survey. The visualizations have been continually
+improved as new sources of data arrive, and in response to interviews with users
+and health experts, usage analytics from the site, and user surveys. Key design
+goals for visualizations displaying COVIDcast data have included representing
+uncertainty and missing data, making features consistent and discoverable
+despite a wide variety of signal types, and smoothly switching between different
+geographic aggregations and time scales.
 

### 2021-06-22 indicators/paper/paper.tex
Fix Yeon Jin Lee's affiliation

@@
 \author[b]{Wichada La Motte-Kerr}
-\author[i]{Yeon Jin Lee}
-\author[k]{Kenneth Lee}
+\author[k]{Yeon Jin Lee}
+\author[l]{Kenneth Lee}
 \author[b]{Zachary C. Lipton}
 \author[i]{Michael X. Liu}
-\author[l]{Lester Mackey}
+\author[m]{Lester Mackey}
 \author[b]{Kathryn Mazaitis}
-\author[m]{Daniel J. McDonald}
+\author[n]{Daniel J. McDonald}
 \author[d]{Phillip McGuinness}
-\author[n,o]{Balasubramanian Narasimhan}
+\author[o,p]{Balasubramanian Narasimhan}
 \author[d]{Michael P. O'Brien}
@@
 \author[b]{Collin A. Politsch}
-\author[n]{Samyak Rajanala}
+\author[o]{Samyak Rajanala}
 \author[e]{Dawn Rucker}
 \author[d]{Chris Scott}
-\author[p]{Nigam H. Shah} 
-\author[q]{Vishnu Shankar} 
-\author[k]{James Sharpnack}
+\author[q]{Nigam H. Shah}
+\author[r]{Vishnu Shankar}
+\author[l]{James Sharpnack}
 \author[b]{Dmitry Shemetov}
-\author[r]{Noah Simon} 
+\author[s]{Noah Simon}
 \author[d]{Benjamin Y. Smith}
 \author[b]{Vishakha Srivastava}
-\author[m]{Shuyi Tan}
-\author[n,o]{Robert Tibshirani}
-\author[n]{Elena Tuzhilina}
+\author[n]{Shuyi Tan}
+\author[o,p]{Robert Tibshirani}
+\author[o]{Elena Tuzhilina}
 \author[b]{Ana Karina Van Nortwick}
@@
 \author[d]{Benjamin Weaver}
-\author[s]{Jeremy C. Weiss} 
+\author[t]{Jeremy C. Weiss}
 \author[d]{Spencer Whitman}
@@
 \affil[j]{School of Natural Sciences and Mathematics, University of Texas at Dallas}
-\affil[k]{Department of Statistics, University of California, Davis}
-\affil[l]{Microsoft Research New England}
-\affil[m]{Department of Statistics, University of British Columbia}
-\affil[n]{Department of Statistics, Stanford University}
-\affil[o]{Department of Biomedical Data Science, Stanford University}
-\affil[p]{Department of Medicine, Stanford University}
-\affil[q]{Program in Immunology, Stanford University School of Medicine}
-\affil[r]{Department of Biostatistics, University of Washington}
-\affil[s]{Heinz College of Information Systems and Public Policy, Carnegie
+\affil[k]{College of Fine Arts and Department of Psychology, Carnegie Mellon University}
+
+\affil[l]{Department of Statistics, University of California, Davis}
+\affil[m]{Microsoft Research New England}
+\affil[n]{Department of Statistics, University of British Columbia}
+\affil[o]{Department of Statistics, Stanford University}
+\affil[p]{Department of Biomedical Data Science, Stanford University}
+\affil[q]{Department of Medicine, Stanford University}
+\affil[r]{Program in Immunology, Stanford University School of Medicine}
+\affil[s]{Department of Biostatistics, University of Washington}
+\affil[t]{Heinz College of Information Systems and Public Policy, Carnegie
   Mellon University}

### 2021-06-22 forecast/paper/paper-medrxiv.tex
Add medRXiv submission doc; fix Addison's middle initial

--- 2021-06-21 forecast/paper/paper.tex
@@
-\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
+\documentclass{article}
 % Use the lineno option to display guide line numbers if required.
 
-\templatetype{pnasresearcharticle} % Choose template 
+% \templatetype{pnasresearcharticle} % Choose template 
 % {pnasresearcharticle} = Template for a two-column research article
@@
 % {pnasinvited} %= Template for a PNAS invited submission
+
+\usepackage{authblk}
+\usepackage{amsmath,amsfonts,amssymb}
+\usepackage[numbers,square,sort&compress]{natbib}
+\usepackage[colorlinks=true, allcolors=blue]{hyperref}
+\usepackage{graphicx}
+\usepackage{booktabs}
+\usepackage[margin=1in]{geometry}
 
@@
 
-% Please give the surname of the lead author for the running footer
-\leadauthor{McDonald}
-
 % Please add a significance statement to explain the relevance of your work
-\significancestatement{Forecasting is a critical element in the public health
-  response to any fast-moving epidemic or pandemic.  Predicting the future
-  spread of some disease process, based on disease activity in the recent
-  past---say, using what is known as autoregressive modeling in the time series
-  literature---is a basic and widely-used method.  In this paper we study
-  whether including auxiliary indicators as additional features in such an
-  autoregressive model can be useful in COVID-19 case forecasting and hotspot
-  prediction.  We find that indicators based on de-identified medical insurance 
-  claims, self-reported symptoms through online surveys, and COVID-related
-  internet search activity all provide a nontrivial boost in accuracy, though
-  their utility generally depends the pandemic's local dynamics (upswing,
-  downswing, or flat period) at prediction time.
-}
-
-% Please include corresponding author, author contribution and author
-% declaration information 
-\authorcontributions{Please provide details of author contributions here.}
-\authordeclaration{Please declare any competing interests here.}
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be
-  addressed. E-mail: daniel@stat.ubc.ca} 
-\equalauthors{\textsuperscript{2}J.B., A.G., and A.J.H. contributed equally to 
-  this work.}
+% \significancestatement{Forecasting is a critical element in the public health
+%   response to any fast-moving epidemic or pandemic.  Predicting the future
+%   spread of some disease process, based on disease activity in the recent
+%   past---say, using what is known as autoregressive modeling in the time series
+%   literature---is a basic and widely-used method.  In this paper we study
+%   whether including auxiliary indicators as additional features in such an
+%   autoregressive model can be useful in COVID-19 case forecasting and hotspot
+%   prediction.  We find that indicators based on de-identified medical insurance 
+%   claims, self-reported symptoms through online surveys, and COVID-related
+%   internet search activity all provide a nontrivial boost in accuracy, though
+%   their utility generally depends the pandemic's local dynamics (upswing,
+%   downswing, or flat period) at prediction time.
+% }
 
@@
 % five keywords, separated by the pipe symbol.
-\keywords{COVID-19 $|$ forecasting $|$ hotspot prediction $|$ time series | 
-  digital surveillance}
+
+\begin{document}
+
+\maketitle
 
@@
   indicators'' of case rates.
+
+  {\bf Keywords:} COVID-19 $|$ forecasting $|$ hotspot prediction $|$ time series
+    $|$ digital surveillance
 \end{abstract}
 
-\dates{This manuscript was compiled on \today}
-\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
+\footnotetext[1]{To whom correspondence should be
+  addressed. E-mail: \href{mailto:daniel@stat.ubc.ca}{daniel@stat.ubc.ca}} 
+\footnotetext[2]{J.B., A.G., and A.J.H. contributed equally to 
+  this work.}
 
-\begin{document}
+\clearpage
 
-\maketitle
-\thispagestyle{firststyle}
-\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
-
-\dropcap{T}racking and forecasting indicators from public health reporting
+Tracking and forecasting indicators from public health reporting
 streams---such as confirmed cases and deaths in the COVID-19 pandemic---is
@@
 \begin{figure}[tb!]
-  \includegraphics[width=\columnwidth]{fig/revisions.pdf}
+  \centering
+  \includegraphics[width=.5\linewidth]{fig/revisions.pdf}
   % RJT: Ensuring we start with a "declarative title" :) 
@@
   this paper.}
-\begin{tabular}{l p{2.75in} p{2.75in}}
+\begin{tabular}{l p{2.25in} p{2.25in}}
   \toprule
@@
 probabilistic forecasting models of the form:  
-\begin{multline}
+\begin{equation}
 \label{eq:forecast_ar_x}
 \Quantile_\tau(Y_{\ell,t+a} \ | \ Y_{\ell,s},\ X_{\ell,s},\ s \leq t) 
-= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j} + 
+= \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j} + 
 \sum_{j=0}^2 \gamma^{a,\tau}_j X_{\ell,t-7j},
-\end{multline}
+\end{equation}
 where $X_{\ell,t}$ denotes any one of the five auxiliary indicators---CHNG-CLI, 
@@
 \begin{figure}[t]
-\includegraphics[width=\columnwidth]{fig/trajectory.pdf}
+  \centering
+\includegraphics[width=.5\columnwidth]{fig/trajectory.pdf}
 \caption{Forecast for the HRR containing New York City from an autoregressive
@@
 logistic models of the form:  
-\begin{multline}
+\begin{equation}
 \label{eq:hotspot_ar_x}
 \logit \big( \P(Z_{\ell,t+a} = 1 \ | \ Y_{\ell,s},\ X_{\ell,s},\ s \leq t) \big)
-= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j} +  
+=  \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j} +  
 \sum_{j=0}^2 \gamma^{a,\tau}_j X^\Delta_{\ell,t-7j},
-\end{multline}
+\end{equation}
 where we use \smash{$X^\Delta_{\ell,t} = (X_{\ell,t} - X_{\ell,
@@
 \begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/leading-only-1.pdf}
+  \centering
+  \includegraphics[width=.5\columnwidth]{fig/leading-only-1.pdf}
   \caption{Correlation of the difference in WIS with the ``leadingness'' of the
@@
 
-\acknow{We thank Matthew Biggerstaff, Logan Brooks, Johannes Bracher, Michael 
+\subsubsection*{Acknowledgements}
+We thank Matthew Biggerstaff, Logan Brooks, Johannes Bracher, Michael  
 Johansson, Evan Ray, Nicholas Reich, and Roni Rosenfeld for several
@@
 RGPIN-2021-02618; and National Science Foundation Graduate Research
-  Fellowship Program (NSF GRFP) award DGE1745016.}
+Fellowship Program (NSF GRFP) award DGE1745016.
 
-\showacknow{} % Display the acknowledgments section
+%\showacknow{} % Display the acknowledgments section
 
 % Bibliography
+\bibliographystyle{unsrtnat}
 \bibliography{../../common/covidcast.bib}
 
+\clearpage
+
+\appendix
+
+\section*{Supplemental information}
+
+\input{supplement-medrxiv.tex}
+
 \end{document}

### 2021-06-22 forecast/paper/supplement-medrxiv.tex
Add medRXiv submission doc; fix Addison's middle initial

--- 2021-06-22 forecast/paper/supplement.tex
@@
-\documentclass[9pt,twoside,lineno]{pnas-new}
-% Use the lineno option to display guide line numbers if required.
-
-\templatetype{pnassupportinginfo}
-
-\input{../../common/prelim.tex}
-
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-\title{Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot  
-  Prediction?}
-
-\author{Daniel J. McDonald, Jacob Bien, Alden Green, Addison Hu, Nat DeFries,
-  Sangwon Hyun, Natalia L. Oliveira, James Sharpnack, Jingjing Tang, Robert
-  Tibshirani, Valerie Ventura, Larry Wasserman, and Ryan J. Tibshirani} 
-\correspondingauthor{Daniel J. McDonald.\\E-mail: daniel@stat.ubc.ca}
-
-\begin{document}
-
-\maketitle
-
-\SItext
-
 \input{supplement-text.tex}
@@
 
-\clearpage
-
-\FloatBarrier
-
-\dataset{predictions.zip}{Archived {\tt .RDS} (R objects) files containing
-  all predictions for forecasting and hotspots using vintage and finalized data.
-Persistent DOI to be added at publication.}
-
-\dataset{evaluations.zip}{Archived {\tt .RDS} (R objects) files containing
-  all evaluations for forecasting and hotspots using vintage and finalized data.
-  Persistent DOI to be added at publication.}
-
-\dataset{analysis.zip}{Archived {\tt .RDS} (R objects) files containing
-  additional data used to produce graphics and conclusions in the manuscript.
-  Persistent DOI to be added at publication.}
-
-\dataset{code.zip}{R script files containing
-  all code used to reproduce the analysis described in the manuscript.
-  Persistent DOI to be added at publication.}
-
 %\bibliography{../../common/covidcast.bib,pnas-materials/pnas-sample.bib}
-
-\end{document}

### 2021-06-22 indicators/paper/supplement.Rmd
Remove supplement placeholders

@@
 <!-- Comment out or remove this line before generating final copy for submission; this will also remove the warning re: "Consecutive odd pages found". -->
-\instructionspage
 
@@
 <!-- Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text. -->
-
-\SItext
-
-All our text goes here. Reference figures with the R chuck label as Figure~\ref{fig:fcast-finalized}.
-
-All figures go on their own page after all of the text...
-
-Edit `pnas-suppl-template.tex` for the correct Author list and Title.
 
@@
 }
-
-cat('Number of unique jurisdictions included in correlation analysis below:\n')
-length(hosp_geo_values)
 ```
@@
 }
-
-cat('Number of common instances among all signals of interest in correlation analysis below:\n')
-nrow(hosp_common_instances)
 ```

### 2021-06-23 forecast/paper/paper.tex
Fix small typo
Add periods to RJT in contributions. Oops.
Remove personal communication with Google reference

@@
 % declaration information
-\authorcontributions{D.J.M., J.B., A.G., A.J.H., and R.J.T wrote the manuscript,
+
+\authorcontributions{D.J.M., J.B., A.G., A.J.H., and R.J.T. wrote the manuscript,
 produced figures, and developed software necessary to perform the analyses. All
-authors performed research and provided intellectual contributions. R.J.T
+authors performed research and provided intellectual contributions. R.J.T.
 acquired funding.}
@@
 September, 2020, we do not have true vintage data before then.  Furthermore, the
-latency of the signal was always at least one week through 2020.  However, this
-signal is never revised after initial publication (confirmed via personal
-communication with the Google team that produces this signal) and furthermore
+latency of the signal was always at least one week through 2020.
+% However, this signal is never revised after initial publication (confirmed via
+% personal communication with the Google team that produces this signal)
+However, unlike the claims-based signals, there is no reason for revisions to
+occur after initial publication, 
+and furthermore,
 the latency of the signal is not an unavoidable property of the data type, so we

### 2021-06-23 indicators/code/indicator-paper-figures.Rmd

---
title: COVIDcast Indicators figures and tables
author: Alex Reinhart, et al.
date: June 23, 2021
---

This code accompanies the paper "An Open Repository of Real-Time COVID-19
Indicators" by Reinhart et al. It requires the [covidcast
package](https://cran.r-project.org/package=covidcast) for R, plus the packages
listed below.

```{r setup, message=FALSE}
library(covidcast)
library(gridExtra)
library(ggplot2)
library(scales)
library(dplyr)
library(purrr)
```

This file can be knit to HTML using
`rmarkdown::render("indicator-paper-figures.Rmd")` from within R, using the
[rmarkdown package](https://cran.r-project.org/package=rmarkdown). By default it
also saves publication-ready PDF versions of figures to the `../paper/fig/`
directory.

## Figure 1

National trends, from April 2020 to April 2021, of four signals in the COVIDcast
API.

```{r figure-1, message=FALSE, cache=TRUE}
SOURCES <- c("jhu-csse", "chng", "chng", "fb-survey")
SIGNALS <- c("confirmed_7dav_incidence_prop",
             "smoothed_adj_outpatient_cli",
             "smoothed_adj_outpatient_covid",
             "smoothed_wcli")
cs <- covidcast_signals(
  SOURCES,
  SIGNALS,
  "2020-04-15", "2021-04-15",
  geo_type = "nation"
)

case_max <- max(cs[[1]]$value)

rescaled <- aggregate_signals(cs, format = "long") %>%
  group_by(signal) %>%
  mutate(value = value / max(value, na.rm = TRUE) * case_max) %>%
  ungroup()

colors <- c("black", scales::hue_pal()(5))
names <- c("Cases", "CHNG-CLI", "CHNG-COVID", "CTIS-CLI",
          "CTIS-CLI-in-community", "Quidel-TPR")
rescaled$signal <- factor(rescaled$signal, labels = names[1:4])

ggplot(rescaled, aes(x = time_value, y = value, color = signal)) +
  geom_line() +
  scale_color_manual(breaks = names[1:4], values = colors[1:4]) +
  labs(x = "Date", y = "Signal value (scaled)") +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
               date_minor_breaks = "1 month") +
  theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank())

ggsave("../paper/fig/time_trends_national.pdf", width = 5, height = 3)
```

# Figure 2

Geo-wise correlations with case rates, from April 15, 2020 to April 15, 2021,
calculated over all counties for which all signals were available and which had
at least 500 cumulative cases by the end of this period.

First, we fetch the required data from the API. This may take a while, since it
fetches the entire county-level history of several signals with large coverage.

```{r figure-2-fetch, message=FALSE, warning=FALSE, cache=TRUE}
sources <- c("chng", "chng", "fb-survey", "fb-survey", "quidel")
signals <- c("smoothed_adj_outpatient_cli",
             "smoothed_adj_outpatient_covid",
             "smoothed_cli",
             "smoothed_whh_cmnty_cli",
             "covid_ag_smoothed_pct_positive")

start_day <- "2020-04-15"
end_day <- "2021-04-15"
as_of_day <- "2021-06-01"

df_signals <- vector("list", length(signals))
for (i in seq_along(signals)) {
  df_signals[[i]] <- covidcast_signal(sources[i], signals[i], start_day,
                                      end_day, as_of = as_of_day)
}

# Fetch JHU-CSSE confirmed case incidence proportion (smoothed with 7-day
# trailing average)
df_cases <- covidcast_signal("jhu-csse", "confirmed_7dav_incidence_prop",
                             start_day, end_day, as_of = as_of_day)
n <- length(signals)
```

Next, we find counties with at least 500 cumulative cases.

```{r figure-2-counties, message=FALSE, cache=TRUE}
case_num <- 500
geo_values <- covidcast_signal("jhu-csse", "confirmed_cumulative_num",
                               max(df_cases$time_value),
                               max(df_cases$time_value),
                               as_of = as_of_day) %>%
  filter(value >= case_num) %>%
  pull(geo_value)

for (i in seq_along(signals)) {
  geo_values <- intersect(geo_values, df_signals[[i]]$geo_value)
}

colors <- c("black", scales::hue_pal()(5))
names <- c("Cases", "CHNG-CLI", "CHNG-COVID", "CTIS-CLI",
           "CTIS-CLI-in-community", "Quidel-TPR")
```

There are `r length(geo_values)` counties that have at least 500 cumulative
cases and have measurements for all the signals desired.

```{r figure-2, cache=TRUE}
df_cor <- vector("list", n)
for (i in 1:n) {
  df_cor[[i]] <- covidcast_cor(df_signals[[i]] %>%
                                 filter(geo_value %in% geo_values),
                               df_cases %>%
                                 filter(geo_value %in% geo_values),
                               by = "time_value", method = "spearman")
  df_cor[[i]]$signal <- names[i]
}

df <- do.call(rbind, df_cor)
df$signal <- factor(df$signal, labels = names[2:6])

ggplot(df, aes(x = time_value, y = value)) +
  geom_line(aes(color = signal)) +
  scale_color_manual(breaks = names[2:6], values = colors[2:6]) +
  ylim(0, NA) +
  labs(x = "Date", y = "Correlation") +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
               date_minor_breaks = "1 month") +
  theme_bw() +
  guides(col = guide_legend(nrow = 2)) +
  theme(legend.position = "bottom", legend.title = element_blank())

ggsave("../paper/fig/geo_wise_corr.pdf", width = 5, height = 3.5)
```

## Figure 3

Time-wise correlations with case rates, from April 15, 2020 to April 15, 2021.

```{r figure-3, warning=FALSE, cache=TRUE}
df_cor <- vector("list", length(signals))

for (i in seq_along(signals)) {
  df_cor[[i]] <- covidcast_cor(df_signals[[i]] %>%
                                 filter(geo_value %in% geo_values),
                               df_cases %>%
                                 filter(geo_value %in% geo_values),
                               by = "geo_value", method = "spearman")
  df_cor[[i]]$signal <- names[i]
}

df <- do.call(rbind, df_cor)
df$signal <- factor(df$signal, labels = names[2:6])

ggplot(df, aes(value)) +
  geom_density(aes(color = signal, fill = signal), alpha = 0.4) +
  labs(x = "Correlation", y = "Density") +
  scale_color_manual(breaks = names[2:6], values = colors[2:6]) +
  scale_fill_manual(breaks = names[2:6], values = colors[2:6]) +
  theme_bw() +
  guides(col = guide_legend(nrow = 2)) +
  theme(legend.position = "bottom", legend.title = element_blank())

ggsave("../paper/fig/time_wise_correlation.pdf", width = 5, height = 3.5)
```

## Figure 4

```{r figure-4, message=FALSE, cache=TRUE}
sa_fips <- "48029"
sa_start <- "2020-06-15"
sa_end <- "2020-08-15"
sa_anomaly_date <- as.Date("2020-07-16")

sa_cases <- covidcast_signal("jhu-csse", "confirmed_7dav_incidence_num",
                             sa_start, sa_end, geo_values = sa_fips)

sa_cli <- covidcast_signal("fb-survey", "smoothed_whh_cmnty_cli",
                            sa_start, sa_end, geo_values = sa_fips)

g1 <- ggplot(sa_cases, aes(x = time_value, y = value)) +
  geom_vline(xintercept = sa_anomaly_date, color = "gray", linetype = "dashed",
             size = 0.75, alpha = 0.75) +
  geom_line() +
  labs(x = "", y = "Cases", title = "Cases") +
  theme_bw()

g2 <- ggplot(sa_cli, aes(x = time_value, y = value)) +
  geom_vline(xintercept = sa_anomaly_date, color = "gray", linetype = "dashed",
             size = 0.75, alpha = 0.75) +
  geom_line() +
  labs(x = "", y = "% CLI-in-community", title = "CTIS-CLI-in-community") +
  scale_y_continuous(labels = label_percent(scale = 1, accuracy = 1),
                     limits = c(NA, 75)) +
  theme_bw()

grid.arrange(g1, g2, ncol = 2)

g <- arrangeGrob(g1, g2, ncol = 2)

ggsave("../paper/fig/bexar_compare.pdf", g, width = 6, height = 4)
```

## Figure 5

Estimated percentage of doctor's visits due to COVID-like illness displayed
across multiple issue dates, with later issue dates adding additional data and
revising past data from prior issue dates.

```{r figure-5, message=FALSE, warning=FALSE, cache=TRUE}
as_ofs <- seq(as.Date("2020-06-08"), as.Date("2020-08-01"), by = "week")
states <- c("az", "ca", "ny", "tx")
dv_as_of <- map_dfr(as_ofs, function(as_of) {
  covidcast_signal(data_source = "doctor-visits", signal = "smoothed_adj_cli",
                   start_day = "2020-06-01", end_day = as_of - 1L,
                   geo_type = "state", geo_values = states, as_of = as_of) %>%
    mutate(as_of = as_of)
})

## Now plot the each "as of" time series curve, faceted by state
dv_as_of %>%
  mutate(geo_label = factor(
    geo_value, levels = states,
    labels = abbr_to_name(states, ignore.case = TRUE))) %>%
  mutate(as_of = factor(format(as.Date(as_of), "%b %d"))) %>%
  ggplot(aes(x = time_value, y = value, color = as_of)) +
  geom_line() +
  facet_wrap(vars(geo_label), nrow = 2) +
  labs(color = "As of:", x = "Date", y = "% doctor's visits due to CLI") +
  expand_limits(y = 0) +
  scale_y_continuous(labels = label_percent(scale = 1, accuracy = 1),
                     breaks = seq(0, 12, by = 4)) +
  theme_bw() +
  theme(legend.pos = "bottom", legend.direction = "horizontal")

ggsave("../paper/fig/dv_as_of.pdf", width = 5, height = 4)
```

## Figure 6

Percentage point increase in SafeGraph estimates of the percentage of people
staying completely at home, based on aggregated mobile device data, between
March 1, 2020 and April 15, 2020.

```{r figure-6, message=FALSE, cache=TRUE}
begin_date <- "2020-03-01"
end_date <- "2020-04-15"

begin_data <- covidcast_signal("safegraph", "completely_home_prop_7dav",
                               begin_date, begin_date, geo_type = "state")
end_data <- covidcast_signal("safegraph", "completely_home_prop_7dav",
                             end_date, end_date, geo_type = "state")

diffs <- begin_data %>%
  inner_join(end_data, by = "geo_value", suffix = c(".begin", ".end")) %>%
  mutate(value = 100 * (value.end - value.begin)) %>%
  select(data_source = data_source.end, signal = signal.end, geo_value, value, issue = issue.end) %>%
  mutate(time_value = end_date) %>%
  as.data.frame %>%
  as.covidcast_signal(signal = "safegraph-change",
                      geo_type = "state")

plot(diffs, plot_type = "choro", range = c(10, 30),
     title = "", choro_params = list(legend_n = 5))

ggsave("../paper/fig/mobility-drop.pdf", width = 5, height = 4)
```

### 2021-07-12 forecast/code/trajectory.R
Light reorganization

@@ -2,8 +2,8 @@ library(covidcast)
 library(evalcast)
 library(tidyverse)
 
-ny_preds <- readRDS("../data/ny_predictions.RDS")
-ny_actuals <- readRDS("../data/ny_actuals.RDS")
+ny_preds <- readRDS("data/ny_predictions.RDS")
+ny_actuals <- readRDS("data/ny_actuals.RDS")
 
 pd <- evalcast:::setup_plot_trajectory(ny_preds, side_truth = ny_actuals)
 pd$truth_df <- pd$truth_df %>% rename(target_end_date = time_value)

### 2021-07-12 indicators/paper/paper.tex
Be more precise with survey name (COVID-19, not COVID)

@@
 \paragraph{Internet-Based Surveys} Conducted in partnership with Facebook,
-Delphi's COVID Trends and Impact Survey receives an average of 50,000 responses
+Delphi's COVID-19 Trends and Impact Survey receives an average of 50,000 responses
 daily, and has received over 20 million responses since April 2020
@@
 \cite{SafeGraphSocial, SafeGraphPatterns}. & 2019-01-01 & County \\
-  COVID Trends and Impact Survey & COVID symptoms, social distancing
+  COVID-19 Trends and Impact Survey & COVID symptoms, social distancing
 behaviors, mental health, economic impact, behavior (e.g., mask-wearing,
@@
 and backfill. There is also a dedicated dashboard for exploring results from the
-COVID Trends and Impact Survey. The visualizations have been continually
+COVID-19 Trends and Impact Survey. The visualizations have been continually
 improved as new sources of data arrive, and in response to interviews with users
@@
   with confirmed COVID-19, based on the same claims data.
-\item COVID Trends and Impact Survey CLI (CTIS-CLI): The estimated percentage
+\item COVID-19 Trends and Impact Survey CLI (CTIS-CLI): The estimated percentage
   of the population with COVID-like illness based on Delphi's surveys of
   Facebook users.
-\item COVID Trends and Impact Survey CLI in the community
+\item COVID-19 Trends and Impact Survey CLI in the community
   (CTIS-CLI-in-community): The estimated percentage of the population who know
@@
 patterns, internet search trends, mask wearing, and vaccine hesitancy, to name
-just a few. (The latter two signals are derived from the COVID Trends and Impact
+just a few. (The latter two signals are derived from the COVID-19 Trends and Impact
 Survey; our companion paper on this survey gives a more detailed view of its

### 2021-07-12 indicators/paper/paper-medrxiv.tex
Add medRxiv version+supplement, remove Google authors temporarily

--- 2021-07-12 indicators/paper/paper.tex
@@
-\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
+\documentclass[11pt]{article}
 % Use the lineno option to display guide line numbers if required.
 
-\templatetype{pnasresearcharticle} % Choose template
+% \templatetype{pnasresearcharticle} % Choose template
 % {pnasresearcharticle} = Template for a two-column research article
@@
 % {pnasinvited} %= Template for a PNAS invited submission
+
+\usepackage{authblk}
+\usepackage{amsmath,amsfonts,amssymb}
+\usepackage[numbers,square,sort&compress]{natbib}
+\usepackage[colorlinks=true, allcolors=blue]{hyperref}
+\usepackage{graphicx}
+\usepackage{geometry}
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
+\usepackage{array}
+\usepackage{lscape}
 
@@
 % Middle authors
-\author[d]{Sumit Agrawal}
+% \author[d]{Sumit Agrawal}
 \author[e]{Wael Al Saeed}
@@
 \author[b]{Brian Clark}
-\author[d]{Sarah Colquhoun}
+% \author[d]{Sarah Colquhoun}
 \author[b]{Nat DeFries}
-\author[d]{David C. Farrow}
+% \author[d]{David C. Farrow}
 \author[i]{Jodi Forlizzi}
-\author[d]{Jed Grabman}
+% \author[d]{Jed Grabman}
 \author[b]{Samuel Gratzl}
@@
 \author[i]{Robin Han}
-\author[d]{Kate Harwood}
+% \author[d]{Kate Harwood}
 \author[a,b]{Addison J. Hu}
-\author[d]{Raphael Hyde}
+% \author[d]{Raphael Hyde}
 \author[h]{Sangwon Hyun}
@@
 \author[n]{Daniel J. McDonald}
-\author[d]{Phillip McGuinness}
+% \author[d]{Phillip McGuinness}
 \author[o,p]{Balasubramanian Narasimhan}
-\author[d]{Michael P. O'Brien}
+% \author[d]{Michael P. O'Brien}
 \author[a,b]{Natalia L. Oliveira}
@@
 \author[e]{Dawn Rucker}
-\author[d]{Chris Scott}
+% \author[d]{Chris Scott}
 \author[q]{Nigam H. Shah}
@@
 \author[s]{Noah Simon}
-\author[d]{Benjamin Y. Smith}
+% \author[d]{Benjamin Y. Smith}
 \author[b]{Vishakha Srivastava}
@@
 \author[a,b]{Larry Wasserman}
-\author[d]{Benjamin Weaver}
+% \author[d]{Benjamin Weaver}
 \author[t]{Jeremy C. Weiss}
-\author[d]{Spencer Whitman}
+% \author[d]{Spencer Whitman}
 \author[i]{Kristin Williams}
@@
 \affil[c]{Computational Biology Department, Carnegie Mellon University}
-\affil[d]{Google.org}
+% \affil[d]{Google.org}
 \affil[e]{Computer Science Department, Carnegie Mellon University}
@@
 
-% Please give the surname of the lead author for the running footer
-\leadauthor{Reinhart}
-
 % Please add a significance statement to explain the relevance of your work
-\significancestatement{To study the COVID-19 pandemic, its effects on society,
-  and measures for reducing its spread, researchers need detailed data on the
-  course of the pandemic. Standard public health data streams suffer
-  inconsistent reporting and frequent, unexpected revisions. They also miss
-  other aspects of a population's behavior that are worthy of consideration. We
-  present an open database of COVID signals in the United States, measured at
-  the county level and updated daily. This includes traditionally reported COVID
-  cases and deaths, and many others: measures of mobility, social distancing,
-  internet search trends, self-reported symptoms, and patterns of COVID-related
-  activity in de-identified medical insurance claims. The database provides all
-  signals in a common, easy-to-use format, empowering both public health
-  research and operational decision-making.
-}
+% \significancestatement{To study the COVID-19 pandemic, its effects on society,
+%   and measures for reducing its spread, researchers need detailed data on the
+%   course of the pandemic. Standard public health data streams suffer
+%   inconsistent reporting and frequent, unexpected revisions. They also miss
+%   other aspects of a population's behavior that are worthy of consideration. We
+%   present an open database of COVID signals in the United States, measured at
+%   the county level and updated daily. This includes traditionally reported COVID
+%   cases and deaths, and many others: measures of mobility, social distancing,
+%   internet search trends, self-reported symptoms, and patterns of COVID-related
+%   activity in de-identified medical insurance claims. The database provides all
+%   signals in a common, easy-to-use format, empowering both public health
+%   research and operational decision-making.
+% }
 
 % Please include corresponding author, author contribution and author declaration information
-\authorcontributions{Author contributions: A.R., L.B., M.J., A.R., J.T., A.C,
-  E.J.C., N.D., D.C.F, A.J.H., S.H., M.J., J.K., W.L.M.K, Z.C.L., L.M., N.L.O.,
-  J.C.W, R.R., and R.J.T. obtained and managed data;
-  A.R., J.B., J.F., Z.C.L., L.M., D.J.M., A.P., S.R., D.R., C.S. N.S., R.T.,
-  V.V., L.W., B.W., J.C.W, S.W., R.R., and R.J.T. designed software and
-  research;
-  A.R., L.B., M.J., A.R., J.T., W.A.S., T.A., A.B., J.B., L.B., A.A.C., A.C.,
-  E.J.C., B.C., S.C., N.D., D.C.F., J.G., S.G., A.G., G.H., R.H., K.H., A.J.H.,
-  S.H., M.J., A.J., J.K., A.K., M.X.L., L.M., K.M., D.J.M., B.N., M.P.O.,
-  N.L.O., P.P., C.A.P., S.R., C.S., V.S, J.S, D.S., N.S., B.Y.S., V.S., S.T.,
-  J.T., R.T., J.C.W, and R.J.T. developed software;
-  S.A., W.A.S., A.C., B.C., D.C.F., S.G., G.H., K.M., D.J.M., B.N., C.S., J.S.,
-  D.S., B.Y.S., and A.K.V.N. managed technical infrastructure and systems;
-  A.R., L.B., M.J., A.R., J.T, J.B., A.C., E.J.C., N.D., D.C.F., A.G., K.H.,
-  A.J.H., R.H., S.H., A.J., J.K., K.L., Y.J.L, Z.C.L., L.M., D.J.M., P.M., B.N.,
-  N.L.O., P.P., C.A.P., S.R., N.S., V.S., J.S., D.S., N.S., S.T., J.T., R.T.,
-  V.V., L.W., J.C.W., R.R., and R.J.T. performed research;
-  R.R. and R.J.T. acquired funding;
-  and A.R., L.B., M.J., A.R., J.T., and R.J.T. wrote the paper.
-}
-\authordeclaration{No competing interests to declare.}
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be
-  addressed. E-mail: areinhar@stat.cmu.edu}
-\equalauthors{\textsuperscript{2}L.B., M.J., A.R., and J.T. contributed equally
-  to this work.}
+% Googlers: SA, SC, DCF, JG, KH, RH, PM, MPO, CS, BYS, BW, SW
+% \authorcontributions{Author contributions: A.R., L.B., M.J., A.R., J.T., A.C,
+%   E.J.C., N.D., A.J.H., S.H., M.J., J.K., W.L.M.K, Z.C.L., L.M., N.L.O.,
+%   J.C.W, R.R., and R.J.T. obtained and managed data;
+%   A.R., J.B., J.F., Z.C.L., L.M., D.J.M., A.P., S.R., D.R., N.S., R.T.,
+%   V.V., L.W., J.C.W, R.R., and R.J.T. designed software and
+%   research;
+%   A.R., L.B., M.J., A.R., J.T., W.A.S., T.A., A.B., J.B., L.B., A.A.C., A.C.,
+%   E.J.C., B.C., N.D., S.G., A.G., G.H., R.H., A.J.H.,
+%   S.H., M.J., A.J., J.K., A.K., M.X.L., L.M., K.M., D.J.M., B.N.,
+%   N.L.O., P.P., C.A.P., S.R., V.S, J.S, D.S., N.S., V.S., S.T.,
+%   J.T., R.T., J.C.W, and R.J.T. developed software;
+%   W.A.S., A.C., B.C., S.G., G.H., K.M., D.J.M., B.N., J.S.,
+%   D.S., and A.K.V.N. managed technical infrastructure and systems;
+%   A.R., L.B., M.J., A.R., J.T, J.B., A.C., E.J.C., N.D.,  A.G.,
+%   A.J.H., R.H., S.H., A.J., J.K., K.L., Y.J.L, Z.C.L., L.M., D.J.M., B.N.,
+%   N.L.O., P.P., C.A.P., S.R., N.S., V.S., J.S., D.S., N.S., S.T., J.T., R.T.,
+%   V.V., L.W., J.C.W., R.R., and R.J.T. performed research;
+%   R.R. and R.J.T. acquired funding;
+%   and A.R., L.B., M.J., A.R., J.T., and R.J.T. wrote the paper.
+% }
+% \authordeclaration{No competing interests to declare.}
+% \correspondingauthor{\textsuperscript{1}To whom correspondence should be
+%   addressed. E-mail: areinhar@stat.cmu.edu}
+% \equalauthors{\textsuperscript{2}L.B., M.J., A.R., and J.T. contributed equally
+%   to this work.}
+\footnotetext[1]{To whom correspondence should be addressed. E-mail:
+  areinhar@stat.cmu.edu}
+\footnotetext[2]{L.B., M.J., A.R., and J.T. contributed equally to this work.}
 
 % At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
-\keywords{COVID-19 $|$ open data $|$ digital surveillance $|$ medical insurance
-  claims $|$ internet surveys}
+% \keywords{COVID-19 $|$ open data $|$ digital surveillance $|$ medical insurance
+%   claims $|$ internet surveys}
+
+\begin{document}
+\maketitle
+
+\clearpage
 
@@
 
-\dates{This manuscript was compiled on \today}
-\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
-
-\begin{document}
-
-\maketitle
-\thispagestyle{firststyle}
-\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
-
-\dropcap{P}ublic health decision makers, healthcare providers, epidemiological
+Public health decision makers, healthcare providers, epidemiological
 researchers, employers, institutions, and the general public benefit from
 promptly and readily accessible data regarding COVID-19 activity levels,
-countermeasures, and pandemic impact.  Real-time indicators of COVID-19
-activity levels, such as statistics on cases, deaths, test positivity, and
+countermeasures, and pandemic impact. Real-time indicators of COVID-19 activity
+levels, such as statistics on cases, deaths, test positivity, and
 hospitalizations, enable reports and interactive dashboard applications for
 situational awareness \cite{Dong:2020, NYTimes, USAFacts}, and are essential for
-most analyses of the pandemic.  These data are available for locations across
-the United States from a number of official sources and independent aggregators
-in varied and inconsistent formats.  Different data types and sources vary in
+most analyses of the pandemic. These data are available for locations across the
+United States from a number of official sources and independent aggregators in
+varied and inconsistent formats. Different data types and sources vary in
 timeliness, based on when measured events occur in the progression of the
@@
 
+\begin{landscape}
 \begin{table*}[t]
@@
   limited form); the second group is mirrored from public sources.}
-\begin{tabular}{>{\raggedright}p{1.2in} p{4.0in} l >{\raggedright\arraybackslash}p{0.5in}}
+\begin{tabular}{>{\raggedright}p{1.2in} p{5.0in} l >{\raggedright\arraybackslash}p{0.5in}}
   \toprule
@@
 \end{tabular}
-\addtabletext{$^*$Available at $>60$\% of counties; $^{**}$available at
+{\raggedright $^*$Available at $>60$\% of counties; $^{**}$available at
   20--60\% of counties; $^{***}$available at $<20$\% of counties.  For some
@@
 \end{table*}
+\end{landscape}
 
@@
 \begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/time_trends_national.pdf}
+  \centering
+  \includegraphics[width=0.6\columnwidth]{fig/time_trends_national.pdf}
   \caption{National trends, from April 2020 to April 2021, of four signals in
@@
 \begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/geo_wise_corr.pdf}
+  \centering
+  \includegraphics[width=0.6\columnwidth]{fig/geo_wise_corr.pdf}
   \caption{Geo-wise correlations with case rates, from April 15, 2020 to April
@@
 \begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/time_wise_correlation.pdf}
+  \centering
+  \includegraphics[width=0.6\columnwidth]{fig/time_wise_correlation.pdf}
   \caption{Time-wise correlations with case rates, from April 15, 2020 to April
@@
 \begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/bexar_compare.pdf}
+  \centering
+  \includegraphics[width=0.6\columnwidth]{fig/bexar_compare.pdf}
   \caption{Reported cases per day in Bexar County, Texas during the summer of
@@
 \begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/dv_as_of.pdf}
+  \centering
+  \includegraphics[width=0.6\columnwidth]{fig/dv_as_of.pdf}
   \caption{Estimated percentage of doctor's visits due to COVID-like illness
@@
   \centering
-  \includegraphics[width=\columnwidth]{fig/mobility-drop.pdf}
+  \includegraphics[width=0.6\columnwidth]{fig/mobility-drop.pdf}
   \caption{Percentage point increase in SafeGraph estimates of the percentage of
@@
 
-\acknow{We thank Carrie Reed, Matt Biggerstaff, Michael Johansson, Rachel
-  Slayton, Velma Lopez, Jo Walker and others on the CDC COVID-19 Modeling Team;
-  Hal Varian, Brett Slatkin, and others on Google's Surveys team;
-  Erin Hattersley, Rasmi Elasmar, and others at Google.org;
-  Evgeniy Gabrilovich and others on Google's Health team;
-  Curtiss Cobb and others on Facebook's  Demography and Survey Science, Data for
-  Good and Health teams;
-  Alex Smola and others at Amazon Web Services;
-  Tim Suther and others at Change Healthcare;
-  Paul Nielsen and others at Optum;
-  John Tamerius and others at Quidel; and
-  Ross Epstein and others at SafeGraph.
-  This material is based on work supported by gifts from Facebook, Google.org,
-  the McCune Foundation, and Optum; Centers for Disease Control and Prevention
-  (CDC) grant U01IP001121; National Science Foundation Graduate Research
-  Fellowship Program (NSF GRFP) award DGE1745016; and the Center for Machine
-  Learning and Health (CMLH) at Carnegie Mellon.}
+\subsection*{Acknowledgments}
 
-\showacknow{} % Display the acknowledgments section
+We thank Carrie Reed, Matt Biggerstaff, Michael Johansson, Rachel Slayton, Velma
+Lopez, Jo Walker and others on the CDC COVID-19 Modeling Team; Hal Varian, Brett
+Slatkin, and others on Google's Surveys team; Erin Hattersley, Rasmi Elasmar,
+and others at Google.org; Evgeniy Gabrilovich and others on Google's Health
+team; Curtiss Cobb and others on Facebook's Demography and Survey Science, Data
+for Good and Health teams; Alex Smola and others at Amazon Web Services; Tim
+Suther and others at Change Healthcare; Paul Nielsen and others at Optum; John
+Tamerius and others at Quidel; and Ross Epstein and others at SafeGraph. This
+material is based on work supported by gifts from Facebook, Google.org, the
+McCune Foundation, and Optum; Centers for Disease Control and Prevention (CDC)
+grant U01IP001121; National Science Foundation Graduate Research Fellowship
+Program (NSF GRFP) award DGE1745016; and the Center for Machine Learning and
+Health (CMLH) at Carnegie Mellon.
 
 % Bibliography
+\bibliographystyle{unsrtnat}
 \bibliography{../../common/covidcast.bib}
 
+\clearpage
+
+\appendix
+
+\section*{Supplemental figures}
+
+\clearpage
+
+\input{supplement-medrxiv.tex}
+
 \end{document}

### 2021-07-12 indicators/paper/supplement-medrxiv.tex
Add medRxiv version+supplement, remove Google authors temporarily

\begin{figure}
\includegraphics[width=\textwidth]{fig/severity-pyramid.pdf}
\caption{The epidemiological ``severity pyramid'' represents the progression of
cases from the public, through infection, through increasingly severe stages of
disease. The annotations here represent the data sources collected by Delphi's
COVIDcast Epidata API.}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/state-trend-grids-1-1}

}

\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. Cases are reported cases per 100,000 population. Other signals are scaled to have the same global maximum across all counties and times, so they can be presented in the same range. (part 1 of 4)}\label{fig:state-trend-grids-1}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/state-trend-grids-2-1}

}

\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. (part 2 of 4)}\label{fig:state-trend-grids-2}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/state-trend-grids-3-1}

}

\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. (part 3 of 4)}\label{fig:state-trend-grids-3}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/state-trend-grids-4-1}

}

\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. (part 4 of 4)}\label{fig:state-trend-grids-4}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/county-trend-grids-1-1}

}

\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. Cases are reported cases per 100,000 population. Other signals are scaled to have the same global maximum across all counties and times, so they can be presented in the same range. (part 1 of 4)}\label{fig:county-trend-grids-1}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/county-trend-grids-2-1}

}

\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 2 of 4)}\label{fig:county-trend-grids-2}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/county-trend-grids-3-1}

}

\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 3 of 4)}\label{fig:county-trend-grids-3}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/county-trend-grids-4-1}

}

\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 4 of 4)}\label{fig:county-trend-grids-4}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/hospitalization_time_trends_national-1}

}

\caption{National trends, from April 2020 to April 2021, of HHS-reported confirmed COVID-19 hospital admissions, plus another hospitalization signal in the COVIDcast API. Here, the HHS data has been limited to values from 2020-08-01 onward due to changes in reporting behavior, and processed into a 7-day trailing average rate per 100,000 resident population using Census Bureau estimates for 2019. HSP-Hosp is the percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners, smoothed in time and adjusted for systematic day-of-week effects (HSP-Hosp is available from an earlier date when working at finer geographical resolutions). HSP-Hosp has been scaled to have the same maximum value as the HHS data.}\label{fig:hospitalization_time_trends_national}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/hosp-correlations-by-time-1}

}

\caption{Geo-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated for all times with sufficient available data within this period, over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which both signals are available.}\label{fig:hosp-correlations-by-time}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/hosp-correlations-by-state-1}

}

\caption{Time-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which both signals are available.}\label{fig:hosp-correlations-by-state}
\end{figure}

### 2021-07-17 forecast/paper/supplement.Rmd
forecasting code refactor

@@
 library(evalcast)
+library(here)
 
-source("../code/eval_funs.R")
-source("../code/hotspot_funs.R")
+source(here("code", "figures", "results_processing_and_plotting"))
+source(here("code", "figures", "eval_funs.R"))
+source(here("code", "figures", "hotspot_funs.R"))
 
-path_to_data <- "~/Documents/COVID-Delphi/fcast-indicators-paper/temp_data/"
 ```
@@
 ```{r data-loading}
-actuals <- readRDS(file.path(path_to_data, "actuals.RDS"))
-fcasts_honest <- readRDS(file.path(path_to_data, "results_honest.RDS")) %>%
+actuals <- readRDS(here("data", "actuals.RDS"))
+fcasts_honest <- readRDS(here("data", "results_honest.RDS")) %>%
   process_res_cases(actuals) %>%
   filter(forecaster != "gs_inherit") 
-fcasts_finalized <- readRDS(file.path(path_to_data, "results_dishonest.RDS")) %>%
+fcasts_finalized <- readRDS(here("data", "results_dishonest.RDS")) %>%
   process_res_cases(actuals) %>%
   filter(forecaster != "gs_inherit")
-hotspots_honest <- readRDS(file.path(path_to_data, "hotspots_honest.RDS")) %>%
+hotspots_honest <- readRDS(here("data", "hotspots_honest.RDS")) %>%
   process_res_hotspots() %>%
   filter(forecaster != "gs_inherit")
-hotspots_finalized <- readRDS(
-  file.path(path_to_data, "hotspots_dishonest.RDS")) %>%
+hotspots_finalized <- readRDS(here("data", "hotspots_dishonest.RDS")) %>%
   process_res_hotspots() %>%
   filter(forecaster != "gs_inherit")
-cases7dav <- readRDS(file.path(path_to_data, "cases7dav.RDS")) %>%
+cases7dav <- readRDS(here("data", "cases7dav.RDS")) %>%
   rename(case_num = value)
@@
 ```{r hrr-names}
-hrr_names <- read_csv(file.path(path_to_data, "Hospital_Referral_Regions.csv"), 
+hrr_names <- read_csv(here("data", "Hospital_Referral_Regions.csv"), 
                       col_types = "_ccc__") %>%
@@
 ```{r bootstrap-loading}
-fcasts_booted <- readRDS(
-  file.path(path_to_data,"results_honest_bootstrapped.RDS")) %>%
+fcasts_booted <- readRDS(here("data","results_honest_bootstrapped.RDS")) %>%
   process_res_cases(actuals) %>%
@@
 }
-hotspots_booted <- readRDS(
-  file.path(path_to_data, "hotspots_honest_bootstrapped.RDS")) %>%
+hotspots_booted <- readRDS(here("data", "hotspots_honest_bootstrapped.RDS")) %>%
   process_res_hotspots() %>%
@@
 ```{r up-down-processing}
-preds <- readRDS(file.path(path_to_data, "predictions_honest.RDS"))
+preds <- readRDS(here("data", "predictions_honest.RDS"))
 preds <- preds %>% 
@@
 
-pct_chng_df <- readRDS(file.path(
-  path_to_data, 
+pct_chng_df <- readRDS(here(
+  "data", 
   "jhu-csse_confirmed_7dav_incidence_num_2021-05-18.RDS")) %>%
@@
 
-all_sigs <- readRDS(file.path(
-  path_to_data, "all_signals_wide_as_of_2020-05-15.RDS"))
+all_sigs <- readRDS(here("data", "all_signals_wide_as_of_2020-05-15.RDS"))
 
@@
 By our leadingness/laggingness metric, DV-CLI is leading (but notlagging) cases over this period."}
-source("../code/ccf-dv-finalized.R")
+source(here("code", "figures", "ccf-dv-finalized.R"))
 gg
@@
 # Figure 1 in the manuscript
-source("../code/revisions.R")
+source(here("code", "figures", "revisions.R"))
 ```
@@
 # Figure 2 in the manuscript
-source("../code/ny-trajectory.R")
+source(here("code", "figures", "trajectory.R"))
 gg
@@
 ```{r traj-data, eval = FALSE}
-source("../code/trajectory_plot_funs.R")
-preds <- readRDS(file.path(path_to_data, "predictions_honest.RDS"))
+source(here("code", "tbd", "trajectory_plot_funs.R"))
+preds <- readRDS(here("data", "predictions_honest.RDS"))
 traj_best <- get_trajectory_plots(fcasts_honest, preds, actuals, hrr_tab, 

### 2021-07-17 forecast/code/deprecated/alden_evalcast.R

@@ -1,0 +1,69 @@
+#--------------------------------------------#
+# My "hack" of evalcast, in order to use already 
+# downloaded data.
+#--------------------------------------------#
+alden_get_predictions <- function(forecaster, name_of_forecaster, signals, forecast_dates, 
+          incidence_period = c("epiweek", "day"), apply_corrections = function(signals) signals, 
+          response_data_source = signals$data_source[1], response_data_signal = signals$signal[1], 
+          forecaster_args = list(),
+          offline_signal_dir = offline_signal_dir) 
+{
+  assert_that(is_tibble(signals), msg = "`signals` should be a tibble.")
+  out <- forecast_dates %>% map(~do.call(alden_get_predictions_single_date, 
+                                         list(forecaster = forecaster, signals = signals, forecast_date = .x, 
+                                              apply_corrections = apply_corrections, forecaster_args = forecaster_args,
+                                              offline_signal_dir = offline_signal_dir))) %>% 
+    bind_rows()
+  names(out$value) = NULL
+  out <- out %>% mutate(forecaster = name_of_forecaster, data_source = response_data_source, 
+                        signal = response_data_signal, target_end_date = get_target_period(.data$forecast_date, 
+                                                                                           incidence_period, .data$ahead)$end, incidence_period = incidence_period) %>% 
+    relocate(.data$forecaster, .before = .data$forecast_date)
+  class(out) <- c("predictions_cards", class(out))
+  out
+}
+
+alden_get_predictions_single_date <- function(forecaster, signals, forecast_date, apply_corrections, 
+          forecaster_args,offline_signal_dir) 
+{
+  # Addison's spoof of download_signal.
+  if (is.null(offline_signal_dir)) {
+    download_signal_function = download_signal
+  } else {
+    download_signal_function = function(data_source,
+                                        signal,
+                                        start_day,
+                                        end_day,
+                                        as_of,
+                                        geo_type,
+                                        geo_values) {
+      signal_fpath = sprintf('%s/%s_%s_%s.RDS',
+                             offline_signal_dir,
+                             data_source,
+                             signal,
+                             end_day)
+      message(sprintf('Reading signal from disk: %s',
+                      signal_fpath))
+      return(readRDS(signal_fpath))
+    }
+  }
+  forecast_date <- lubridate::ymd(forecast_date)
[…]

### 2021-07-17 forecast/code/deprecated/demo.R

@@ -1,0 +1,59 @@
+library(dplyr)
+library(tibble)
+library(covidcast)
+library(evalcast)
+
+response_source <- "jhu-csse"
+response_signal <- "confirmed_7dav_incidence_prop"
+incidence_period <- "day"
+ahead <- 7:21
+geo_type <- "hrr" 
+ntrain = 21
+forecast_dates <- seq(as.Date("2020-06-09"), as.Date("2020-06-10"), by = "day")
+
+start_day_baseline <- function(forecast_date) {
+  return(as.Date(forecast_date) - ntrain - 4 + 1)
+}
+
+# Example
+signals_baseline = tibble::tibble(
+                      data_source = response_source, 
+                      signal = response_signal,
+                      start_day = list(start_day_baseline),
+                      as_of = list(function(x) x), # Warning: ignored because reading from disk
+                             # (Specify honest vs dishonest by pointing to the each directory)
+                      geo_values='*',
+                      geo_type='hrr') # also ignored because reading from disk.
+
+preds = get_predictions(baseline_forecaster,
+                        'Baseline',
+                        signals_baseline,
+                        forecast_dates,
+                        incidence_period='day',
+                        offline_signal_dir='./honest_as_of/',
+                        forecaster_args=list(
+                            incidence_period='day',
+                            ahead=ahead)
+                        )
+                        
+# Example: only train on subset, e.g., for Google Symptoms
+signals_baseline_subset = tibble::tibble(
+                      data_source = response_source, 
+                      signal = response_signal,
+                      start_day = list(start_day_baseline),
+                      as_of = list(function(x) x),
+                      geo_values=list(c('1', '10')),
+                      geo_type='hrr')
+
+preds_subset = get_predictions(baseline_forecaster,
+                        'Baseline',
+                        signals_baseline_subset,
+                        forecast_dates,
+                        incidence_period='day',
+                        offline_signal_dir='./honest_as_of/',
+                        forecaster_args=list(
+                            incidence_period='day',
+                            ahead=ahead)
+                        )
+                        
+

### 2021-07-17 forecast/code/deprecated/process_results.R

@@ -1,0 +1,182 @@
+library(covidcast)
+library(dplyr)
+library(tibble)
+
+train_types = c('honest', 'dishonest')
+idx = 2
+train_type = train_types[idx]
+
+results = readRDS(sprintf('results/results_%s.RDS', train_type))
+actuals = readRDS('actuals.RDS')
+
+results = results %>% left_join (
+    actuals,
+    by=c('geo_value', 'target_end_date'),
+  )
+
+print(results %>% group_by (
+    forecaster,
+  ) %>% summarize (
+    count = n()
+  ))
+
+results %>% summarize (
+    wis_na = sum(is.na(wis)),
+    ae_na = sum(is.na(ae)),
+    actual_na = sum(is.na(actual)),
+  ) %>% print
+
+# drop unnecessary columns
+results = results %>% select (
+    -target_end_date,
+  )
+
+dim(results)
+
+results %>% group_by (
+    forecaster,
+  ) %>% summarize (
+    count = n(),
+  )
+
+# create google subset forecaster that inherits AR3 where missing
+gs_results = results %>% filter (
+    forecaster == 'AR3GSSAA3_Subset',
+  )
+AR3_results = results %>% filter (
+    forecaster == 'AR3',
+  )
+inherited_results = AR3_results %>% left_join (
+    gs_results %>% select(
+        -forecaster,
+        -actual,
+      ) %>% rename (
+        wis_gs = wis,
+        ae_gs = ae,
+      ),
+    by=c('geo_value', 'ahead', 'forecast_date')
+  ) %>% mutate (
+    forecaster = 'AR3_GSSAA3_Subset_InheritAR3',
+    wis = coalesce(wis_gs, wis),
+    ae = coalesce(ae_gs, ae),
+  ) %>% select (
+    -wis_gs,
+    -ae_gs,
+  )
+results = bind_rows(results, inherited_results)
+
+results %>% group_by (
+    forecaster,
+  ) %>% summarize (
+    count = n(),
+  )
+    
+na_results = results %>% group_by (
+    forecaster,
+  ) %>% summarize (
+    na_wis = sum(is.na(wis)),
+    na_ae = sum(is.na(ae)),
+  )
+print(na_results)
+
+dim(results)
+
+min_time = min(results$forecast_date)
+max_time = max(results$forecast_date)
+# get forecast_date case rates
+cases_df = covidcast_signal(data_source = "jhu-csse",
[…]

### 2021-07-17 forecast/code/download_signals/download_signals.R

@@ -1,0 +1,126 @@
+library(dplyr)
+library(tibble)
+library(covidcast)
+library(evalcast)
+library(here)
+
+# Parameters for determining which data (and how much) we need --------
+geo_type <- "hrr"
+ntrain = 21
+ahead = 7:21
+lags = c(0, 7, 14)
+hotspot_lag = 7
+end_forecast_date = '2021-03-31'
+named_function = function(x) lubridate::mdy("2021-05-18") # dishonest as_of
+
+honest_as_of_signals = tribble(
+  ~data_source,         ~signal,                              ~start_forecast_date,
+  'jhu-csse',           'confirmed_7dav_incidence_prop',      '2020-06-09',
+  'jhu-csse',           'confirmed_7dav_incidence_num',       '2020-06-09',
+  'fb-survey',          'smoothed_hh_cmnty_cli',              '2020-06-09',
+  'doctor-visits',      'smoothed_adj_cli',                   '2020-06-09',
+  'chng',               'smoothed_adj_outpatient_cli',        '2021-02-22',
+  'chng',               'smoothed_adj_outpatient_covid',      '2021-02-22',
+)
+
+dishonest_as_of_signals = tribble(
+  ~data_source,         ~signal,                              ~start_forecast_date,
+  'jhu-csse',           'confirmed_7dav_incidence_prop',      '2020-06-09',
+  'jhu-csse',           'confirmed_7dav_incidence_num',       '2020-06-09',
+  'fb-survey',          'smoothed_hh_cmnty_cli',              '2020-06-09',
+  'doctor-visits',      'smoothed_adj_cli',                   '2020-06-09',
+  'chng',               'smoothed_adj_outpatient_cli',        '2020-06-09',
+  'chng',               'smoothed_adj_outpatient_covid',      '2020-06-09',
+  'google-symptoms',    'sum_anosmia_ageusia_smoothed_search','2020-06-09',
+)
+
+
+
+
+# Functions ---------------------------------------------------------------
+
+signal_downloader <- function(df_list,
+                              forecast_date,
+                              offline_signal_dir = NULL) {
+  if (class(df_list)[1] == "list") df_list <- df_list[[1]]
+  data_source = unique(df_list$data_source)
+  signal = unique(df_list$signal)
+  if ((length(data_source) > 1) || (length(signal) > 1)) {
+    stop('Data source / signal not unique.')
+  }
+  saveRDS(
+    df_list,
[…]

### 2021-07-17 forecast/code/figures/eval_funs.R

@@ -1,0 +1,23 @@
+fcast_colors <- c(RColorBrewer::brewer.pal(5, "Set1"), "#000000")
+names(fcast_colors) <- c("CHNG-CLI", "CHNG-COVID", "CTIS-CLIIC", "DV-CLI",
+                         "Google-AA", "AR")
+
+
+Mean <- function(x) mean(x, na.rm = TRUE)
+Median <- function(x) median(x, na.rm = TRUE)
+TrimMean01 <- function(x) mean(x, trim = .01, na.rm = TRUE)
+TrimMean025 <- function(x) mean(x, trim = .025, na.rm = TRUE)
+TrimMean05 <- function(x) mean(x, trim = .05, na.rm = TRUE)
+TrimMean1 <- function(x) mean(x, trim = .1, na.rm = TRUE)
+GeoMean <- function(x) exp(mean(log(x), na.rm = TRUE))
+
+
+pct_change <- function(x, n = 14, col_name = "pct_change"){
+  # courtesy of Alden
+  dt <- c(0,-n)
+  y = covidcast:::apply_shifts(x,dt) %>%
+    rename(value = tidyselect::matches("value\\+"),
+           lag_value = tidyselect::matches("value\\-")) %>%
+    mutate(`:=`(!!col_name,(value - lag_value) / lag_value))
+  return(y)
+}

### 2021-07-17 forecast/code/figures/results_processing_and_plotting.R

@@ -1,15 +1,3 @@
-Mean <- function(x) mean(x, na.rm = TRUE)
-Median <- function(x) median(x, na.rm = TRUE)
-TrimMean01 <- function(x) mean(x, trim = .01, na.rm = TRUE)
-TrimMean025 <- function(x) mean(x, trim = .025, na.rm = TRUE)
-TrimMean05 <- function(x) mean(x, trim = .05, na.rm = TRUE)
-TrimMean1 <- function(x) mean(x, trim = .1, na.rm = TRUE)
-GeoMean <- function(x) exp(mean(log(x), na.rm = TRUE))
-
-fcast_colors <- c(RColorBrewer::brewer.pal(5, "Set1"), "#000000")
-names(fcast_colors) <- c("CHNG-CLI", "CHNG-COVID", "CTIS-CLIIC", "DV-CLI",
-                         "Google-AA", "AR")
-
 summarizer <- function(df, y, centerer, scaler, aggr, order_of_operations) {
   for (i in seq_along(order_of_operations)) {
     df <- switch(
@@ -27,7 +15,7 @@ summarizer <- function(df, y, centerer, scaler, aggr, order_of_operations) {
 
 
 plotter <- function(df, y, aggr,
-                    centerer = NULL, scaler = NULL, 
+                    centerer = NULL, scaler = NULL,
                     facet_by_period = FALSE,
                     order_of_operations = c("center", "scale", "aggr")) {
   df <- df %>% group_by(ahead, forecaster)
@@ -36,17 +24,17 @@ plotter <- function(df, y, aggr,
     period_labels <- c(jj = "June - July", as = "Aug - Sep", o = "Oct",
                        nd = "Nov - Dec", jm = "Jan - Mar")
     df <- df %>% group_by(period, .add = TRUE)
-    facet_layer <- facet_wrap(~period, scales = "free_y", 
+    facet_layer <- facet_wrap(~period, scales = "free_y",
                               labeller = labeller(period = period_labels))
   }
-  
+
   df <- summarizer(df, y, centerer, scaler, aggr, order_of_operations)
-  
-  
-  ggplot(df, aes(ahead, wis, color = forecaster)) + 
+
+
+  ggplot(df, aes(ahead, wis, color = forecaster)) +
     geom_line() + geom_point() +
     facet_layer +
-    theme_bw() + 
+    theme_bw() +
     xlab("Days ahead") +
[…]

### 2021-07-17 forecast/code/hotspots/auc.R

@@ -1,0 +1,73 @@
+# Compute area under curve.
+auc <- function(preds,actual,alpha = seq(.01,1,by = .01)){
+  n_pos = sum(actual == 1,na.rm = T)
+  n_neg = sum(actual == 0,na.rm = T)
+  
+  pred_pos <- preds[actual == 1]
+  pred_neg <- preds[actual == 0]
+  
+  power <- roc(preds,actual,alpha) %>% arrange(alpha) 
+  
+  slider::slide_dbl(power,
+                    .before = 1,
+                    .f = ~ .x$result[2] * (.x$alpha[2] - .x$alpha[1])) %>%
+    sum(na.rm = T)
+}
+
+
+# Compute a receiver operator curve (ROC curve).
+# The ROC curve describes the power for a given level of type I error.
+# alpha: level of type I error
+roc <- function(preds,actual,alpha){
+  pred_pos <- sort(preds[actual == 1])
+  pred_neg <- sort(preds[actual == 0])
+  n_pos = length(pred_pos)
+  n_neg = length(pred_neg)
+  
+  power <- numeric(length(alpha))
+  for(ii in 1:length(alpha)){
+    a <- alpha[ii]
+    t = if(a == 1) 0 else pred_neg[floor(n_neg * (1 - a))]
+    power[ii] <- sum(pred_pos > t)/n_pos
+  }
+  return(data.frame(alpha = alpha,result = power))
+}
+
+# Aggregation function is area under the receiver-operator curve (ROC) curve,
+# aka area under curve (AUC). x is a vector of (signed) errors, i.e.
+# predicted values minus actual values, so that
+# x < 0 means the truth was 1 and 1 + x was the prediction, and
+# x > 0 means the truth was 0 and x was the prediction.
+auc_aggr <- function(x){
+  pred_pos <- sort(1 + x[x < 0])
+  pred_neg <- sort(x[x > 0])
+  
+  n_pos = length(pred_pos)
+  n_neg = length(pred_neg)
+  
+  if(n_pos == 0 | n_neg == 0){NA} else{
+    # Same as sum(outer(pred_neg,pred_pos, "<")).
+    # Slower, but takes less memory.
+    sapply(pred_pos,FUN = function(p){max(which(p >= pred_neg))} ) %>%
+      pmax(.,0) %>% sum() / (n_pos*n_neg)
+  }
+}
+
+roc_aggr <- function(x,alpha){
[…]

### 2021-07-17 forecast/code/hotspots/evaluate_hotspots.R

@@ -1,0 +1,80 @@
+
+## Evaluate hotspots
+library(dplyr)
+library(purrr)
+library(tibble)
+source("hotspot_detection.R")
+source("misc.R")
+source("utils.R")
+detect_hotspots = detect_upswings_basic
+
+# Compute ``ground truth'' hotspots, i.e. whether or not a hotspot
+# actually occurred. This needs to be done independently, because the
+# ground truth hotspots are not available as a covidcast signal.
+forecast_dates <- seq(as.Date("2020-06-16"), as.Date("2021-03-31"), by = "day")
+ahead <- 7:21
+AR_models = c(
+  'AR3',
+  'AR3FB3',
+  'AR3DVCLI3',
+  'AR3CHCLI3',
+  'AR3CHCOV3',
+  'AR3GG3',
+  'AR3GG3_imputed')
+
+
+
+target_dates_df <- tidyr::expand_grid(forecast_date = forecast_dates, ahead) %>%
+  mutate(target_date = forecast_date + ahead)
+
+response_df <- readRDS(here::here("data", "actuals.RDS"))
+hotspots_df <- detect_hotspots(response_df) %>%
+  filter(time_value %in% unique(target_dates_df$target_date)) %>%
+  left_join(target_dates_df, by = c("time_value" = "target_date")) %>%
+  select(geo_value,value,ahead,forecast_date,data_source,signal) %>%
+  rename(actual = value)
+
+
+# Ideally, the next step is
+#
+# evals <- evaluate_predictions(
+# predictions_cards = bind_rows(preds),
+# err_measures = list(err =  function(quantile,value,actual) value - actual),
+# side_truth = hotspots_df,
+# backfill_buffer = 0
+# )
+#
+# However,
+#   (1) that's way too cumbersome, and in fact it errors out.
+#   (2) the way evalcast is currently set up, you cannot supply side truth
+#       AND get back both preds and actuals columns.
+# So I do the following instead.
+#
+train_types = c('honest', 'dishonest', 'honest_bootstrapped')
+
+for (tt in train_types) {
+  # Read baseline
+  train_type <- stringr::str_extract(tt, "[^_]*")
+  # Ingest AR predictions
+  preds_list = vector('list', length(AR_models)+1)
+
+  for (idx in 1:length(AR_models)) {
+    preds_list[[idx]] = readRDS(
+      here::here("data","predictions", "hotspots",
[…]

### 2021-07-17 forecast/code/hotspots/hotspot_signals.R

@@ -1,0 +1,138 @@
+#------------------------------------------------------------------------------#
+# Experiments to evaluate the quality of DELPHI's various indicators,
+# by fitting auto-regressive models to predict hotspots.
+#
+# Hotspots are defined according to a hotspot detection function.
+# Two options:
+#  -- hotspot by simple thresholding,
+#  -- hotspot by test for statistical significance.
+#
+# Several different models are then fit.
+# -- The most basic is an AR3 model.
+# -- All others are AR3 + 3 lagged values of a particular signal.
+#
+# Fitting is done using logistic regression.
+#
+# Where possible I follow Ryan's quantgen-forecast vignette, which can be found
+# in the modeltools package.
+#
+# TODO:
+#  (1) Integrate logisticlasso_forecaster, hotspot_detector in modeltools pkg.
+#------------------------------------------------------------------------------#
+
+# We use the following versions of DELPHI packages.
+# covidcast from CRAN
+# devtools::install_github("cmu-delphi/covidcast", ref = "evalcast",
+#                          subdir = "R-packages/evalcast")
+# devtools::install_local(c("../covidcast/R-packages/modeltools"))
+
+## Source
+library(assertthat)
+library(covidcast)
+library(dplyr)
+library(evalcast)
+library(glmnet)
+library(purrr)
+library(tibble)
+library(modeltools)
+source("auc.R")
+source("alden_evalcast.R") # Used to get predictions from already downloaded data
+source("cv_logisticlasso.R")
+source("hotspot_detection.R")
+source("logisticlasso.R")
+source("logisticlasso_hotspot.R")
+source("misc.R")
+source("utils.R")
+
+# What type of analysis are we doing
+train_types <- c("honest", "dishonest", "honest_bootstrapped")
+
+# What are we forecasting?
+response_source <- "jhu-csse"
+response_signal <- "confirmed_7dav_incidence_num"
+incidence_period <- "day"
+geo_type <- "hrr"
+forecast_dates <- seq(as.Date("2020-06-16"), as.Date("2021-03-31"), by = "day")
+
+# How do we detect a hotspot?
+hotspot_lag = 7            # Number of days in the past to compare to.
+upswing_threshold = .25     # Threshold for null hypothesis.
+min_threshold = 30
+detect_hotspots = detect_upswings_basic
[…]

### 2021-07-17 forecast/code/hotspots/logisticlasso.R

@@ -1,0 +1,32 @@
+#------------------------------------------------------------------------------#
+# A wrapper around glmnet::glmnet.
+#
+# Parameters
+# -- x Matrix of predictors.
+# -- y Vector of responses. 
+#
+# TODO:
+# -- (1) Fix this to match cross validated logistic lasso returned object.
+#------------------------------------------------------------------------------#
+lgstlasso <- function(x, y, lambda, penalty.factor = rep(1,ncol(x)), weights = NULL, intercept = TRUE,
+                       standardize = TRUE,verbose = FALSE){
+  
+  # Setup up x,y, weights
+  a = setup_xy(x,y, weights)
+  x = a$x
+  y = a$y
+  weights = a$weights
+  
+  # Solve the logistic lasso problem
+  glmnet.control(fdev = 0,devmax = 1) # Solve exactly the problem we're asking for.
+  train_obj <- glmnet(x = x, y = y, family = "binomial", lambda = lambda,
+                         penalty.factor = penalty.factor,weights = weights, 
+                         standardize = standardize, intercept = intercept)
+  return(train_obj)
+}
+
+predict.lgstlasso <- function(object,newx,s = NULL,type = "response"){
+  # TODO: figure out why using predict::predict.glmnet here
+  #       gives predictions on the linear scale.
+  predict(object,newx,s,type)
+}

### 2021-07-17 forecast/code/hotspots/logisticlasso_hotspot.R

@@ -1,0 +1,296 @@
+#------------------------------------------------------------------------------#
+# A simple autoregressive hotspot forecaster using logistic lasso.
+#
+# Parameters
+# -- df Data frame of signal values to use for forecasting, of the format
+#       returned by [covidcast::covidcast_signals()].
+# -- forecast_date
+# -- signals. Tibble with columns `data_source` and `signal` that specifies
+#     which variables are being fetched from the COVIDcast API. The first
+#     row of signal is taken to be the response.
+#     Optional column of `start_day`, which can be a Date or a string in 
+#     YYYY-MM-DD form, indicating the earliest date from which data should be
+#     taken. Importantly, `start_day` can also be a function (represented as a list
+#     column) that takes a forecast date and returns a start date for model
+#     training (again, Date object or string in the form "YYYY-MM-DD").
+# -- incidence_period
+# -- ahead 
+# -- n Size of the training window.
+# -- lags
+# -- hotspot_detector. (???)
+# -- cv_type One of "forward" or "random", indicating the type of
+#     cross-validation to perform.
+# -- model_dir. If not NULL, then a relative path to a directory where fitted 
+#               models should be saved.
+# -- verbose
+# -- ... Additional arguments. lambda, 
+#
+# TODO
+#   (1) Compare with `quantgen_forecaster()` to make sure nothing has been
+#       unintentionally changed.
+#------------------------------------------------------------------------------#
+
+lgstlasso_hotspot_forecaster <- function(df_list, forecast_date,
+                                         ahead,
+                                         hotspot_detector,
+                                         n = 28,
+                                         lags = 0,
+                                         feature_type = c("both", "pct_difference","raw"),
+                                         cv_type = c("forward","random"),
+                                         zero_impute= NULL,
+                                         resample=NULL,
+                                         verbose = F, model_dir = NULL,
+                                         ...){
+  
+  # TODO: Check lags vector or list.
+  if (!is.list(lags)) lags = rep(list(lags), length(df_list))
+  
[…]

### 2021-07-17 forecast/code/hotspots/offline_bootstrap_logistic_regression_hotspot_prediction_pipeline.R

@@ -1,0 +1,207 @@
+#------------------------------------------------------------------------------#
+# Experiments to evaluate the quality of DELPHI's various indicators,
+# by fitting auto-regressive models to predict hotspots.
+#
+# Hotspots are defined according to a hotspot detection function.
+# Two options:
+#  -- hotspot by simple thresholding,
+#  -- hotspot by test for statistical significance.
+#
+# Several different models are then fit.
+# -- The most basic is an AR3 model.
+# -- All others are AR3 + 3 lagged values of a particular signal.
+#
+# Fitting is done using logistic regression.
+#
+# Where possible I follow Ryan's quantgen-forecast vignette, which can be found
+# in the modeltools package.
+#
+# TODO:
+#  (1) Integrate logisticlasso_forecaster, hotspot_detector in modeltools pkg.
+#------------------------------------------------------------------------------#
+
+# We use the following versions of DELPHI packages.
+# covidcast from CRAN
+# devtools::install_github("cmu-delphi/covidcast", ref = "evalcast",
+#                          subdir = "R-packages/evalcast")
+# devtools::install_local(c("../covidcast/R-packages/modeltools"))
+
+## Source
+library(assertthat)
+library(covidcast)
+library(dplyr)
+library(evalcast)
+library(glmnet)
+library(purrr)
+library(tibble)
+library(modeltools)
+source("auc.R")
+source("alden_evalcast.R") # Used to get predictions from already downloaded data
+source("cv_logisticlasso.R")
+source("hotspot_detection.R")
+source("logisticlasso.R")
+source("logisticlasso_hotspot.R")
+source("misc.R")
+source("utils.R")
+
+## Setup
+save_model <- FALSE # Should we save trained model to file?
+
+# What type of analysis are we doing
+analysis_type <- "logistic-regression-basic-hotspot_dishonest_as_of"
+
+offline_signal_dir <- "data/offline_signals/honest_as_of"
+
+
+
+
+
+# Load data, if necessary.
+if(load_preds)
+{
+  analysis_dir <- file.path("data",geo_type,analysis_type)
+  load(file.path(analysis_dir, paste(analysis_type,"prediction.rda",sep = "-")))
+}
+
+# What are we forecasting?
+response_source <- "jhu-csse"
+response_signal <- "confirmed_7dav_incidence_num"
[…]

### 2021-07-17 forecast/code/qr_forecast/evaluate_predictions.R

@@ -1,0 +1,71 @@
+library(tibble)
+library(dplyr)
+library(evalcast)
+library(covidcast)
+
+actuals = readRDS(here::here("data", "actuals.RDS"))
+
+tau = c(0.025, 0.1, 0.25, 0.5, 0.75, 0.9, 0.975)
+
+AR_models = c(
+      'AR3',
+      'AR3FBCLI3',
+      'AR3DVCLI3',
+      'AR3CHCLI3',
+      'AR3CHCOV3',
+      'AR3GSSAA3_Zero',
+      'AR3GSSAA3_Subset')
+
+
+# Set train type
+train_types = c('honest', 'dishonest', 'honest_bootstrapped')
+
+for (tt in train_types) {
+  # Read baseline
+  train_type <- stringr::str_extract(tt, "[^_]*")
+  baseline_preds = readRDS(
+    here::here("data", "predictions", "quantreg", sprintf('Baseline_%s.RDS', train_type))
+  )
+
+  # Subset baseline quantiles (properly!)
+  baseline_preds_subset = baseline_preds %>%
+    filter (as.character(quantile*1000) %in% as.character(tau*1000))
+
+
+  # Ingest AR predictions
+  preds_list = vector('list', length(AR_models)+1)
+
+  for (idx in 1:length(AR_models)) {
+    preds_list[[idx]] = readRDS(
+      here::here("data","predictions", "quantreg",
+                 sprintf('%s_%s.RDS', AR_models[[idx]], tt))
+    )
+  }
+
+  preds_list[[length(preds_list)]] = baseline_preds_subset
+
+
+
+  # Only NAs for the subset training, as expected.  Hence
+  # we filter out NAs for that one and proceed.
+
+  preds_list[[7]] = preds_list[[7]] %>% filter(!is.na(value))
+  preds = bind_rows(preds_list)
+
+  rm(preds_list)
+
+  if (tt == "honest") {
+    saveRDS(preds, here::here("data", sprintf('predictions_%s.RDS', tt)))
+  }
+
+  results = evaluate_predictions(
+    preds,
+    actuals,
+    err_measures = list(ae=absolute_error,
+                        wis=weighted_interval_score),
+    grp_vars = c("forecaster", "forecast_date", "ahead", "geo_value"))
+  results <- results %>% select(-data_source, -signal, -incidence_period)
+
+  saveRDS(results, here::here("data", sprintf('results_%s.RDS', tt)))
+}
+

### 2021-07-17 forecast/code/qr_forecast/quantgen.R

@@ -1,0 +1,241 @@
+library(quantgen)
+library(tidyr)
+
+#' Helper functions
+
+make_newx = function(df_features) {
+    # From Alden; actually is doing "latest" as well.
+    # I add a suppressWarnings() here beacuse we know that max(time_value[!is.na(.x)])
+    # will have an empty argument for Google Symptoms and I want to separate these
+    # out from unexpected warnings.
+    suppressWarnings(newx_info <- df_features %>%
+      group_by(geo_value) %>%
+      summarise(across(tidyselect::matches("^value(\\+0|-)"),
+                       ~ max(time_value[!is.na(.x)])), .groups = "drop_last") %>%
+      tidyr::pivot_longer(-geo_value,names_to = "feature_name", values_to = "time_value"))
+    # Test features
+    df_features_test = left_join(newx_info,
+                                 df_features %>%
+                                   filter(time_value %in% unique(newx_info$time_value)) %>%
+                                   tidyr::pivot_longer(-c(geo_value,time_value),
+                                   names_to = "feature_name", values_to = "value"),
+                                 by = c("feature_name","geo_value","time_value")
+    ) %>% select(-time_value) %>%
+      tidyr::pivot_wider(names_from = "feature_name",values_from = "value")
+    test_geo_value = df_features_test %>% pull(geo_value)
+    newx = df_features_test %>% select(-geo_value) %>% as.matrix()
+    # Note: this will be deferred to post hoc analysis.
+    #newx_max_latency = (lubridate::ymd(forecast_date)
+    #                - min(lubridate::ymd(newx_info$time_value), na.rm=TRUE))
+    return(list(newx=newx,
+                test_geo_value=test_geo_value,
+                newx_info=newx_info))
+}
+
+make_newx_latest_data = function(df_features) {
+    df_feat = tibble(expand.grid(geo_value=unique(df_features$geo_value),
+                                 time_value=unique(df_features$time_value)))
+    df_feat$geo_value = as.character(df_feat$geo_value)
+    df_feat = left_join(df_feat, df_features, by=c('geo_value', 'time_value'))
+    df_feat = df_feat %>% arrange (
+        time_value,
+      ) %>% group_by (
+        geo_value,
+      ) %>% group_modify (
+        ~zoo::na.locf(.x, na.rm=FALSE)
+      ) %>% ungroup
+    test_geo_value = df_feat %>%
[…]

### 2021-07-17 forecast/code/qr_forecast/train_baseline.R

@@ -1,0 +1,62 @@
+library(furrr)
+future::plan(multisession)
+
+library(progressr)
+
+library(dplyr)
+library(tibble)
+library(covidcast)
+library(evalcast)
+
+
+geo_type <- "hrr"
+ntrain = 21
+lags = c(0, 7, 14)
+forecast_dates <- seq(as.Date('2020-06-09'),
+                      as.Date('2021-03-31'),
+                      by = "day")
+ahead = 7:21
+response_data_source = 'jhu-csse'
+response_signal = 'confirmed_7dav_incidence_prop'
+
+# To user with future_map, we must evaluate these globals ahead of time
+make_start_day_baseline = function(ntrain) {
+  offset = eval(1 - ntrain - 4)
+  start_day_baseline = function(forecast_date) {
+    return(as.Date(forecast_date) + offset)
+  }
+  return(start_day_baseline)
+}
+
+start_day_baseline = make_start_day_baseline(ntrain)
+
+signals_baseline = tibble::tibble(
+                      data_source = response_data_source,
+                      signal = response_signal,
+                      start_day = list(start_day_baseline),
+                      geo_values='*',
+                      geo_type='hrr') # also ignored because reading from disk.
+
+preds_dir <- here::here("data", "predictions", "quantreg")
+if (!dir.exists(preds_dir)) dir.create(preds_dir)
+
+
+for (train_type in c('honest', 'dishonest')) {
+  offline_signal_dir = here::here(
+    "data", "offline_signals", sprintf('%s_as_of', train_type))
+  t0 = Sys.time()
+  with_progress(preds <- get_predictions(baseline_forecaster,
+                                         'Baseline',
+                                         signals_baseline,
+                                         forecast_dates,
+                                         incidence_period='day',
+                                         offline_signal_dir=offline_signal_dir,
+                                         forecaster_args=list(
+                                           incidence_period='day',
+                                           ahead=ahead)
+  ))
+  t1 = Sys.time()
+  print(t1-t0)
+  saveRDS(preds,
+          here::here(preds_dir, sprintf('%s_%s.RDS', 'Baseline', train_type)))
+}

### 2021-07-17 forecast/code/qr_forecast/train_signals.R

@@ -1,0 +1,103 @@
+library(furrr)
+future::plan(multisession)
+
+library(progressr)
+
+library(dplyr)
+library(tibble)
+library(covidcast)
+library(evalcast)
+source(here::here("qr_forecast", "quantgen.R"))
+
+
+geo_type <- "hrr"
+ntrain = 21
+lags = c(0, 7, 14)
+forecast_dates <- seq(as.Date('2020-06-09'),
+                      as.Date('2021-03-31'),
+                      by = "day")
+ahead = 7:21
+tau = c(0.025, 0.1, 0.25, 0.5, 0.75, 0.9, 0.975)
+response_data_source = 'jhu-csse'
+response_signal = 'confirmed_7dav_incidence_prop'
+signals_df_1 = tribble(
+  ~data_source,         ~signal,
+  response_data_source, response_signal,
+  'fb-survey',          'smoothed_hh_cmnty_cli',
+  'doctor-visits',      'smoothed_adj_cli',
+  'chng',               'smoothed_adj_outpatient_cli',
+  'chng',               'smoothed_adj_outpatient_covid',
+  'google-symptoms',    'sum_anosmia_ageusia_smoothed_search',
+  'google-symptoms',    'sum_anosmia_ageusia_smoothed_search',
+  )
+signals_df_2 = tribble(
+  ~geo_values,  ~name,              ~zero_impute,
+  '*',          'AR3',              NULL,
+  '*',          'AR3FBCLI3',        NULL,
+  '*',          'AR3DVCLI3',        NULL,
+  '*',          'AR3CHCLI3',        NULL,
+  '*',          'AR3CHCOV3',        NULL,
+  '*',          'AR3GSSAA3_Zero',   'anosmia',
+  '*',          'AR3GSSAA3_Subset', NULL,
+  )
+signals_df = bind_cols(signals_df_1, signals_df_2)
+
+# To user with future_map, we must evaluate these globals ahead of time
+make_start_day_ar = function(ahead, ntrain, lags) {
+  offset = eval(1 - max(ahead) - ntrain - max(lags))
+  start_day_ar = function(forecast_date) {
+    return(as.Date(forecast_date) + offset)
+  }
+  return(start_day_ar)
+}
+
+start_day_ar = make_start_day_ar(ahead, ntrain, lags)
+preds_dir <- here::here("data", "predictions", "quantreg")
+if (!dir.exists(preds_dir)) dir.create(preds_dir)
+
[…]

### 2021-07-19 forecast/code/download_signals/download_signals.R
code refactor complete

@@ -68,6 +68,7 @@ start_day_download <- function(forecast_date) {
 
 signals_df = honest_as_of_signals
 offline_signal_dir = here("data", "offline_signals", "honest_as_of")
+if (!dir.exists(offline_signal_dir)) dir.create(offline_signal_dir)
 for (idx in 1:nrow(honest_as_of_signals)) {
   forecast_dates <- seq(as.Date(signals_df$start_forecast_date[idx]),
                         as.Date(end_forecast_date), by = "day")
@@ -95,6 +96,7 @@ for (idx in 1:nrow(honest_as_of_signals)) {
 
 signals_df = dishonest_as_of_signals
 offline_signal_dir = here("data", "offline_signals", "dishonest_as_of")
+if (!dir.exists(offline_signal_dir)) dir.create(offline_signal_dir)
 for (idx in 1:nrow(dishonest_as_of_signals)) {
     forecast_dates <- seq(as.Date(signals_df$start_forecast_date[idx]),
                           as.Date(end_forecast_date), by = "day")

### 2021-07-19 forecast/code/hotspots/evaluate_hotspots.R

@@ -1,18 +1,16 @@
 
 ## Evaluate hotspots
-library(dplyr)
-library(purrr)
-library(tibble)
-source("hotspot_detection.R")
-source("misc.R")
-source("utils.R")
+source(here::here("code", "pkgs_and_common.R"))
+source(here::here("code", "hotspots", "hotspot_detection.R"))
+source(here::here("code", "hotspots", "misc.R"))
+source(here::here("code", "hotspots", "utils.R"))
 detect_hotspots = detect_upswings_basic
 
 # Compute ``ground truth'' hotspots, i.e. whether or not a hotspot
 # actually occurred. This needs to be done independently, because the
 # ground truth hotspots are not available as a covidcast signal.
-forecast_dates <- seq(as.Date("2020-06-16"), as.Date("2021-03-31"), by = "day")
-ahead <- 7:21
+forecast_dates <- hotspot_forecast_dates
+
 AR_models = c(
   'AR3',
   'AR3FB3',
@@ -50,7 +48,6 @@ hotspots_df <- detect_hotspots(response_df) %>%
 #       AND get back both preds and actuals columns.
 # So I do the following instead.
 #
-train_types = c('honest', 'dishonest', 'honest_bootstrapped')
 
 for (tt in train_types) {
   # Read baseline

### 2021-07-19 forecast/code/hotspots/hotspot_signals.R

@@ -27,71 +27,37 @@
 # devtools::install_local(c("../covidcast/R-packages/modeltools"))
 
 ## Source
-library(assertthat)
-library(covidcast)
-library(dplyr)
-library(evalcast)
-library(glmnet)
-library(purrr)
-library(tibble)
-library(modeltools)
-source("auc.R")
-source("alden_evalcast.R") # Used to get predictions from already downloaded data
-source("cv_logisticlasso.R")
-source("hotspot_detection.R")
-source("logisticlasso.R")
-source("logisticlasso_hotspot.R")
-source("misc.R")
-source("utils.R")
-
-# What type of analysis are we doing
-train_types <- c("honest", "dishonest", "honest_bootstrapped")
+source(here::here("code", "pkgs_and_common.R"))
+source(here::here("code", "hotspots", "logisticlasso_hotspot.R"))
+source(here::here("code", "hotspots", "logisticlasso.R"))
+source(here::here("code", "hotspots", "misc.R"))
+source(here::here("code", "hotspots", "utils.R"))
+source(here::here("code", "hotspots", "hotspot_detection.R"))
 
 # What are we forecasting?
-response_source <- "jhu-csse"
-response_signal <- "confirmed_7dav_incidence_num"
-incidence_period <- "day"
-geo_type <- "hrr"
-forecast_dates <- seq(as.Date("2020-06-16"), as.Date("2021-03-31"), by = "day")
+forecast_dates <- hotspot_forecast_dates
 
 # How do we detect a hotspot?
 hotspot_lag = 7            # Number of days in the past to compare to.
 upswing_threshold = .25     # Threshold for null hypothesis.
 min_threshold = 30
 detect_hotspots = detect_upswings_basic
+
+
 # Arguments for our forecaster to take.
 forecaster_args <- list(
-  ahead = 7:21,
-  n = 21, # training set size
-  lags = c(0, 7, 14), # in days for features
+  ahead = ahead,
+  n = ntrain, # training set size
+  lags = lags, # in days for features
   hotspot_detector = detect_hotspots,
   feature_type = "pct_difference",
   lambda = 0 # no penalty
 )
 
+# different names for hotspots
+signals_df$name <- c("AR3", "AR3FB3", "AR3DV3", "AR3CHCLI3", "AR3CHCOV3",
+                     "AR3GG3_imputed", "AR3GG3")
 
-# What signals to use
-signals_df_1 = tribble(
[…]

### 2021-07-19 forecast/code/hotspots/logisticlasso.R

@@ -3,24 +3,24 @@
 #
 # Parameters
 # -- x Matrix of predictors.
-# -- y Vector of responses. 
+# -- y Vector of responses.
 #
 # TODO:
 # -- (1) Fix this to match cross validated logistic lasso returned object.
 #------------------------------------------------------------------------------#
 lgstlasso <- function(x, y, lambda, penalty.factor = rep(1,ncol(x)), weights = NULL, intercept = TRUE,
                        standardize = TRUE,verbose = FALSE){
-  
+
   # Setup up x,y, weights
   a = setup_xy(x,y, weights)
   x = a$x
   y = a$y
   weights = a$weights
-  
+
   # Solve the logistic lasso problem
   glmnet.control(fdev = 0,devmax = 1) # Solve exactly the problem we're asking for.
   train_obj <- glmnet(x = x, y = y, family = "binomial", lambda = lambda,
-                         penalty.factor = penalty.factor,weights = weights, 
+                         penalty.factor = penalty.factor,weights = weights,
                          standardize = standardize, intercept = intercept)
   return(train_obj)
 }

### 2021-07-19 forecast/code/hotspots/logisticlasso_hotspot.R

@@ -8,22 +8,22 @@
 # -- signals. Tibble with columns `data_source` and `signal` that specifies
 #     which variables are being fetched from the COVIDcast API. The first
 #     row of signal is taken to be the response.
-#     Optional column of `start_day`, which can be a Date or a string in 
+#     Optional column of `start_day`, which can be a Date or a string in
 #     YYYY-MM-DD form, indicating the earliest date from which data should be
 #     taken. Importantly, `start_day` can also be a function (represented as a list
 #     column) that takes a forecast date and returns a start date for model
 #     training (again, Date object or string in the form "YYYY-MM-DD").
 # -- incidence_period
-# -- ahead 
+# -- ahead
 # -- n Size of the training window.
 # -- lags
 # -- hotspot_detector. (???)
 # -- cv_type One of "forward" or "random", indicating the type of
 #     cross-validation to perform.
-# -- model_dir. If not NULL, then a relative path to a directory where fitted 
+# -- model_dir. If not NULL, then a relative path to a directory where fitted
 #               models should be saved.
 # -- verbose
-# -- ... Additional arguments. lambda, 
+# -- ... Additional arguments. lambda,
 #
 # TODO
 #   (1) Compare with `quantgen_forecaster()` to make sure nothing has been
@@ -41,17 +41,17 @@ lgstlasso_hotspot_forecaster <- function(df_list, forecast_date,
                                          resample=NULL,
                                          verbose = F, model_dir = NULL,
                                          ...){
-  
+
   # TODO: Check lags vector or list.
   if (!is.list(lags)) lags = rep(list(lags), length(df_list))
-  
-  # Define dt by flipping the sign of lags, include dt = +ahead as a response 
-  # shift, for each ahead value, for convenience later 
+
+  # Define dt by flipping the sign of lags, include dt = +ahead as a response
+  # shift, for each ahead value, for convenience later
   dt <- lapply(lags,"-")
-  
+
   # Data frame with hotspots, defined in terms of the response.
   df_hotspot <- hotspot_detector(df_list[[1]])
-  
+
   # Data frames with pct changes.
   feature_type = match.arg(feature_type)
   if(feature_type %in% c("both","pct_difference")){
@@ -73,30 +73,30 @@ lgstlasso_hotspot_forecaster <- function(df_list, forecast_date,
       df <- df_pct_change
     }
   }
-  
+
[…]

### 2021-07-19 forecast/code/pkgs_and_common.R

@@ -1,0 +1,139 @@
+
+
+# Library loading ---------------------------------------------------------
+
+
+library(furrr)
+future::plan(multisession)
+
+library(progressr)
+library(purrr)
+library(dplyr)
+library(tibble)
+library(covidcast)
+library(evalcast)
+library(glmnet)
+
+
+# User parameters ---------------------------------------------------------
+
+
+
+geo_type <- "hrr"
+ntrain = 21
+ahead = 7:21
+response_data_source = 'jhu-csse'
+response_signal = 'confirmed_7dav_incidence_prop'
+incidence_period = "day"
+
+lags = c(0, 7, 14)
+qr_forecast_dates <- seq(as.Date('2020-06-09'), as.Date('2021-03-31'), by = "day")
+hotspot_forecast_dates <- seq(as.Date("2020-06-16"), as.Date("2021-03-31"), by = "day")
+
+train_types <- c("honest", "dishonest", "honest_bootstrapped")
+lp_solver = "gurobi"
+
+# Signals -----------------------------------------------------------------
+
+
+
+signals_df_1 = tribble(
+  ~data_source,         ~signal,
+  response_data_source, response_signal,
+  'fb-survey',          'smoothed_hh_cmnty_cli',
+  'doctor-visits',      'smoothed_adj_cli',
+  'chng',               'smoothed_adj_outpatient_cli',
+  'chng',               'smoothed_adj_outpatient_covid',
+  'google-symptoms',    'sum_anosmia_ageusia_smoothed_search',
+  'google-symptoms',    'sum_anosmia_ageusia_smoothed_search',
+)
+signals_df_2 = tribble(
+  ~geo_values,  ~name,              ~zero_impute,
+  '*',          'AR3',              NULL,
+  '*',          'AR3FBCLI3',        NULL,
+  '*',          'AR3DVCLI3',        NULL,
+  '*',          'AR3CHCLI3',        NULL,
+  '*',          'AR3CHCOV3',        NULL,
+  '*',          'AR3GSSAA3_Zero',   'anosmia',
+  '*',          'AR3GSSAA3_Subset', NULL,
+)
+signals_df = bind_cols(signals_df_1, signals_df_2)
+
+
+#--------------------------------------------#
+# "hack" of evalcast, in order to use already
+# downloaded data.
+#--------------------------------------------#
+offline_get_predictions <- function(forecaster, name_of_forecaster, signals,
+                                    forecast_dates,
+                                    incidence_period = c("epiweek", "day"),
[…]

### 2021-07-19 forecast/code/qr_forecast/train_baseline.R

@@ -1,23 +1,8 @@
-library(furrr)
-future::plan(multisession)
-
-library(progressr)
-
-library(dplyr)
-library(tibble)
-library(covidcast)
-library(evalcast)
+source(here::here("code", "pkgs_and_common.R"))
+lags = qr_lags
+forecast_dates <- qr_forecast_dates
 
 
-geo_type <- "hrr"
-ntrain = 21
-lags = c(0, 7, 14)
-forecast_dates <- seq(as.Date('2020-06-09'),
-                      as.Date('2021-03-31'),
-                      by = "day")
-ahead = 7:21
-response_data_source = 'jhu-csse'
-response_signal = 'confirmed_7dav_incidence_prop'
 
 # To user with future_map, we must evaluate these globals ahead of time
 make_start_day_baseline = function(ntrain) {

### 2021-07-19 forecast/code/qr_forecast/train_signals.R

@@ -1,46 +1,11 @@
-library(furrr)
-future::plan(multisession)
-
-library(progressr)
-
-library(dplyr)
-library(tibble)
-library(covidcast)
-library(evalcast)
-source(here::here("qr_forecast", "quantgen.R"))
+source(here::here("code", "pkgs_and_common.R"))
+source(here::here("code", "qr_forecast", "quantgen.R"))
 
 
-geo_type <- "hrr"
-ntrain = 21
-lags = c(0, 7, 14)
-forecast_dates <- seq(as.Date('2020-06-09'),
-                      as.Date('2021-03-31'),
-                      by = "day")
-ahead = 7:21
+lags = qr_lags
+forecast_dates <- qr_forecast_dates
 tau = c(0.025, 0.1, 0.25, 0.5, 0.75, 0.9, 0.975)
-response_data_source = 'jhu-csse'
-response_signal = 'confirmed_7dav_incidence_prop'
-signals_df_1 = tribble(
-  ~data_source,         ~signal,
-  response_data_source, response_signal,
-  'fb-survey',          'smoothed_hh_cmnty_cli',
-  'doctor-visits',      'smoothed_adj_cli',
-  'chng',               'smoothed_adj_outpatient_cli',
-  'chng',               'smoothed_adj_outpatient_covid',
-  'google-symptoms',    'sum_anosmia_ageusia_smoothed_search',
-  'google-symptoms',    'sum_anosmia_ageusia_smoothed_search',
-  )
-signals_df_2 = tribble(
-  ~geo_values,  ~name,              ~zero_impute,
-  '*',          'AR3',              NULL,
-  '*',          'AR3FBCLI3',        NULL,
-  '*',          'AR3DVCLI3',        NULL,
-  '*',          'AR3CHCLI3',        NULL,
-  '*',          'AR3CHCOV3',        NULL,
-  '*',          'AR3GSSAA3_Zero',   'anosmia',
-  '*',          'AR3GSSAA3_Subset', NULL,
-  )
-signals_df = bind_cols(signals_df_1, signals_df_2)
+
 
 # To user with future_map, we must evaluate these globals ahead of time
 make_start_day_ar = function(ahead, ntrain, lags) {
@@ -55,49 +20,46 @@ start_day_ar = make_start_day_ar(ahead, ntrain, lags)
 preds_dir <- here::here("data", "predictions", "quantreg")
 if (!dir.exists(preds_dir)) dir.create(preds_dir)
 
[…]

### 2021-07-20 forecast/code/hotspots/evaluate_hotspots.R
fix num/prop issue
forecasting uses the prop signals while hotspots needs the num signals. This means we need both at evaluation time. Be careful here.

@@ -5,7 +5,7 @@ source(here::here("code", "hotspots", "hotspot_detection.R"))
 source(here::here("code", "hotspots", "misc.R"))
 source(here::here("code", "hotspots", "utils.R"))
 detect_hotspots = detect_upswings_basic
-
+source(here::here("code", "hotspots", "hotspot_params.R"))
 # Compute ``ground truth'' hotspots, i.e. whether or not a hotspot
 # actually occurred. This needs to be done independently, because the
 # ground truth hotspots are not available as a covidcast signal.

### 2021-07-20 forecast/code/hotspots/hotspot_signals.R

@@ -38,25 +38,13 @@ source(here::here("code", "hotspots", "hotspot_detection.R"))
 forecast_dates <- hotspot_forecast_dates
 
 # How do we detect a hotspot?
-hotspot_lag = 7            # Number of days in the past to compare to.
-upswing_threshold = .25     # Threshold for null hypothesis.
-min_threshold = 30
-detect_hotspots = detect_upswings_basic
+source(here::here("code", "hotspots", "hotspot_params.R"))
 
-
-# Arguments for our forecaster to take.
-forecaster_args <- list(
-  ahead = ahead,
-  n = ntrain, # training set size
-  lags = lags, # in days for features
-  hotspot_detector = detect_hotspots,
-  feature_type = "pct_difference",
-  lambda = 0 # no penalty
-)
 
 # different names for hotspots
 signals_df$name <- c("AR3", "AR3FB3", "AR3DV3", "AR3CHCLI3", "AR3CHCOV3",
                      "AR3GG3_imputed", "AR3GG3")
+signals_df$signal[1] <- response_signal
 
 
 debug_dir <- here::here("data", "debug_results", "hotspots")

### 2021-07-20 forecast/paper/supplement.Rmd
metadata for the dataset added, commit small datasets
Remains to add links to persistent storage. Remove .Rhistory from tracking.

@@
 
-source(here("code", "figures", "results_processing_and_plotting"))
+source(here("code", "figures", "results_processing_and_plotting.R"))
 source(here("code", "figures", "eval_funs.R"))
@@
 ```{r data-loading}
-actuals <- readRDS(here("data", "actuals.RDS"))
+actuals <- readRDS(here("data", "confirmed_7dav_incidence_prop.RDS"))
 fcasts_honest <- readRDS(here("data", "results_honest.RDS")) %>%
@@
   filter(forecaster != "gs_inherit")
-cases7dav <- readRDS(here("data", "cases7dav.RDS")) %>%
+cases7dav <- readRDS(here("data", "confirmed_7dav_incidence_num.RDS")) %>%
   rename(case_num = value)
@@
   "data", 
-  "jhu-csse_confirmed_7dav_incidence_num_2021-05-18.RDS")) %>%
+  "confirmed_7dav_incidence_num.RDS")) %>%
   pct_change(n = 7) %>%
@@
 
-all_sigs <- readRDS(here("data", "all_signals_wide_as_of_2020-05-15.RDS"))
+all_sigs <- readRDS(here("data", "all_signals_wide_as_of_2020-05-18.RDS"))
 

### 2021-07-20 forecast/code/download_signals/download_signals.R

@@ -11,7 +11,7 @@ ahead = 7:21
 lags = c(0, 7, 14)
 hotspot_lag = 7
 end_forecast_date = '2021-03-31'
-named_function = function(x) lubridate::mdy("2021-05-18") # dishonest as_of
+named_function = function(x) lubridate::ymd("2021-05-18") # finalized as_of
 
 honest_as_of_signals = tribble(
   ~data_source,         ~signal,                              ~start_forecast_date,
@@ -125,4 +125,23 @@ for (idx in 1:nrow(dishonest_as_of_signals)) {
 
 google_sigs <- list.files(offline_signal_dir, "google-symptoms")
 file.copy(google_sigs, here("data", "offline_signals", "honest_as_of"))
+
+
+# Create case / prop actuals ----------------------------------------------
+df <- covidcast_signals(
+  data_source = "jhu-csse",
+  signal = c("confirmed_7dav_incidence_prop", "confirmed_7dav_incidence_num"),
+  start_day = "2020-06-01",
+  end_day = "2021-05-01",
+  as_of = named_function(1)
+)
+
+saveRDS(df[[1]] %>%
+          select(geo_value, time_value, value) %>%
+          rename(target_end_date = time_value, actual = value),
+        here("data", "confirmed_7dav_incidence_prop.RDS"))
+
+saveRDS(df[[1]] %>%
+          select(geo_value, time_value, value),
+        here("data", "confirmed_7dav_incidence_num.RDS"))
 

### 2021-07-20 forecast/code/figures/trajectory.R

@@ -2,8 +2,8 @@ library(covidcast)
 library(evalcast)
 library(tidyverse)
 
-ny_preds <- readRDS("../data/ny_predictions.RDS")
-ny_actuals <- readRDS("../data/ny_actuals.RDS")
+ny_preds <- readRDS(here::here("data", "ny_predictions.RDS"))
+ny_actuals <- readRDS(here::here("data", "ny_actuals.RDS"))
 
 pd <- evalcast:::setup_plot_trajectory(ny_preds, side_truth = ny_actuals)
 pd$truth_df <- pd$truth_df %>% rename(target_end_date = time_value)
@@ -28,4 +28,4 @@ ggplot(pd$truth_df, mapping = aes(x = target_end_date)) +
   ylab("Cases per 100,000 people") + 
   xlab("Date")
 
-ggsave("../paper/fig/trajectory.pdf", width = 6, height = 4)
+ggsave(here::here("paper", "fig", "trajectory.pdf"), width = 6, height = 4)

### 2021-07-20 survey/code/paper-tables.Rmd

---
title: Survey paper tables and figures
author: Alex Reinhart, Alyssa Bilinski, and coauthors
date: July 12, 2021
---

This file contains the code used to generate Tables 2 and 3 and Figures 1, 3, 4,
and 5 of "COVID Trends and Impact Survey in the United States, 2020-2021:
Continuous real-time measurement of COVID-19 symptoms, risks, protective
behaviors, testing and vaccination" by Salomon et al. Figure 2 is generated
separately in `survey-paper-exhibits.ipynb`.

To knit this file: `rmarkdown::render("paper-tables.Rmd")`. This file requires
the packages listed below, plus unified versions of the complete survey
microdata; microdata is only available under [data use
agreement](https://cmu-delphi.github.io/delphi-epidata/symptom-survey/data-access.html).

```{r setup, message=FALSE}
library(data.table)
library(dplyr)
library(tidyr)
library(epitools)
library(tidycensus)

library(ggplot2)
library(gridExtra)
library(scales)
library(geofacet)

library(knitr)
library(lubridate)
library(covidcast)

source("survey-utils.R")
```

## Setup

These categories are in order based on their coding in Q64, so indices
correspond to coding:

```{r more-setup}
occupations <- c(
  "Community and social service",
  "Education, training, and library",
  "Arts, design, entertainment, sports, and media",
  "Healthcare practitioners and technicians",
  "Healthcare support",
  "Protective service",
  "Food preparation and serving related",
  "Building and grounds cleaning and maintenance",
  "Personal care and service",
  "Sales and related",
  "Office and administrative support",
  "Construction and extraction",
  "Installation, maintenance, and repair",
  "Production",
  "Transportation and material moving",
  "Other"
)
```

We load the replication data files. Each object takes quite a lot of RAM, so
removing them and garbage-collecting is essential for this to run in the
available RAM on our server.

```{r load-data}
d1 <- fread("unified-1-year-part-1.csv.gz")
d2 <- fread("unified-1-year-part-2.csv.gz")

data <- rbind(d1, d2, fill = TRUE)

rm(d1, d2)
gc()
```

## Metadata

This is based on data in the following time range:

```{r date-range}
range(data$StartDatetime)
```

Each wave has received this many responses:

```{r wave-count}
data %>%
  select(wave) %>%
  group_by(wave) %>%
  summarize(n = n()) %>%
  arrange(wave) %>%
  kable(col.names = c("Wave", "N"))
```

In total, that is:

```{r sample-size}
nrow(data)
```

Design effects:

```{r deff}
deff <- data %>%
  select(StartDatetime, weight) %>%
  mutate(StartDatetime = as.Date(StartDatetime),
         week = floor_date(StartDatetime, unit = "week"),
         epiweek = epiweek(StartDatetime),
         epiyear = epiyear(StartDatetime)) %>%
  group_by(week, epiyear, epiweek) %>%
  summarize(deff = mean(weight**2) / (mean(weight)**2),
            n = n(),
            .groups = "keep") %>%
  arrange(week) %>%
  select(week, epiyear, epiweek, deff, n)

write_csv(deff, "deff.csv")

ggplot(deff, aes(x = week, y = deff)) +
  geom_line() +
  labs(x = "", y = "Design effect") +
  scale_x_date(date_breaks = "2 months", date_labels = "%b %Y",
               date_minor_breaks = "1 month") +
  scale_y_continuous(limits = c(0, NA)) +
  theme_bw()
```

## Respondent demographics

We'll produce unweighted and weighted proportions for each demographic variable.

```{r demo-setup}
proportionate <- function(data, var) {
  var <- ensym(var)

  data %>%
    filter(!is.na(!!var)) %>%
    group_by(!!var) %>%
    summarize(n = n(), sum_weight = sum(weight)) %>%
    mutate(unweighted_prop = n / sum(n) * 100,
           weighted_prop = sum_weight / sum(sum_weight) * 100) %>%
    select(!!var, n, unweighted_prop, weighted_prop)
}
```

### Gender

```{r demo-gender}
data %>%
  select(D1, weight) %>%
  mutate(gender = case_when(
    D1 == "1" ~ "Male",
    D1 == "2" ~ "Female",
    D1 == "3" | D1 == "4" ~ "Non-binary/self-describe",
    D1 == "5" | is.na(D1) ~ NA_character_
  )) %>%
  proportionate(gender) %>%
  kable(col.names = c("Gender", "Number", "Unweighted %", "Weighted %"),
        digits = 1, format.args = list(big.mark = ","))
```

### Age

```{r demo-age}
gc()

data %>%
  select(D2, weight) %>%
  mutate(age = case_when(
    D2 == "1" ~ "18-24",
    D2 == "2" ~ "25-34",
    D2 == "3" ~ "35-44",
    D2 == "4" ~ "45-54",
    D2 == "5" ~ "55-64",
    D2 == "6" ~ "65-74",
    D2 == "7" ~ "75 plus",
    is.na(D2) ~ NA_character_
  )) %>%
  proportionate(age) %>%
  kable(col.names = c("Age", "Number", "Unweighted %", "Weighted %"),
        digits = 1, format.args = list(big.mark = ","))
```

### Region

```{r demo-region}
data %>%
  select(fips, weight) %>%
  mutate(
    fips = ifelse(is.na(fips), NA_character_, sprintf("%05d", fips)),
    state_abbr = toupper(name_to_abbr(state_fips_to_name(fips)))) %>%
  inner_join(select(state_census, REGION, ABBR),
             by = c("state_abbr" =  "ABBR")) %>%
  mutate(REGION = case_when(
    REGION == 1 ~ "Northeast",
    REGION == 2 ~ "Midwest",
    REGION == 3 ~ "South",
    REGION == 4 ~ "West"
  )) %>%
  proportionate(REGION) %>%
  kable(col.names = c("Region", "Number", "Unweighted %", "Weighted %"),
        digits = 1, format.args = list(big.mark = ","))
```

### Household size

```{r demo-hh-size}
gc()

data %>%
  select(A2b, A5_1, A5_2, A5_3, wave, weight) %>%
  mutate(
    A2b = as.integer(A2b),
    A5_1 = as.integer(A5_1),
    A5_2 = as.integer(A5_2),
    A5_3 = as.integer(A5_3),
    hh_size = case_when(
      wave < 4 ~ A2b,
      wave >= 4 ~ ifelse(is.na(A5_1) + is.na(A5_2) + is.na(A5_3) < 3,
                         ifelse(is.na(A5_1), 0L, A5_1) +
                           ifelse(is.na(A5_2), 0L, A5_2) +
                           ifelse(is.na(A5_3), 0L, A5_3),
                         NA_integer_)
    ),
    hh_size_bucket = case_when(
      hh_size == 1 ~ "1",
      hh_size == 2 ~ "2",
      hh_size >= 3 & hh_size <= 5 ~ "3-5",
      hh_size >= 6 & hh_size <= 10 ~ "6-10",
      hh_size > 10 ~ ">10"
      )) %>%
  proportionate(hh_size_bucket) %>%
  kable(col.names = c("Household size", "Number", "Unweighted %", "Weighted %"),
        digits = 1, format.args = list(big.mark = ","))
```

### Date of completion

```{r demo-completion-date}
gc()

data %>%
  select(StartDatetime, weight) %>%
  mutate(date = as.Date(StartDatetime),
         quarter = paste0(year(date), "Q", quarter(date))) %>%
  proportionate(quarter) %>%
  kable(col.names = c("Quarter", "Number", "Unweighted %", "Weighted %"),
        digits = 1, format.args = list(big.mark = ","))
```

### Education

ACS educational attainment data is for adults age 25 and older.

```{r demo-education}
data %>%
  filter(D2 >= 2) %>%
  select(D8, weight) %>%
  mutate(education = case_when(
    D8 == 1 ~ "Less than high school",
    D8 == 2 ~ "High school or equivalent",
    D8 == 3 ~ "Some college, no degree",
    D8 == 4 ~ "Associate's degree",
    D8 == 5 ~ "Bachelor's degree",
    D8 %in% 6:8 ~ "Graduate or professional degree"
  )) %>%
  proportionate(education) %>%
  kable(col.names = c("Education", "Number", "Unweighted %", "Weighted %"),
        digits = 1, format.args = list(big.mark = ","))
```

This does not match the ACS data (our respondents are more educated), so it may
be valuable to look at this breakdown monthly. If the bias is consistent over
time, at least we know it is not messing up time trends.

```{r demo-education-time}
data %>%
  filter(D2 >= 2) %>%
  select(D8, StartDatetime, weight) %>%
  mutate(date = as.Date(StartDatetime),
         month = paste0(year(date), "M", month(date)),
         education = case_when(
           D8 == 1 ~ "Less than high school",
           D8 == 2 ~ "High school or equivalent",
           D8 == 3 ~ "Some college, no degree",
           D8 == 4 ~ "Associate's degree",
           D8 == 5 ~ "Bachelor's degree",
           D8 %in% 6:8 ~ "Graduate or professional degree"
         )) %>%
  group_by(education, month) %>%
  summarize(sum_weight = sum(weight), .groups = "drop") %>%
  group_by(month) %>%
  mutate(prop = 100 * sum_weight / sum(sum_weight)) %>%
  ungroup() %>%
  select(month, education, prop) %>%
  pivot_wider(names_from = month, values_from = prop)
```

### Comorbidities

The paper only reports on "at least 1 comorbidity".

```{r comorbidities}
data %>%
  filter(!is.na(C1)) %>%
  select(C1, weight) %>%
  mutate(at_least_one = case_when(
    C1 == "9" ~ "None",
    TRUE ~ "At least one comorbidity"
  )) %>%
  proportionate(at_least_one) %>%
  kable(col.names = c("Comorbidities", "Number", "Unweighted %", "Weighted %"),
        digits = 1, format.args = list(big.mark = ","))
```

### Tested given symptoms

## Occupation

Table columns:

- % tested positive: The % of respondents in each occupation category who have
  ever tested positive. P(ever tested positive | occupation group). Note no
  conditioning on testing, so respondents who were never tested count as not
  having tested positive.
- % working with symptoms: In each occupation category, % of respondents who (a)
  have worked outside their home in the past few days but (b) have unusual
  symptoms. That is, P(symptoms & worked outside home | occupation group)
- % working outside/not tested: In each occupation category, % of respondents
  who (a) have worked outside their home in the past 4 weeks but (b) have never
  been tested. (There are some pretty stark disparities in this table). That is,
  P(never tested & worked outside home | occupation group).

Calculated on all responses in January 2020 with the relevant items completed.

**Note:** From here on, all analyses will be for Wave 4 onward (since Wave 4 has
the needed variables).

```{r occupation}
data <- data %>%
  filter(wave >= 4)

data$C13_logic <- split_options(data$C13)

data %>%
  filter(!is.na(Q64),
         StartDatetime >= "2021-01-01 00:00:00",
         StartDatetime < "2021-02-01 00:00:00") %>%
  mutate(
    ever_pos = case_when(
      B10a == "1" ~ TRUE,
      B11 == "1" ~ TRUE,
      B11 == "2" ~ FALSE,
      B8 == "2" ~ FALSE,
      TRUE ~ NA
    ),
    work_sym = case_when(
      is.na(C13) ~ NA, # No info on activities outside home
      is.na(B2) ~ NA, # No info on symptoms, unusual or not
      B2 == "15" ~ FALSE, # No symptoms, unusual or not
      is.na(B2c) ~ FALSE, # No unusual symptoms listed
      !is_selected(C13_logic, "1") ~ FALSE, # Did not work outside home
      is_selected(C13_logic, "1") ~ TRUE, # Worked outside home
      TRUE ~ NA
    ),
    work_not_tested = case_when(
      D10 == "2" ~ FALSE,
      B8 == "1" ~ FALSE,
      D10 == "1" & B8 == "2" ~ TRUE,
      TRUE ~ NA
    ),
    occupation = occupations[Q64]) %>%
  select(occupation, ever_pos, work_sym, work_not_tested, weight) %>%
  group_by(occupation) %>%
  summarize(work_pos_rate = 100 * weighted.mean(ever_pos, weight, na.rm = TRUE),
            work_sym_rate = 100 * weighted.mean(work_sym, weight, na.rm = TRUE),
            work_not_tested_rate = 100 * weighted.mean(work_not_tested, weight, na.rm = TRUE),
            n = n(),
            .groups = "drop_last") %>%
  kable(col.names = c("Occupation group", "% tested positive",
                      "% working with symptoms", "% working outside/not tested",
                      "N"),
        digits = 1, format.args = list(big.mark = ","))
```

## Symptoms

```{r symptom-rates}
data$positive_14d <- case_when(
  data$B10a == 1 ~ "Tested positive",
  TRUE ~ "All others"
)

symptoms <- list(
  "1" = "Fever",
  "2" = "Cough",
  "3" = "Shortness of breath",
  "4" = "Difficulty breathing",
  "5" = "Tiredness or exhaustion",
  "6" = "Nasal congestion",
  "7" = "Runny nose",
  "8" = "Muscle or joint aches",
  "9" = "Sore throat",
  "10" = "Chest pain",
  "11" = "Nausea or vomiting",
  "12" = "Diarrhea",
  "13" = "Loss of smell/taste",
  "16" = "Eye pain",
  "17" = "Chills",
  "18" = "Headaches",
  "19" = "Changes in sleep",
  "14" = "Other"
)

## P(unusual symptoms | tested positive in past 14 days, age group)
symptom_data <- data %>%
  select(positive_14d, B2, B2c, weight)

symptom_data$B2c_logic <- split_options(symptom_data$B2c)

for (ii in names(symptoms)) {
  sym_name <- paste0("sym_", symptoms[[ii]])

  symptom_data[[sym_name]] <- case_when(
    symptom_data$B2 == "15" ~ FALSE, # no symptoms, unusual or not
    is.na(symptom_data$B2) ~ NA, # left symptoms blank
    is.na(symptom_data$B2c) ~ FALSE, # symptom given in B2, no unusual symptoms
    TRUE ~ is_selected(symptom_data$B2c_logic, ii)
  )
}

symptom_rates_overall <- symptom_data %>%
  summarize(across(starts_with("sym_"), ~ weighted.mean(.x, weight, na.rm = TRUE))) %>%
  pivot_longer(cols = starts_with("sym_"), names_to = "symptom",
               names_prefix = "sym_", values_to = "rate")

symptom_rates <- symptom_data %>%
  group_by(positive_14d) %>%
  summarize(across(starts_with("sym_"), ~ weighted.mean(.x, weight, na.rm = TRUE))) %>%
  pivot_longer(cols = starts_with("sym_"), names_to = "symptom",
               names_prefix = "sym_", values_to = "rate")
```

Here are overall symptoms among all respondents:

```{r symptoms-overall-table}
symptom_rates_overall %>%
  mutate(rate = 100 * rate) %>%
  arrange(desc(rate)) %>%
  kable(col.names = c("Symptom", "% with symptom"),
        digits = 1)
```

Here we have unusual symptoms, split by whether the respondent reported testing
positive in the past 14 days. Calculate risk ratios and their CIs.

```{r symptoms-compare}
# Reorder factor levels to order by symptoms among those who tested positive
ordered_symptoms <- symptom_rates %>%
  filter(positive_14d == "Tested positive") %>%
  arrange(desc(rate)) %>%
  pull(symptom)
symptom_rates$symptom <- factor(
  symptom_rates$symptom,
  levels = ordered_symptoms)

# Get relative risks using epitools. We need raw counts, not weighted estimates,
# unfortunately.
symptom_counts <- symptom_data %>%
  group_by(positive_14d) %>%
  summarize(across(starts_with("sym_"), ~ sum(.x, na.rm = TRUE))) %>%
  pivot_longer(cols = starts_with("sym_"), names_to = "symptom",
               names_prefix = "sym_", values_to = "rate")

n_positive <- sum(symptom_data$positive_14d == "Tested positive", na.rm = TRUE)
n_negative <- sum(symptom_data$positive_14d == "All others", na.rm = TRUE)

symptom_rrs <- symptom_counts %>%
  pivot_wider(id_cols = symptom, names_from = positive_14d, values_from = rate)

symptom_rrs$rr <- NA
symptom_rrs$rr_lower <- NA
symptom_rrs$rr_upper <- NA
for (ii in seq_len(nrow(symptom_rrs))) {
  res <- riskratio(c(
    n_negative - symptom_rrs[[ii, "All others"]], # without symptom, negative
    symptom_rrs[[ii, "All others"]], # with symptom, negative
    n_positive - symptom_rrs[[ii, "Tested positive"]], # without symptom, positive
    symptom_rrs[[ii, "Tested positive"]]))$measure

  symptom_rrs$rr[ii] <- res["Exposed2", "estimate"]
  symptom_rrs$rr_lower[ii] <- res["Exposed2", "lower"]
  symptom_rrs$rr_upper[ii] <- res["Exposed2", "upper"]
}
```

Now plot:

```{r symptoms-compare-plot}
rr_scale_factor <- 70

ggplot(symptom_rates, aes(x = symptom, y = rate, fill = positive_14d)) +
  geom_col(position = position_dodge()) +
  geom_pointrange(
    aes(x = symptom, y = rr / rr_scale_factor,
        ymin = rr_lower / rr_scale_factor,
        ymax = rr_upper / rr_scale_factor),
    data = symptom_rrs,
    inherit.aes = FALSE) +
  scale_y_continuous(labels = label_percent(accuracy = 1),
                     sec.axis = sec_axis(~ . * rr_scale_factor,
                                         name = "Relative risk")) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(x = "", y = "Frequency among respondents",
       fill = "Group") +
  theme_bw() +
  theme(legend.position = "bottom")

ggsave("figures/symptoms-comparison.pdf")
```

In table form, so we can quote it:

```{r symptoms-table}
symptom_rates %>%
  mutate(rate = 100 * rate) %>%
  pivot_wider(names_from = "positive_14d", values_from = "rate") %>%
  select(symptom, positive = `Tested positive`, others = `All others`) %>%
  arrange(desc(positive)) %>%
  kable(col.names = c("Symptom", "Tested positive", "All others"),
        digits = 1)
```

And the table of relative risks and their CIs:

```{r rr-table}
symptom_rrs %>%
  select(symptom, rr, rr_lower, rr_upper) %>%
  arrange(desc(rr)) %>%
  kable(col.names = c("Symptom", "RR", "lower", "upper"),
        digits = 2)
```

## Cumulative test positivity

Based on code from Minttu Rönn.

We use 7-day average cumulative incidence (per 100,000 population) to match the
7-day averaging used in the survey estimates. The survey data will have one
estimate per week, rather than per day, and so when we join the two data frames
we'll get one JHU case rate per week.

```{r case-rate-compare}
# State adult populations, from Census data
state_pops <- read.csv("state-age-pops.csv")
# State-level age distributions of cases, from CDC data
case_age_dist <- read.csv("case-age-distribution.csv")

start_day <- floor_date(as.Date(min(data$StartDatetime)), unit = "week")
state_case_rate <- covidcast_signal("jhu-csse", "confirmed_7dav_cumulative_num",
                                    start_day = start_day,
                                    end_day = as.Date(max(data$StartDatetime)),
                                    geo_type = "state") %>%
  mutate(geo_value = toupper(geo_value),
         year = year(time_value),
         month = month(time_value)) %>%
  select(state = geo_value, cumulative_cases = value, time_value, year, month) %>%
  inner_join(state_pops, by = "state") %>%
  inner_join(case_age_dist, by = c("year", "month")) %>%
  mutate(case_rate_adult = cumulative_cases * frac_adult_cases / adult_pop * 100)

state_case_compare <- data %>%
  filter(!is.na(fips)) %>%
  mutate(
    ever_pos = case_when(
      B10a == "1" ~ TRUE,
      B11 == "1" ~ TRUE,
      B11 == "2" ~ FALSE,
      B8 == "2" ~ FALSE,
      TRUE ~ NA
    ),
    week = floor_date(as.Date(StartDatetime), unit = "week"),
    fips = sprintf("%05d", fips),
    state_abbr = name_to_abbr(state_fips_to_name(fips))
  ) %>%
  group_by(week, state_abbr) %>%
  summarize(ever_pos = 100 * weighted.mean(ever_pos, weight, na.rm = TRUE),
            .groups = "drop") %>%
  left_join(state_case_rate,
            by = c("state_abbr" = "state", "week" = "time_value"))

g <- ggplot(state_case_compare) +
  facet_geo(~ state_abbr, grid = "us_state_with_DC_PR_grid1", label = "code") +
  geom_line(aes(x = week, y = case_rate_adult, color = "Surveillance"),
            linetype = 1, size = 0.5) +
  geom_line(aes(x = week, y = ever_pos, color = "CTIS"),
            linetype = 1, size = 0.5) +
  scale_color_manual(values = c("lightblue", "gray50")) +
  scale_y_continuous(breaks = c(0, 10),
                     minor_breaks = c(5, 15),
                     labels = label_percent(scale = 1,
                                            accuracy = 1)) +
  scale_x_date(date_breaks = "4 months", date_labels = "%b",
               date_minor_breaks = "1 month") +
  labs(x = "", y = "", color = "") +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave("figures/cumulative-case-compare.pdf", g, width = 10, height = 7)

g
```

We should also calculate correlations between state reports and survey estimates
during the last week of data:

```{r}
max(state_case_compare$week)

final_week <- state_case_compare %>%
  filter(week == max(week))

# States with the lowest proportion of ever-diagnosed in survey data
final_week %>%
  arrange(ever_pos) %>%
  select(state_abbr, ever_pos, case_rate_adult) %>%
  head()

# States with the highest proportion
final_week %>%
  arrange(desc(ever_pos)) %>%
  select(state_abbr, ever_pos, case_rate_adult) %>%
  head()

# Pearson correlation between surveillance and CTIS
cor(final_week$case_rate_adult, final_week$ever_pos)

# Spearman (rank) correlation between surveillance and CTIS
cor(final_week$case_rate_adult, final_week$ever_pos,
    method = "spearman")
```

## Contacts, masks, and public transport

Based on code from Alyssa Bilinski.

```{r external-data}
make_df <- function(f, var = SVI_cat) {
  var <- ensym(var)

  f %>%
    select(!!var, week, C13_options, C14, contacts_tot, weight) %>%
    filter(!is.na(!!var)) %>%
    group_by(!!var, week) %>%
    summarize(
      pub_trans = 100 * weighted.mean(is_selected(C13_options, "6"),
                                      weight, na.rm = TRUE),
      masks_not_always = 100 * weighted.mean(
        case_when(
          C14 == 1 ~ FALSE,
          C14 %in% 2:5 ~ TRUE,
          C14 == 6 ~ NA), # exclude those not in public in past 5/7 days
        weight, na.rm = TRUE),
      contacts_avg_mod = weighted.mean(contacts_tot, weight, na.rm = TRUE)) %>%
    pivot_longer(c(contacts_avg_mod, pub_trans, masks_not_always),
                 names_to = "var", values_to = "value") %>%
    mutate(var_cat = case_when(
      var == "contacts_avg_mod" ~ "Average contacts (past 24h)",
      var == "pub_trans" ~ "Rode public transit (past 24h, %)",
      var == "masks_not_always" ~ "Not always masked in public (%)"))
}

pops <- read.csv("data/population.csv")

# SVI
svi <- read.csv("data/SVI2018_US_COUNTY.csv") %>%
  mutate(SVI_cat = case_when(RPL_THEMES >= 0.75 ~ "1st quartile",
                             RPL_THEMES >= 0.5 & RPL_THEMES < 0.75 ~ "2nd quartile",
                             RPL_THEMES >= 0.25 & RPL_THEMES < 0.5 ~ "3rd quartile",
                             RPL_THEMES < 0.25 ~ "4th quartile")) %>%
  select(FIPS, SVI_cat)
```

Produce the variables and data:

```{r contacts-data}
contacts_data <- data %>%
  # Exclude ludicrous numbers of contacts, since they screw up the means
  filter(chk_work_contacts & chk_shopping_contacts & chk_social_contacts & chk_other_contacts) %>%
  mutate(
    C13 = ifelse(!is.na(C13), C13, C13b), # Replaced in Wave 10; treat as identical
    C14 = ifelse(!is.na(C14), C14, C14a), # Replaced in Wave 8; treat as identical
    C13_options = split_options(C13),
    state = state_fips_to_name(ifelse(is.na(fips), NA_character_, sprintf("%05d", fips))),
    week = floor_date(as.Date(StartDatetime), unit = "week"),
    any_work = ifelse(is.na(C10_1_1), 0, C10_1_1),
    any_shopping = ifelse(is.na(C10_2_1), 0, C10_2_1),
    any_social = ifelse(is.na(C10_3_1), 0, C10_3_1),
    any_other = ifelse(is.na(C10_4_1), 0, C10_4_1),
    chk = !is.na(C10_1_1) | !is.na(C10_2_1) | !is.na(C10_3_1) | !is.na(C10_4_1),
    work_mod2 = ifelse(chk, any_work, NA),
    social_mod2 = ifelse(chk, any_social, NA),
    shopping_mod2 = ifelse(chk, any_shopping, NA),
    other_mod2 = ifelse(chk, any_other, NA),
    contacts_tot = (as.numeric(work_mod2) + as.numeric(shopping_mod2) +
                      as.numeric(social_mod2) + as.numeric(other_mod2)),
    age_cat = case_when(D2 <= 2 ~ "18-34",
                        D2 %in% 3:5 ~ "35-64",
                        D2 > 5 ~ "65+")
  ) %>%
  left_join(svi, by = c("fips" = "FIPS")) %>%
  left_join(pops, by = c("state" = "StateName"))
```

Produce the plots:

```{r contacts-plots}
g1 <- make_df(contacts_data, var = SVI_cat)
g2 <- make_df(contacts_data, var = age_cat)
g3 <- make_df(contacts_data, var = region)

# TODO See if we can merge these into one big facet_grid so the widths match
# (not sure how to deal with legends?)
a <- ggplot(g1, aes(x = week, y = value, col = SVI_cat, group = SVI_cat)) +
  geom_line() +
  facet_wrap(. ~ var_cat, scales = "free", ncol = 4) +
  scale_color_discrete(name = "SVI") +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
               date_minor_breaks = "1 month") +
  labs(x = "", y = "") +
  ylim(0,  NA) +
  theme_bw()

b <- ggplot(g2, aes(x = week, y = value, col = age_cat, group = age_cat)) +
  geom_line() +
  facet_wrap(. ~ var_cat, scales = "free", ncol = 4) +
  scale_color_discrete(name = "Age") +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
               date_minor_breaks = "1 month") +
  labs(x = "", y = "") +
  ylim(0, NA) +
  theme_bw()

c <- ggplot(g3, aes(x = week, y = value, col = region, group = region)) +
  geom_line() +
  facet_wrap(.~var_cat, scales = "free", ncol = 4) +
  scale_color_discrete(name = "Region")  +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
               date_minor_breaks = "1 month") +
  labs(x = "", y = "") +
  ylim(0, NA) +
  theme_bw()

g <- arrangeGrob(a, b, c, nrow = 3)

ggsave("figures/contacts-masks-transport.pdf", g, width = 8, height = 5)
```

## Vaccine acceptance and hesitancy

Based on code from Marissa Reitsma.

```{r vaccine-hesitancy-map}
# Get data into the form Marissa's code expects
df <- data %>%
  select(StartDatetime, fips, V1, V3, weight) %>%
  mutate(date = as.Date(StartDatetime)) %>%
  filter(!is.na(fips),
         date >= as.Date("2021-03-01"),
         date <= as.Date("2021-04-05")) %>%
  mutate(
    hesitant = case_when(
      V1 == 1 ~ FALSE,
      V3 %in% 1:2 ~ FALSE,
      V3 %in% 3:4 ~ TRUE),
    fips = sprintf("%05d", fips))

# Get all FIPSen
fips <- as.data.table(tidycensus::fips_codes)
fips <- fips[, fips:=paste0(state_code, county_code)]

# Collapse into summaries. This uses megacounty logic to get summaries by state
# if sample size per county is inadequate
df <- merge(df, fips, by = "fips", all.y = TRUE)

# Compute percent vaccinated or accepting
df <- df[, sample_size := .N, by = c("fips")]
df <- df[, geo_id := ifelse(sample_size >= 50, fips, state_code)]

df <- df[, pct_hesitant := weighted.mean(hesitant, weight, na.rm = TRUE), by = c("geo_id")]
df <- df[, pct_hesitant := mean(pct_hesitant, na.rm = TRUE), by = c("geo_id")]

# Collapse the dataset
df <- unique(df[, .(pct_hesitant, geo_id, fips, sample_size)])

# Code the colors
df <- df[pct_hesitant < 0.1, plot_var := "<10%"]
df <- df[pct_hesitant >= 0.1 & pct_hesitant < 0.14, plot_var := "10-14%"]
df <- df[pct_hesitant >= 0.14 & pct_hesitant < 0.18, plot_var := "14-18%"]
df <- df[pct_hesitant >= 0.18 & pct_hesitant < 0.22, plot_var := "18-22%"]
df <- df[pct_hesitant >= 0.22 & pct_hesitant < 0.26, plot_var := "22-26%"]
df <- df[pct_hesitant >= 0.26 & pct_hesitant < 0.3, plot_var := "26-30%"]
df <- df[pct_hesitant >= 0.3, plot_var := "30+%"]

df <- df[, plot_var := factor(plot_var, levels = c("<10%", "10-14%", "14-18%", "18-22%", "22-26%", "26-30%", "30+%"))]

df <- df[, GEOID := fips]

counties <- get_acs(geography = "county", geometry = TRUE,
                    shift_geo = TRUE, variables = "B19013_001")
states <- get_acs(geography = "state", geometry = TRUE,
                  shift_geo = TRUE, variables = "B19013_001")

plot_data <- merge(counties, df, by = "GEOID", all.x = TRUE)

g <- ggplot(data = plot_data, aes(fill = plot_var)) +
  geom_sf(data = plot_data[plot_data$geo_id == plot_data$fips, ],
          size = 0.2, color = "gray") +
  geom_sf(data = plot_data[plot_data$geo_id != plot_data$fips, ],
          size = 0.2, color = "NA") +
  geom_sf(data = states, fill = NA, color = "#444444") +
  scale_fill_manual(
    values = c("#0d2c84", "#225ea8", "#1d91c0", "#41b6c4",
               "#7fcdbb", "#c7e9b4", "#ffffcc", "#dedede"),
    na.value = "#dedede") +
  theme_void() +
  labs(fill = "% hesitant", title = "")

ggsave("figures/hesitancy-county-map.pdf", g, width = 6, height = 4)

g
```

### 2021-07-21 forecast/paper/supplement.Rmd

@@
 ```{r traj-data, eval = FALSE}
-source(here("code", "tbd", "trajectory_plot_funs.R"))
+source(here("code", "deprecated", "trajectory_plot_funs.R"))
 preds <- readRDS(here("data", "predictions_honest.RDS"))

### 2021-08-02 indicators/paper/paper.tex
Re-add Googlers now that approval has been granted

@@
 % Please include corresponding author, author contribution and author declaration information
-\authorcontributions{Author contributions: A.R., L.B., M.J., A.R., J.T., A.C,
-  E.J.C., N.D., D.C.F, A.J.H., S.H., M.J., J.K., W.L.M.K, Z.C.L., L.M., N.L.O.,
-  J.C.W, R.R., and R.J.T. obtained and managed data;
-  A.R., J.B., J.F., Z.C.L., L.M., D.J.M., A.P., S.R., D.R., C.S. N.S., R.T.,
-  V.V., L.W., B.W., J.C.W, S.W., R.R., and R.J.T. designed software and
+% Googlers: SA, SC, DCF, JG, KH, RH, PM, MPO, CS, BYS, BW, SW
+\authorcontributions{Author contributions: A.R., L.B., M.J., A.R., J.T., A.C.,
+  E.J.C., N.D., D.C.F., A.J.H., S.H., M.J., J.K., W.L.M.K., Z.C.L., L.M., N.L.O.,
+  J.C.W., R.R., and R.J.T. obtained and managed data;
+  A.R., J.B., J.F., Z.C.L., L.M., D.J.M., A.P., S.R., D.R., C.S., N.S., R.T.,
+  V.V., L.W., B.W., J.C.W,, S.W., R.R., and R.J.T. designed software and
   research;
@@
   S.H., M.J., A.J., J.K., A.K., M.X.L., L.M., K.M., D.J.M., B.N., M.P.O.,
-  N.L.O., P.P., C.A.P., S.R., C.S., V.S, J.S, D.S., N.S., B.Y.S., V.S., S.T.,
-  J.T., R.T., J.C.W, and R.J.T. developed software;
+  N.L.O., P.P., C.A.P., S.R., C.S., V.S., J.S., D.S., N.S., B.Y.S., V.S., S.T.,
+  J.T., R.T., J.C.W., and R.J.T. developed software;
   S.A., W.A.S., A.C., B.C., D.C.F., S.G., G.H., K.M., D.J.M., B.N., C.S., J.S.,
   D.S., B.Y.S., and A.K.V.N. managed technical infrastructure and systems;
-  A.R., L.B., M.J., A.R., J.T, J.B., A.C., E.J.C., N.D., D.C.F., A.G., K.H.,
-  A.J.H., R.H., S.H., A.J., J.K., K.L., Y.J.L, Z.C.L., L.M., D.J.M., P.M., B.N.,
+  A.R., L.B., M.J., A.R., J.T., J.B., A.C., E.J.C., N.D., D.C.F., A.G., K.H.,
+  A.J.H., R.H., S.H., A.J., J.K., K.L., Y.J.L., Z.C.L., L.M., D.J.M., P.M., B.N.,
   N.L.O., P.P., C.A.P., S.R., N.S., V.S., J.S., D.S., N.S., S.T., J.T., R.T.,

### 2021-08-02 indicators/paper/paper.tex
Disambiguate the two ARs

@@
 % Googlers: SA, SC, DCF, JG, KH, RH, PM, MPO, CS, BYS, BW, SW
-\authorcontributions{Author contributions: A.R., L.B., M.J., A.R., J.T., A.C.,
+\authorcontributions{Author contributions: A.Re., L.B., M.J., A.Ru., J.T., A.C.,
   E.J.C., N.D., D.C.F., A.J.H., S.H., M.J., J.K., W.L.M.K., Z.C.L., L.M., N.L.O.,
   J.C.W., R.R., and R.J.T. obtained and managed data;
-  A.R., J.B., J.F., Z.C.L., L.M., D.J.M., A.P., S.R., D.R., C.S., N.S., R.T.,
+  A.Re., J.B., J.F., Z.C.L., L.M., D.J.M., A.P., S.R., D.R., C.S., N.S., R.T.,
   V.V., L.W., B.W., J.C.W,, S.W., R.R., and R.J.T. designed software and
   research;
-  A.R., L.B., M.J., A.R., J.T., W.A.S., T.A., A.B., J.B., L.B., A.A.C., A.C.,
+  A.Re., L.B., M.J., A.Ru., J.T., W.A.S., T.A., A.B., J.B., L.B., A.A.C., A.C.,
   E.J.C., B.C., S.C., N.D., D.C.F., J.G., S.G., A.G., G.H., R.H., K.H., A.J.H.,
@@
   D.S., B.Y.S., and A.K.V.N. managed technical infrastructure and systems;
-  A.R., L.B., M.J., A.R., J.T., J.B., A.C., E.J.C., N.D., D.C.F., A.G., K.H.,
+  A.Re., L.B., M.J., A.Ru., J.T., J.B., A.C., E.J.C., N.D., D.C.F., A.G., K.H.,
   A.J.H., R.H., S.H., A.J., J.K., K.L., Y.J.L., Z.C.L., L.M., D.J.M., P.M., B.N.,
@@
   R.R. and R.J.T. acquired funding;
-  and A.R., L.B., M.J., A.R., J.T., and R.J.T. wrote the paper.
+  and A.Re., L.B., M.J., A.Ru., J.T., and R.J.T. wrote the paper.
 }
@@
   addressed. E-mail: areinhar@stat.cmu.edu}
-\equalauthors{\textsuperscript{2}L.B., M.J., A.R., and J.T. contributed equally
+\equalauthors{\textsuperscript{2}L.B., M.J., A.Ru., and J.T. contributed equally
   to this work.}

### 2021-09-07 forecast/code/deprecated/revision-draft.R
additional analyses for the revision

@@ -1,0 +1,113 @@
+# Cumulative mean prediction error plot -----------------------------------
+
+tt <- fcasts_honest %>%
+  group_by(forecaster, forecast_date) %>%
+  summarise(w = sum(wis) / sum(strawman_wis)) %>%
+  arrange(forecast_date) %>%
+  mutate(n = n(), w = cummean(w))
+
+
+fcast_colors <- c(RColorBrewer::brewer.pal(5, "Set1"), "#000000")
+names(fcast_colors) <- c("CHNG-CLI", "CHNG-COVID", "CTIS-CLIIC", "DV-CLI",
+                         "Google-AA", "AR")
+
+
+ggplot(tt %>% filter(forecast_date < "2021-01-01"), 
+       aes(forecast_date, w, color = forecaster)) + 
+  geom_line() +
+  geom_hline(yintercept = 1) +
+  xlab("Days ahead") +
+  ylab("cummean(sum(wis_forecaster[t]) /\n sum(wis_strawman[t]))") +
+  scale_color_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
+  theme_bw() +
+  theme(legend.position = "bottom", legend.title = element_blank())
+
+
+# Sign test ---------------------------------------------------------------
+
+fcast_colors2 <- fcast_colors[names(fcast_colors) != "AR"]
+
+st <- fcasts_honest %>%
+  mutate(wis = wis / strawman_wis) %>%
+  select(forecaster, geo_value, ahead, forecast_date, wis) %>%
+  pivot_wider(names_from = forecaster, values_from = wis) %>%
+  mutate(across(AR:`Google-AA`, ~ AR -.x)) %>%
+  select(-AR) %>%
+  pivot_longer(`CHNG-CLI`:`Google-AA`, names_to = "forecaster", values_to = "dif") %>%
+  group_by(forecaster, forecast_date, geo_value) %>%
+  summarise(dif = sum(dif)) %>%
+  group_by(forecaster, forecast_date) %>%
+  summarise(p = binom.test(
+    x = sum(dif > 0),
+    n = n(),
+    alternative = "greater")$p.val)
+
+ggplot(st, aes(p)) +
+  geom_histogram(
+    aes(y = ..count.. / sum(..count..), color = forecaster, fill = forecaster), 
+    alpha = 0.4,
+    bins = 40) +
[…]

### 2021-09-08 forecast/code/deprecated/revision-draft.R
fix xlab

@@ -16,7 +16,7 @@ ggplot(tt %>% filter(forecast_date < "2021-01-01"),
        aes(forecast_date, w, color = forecaster)) + 
   geom_line() +
   geom_hline(yintercept = 1) +
-  xlab("Days ahead") +
+  xlab("forecast date") +
   ylab("cummean(sum(wis_forecaster[t]) /\n sum(wis_strawman[t]))") +
   scale_color_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
   theme_bw() +

### 2021-09-09 indicators/paper/responses-to-reviews.tex
Address Reviewer 1 comment and add response template

\documentclass[11pt]{article}

\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{microtype}

\usepackage{newtxmath}
\usepackage{newtxtext}

\title{Response to reviewers}
\author{}
\date{\today}

\begin{document}
\maketitle

\section*{Editor}

\begin{quote}
  The reviewers both saw high value in the database and regarded the paper as
  important and relevant. Reviewer 2 had several high-level comments about the
  current framing of the results and offered a number of suggestions for the
  authors to consider additional or modified analyses that would strengthen the
  paper. I would like to invite the authors to submit a revised version of the
  manuscript including a detailed point-by-point response to the reviewer
  comments, particularly where Reviewer 2 has suggested need for either further
  analyses or pointed to notable limitations. While the authors should use their
  discretion regarding whether additional analyses would be within scope for the
  revision, it will be important to respond to the suggestions in either case
  and at least to expand on limitations as relevant in the revised manuscript.
\end{quote}

TODO

\section*{Reviewer 1}

\begin{quote}
  In section 1.B, you mention non-trivial signal processing and direct the
  reader to github to learn more. I would like to see more discussion of how
  each source was processed. If this is not possible due to length constraints,
  where can the reader find more details? Is there documentation at the Delphi
  website that explains how the data from each source is processed? In a quick
  search at the Delphi website, I was not able to find information about how the
  survey data (for example) were weighted. These details will be important to
  those using the data.
\end{quote}

Thanks, this is an important point. We have added a sentence to Section 1.B to
make clear that detailed documentation, including such details as the survey
weighting procedures and the day-of-week adjustments used in other data sources,
is available on the public documentation site.

\section*{Reviewer 2}

\begin{quote}
  This paper describes a diverse public database of signals related to COVID-19
  prevalence, including symptom survey data, mobility, at-home tests, and
  hospitalizations. The database is unique, important, and useful, especially in
  the current surveillance-limited context of the United States. However, the
  main analyses presented in the results section did not feel cohesive, and the
  ``story'' of the paper felt incomplete. While I respect the challenge of
  dividing work among multiple related papers, I think that a key goal of
  revisions might be to add further independent analyses of the data described
  that emphasize its utility beyond comparisons to cases or add additional
  analyses to build a unique cohesive story for this paper.
\end{quote}

TODO

\begin{quote}
  The main value of the data discussed in the paper seems to focus on its
  potential utility in forecasting, which guides the main comparisons to
  observed case burden and primary outcome of ``signal value'' in results and in
  the appendix. However, there are a number of important limitations to the
  results presented that impede this message.
\end{quote}

TODO

\begin{quote}
  Most concrete results on this point seem to be presented in another paper.
  Correlations in the paper are focused on current case burden (signal vs. cases
  both at time t), rather than prediction for forecasts.
\end{quote}

TODO

\begin{quote}
  Even from a nowcasting perspective, comparisons are all benchmarked against
  cases which are observable in real-time.
\end{quote}

TODO

\begin{quote}
  The strength of observed correlations change (non-monotonically) over time,
  and the implications of this are not discussed.
\end{quote}

TODO

\begin{quote}
  Signal strength is compared only to cases (rather than downstream outcomes
  like hospitalization or death).
\end{quote}

TODO

\begin{quote}
  The case studies of signal utility that are provided seem stylized.
  Policymakers would *know* in real-time when backlogged cases are reported and
  can adjust case numbers accordingly or when cases are not being reported due
  to holiday backlogs.
\end{quote}

TODO

\begin{quote}
  As a result, it might be useful to focus quantitative analyses in this paper
  on other attributes of the data or expand analyses to include, e.g.:
  \begin{itemize}
  \item descriptive statistics from each data source, expressed in the units of
    that data source
  \item analysis of how signals differ
  \item analysis of case correlations over time due to discussion of how they
    have evolved over time -- e.g. due to changes in vaccination, testing, etc.
  \item for case studies: I think it would be more powerful to either highlight
    an instance issues were not apparent in the absence of other signals or to
    provide more in depth analysis about a specific policy conclusion that would
    result from use of these signals (or specific examples of
    geographic/temporal variation that the signal data can uniquely inform).
  \end{itemize}
\end{quote}

TODO

\begin{quote}
  A compelling framing for the results section might be to include general
  descriptive statistics and then pick case studies that illustrate 2-3 uses of
  the data.
\end{quote}

TODO

\begin{quote}
  I think that descriptions in results of ``what the data could be used for''
  are less compelling than more specific examples would be. For instance, the
  section on revised estimates over time seems to echo what was in the methods
  and implications don't really hit home without connection to changes in
  specific forecasts.
\end{quote}

TODO

\begin{quote}
  Some of the verb tenses (e.g. line 33) feel awkward: ``The Delphi Group
  **works** (worked) with partner organizations and34 public data sets to build
  a massive database of indicators35 tracking COVID-19 activity and other
  relevant phenomena36 in the United States, which has been publicly available
  and37 continuously updated since April 2020.''
\end{quote}

TODO

\begin{quote}
  It would be helpful to end the introduction with a description of the
  contribution of this paper, rather than other papers.
\end{quote}

TODO

\begin{quote}
  I think it would be useful to replace ``massive'' with a more precise
  adjective unless it is standard in the literature.
\end{quote}

TODO

\begin{quote}
  Similarly, the verb ``ingest'' also feels odd to refer to data sources.
\end{quote}

TODO

\begin{quote}
  I'm not sure that it is common practice to repeatedly refer to a sponsoring
  research group throughout an academic article.
\end{quote}

TODO

\begin{quote}
  Descriptions of statistical analyses (e.g. correlation methods) feel more
  appropriate for the methods section, not the results section.
\end{quote}
TODO

\end{document}

### 2021-09-09 indicators/paper/responses-to-reviews.tex
Some initial responses to Reviewer 2

@@
 
-TODO
+We appreciate this point and agree the message may have gotten split among
+manuscripts unnecessarily. We have made some changes to the Results section
+accordingly. Most notably, we have replaced Figure 6 with one that more clearly
+shows a potential use case of the data presented beyond forecasting of cases; it
+now shows now data sources in the API could have aided in planning for the COVID
+vaccination campaign.
+
+TODO any additional changes?
 
@@
   results presented that impede this message.
-\end{quote}
 
-TODO
-
-\begin{quote}
   Most concrete results on this point seem to be presented in another paper.
@@
 
-TODO
+Unfortunately this is not the case: Backlogs and reporting delays are generally
+poorly understood and can happen unexpectedly, and users of these datasets often
+have to make informed guesses about changes. For instance, many backlogs are due
+to retrospective audits of records that uncover cases and deaths that were
+incorrectly classified, and can affect case and death estimates months in the
+past. Until the audits are complete, policymakers would not know about the
+corrections, their magnitude, or the affects they'd have on policy decisions.
+
+Even for predictable backlogs, such as those that occur during major holidays,
+the backlog still means policymakers do not have an important signal: If case
+and death reporting is suspended for the holiday, knowing that the suspension is
+due to the holiday does not help a policymaker judge whether cases are
+increasing or decreasing and whether urgent action is required. Outside data
+sources would be needed to inform that decision.
+
+TODO Should we add anything to 2.C to make this point more obvious?
 

### 2021-09-09 indicators/code/indicator-paper-figures.Rmd
Some initial responses to Reviewer 2

@@
 ```
+
+County-level vaccination data obtained [from the
+CDC](https://data.cdc.gov/Vaccinations/COVID-19-Vaccinations-in-the-United-States-County/8xkx-amqh).
+Data downloaded on August 2, 2021.
+
+```{r figure-6-redux, message=FALSE, cache=TRUE}
+cdc <- read.csv("COVID-19_Vaccinations_in_the_United_States_County.csv") %>%
+  filter(Date == "07/20/2021")
+
+cdc_sm <- cdc %>%
+  select(FIPS, Series_Complete_18PlusPop_Pct, Administered_Dose1_Recip_18PlusPop_Pct) %>%
+  mutate(best_pop = pmax(Series_Complete_18PlusPop_Pct, Administered_Dose1_Recip_18PlusPop_Pct))
+
+vacc <- covidcast_signal("fb-survey", "smoothed_wcovid_vaccinated_or_accept",
+                         geo_type = "county", start_day = "2021-01-20",
+                         end_day = "2021-01-20")
+
+vacc_on_inauguration <- vacc %>%
+  mutate(location = abbr_to_name(geo_value, ignore.case = TRUE)) %>%
+  filter(time_value == "2021-01-20")
+
+w_cdc <- vacc_on_inauguration %>%
+  left_join(county_census, by = c("geo_value" = "FIPS")) %>%
+  left_join(cdc_sm, by = c("geo_value" = "FIPS")) %>%
+  mutate(REGION = case_when(
+    REGION == 1 ~ "Northeast",
+    REGION == 2 ~ "Midwest",
+    REGION == 3 ~ "South",
+    REGION == 4 ~ "West"))
+
+w_cdc %>%
+  filter(best_pop > 0, sample_size >= 250) %>%
+  ggplot(aes(x = value, y = Series_Complete_18PlusPop_Pct, color = REGION)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE) +
+  scale_x_continuous(labels = label_percent(scale = 1)) +
+  scale_y_continuous(labels = label_percent(scale = 1)) +
+  theme_bw() +
+  labs(x = "Estimated % willing to be vaccinated, 2021-01-20",
+       y = "% vaccinated on 2021-07-20", color = "Region") +
+  theme(legend.pos = "bottom", legend.direction = "horizontal")
+
+ggsave("../paper/fig/vaccine-intent-compare.pdf", width = 5, height = 4)
+```

### 2021-09-09 indicators/paper/paper.tex
Address Reviewer 1 comment and add response template
Some initial responses to Reviewer 2

@@
 the data. For example, data on visits to doctors' offices is subject to strong
-day-of-week effects, and so regression is used to adjust for these
-effects. Other indicators are available in raw versions and versions smoothed
-with a 7-day trailing average. All processing is done using open-source code
-written primarily in Python and R, and available publicly at
-\url{https://github.com/cmu-delphi/covidcast-indicators/}.
+day-of-week effects, and so regression is used to adjust for these effects.
+Other indicators are available in raw versions and versions smoothed with a
+7-day trailing average. All processing is done using open-source code written
+primarily in Python and R, and available publicly at
+\url{https://github.com/cmu-delphi/covidcast-indicators/}. The processing steps
+used for each signal are publicly documented on their respective pages at
+\url{https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html}.
 
@@
   \centering
-  \includegraphics[width=\columnwidth]{fig/mobility-drop.pdf}
-  \caption{Percentage point increase in SafeGraph estimates of the percentage of
-  people staying completely at home, based on aggregated mobile device data,
-  between March 1, 2020 and April 15, 2020.}
-  \label{fig:mobility-drop}
+  \includegraphics[width=\columnwidth]{fig/vaccine-intent-compare.pdf}
+  \caption{CTIS estimates of the percentage of people willing to get vaccinated
+    on January 20, 2021, compared to CDC reporting of the percentage of people
+    vaccinated in each county by July 20, 2021. Points are colored by the United
+    States Census region containing each county. Restricted to counties with at
+    least 250 survey responses from January 14--20, 2021.}
+  \label{fig:vaccine-intent-compare}
 \end{figure}
 
-As an illustration, Figure~\ref{fig:mobility-drop} maps the sharp increase in
-rates of people staying at home between March 1 and April 15, 2020. Using
-SafeGraph data on the fraction of mobile devices included in SafeGraph's panel
-that did not leave the immediate area of their home, it illustrates the sharp
-drop in travel and work outside the home that occurred in the early stages of
-the pandemic. It also shows that this drop was much more pronounced in some
-states than others, enabling analysis of policy impacts and disease spread.
-Similar maps can be quickly constructed for any signal in the COVIDcast API.
+As an illustration, Figure~\ref{fig:vaccine-intent-compare} illustrates how
+CTIS results in January 2021, based on an item asking survey respondents whether
+they would accept a COVID-19 vaccine if one were vaccinated today, predicted
+actual uptake of COVID-19 vaccines by July 2021. It also illustrates a
+geographical disparity: in the Northeast, actual vaccination rates more closely
+match vaccine acceptance in January than they do in the South, where vaccination
+rates lag overall. While these results must be interpreted carefully due to
+potential sampling biases in the survey, they illustrate the potential of data
+in the COVIDcast API to inform public health decision-making.
 

### 2021-09-10 indicators/paper/responses-to-reviews.tex
Another Reviewer 2 comment
Reviewer 2 improvements

@@
 
-TODO
+While it is difficult to fully explain the changes in correlation---they are
+likely due to a variety of changes in reporting practices, health-seeking
+behavior, survey response patterns, and so on---one key point is that the
+strength of the correlation depends on the overall differences between counties.
+That is, if counties have very similar COVID case rates, it is difficult to
+achieve a high Spearman correlation between case rates and signals, because
+getting the rank correct is difficult. If counties have widely varying COVID
+case rates, a signal that provides information about cases will be able to
+achieve a higher Spearman correlation.
+
+We have slightly expanded our discussion of Figure 2 to make this point, though
+we have tried to keep the discussion brief so as not to sidetrack the main
+discussion of the manuscript.
 
@@
   \end{itemize}
-\end{quote}
-
-TODO
-
-\begin{quote}
   A compelling framing for the results section might be to include general
@@
 
-TODO
+We appreciate the interest in additional analyses, and our replacement of Figure
+6 is partly intended to address this concern about choice of case studies. It
+represents an example showing how these data sources can reveal points not
+available in other standard data sources, such as case and death data, and serve
+to aid public health decision-making.
+
+We have chosen not to include additional descriptive statistics or comparisons
+between signals for two reasons. First, given the large number of signals
+available in our API, a table of descriptive statistics would be prohibitively
+large; this is particularly difficult if one accounts for time and geography by
+presenting statistics over time or across space. Second, it's not clear to us
+what a reader would gain from such statistics. Because the data is freely
+available, a reader interested in any specific signal can quickly investigate it
+and explore whatever aspects of it are of interest.
 
@@
 
-TODO
+We have adjusted the introduction to conclude with the contribution of this
+paper, to set it in context against the other papers.
 

### 2021-09-10 indicators/paper/paper.tex
Another Reviewer 2 comment
Reviewer 2 improvements

@@
 public surveys, run in partnership with Facebook and available in aggregate form
-in the COVIDcast API.
+in the COVIDcast API. This paper focuses on the COVID-19 activity indicators
+themselves, describing the data streams, how they are processed and made
+publicly available, and insights that can be gained by combining novel data
+sources with standard public health surveillance datasets.
 
@@
 performed over all counties with at least 500 cumulative cases by the end of
-this period, and at which all indicators are available (956 counties in
-total). The large positive correlations suggest that these signals could be
-useful in hotspot detection (identifying counties that have relatively high
-COVID activity, at a given time). Somewhat surprisingly, the survey-based
+this period, and at which all indicators are available (956 counties in total).
+The large positive correlations suggest that these signals could be useful in
+hotspot detection (identifying counties that have relatively high COVID
+activity, at a given time). Somewhat surprisingly, the survey-based
 CLI-in-community signal shows the strongest correlations for much of the time
 period. This clearly demonstrates the value of a large-scale survey such as CTIS
-for tracking symptoms and case trends, especially when other data is unavailable.
+for tracking symptoms and case trends, especially when other data is
+unavailable. The correlations vary over time in complex ways; while some of this
+variation is doubtless due to complex changes in public behavior, reporting and
+testing practices, and so on, some is also related to changes in overall
+COVID-19 case rates. For example, comparing against
+Figure~\ref{fig:time_trends_national}, we observe that correlations drop in
+February 2021 for several of the signals, matching when COVID cases also sharply
+decline. This decline likely caused the heterogeneity in case rates by county to
+decline, making it more difficult for any signal to achieve a high correlation
+by county.
 

### 2021-09-15 forecast/code/deprecated/revision-draft.R
add DM test

@@ -6,6 +6,12 @@ tt <- fcasts_honest %>%
   arrange(forecast_date) %>%
   mutate(n = n(), w = cummean(w))
 
+tt <- fcasts_honest %>%
+  group_by(forecaster, forecast_date) %>%
+  summarise(w = sum(wis) / sum(strawman_wis)) %>%
+  arrange(forecast_date) %>%
+  mutate(n = n(), w = cumsum(w))
+
 
 fcast_colors <- c(RColorBrewer::brewer.pal(5, "Set1"), "#000000")
 names(fcast_colors) <- c("CHNG-CLI", "CHNG-COVID", "CTIS-CLIIC", "DV-CLI",

### 2021-09-16 indicators/paper/responses-to-reviews.tex
Language tweaks requested by Reviewer 2

@@
 
-TODO
+We have tweaked this sentence and a few others, though we note that the present
+tense is accurate in many cases, because data collection and aggregation is
+ongoing.
 
@@
 
-TODO
+We have adjusted the language used throughout, so references to the research
+group are limited to where they are most useful to the reader.
 
@@
 \end{quote}
-TODO
+We are open to changes if the Editor feels they're necessary to match typical
+\textit{PNAS} style, but we feel that in our paper, it makes the most sense for
+the Methods to describe the methods underlying the data we gather and aggregate,
+since that data is the main focus of the article. The Results focus on
+illustrations of the data's utility, including their correlations with other
+COVID indicators.
 

### 2021-09-16 indicators/paper/paper.tex
Language tweaks requested by Reviewer 2

@@
 
-The Delphi Group works with partner organizations and public data sets to build
+The Delphi Group worked with partner organizations and public data sets to build
 a massive database of indicators tracking COVID-19 activity and other relevant
@@
 \centering
-\caption{Data sources available in Delphi's COVIDcast API \cite{CovidcastAPI},
-  as of date of publication. The first group of data sources are produced by
-  Delphi from data not otherwise available publicly (or only available in
-  limited form); the second group is mirrored from public sources.}
+\caption{Data sources available in the COVIDcast API \cite{CovidcastAPI}, as of
+  date of publication. The first group of data sources are produced from data
+  not otherwise available publicly (or only available in limited form); the
+  second group is mirrored from public sources.}
 \begin{tabular}{>{\raggedright}p{1.2in} p{4.0in} l >{\raggedright\arraybackslash}p{0.5in}}
@@
 the \textit{time value}, the date the underlying events (such as tests or
-doctor's visits) occurred, and the \textit{issue date} when Delphi aggregated
+doctor's visits) occurred, and the \textit{issue date} when we aggregated
 and reported the data for that time value. Importantly, there can be multiple
 observations for a single time value with different issue dates, for example if
-data is revised or claims records arrive late. Delphi tracks revisions to all
+data is revised or claims records arrive late. We track revisions to all
 data sources we ingest, including external data sources (such as sources
@@
 
-Since April 2020, Delphi has been maintaining and continually improving various
+Since April 2020, we have been maintaining and continually improving various
 online visualization tools for the COVIDcast indicators \cite{CovidcastViz}.
@@
 
-Since July 2020, Delphi has been regularly submitting short-term forecasts of
+Since July 2020, we have been regularly submitting short-term forecasts of
 COVID-19 case and death incidence, at the state and county levels, to the

### 2021-09-16 indicators/paper/responses-to-reviews.tex
Remove "ingest"
Be more explicit about standard reporting problems

@@
 
-TODO Should we add anything to 2.C to make this point more obvious?
+We have adjusted the language in Section 2.C to make these points clearer.
 
@@
 
-TODO
+``Ingest'' is commonly used in the data science community to refer to the
+process of extracting data and loading it into a central database; but we agree
+this use of the term may not be widespread among our audience, so we have
+adjusted the language to avoid using field-specific jargon.
 

### 2021-09-16 indicators/paper/paper.tex
Remove "ingest"
Be more explicit about standard reporting problems

@@
 the \textit{time value}, the date the underlying events (such as tests or
-doctor's visits) occurred, and the \textit{issue date} when we aggregated
-and reported the data for that time value. Importantly, there can be multiple
+doctor's visits) occurred, and the \textit{issue date} when we aggregated and
+reported the data for that time value. Importantly, there can be multiple
 observations for a single time value with different issue dates, for example if
-data is revised or claims records arrive late. We track revisions to all
-data sources we ingest, including external data sources (such as sources
+data is revised or claims records arrive late. We track revisions to all data
+sources included in the API, including external data sources (such as sources
 tracking cases and deaths). Many external sources do not keep a public or
@@
 death reporting (for this reason, the artificial dips persist at the state- and
-county-level as well). The CLI signal from the survey, on the other hand,
-displays no such dips. The claims signals actually display holiday effects going
-in the other direction: they exhibit \textit{spikes} around Thanksgiving and New
-Year's. This is because they measure the fraction of all outpatient visits with
-a certain condition, and the denominator here (total outpatient visits) drops
-disproportionately during holiday periods, as people are likely less willing to
-go to the doctor for more routine issues. Fortunately, in principle, the holiday
-effects in claims signals should be correctable: they are mainly due to
-\textit{overall} changes in medical seeking behavior during holiday, periods,
-and we can estimate such effects using historical claims data.
+county-level as well). This delay denies public health officials timely signals
+of current trends for the duration of the holidays. The CLI signal from the
+survey, on the other hand, displays no such dips. The claims signals actually
+display holiday effects going in the other direction: they exhibit
+\textit{spikes} around Thanksgiving and New Year's. This is because they measure
+the fraction of all outpatient visits with a certain condition, and the
+denominator (total outpatient visits) drops disproportionately during holiday
+periods, as people are likely less willing to go to the doctor for more routine
+issues. Fortunately, in principle, the holiday effects in claims signals should
+be correctable: they are mainly due to \textit{overall} changes in medical
+seeking behavior during holiday, periods, and we can estimate such effects using
+historical claims data.
 
@@
 reporting problems have occurred in many jurisdictions across the United States,
-making it valuable to cross-check against external sources not part of the same
-reporting systems.
+and audits have regularly discovered misclassified or unreported cases and
+deaths, making it valuable to cross-check against external sources not part of
+the same reporting systems.
 

### 2021-09-16 indicators/paper/responses-to-reviews.tex
Don't indent after quoting reviews

--- 2021-09-16 indicators/paper/responses-to-reviews.tex
@@
 \end{quote}
-
 TODO
@@
 \end{quote}
-
 Thanks, this is an important point. We have added a sentence to Section 1.B to
@@
 \end{quote}
-
 We appreciate this point and agree the message may have gotten split among
@@
 \end{quote}
-
 While it is difficult to fully explain the changes in correlation---they are
@@
 \end{quote}
-
 Unfortunately this is not the case: Backlogs and reporting delays are generally
@@
 \end{quote}
-
 We appreciate the interest in additional analyses, and our replacement of Figure
@@
 \end{quote}
-
 We have tweaked this sentence and a few others, though we note that the present
@@
 \end{quote}
-
 We have adjusted the introduction to conclude with the contribution of this
@@
 \end{quote}
-
-TODO
+``Ingest'' is commonly used in the data science community to refer to the
+process of extracting data and loading it into a central database; but we agree
+this use of the term may not be widespread among our audience, so we have
+adjusted the language to avoid using field-specific jargon.
 
@@
 \end{quote}
-
 We have adjusted the language used throughout, so references to the research

### 2021-09-16 indicators/paper/paper.tex
Additional correlation analyses with hospitalization, for supplement

@@
 trailing average) over a year of the pandemic (April 15, 2020 to April 15,
-2021), illustrating how they track national trends quite well.  Importantly,
-this same relationship persists across multiple resolutions of the data, down to
+2021), illustrating how they track national trends quite well. Importantly, this
+same relationship persists across multiple resolutions of the data, down to
 smaller geographic regions such as states and counties, as shown in the
-supplement.  This will also be illustrated in a more detailed correlation
-analysis in the next subsection.
+Supplementary Information. This will also be illustrated in a more detailed
+correlation analysis in the next subsection.
 
@@
 \end{itemize}
+The Supplementary Information contains additional correlation analyses that
+compare these COVID-related signals to COVID hospitalizations reported by the
+Department of Health and Human Services. These show similar results,
+illustrating that the signals are useful for tracking key health outcomes.
 

### 2021-09-16 indicators/paper/supplement.Rmd
Additional correlation analyses with hospitalization, for supplement

@@
 \includegraphics[width=\textwidth]{fig/severity-pyramid.pdf}
-\caption{The epidemiological "severity pyramid" represents the progression of
+\caption{The epidemiological ``severity pyramid'' represents the progression of
 cases from the public, through infection, through increasingly severe stages of
 disease. The annotations here represent the data sources collected by Delphi's
-COVIDcast Epidata API. TODO Edit to use consistent language to paper}
+COVIDcast Epidata API.}
 \end{figure}
@@
 
-```{r hospitalization_time_trends_national, fig.cap="National trends, from April 2020 to April 2021, of HHS-reported confirmed COVID-19 hospital admissions, plus another hospitalization signal in the COVIDcast API.
-Here, the HHS data has been limited to values from 2020-08-01 onward due to changes in reporting behavior, and processed into a 7-day trailing average rate per 100,000 resident population using Census Bureau estimates for 2019.
+```{r hospitalization_time_trends_national, fig.cap="National trends, from August 2020 to August 2021, of HHS-reported confirmed COVID-19 hospital admissions, plus comparison signals in the COVIDcast API.
+(HHS data was not consistently reported before August 2020.) Here, the HHS data has been processed into a 7-day trailing average rate per 100,000 resident population using Census Bureau estimates for 2019.
 HSP-Hosp is the percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners, smoothed in time and adjusted for systematic day-of-week effects (HSP-Hosp is available from an earlier date when working at finer geographical resolutions).
-HSP-Hosp has been scaled to have the same maximum value as the HHS data."}
-start_day = "2020-04-15"
-end_day = "2021-04-15"
-as_of_day = "2021-06-01"
+HSP-Hosp, CTIS-CLI-in-community, and CHNG-CLI have been scaled to have the same maximum value as the HHS data."}
+start_day = "2020-08-15"
+end_day = "2021-08-15"
+as_of_day = "2021-09-01"
 
-HOSP_SOURCES <- c("hhs", "hospital-admissions")
-HOSP_SIGNALS <- c("confirmed_admissions_covid_1d", "smoothed_adj_covid19_from_claims")
-hosp_names = c("Hospitalizations", "HSP-Hosp")
-hosp_signal_has_national = c(TRUE, TRUE)
+HOSP_SOURCES <- c("hhs", "jhu-csse", "hospital-admissions", "fb-survey", "chng")
+HOSP_SIGNALS <- c("confirmed_admissions_covid_1d", "confirmed_7dav_incidence_prop",
+                  "smoothed_adj_covid19_from_claims",
+                  "smoothed_whh_cmnty_cli", "smoothed_adj_outpatient_covid")
+hosp_names = c("Hospitalizations", "Cases", "HSP-Hosp",
+               "CTIS-CLI-in-community", "CHNG-CLI")
+hosp_signal_has_national = c(TRUE, TRUE, TRUE, TRUE, TRUE)
 hhs_start_date = as.Date("2020-08-01")
@@
 
-```{r hosp-correlations-by-time, fig.cap="Geo-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated for all times with sufficient available data within this period, over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which both signals are available."}
+```{r hosp-correlations-by-time, fig.cap="Geo-wise correlations with hospitalization rates derived from HHS data, from August 15, 2020 to August 15, 2021, calculated for all times with sufficient available data within this period, over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which all signals are available."}
 n_hosp_signals = length(HOSP_SIGNALS)
@@
 
-```{r hosp-correlations-by-state, fig.cap="Time-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which both signals are available."}
+```{r hosp-correlations-by-state, fig.cap="Time-wise correlations with hospitalization rates derived from HHS data, from August 15, 2020 to August 15, 2021, calculated over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which all signals are available."}
 df_list_hosp_timewise_cor = vector("list", length(HOSP_SIGNALS))

### 2021-09-16 indicators/paper/responses-to-reviews.tex
Point Reviewer 2 at the supplement

@@
 \end{quote}
-TODO
+We have extended the Supplementary Information to include more detailed
+comparisons between our signals and reported COVID hospitalizations, including
+comparisons of correlations. These show broadly similar results as the COVID
+case analyses. Section 2.B now points to the Supplement for interested readers
+to examine these analyses.
 

### 2021-09-16 indicators/paper/responses-to-reviews.tex
Point out that case lag means correlations are not contemporaneous

@@
 \end{quote}
-TODO
+We have extended the analysis to include comparisons against hospitalizations as
+well; see below.
+
+But it is worth emphasizing that cases are not actually observable in real time.
+Case reporting typically occurs with lag; lag was often high early in the
+pandemic when testing was limited and health departments were struggling to
+scale up reporting systems, so case reports for a particular date may not have
+been available until several days or a week later. This gradually changed until
+most states reported cases daily---though these reports were still subject to
+lag and referred to cases that may have been diagnosed several days earlier.
+Now, some states have returned to a weekly reporting schedule, so case reports
+only become available once a week, and so cases may take up to seven days to be
+reported.
+
+When our correlation analyses show a high correlation between a signal and case
+reports on a specific date, that signal may have been available several days
+before the case reports were, depending on the time this occurred during the
+pandemic and the state's reporting practices. We have expanded Section 2.B to
+make this point clearer.
 

### 2021-09-16 indicators/paper/paper.tex
Point out that case lag means correlations are not contemporaneous

@@
 CTIS-CLI-in-community signal is a close second, and the CHNG-CLI signal is
-third. There are two noteworthy points:
+third. There are three noteworthy points:
 
@@
   a confirmed case passing through the outpatient system.
+
+\item As we will discuss below in more detail, public health reporting is often
+  subject to delays and backlogs that cause data to be reported with lag.
+  Meanwhile, signals such as CTIS-CLI-in-community can be reported with as
+  little as 1--2 days of lag. These high correlations hence indicate a high
+  correlation with a case value that may have only been officially reported days
+  after these signals.
 \end{itemize}

### 2021-09-16 indicators/paper/responses-to-reviews.tex
De-massivate

@@
   both at time t), rather than prediction for forecasts.
-\end{quote}
-TODO
 
-\begin{quote}
   Even from a nowcasting perspective, comparisons are all benchmarked against
@@
 \end{quote}
-
-TODO
+We have slightly adjusted our language here. We still use ``massive'' to refer
+to our online surveys, in line with other uses of the term, such as ``massive
+open online courses;'' we also feel this reflects its truly unprecedented scale,
+as the survey is the largest survey ever conducted (except for censuses).
 

### 2021-09-16 indicators/paper/paper.tex
De-massivate

@@
 The Delphi Group worked with partner organizations and public data sets to build
-a massive database of indicators tracking COVID-19 activity and other relevant
-phenomena in the United States, which has been publicly available and
+a large-scale database of indicators tracking COVID-19 activity and other
+relevant phenomena in the United States, which has been publicly available and
 continuously updated since April 2020. Alongside public data on reported cases

### 2021-09-17 indicators/paper/responses-to-reviews.tex
Fix typos

@@
 shows a potential use case of the data presented beyond forecasting of cases; it
-now shows now data sources in the API could have aided in planning for the COVID
+now shows how data sources in the API could have aided in planning for the COVID
 vaccination campaign.
@@
 past. Until the audits are complete, policymakers would not know about the
-corrections, their magnitude, or the affects they'd have on policy decisions.
+corrections, their magnitude, or the effects they'd have on policy decisions.
 

### 2021-09-17 survey/code/paper-tables.Rmd
WIP Figure 2 reimplementation in R

@@
 library(tidycensus)
+library(zoo)
 
@@
 
+## Trends in each signal, by Census region
+
+```{r fetch-signal-data}
+start_day <- "2020-04-06"
+end_day <- "2021-04-05"
+
+anosmia <- data %>%
+  mutate(date = as.Date(StartDatetime),
+         anosmia = is_selected(B2, "13"),
+         state = toupper(name_to_abbr(state_fips_to_name(fips)))) %>%
+  group_by(date, state) %>%
+  summarize(anosmia = weighted.mean(anosmia, weight, na.rm = TRUE)) %>%
+  select(time_value = date, value = anosmia, geo_value = state) %>%
+  as.covidcast_signal(signal = "anosmia")
+
+signals <- covidcast_signals(
+  c("fb-survey", "fb-survey", "jhu-csse", "hhs", "jhu-csse"),
+  c("raw_wcli", "raw_whh_cmnty_cli", "confirmed_incidence_prop",
+    "confirmed_admissions_covid_1d_prop", "deaths_incidence_prop"),
+  start_day, end_day,
+  geo_type = "state") %>%
+  aggregate_signals(format = "long")
+
+signals <- append(signals, anosmia)
+
+signals$signal <- factor(
+  signals$signal,
+  levels = c("anosmia", "raw_wcli", "raw_whh_cmnty_cli", "confirmed_incidence_prop",
+             "confirmed_admissions_covid_1d_prop", "deaths_incidence_prop"),
+  labels = c("Anosmia %", "CLI %", "Community CLI %", "Cases per 100,000",
+             "Hospitalizations per 100,000", "Deaths per 100,000"))
+
+to_plot <- signals %>%
+  mutate(geo_value = toupper(geo_value)) %>%
+  inner_join(select(state_census, ABBR, REGION, POPESTIMATE2019),
+             by = c("geo_value" = "ABBR")) %>%
+  group_by(REGION, time_value, signal) %>%
+  summarize(value = weighted.mean(value,
+                                  ifelse(data_source %in% c("jhu-csse", "hhs"),
+                                         POPESTIMATE2019, sample_size),
+                                  na.rm = TRUE))
+
+to_plot <- to_plot %>%
+  group_by(REGION, signal) %>%
+  arrange(time_value) %>%
+  mutate(value = rollmean(value, 7, fill = NA))
+
+ggplot(to_plot, aes(x = time_value, y = value, color = REGION)) +
+  geom_line() +
+  facet_wrap(vars(signal), nrow = 2, ncol = 3, scales = "free_y") +
+  labs(x = "Date", y = "Value") +
+  scale_x_date(date_breaks = "3 months",
+               date_labels = "%b %Y",
+               date_minor_breaks = "1 month") +
+  theme_bw() +
+  theme(legend.position = "bottom")
+```
+
 ## Cumulative test positivity

### 2021-09-20 survey/code/paper-tables.Rmd
Tidy up code and figures
1. Convert everything to put figures in the right directory.

2. Use consistent population data.

3. Move legends on contacts plot so the plots have consistent x axes.

4. Finish signal time series plot.

--- 2021-07-20 survey/code/paper-tables.Rmd
@@
 gc()
+```
+
+Some essential cleaning. We use 5-digit FIPS codes with leading zeros by
+convention, and we'll want region names for plots:
+
+```{r clean-data}
+data$fips <- ifelse(is.na(data$fips), NA_character_, sprintf("%05d", data$fips))
+
+state_census$REGION <- case_when(
+  state_census$REGION == 1 ~ "Northeast",
+  state_census$REGION == 2 ~ "Midwest",
+  state_census$REGION == 3 ~ "South",
+  state_census$REGION == 4 ~ "West"
+)
 ```
@@
   select(fips, weight) %>%
-  mutate(
-    fips = ifelse(is.na(fips), NA_character_, sprintf("%05d", fips)),
-    state_abbr = toupper(name_to_abbr(state_fips_to_name(fips)))) %>%
+  mutate(state_abbr = toupper(name_to_abbr(state_fips_to_name(fips)))) %>%
   inner_join(select(state_census, REGION, ABBR),
              by = c("state_abbr" =  "ABBR")) %>%
-  mutate(REGION = case_when(
-    REGION == 1 ~ "Northeast",
-    REGION == 2 ~ "Midwest",
-    REGION == 3 ~ "South",
-    REGION == 4 ~ "West"
-  )) %>%
   proportionate(REGION) %>%
@@
 
-### Tested given symptoms
+## Trends in each signal, by Census region
+
+Figure 2:
+
+```{r signal-anosmia-trends, message=FALSE}
+gc()
+
+start_day <- "2020-04-06"
+end_day <- "2021-04-05"
+
+anosmia <- data %>%
+  select(StartDatetime, B2, fips, weight) %>%
+  filter(!is.na(fips)) %>%
+  mutate(date = as.Date(StartDatetime),
+         anosmia = is_selected(split_options(B2), "13"),
+         state = toupper(name_to_abbr(state_fips_to_name(fips)))) %>%
+  group_by(date, state) %>%
+  summarize(anosmia = 100 * weighted.mean(anosmia, weight, na.rm = TRUE),
+            sample_size = sum(!is.na(anosmia))) %>%
+  select(time_value = date, value = anosmia, geo_value = state, sample_size) %>%
+  as.covidcast_signal(signal = "anosmia", geo_type = "state")
+
+signals <- covidcast_signals(
+  c("fb-survey", "fb-survey", "jhu-csse", "hhs", "jhu-csse"),
+  c("raw_wcli", "raw_whh_cmnty_cli", "confirmed_incidence_prop",
+    "confirmed_admissions_covid_1d_prop", "deaths_incidence_prop"),
+  start_day, end_day,
+  geo_type = "state")
+
+signals <- append(signals, list(anosmia)) %>%
+  aggregate_signals(format = "long")
+
+signals$signal <- factor(
+  signals$signal,
+  levels = c("anosmia", "raw_wcli", "raw_whh_cmnty_cli", "confirmed_incidence_prop",
+             "confirmed_admissions_covid_1d_prop", "deaths_incidence_prop"),
+  labels = c("Anosmia %", "CLI %", "Community CLI %", "Cases per 100,000",
+             "Hospitalizations per 100,000", "Deaths per 100,000"))
+
+to_plot <- signals %>%
+  mutate(geo_value = toupper(geo_value)) %>%
+  inner_join(select(state_census, ABBR, REGION, POPESTIMATE2019),
+             by = c("geo_value" = "ABBR")) %>%
+  group_by(REGION, time_value, signal) %>%
+  summarize(value = weighted.mean(value,
+                                  ifelse(data_source %in% c("jhu-csse", "hhs"),
+                                         POPESTIMATE2019, sample_size),
+                                  na.rm = TRUE))
+
+to_plot <- to_plot %>%
+  filter(REGION != "X") %>%
+  group_by(REGION, signal) %>%
+  arrange(time_value) %>%
+  mutate(value = rollmean(value, 7, fill = NA))
+
+ggplot(to_plot, aes(x = time_value, y = value, color = REGION)) +
+  geom_line() +
+  facet_wrap(vars(signal), nrow = 2, ncol = 3, scales = "free_y") +
+  labs(x = "Date", y = "Value", color = "Census Region") +
+  scale_x_date(date_breaks = "3 months",
+               date_labels = "%b %Y",
+               date_minor_breaks = "1 month",
+               guide = guide_axis(angle = 45)) +
+  theme_bw() +
+  theme(legend.position = "bottom")
+
+ggsave("../paper/fig/signal-time-series.pdf",
+       width = 7, height = 5)
+```
 
@@
 
-Now plot:
+Now plot (Figure 1):
 
@@
 
-ggsave("figures/symptoms-comparison.pdf")
+ggsave("../paper/fig/symptoms-comparison.pdf")
 ```
@@
 
-Based on code from Minttu Rönn.
+Figure 3. Based on code from Minttu Rönn.
 
@@
 
-```{r case-rate-compare}
-# State adult populations, from Census data
+```{r case-rate-compare, message=FALSE}
+# Estimates of the Total Resident Population and Resident Population Age 18
+# Years and Older for the United States, States, and Puerto Rico: July 1, 2019
+# (SCPRC-EST2019-18+POP-RES)
 state_pops <- read.csv("state-age-pops.csv")
+
 # State-level age distributions of cases, from CDC data
@@
     week = floor_date(as.Date(StartDatetime), unit = "week"),
-    fips = sprintf("%05d", fips),
     state_abbr = name_to_abbr(state_fips_to_name(fips))
@@
 
-ggsave("figures/cumulative-case-compare.pdf", g, width = 10, height = 7)
+ggsave("../paper/fig/cumulative-case-compare.pdf", g, width = 10, height = 7)
 
@@
 
-```{r}
+```{r case-rate-correlations}
 max(state_case_compare$week)
@@
 
-Based on code from Alyssa Bilinski.
+Figure 4. Based on code from Alyssa Bilinski.
 
@@
 
-pops <- read.csv("data/population.csv")
+# We don't actually need populations -- just region codes
+pops <- state_census %>%
+  select(state = NAME, region = REGION)
 
@@
                              RPL_THEMES >= 0.25 & RPL_THEMES < 0.5 ~ "3rd quartile",
-                             RPL_THEMES < 0.25 ~ "4th quartile")) %>%
+                             RPL_THEMES < 0.25 ~ "4th quartile"),
+         FIPS = sprintf("%05d", FIPS)) %>%
   select(FIPS, SVI_cat)
@@
     C13_options = split_options(C13),
-    state = state_fips_to_name(ifelse(is.na(fips), NA_character_, sprintf("%05d", fips))),
+    state = state_fips_to_name(fips),
     week = floor_date(as.Date(StartDatetime), unit = "week"),
@@
   left_join(svi, by = c("fips" = "FIPS")) %>%
-  left_join(pops, by = c("state" = "StateName"))
+  left_join(pops, by = c("state"))
 ```
@@
 
-# TODO See if we can merge these into one big facet_grid so the widths match
-# (not sure how to deal with legends?)
+# Pull the legends up towards the plots to eliminate wasted space. ggplot's
+# default unit is half a line height, and a line height is 11 pt, so work in
+# multiples of this.
+legend_top_margin <- -11 * 3 / 2
+
 a <- ggplot(g1, aes(x = week, y = value, col = SVI_cat, group = SVI_cat)) +
@@
   ylim(0,  NA) +
-  theme_bw()
+  theme_bw() +
+  theme(legend.position = "bottom",
+        legend.margin = margin(legend_top_margin, 0, 11 / 2, 0),
+        plot.margin = margin(11 / 2, 11 / 2, 0, 11 / 2))
 
@@
   ylim(0, NA) +
-  theme_bw()
+  theme_bw() +
+  theme(legend.position = "bottom",
+        legend.margin = margin(legend_top_margin, 0, 11 / 2, 0),
+        plot.margin = margin(11 / 2, 11 / 2, 0, 11 / 2))
 
@@
   geom_line() +
-  facet_wrap(.~var_cat, scales = "free", ncol = 4) +
+  facet_wrap(. ~ var_cat, scales = "free", ncol = 4) +
   scale_color_discrete(name = "Region")  +
@@
 
-ggsave("figures/contacts-masks-transport.pdf", g, width = 8, height = 5)
+ggsave("../paper/fig/contacts-masks-transport.pdf", g, width = 8, height = 5)
 ```
@@
 
-Based on code from Marissa Reitsma.
+Figure 5B. Based on code from Marissa Reitsma.
 
@@
          date <= as.Date("2021-04-05")) %>%
-  mutate(
-    hesitant = case_when(
+  mutate(hesitant = case_when(
              V1 == 1 ~ FALSE,
              V3 %in% 1:2 ~ FALSE,
-      V3 %in% 3:4 ~ TRUE),
-    fips = sprintf("%05d", fips))
+             V3 %in% 3:4 ~ TRUE))
 
@@
 
-ggsave("figures/hesitancy-county-map.pdf", g, width = 6, height = 4)
+ggsave("../paper/fig/hesitancy-county-map.pdf", g, width = 6, height = 4)
 

### 2021-09-20 survey/code/vaccine-acceptance-race.R
Add missing code

@@ -1,0 +1,105 @@
+# This code calculates vaccine acceptance by race/ethnicity, using data files
+# that are not yet available to users with a data use agreement. Information
+# about obtaining access is available on our website, and will be updated when
+# race/ethnicity data becomes available:
+# https://cmu-delphi.github.io/delphi-epidata/symptom-survey/data-access.html
+
+# Code by Marissa Reitsma.
+
+library(dplyr)
+library(data.table)
+library(ggplot2)
+library(tidycensus)
+library(scales)
+
+df <- fread("2021-03-race-ethnicity-microdata.csv.gz") %>%
+  select(StartDatetime, fips, V1, V3, D6, D7, D2, D1, weight) %>%
+  filter(!is.na(fips)) %>%
+  mutate(fips = sprintf("%05d", fips),
+         accepting = case_when(
+           V1 == 1 ~ TRUE,
+           V3 %in% 1:2 ~ TRUE,
+           V3 %in% 3:4 ~ FALSE))
+
+regions <- fread("data/regions.csv")
+
+fips <- as.data.table(tidycensus::fips_codes)
+fips <- fips[, fips := paste0(state_code, county_code)]
+
+############################
+## Prepare the indicators ##
+############################
+
+df <- merge(df, fips, by = "fips")
+df <- merge(df, regions[, .(state_name, Region)], by = "state_name") # This excludes Puerto Rico
+
+## Race/ethnicity
+df <- df[D6 == 1, race_grp:="Hispanic"]
+df <- df[D7 %like% "," & is.na(race_grp), race_grp := "Multiracial"]
+df <- df[D7 %like% "1" & is.na(race_grp), race_grp := "American Indian or Alaska Native"]
+df <- df[D7 %like% "2" & is.na(race_grp), race_grp := "Asian"]
+df <- df[D7 %like% "3" & is.na(race_grp), race_grp := "Black"]
+df <- df[D7 %like% "4" & is.na(race_grp), race_grp := "Native Hawaiian or other Pacific Islander"]
+df <- df[D7 %like% "5" & is.na(race_grp), race_grp := "White"]
+df <- df[D7 %like% "6" & is.na(race_grp), race_grp := "Other"]
+
+## Age
+df <- df[D2 == 1, age_cat_fb := "18-24"]
+df <- df[D2 == 2, age_cat_fb := "25-34"]
[…]

### 2021-09-20 survey/code/survey-utils.R
Add missing code

@@ -1,0 +1,97 @@
+library(readr)
+library(purrr)
+library(dplyr)
+
+#' Fetch all survey data in a chosen directory.
+#'
+#' This function should be used on the monthly rollup CSVs available to data
+#' users, with names like YYYY-MM.csv.gz.
+#'
+#' This function extracts the date from each file and produces a single data
+#' frame representing the most recent data available for each day. It can read
+#' gzip-compressed CSV files, such as those on the SFTP site, using
+#' `readr::read_csv`.
+#'
+#' @param directory Directory in which to look for survey CSV files, relative to
+#'   the current working directory.
+#' @param pattern Regular expression indicating which files in that directory to
+#'   open. By default, selects all `.csv.gz` files, such as those downloaded
+#'   from the SFTP site.
+#' @return A single data frame containing all survey responses. Note that this
+#'   data frame may have millions of rows and use gigabytes of memory, if this
+#'   function is run on *all* survey responses.
+get_survey_df <- function(directory, pattern = "*.csv.gz$") {
+  files <- list.files(directory, pattern = pattern)
+
+  big_df <- map_dfr(
+    files,
+    function(f) {
+      # stop readr from thinking commas = thousand separators,
+      # and from inferring column types incorrectly
+      read_csv(file.path(directory, f),
+               locale = locale(grouping_mark = ""),
+               col_types = cols(
+                 A2b = col_number(),
+                 A3 = col_character(),
+                 B2 = col_character(),
+                 B2_14_TEXT = col_character(),
+                 B2c = col_character(),
+                 B2c_14_TEXT = col_character(),
+                 B7 = col_character(),
+                 B10b = col_character(),
+                 B12a = col_character(),
+                 V5a = col_character(),
+                 V5b = col_character(),
+                 V5c = col_character(),
+                 V5d = col_character(),
+                 V6 = col_character(),
+                 C1 = col_character(),
+                 C13 = col_character(),
+                 C13a = col_character(),
+		 C13b = col_character(),
+                 D1_4_TEXT = col_character(),
+                 D7 = col_character(),
+                 fips = col_character(),
+                 UserLanguage = col_character(),
+                 StartDatetime = col_character(),
+                 EndDatetime = col_character(),
+                 .default = col_number()))
+    }
+  )
+  return(big_df)
+}
+
+#' Split multiselect options into codable form
+#'
+#' Multiselect options are coded by Qualtrics as a comma-separated string of
[…]

### 2021-09-23 indicators/paper/paper.tex
Lowercase group

--- 2021-09-16 indicators/paper/paper.tex
@@
 
-The Delphi Group worked with partner organizations and public data sets to build
-a massive database of indicators tracking COVID-19 activity and other relevant
-phenomena in the United States, which has been publicly available and
+The Delphi group worked with partner organizations and public data sets to build
+a large-scale database of indicators tracking COVID-19 activity and other
+relevant phenomena in the United States, which has been publicly available and
 continuously updated since April 2020. Alongside public data on reported cases

### 2021-09-23 indicators/paper/paper.tex
Additional research reference

@@
 the pandemic \cite{Adjodah:2021, Pierri:2021, Jewell:2021, Chakrabarti:2020,
-  Doerr:2021}. The API currently serves hundreds of thousands of requests to
-thousands of users every day.
+  Doerr:2021, Rebeiro:2021}. The API currently serves hundreds of thousands of
+requests to thousands of users every day.
 

### 2021-09-23 indicators/paper/responses-to-reviews.tex
Plot location, summary of review responses

@@
 
-\title{Response to reviewers}
+\title{Response to peer reviews}
 \author{}
@@
 \end{quote}
-TODO
+We thank the editor and reviewers for their comments and the opportunity to make
+revisions. We have made a number of revisions, described in detail below; in
+summary, these revisions include:
+\begin{itemize}
+\item Additional graphics in the Supplementary Information, and additional
+  references to it where appropriate, to provide additional useful ways to view
+  the data and comparisons to outcomes other than COVID-19 cases
+
+\item Clarification of limitations, both in our data and in other publicly
+  available data sources, so the strengths and weaknesses of our data compared
+  to other sources are clearer
+
+\item Improved analyses in the Results section, including more pertinent figures
+  and clearer examples.
+\end{itemize}
+We believe these changes have strengthened the manuscript, and describe them in
+more detail below.
 
@@
 
+Our Supplementary Information does include additional exploratory analyses of
+the signals, including state- and county-level time series plots that reveal
+interesting signal behavior.
+
 \begin{quote}

### 2021-09-23 indicators/paper/paper.tex
Plot location, summary of review responses

@@
 
+% TODO Recount!
 Altogether, we collect 110 signals from 12 distinct sources, aggregate and
@@
 
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{fig/time_wise_correlation.pdf}
+  \caption{Time-wise correlations with case rates, from April 15, 2020 to April
+    15, 2021, calculated over all counties for which all signals were available
+    and which had at least 500 cumulative cases by the end of this period.}
+  \label{fig:time_wise_correlation}
+\end{figure}
+
 \begin{itemize}
@@
 historical claims data.
-
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/time_wise_correlation.pdf}
-  \caption{Time-wise correlations with case rates, from April 15, 2020 to April
-    15, 2021, calculated over all counties for which all signals were available
-    and which had at least 500 cumulative cases by the end of this period.}
-  \label{fig:time_wise_correlation}
-\end{figure}
 
@@
 
-As an illustration, Figure~\ref{fig:vaccine-intent-compare} illustrates how
-CTIS results in January 2021, based on an item asking survey respondents whether
-they would accept a COVID-19 vaccine if one were vaccinated today, predicted
-actual uptake of COVID-19 vaccines by July 2021. It also illustrates a
-geographical disparity: in the Northeast, actual vaccination rates more closely
-match vaccine acceptance in January than they do in the South, where vaccination
-rates lag overall. While these results must be interpreted carefully due to
-potential sampling biases in the survey, they illustrate the potential of data
-in the COVIDcast API to inform public health decision-making.
+As an illustration, Figure~\ref{fig:vaccine-intent-compare} illustrates how CTIS
+results in January 2021, based on an item asking survey respondents whether they
+would accept a COVID-19 vaccine if one were vaccinated today, predicted actual
+uptake of COVID-19 vaccines by July 2021 (as reported by the Centers for Disease
+Control and Prevention). It also illustrates a geographical disparity: in the
+Northeast, actual vaccination rates more closely match vaccine acceptance in
+January than they do in the South, where vaccination rates lag overall. While
+these results must be interpreted carefully due to potential sampling biases in
+the survey, they illustrate the potential of data in the COVIDcast API to inform
+public health decision-making.
 

### 2021-09-24 forecast/paper/supplement-text.tex
add statistical significance section to forecast paper supplement
- Some supplemental figures need some cleanup.
- Check on the naming of "all_signals_wide" in the Data Store.

@@
 for \chngcli.
+
+\section{Statistical significance}
+
+In the Introduction of the main paper, we give some reasons that we avoid making
+formal statements about statistical significance, preferring instead to examine
+the stability of our results in different contexts. There are strong reasons to
+avoid model-based significance tests because the necessary assumptions about
+stationarity, independence, and the  model being true (or at least approximately
+true) are certainly violated. With those caveats in mind, we undertake two
+relatively assumption-lean investigations in this section. The first is a
+sign-test for whether the difference between the AR model’s relative WIS  and
+each other model’s relative WIS is centered at zero. By ``relative WIS'' we mean
+scaled by the strawman as displayed in Figure 3. To mitigate the dependence
+across time (which intuitively seems to matter more than that across space), we
+computed these tests in a stratified way, where for each forecast date we run a
+sign test on the scaled errors between two models over all 440 counties. The
+results are plotted as histograms in Figure \ref{fig:sign-test}. In this case,
+we use the total relative WIS over all aheads, but the histograms are largely
+similar for individual target horizons. If there were no difference, we would
+expect to see a uniform distribution. However, for all indicator-assisted
+models, we see many more small $P$-values than would be expected if the null
+hypothesis (that the AR model is better) were true.
+
+Another relatively assumption-lean method of testing for differences in forecast
+accuracy is the Diebold-Mariano test \cite{Diebold:2002,Diebold:2015,harvey1997testing}.
+Essentially, the differences between forecast losses are assumed to have a
+constant mean and a covariance that depends on time. Under these conditions, the
+asymptotic distribution for the standardized mean of the differences is limiting
+normal provided that a heteroskedasticity and autocorrelation robust estimate of
+the variance is used. Using the loss as weighted interval score across all HRRs
+and horizons (7 to 21 days ahead), we perform the DM test using both the mean
+relative to the strawman (as reported in the manuscript) and the geometric mean
+relative to the strawman as described above. The first two rows of Table
+\ref{tab:dm-test} displays 
+$P$-values for the test that the indicator-assisted model is no-better than the
+AR model. Only the CHNG-CLI model's $P$-value exceeds conventional statistical
+significance thresholds.
 

### 2021-09-24 forecast/paper/supplement.Rmd
add statistical significance section to forecast paper supplement
- Some supplemental figures need some cleanup.
- Check on the naming of "all_signals_wide" in the Data Store.

@@
 ---
-bibliography: [../../common/covidcast.bib, pnas-materials/pnas-sample.bib]
+bibliography: [../../common/covidcast.bib]
 output: 
@@
   hotspots_honest %>%
+    filter(period != "jm") %>%
     group_by(forecaster, ahead) %>%
@@
   hotspots_finalized %>%
+    filter(period != "jm") %>%
     group_by(forecaster, ahead) %>%
@@
 \clearpage
+
+<!-- Statistical significance -->
+
+```{r sign-test, fig.cap="P-values for a sign test that the WIS of the AR forecaster is smaller than that of the indicator-assisted model. Each P-value corresponds a particular forecast date."}
+fcast_colors2 <- fcast_colors[names(fcast_colors) != "AR"]
+
+st <- fcasts_honest %>%
+  filter(period != "jm") %>%
+  mutate(wis = wis / strawman_wis) %>%
+  select(forecaster, geo_value, ahead, forecast_date, wis) %>%
+  pivot_wider(names_from = forecaster, values_from = wis) %>%
+  mutate(across(AR:`Google-AA`, ~ AR - .x)) %>%
+  select(-AR) %>%
+  pivot_longer(`CHNG-CLI`:`Google-AA`, names_to = "forecaster", values_to = "dif") %>%
+  group_by(forecaster, forecast_date, geo_value) %>%
+  summarise(dif = sum(dif)) %>%
+  group_by(forecaster, forecast_date) %>%
+  summarise(p = binom.test(
+    x = sum(dif > 0),
+    n = n(),
+    alternative = "greater")$p.val)
+
+ggplot(st, aes(p)) +
+  geom_histogram(
+    aes(y = after_stat(density), color = forecaster, fill = forecaster), 
+    alpha = 0.4,
+    bins = 40) +
+  scale_color_manual(values = fcast_colors2, guide = guide_legend(nrow = 1)) +
+  scale_fill_manual(values = fcast_colors2, guide = guide_legend(nrow = 1)) +
+  facet_wrap(~forecaster) +
+  ylab("Frequency") +
+  xlab("P-value") +
+  scale_y_continuous(
+    labels = scales::percent_format(scale = 1, accuracy = 1)) +
+  theme_bw() +
+  theme(legend.position = "bottom")
+
+```
+
+```{r dm-test}
+errs <- fcasts_honest %>% 
+  filter(period != "jm") %>%
+  group_by(forecaster, forecast_date) %>%
+  summarise(
+    `Mean Relative WIS` = mean(wis) / mean(strawman_wis),
+    # mwis = mean(wis), 
+    `Geometric Mean Relative WIS` = GeoMean(wis / strawman_wis) #,
+    # gmwis = GeoMean(wis)
+    ) %>%
+  ungroup()
+ar <- errs %>% 
+  filter(forecaster == "AR") %>% 
+  select(-forecaster) %>%
+  pivot_longer(-forecast_date, names_to = "metric", values_to = "AR")
+notar <- errs %>%
+  filter(forecaster != "AR") %>% 
+  pivot_longer(-c("forecaster", "forecast_date"), names_to = "metric", values_to = "err")
+
+all_errs <- left_join(notar, ar)
+dms <- all_errs %>% 
+  group_by(forecaster, metric) %>% 
+  summarise(dm = forecast::dm.test(AR, err, "greater", h = 7 , power = 1)$p.value)
+
+kableExtra::kable(
+  bind_rows(dms) %>% 
+    pivot_wider(names_from = forecaster, values_from = dm), 
+  digits = 3, booktabs = TRUE, caption = "$P$-values for a Diebold-Mariano test for differences in forecast error.
+The test is for the null hypothesis of equal performance against the alternative that the indicator assisted model is better.")
+```
 

### 2021-09-24 forecast/paper/supplement.tex
add statistical significance section to forecast paper supplement
- Some supplemental figures need some cleanup.
- Check on the naming of "all_signals_wide" in the Data Store.

@@
 
-\author{Daniel J. McDonald, Jacob Bien, Alden Green, Addison Hu, Nat DeFries,
+\author{Daniel J. McDonald, Jacob Bien, Alden Green, Addison J. Hu, Nat DeFries,
   Sangwon Hyun, Natalia L. Oliveira, James Sharpnack, Jingjing Tang, Robert
@@
 \clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/sign-test-1}
+
+}
+
+\caption{P-values for a sign test that the WIS of the AR forecaster is smaller than that of the indicator-assisted model. Each P-value corresponds a particular forecast date.}\label{fig:sign-test}
+\end{figure}
+
+\begin{table}
+
+\caption{\label{tab:dm-test}$P$-values for a Diebold-Mariano test for differences in forecast error.
+The test is for the null hypothesis of equal performance against the alternative that the indicator assisted model is better.}
+\centering
+\begin{tabular}[t]{lrrrrr}
+\toprule
+metric & CHNG-CLI & CHNG-COVID & CTIS-CLIIC & DV-CLI & Google-AA\\
+\midrule
+Geometric Mean Relative WIS & 0.141 & 0.076 & 0.000 & 0.059 & 0.008\\
+Mean Relative WIS & 0.186 & 0.137 & 0.002 & 0.074 & 0.064\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
-%\bibliography{../../common/covidcast.bib,pnas-materials/pnas-sample.bib}
+\bibliography{../../common/covidcast.bib}
 

### 2021-09-24 forecast/code/deprecated/revision-draft.R

@@ -1,16 +1,12 @@
-# Cumulative mean prediction error plot -----------------------------------
+# Cumulative sum/mean prediction error plot -----------------------------------
 
-tt <- fcasts_honest %>%
+cumulatives <- fcasts_honest %>%
   group_by(forecaster, forecast_date) %>%
   summarise(w = sum(wis) / sum(strawman_wis)) %>%
   arrange(forecast_date) %>%
-  mutate(n = n(), w = cummean(w))
-
-tt <- fcasts_honest %>%
-  group_by(forecaster, forecast_date) %>%
-  summarise(w = sum(wis) / sum(strawman_wis)) %>%
-  arrange(forecast_date) %>%
-  mutate(n = n(), w = cumsum(w))
+  mutate(`Cumulative Mean` = cummean(w),
+         `Cumulative Sum` = cumsum(w),
+         `14 day trailing average` = RcppRoll::roll_meanl(w, n = 14L))
 
 
 fcast_colors <- c(RColorBrewer::brewer.pal(5, "Set1"), "#000000")
@@ -18,15 +14,15 @@ names(fcast_colors) <- c("CHNG-CLI", "CHNG-COVID", "CTIS-CLIIC", "DV-CLI",
                          "Google-AA", "AR")
 
 
-ggplot(tt %>% filter(forecast_date < "2021-01-01"), 
-       aes(forecast_date, w, color = forecaster)) + 
-  geom_line() +
+ggplot(cumulatives %>% filter(forecast_date < "2021-01-01"), 
+       aes(forecast_date, color = forecaster)) + 
+  geom_line(aes(y = `14 day trailing average`)) +
   geom_hline(yintercept = 1) +
   xlab("forecast date") +
-  ylab("cummean(sum(wis_forecaster[t]) /\n sum(wis_strawman[t]))") +
   scale_color_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
   theme_bw() +
   theme(legend.position = "bottom", legend.title = element_blank())
+
 
 
 # Sign test ---------------------------------------------------------------
@@ -37,7 +33,7 @@ st <- fcasts_honest %>%
   mutate(wis = wis / strawman_wis) %>%
   select(forecaster, geo_value, ahead, forecast_date, wis) %>%
   pivot_wider(names_from = forecaster, values_from = wis) %>%
-  mutate(across(AR:`Google-AA`, ~ AR -.x)) %>%
[…]

### 2021-09-24 forecast/code/figures/eval_funs.R

@@ -15,7 +15,15 @@ GeoMean <- function(x) exp(mean(log(x), na.rm = TRUE))
 pct_change <- function(x, n = 14, col_name = "pct_change"){
   # courtesy of Alden
   dt <- c(0,-n)
-  y = covidcast:::apply_shifts(x,dt) %>%
+  # y = covidcast:::apply_shifts(x,dt) %>%  clobbered in recent upgrade
+  x = x %>%
+    dplyr::group_by(geo_value) %>%
+    dplyr::arrange(time_value)
+  for (n in dt) {
+    varname <- sprintf("value%+d", n)
+    x <- x %>% dplyr::mutate(!!varname := covidcast:::shift(value, n))
+  }
+  y <- dplyr::select(x, -value) %>%
     rename(value = tidyselect::matches("value\\+"),
            lag_value = tidyselect::matches("value\\-")) %>%
     mutate(`:=`(!!col_name,(value - lag_value) / lag_value))

### 2021-09-27 forecast/paper/response-to-reviewers.tex
more changes for PR
* added text for the cumulative plot in supplement
* finalized response text (for comments)
* fixed a few figs in the supplement
* made up some spatial error plots.

\documentclass[11pt]{article}

\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{microtype}

\usepackage{newtxmath}
\usepackage{newtxtext}

\usepackage{xcolor}
\definecolor{cobalt}{rgb}{0.0, 0.28, 0.67}

\newenvironment{resp}{\begin{quote}\color{cobalt}}{\end{quote}}
\newcommand{\sresp}[1]{\textcolor{cobalt}{#1}}

\setlength{\parindent}{0in}
\setlength{\parskip}{11pt}

\title{Response to peer reviews}
\author{}
\date{\today}

\begin{document}
\maketitle

\section*{Editor}

  Reviews are generally supportive, but bring up several issues that need to be
  addressed. Once addressed, please submit the revised manuscript with an
  item-by-item reply/rebuttal.

\begin{resp}
  We thank the editor and reviewers for their comments and the opportunity to make
  revisions. We have made a number of revisions, described in detail below.
  We believe these changes have strengthened the manuscript, and describe them in
  more detail below.
  We quote below the Reviewer comment and provide our response in this color.
\end{resp}

\section*{Reviewer 1}

  1. In their presentation of of evaluation metrics they do not discuss
  stability. For instance in [7] does the choice of the weights, using $\tau$ for
  $x\geq 0$ rather than some monotone function say some power $\tau^\alpha$ make a
  difference? Since as they indicate the measure is used for fitting as well as
  ascertainment, some stability investigation seems appropriate. I suspect here
  is no great effect but...

\begin{resp}
  The reviewer is correct: the choice of $\tau$ rather than a monotone function of
  $\tau$ will make a difference. But it seems that we should have been more clear
  in our description of WIS. The weights used here are those used to evaluate
  forecasts submitted to the COVID-19 Forecast Hub and the CDC. More broadly,
  this is a standard metric in the forecasting community (just as Mean Squared
  Forecast Error is standard), rather than an ``author decision''.  In principle,
  we would expect that different weights may alter the conclusions, though
  likely not substantively. Applying a monotone function to tau only effectively
  changes the quantile of interest, resulting in a mismatch between coverage you
  tried to get and the coverage WIS is evaluating. Applying the function to $\tau$
  and $(1-\tau)$ results in an asymmetry, penalizing forecasts that miss on one
  side more than on the other.

  $\bullet$ We have added a sentence in the Evaluation Metrics section emphasizing that
  this metric is standard.
\end{resp}

  2. They do not discuss the possible effects of vaccination rates and other
  factors, though this may be one of the areas of localization they propose to
  investigate.

\begin{resp}
  During the period discussed in the main paper, vaccinations hadn’t really
  taken off. On Dec 31 2000, when the main evaluation period ends,  according to
  CDC data, about 1\% of the population 
  had received 1 dose. That said, we should probably be more specific in the
  paragraph on lines 31-40: our goal is to use very simple models that do not
  account for extra information (vaccinations but also strains of the virus,
  ``super-spreader'' events, etc.).

  $\bullet$ We have added a sentence around lines 30--40 emphasizing the choice
  to ignore such factors.

  $\bullet$ We also added some text to the Prediction Period section stating
  that our end 
  date (31 December) coincides with the introduction of vaccinations.
\end{resp}

  3. This points to what is perhaps a more major issue. Can they identify in
  quantitative terms the consequences of the additional improvements in
  forecasting provided by these methods in terms of gains in resources available
  to deal with the epidemic? Unfortunately as they state, in the crucial periods
  of sharply rising cases, the additional benefit seems least.

\begin{resp}
  This is a good point, and we have made a few modifications described in more
  detail below (Reviewer 2, Point 3).
\end{resp}

  4. Despite the fact that a detailed description of terms in the text is
  impossible, it would I think be helpful to say if there is any significance to
  the names they associate with their first 2 and 4th indicators.

\begin{resp}
  It seems that the Reviewer is referring to CHNG-COVID and DV-CLI. These are
  described in the Methods Section, Signals and Locations Subsection. To be more
  specific, Change 
  Healthcare is a large healthcare insurance claims processor. The difference
  between ``-CLI'' and ``-COVID'' is that the first measures insurance
  claims related to symptoms
  consistent with COVID while the second requires a medical diagnosis (through
  testing or presumed positive by the medical provider). The ``DV'' modifier
  corresponds to different claims processors (not Change Healthcare) who wish to
  remain anonymous.

  TODO Do we want to alter the text at all?
\end{resp}

  5. What are Change Healthcare claims..as opposed to others?

\begin{resp}
  Hopefully the discussion to point 4 above has cleared up any confusion.
\end{resp}

\section*{Reviewer 2}
\begin{quote}
  \emph{(items numbered by the editor for convenience of reference)}
\end{quote}

\subsection*{Statistical or Methodological Comments:}

1. Is this really a stationary phenomenon? Over the timerange extending from
pre-lockdown to today, it certainly isn't. The method in this paper of
presenting model prediction errors makes it impossible to see this. People in
econometrics look at cumsums of normalized prediction errors. You can see change
points and trends very easily.

\begin{resp}
  There is little reason to believe that this data is stationary. The models
  however are trained and evaluated over reasonably short periods (2-3 months),
  over which time the behavior may well be stationary. We have added a section
  to the appendix that uses the cumulative mean as suggested (the sum is
  difficult to see due to the scaling). With some exceptions at the beginning of
  the evaluation period, the relative ordering remains fairly consistent. The AR
  model is easily the worst.

  TODO I have added the code and a caption for the figure to the Supplement, but
  no corresponding 
  text. It seems a fairly minor addition, not necessarily deserving its own
  section. 
\end{resp}

2. Are the results in this paper statistically significant? I see no discussion
of this question, which seems unbelievable for me to be saying, given the
authorship. And yet here I am.

\begin{resp}
  This is an important point that we should have addressed more fully.
  Conventional methods for assessing statistical significance that we explored
  were overconfident, to the point of being unbelievable. To a large degree,
  this is a ``huge n, tiny p'' problem so even bootstrap analyses result in
  extremely narrow intervals. Essentially, ``everything is significant''.

  We also have strong reasons for not using model-based approaches (one being the
  nonstationarity the Reviewer mentions above, but also the belief that these
  models are wrong). We felt that the related investigation in 2B was more
  pertinent.

  To respond more directly:

  $\bullet$ We have added a section to the Supplement (and point
  to it around 
  line 85 in the main text), that performs a Sign Test for differences in
  accuracy. P-values there tend to cluster near zero suggesting again that these
  improvements are statistically significant.

  $\bullet$ We also use a Diebold-Mariano test
  for comparing forecast accuracy. The results for this test are more mixed,
  likely due to nonstationarity the Reviewer mentions above (the indicators are
  more helpful in some periods than others, see also the upswing/downswing
  discussion). 
\end{resp}

3. Are the results in this paper practically significant? I see no discussion of
this question, which seems unbelievable for me to be saying, given that the
authors refuse to consider statistical significance. And yet here I am.

\begin{resp}
  See also Reviewer 1 point 3. And the general comment below (Point 8).
  Practically, on average, if our goal is to predict 11 days ahead, when the AR
  model is at its best, CHNG-CLI
  doesn’t help (at least with these simple models). However, CHNG-COVID,
  DV-CLI, and Google-AA ``buy'' 4 extra days for the 
  same accuracy. CTIS-CLIIC ``buys'' about 5 extra days. What we mean here is
  that if we have a certain tolerance for inaccuracy, adding these signals increases the
  horizon at which we can predict within our tolerance. As for the
  practical significance of 4-5 extra days, this would of course depend on the use case.  However, cases grow
  exponentially, so making public policy changes 15 or 16 days ahead rather than
  11 can mean real gains. There also seems to be a fairly constant mapping
  between cases and future hospitalizations. Giving
  hospitals a few extra days to accumulate supplies and find personnel is
  meaningful. For predicting hotspots, we see similar results (gains of a few
  days) though with different indicators. 
\end{resp}

4. For the cumsum plots I mentioned above, and other related techniques, we can
plot all the model prediction error cumsum curves versus time and thereby
compare different models for size, nonstationarity and significance of
prediction improvement. The plots that have been presented in Figure 4 say do
none of those tasks.

\begin{resp}
  TODO Any ideas here? It's not clear to me that this is different from above.
\end{resp}

5. The idea of using Asof data is of course important but I have known about it
for twenty years and practiced it routinely; it has apparently been in routine
use in econometrics for 40 years.

\begin{resp}
  We appreciate the reviewer's perspective on the ubiquity of Asof data in
  econometrics. But the epidemic modelling community is less aware. And it
  has a real and meaningful importance for model selection going forward. We
  hope that this paper will lead to better behaviour in this community.
\end{resp}

\subsection*{Comments on Significance Statement:}

6. I don't know what ``all provide a nontrivial boost in accuracy'', really
means. I guess that the general audience won't either.

\begin{resp}
  We have changed the clause to ``all provide a nontrivial improvement in
  forecast accuracy''
\end{resp}

7. The sentence following is missing a word: depends ***on*** the pandemic...

\begin{resp}
  Fixed. Thank you for catching this.
\end{resp}

\subsection*{General Comments: }

8. In my understanding there are some standard models which have been relied
upon I guess millions of times during the pandemic. For example IHME's model.
The authors say a sentence or so about their choice of baseline, saying what
seems to me quite vague about the relation between their baseline and the
heavily used existing models. Wouldn't it be better to identify a specific
widely-used model explicitly as a baseline and then develop measures of
improvement over that specific widely used model? Also, might the authors
explain how the improvements observed by the authors stack up in terms of
concrete differences over the widely used models that people have been using.
Are there major misses that have taken place and could have been avoided?

\begin{resp}
  We have tried to focus mainly on simple time series models in this paper
  rather than comparing to other more elaborate models. A careful
  (out-of-sample) comparison of forecasting performance for state-of-the-art
  models can be found in reference 6 (Cramer et al.). That paper focuses on
  forecasting deaths rather than cases and at the state-level rather than
  disaggregated HRRs. However, their ranking (Figure 2) puts the
  ``COVIDhub-Baseline'' about 10th out of 27 models. As we discuss in the
  manuscript, that baseline is 
  the
  same as the baseline we use for forecasting in this paper: it propagates today
  forward and resamples the residuals to create quantiles (we have expanded the
  description on page 6 to be more explicit). All our ``simplistic'' models beat
  this baseline (the AR(3) does as well). For comparison, IHME (along with 17 other
  ``state-of-the-art'' models) does not. We would argue that using any of the
  models in this paper would be preferable to relying on many of the
  ``widely-used'' models if the goal is forecast accuracy.

  $\bullet$ We have added a sentence to the introduction (around line 40) that
  describes the performance of simple models relative to widely-used models as
  shown in Cramer et al.

  $\bullet$ We have expanded our description of our baseline model (around line
  457). In particular, we emphasize that it is the same as that used by the
  COVID-19 Forecast Hub. 
\end{resp}

\end{document}

### 2021-09-27 forecast/paper/paper.tex
more changes for PR
* added text for the cumulative plot in supplement
* finalized response text (for comments)
* fixed a few figs in the supplement
* made up some spatial error plots.

@@
   claims, self-reported symptoms through online surveys, and COVID-related
-  internet search activity all provide a nontrivial boost in accuracy, though
-  their utility generally depends the pandemic's local dynamics (upswing,
+  internet search activity
+  % all provide a nontrivial boost in accuracy,
+  all provide a nontrivial improvement in forecast accuracy,
+  though
+  their utility generally depends on the pandemic's local dynamics (upswing,
   downswing, or flat period) at prediction time.
@@
 correction of outliers/anomalies in the data, and inclusion of regularization or
-nonlinearities. That said, analyses of forecasts submitted to the COVID-19
+nonlinearities.
+% Addition in response to Reviewer 1
+For the same reason, we do not account for any environmental factors that may
+well aid in forecasting---age structure, variants, holidays,
+reporting patterns, vaccination, public health measures, etc.
+% End addition
+That said, analyses of forecasts submitted to the COVID-19
 Forecast Hub \cite{ForecastHub} by a large community of modelers have shown that
 simple, robust models have consistently been among the best-performing over the
-pandemic \cite{Cramer:2021}, including time series models similar to those we
+pandemic, including time series models similar to those we
 consider in what follows.
+% Added in response to Reviewer 2
+\cite{Cramer:2021} investigates the performance of submitted forecasts for
+state-level deaths and shows that time series models perform among the top 5.
+Our internal evaluations suggests that the same is true (at least, if not to a
+greater extent) for case forecasting.
 
@@
 conclusions with respect to numerous modes of analysis.
+% Addition upon review
+We have included a section in the Supplement that examines two assumption-lean
+methods for assessing the statistical significance of our main results.
 
@@
 when the survey was first launched). The end date was chosen again with a 
-consideration to align both tasks as best as possible, and because few hotspots
-exist post December 31, 2020, due to the general and gradual decline of the
-pandemic in 2021.
+consideration to align both tasks as best as possible,
+% added to emphasize 
+and to coincide with the
+introduction of vaccinations.
+% Removed: this is no longer true unfortunately.
+% because few hotspots
+% exist post December 31, 2020, due to the general and gradual decline of the
+% pandemic in 2021.
 
@@
 the main evaluation metric used in the COVID-19 Forecast Hub.
+% Addition in response to Reviewer 1
+More broadly, the specific form of WIS used here is a standard metric in the
+forecasting community for evaluating probabilistic interval forecasts, just as
+Mean Squared Forecast Error is standard for point forecasts.
 
@@
 dates.  Shown is the average WIS for each forecast model divided by that from a
-baseline model, which is basically a flat-line forecaster (its median forecast
-for $Y_{\ell,t+a}$ is always $Y_{\ell,t}$, with predicted quantiles defined
-around this based on historical variation). This is the same baseline model as
+% change since review
+baseline model.
+The baseline model is a flat-line forecaster which forms its median forecast by
+using the most recent value $Y_{\ell,t}$ as the median forecast for all aheads
+$Y_{\ell,t+a}$ . To create
+the quantiles, we use the empirical quantiles of the residuals from this
+forecast over the available history. 
+% old version
+% baseline model,
+% which is basically a flat-line forecaster (its median forecast
+% for $Y_{\ell,t+a}$ is always $Y_{\ell,t}$, with predicted quantiles defined
+% around this based on historical variation).
+This is the same baseline model as
 in the COVID-19 Forecast Hub. Here, we use the baseline model in order to scale

### 2021-09-27 forecast/paper/supplement-text.tex
more changes for PR
* added text for the cumulative plot in supplement
* finalized response text (for comments)
* fixed a few figs in the supplement
* made up some spatial error plots.

@@
 common for there to be many fewer hotspots in training.
+
+\section{Examining accumulated error over time and space}
+
+Following the suggestion of an anonymous reviewer, we investigate two
+disaggregated versions of 
+the main forecasting result shown in Figure 3 on the manuscript. The first
+(Figure \ref{fig:cumulative-mean})
+displays the cumulative error (summed over all horizons) of each forecaster
+divided by the cumulative error 
+of the baseline.
+This perspective should illustrate how
+the models perform over time, drawing attention to any nonstationary behavior.
+During the initial increase in cases in July 2020, CTIS-CLIIC and Google-AA gain
+a lot relative to the AR model. All the indicators do much better than the AR
+model during the following downturn (the ebb of the second wave). The AR model
+actually improves over the indicators in October 2020, before losing a bit in
+late November.
+
+Figure \ref{fig:errs-in-space} examines the spatial behavior over the entire
+period of the indicator-assisted models relative to the AR. For ease of
+comparison, we show the percent improvement in each HRR. Negative numbers (blue)
+mean that the indicator helped while positives (red) mean that the indicator
+hurt forecasting performance. The clear pattern is that in most HRRs, the
+indicators improved performance, though usually by small amounts (2.5\%--10\%).
+In some isolated HRRs, performance was markedly worse, though there does not
+appear to be any particular pattern to these locations.

### 2021-09-27 forecast/paper/supplement.Rmd
more changes for PR
* added text for the cumulative plot in supplement
* finalized response text (for comments)
* fixed a few figs in the supplement
* made up some spatial error plots.

@@
            stat = "identity") +
-  scale_fill_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
+  scale_fill_manual(values = fcast_colors[1:5], guide = guide_legend(nrow = 1)) +
   scale_y_continuous() +
@@
            stat = "identity") +
-  scale_fill_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
+  scale_fill_manual(values = fcast_colors[1:5], guide = guide_legend(nrow = 1)) +
   scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
@@
            stat = "identity") +
-  scale_fill_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
+  scale_fill_manual(values = fcast_colors[1:5], guide = guide_legend(nrow = 1)) +
   theme_bw() +
@@
   geom_hline(yintercept = 0) +
-  scale_fill_manual(values = fcast_colors) +
+  scale_fill_manual(values = fcast_colors[1:5], guide = guide_legend(nrow = 1)) +
   theme_bw() +
@@
   geom_hline(yintercept = 0) +
-  scale_fill_manual(values = fcast_colors) +
+  scale_fill_manual(values = fcast_colors[1:5]) +
   theme_bw() +
@@
 
+<!-- Cumulative Mean -->
+
+```{r cumulative-mean, fig.cap="This figure shows the cumulative sum of WIS for each forecaster divided by the cumulative sum of WIS for the baseline model. The shaded background shows" }
+cumulatives <- fcasts_honest %>%
+  group_by(forecaster, forecast_date) %>%
+  summarise(mw = Mean(wis),
+            smw = Mean(strawman_wis),
+            gmw = GeoMean(wis / strawman_wis)) %>%
+  arrange(forecast_date) %>%
+  mutate(`Cumulative Mean` = cummean(mw) / cummean(smw),
+         `Cumulative Sum` = cumsum(mw) / cumsum(smw),
+         `Cumulative Geo Mean` = exp(cummean(log(gmw))),
+         `GM Regret` = cumsum(gmw) - cumsum(smw/smw),
+         `AM Regret` = cumsum(mw) - cumsum(smw),
+         `14 day trailing average` = RcppRoll::roll_meanl(mw / smw, n = 14L))
+
+us_cases <- covidcast::covidcast_signal(
+  "jhu-csse", "confirmed_7dav_incidence_num", "2020-03-01", "2021-01-01",
+  geo_type = "nation") %>% as_tibble() %>%
+  mutate(value = (value / max(value)) * .25 +  .75) %>%
+  filter(time_value > "2020-06-01")
+
+ggplot(cumulatives %>% filter(forecast_date < "2021-01-01"), 
+       aes(forecast_date, color = forecaster)) + 
+  geom_line(aes(y = `Cumulative Sum`)) +
+  annotate(x = us_cases$time_value, ymax = us_cases$value, ymin = -Inf,
+           geom = "ribbon", alpha = .1) +
+  geom_hline(yintercept = 1) +
+  xlab("forecast date") +
+  scale_color_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
+  theme_bw() +
+  theme(legend.position = "bottom", legend.title = element_blank())
+```
+
+```{r errs-in-space, fig.cap="Percent improvement in WIS relative to the AR forecaster by region (negative numbers indicate improved performance, positives indicate worsening)."}
+# fig.height = 9.6, fig.width = 6, out.width = "5in", out.height="8in", 
+by_hrr <- fcasts_honest %>% 
+  filter(period != "jm") %>%
+  group_by(forecaster, geo_value) %>% 
+  summarize(value = GeoMean(wis)) %>%
+  mutate(time_value = Sys.Date()) %>%
+  ungroup()
+by_hrr <- left_join(
+  by_hrr %>% filter(forecaster != "AR"), 
+  by_hrr %>% 
+    filter(forecaster == "AR") %>% 
+    rename(AR = value) %>%
+    select(geo_value, AR)) %>%
+  mutate(value = (value - AR) / AR * 100) %>%
+  group_by(forecaster) %>%
+  group_split()
+
+cc_sigs <- purrr::map(by_hrr,
+                   ~ as.covidcast_signal(
+                     .x, signal = "Improvement relative to AR",
+                     data_source = .x$forecaster[1],
+                     geo_type = "hrr"))
+br <- c(2.5, 5, 10, 25, 50)
+br <- c(-rev(br), 0, br)
+colours <- brewer_pal(palette = "RdBu")
+
+maps <- purrr::map(
+  cc_sigs, 
+  ~plot(.x, title = .x$forecaster[1], 
+        choro_params = list(breaks = br), 
+        choro_col = rev(colours(11))) +
+    theme(legend.position = "none"))
+
+map_plot <- cowplot::plot_grid(plotlist = maps, ncol = 3)
+map_leg <- get_legend(maps[[1]] + theme(legend.position = "bottom"))
+cowplot::plot_grid(map_plot, map_leg, ncol = 1, rel_heights = c(1, .1))
+```
+
+\clearpage
+
 <!-- All code below is deprecated -->

### 2021-09-27 forecast/paper/supplement.tex
more changes for PR
* added text for the cumulative plot in supplement
* finalized response text (for comments)
* fixed a few figs in the supplement
* made up some spatial error plots.

@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/cumulative-mean-1}
+
+}
+
+\caption{This figure shows the cumulative sum of WIS for each forecaster divided by the cumulative sum of WIS for the baseline model. The shaded background shows}\label{fig:cumulative-mean}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/errs-in-space-1}
+
+}
+
+\caption{Percent improvement in WIS relative to the AR forecaster by region (negative numbers indicate improved performance, positives indicate worsening).}\label{fig:errs-in-space}
+\end{figure}
+
+\clearpage
+
 \FloatBarrier

### 2021-09-27 forecast/code/deprecated/revision-draft.R

@@ -2,21 +2,28 @@
 
 cumulatives <- fcasts_honest %>%
   group_by(forecaster, forecast_date) %>%
-  summarise(w = sum(wis) / sum(strawman_wis)) %>%
+  summarise(mw = Mean(wis),
+            smw = Mean(strawman_wis),
+            gmw = GeoMean(wis / strawman_wis)) %>%
   arrange(forecast_date) %>%
-  mutate(`Cumulative Mean` = cummean(w),
-         `Cumulative Sum` = cumsum(w),
-         `14 day trailing average` = RcppRoll::roll_meanl(w, n = 14L))
+  mutate(`Cumulative Mean` = cummean(mw) / cummean(smw),
+         `Cumulative Sum` = cumsum(mw),
+         `Cumulative Geo Mean` = exp(cummean(log(gmw))),
+         `GM Regret` = cumsum(gmw) - cumsum(smw/smw),
+         `AM Regret` = cumsum(mw) - cumsum(smw),
+         `14 day trailing average` = RcppRoll::roll_meanl(mw / smw, n = 14L))
 
-
+  
+  
 fcast_colors <- c(RColorBrewer::brewer.pal(5, "Set1"), "#000000")
 names(fcast_colors) <- c("CHNG-CLI", "CHNG-COVID", "CTIS-CLIIC", "DV-CLI",
                          "Google-AA", "AR")
 
 
+
 ggplot(cumulatives %>% filter(forecast_date < "2021-01-01"), 
        aes(forecast_date, color = forecaster)) + 
-  geom_line(aes(y = `14 day trailing average`)) +
+  geom_line(aes(y = `Cumulative Geo Mean`)) +
   geom_hline(yintercept = 1) +
   xlab("forecast date") +
   scale_color_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
@@ -95,18 +102,21 @@ knitr::kable(dms %>% pivot_wider(names_from = forecaster, values_from = dm), dig
 
 days_gained <- function(forecaster, reference, aheads, ahead = 14L, 
                         smaller_is_better = TRUE) {
+  n <- length(forecaster)
   # check for non-monotonicity
   if (!any(aheads == ahead)) return(NA)
   if (smaller_is_better) {
-    if (is.unsorted(forecaster) || is.unsorted(reference)) return(NA) 
+    if (is.unsorted(forecaster)) return(NA) 
   } else {
[…]

### 2021-09-27 indicators/code/indicator-paper-figures.Rmd
Remove code for old SafeGraph mobility map
Replaced with the vaccination acceptance/uptake graph.

@@
 
-## Now plot the each "as of" time series curve, faceted by state
+## Now plot each "as of" time series curve, faceted by state
 dv_as_of %>%
@@
 ## Figure 6
-
-Percentage point increase in SafeGraph estimates of the percentage of people
-staying completely at home, based on aggregated mobile device data, between
-March 1, 2020 and April 15, 2020.
-
-```{r figure-6, message=FALSE, cache=TRUE}
-begin_date <- "2020-03-01"
-end_date <- "2020-04-15"
-
-begin_data <- covidcast_signal("safegraph", "completely_home_prop_7dav",
-                               begin_date, begin_date, geo_type = "state")
-end_data <- covidcast_signal("safegraph", "completely_home_prop_7dav",
-                             end_date, end_date, geo_type = "state")
-
-diffs <- begin_data %>%
-  inner_join(end_data, by = "geo_value", suffix = c(".begin", ".end")) %>%
-  mutate(value = 100 * (value.end - value.begin)) %>%
-  select(data_source = data_source.end, signal = signal.end, geo_value, value, issue = issue.end) %>%
-  mutate(time_value = end_date) %>%
-  as.data.frame %>%
-  as.covidcast_signal(signal = "safegraph-change",
-                      geo_type = "state")
-
-plot(diffs, plot_type = "choro", range = c(10, 30),
-     title = "", choro_params = list(legend_n = 5))
-
-ggsave("../paper/fig/mobility-drop.pdf", width = 5, height = 4)
-```
 

### 2021-09-27 indicators/code/indicator-paper-figures.Rmd
New backfill figure

@@
 library(dplyr)
+library(tidyr)
 library(purrr)
@@
 
+Amount of backfill of key signals from various sources, for comparison:
+
+```{r figure-6, message=FALSE, cache=TRUE}
+SOURCES <- c("chng", "fb-survey", "jhu-csse", "jhu-csse")
+SIGNALS <- c("smoothed_outpatient_cli", "smoothed_whh_cmnty_cli",
+             "confirmed_incidence_prop", "deaths_incidence_prop")
+NAMES <- c("lag", "CHNG-CLI", "CTIS-CLI-in-community", "Cases", "Deaths")
+
+calc_pct_abs_diff <- function(source, signal, start_date, end_date,
+                              final_lag, lags) {
+  pctile_abs_change <- numeric(length(lags))
+
+  dates <- seq.Date(as.Date(start_date),
+                    as.Date(end_date),
+                    by = "1 day")
+
+  # finalized value
+  df_final <- covidcast_signal(source, signal, start_date, end_date,
+                               as_of = end_date + final_lag,
+                               geo_type = "state") %>%
+    select(geo_value, time_value, value)
+
+  for (ii in seq_along(lags)) {
+    lag <- lags[ii]
+    errs <- c()
+
+    for (jj in seq_along(dates)) {
+      date <- dates[jj]
+
+      df_prelim <- covidcast_signal(source, signal, date, date,
+                                    as_of = date + lag,
+                                    geo_type = "state")
+      if (nrow(df_prelim) == 0) {
+        next
+      }
+
+      df_prelim <- df_prelim %>%
+        select(geo_value, time_value, value) %>%
+        mutate(lag = lag)
+
+      errs_this_date <- df_final %>%
+        inner_join(df_prelim,
+                   by = c("geo_value", "time_value"),
+                   suffix = c(".final", ".prelim")) %>%
+        mutate(abs_diff = abs(value.prelim - value.final) / value.final) %>%
+        pull(abs_diff)
+
+      errs <- c(errs, errs_this_date)
+    }
+
+    pctile_abs_change[ii] <- quantile(errs, probs = 0.95, na.rm = TRUE,
+                                      names = FALSE)
+  }
+
+  return(pctile_abs_change)
+}
+
+format_df <- function(df) {
+  names(df) <- NAMES
+
+  df <- df %>%
+    pivot_longer(!lag, names_to = "signal", values_to = "value")
+
+  return(df)
+}
+
+start_date <- as.Date("2020-10-15")
+end_date <- as.Date("2021-04-15")
+final_lag <- 120
+lags <- c(90, 70, 50, 30, 10)
+
+pctile_df <- data.frame(lag = lags)
+
+for (ii in seq_along(SOURCES)) {
+  source <- SOURCES[ii]
+  signal <- SIGNALS[ii]
+
+  pctile_df[[signal]] <- calc_pct_abs_diff(source, signal, start_date, end_date,
+                                           final_lag, lags)
+}
+
+pctile_df <- format_df(pctile_df)
+
+pctile_df %>%
+  ggplot(aes(x = lag, y = value, group = signal, color = signal)) +
+  geom_line() +
+  geom_text(data = filter(pctile_df, lag == min(lags)),
+            aes(x = lag, y = value, label = signal, hjust = "left"),
+            inherit.aes = FALSE, nudge_y = 0.015) +
+  labs(x = "Days since initial issue",
+       y = "95th percentile of absolute percent error",
+       color = "Signal") +
+  scale_y_continuous(labels = label_percent()) +
+  scale_x_continuous(breaks = lags) +
+  theme_bw() +
+  theme(legend.position = "none")
+
+ggsave("../paper/fig/backfill-compare.pdf", width = 5, height = 3.5)
+```
+
+## Figure 7
+
 County-level vaccination data obtained [from the
@@
 
-```{r figure-6-redux, message=FALSE, cache=TRUE}
+```{r figure-7, message=FALSE, cache=TRUE}
 cdc <- read.csv("COVID-19_Vaccinations_in_the_United_States_County.csv") %>%

### 2021-09-27 indicators/paper/paper.tex
New backfill figure

@@
 to revision as death certificates are audited and backlogs cleared, resulting in
-thousands of cases and deaths being added or removed. This process is much more
-difficult to predict, and thus claims data and other sources may be a useful
-stand-in while public health reports are aggregated and corrected.
+thousands of cases and deaths being added or removed.
+Figure~\ref{fig:backfill-compare} shows the amount of revision for several
+signals, in terms of the 95th percentile of relative change between initial
+estimates and revised estimates issued later; this shows that reported deaths
+can have large revisions comparable to revisions in claims-based signals. This
+backfill is much more difficult to predict, and thus claims data and other
+sources may be a useful stand-in while public health reports are aggregated and
+corrected.
 
-To reiterate a previous point, when training and validating forecast models (on
-historical data), users will want to use data that was known \textit{as of} the
-forecast date, not revised versions that only became available much later. The
-COVIDcast API makes all historical versions available and easily accessible for
-this purpose; and this feature plays a prominent role in our own analysis of
-forecasting and hotspot prediction models appearing in a companion paper.
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{fig/backfill-compare.pdf}
+  \caption{95th percentile relative changes between initial values of key
+    signals, compared to the final values reported much later. Calculated using
+    each day's state-level estimates for the six months between October 15, 2020
+    and April 15, 2021; final values were those reported 120 days after the end
+    of this period. Even officially reported case and death data can have large
+    revisions 30--60 days after initial reporting; much of this is driven by
+    individual large revisions affecting specific states and dates, rather than
+    systematic changes affecting all states and dates.}
+  \label{fig:backfill-compare}
+\end{figure}
+
+To reiterate a previous point, when training and validating forecast models that
+use signals (on historical data), users will want to use data that was known
+\textit{as of} the forecast date, not revised versions that only became
+available much later. The COVIDcast API makes all historical versions available
+and easily accessible for this purpose; and this feature plays a prominent role
+in our own analysis of forecasting and hotspot prediction models appearing in a
+companion paper. Revisions to forecast targets, such as cases and deaths, also
+complicate the evaluation of forecast models: comparisons of forecasts to the
+ground truth may shift when the ground truth is revised weeks later. Only by
+systematically tracking revisions can this effect be monitored.
 

### 2021-09-27 indicators/paper/responses-to-reviews.tex
Move point back to review responses and finish them off

@@
 now shows how data sources in the API could have aided in planning for the COVID
-vaccination campaign.
-
-TODO any additional changes?
+vaccination campaign. We have also introduced a new figure (discussed below) to
+make the value of revision data more clear, and adjusted language throughout.
 
@@
 before the case reports were, depending on the time this occurred during the
-pandemic and the state's reporting practices. We have expanded Section 2.B to
-make this point clearer.
+pandemic and the state's reporting practices. Meanwhile, signals such as
+CTIS-CLI-in-community can be reported within 1--2 days of the survey responses.
+High correlations with case and death data hence indicate a high correlation
+with a case value that may have only been officially reported days after these
+signals.
+
+But this point is somewhat subtle; ideally, one would like to correlate our
+surveys against the number of new infections occurring on a date, rather than
+the number of cases reported on that date. These could be very different, since
+infections occur days or weeks before cases are reported, but ground-truth
+infection data is unavailable.
+
+Instead, we have added a new figure to illustrate a related point. Because we
+track revisions of signals, we can see how often officially reported case and
+death data is revised, and how long these revisions can occur after initial
+reporting. The new figure (now Figure 6) shows that case and death data can be
+subject to quite large revisions even 30--60 days after initial reporting,
+making it important to have supplementary data sources to check against.
 
@@
 We have extended the Supplementary Information to include more detailed
-comparisons between our signals and reported COVID hospitalizations, including
-comparisons of correlations. These show broadly similar results as the COVID
-case analyses. Section 2.B now points to the Supplement for interested readers
-to examine these analyses.
+comparisons between our signals and reported COVID hospitalizations (as compiled
+by the Department of Health and Human Services), including comparisons of
+correlations. These show broadly similar results as the COVID case analyses.
+Section 2.B now points to the Supplement for interested readers to examine these
+analyses.
 
@@
 corrections, their magnitude, or the effects they'd have on policy decisions.
+The new Figure 6 illustrates this by showing the magnitude of revisions to case
+and death reports.
 
@@
 We appreciate the interest in additional analyses, and our replacement of Figure
-6 is partly intended to address this concern about choice of case studies. It
-represents an example showing how these data sources can reveal points not
-available in other standard data sources, such as case and death data, and serve
-to aid public health decision-making.
+6 (with the new Figure 7) is partly intended to address this concern about
+choice of case studies. It represents an example showing how these data sources
+can reveal points not available in other standard data sources, such as case and
+death data, and serve to aid public health decision-making.
 
@@
 \end{quote}
-TODO
+We have extended the discussion of revisions, as mentioned above, to give
+another illustration of why having revision information for external data
+sources is valuable. Given the challenges of COVID forecasting, it did not seem
+in-scope for this paper to explore revision information and its impact on
+forecasting in great depth. We have also replaced the example in Section 2.E
+(now Figure 7) to be more topical and illustrate the unique value of the data
+more strongly.
 

### 2021-09-27 indicators/paper/paper.tex
I demand a recount
Move point back to review responses and finish them off

@@
 
-% TODO Recount!
-Altogether, we collect 110 signals from 12 distinct sources, aggregate and
+Altogether, we collect over 170 signals from 12 distinct sources, aggregate and
 process them, and provide them in a common format for public access. This
@@
 CTIS-CLI-in-community signal is a close second, and the CHNG-CLI signal is
-third. There are three noteworthy points:
+third. There are two noteworthy points:
 
@@
   a confirmed case passing through the outpatient system.
-
-\item As we will discuss below in more detail, public health reporting is often
-  subject to delays and backlogs that cause data to be reported with lag.
-  Meanwhile, signals such as CTIS-CLI-in-community can be reported with as
-  little as 1--2 days of lag. These high correlations hence indicate a high
-  correlation with a case value that may have only been officially reported days
-  after these signals.
 \end{itemize}

### 2021-09-29 forecast/paper/response-to-reviewers.tex
Pre-meeting modifications, responding to @jacobbien @huisaddison

@@
   we would expect that different weights may alter the conclusions, though
-  likely not substantively. Applying a monotone function to tau only effectively
-  changes the quantile of interest, resulting in a mismatch between coverage you
+  likely not substantively. Applying a monotone function to $\tau$ only effectively
+  changes the quantile of interest, resulting in a mismatch between the coverage you
   tried to get and the coverage WIS is evaluating. Applying the function to $\tau$
@@
 \begin{resp}
-  It seems that the Reviewer is referring to CHNG-COVID and DV-CLI. These are
+  It seems that the Reviewer is referring to CHNG-CLI, CHNG-COVID and DV-CLI. These are
   described in the Methods Section, Signals and Locations Subsection. To be more
@@
   between ``-CLI'' and ``-COVID'' is that the first measures insurance
-  claims related to symptoms
-  consistent with COVID while the second requires a medical diagnosis (through
+  claims involving symptoms
+  associated with COVID-like-illness  while the second requires a medical
+  diagnosis (through 
   testing or presumed positive by the medical provider). The ``DV'' modifier
@@
 
-  TODO I have added the code and a caption for the figure to the Supplement, but
-  no corresponding 
-  text. It seems a fairly minor addition, not necessarily deserving its own
-  section. 
 \end{resp}
@@
   more helpful in some periods than others, see also the upswing/downswing
-  discussion). 
+  discussion in Section 2C). 
 \end{resp}
@@
 \begin{resp}
-  TODO Any ideas here? It's not clear to me that this is different from above.
+  We added the cumulative-sum plot as described above. Additionally, the
+  Reviewer's point aligns with much of  our motivation for Fig 4. If we were
+  predicting a single location's 
+  time series, then the cum-sum plot would provide all the information.
+  However, we have a separate time series prediction problem for each HRR, and
+  the upswings/downswings occur at different dates in different HRRs, so
+  conditioning on forecast date for the comparison (as is done in a
+  cumulative-sum plot) ends up adding together errors from different phases
+  of the pandemic. In Figure 4, we instead condition on phase of the pandemic,
+  which we think is more meaningful than calendar date. 
 \end{resp}
@@
 \begin{resp}
-  We appreciate the reviewer's perspective on the ubiquity of Asof data in
+  We appreciate the Reviewer's perspective on the ubiquity of Asof data in
   econometrics. But the epidemic modelling community is less aware. And it
   has a real and meaningful importance for model selection going forward. We
-  hope that this paper will lead to better behaviour in this community.
+  hope that this paper will lead to greater appreciation of this issue in this
+  community. 
 \end{resp}
@@
   forecasting deaths rather than cases and at the state-level rather than
-  disaggregated HRRs. However, their ranking (Figure 2) puts the
+  disaggregated HRRs. However, their ranking (Figure 2 of Cramer et al.) puts the
   ``COVIDhub-Baseline'' about 10th out of 27 models. As we discuss in the

### 2021-09-29 forecast/paper/paper.tex
Pre-meeting modifications, responding to @jacobbien @huisaddison

@@
 state-level deaths and shows that time series models perform among the top 5.
-Our internal evaluations suggests that the same is true (at least, if not to a
+Our internal evaluations suggest that the same is true (at least, if not to a
 greater extent) for case forecasting.
@@
   Google search volume for queries that relate to anosmia or ageusia (loss of
-  smell or taste), based on Google's COVID-19 Search Trends data set.  
+  smell or taste), based on Google's COVID-19 Search Trends data
+  set~\cite{Wahltinez2020}.   
 \end{itemize}

### 2021-09-29 forecast/paper/supplement.Rmd
Pre-meeting modifications, responding to @jacobbien @huisaddison

@@
     pivot_wider(names_from = forecaster, values_from = dm), 
-  digits = 3, booktabs = TRUE, caption = "$P$-values for a Diebold-Mariano test for differences in forecast error.
+  digits = 3, booktabs = TRUE, caption = "$P$-values for a one-sided Diebold-Mariano test for improvement in forecast error.
 The test is for the null hypothesis of equal performance against the alternative that the indicator assisted model is better.")
@@
 
-```{r cumulative-mean, fig.cap="This figure shows the cumulative sum of WIS for each forecaster divided by the cumulative sum of WIS for the baseline model. The shaded background shows" }
+```{r cumulative-mean, fig.cap="This figure shows the cumulative sum of WIS for each forecaster divided by the cumulative sum of WIS for the baseline model.
+The shaded background shows total U.S.
+cases as reported by JHU-CSSE." }
 cumulatives <- fcasts_honest %>%

### 2021-09-30 indicators/paper/paper.tex
Cite our companion papers

@@
 In a companion paper, we analyze the utility provided by a core set of the
-indicators in COVID-19 forecasting and hotspot prediction models.  In another
-companion paper, we elaborate on our research group's (Delphi's) large-scale
-public surveys, run in partnership with Facebook and available in aggregate form
-in the COVIDcast API. This paper focuses on the COVID-19 activity indicators
-themselves, describing the data streams, how they are processed and made
-publicly available, and insights that can be gained by combining novel data
-sources with standard public health surveillance datasets.
+indicators in COVID-19 forecasting and hotspot prediction models
+\cite{McDonald:2021}. In another companion paper, we elaborate on our research
+group's (Delphi's) large-scale public surveys, run in partnership with Facebook
+and available in aggregate form in the COVIDcast API \cite{Salomon:2021}. This
+paper focuses on the COVID-19 activity indicators themselves, describing the
+data streams, how they are processed and made publicly available, and insights
+that can be gained by combining novel data sources with standard public health
+surveillance datasets.
 
@@
 daily, and has received over 20 million responses since April 2020
-\cite{DelphiSurvey, Kreuter:2020}. From the surveys, we construct indicators on
+\cite{Salomon:2021, Kreuter:2020}. From the surveys, we construct indicators on
 symptoms, social distancing, vaccination, and other attitudes and behaviors
@@
 vaccination attitudes), and COVID testing signals based on daily surveys
-conducted nationally by Delphi through Facebook \cite{DelphiSurvey,
+conducted nationally by Delphi through Facebook \cite{Salomon:2021,
 Kreuter:2020}. & 2020-04-06 & County$^{**}$ \\
@@
 finalized data yields overly optimistic estimates of true test-time
-performance.
+performance \cite{McDonald:2021}.
 
@@
 be used to improve forecasts of future COVID case trends, as investigated in a
-companion paper.
+companion paper \cite{McDonald:2021}.
 
@@
 in our own analysis of forecasting and hotspot prediction models appearing in a
-companion paper. Revisions to forecast targets, such as cases and deaths, also
-complicate the evaluation of forecast models: comparisons of forecasts to the
-ground truth may shift when the ground truth is revised weeks later. Only by
-systematically tracking revisions can this effect be monitored.
+companion paper \cite{McDonald:2021}. Revisions to forecast targets, such as
+cases and deaths, also complicate the evaluation of forecast models: comparisons
+of forecasts to the ground truth may shift when the ground truth is revised
+weeks later. Only by systematically tracking revisions can this effect be
+monitored.
 
@@
 tasks such as forecasting and nowcasting. Our companion paper on forecasting
-discusses this in more detail.
+discusses this in more detail \cite{McDonald:2021}.
 
@@
 patterns, internet search trends, mask wearing, and vaccine hesitancy, to name
-just a few. (The latter two signals are derived from the COVID-19 Trends and Impact
-Survey; our companion paper on this survey gives a more detailed view of its
-features and capabilities.) These signals have already supported pandemic
-research and  policy-making.
+just a few. (The latter two signals are derived from the COVID-19 Trends and
+Impact Survey; our companion paper on this survey gives a more detailed view of
+its features and capabilities \cite{Salomon:2021}.) These signals have already
+supported pandemic research and policy-making.
 

### 2021-09-30 indicators/code/indicator-paper-figures.Rmd
Mention data license
I haven't committed the data because it's quite large, but if we
deposit the code and data somewhere, it's good to know it's public
domain.
Improve description of Figure 6 (backfill)

@@
 
-```{r figure-6, message=FALSE, cache=TRUE}
+```{r figure-6-data, message=FALSE, cache=TRUE}
 SOURCES <- c("chng", "fb-survey", "jhu-csse", "jhu-csse")
@@
 pctile_df <- format_df(pctile_df)
+```
 
+```{r figure-6, cache=TRUE}
 pctile_df %>%
@@
             inherit.aes = FALSE, nudge_y = 0.015) +
-  labs(x = "Days since initial issue",
+  labs(x = "Lag after events",
        y = "95th percentile of absolute percent error",
@@
 CDC](https://data.cdc.gov/Vaccinations/COVID-19-Vaccinations-in-the-United-States-County/8xkx-amqh).
-Data downloaded on August 2, 2021.
+Data downloaded on August 2, 2021. This data is in the public domain as a United
+States government work.
 

### 2021-09-30 indicators/paper/paper.tex
Improve description of Figure 6 (backfill)

@@
 thousands of cases and deaths being added or removed.
-Figure~\ref{fig:backfill-compare} shows the amount of revision for several
-signals, in terms of the 95th percentile of relative change between initial
-estimates and revised estimates issued later; this shows that reported deaths
-can have large revisions comparable to revisions in claims-based signals. This
-backfill is much more difficult to predict, and thus claims data and other
-sources may be a useful stand-in while public health reports are aggregated and
-corrected.
+Figure~\ref{fig:backfill-compare} shows relative differences between early
+signal values, reported 10--90 days after the underlying events, and later
+versions at least four months later. As the distribution of these revision
+amounts is highly skewed, the figure plots the 95th percentile of relative
+change, showing that reported deaths can have large revisions comparable to
+revisions in claims-based signals. This backfill is much more difficult to
+predict, and thus claims data and other sources may be a useful stand-in while
+public health reports are aggregated and corrected.
 
@@
   \includegraphics[width=\columnwidth]{fig/backfill-compare.pdf}
-  \caption{95th percentile relative changes between initial values of key
-    signals, compared to the final values reported much later. Calculated using
-    each day's state-level estimates for the six months between October 15, 2020
-    and April 15, 2021; final values were those reported 120 days after the end
-    of this period. Even officially reported case and death data can have large
-    revisions 30--60 days after initial reporting; much of this is driven by
-    individual large revisions affecting specific states and dates, rather than
-    systematic changes affecting all states and dates.}
+  \caption{The 95th percentiles of relative changes between early reported
+    values of key signals and final values reported much later. For each date
+    between October 15, 2020 and April 15, 2021, the values for each state
+    reported between 10 and 90 days later are compared to ``final'' versions
+    recorded as of August 13, 2021, 120 days after April 15. Even officially
+    reported case and death data can have large revisions 30--60 days or more
+    after initial reporting; much of this is driven by individual large
+    revisions affecting specific states and dates, rather than systematic
+    changes affecting all states and dates.}
   \label{fig:backfill-compare}

### 2021-09-30 forecast/paper/response-to-reviewers.tex
modifications after the first revision meeting.

@@
 
-  TODO Do we want to alter the text at all?
 \end{resp}
@@
   COVID-19 Forecast Hub.
+
+  $\bullet$ We have added two figures to the supplement along with a section of
+  text (Section 9) that compares our forecasts with state-of-the-art forecasts
+  submitted to the COVID-19 Forecast Hub. The first (Figure S26) overlays the
+  scores from the submissions on top of Figure 3 from the manuscript. We should
+  note that the submitted forecasts are at the state level, while ours are at
+  the HRR level, but once scaled by the baseline, should be roughly comparable.
+  Only those forecasters that submitted at least 300 forecasts (roughly 50
+  states over 6 weeks during the 6 months examined) are shown. From this Figure
+  it is clear that all of the forecasters considered in this paper beat most
+  submissions. 
 \end{resp}

### 2021-09-30 forecast/paper/paper.tex
modifications after the first revision meeting.

@@
 state-level deaths and shows that time series models perform among the top 5.
-Our internal evaluations suggest that the same is true (at least, if not to a
-greater extent) for case forecasting.
+In the Supplement, we compare the time series models investigated in this paper
+to the submitted case forecasts over the same period. The comparison is not
+particularly flattering: the simple models examined here outperform most
+submitted models and approach the accuracy of the COVIDHub Ensemble used by the
+CDC.
 
@@
 % Addition upon review
-We have included a section in the Supplement that examines two assumption-lean
+In the Supplement we examine
+two assumption-lean
 methods for assessing the statistical significance of our main results.

### 2021-09-30 forecast/paper/supplement-text.tex
modifications after the first revision meeting.

@@
 appear to be any particular pattern to these locations.
+
+\section{Comparisons to state-of-the-art forecasters}
+
+Since mid-July 2020, a number of forecasting teams (48 in total) have submitted
+1--4 week ahead state-level case forecasts to the COVID19 Forecast Hub
+\cite{ForecastHub}. To compare those forecasts to those created in this paper,
+we selected only those forecasters who submitted at least 300 forecasts for the
+2-week-ahead target before
+January 1, 2021 (approximately 1 forecast for each state for 6 weeks). This left
+32 forecasters, one of which, the COVIDhub-4-week-ensemble, was an exact
+duplicate of another. To make the comparison fair, we scaled by the
+COVIDhub-baseline as in Figure 3 of the manuscript. Figure
+\ref{fig:compare-to-hub-mean} shows all submitted forecasts (available 7, 14,
+and 21 days ahead) overlaid on Figure 3. We have highlighted the
+COVIDhub-ensemble in light blue for comparison. All the models developed in this
+paper beat the baseline, and would rank in the top 8 of 31 submitted forecasts
+over this period.
+
+A more fair comparison that directly adjusts for spatiotemporal difficulty is to
+use the geometric mean scaled by the baseline. This is the same idea shown in
+\ref{fig:fcast-adjusted}. We have also remade this figure overlaying submitted
+forecasts available on the Forecast Hub. It occasionally happens that a
+submitted forecast achieved a WIS of 0, so in this case we use the WIS$+ 1$
+rather than WIS for all forecasters. The result is shown in
+\ref{fig:compare-to-hub-geomean}. By this metric, the methods employed in this
+paper would rank in the top 2, beating the COVIDhub-ensemble at all forecast
+horizons. The improvement of the best indicator assisted model over the AR model
+is roughly the same as the improvement of the AR model over the COVIDhub-ensemble.

### 2021-09-30 forecast/code/deprecated/compare-to-hub.R

@@ -1,0 +1,95 @@
+library(aws.s3)
+Sys.setenv("AWS_DEFAULT_REGION" = "us-east-2")
+s3bucket <- get_bucket("forecast-eval")
+n_keeper_weeks <- 6L
+n_keeper_locs <- 50L
+case_scores <- s3readRDS("score_cards_state_cases.rds", s3bucket)
+case_scores <- case_scores %>% 
+  mutate(forecast_date = target_end_date - ahead * 7) %>% 
+  # fix weirdnesses about submission dates
+  filter(forecast_date < "2021-01-01") %>%
+  select(ahead, geo_value, forecaster, target_end_date, wis, forecast_date)
+
+strawman <- case_scores %>% filter(forecaster == "COVIDhub-baseline")
+case_scores <- left_join(
+  case_scores, 
+  strawman %>% 
+    select(forecast_date, target_end_date, geo_value, wis) %>% 
+    rename(strawman_wis = wis)
+  )
+case_scores <- case_scores %>% 
+  filter(forecaster != "COVIDhub-baseline")
+n_submitted <- case_scores %>% 
+  filter(ahead == 2) %>% 
+  group_by(forecaster) %>% 
+  summarise(nfcasts = n())
+keepers <- n_submitted %>% 
+  filter(nfcasts / n_keeper_locs > n_keeper_weeks - .0001) %>% 
+  # submitted at least z weeks for x locations
+  pull(forecaster)
+
+case_scores <- case_scores %>% filter(forecaster %in% keepers)
+all_time_performance <- case_scores %>%
+  group_by(forecaster, ahead) %>%
+  summarise(rel_wis = Mean(wis) / Mean(strawman_wis),
+            geo_wis1 = GeoMean((wis + 1) / (strawman_wis + 1)),
+            geo_wis = GeoMean(wis / strawman_wis))
+
+all_time_performance %>%
+  pivot_longer(contains("wis")) %>%
+  ggplot(aes(ahead, value, color = forecaster)) +
+  theme_bw() +
+  geom_point() +
+  geom_line() +
+  scale_color_viridis_d() +
+  facet_wrap(~name) +
+  geom_hline(yintercept = 1, color = "red") +
+  theme(legend.position = "bottom", legend.title = element_blank()) +
+  scale_y_log10()
+
+df <- summarizer(fcasts_honest %>% filter(period != "jm") %>%
+                   group_by(ahead, forecaster), "wis", NULL, 
[…]

### 2021-09-30 forecast/code/deprecated/revision-draft.R

@@ -7,13 +7,26 @@ cumulatives <- fcasts_honest %>%
             gmw = GeoMean(wis / strawman_wis)) %>%
   arrange(forecast_date) %>%
   mutate(`Cumulative Mean` = cummean(mw) / cummean(smw),
-         `Cumulative Sum` = cumsum(mw),
+         `Cumulative Sum` = cumsum(mw) / cumsum(smw),
          `Cumulative Geo Mean` = exp(cummean(log(gmw))),
          `GM Regret` = cumsum(gmw) - cumsum(smw/smw),
          `AM Regret` = cumsum(mw) - cumsum(smw),
          `14 day trailing average` = RcppRoll::roll_meanl(mw / smw, n = 14L))
 
-  
+reltoAR <- cumulatives %>% 
+  select(forecaster, forecast_date, mw, gmw) %>%
+  ungroup()
+reltoAR <- left_join(
+  reltoAR %>% filter(forecaster != "AR"),
+  reltoAR %>% filter(forecaster == "AR") %>% select(-forecaster) %>%
+    rename(armw = mw, argmw = gmw)
+)  
+
+reltoAR <- reltoAR %>%
+  group_by(forecaster) %>%
+  arrange(forecast_date) %>%
+  mutate(cm = (mw - armw) / armw)
+
   
 fcast_colors <- c(RColorBrewer::brewer.pal(5, "Set1"), "#000000")
 names(fcast_colors) <- c("CHNG-CLI", "CHNG-COVID", "CTIS-CLIIC", "DV-CLI",
@@ -23,13 +36,22 @@ names(fcast_colors) <- c("CHNG-CLI", "CHNG-COVID", "CTIS-CLIIC", "DV-CLI",
 
 ggplot(cumulatives %>% filter(forecast_date < "2021-01-01"), 
        aes(forecast_date, color = forecaster)) + 
-  geom_line(aes(y = `Cumulative Geo Mean`)) +
+  geom_line(aes(y = `Cumulative Sum`)) +
   geom_hline(yintercept = 1) +
   xlab("forecast date") +
   scale_color_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
   theme_bw() +
   theme(legend.position = "bottom", legend.title = element_blank())
 
+ggplot(reltoAR %>% filter(forecast_date < "2021-01-01"), 
+       aes(forecast_date, color = forecaster)) + 
+  geom_line(aes(y = cm)) +
+  geom_hline(yintercept = 0) +
+  xlab("forecast date") +
[…]

### 2021-09-30 forecast/paper/response-to-reviewers.tex

@@
   this is a standard metric in the forecasting community (just as Mean Squared
-  Forecast Error is standard), rather than an ``author decision''.  In principle,
+  Forecast Error is standard), rather than an ``author decision''.  With the
+  weights as used, it is equivalent to quantile loss and therefore is a discrete
+  version of CRPS (see [7]).
+
+  In principle,
   we would expect that different weights may alter the conclusions, though
@@
   over which time the behavior may well be stationary. We have added a section
-  to the appendix that uses the cumulative mean as suggested (the sum is
-  difficult to see due to the scaling). With some exceptions at the beginning of
+  to the appendix that uses the cumulative sum normalized by baseline as
+  suggested. With some exceptions at the beginning of 
   the evaluation period, the relative ordering remains fairly consistent. The AR
@@
   Conventional methods for assessing statistical significance that we explored
-  were overconfident, to the point of being unbelievable. To a large degree,
-  this is a ``huge n, tiny p'' problem so even bootstrap analyses result in
-  extremely narrow intervals. Essentially, ``everything is significant''.
+  were overconfident, to the point of being unbelievable due to the
+  spatiotemporal correlation.
 
@@
 
-  To respond more directly:
+  To respond more directly we have made a number of additions.:
 
@@
   line 85 in the main text), that performs a Sign Test for differences in
-  accuracy. P-values there tend to cluster near zero suggesting again that these
-  improvements are statistically significant.
+  accuracy. P-values there tend to cluster near zero and one suggesting some
+  periods have significantly improved accuracy for the indicator-assisted models
+  while in others, the reverse is true.
 
   $\bullet$ We also use a Diebold-Mariano test
-  for comparing forecast accuracy. The results for this test are more mixed,
-  likely due to nonstationarity the Reviewer mentions above (the indicators are
+  for comparing forecast accuracy. The results for this test are mixed, as would
+  be expected given the sign test and
+  the nonstationarity the Reviewer mentions above. The indicators are
   more helpful in some periods than others, see also the upswing/downswing
-  discussion in Section 2C). 
+  discussion in Section 2C. 
 \end{resp}
@@
 \begin{resp}
-  See also Reviewer 1 point 3. And the general comment below (Point 8).
+  See also Reviewer 1 point 3. And the general comment below (Point 8) where we
+  outline some specific changes we have made that bare on this point.
+
   Practically, on average, if our goal is to predict 11 days ahead, when the AR
@@
   it is clear that all of the forecasters considered in this paper beat most
-  submissions. 
+  submissions. In the second figure, we use the Geometric Mean scaled by the
+  baseline, which should be a more meaningful comparison, as argued elsewhere in
+  the supplement. By this metric, the methods employed in this
+  paper would rank in the top 2 of all submitted forecasts, beating the
+  COVIDhub-ensemble at all forecast 
+  horizons. The improvement of the best indicator assisted model over the AR model
+  is roughly the same as the improvement of the AR model over the COVIDhub-ensemble.
 \end{resp}

### 2021-09-30 forecast/paper/supplement.Rmd

@@
 ```{r upswing-summary, fig.cap="Average difference between the WIS of the AR model and the WIS of the other forecasters. The indicator-assisted forecasters do best during down and flat periods."}
-# Figure 5 in the manuscript
 up_down_df_summary <- up_down_df %>%
@@
   ylab("Mean of WIS(AR) - WIS(F)") +
+  xlab("") +
+  theme(legend.position = "bottom", legend.title = element_blank())
+```
+
+```{r upswing-summary-remake, fig.cap="REMAKE: Percent change in average WIS of the forecaster (AR or indicator assisted) relative to the baseline.
+All models perform poorly durning down periods, but the indicators help.
+During flat periods, the indicators improve slightly over the AR.
+During up periods, all forecasters do much better than the baseline, but only some do as well as AR."}
+up_down_df_summary <- inner_join(
+  left_join(
+    corr_df, 
+    fcasts_honest %>% select(ahead, geo_value, forecast_date, strawman_wis)),
+  up_down, by = c("geo_value","target_end_date")) %>%
+  filter(period != "jm") %>%
+  group_by(forecaster, udf) %>%
+  summarise(mean = Mean(wis) / Mean(strawman_wis) - 1)
+up_down_df_summary %>%
+  ggplot(aes(udf, mean, fill = forecaster)) +
+  geom_bar(width = 0.6, position = position_dodge(width=0.6),
+           stat = "identity") +
+  scale_fill_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
+  theme_bw() +
+  scale_y_continuous(labels = scales::percent_format()) +
+  geom_hline(yintercept = 0) +
+  ylab("% change in Forecaster WIS\n relative to Baseline") +
   xlab("") +
@@
   ylab("Change relative to AR") +
+  xlab("") +
+  theme(legend.position = "bottom", legend.title = element_blank())
+```
+
+```{r hotspots-upswing-downswing-remake, fig.cap="REMAKE: Classification and loglikelihood separated into periods of upswing, downswing, and flat cases.
+The indicator assisted models always have lower classification error relative to the null classifier, while the AR actually does worse in down and flat periods.
+In terms of loglikelihood, all forecasters have lower likelihood than the null classifier during up periods.
+In down and flat periods, the indicators generally improve over the AR, while the are worse during up periods."}
+hot_udf <- inner_join(
+  hotspots_honest %>% filter(period != "jm"),
+  up_down %>% select(geo_value, target_end_date, udf))
+
+cutoff <- mean(hot_udf %>% filter(forecaster == "AR") %>% pull(actual))
+
+con_tab <- hot_udf %>% 
+  filter(!is.na(actual)) %>%
+  mutate(pred = value > cutoff) %>%
+  group_by(forecaster, udf) %>%
+  summarise(m = mean(pred != actual) - .5) %>%
+  ungroup()
+
+# con_tab <- left_join(
+#   con_tab %>% filter(forecaster != "AR"),
+#   con_tab %>% filter(forecaster == "AR") %>%
+#     select(-forecaster) %>%
+#     rename(mar = m))
+
+llike_tab <- hot_udf %>%
+  filter(!is.na(actual)) %>%
+  mutate( # kill prob 0-1 predictions
+    value = case_when(
+      value < 1e-8 ~ 1e-8,
+      value > 1-1e-8 ~ 1-1e-8,
+      TRUE ~ value
+    ),
+    llike = (actual == 1) * log(value) + (actual == 0) * log(1 - value),
+    nulldev = (actual == 1) * log(.5) + (actual == 0) * log(1 - .5)) %>%
+  group_by(forecaster, udf) %>%
+  summarise(m = mean(llike) - mean(nulldev)) %>%
+  ungroup()
+
+# llike_tab <- left_join(
+#   llike_tab %>% filter(forecaster != "AR"),
+#   llike_tab %>% filter(forecaster == "AR") %>%
+#     select(-forecaster) %>%
+#     rename(mar = m))
+
+bind_rows(con_tab %>% mutate(err = "Classification error"), 
+          llike_tab %>% mutate(err = "Log likelihood")) %>%
+  ggplot(aes(udf, m , fill = forecaster)) +
+  geom_bar(width = 0.6, position = position_dodge(width=0.6),
+           stat = "identity") +
+  scale_fill_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
+  #scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
+  theme_bw() +
+  geom_hline(yintercept = 0) +
+  facet_wrap(~err, scales = "free_y") +
+  ylab("Improvement relative to the null classifier") +
   xlab("") +
@@
 The shaded background shows total U.S.
-cases as reported by JHU-CSSE." }
+cases as reported by JHU-CSSE for the 14-day ahead target.
+Hashes along the $x$-axis denote weeks." }
 cumulatives <- fcasts_honest %>%
@@
 us_cases <- covidcast::covidcast_signal(
-  "jhu-csse", "confirmed_7dav_incidence_num", "2020-03-01", "2021-01-01",
+  "jhu-csse", "confirmed_7dav_incidence_num", "2020-03-01", "2021-02-01",
   geo_type = "nation") %>% as_tibble() %>%
@@
   filter(time_value > "2020-06-01")
-
-ggplot(cumulatives %>% filter(forecast_date < "2021-01-01"), 
-       aes(forecast_date, color = forecaster)) + 
-  geom_line(aes(y = `Cumulative Sum`)) +
-  annotate(x = us_cases$time_value, ymax = us_cases$value, ymin = -Inf,
+ticks <- seq(min(us_cases$time_value), max(us_cases$time_value), by = 7) - 14
+
+ggplot(cumulatives %>% filter(forecast_date < "2021-01-01")) + 
+  geom_line(aes(forecast_date, y = `Cumulative Sum`, color = forecaster)) +
+  annotate(x = us_cases$time_value - 14, ymax = us_cases$value, ymin = -Inf,
            geom = "ribbon", alpha = .1) +
+  annotate("segment", x=ticks, y=-Inf, xend = ticks, yend = .77, 
+               color = "grey70") +
   geom_hline(yintercept = 1) +
   xlab("forecast date") +
+  coord_cartesian(#ylim = c(.76, 1), 
+                  xlim = ymd(c("2020-06-01", "2021-01-01"))) +
+  scale_x_date(date_breaks = "2 months", date_minor_breaks = "1 week",
+               date_labels = "%b %Y") +
   scale_color_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
   theme_bw() +
-  theme(legend.position = "bottom", legend.title = element_blank())
-```
+  theme(legend.position = "bottom", legend.title = element_blank(),
+        panel.grid.minor.x = element_blank())
+```
+
+\clearpage
+
+```{r cumulative-geo-mean, fig.cap="This figure shows the cumulative geometric mean of WIS for each forecaster divided by the WIS for the baseline model.
+The shaded background shows total U.S.
+cases as reported by JHU-CSSE for the 14-day ahead target.
+Hashes along the $x$-axis denote weeks." }
+ggplot(cumulatives %>% filter(forecast_date < "2021-01-01")) + 
+  geom_line(aes(forecast_date, y = `Cumulative Geo Mean`, color = forecaster)) +
+  annotate(x = us_cases$time_value - 14, ymax = us_cases$value, ymin = -Inf,
+           geom = "ribbon", alpha = .1) +
+  annotate("segment", x=ticks, y=-Inf, xend = ticks, yend = .77, 
+               color = "grey70") +
+  geom_hline(yintercept = 1) +
+  xlab("forecast date") + ylab("Cumulative Geometric Mean") +
+  coord_cartesian(#ylim = c(.76, 1), 
+                  xlim = ymd(c("2020-06-01", "2021-01-01"))) +
+  scale_x_date(date_breaks = "2 months", date_minor_breaks = "1 week",
+               date_labels = "%b %Y") +
+  scale_color_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
+  theme_bw() +
+  theme(legend.position = "bottom", legend.title = element_blank(),
+        panel.grid.minor.x = element_blank())
+```
+
+\clearpage
 
@@
 
+<!-- Comparison to submitted forecasts -->
+
+```{r grab-submitted-case-fcasts}
+library(aws.s3)
+Sys.setenv("AWS_DEFAULT_REGION" = "us-east-2")
+s3bucket <- get_bucket("forecast-eval")
+# ensure that forecasters have submitted a reasonable amount
+n_keeper_weeks <- 6L
+n_keeper_locs <- 50L
+case_scores <- s3readRDS("score_cards_state_cases.rds", s3bucket) %>% 
+  mutate(forecast_date = target_end_date - ahead * 7) %>% 
+  # fix weirdnesses about submission dates, 
+  # this is generous to the forecaster
+  filter(forecast_date < "2021-01-01", ahead < 4) %>% 
+  select(ahead, geo_value, forecaster, target_end_date, wis, forecast_date)
+
+strawman <- case_scores %>% filter(forecaster == "COVIDhub-baseline")
+case_scores <- left_join(
+  case_scores, 
+  strawman %>% 
+    select(forecast_date, target_end_date, geo_value, wis) %>% 
+    rename(strawman_wis = wis)
+) %>% filter(forecaster != "COVIDhub-baseline")
+
+# Subset to those forecasters that submitted enough 14-day ahead forecasts
+n_submitted <- case_scores %>% 
+  filter(ahead == 2) %>% 
+  group_by(forecaster) %>% 
+  summarise(nfcasts = n())
+keepers <- n_submitted %>% 
+  filter(nfcasts / n_keeper_locs > n_keeper_weeks - .0001, 
+         forecaster != "COVIDhub-4_week_ensemble") %>% # same as CH-Ensemble 
+  pull(forecaster)
+
+case_scores <- case_scores %>% 
+  filter(forecaster %in% keepers) %>%
+  group_by(forecaster, ahead) %>%
+  summarise(rel_wis = Mean(wis) / Mean(strawman_wis),
+            geo_wis1 = GeoMean((wis + 1) / (strawman_wis + 1)),
+            geo_wis = GeoMean(wis / strawman_wis))
+```
+
+```{r compare-to-hub-mean, fig.height = 8.5, fig.cap="This figure reproduces Figure 3 in the main paper but overlays scores for the forecasts submitted to the COVID-19 Forecast Hub.
+Grey lines correspond to the various teams that submitted during period our evaluation period.
+We have highlighted the COVIDhub-ensemble, which is the official forecast of the CDC."}
+# Same as Fig 3 in the paper
+df <- summarizer(fcasts_honest %>% filter(period != "jm") %>%
+                   group_by(ahead, forecaster), "wis", NULL, 
+                 "strawman_wis", Mean, c("aggr","scale"))
+ggplot(df) +
+  geom_line(aes(ahead, wis, color = forecaster)) + 
+  geom_point(aes(ahead, wis, color = forecaster)) +
+  theme_bw() +
+  geom_hline(yintercept = 1, size = 1.5) +
+  xlab("Days ahead") +
+  ylab("Mean WIS (relative to baseline)") +
+  scale_color_manual(values = c(fcast_colors, "CH-ensemble" = "lightblue"),
+                     guide = guide_legend(nrow = 2)) +
+  theme(legend.position = "bottom", legend.title = element_blank()) +
+  geom_line(data = case_scores, 
+            aes(ahead * 7, rel_wis, group = forecaster),
+            color = "grey70") +
+  geom_line(data = case_scores %>% filter(forecaster == "COVIDhub-ensemble"),
+            aes(ahead * 7, rel_wis), color = "lightblue", size = 1.5) +
+  scale_y_log10()
+```
+
+\clearpage
+
+```{r compare-to-hub-geomean, fig.height = 8.5, fig.cap="This figure is similar to Figure \\ref{fig:fcast-adjusted}.
+In this case, we add 1 to both the forecaster WIS and the baseline WIS before scaling (to allow forecasters that achieve 0 error to appear), and we overlay scores for the forecasts submitted to the COVID-19 Forecast Hub.
+Grey lines correspond to the various teams that submitted during period our evaluation period.
+We have highlighted the COVIDhub-ensemble, which is the official forecast of the CDC."}
+# Like supplement GeoMean figure, but adding 1
+df2 <- fcasts_honest %>% 
+  filter(period != "jm") %>%
+  group_by(ahead, forecaster) %>%
+  summarise(wis = GeoMean((wis + 1) / (strawman_wis + 1)))
+
+ggplot(df2) +
+  geom_line(aes(ahead, wis, color = forecaster)) + 
+  geom_point(aes(ahead, wis, color = forecaster)) +
+  theme_bw() +
+  geom_hline(yintercept = 1, size = 1.5) +
+  xlab("Days ahead") +
+  ylab("Geometric mean of WIS (relative to baseline)") +
+  scale_color_manual(values = c(fcast_colors, "CH-ensemble" = "lightblue"),
+                     guide = guide_legend(nrow = 2)) +
+  theme(legend.position = "bottom", legend.title = element_blank()) +
+  geom_line(data = case_scores, 
+            aes(ahead * 7, geo_wis1, group = forecaster),
+            color = "grey70") +
+  geom_line(data = case_scores %>% filter(forecaster == "COVIDhub-ensemble"),
+            aes(ahead * 7, geo_wis1), color = "lightblue", size = 1.5) +
+  scale_y_log10()
+```
+
 <!-- All code below is deprecated -->

### 2021-09-30 forecast/paper/supplement.tex

@@
 
-\caption{\label{tab:dm-test}$P$-values for a Diebold-Mariano test for differences in forecast error.
+\caption{\label{tab:dm-test}$P$-values for a one-sided Diebold-Mariano test for improvement in forecast error.
 The test is for the null hypothesis of equal performance against the alternative that the indicator assisted model is better.}
@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/upswing-summary-remake-1}
+
+}
+
+\caption{REMAKE: Percent change in average WIS of the forecaster (AR or indicator assisted) relative to the baseline.
+All models perform poorly durning down periods, but the indicators help.
+During flat periods, the indicators improve slightly over the AR.
+During up periods, all forecasters do much better than the baseline, but only some do as well as AR.}\label{fig:upswing-summary-remake}
+\end{figure}
+
 \clearpage
@@
 Like the analysis of the forecasting task in the main paper (see Figure 7), performance is better during down and flat periods.}\label{fig:hotspots-upswing-downswing}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/hotspots-upswing-downswing-remake-1}
+
+}
+
+\caption{REMAKE: Classification and loglikelihood separated into periods of upswing, downswing, and flat cases.
+The indicator assisted models always have lower classification error relative to the null classifier, while the AR actually does worse in down and flat periods.
+In terms of loglikelihood, all forecasters have lower likelihood than the null classifier during up periods.
+In down and flat periods, the indicators generally improve over the AR, while the are worse during up periods.}\label{fig:hotspots-upswing-downswing-remake}
 \end{figure}
@@
 
-\caption{This figure shows the cumulative sum of WIS for each forecaster divided by the cumulative sum of WIS for the baseline model. The shaded background shows}\label{fig:cumulative-mean}
+\caption{This figure shows the cumulative sum of WIS for each forecaster divided by the cumulative sum of WIS for the baseline model.
+The shaded background shows total U.S.
+cases as reported by JHU-CSSE for the 14-day ahead target.
+Hashes along the $x$-axis denote weeks.}\label{fig:cumulative-mean}
 \end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/cumulative-geo-mean-1}
+
+}
+
+\caption{This figure shows the cumulative geometric mean of WIS for each forecaster divided by the WIS for the baseline model.
+The shaded background shows total U.S.
+cases as reported by JHU-CSSE for the 14-day ahead target.
+Hashes along the $x$-axis denote weeks.}\label{fig:cumulative-geo-mean}
+\end{figure}
+
+\clearpage
 
@@
 \clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/compare-to-hub-mean-1}
+
+}
+
+\caption{This figure reproduces Figure 3 in the main paper but overlays scores for the forecasts submitted to the COVID-19 Forecast Hub.
+Grey lines correspond to the various teams that submitted during period our evaluation period.
+We have highlighted the COVIDhub-ensemble, which is the official forecast of the CDC.}\label{fig:compare-to-hub-mean}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/compare-to-hub-geomean-1}
+
+}
+
+\caption{This figure is similar to Figure \ref{fig:fcast-adjusted}.
+In this case, we add 1 to both the forecaster WIS and the baseline WIS before scaling (to allow forecasters that achieve 0 error to appear), and we overlay scores for the forecasts submitted to the COVID-19 Forecast Hub.
+Grey lines correspond to the various teams that submitted during period our evaluation period.
+We have highlighted the COVIDhub-ensemble, which is the official forecast of the CDC.}\label{fig:compare-to-hub-geomean}
+\end{figure}
 

### 2021-10-01 survey/code/paper-tables.Rmd
Figure tweaks suggested by Josh

@@
 signals <- append(signals, list(anosmia)) %>%
-  aggregate_signals(format = "long")
+  aggregate_signals(format = "long") %>%
+  filter(!(signal == "confirmed_admissions_covid_1d_prop" &
+             time_value < as.Date("2020-07-15")))
 
@@
              "confirmed_admissions_covid_1d_prop", "deaths_incidence_prop"),
-  labels = c("Anosmia %", "CLI %", "Community CLI %", "Cases per 100,000",
+  labels = c("Anosmia (%)", "CLI (%)", "Community CLI (%)", "Cases per 100,000",
              "Hospitalizations per 100,000", "Deaths per 100,000"))
@@
   facet_wrap(vars(signal), nrow = 2, ncol = 3, scales = "free_y") +
-  labs(x = "Date", y = "Value", color = "Census Region") +
+  labs(x = "Date", y = "", color = "Census Region") +
   scale_x_date(date_breaks = "3 months",
@@
                guide = guide_axis(angle = 45)) +
+  scale_y_continuous(minor_breaks = NULL) +
   theme_bw() +
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom",
+        strip.background = element_rect(color = "black", fill = "white"))
 
@@
 )
+
+# Force factor ordering
+data$positive_14d <- factor(
+  data$positive_14d,
+  levels = c("Tested positive", "All others"))
 
@@
                      sec.axis = sec_axis(~ . * rr_scale_factor,
-                                         name = "Relative risk")) +
+                                         name = "Relative risk"),
+                     minor_breaks = NULL) +
   scale_x_discrete(guide = guide_axis(angle = 45)) +
@@
   theme_bw() +
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom",
+        panel.grid.major.x = element_blank())
 
@@
   inner_join(case_age_dist, by = c("year", "month")) %>%
-  mutate(case_rate_adult = cumulative_cases * frac_adult_cases / adult_pop * 100)
+  mutate(case_rate_adult = cumulative_cases * frac_adult_cases / adult_pop * 100) %>%
+  filter(state != "PR")
 
@@
             by = c("state_abbr" = "state", "week" = "time_value"))
+```
 
+```{r case-rate-compare-plot}
 g <- ggplot(state_case_compare) +
-  facet_geo(~ state_abbr, grid = "us_state_with_DC_PR_grid1", label = "code") +
+  facet_geo(~ state_abbr, grid = "us_state_grid1", label = "code") +
   geom_line(aes(x = week, y = case_rate_adult, color = "Surveillance"),
@@
   scale_color_manual(values = c("lightblue", "gray50")) +
-  scale_y_continuous(breaks = c(0, 10),
+  scale_y_continuous(limits = c(0, 20), breaks = c(0, 10, 20),
                      minor_breaks = c(5, 15),
@@
                                             accuracy = 1)) +
-  scale_x_date(date_breaks = "4 months", date_labels = "%b",
-               date_minor_breaks = "1 month") +
+  scale_x_date(breaks = as.Date(c("2020-10-01", "2021-04-01")),
+               date_labels = "%b", minor_breaks = NULL) +
   labs(x = "", y = "", color = "") +
   theme_minimal() +
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom",
+        axis.text = element_text(size = rel(0.5)))
 
@@
   scale_color_discrete(name = "SVI") +
+  scale_y_continuous(minor_breaks = NULL, limits = c(0, NA)) +
   scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
@@
   labs(x = "", y = "") +
-  ylim(0,  NA) +
   theme_bw() +
@@
         legend.margin = margin(legend_top_margin, 0, 11 / 2, 0),
-        plot.margin = margin(11 / 2, 11 / 2, 0, 11 / 2))
+        plot.margin = margin(11 / 2, 11 / 2, 0, 11 / 2),
+        strip.background = element_rect(color = "black", fill = "white"))
 
@@
   scale_color_discrete(name = "Age") +
+  scale_y_continuous(minor_breaks = NULL, limits = c(0, NA)) +
   scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
@@
   labs(x = "", y = "") +
-  ylim(0, NA) +
   theme_bw() +
@@
   scale_color_discrete(name = "Region")  +
+  scale_y_continuous(minor_breaks = NULL, limits = c(0, NA)) +
   scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
@@
 
-ggsave("../paper/fig/contacts-masks-transport.pdf", g, width = 8, height = 5)
+ggsave("../paper/fig/contacts-masks-transport.pdf", g, width = 7, height = 6)
 ```

### 2021-10-01 survey/code/vaccine-acceptance-race.R

@@ -89,17 +89,18 @@ ggplot(df[race_grp %in% c("Asian", "Black", "Hispanic", "White") &
             gender %in% c("Female", "Male") &
             sample_size >= 50],
        aes(x = age_cat_fb, y = pct_accepting, color = race_grp, group = race_grp)) +
-  geom_point(size = 2) +
-  geom_line(size = 1.2) +
+  geom_point() +
+  geom_line() +
   facet_grid(cols = vars(Region), rows = vars(gender)) +
   theme_bw() +
-  labs(x = "Age Group", y = "Vaccinated or accepting",
+  labs(x = "Age group (years)", y = "Vaccinated or accepting",
        color = "") +
   scale_y_continuous(labels = label_percent(accuracy = 1.0)) +
   scale_color_manual(breaks = c("Hispanic", "White", "Black", "Asian"),
                      values = c("Hispanic" = "#c42e31", "White" = "#832543",
                                 "Asian" = "#6399AC", "Black" = "#e5a825")) +
   theme(text = element_text(size = 12), legend.position = "bottom",
-        axis.text.x = element_text(angle = 45, hjust = 1))
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        strip.background = element_rect(color = "black", fill = "white"))
 
 ggsave("../paper/fig/age-race-vaccine-acceptance.pdf", width = 8, height = 5)

### 2021-10-03 indicators/paper/paper.tex
Tweak last intro paragraph

@@
 previous (historical) versions of the signals, a key feature that exposes the
-effects of data revisions. Lastly, we provide R \cite{CovidcastR} and Python
+effects of data revisions. Moreover, we provide R \cite{CovidcastR} and Python 
 \cite{CovidcastPy} packages to facilitate interaction with the API, and an
@@
 In a companion paper, we analyze the utility provided by a core set of the
-indicators in COVID-19 forecasting and hotspot prediction models
+indicators in short-term COVID-19 forecasting and hotspot prediction models 
 \cite{McDonald:2021}. In another companion paper, we elaborate on our research
@@
 and available in aggregate form in the COVIDcast API \cite{Salomon:2021}. This
-paper focuses on the COVID-19 activity indicators themselves, describing the
+paper focuses on the COVID-19 indicators themselves, describing the
 data streams, how they are processed and made publicly available, and insights
 that can be gained by combining novel data sources with standard public health
-surveillance datasets.
+surveillance data.
 

### 2021-10-03 indicators/paper/paper.tex
Tweak last paragraph in Sect 1A

@@
 
-Altogether, we collect over 170 signals from 12 distinct sources, aggregate and
-process them, and provide them in a common format for public access. This
-unifies both unique (not available anywhere else) and standard COVID data
-streams into a single common format, enabling efficient comparison and modeling.
-A summary of the data sources and signals in the API is in
-Table~\ref{tab:sources_signals}, and detailed documentation is available at
+Altogether, we produce over 170 signals from 12 distinct sources, and provide
+them in a common format for public access. This unifies both unique (unavailable
+anywhere else) and standard COVID data streams into a single common 
+format, enabling efficient comparison and modeling. A summary of the data
+sources and signals in the API is in Table~\ref{tab:sources_signals}, and
+detailed documentation is available online at
 \url{https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html}.

### 2021-10-04 indicators/paper/paper.tex
Commenting out key design goals

@@
 improved as new sources of data arrive, and in response to interviews with users
-and health experts, usage analytics from the site, and user surveys. Key design
-goals for visualizations displaying COVIDcast data have included representing
-uncertainty and missing data, making features consistent and discoverable
-despite a wide variety of signal types, and smoothly switching between different
-geographic aggregations and time scales.
+and health experts, usage analytics from the site, and user surveys.% Key design 
+% goals for visualizations displaying COVIDcast data have included representing
+% uncertainty and missing data, making features consistent and discoverable
+% despite a wide variety of signal types, and smoothly switching between
+% different geographic aggregations and time scales.
 

### 2021-10-04 indicators/paper/paper.tex
Edit text around Fig 6

@@
 Since July 2020, we have been regularly submitting short-term forecasts of
-COVID-19 case and death incidence, at the state and county levels, to the
+COVID-19 case and death incidence, at both the state and county levels, to the 
 COVID-19 Forecast Hub \cite{ForecastHub}, with ``CMU-TimeSeries'' as the
@@
 Besides tracking contemporaneous COVID activity, these and other indicators can
-be used to improve forecasts of future COVID case trends, as investigated in a
+be used to improve forecasts of future COVID case trends, as investigated in our
 companion paper \cite{McDonald:2021}.
@@
 for tracking symptoms and case trends, especially when other data is
-unavailable. The correlations vary over time in complex ways; while some of this
-variation is doubtless due to complex changes in public behavior, reporting and
-testing practices, and so on, some is also related to changes in overall
-COVID-19 case rates. For example, comparing against
-Figure~\ref{fig:time_trends_national}, we observe that correlations drop in
-February 2021 for several of the signals, matching when COVID cases also sharply
-decline. This decline likely caused the heterogeneity in case rates by county to
-decline, making it more difficult for any signal to achieve a high correlation
-by county.
+unavailable.
+
+Also notable is the fact that the correlations fluctuate over time in complex
+ways; while some of this variation is likely due to changes in public behavior,
+reporting and testing practices, and so on, some is also related to changes in
+overall COVID-19 case trends. For example, comparing against
+Figure~\ref{fig:time_trends_national}, we see that correlations drop in February
+2021 for many of the signals, roughly matching the point in time when COVID
+cases also sharply decline. This decline likely caused the heterogeneity in case
+rates by county to decline, making it more difficult for any signal to achieve a
+high geo-wise correlation.
 
@@
 population who know someone in their local community with COVID-Like Illness
-(CLI). As we can see in the right-hand panel of the figure, this signal was not
-affected by Bexar County's reporting problems and, as shown in the last
+(CLI). As we can see in the right-hand panel of the figure, this indicator was 
+not affected by Bexar County's reporting problems and, as shown in the last 
 subsection, it is (in general) highly correlated with case rates, providing an
-alternate stream of data about COVID activity unaffected by backlogs. Similar
-reporting problems have occurred in many jurisdictions across the United States,
-and audits have regularly discovered misclassified or unreported cases and
-deaths, making it valuable to cross-check against external sources not part of
-the same reporting systems.
+alternate stream of data about COVID activity unaffected by backlogs. In
+general, reporting problems have occurred in many jurisdictions across the
+United States, and audits have regularly discovered misclassified or unreported
+cases and deaths, making it valuable to cross-check against external sources not
+part of the same reporting systems.
 
@@
 suggests that statistical models could be fit (potentially separately for each
-location) to estimate the final values from preliminary reports. On the other
-hand, official public health reporting of COVID cases and deaths can be subject
-to revision as death certificates are audited and backlogs cleared, resulting in
-thousands of cases and deaths being added or removed.
-Figure~\ref{fig:backfill-compare} shows relative differences between early
-signal values, reported 10--90 days after the underlying events, and later
-versions at least four months later. As the distribution of these revision
-amounts is highly skewed, the figure plots the 95th percentile of relative
-change, showing that reported deaths can have large revisions comparable to
-revisions in claims-based signals. This backfill is much more difficult to
-predict, and thus claims data and other sources may be a useful stand-in while
-public health reports are aggregated and corrected.
+location) to estimate the final values from preliminary reports.
 
@@
     reported between 10 and 90 days later are compared to ``final'' versions
-    recorded as of August 13, 2021, 120 days after April 15. Even officially
-    reported case and death data can have large revisions 30--60 days or more
-    after initial reporting; much of this is driven by individual large
-    revisions affecting specific states and dates, rather than systematic
-    changes affecting all states and dates.}
+    recorded as of August 13, 2021. Even officially reported case and death data
+    can have large revisions 30--60 days or more after initial reporting; much
+    of this is driven by individual large revisions affecting specific states
+    and dates, rather than systematic changes affecting all states and dates.} 
   \label{fig:backfill-compare}
 \end{figure}
+
+Figure~\ref{fig:backfill-compare} shows relative differences between early
+indicator values, reported 10--90 days after the underlying events, and later 
+versions at least four months later. As the distribution of these revision
+amounts is highly skewed, the figure plots the 95th percentile of relative
+change, showing that reported deaths can incur large relative error in initial
+values, comparable to that in claims-based signals.  However, for deaths, as
+well as cases, the revision pattern is not very systematic, with large
+corrections typically occuring at a sparse subset of locations and times
+(e.g., due to audits or backlogs being cleared, which can result in thousands of
+cases or deaths can be  added or removed all at once).  This backfill is much
+more difficult to predict than that of claims data, thus the latter (and other 
+sources) may be useful for nowcasting cases and deaths while public health 
+reports are being aggregated and corrected.
 

### 2021-10-04 indicators/paper/paper.tex
Comment out last revisions paragraph

@@
 in our own analysis of forecasting and hotspot prediction models appearing in a
-companion paper \cite{McDonald:2021}. Revisions to forecast targets, such as
-cases and deaths, also complicate the evaluation of forecast models: comparisons
-of forecasts to the ground truth may shift when the ground truth is revised
-weeks later. Only by systematically tracking revisions can this effect be
-monitored.
+companion paper \cite{McDonald:2021}.% Revisions to forecast targets, such as 
+% cases and deaths, also complicate the evaluation of forecast models:
+% comparisons of forecasts to the ground truth may shift when the ground truth
+% is revised weeks later. Only by systematically tracking revisions can this
+% effect be monitored.
 

### 2021-10-04 indicators/paper/paper.tex
Edit text around Fig 7

@@
   \includegraphics[width=\columnwidth]{fig/vaccine-intent-compare.pdf}
-  \caption{CTIS estimates of the percentage of people willing to get vaccinated
-    on January 20, 2021, compared to CDC reporting of the percentage of people
-    vaccinated in each county by July 20, 2021. Points are colored by the United
-    States Census region containing each county. Restricted to counties with at
-    least 250 survey responses from January 14--20, 2021.}
+  \caption{CTIS estimates of the percentage of people willing to get
+    vaccinated, back on January 20, 2021, compared to CDC reporting of the  
+    percentage of people vaccinated, on July 20, 2021. Each point is a county
+    (with at least 250 survey responses between January 14--20, 2021), colored
+    by its parent United States Census region.}     
   \label{fig:vaccine-intent-compare}
@@
 uptake of COVID-19 vaccines by July 2021 (as reported by the Centers for Disease
-Control and Prevention). It also illustrates a geographical disparity: in the
-Northeast, actual vaccination rates more closely match vaccine acceptance in
-January than they do in the South, where vaccination rates lag overall. While
+Control and Prevention). It also reveals a geographical disparity: in the
+Northeast, actual vaccination rates more closely match vaccine willingness rates
+in January than they do in the South, where vaccination rates lag overall. While 
 these results must be interpreted carefully due to potential sampling biases in

### 2021-10-04 indicators/code/indicator-paper-figures.Rmd
Axes labels for Fig 7

@@
   theme_bw() +
-  labs(x = "Estimated % willing to be vaccinated, 2021-01-20",
-       y = "% vaccinated on 2021-07-20", color = "Region") +
+  labs(x = "% willing to be vaccinated on January 20, 2021",
+       y = "% vaccinated on July 20, 2021", color = "Region") +
   theme(legend.pos = "bottom", legend.direction = "horizontal")

### 2021-10-04 forecast/paper/paper.tex
Small edits - Shorten significance statement (it was over 120 words) - Use GoogleSymptoms ref instead of Wahltinez2020 ref; their git   repo suggests we were citing it correctly in the first place,   and also the Wahltinez2020 was giving me compilation errors

@@
   spread of some disease process, based on disease activity in the recent
-  past---say, using what is known as autoregressive modeling in the time series
-  literature---is a basic and widely-used method.  In this paper we study
-  whether including auxiliary indicators as additional features in such an
-  autoregressive model can be useful in COVID-19 case forecasting and hotspot
-  prediction.  We find that indicators based on de-identified medical insurance 
-  claims, self-reported symptoms through online surveys, and COVID-related
-  internet search activity
-  % all provide a nontrivial boost in accuracy,
-  all provide a nontrivial improvement in forecast accuracy,
-  though
-  their utility generally depends on the pandemic's local dynamics (upswing,
-  downswing, or flat period) at prediction time.
+  past---say, using autoregressive modeling from the time series literature---is 
+  a basic and widely-used method.  In this paper we study whether including
+  auxiliary indicators as additional features in such an autoregressive model
+  can be useful in COVID-19 case forecasting and hotspot prediction.  We find
+  that indicators based on de-identified medical insurance claims, self-reported
+  symptoms through online surveys, and COVID-related internet search activity
+  all provide a nontrivial improvement in forecast accuracy, though their
+  utility depends on the pandemic's local dynamics (upswing, downswing, or flat
+  period) at prediction time. 
 }
@@
   smell or taste), based on Google's COVID-19 Search Trends data
-  set~\cite{Wahltinez2020}.   
+  set~\cite{GoogleSymptoms}.   
 \end{itemize}

### 2021-10-04 forecast/paper/paper.tex
Edit baseline model text

@@
   Google search volume for queries that relate to anosmia or ageusia (loss of
-  smell or taste), based on Google's COVID-19 Search Trends data
-  set~\cite{GoogleSymptoms}.   
+  smell or taste), based on Google's COVID-19 Search Trends data set.
 \end{itemize}
@@
 dates.  Shown is the average WIS for each forecast model divided by that from a
-% change since review
-baseline model.
-The baseline model is a flat-line forecaster which forms its median forecast by
-using the most recent value $Y_{\ell,t}$ as the median forecast for all aheads
-$Y_{\ell,t+a}$ . To create
-the quantiles, we use the empirical quantiles of the residuals from this
-forecast over the available history. 
-% old version
-% baseline model,
-% which is basically a flat-line forecaster (its median forecast
-% for $Y_{\ell,t+a}$ is always $Y_{\ell,t}$, with predicted quantiles defined
-% around this based on historical variation).
-This is the same baseline model as
-in the COVID-19 Forecast Hub. Here, we use the baseline model in order to scale
-mean WIS so that it gets put on an interpretable, unitless scale.  In the
-figure, we can see that all curves are below 1, which means (smaller WIS is
-better) that all of the models, including AR, outperform the baseline on average
-over the forecasting period.  On the other hand, the models deliver at best an
-improvement of about 20\% in average WIS over the baseline model, with this gap
-narrowing to about 10\% at the largest ahead values, illustrating the difficulty 
-of the forecasting problem.
+baseline model. The baseline model is a flat-line forecaster which forms its
+median forecast by using the most recent value $Y_{\ell,t}$ for all aheads
+$Y_{\ell,t+a}$, with predicted quantiles defined by the empirical distribution
+of the residuals from this median forecast over recent history. This is the same
+baseline model as in the COVID-19 Forecast Hub. Here, we use the baseline model
+in order to scale mean WIS to put it on an interpretable, unitless
+scale.  In the figure, we can see that all curves are below 1, which means
+(smaller WIS is better) that all of the models, including AR, outperform the
+baseline on average over the forecasting period.  On the other hand, the models
+deliver at best an improvement of about 20\% in average WIS over the baseline
+model, with this gap narrowing to about 10\% at the largest ahead values,
+illustrating the difficulty of the forecasting problem.
 

### 2021-10-04 forecast/paper/supplement-text.tex
Update reference

@@
 Another relatively assumption-lean method of testing for differences in forecast
-accuracy is the Diebold-Mariano test \cite{Diebold:2002,Diebold:2015,harvey1997testing}.
-Essentially, the differences between forecast losses are assumed to have a
-constant mean and a covariance that depends on time. Under these conditions, the
-asymptotic distribution for the standardized mean of the differences is limiting
-normal provided that a heteroskedasticity and autocorrelation robust estimate of
-the variance is used. Using the loss as weighted interval score across all HRRs
-and horizons (7 to 21 days ahead), we perform the DM test using both the mean
-relative to the strawman (as reported in the manuscript) and the geometric mean
-relative to the strawman as described above. The first two rows of Table
-\ref{tab:dm-test} displays 
-$P$-values for the test that the indicator-assisted model is no-better than the
-AR model. Only the CHNG-CLI model's $P$-value exceeds conventional statistical
-significance thresholds.
+accuracy is the Diebold-Mariano test \cite{Diebold:2002, Diebold:2015,
+  Harvey:1997}. Essentially, the differences between forecast losses are assumed
+to have a constant mean and a covariance that depends on time. Under these
+conditions, the asymptotic distribution for the standardized mean of the
+differences is limiting normal provided that a heteroskedasticity and
+autocorrelation robust estimate of the variance is used. Using the loss as
+weighted interval score across all HRRs and horizons (7 to 21 days ahead), we
+perform the DM test using both the mean relative to the strawman (as reported in
+the manuscript) and the geometric mean relative to the strawman as described
+above. The first two rows of Table \ref{tab:dm-test} displays $P$-values for the
+test that the indicator-assisted model is no-better than the AR model. Only the
+CHNG-CLI model's $P$-value exceeds conventional statistical significance
+thresholds.
 

### 2021-10-04 forecast/paper/paper.tex
Add companion refs

@@
 crucial for understanding disease spread, formulating public policy responses,
-and anticipating future public health resource needs.  In a companion paper, we
-describe our research group's (Delphi's) efforts in curating and maintaining a 
-database of real-time indicators that track COVID-19 activity and other relevant
-phenomena. The signals (a term we use synonomously with ``indicators'') in this
-database are accessible through the COVIDcast API \cite{CovidcastAPI}, with
-associated R \cite{CovidcastR} and Python \cite{CovidcastPy} packages for
-convenient data fetching and processing tools. In the current paper, we aim to
-quantify the utility provided by a core set of these indicators for two
-fundamental prediction tasks: probabilistic forecasting of COVID-19 case 
-rates and prediction of future COVID-19 case hotspots (defined by the event that
-a relative increase in COVID-19 cases exceeds a certain threshold).
+and anticipating future public health resource needs.  A companion paper
+\cite{Reinhart:2021} describes our research group's (Delphi's) efforts in
+curating and maintaining a database of real-time indicators that track COVID-19
+activity and other relevant phenomena. The signals (a term we use synonomously
+with ``indicators'') in this database are accessible through the COVIDcast API
+\cite{CovidcastAPI}, with associated R \cite{CovidcastR} and Python
+\cite{CovidcastPy} packages for convenient data fetching and processing
+tools. In the current paper, we aim to quantify the utility provided by a core
+set of these indicators for two fundamental prediction tasks: probabilistic
+forecasting of COVID-19 case rates and prediction of future COVID-19 case
+hotspots (defined by the event that a relative increase in COVID-19 cases
+exceeds a certain threshold).
 
@@
 
-In our companion paper, we analyze correlations between various indicators and   
-COVID case rates. These correlations are natural summaries of the
-contemporaneous association between an indicator and COVID cases, but they fall 
-short of delivering a satisfactory answer to the question that motivates the
-current article: is the information contained in an indicator demonstrably
-useful for the prediction tasks we care about? For such a question, lagged
-correlations (e.g., measuring the correlation between an indicator and COVID
-case rates several days in the future) move us in the right direction, but still
-do not move us all the way there. The question about \textit{utility for 
-  prediction} is focused on a much higher standard than simply asking 
-about correlations; to be useful in forecast or hotspot models, an indicator
-must provide relevant information that is not otherwise contained in past values
-of the case rate series itself. We will assess this in the most direct way
+In the companion paper \cite{Reinhart:2021}, we analyze correlations between
+various indicators and COVID case rates. These correlations are natural
+summaries of the contemporaneous association between an indicator and COVID
+cases, but they fall short of delivering a satisfactory answer to the question
+that motivates the current article: is the information contained in an indicator
+demonstrably useful for the prediction tasks we care about? For such a question,
+lagged correlations (e.g., measuring the correlation between an indicator and
+COVID case rates several days in the future) move us in the right direction, but
+still do not move us all the way there. The question about \textit{utility for
+  prediction} is focused on a much higher standard than simply asking about 
+correlations; to be useful in forecast or hotspot models, an indicator must
+provide relevant information that is not otherwise contained in past values of
+the case rate series itself. We will assess this in the most direct way
 possible: by inspecting the difference in predictive performance of simple
@@
 \url{https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html},
-which documents all of the signals in the COVIDcast API, and our companion paper
-on the API and database. For CTIS in particular, we refer to our companion paper
-on this survey. For the Google COVID-19 Search Trends data set, see
+and the companion paper on the COVIDcast API and its signals
+\cite{Reinhart:2021}. For CTIS in particular, we refer to the companion paper
+\cite{Salomon:2021}. For the Google COVID-19 Search Trends data set, see
 \cite{GoogleSymptoms}; see also \cite{Klopfen:2020, Vaira:2020} for a
-justification of the relevance of anosmia or ageusia to COVID-19 infection.
+justification of the relevance of anosmia or ageusia to COVID infection.
 
@@
 point in time.)  This is possible because of the COVIDcast API's comprehensive
-data versioning system (described in more detail in our companion paper).  We
-also use the \texttt{evalcast} R package \cite{EvalcastR}, which streamlines the 
-process of training arbitrary prediction models over a sequence of prediction
-dates, by constructing the proper sequence of vintage training data sets.
+data versioning system (described in more detail in the companion paper
+\cite{Reinhart:2021}).  We also use the \texttt{evalcast} R package
+\cite{EvalcastR}, which streamlines the process of training arbitrary
+prediction models over a sequence of prediction dates, by constructing the
+proper sequence of vintage training data sets.
 

### 2021-10-04 forecast/paper/paper.tex
Copyedit claims about WIS

@@
 in COVID-19 forecast scoring is discussed in \cite{Bracher:2021}; WIS is also
-the main evaluation metric used in the COVID-19 Forecast Hub.
-% Addition in response to Reviewer 1
-More broadly, the specific form of WIS used here is a standard metric in the
-forecasting community for evaluating probabilistic interval forecasts, just as
-Mean Squared Forecast Error is standard for point forecasts.
+the main evaluation metric used in the COVID-19 Forecast Hub. More broadly, the
+specific form of WIS used here is a standard metric in the forecasting community
+for evaluating quantile-based probabilistic forecasts, just as mean squared
+forecast error is standard for point forecasts.
 

### 2021-10-04 forecast/paper/response-to-reviewers.tex
clean up figure 5, move hub comparisons out

@@
 \usepackage{newtxtext}
+\usepackage{graphicx}
 
@@
 
-  $\bullet$ We have added two figures to the supplement along with a section of
-  text (Section 9) that compares our forecasts with state-of-the-art forecasts
-  submitted to the COVID-19 Forecast Hub. The first (Figure S26) overlays the
+  $\bullet$ Here are adding two figures
+    that compare our forecasts with state-of-the-art forecasts
+  submitted to the COVID-19 Forecast Hub. The first overlays the
   scores from the submissions on top of Figure 3 from the manuscript. We should
@@
   horizons. The improvement of the best indicator assisted model over the AR model
-  is roughly the same as the improvement of the AR model over the COVIDhub-ensemble.
+  is roughly the same as the improvement of the AR model over the
+  COVIDhub-ensemble.
+
+  \begin{figure}[tb]
+    \centering
+    \includegraphics[width=.9\textwidth]{fig/compare-to-hub-mean-1.pdf}
+    \caption{This figure reproduces Figure 3 in the main paper but overlays
+      scores for the forecasts submitted to the COVID-19 Forecast Hub. Grey
+      lines correspond to the various teams that submitted during period our
+      evaluation period. We have highlighted the COVIDhub-ensemble, which is the
+      official forecast of the CDC.}
+    \label{fig:compare-to-hub-mean}
+  \end{figure}
+
+  \begin{figure}[tb]
+    \centering
+    \includegraphics[width=.9\textwidth]{fig/compare-to-hub-geomean-1.pdf}
+    \caption{This figure is more like Figure S6 in the Supplement. In this case,
+      we add 1 to both the forecaster WIS and the baseline WIS before scaling
+      (to allow forecasters that achieve 0 error to appear), and we overlay
+      scores for the forecasts submitted to the COVID-19 Forecast Hub. Grey
+      lines correspond to the various teams that submitted during period our
+      evaluation period. We have highlighted the COVIDhub-ensemble, which is the
+      official forecast of the CDC.}
+    \label{fig:compare-to-hub-geomean}
+  \end{figure}
 \end{resp}

### 2021-10-04 forecast/paper/paper.tex
clean up figure 5, move hub comparisons out

--- 2021-10-04 forecast/paper/paper.tex
@@
   spread of some disease process, based on disease activity in the recent
-  past---say, using autoregressive modeling from the time series literature---is 
-  a basic and widely-used method.  In this paper we study whether including
-  auxiliary indicators as additional features in such an autoregressive model
-  can be useful in COVID-19 case forecasting and hotspot prediction.  We find
-  that indicators based on de-identified medical insurance claims, self-reported
-  symptoms through online surveys, and COVID-related internet search activity
-  all provide a nontrivial improvement in forecast accuracy, though their
-  utility depends on the pandemic's local dynamics (upswing, downswing, or flat
-  period) at prediction time. 
+  past---say, using what is known as autoregressive modeling in the time series
+  literature---is a basic and widely-used method.  In this paper we study
+  whether including auxiliary indicators as additional features in such an
+  autoregressive model can be useful in COVID-19 case forecasting and hotspot
+  prediction.  We find that indicators based on de-identified medical insurance 
+  claims, self-reported symptoms through online surveys, and COVID-related
+  internet search activity
+  % all provide a nontrivial boost in accuracy,
+  all provide a nontrivial improvement in forecast accuracy,
+  though
+  their utility generally depends on the pandemic's local dynamics (upswing,
+  downswing, or flat period) at prediction time.
 }
@@
 state-level deaths and shows that time series models perform among the top 5.
-In the Supplement, we compare the time series models investigated in this paper
-to the submitted case forecasts over the same period. The comparison is not
-particularly flattering: the simple models examined here outperform most
-submitted models and approach the accuracy of the COVIDHub Ensemble used by the
-CDC.
+Our internal evaluations suggest that the same is true (at least, if not to a
+greater extent) for case forecasting.
 
@@
   Google search volume for queries that relate to anosmia or ageusia (loss of
-  smell or taste), based on Google's COVID-19 Search Trends data set.
+  smell or taste), based on Google's COVID-19 Search Trends data
+  set~\cite{Wahltinez2020}.   
 \end{itemize}
@@
 dates.  Shown is the average WIS for each forecast model divided by that from a
-baseline model. The baseline model is a flat-line forecaster which forms its
-median forecast by using the most recent value $Y_{\ell,t}$ for all aheads
-$Y_{\ell,t+a}$, with predicted quantiles defined by the empirical distribution
-of the residuals from this median forecast over recent history. This is the same
-baseline model as in the COVID-19 Forecast Hub. Here, we use the baseline model
-in order to scale mean WIS to put it on an interpretable, unitless
-scale.  In the figure, we can see that all curves are below 1, which means
-(smaller WIS is better) that all of the models, including AR, outperform the
-baseline on average over the forecasting period.  On the other hand, the models
-deliver at best an improvement of about 20\% in average WIS over the baseline
-model, with this gap narrowing to about 10\% at the largest ahead values,
-illustrating the difficulty of the forecasting problem.
+% change since review
+baseline model.
+The baseline model is a flat-line forecaster which forms its median forecast by
+using the most recent value $Y_{\ell,t}$ as the median forecast for all aheads
+$Y_{\ell,t+a}$ . To create
+the quantiles, we use the empirical quantiles of the residuals from this
+forecast over the available history. 
+% old version
+% baseline model,
+% which is basically a flat-line forecaster (its median forecast
+% for $Y_{\ell,t+a}$ is always $Y_{\ell,t}$, with predicted quantiles defined
+% around this based on historical variation).
+This is the same baseline model as
+in the COVID-19 Forecast Hub. Here, we use the baseline model in order to scale
+mean WIS so that it gets put on an interpretable, unitless scale.  In the
+figure, we can see that all curves are below 1, which means (smaller WIS is
+better) that all of the models, including AR, outperform the baseline on average
+over the forecasting period.  On the other hand, the models deliver at best an
+improvement of about 20\% in average WIS over the baseline model, with this gap
+narrowing to about 10\% at the largest ahead values, illustrating the difficulty 
+of the forecasting problem.
 

### 2021-10-04 forecast/paper/supplement.Rmd
clean up figure 5, move hub comparisons out

@@
 ```{r upswing-histogram, include=FALSE, fig.width=9, fig.height=4.5}
-# Figure 4
+# Fgure 4
 up_down_df %>% 
@@
 In down and flat periods, the indicators generally improve over the AR, while the are worse during up periods."}
+# mycummean <- function(x) {
+#   isgood <- ! is.na(x)
+#   denom <- cumsum(isgood)
+#   x[!isgood] <- 0
+#   num <- cumsum(x)
+#   y <- num / denom
+#   y[is.na(y)] <- .5 # deal with empties at the beginning
+#   y
+# }
+# 
+# null_classifier <- readRDS(here("data", 
+#                                 "all_signals_wide_as_of_2020-05-18.RDS")) %>%
+#   rename(jhu = `value+0:jhu-csse_confirmed_7dav_incidence_prop`) %>%
+#   select(geo_value, time_value, jhu) %>%
+#   group_by(geo_value) %>%
+#   arrange(time_value) %>%
+#   mutate(lag_value = lag(jhu, 7),
+#          apc = (jhu - lag_value) / lag_value,
+#          hot = case_when(
+#            apc > .25 & lag_value > 30 ~ 1,
+#            apc <= .25 & lag_value > 30 ~ 0),
+#          null_class = mycummean(hot))
+
 hot_udf <- inner_join(
@@
   group_by(forecaster, udf) %>%
-  summarise(m = mean(pred != actual) - .5) %>%
+  summarise(m = (mean(pred != actual) - cutoff) / cutoff,
+            err = mean(pred != actual)) %>%
   ungroup()
@@
   filter(!is.na(actual)) %>%
+  # left_join(null_classifier %>% select(geo_value, time_value, null_class) %>%
+  #             rename(target_end_date = time_value)) %>%
   mutate( # kill prob 0-1 predictions
@@
       value > 1-1e-8 ~ 1-1e-8,
-      TRUE ~ value
-    ),
+      TRUE ~ value),
+    # null_class = case_when(
+    #   null_class < 1e-8 ~ 1e-8,
+    #   null_class > 1-1e-8 ~ 1-1e-8,
+    #   TRUE ~ null_class),
     llike = (actual == 1) * log(value) + (actual == 0) * log(1 - value),
-    nulldev = (actual == 1) * log(.5) + (actual == 0) * log(1 - .5)) %>%
-  group_by(forecaster, udf) %>%
-  summarise(m = mean(llike) - mean(nulldev)) %>%
+    nulldev = (actual == 1) * log(cutoff) + 
+      (actual == 0) * log(1 - cutoff)) %>%
+  group_by(forecaster, udf) %>%
+  summarise(m = (mean(llike) - mean(nulldev)) / abs(mean(nulldev)),
+            mm = mean(llike),
+            mmm = mean(nulldev)) %>%
   ungroup()
@@
   geom_hline(yintercept = 0) +
+  scale_y_continuous(labels = scales::percent_format()) +
   facet_wrap(~err, scales = "free_y") +
@@
 ```{r leading-only, include=FALSE, fig.width=6, fig.height=4}
-# Figure 6 in the manuscript
+# Figure 5 in the manuscript
 df2 %>%
@@
   geom_hline(yintercept = 0) +
-  scale_fill_manual(values = fcast_colors) +
+  scale_fill_manual(values = fcast_colors[1:5]) +
   guides(fill = guide_legend(nrow = 1)) +
@@
 Grey lines correspond to the various teams that submitted during period our evaluation period.
-We have highlighted the COVIDhub-ensemble, which is the official forecast of the CDC."}
+We have highlighted the COVIDhub-ensemble, which is the official forecast of the CDC.", include=FALSE}
 # Same as Fig 3 in the paper
@@
 Grey lines correspond to the various teams that submitted during period our evaluation period.
-We have highlighted the COVIDhub-ensemble, which is the official forecast of the CDC."}
+We have highlighted the COVIDhub-ensemble, which is the official forecast of the CDC.", include=FALSE}
 # Like supplement GeoMean figure, but adding 1

### 2021-10-04 indicators/paper/paper.tex
Update DV backfill claims
The original DV backfill numbers were calculated with code that called
covidcast_signal() repeatedly with different lag arguments. However,
specifying lag = k selects *only* those values issued exactly k days
after their time_value, meaning this only selects those values reissued
at a specific lag. If an estimate was reported once and never reissued,
it would not be included in this calculation, because it would only be
reported at one lag.

The new code uses as_of to get all estimates as of a specific lag, and
avoids this problem. This results in lower median backfil

@@
 processed and errors are corrected; the median relative error between initial
-reports and final values is over 20\% for such data, and only after roughly 35
+reports and final values is over 10\% for such data, and only after roughly 30
 days do estimates typically match finalized values within 5\%. However, the

### 2021-10-04 indicators/code/indicator-paper-figures.Rmd
Switch to DV-CLI instead of CHNG-CLI
Historical CHNG data is currently suspect due to an issue in
reconstructing issue dates of the CHNG signals; rather than relying on
CHNG data here, we use DV-CLI, which gives similar results with less
concern about incorrect issue date information.

@@
 ```{r figure-6-data, message=FALSE, cache=TRUE}
-SOURCES <- c("chng", "fb-survey", "jhu-csse", "jhu-csse")
-SIGNALS <- c("smoothed_outpatient_cli", "smoothed_whh_cmnty_cli",
+SOURCES <- c("doctor-visits", "fb-survey", "jhu-csse", "jhu-csse")
+SIGNALS <- c("smoothed_cli", "smoothed_whh_cmnty_cli",
              "confirmed_incidence_prop", "deaths_incidence_prop")
-NAMES <- c("lag", "CHNG-CLI", "CTIS-CLI-in-community", "Cases", "Deaths")
+NAMES <- c("lag", "DV-CLI", "CTIS-CLI-in-community", "Cases", "Deaths")
 

### 2021-10-04 indicators/paper/paper.tex
Switch to DV-CLI instead of CHNG-CLI
Historical CHNG data is currently suspect due to an issue in
reconstructing issue dates of the CHNG signals; rather than relying on
CHNG data here, we use DV-CLI, which gives similar results with less
concern about incorrect issue date information.

@@
 claims signal evolved as it was revised across multiple issue dates, in four
-different states, between June 1 and August 1, 2020. In each plot, the rightmost
-ends of the lines correspond to estimates for the last day that data are
-available for each issue date, which are generally the most tentative estimates,
-and appear to be significantly biased upward in Arizona in June 2020, and
-significantly biased downward in New York throughout June and July 2020.
+different states, between June 1 and August 1, 2020. This signal, known as
+DV-CLI, is similar to CHNG-CLI and measures outpatient doctor visits about
+COVID-related symptoms, but is provided by a different data partner. In each
+plot, the rightmost ends of the lines correspond to estimates for the last day
+that data are available for each issue date, which are generally the most
+tentative estimates, and appear to be significantly biased upward in Arizona in
+June 2020, and significantly biased downward in New York throughout June and
+July 2020.
 
@@
   \caption{Estimated percentage of doctor's visits due to COVID-like illness
-    displayed across multiple issue dates, with later issue dates adding
-    additional data and revising past data from prior issue dates.}
+    (DV-CLI) displayed across multiple issue dates, with later issue dates
+    adding additional data and revising past data from prior issue dates.}
   \label{fig:dv_as_of}

### 2021-10-04 indicators/paper/paper.tex
Pass over Sect 2D

--- 2021-10-04 indicators/paper/paper.tex
@@
 The revision tracking feature in the API assists in model-building and
-evaluation. Figure~\ref{fig:dv_as_of} illustrates how one COVIDcast medical
-claims signal evolved as it was revised across multiple issue dates, in four
-different states, between June 1 and August 1, 2020. In each plot, the rightmost
-ends of the lines correspond to estimates for the last day that data are
-available for each issue date, which are generally the most tentative estimates,
-and appear to be significantly biased upward in Arizona in June 2020, and
-significantly biased downward in New York throughout June and July 2020.
+evaluation. Figure~\ref{fig:dv_as_of} illustrates how DV-CLI, a medical claims
+signal, evolved as it was revised across multiple issue dates, in four different
+states, between June 1 and August 1, 2020. DV-CLI is similar to CHNG-CLI and
+reflects outpatient visits with COVID-related symptoms (the two signals are
+based on claims data provided by different data partners, which cover different
+hospital systems). In each panel, the rightmost end of each colored line
+corresponds to an estimate for the last day of available data for a given issue
+date, which we can see tends to be significantly biased upward in Arizona in
+June 2020, and significantly biased downward in New York throughout June and
+July 2020.
 
@@
   \includegraphics[width=\columnwidth]{fig/dv_as_of.pdf}
-  \caption{Estimated percentage of doctor's visits due to COVID-like illness
-    displayed across multiple issue dates, with later issue dates adding
-    additional data and revising past data from prior issue dates.}
+  \caption{Estimated percentage of outpatient visits due to COVID-like illness 
+    (DV-CLI) displayed across multiple issue dates, with later issue dates
+    adding additional data and revising past data from prior issue dates.}
   \label{fig:dv_as_of}
@@
 well as cases, the revision pattern is not very systematic, with large
-corrections typically occuring at a sparse subset of locations and times
-(e.g., due to audits or backlogs being cleared, which can result in thousands of
-cases or deaths can be  added or removed all at once).  This backfill is much
-more difficult to predict than that of claims data, thus the latter (and other 
+corrections typically occuring at a sparse subset of locations and times (e.g.,
+due to audits or backlogs being cleared, which can result in thousands of cases
+or deaths can be added or removed all at once).  This backfill is much more
+difficult to predict than that of claims data, therefore the latter (and other
 sources) may be useful for nowcasting cases and deaths while public health
@@
 
-To reiterate a previous point, when training and validating forecast models that
-use signals (on historical data), users will want to use data that was known
-\textit{as of} the forecast date, not revised versions that only became
-available much later. The COVIDcast API makes all historical versions available
-and easily accessible for this purpose; and this feature plays a prominent role
-in our own analysis of forecasting and hotspot prediction models appearing in a
-companion paper \cite{McDonald:2021}.% Revisions to forecast targets, such as 
-% cases and deaths, also complicate the evaluation of forecast models:
-% comparisons of forecasts to the ground truth may shift when the ground truth
-% is revised weeks later. Only by systematically tracking revisions can this
-% effect be monitored.
+To reiterate a previous point, when building forecast models (on historical
+data) for retrospective evaluation purposes, users will want to use data that
+was known \textit{as of} the forecast date, not revised versions that only
+became available at a later time. Note that not only model training, but also
+model assessment, can be affected by the revision process (comparisons of
+forecasts to the ground truth may shift when the ground truth is revised weeks
+later). Only by systematically tracking revisions can all these effects be
+monitored and properly accounted for.  The COVIDcast API makes all historical
+versions available and easily accessible for this purpose; and this feature
+plays a prominent role in our own analysis of forecasting and hotspot prediction
+models appearing in a companion paper \cite{McDonald:2021}.
 

### 2021-10-04 forecast/paper/response-to-reviewers.tex

@@
 
-  $\bullet$ Here are adding two figures
+  $\bullet$ To directly address the Reviewer's point about IHME, we include two
+  figures here
   that compare our forecasts with state-of-the-art forecasts
@@
   submissions. In the second figure, we use the Geometric Mean scaled by the
-  baseline, which should be a more meaningful comparison, as argued elsewhere in
-  the supplement. By this metric, the methods employed in this
+  baseline as shown in Figure S6 in the Supplement. By this metric, the methods employed in this
   paper would rank in the top 2 of all submitted forecasts, beating the

### 2021-10-04 forecast/paper/supplement-text.tex
clean up figure 5, move hub comparisons out

--- 2021-09-27 forecast/paper/supplement-text.tex
@@
 appear to be any particular pattern to these locations.
+
+% \section{Comparisons to state-of-the-art forecasters}
+
+% Since mid-July 2020, a number of forecasting teams (48 in total) have submitted
+% 1--4 week ahead state-level case forecasts to the COVID19 Forecast Hub
+% \cite{ForecastHub}. To compare those forecasts to those created in this paper,
+% we selected only those forecasters who submitted at least 300 forecasts for the
+% 2-week-ahead target before
+% January 1, 2021 (approximately 1 forecast for each state for 6 weeks). This left
+% 32 forecasters, one of which, the COVIDhub-4-week-ensemble, was an exact
+% duplicate of another. To make the comparison fair, we scaled by the
+% COVIDhub-baseline as in Figure 3 of the manuscript. Figure
+% \ref{fig:compare-to-hub-mean} shows all submitted forecasts (available 7, 14,
+% and 21 days ahead) overlaid on Figure 3. We have highlighted the
+% COVIDhub-ensemble in light blue for comparison. All the models developed in this
+% paper beat the baseline, and would rank in the top 8 of 31 submitted forecasts
+% over this period.
+
+% A more fair comparison that directly adjusts for spatiotemporal difficulty is to
+% use the geometric mean scaled by the baseline. This is the same idea shown in
+% \ref{fig:fcast-adjusted}. We have also remade this figure overlaying submitted
+% forecasts available on the Forecast Hub. It occasionally happens that a
+% submitted forecast achieved a WIS of 0, so in this case we use the WIS$+ 1$
+% rather than WIS for all forecasters. The result is shown in
+% \ref{fig:compare-to-hub-geomean}. By this metric, the methods employed in this
+% paper would rank in the top 2, beating the COVIDhub-ensemble at all forecast
+% horizons. The improvement of the best indicator assisted model over the AR model
+% is roughly the same as the improvement of the AR model over the COVIDhub-ensemble.

### 2021-10-04 forecast/paper/supplement.tex

@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{fig/compare-to-hub-mean-1}
-
-}
-
-\caption{This figure reproduces Figure 3 in the main paper but overlays scores for the forecasts submitted to the COVID-19 Forecast Hub.
-Grey lines correspond to the various teams that submitted during period our evaluation period.
-We have highlighted the COVIDhub-ensemble, which is the official forecast of the CDC.}\label{fig:compare-to-hub-mean}
-\end{figure}
-
 \clearpage
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{fig/compare-to-hub-geomean-1}
-
-}
-
-\caption{This figure is similar to Figure \ref{fig:fcast-adjusted}.
-In this case, we add 1 to both the forecaster WIS and the baseline WIS before scaling (to allow forecasters that achieve 0 error to appear), and we overlay scores for the forecasts submitted to the COVID-19 Forecast Hub.
-Grey lines correspond to the various teams that submitted during period our evaluation period.
-We have highlighted the COVIDhub-ensemble, which is the official forecast of the CDC.}\label{fig:compare-to-hub-geomean}
-\end{figure}
 

### 2021-10-04 indicators/paper/supplement.tex

--- 2021-07-12 indicators/paper/supplement-medrxiv.tex
@@
+\documentclass[9pt,twoside,lineno]{pnas-new}
+% Use the lineno option to display guide line numbers if required.
+
+\templatetype{pnassupportinginfo}
+
+\input{../../common/prelim.tex}
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+\title{An Open Repository of Real-Time COVID-19 Indicators}
+\author{Alex Reinhart, Logan Brooks, Maria Jahja, Aaron Rumack, Jingjing Tang,
+  et al.}
+\correspondingauthor{Alex Reinhart.\\E-mail: areinhar@stat.cmu.edu}
+
+\begin{document}
+
+\maketitle
+
 \begin{figure}
@@
 
-\caption{National trends, from April 2020 to April 2021, of HHS-reported confirmed COVID-19 hospital admissions, plus another hospitalization signal in the COVIDcast API.
-Here, the HHS data has been limited to values from 2020-08-01 onward due to changes in reporting behavior, and processed into a 7-day trailing average rate per 100,000 resident population using Census Bureau estimates for 2019.
+\caption{National trends, from August 2020 to August 2021, of HHS-reported confirmed COVID-19 hospital admissions, plus comparison signals in the COVIDcast API.
+(HHS data was not consistently reported before August 2020.) Here, the HHS data has been processed into a 7-day trailing average rate per 100,000 resident population using Census Bureau estimates for 2019.
 HSP-Hosp is the percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners, smoothed in time and adjusted for systematic day-of-week effects (HSP-Hosp is available from an earlier date when working at finer geographical resolutions).
-HSP-Hosp has been scaled to have the same maximum value as the HHS data.}\label{fig:hospitalization_time_trends_national}
+HSP-Hosp, CTIS-CLI-in-community, and CHNG-CLI have been scaled to have the same maximum value as the HHS data.}\label{fig:hospitalization_time_trends_national}
 \end{figure}
@@
 
-\caption{Geo-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated for all times with sufficient available data within this period, over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which both signals are available.}\label{fig:hosp-correlations-by-time}
+\caption{Geo-wise correlations with hospitalization rates derived from HHS data, from August 15, 2020 to August 15, 2021, calculated for all times with sufficient available data within this period, over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which all signals are available.}\label{fig:hosp-correlations-by-time}
 \end{figure}
@@
 
-\caption{Time-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which both signals are available.}\label{fig:hosp-correlations-by-state}
+\caption{Time-wise correlations with hospitalization rates derived from HHS data, from August 15, 2020 to August 15, 2021, calculated over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which all signals are available.}\label{fig:hosp-correlations-by-state}
 \end{figure}
+
+\FloatBarrier
+
+\bibliography{../../common/covidcast.bib,pnas-materials/pnas-sample.bib}
+
+\end{document}

### 2021-10-05 forecast/paper/paper.tex
Pass over paper
- Mostly just small copyedits
- Moved reference to the supplement's new statistical significance
  tests the resuls section

@@
 While forecasting carries a rich literature that offers a wide range of
-techniques, see e.g., \cite{Hyndman:2018}, we purposely constrain ourselves to
+techniques, see, e.g., \cite{Hyndman:2018}, we purposely constrain ourselves to 
 very simple models, avoiding common enhancements such as order selection,
 correction of outliers/anomalies in the data, and inclusion of regularization or
-nonlinearities.
-% Addition in response to Reviewer 1
-For the same reason, we do not account for any environmental factors that may
-well aid in forecasting---age structure, variants, holidays,
-reporting patterns, vaccination, public health measures, etc.
-% End addition
-That said, analyses of forecasts submitted to the COVID-19
-Forecast Hub \cite{ForecastHub} by a large community of modelers have shown that
-simple, robust models have consistently been among the best-performing over the
-pandemic, including time series models similar to those we
-consider in what follows.
-% Added in response to Reviewer 2
-\cite{Cramer:2021} investigates the performance of submitted forecasts for
-state-level deaths and shows that time series models perform among the top 5.
-Our internal evaluations suggest that the same is true (at least, if not to a
-greater extent) for case forecasting.
+nonlinearities. For the same reason, we do not account for other factors that
+may well aid in forecasting, such as age-specific effects, holiday adjustments,
+public health mandates, and so on. That said, analyses of forecasts submitted
+to the COVID-19 Forecast Hub \cite{ForecastHub} by a large community of modelers 
+have shown that simple, robust models have consistently been among the
+best-performing over the pandemic, including time series models similar the ones
+we use in this paper. See, e.g., \cite{Cramer:2021}, for a detailed study
+of death forecasting. Our internal evaluations suggest that the same is true (at
+least, if not to a greater extent) for case forecasting.
 
@@
 conclusions with respect to numerous modes of analysis.
-% Addition upon review
-In the Supplement we examine
-two assumption-lean
-methods for assessing the statistical significance of our main results.
 
@@
 
-\paragraph{Dynamic Re-Training}
+\paragraph{Dynamic Retraining}
 
@@
 when the survey was first launched). The end date was chosen again with a 
-consideration to align both tasks as best as possible,
-% added to emphasize 
-and to coincide with the
-introduction of vaccinations.
-% Removed: this is no longer true unfortunately.
-% because few hotspots
-% exist post December 31, 2020, due to the general and gradual decline of the
-% pandemic in 2021.
+consideration to align both tasks as best as possible, and because very few
+hotspots exist for the several few months of 2021, due to the general and
+gradual decline of the pandemic during this period.
 
@@
 quantile models in \eqref{eq:forecast_ar} and \eqref{eq:forecast_ar_x} using
-\textit{weighted interval score} (WIS), a quantile-based scoring rule; see e.g.,
-\cite{Gneiting:2007}.  WIS is a proper score, which means that its expectation 
-is minimized by the population quantiles of the target variable.  The use of WIS
-in COVID-19 forecast scoring is discussed in \cite{Bracher:2021}; WIS is also
-the main evaluation metric used in the COVID-19 Forecast Hub. More broadly, the
-specific form of WIS used here is a standard metric in the forecasting community
-for evaluating quantile-based probabilistic forecasts, just as mean squared
-forecast error is standard for point forecasts.
+\textit{weighted interval score} (WIS), a quantile-based scoring rule; see,
+e.g., \cite{Gneiting:2007}.  WIS is a proper score, which means that its
+expectation is minimized by the population quantiles of the target variable.
+The use of WIS in COVID-19 forecast scoring is discussed in \cite{Bracher:2021};
+WIS is also the main evaluation metric used in the COVID-19 Forecast Hub. More
+broadly, the specific form of WIS used here is a standard metric in the
+forecasting community for evaluating quantile-based probabilistic forecasts,
+just as mean squared forecast error is standard for point forecasts.
 
@@
 \eqref{eq:forecast_ar_x} that sets $X_{\ell,t}$ to be the value of the CHNG-CLI
-indicator at location $\ell$ and time $t$.
-% fits the coefficients based on the training data, and makes its predictions by
-% setting $X_{\ell,t}$ to the appropriate CHNG-CLI values.  
-Finally, we use the term ``indicator model'' to refer to any one of the ten
-models of the form \eqref{eq:forecast_ar_x} or \eqref{eq:hotspot_ar_x} (five
-from each of the forecasting and hotspot prediction tasks).
+indicator at location $\ell$ and time $t$.  Finally, we use the term ``indicator
+model'' to refer to any one of the ten models of the form
+\eqref{eq:forecast_ar_x} or \eqref{eq:hotspot_ar_x} (five from each of the
+forecasting and hotspot prediction tasks).
 
@@
   indicator models tend to be better than AR when case rates are flat or
-  decreasing (most notable in CHNG-COVID and CTIS-CLI-in-community), but they
-  are worse than AR when case rates are increasing (this is most notable in
-  CHNG-CLI and DV-CLI). More rarely does an indicator model tend to beat AR when
-  case rates are increasing, but there appears to be some evidence of this for
-  the Google-AA model. 
+  decreasing (most notable in CHNG-COVID and CTIS-CLI-in-community), but can be
+  worse than AR when case rates are increasing (this is most notable in CHNG-CLI
+  and DV-CLI). More rarely does an indicator model tend to beat AR when case
+  rates are increasing, but there appears to be some evidence of this for the
+  Google-AA model.  
   % In an increasing period, when an indicator model is worse than AR, it tends
@@
 proper vintage data, we ran a parallel set of forecasting and hotspot prediction
-experiments using finalized data. The results, given in the supplement, show
-that training and making predictions on finalized data can result in overly
-optimistic estimates of true test-time performance (up to 10\% better in terms
-of average WIS or AUC). Furthermore, since indicators can have greatly different
-backfill profiles, the use of finalized data in retrospective evaluations
-changes the relative ranking of models.  For example, CHNG-CLI and DV-CLI,
-when trained on finalized data, perform very similarly in forecasting.
+experiments using finalized data. The results, given in the Supplementary
+Materials, show that training and making predictions on finalized data can
+result in overly optimistic estimates of true test-time performance (up to 10\%
+better in terms of average WIS or AUC). Furthermore, since indicators can have
+greatly different backfill profiles, the use of finalized data in retrospective
+evaluations changes the relative ranking of models.  For example, CHNG-CLI and
+DV-CLI, when trained on finalized data, perform very similarly in forecasting.
 This makes sense since they are both claims-based indicators that are supposedly
-measuring the same thing.  However, DV-CLI outperforms CHNG-CLI on vintage
-data, reflecting its has a less severe backfill profile.
+measuring the same thing.  However, DV-CLI outperforms CHNG-CLI on vintage data,
+reflecting its has a less severe backfill profile.
+
+The Supplement provides a number of other additional analyses; for example, we
+examine two assumption-lean methods for assessing the statistical significance
+of our main results.
 
@@
 \begin{figure*}[t]
-  \includegraphics[width=\textwidth]{fig/fcast-hot-combo-1.pdf}
+  \centering
+  \includegraphics[width=0.95\textwidth]{fig/fcast-hot-combo-1.pdf}
   \caption{Main results for both tasks. Left: average WIS for each forecast
@@
 this explanation is twofold: (1) the CHNG-CLI model benefits from a more robust
-method of aggregating WIS (geometric mean; shown in the supplement); and (2)
-when we train and make predictions on finalized data, it handily beats AR, on
-par with the best-performing models (also shown in the supplement).
+method of aggregating WIS (geometric mean; shown in the Supplementary
+Materials); and (2) when we train and make predictions on finalized data, it
+handily beats AR, on par with the best-performing models (also shown in the
+Supplement).
 
@@
 prediction, again stratified by ahead value and averaged over all HRRs and
-prediction dates.  We can see many similarities to the forecasting results (now
-larger AUC is better). For example, CHNG-COVID and DV-CLI offer the biggest
-improvement over AR, and all models, including AR, degrade in performance
-towards the baseline (in this context, a classifier based on random guessing,
-which achieves an AUC of 0.5) as the ahead values grow, illustrating the
-difficulty of the hotspot prediction problem. A clear difference, however, is
-that the CHNG-CLI model performs quite well in hotspot prediction, close to the
-best-performing indicator models for many of the ahead values. This may be
+prediction dates.  We can see many similarities to the forecasting results
+(note: now larger AUC is better). For example, CHNG-COVID and DV-CLI offer the
+biggest improvement over AR, and all models, including AR, degrade in
+performance towards the baseline (in this context, a classifier based on random
+guessing, which achieves an AUC of 0.5) as the ahead values grow, illustrating
+the difficulty of the hotspot prediction problem. A clear difference, however,
+is that the CHNG-CLI model performs quite well in hotspot prediction, close to
+the best-performing indicator models for many of the ahead values. This may be
 because volatility in the CHNG-CLI indicator plays less of a role in the
@@
 (bootstrap samples from a signal's history).  The results, shown and explained
-more precisely in the supplement, are vastly different (worse) than the original
-set of results.  In both forecasting and hotspot prediction, the ``fake''
-indicator models offered essentially no improvement over the pure AR model,
-which strongly rejects (informally speaking) the implicit regularization
+more precisely in the Supplementary Materials, are vastly different (worse) than
+the original set of results.  In both forecasting and hotspot prediction, the
+``fake'' indicator models offered essentially no improvement over the pure AR
+model, which strongly rejects (informally speaking) the implicit regularization
 hypothesis.
@@
 
-The supplement pursues this analysis further.  For example, we examine
+The Supplement pursues this analysis further.  For example, we examine
 classification accuracy and log-likelihood for the hotspot task and find a

### 2021-10-05 forecast/paper/supplement-text.tex
Pass over supplement
- Just formatting and copyedits
- Comment out all eval results from 2021 (but I can't seem to
  compile the Rmd because I'm missing the data files)

@@
-\section{Examining the relative advantage of using finalized rather than vintage data}
+\section{Finalized Versus Vintage Data}
 
-The goal of this section is to quantify the effect of
-not properly accounting for the question of ``what was known when'' in
-performing retrospective evaluations of forecasters.   Figures
-\ref{fig:fcast-finalized} and \ref{fig:hot-finalized} show 
-what Figures 3 and 4 in the main paper would have looked like if we
-had simply trained all models using the finalized data rather than
-using vintage data.  This comparison can be seen more
-straightforwardly in Figures \ref{fig:fcast-honest-v-finalized} and
-\ref{fig:hot-honest-v-finalized}, which show the ratio in performance between
-the vintage and finalized 
-versions.  When methods are 
-given the finalized version of the data rather than the version available at the
-time that the forecast would have been made, all methods 
-appear (misleadingly) to have better performance than they would
-have had if run prospectively.  For example, for forecasting case rates
-7-days ahead, the WIS of all methods is at least 8\% larger than what would have been
-achieved using finalized data.  This effect
-diminishes as the forecasting horizon increases, reflecting the fact
-that longer-horizon forecasters rely less heavily on recent data than very
-short-horizon forecasters.  Crucially, some methods are
-``helped'' more than others by the less scrupulous retrospective
-evaluation, underscoring the difficulty of avoiding misleading
-conclusions when performing retrospective evaluations of forecasters.
+The goal of this section is to quantify the effect of not properly accounting
+for the question of ``what was known when'' in performing retrospective
+evaluations of forecasters.  Figures \ref{fig:fcast-finalized} and
+\ref{fig:hot-finalized} show what Figures 3 and 4 in the main paper would have
+looked like if we had simply trained all models using the finalized data rather
+than using vintage data.  This comparison can be seen more straightforwardly in
+Figures \ref{fig:fcast-honest-v-finalized} and \ref{fig:hot-honest-v-finalized},
+which show the ratio in performance between the vintage and finalized versions.
+When methods are given the finalized version of the data rather than the version
+available at the time that the forecast would have been made, all methods appear
+(misleadingly) to have better performance than they would have had if run
+prospectively.  For example, for forecasting case rates 7-days ahead, the WIS of
+all methods is at least 8\% larger than what would have been achieved using
+finalized data.  This effect diminishes as the forecasting horizon increases,
+reflecting the fact that longer-horizon forecasters rely less heavily on recent
+data than very short-horizon forecasters.  Crucially, some methods are
+``helped'' more than others by the less scrupulous retrospective evaluation,
+underscoring the difficulty of avoiding misleading conclusions when performing
+retrospective evaluations of forecasters.
 
-\chngcli~(and, to a lesser extent, the other claims-based
-signals) is the most  
-affected by this distinction, reflecting the latency in claims-based
-reporting.  This underscores the importance of efforts to provide ``nowcasts'' for
-claims signals 
-(which corresponds to a 0-ahead forecast of what the claims
-signal's value will be once all data has been collected). Looking at
-the \chngcli~and \dv~curves in Figure \ref{fig:fcast-finalized}, we
-can see that they perform very similarly when trained on the finalized data.
-This is reassuring 
-because they are, in principle, measuring the same thing (namely, the percentage of
-outpatient visits that are primarily about COVID-related symptoms).
-The substantial difference in their curves in Figure 3 of the
-main paper must therefore reflect their having very different
-backfill profiles.
+\chngcli~(and, to a lesser extent, the other claims-based signals) is the most
+affected by this distinction, reflecting the latency in claims-based reporting.
+This underscores the importance of efforts to provide ``nowcasts'' for claims
+signals (which corresponds to a 0-ahead forecast of what the claims signal's
+value will be once all data has been collected). Looking at the \chngcli~and
+\dv~curves in Figure \ref{fig:fcast-finalized}, we can see that they perform
+very similarly when trained on the finalized data.  This is reassuring because
+they are, in principle, measuring the same thing (namely, the percentage of
+outpatient visits that are primarily about COVID-related symptoms).  The
+substantial difference in their curves in Figure 3 of the main paper must
+therefore reflect their having very different backfill profiles.
 
-While using finalized rather than vintage data affects \dv~the least for forecasting,
-it is one of the most affected methods for the
-hotspot problem. 
+While using finalized rather than vintage data affects \dv~the least for
+forecasting, it is one of the most affected methods for the hotspot problem.
 This is a reminder that the forecasting and hotspot problems are fundamentally
-distinct.  For example, the hotspot problem does not measure
-the ability to distinguish between flat and downward trends.
+distinct.  For example, the hotspot problem does not measure the ability to
+distinguish between flat and downward trends.
 
-Even the \ar~model is affected by this distinction, reflecting the
-fact that the case rates themselves (i.e., the response values) are also
-subject to revision.  The forecasters based on indicators are thus
-affected both by revisions to the indicators and by revisions to the
-case rates.  In the case of the \gs~model, in which we only used finalized
-values for the \gs~indicator, the difference in performance can be
-wholly attributed to revisions of case rates.
+Even the \ar~model is affected by this distinction, reflecting the fact that the
+case rates themselves (i.e., the response values) are also subject to revision.
+The forecasters based on indicators are thus affected both by revisions to the
+indicators and by revisions to the case rates.  In the case of the \gs~model, in
+which we only used finalized values for the \gs~indicator, the difference in
+performance can be wholly attributed to revisions of case rates.
 
-\section{Aggregating with geometric mean}
+\section{Aggregation with Geometric Mean}
 
-In this section, we consider using the geometric mean instead of the
-arithmetic mean when aggregating the weighted interval score (WIS) across
-location-time pairs. 
-There are three reasons why using the geometric mean may be desirable.
+In this section, we consider using the geometric mean instead of the arithmetic
+mean when aggregating the weighted interval score (WIS) across location-time
+pairs.  There are three reasons why using the geometric mean may be desirable.
 \begin{enumerate}
-\item  WIS is right-skewed, being bounded below by zero and having
-  occasional very large values.   Figure \ref{fig:wis-densities} illustrates that the
-  densities appear roughly log-Gaussian.  The geometric mean is a
-  natural choice in such a context since the relative ordering of forecasters is
-  determined by the arithmetic mean of the {\em logarithm} of their WIS
-  values.
-\item In the main paper, we report the ratio of the mean WIS of a
-  forecaster to the mean WIS of the baseline forecaster. Another
-  choice could be to take the mean of the ratio of WIS values for the
-  two methods. This latter choice would penalize a method less for
-  doing poorly where the baseline forecaster also does poorly.
-  Using instead the geometric mean makes the order of aggregation and
-  scaling immaterial since the ratio of geometric means is the same as
-  the geometric mean of ratios.
-\item If one imagines that a forecaster's WIS is composed of
-  multiplicative space-time effects $S_{\ell,t}$ shared across all forecasters,
-  i.e. $\WIS(F_{\ell,t,f},Y_{\ell,t})=S_{\ell,t}E_{f,t}$ with $E_{f,t}$ a
-  forecaster-specific error, then
-  taking the ratio of two forecasters' geometric mean WIS values will
-  effectively cancel these space-time effects.
+\item WIS is right-skewed, being bounded below by zero and having occasional 
+  very large values.  Figure \ref{fig:wis-densities} illustrates that the
+  densities appear roughly log-Gaussian.  The geometric mean is a natural choice  
+  in such a context since the relative ordering of forecasters is determined by
+  the arithmetic mean of the {\em logarithm} of their WIS values.
+\item In the main paper, we report the ratio of the mean WIS of a forecaster to
+  the mean WIS of the baseline forecaster. Another choice could be to take the
+  mean of the ratio of WIS values for the two methods. This latter choice would
+  penalize a method less for doing poorly where the baseline forecaster also
+  does poorly.  Using instead the geometric mean makes the order of aggregation
+  and scaling immaterial since the ratio of geometric means is the same as the 
+  geometric mean of ratios.
+\item If one imagines that a forecaster's WIS is composed of multiplicative
+  space-time effects $S_{\ell,t}$ shared across all forecasters,
+  i.e., $\WIS(F_{\ell,t,f},Y_{\ell,t})=S_{\ell,t}E_{f,t}$ with $E_{f,t}$ a
+  forecaster-specific error, then taking the ratio of two forecasters' geometric
+  mean WIS values will effectively cancel these space-time effects. 
 \end{enumerate}
 
-Figure \ref{fig:fcast-adjusted} uses the geometric
-mean for aggregation.  Comparing this with Figure 3 of the main paper,
-we see that the main conclusions are largely unchanged; however,
-\chngcli~now appears better than \ar.  This behavior would be
-expected if \chngcli's poor performance is attributable to a
-relatively small number of large errors (as opposed to a large number
-of moderate errors).  Indeed, Figure 5 of the main paper further
-corroborates this, in which we see the heaviest left tails occuring
-for \chngcli.
+Figure \ref{fig:fcast-adjusted} uses the geometric mean for aggregation.
+Comparing this with Figure 3 of the main paper, we see that the main conclusions
+are largely unchanged; however, \chngcli~now appears better than \ar.  This
+behavior would be expected if \chngcli's poor performance is attributable to a
+relatively small number of large errors (as opposed to a large number of
+moderate errors).  Indeed, Figure 5 of the main paper further corroborates this,
+in which we see the heaviest left tails occuring for \chngcli.
 
-\section{Statistical significance}
+\section{Statistical Significance}
 
@@
 expect to see a uniform distribution. However, for all indicator-assisted
-models, we see many more small $P$-values than would be expected if the null
+models, we see many more small p-values than would be expected if the null
 hypothesis (that the AR model is better) were true.
@@
 the manuscript) and the geometric mean relative to the strawman as described
-above. The first two rows of Table \ref{tab:dm-test} displays $P$-values for the
+above. The first two rows of Table \ref{tab:dm-test} displays p-values for the
 test that the indicator-assisted model is no-better than the AR model. Only the
-CHNG-CLI model's $P$-value exceeds conventional statistical significance
+CHNG-CLI model's p-value exceeds conventional statistical significance
 thresholds.
 
-\section{Bootstrap results}
+\section{Bootstrap Results}
 
-As explained in Section 2.B. of the main paper, a (somewhat cynical)
-hypothesis for why we see benefits in forecasting and hotspot
-prediction is that the indicators are not actually providing useful
-information but they are instead acting as a sort of ``implicit
-regularization,''  leading to shrinkage on the autoregressive
-coefficients and therefore to less volatile predictions.  To investigate
-this hypothesis, we consider fitting  ``noise features'' that in truth
-should have zero coefficients.  Recall (from the main paper) that at each
-forecast date, we 
-train a model on 6,426 location-time pairs.  Indicator models are
-based on six features, corresponding to the three autoregressive terms
-and the three lagged indicator values.  To form noise indicator features,
-we replace their values with those from a randomly chosen time-space pair
-(while keeping the autoregressive 
-features fixed).  In particular, at each location $\ell$ and
-time $t$, for the forecasting task we replace the triplet $(X_{\ell,t}, X_{\ell,t-7},
-X_{\ell,t-14})$ in Eq. (3) of the main paper with the triplet $(X_{\ell^*,t^*},
-X_{\ell^*,t^*-7}, 
-X_{\ell^*,t^*-14})$, where $(\ell^*,t^*)$ is a location-time pair
-sampled with replacement from the 6,426 location-time pairs.  
-Likewise in the hotspot prediction task, we replace the triplet
-$(X_{\ell,t}^\Delta, X_{\ell,t-7}^\Delta, 
-X_{\ell,t-14}^\Delta)$ in Eq. (5) of the main paper with
-$(X_{\ell^*,t^*}^\Delta, X_{\ell^*,t^*-7}^\Delta, 
-X_{\ell^*,t^*-14}^\Delta)$.
-Figures
-\ref{fig:fcast-booted}--\ref{fig:hot-booted} show the results.  No
-method exhibits a noticeable performance gain over the \ar~method,
-leading us to dismiss the implicit regularization hypothesis.
+As explained in Section 2.B. of the main paper, a (somewhat cynical) hypothesis
+for why we see benefits in forecasting and hotspot prediction is that the
+indicators are not actually providing useful information but they are instead
+acting as a sort of ``implicit regularization,'' leading to shrinkage on the
+autoregressive coefficients and therefore to less volatile predictions.  To
+investigate this hypothesis, we consider fitting ``noise features'' that in
+truth should have zero coefficients.  Recall (from the main paper) that at each
+forecast date, we train a model on 6,426 location-time pairs.  Indicator models
+are based on six features, corresponding to the three autoregressive terms and
+the three lagged indicator values.  To form noise indicator features, we replace
+their values with those from a randomly chosen time-space pair (while keeping
+the autoregressive features fixed).  In particular, at each location $\ell$ and
+time $t$, for the forecasting task we replace the triplet $(X_{\ell,t},
+X_{\ell,t-7}, X_{\ell,t-14})$ in Eq. (3) of the main paper with the triplet
+$(X_{\ell^*,t^*}, X_{\ell^*,t^*-7}, X_{\ell^*,t^*-14})$, where $(\ell^*,t^*)$ is
+a location-time pair sampled with replacement from the 6,426 location-time
+pairs.  Likewise in the hotspot prediction task, we replace the triplet
+$(X_{\ell,t}^\Delta, X_{\ell,t-7}^\Delta, X_{\ell,t-14}^\Delta)$ in Eq. (5) of
+the main paper with $(X_{\ell^*,t^*}^\Delta, X_{\ell^*,t^*-7}^\Delta,
+X_{\ell^*,t^*-14}^\Delta)$.  Figures
+\ref{fig:fcast-booted}--\ref{fig:hot-booted} show the results.  No method
+exhibits a noticeable performance gain over the \ar~method, leading us to
+dismiss the implicit regularization hypothesis.
 
@@
 In this section we provide extra details about the upswing / flat / downswing
-analysis described in the main text. Figure \ref{fig:upswing-summary} shows
-the overall results, examining the average difference $\WIS(AR) - \WIS(F)$ in 
+analysis described in the main text. Figure \ref{fig:upswing-summary} shows the
+overall results, examining the average difference $\WIS(AR) - \WIS(F)$ in
 period. Figure \ref{fig:hotspots-upswing-downswing} shows the same information
 for the hotspot task. On average, during downswings and flat periods, the
-indicator-assisted models have lower classification error and higher 
-log likelihood than the AR model. For hotspots, both Google-AA and CTIS-CLIIC
+indicator-assisted models have lower classification error and higher log
+likelihood than the AR model. For hotspots, both Google-AA and CTIS-CLIIC
 perform better than the AR model during upswings, in contrast to the forecasting
 task, where only Google-AA improves. For a related analysis, Figure
-\ref{fig:cor-wis-ratio} shows histograms of the Spearman
-correlation (Spearman's $\rho$, a rank-based measure of association) between
-the $\WIS(F)/\WIS(AR)$ and the magnitude of the swing. Again we see that case
-rate increases are positively related to diminished performance of the 
-indicator models.
+\ref{fig:cor-wis-ratio} shows histograms of the Spearman correlation (Spearman's
+$\rho$, a rank-based measure of association) between the $\WIS(F)/\WIS(AR)$ and
+the magnitude of the swing. Again we see that case rate increases are positively
+related to diminished performance of the indicator models.
 
-One hypothesis for diminished relative performance during upswings is that 
-the AR model tends to overpredict downswings and underpredict upswings. Adding
+One hypothesis for diminished relative performance during upswings is that the
+AR model tends to overpredict downswings and underpredict upswings. Adding
 indicators appears to help avoid this behavior on the downswing but not as much
@@
 between $\WIS(AR) - \WIS(F)$ and the difference of their median forecasts.
-During 
-downswings, this correlation is large, implying that improved relative
-performance 
-of $F$ is related to making lower forecasts than the AR model. The opposite is
-true during upswings. This is largely to be expected. However, the relationship
-attenuates in flat periods and during upswings. That is, when performance is
-better 
-in those cases, it may be due to other factors than simply making predictions
-in the correct direction, for example, narrower confidence intervals.
+During downswings, this correlation is large, implying that improved relative
+performance of $F$ is related to making lower forecasts than the AR model. The
+opposite is true during upswings. This is largely to be expected. However, the
+relationship attenuates in flat periods and during upswings. That is, when
+performance is better in those cases, it may be due to other factors than simply
+making predictions in the correct direction, for example, narrower confidence
+intervals.
 
-\section{Leadingness and laggingness}
+\section{Leadingness and Laggingness}
 
@@
 of leadingness or laggingness at location $\ell$, we use the cross correlation
-function (CCF) 
-between the two time series. The $\CCF_\ell(a)$ of an indicator $X_{\ell}$ and case
-rates $Y_{\ell}$ is defined as
+function (CCF) between the two time series. The $\CCF_\ell(a)$ of an indicator
+$X_{\ell}$ and case rates $Y_{\ell}$ is defined as
 % \[
@@
 % standard deviation of $X_{\ell,t}$ (respectively $Y_{\ell,t}$). The result is a
-their Pearson correlation where $X_\ell$ has been
-% sequence of Pearson correlations for each selected $a$, where we
-aligned
-with the values of $Y_{\ell}$ that occurred $a$ days earlier.
-Thus, for any $a>0$, $\CCF_{\ell}(a)>0$ indicates that
-$Y_{\ell,t}$ is moving together with $X_{\ell,t+a}$.
-% days in the future.
-In this case we say that $X_{\ell}$ is lagging
-$Y_{\ell}$. For $a<0$, $\CCF_{\ell}(a)>0$ means that $Y_{\ell,t}$ is
-positively correlated with $X_{\ell,t-a}$, so we say that
-$X_{\ell}$ leads $Y_{\ell}$.
+their Pearson correlation where $X_\ell$ has been aligned with the values of
+$Y_{\ell}$ that occurred $a$ days earlier.  Thus, for any $a>0$,
+$\CCF_{\ell}(a)>0$ indicates that $Y_{\ell,t}$ is moving together with
+$X_{\ell,t+a}$.  In this case we say that $X_{\ell}$ is lagging $Y_{\ell}$. For
+$a<0$, $\CCF_{\ell}(a)>0$ means that $Y_{\ell,t}$ is positively correlated with
+$X_{\ell,t-a}$, so we say that $X_{\ell}$ leads $Y_{\ell}$.
 
@@
 September. These are the same signals shown in Figure 1 in the manuscript but
-using finalized data. To
-define ``leadingness'' we compute $\CCF_{\ell}(a)$ (as implemented with the R
-function {\tt ccf()}) for each $a\in
-\{-15,\ldots,15\}$ using the 56 days leading up to the target date. This is the same
-amount of data used to train the forecasters: 21 days of training data, 21 days
-to get the response at $a=21$, and 14 days for the longest lagged value. The
-orange dashed horizontal line represents the 95\% significance threshold for
-correlations based on 56 observations. Any correlations larger in magnitude than
-this value are considered statistically significant under the null hypothesis of
-no relationship. We define leadingness to be the sum of the significant
+using finalized data. To define ``leadingness'' we compute $\CCF_{\ell}(a)$ (as
+implemented with the R function {\tt ccf()}) for each $a\in \{-15,\ldots,15\}$
+using the 56 days leading up to the target date. This is the same amount of data
+used to train the forecasters: 21 days of training data, 21 days to get the
+response at $a=21$, and 14 days for the longest lagged value. The orange dashed
+horizontal line represents the 95\% significance threshold for correlations
+based on 56 observations. Any correlations larger in magnitude than this value
+are considered statistically significant under the null hypothesis of no
+relationship. We define leadingness to be the sum of the significant
 correlations that are leading (those above the dashed line with $a<0$) while
@@
 significant correlations on the ``leading'' side (at $a = -5, -4, -3$), so
-leadingness will be the sum of those values while laggingness is 0:
-%correlations, so both scores will be 0:
-on September 28 in Charlotte, \dv~is leading cases
-leading but not lagging.
+leadingness will be the sum of those values while laggingness is 0: on September
+28 in Charlotte, \dv~is leading cases leading but not lagging.
 
@@
 difference in indicator WIS and AR WIS. Unlike leadingness (Figure 5 in the
-manuscript) there is no obvious relationship that holds consistently
-across indicators. This is heartening as laggingness
-should not aid forecasting performance. On the other hand, if an indicator is
-more lagging than it is leading, this may suggest diminished performance.
+manuscript) there is no obvious relationship that holds consistently across
+indicators. This is heartening as laggingness should not aid forecasting
+performance. On the other hand, if an indicator is more lagging than it is
+leading, this may suggest diminished performance.
 Figure~\ref{fig:diff-in-lead-lag} shows the correlation of the difference in
@@
 In calculating the CCF and the associated leadingness and laggingness scores, we
-have used the finalized data, and we look at the behavior at the 
-target date of the forecast. That is we are using the same data to
-evaluate predictive accuracy as to determine
-leadingness and laggingness. It should be noted that the leadingness of the
-indicator at the time the model is trained may also be important. Thus, we could
-calculate separate leadingness and laggingness scores for the trained model and
-for the evaluation data and examine their combination in some way. We do not
-pursue this combination further and leave this investigation for future work.
+have used the finalized data, and we look at the behavior at the target date of
+the forecast. That is we are using the same data to evaluate predictive accuracy
+as to determine leadingness and laggingness. It should be noted that the
+leadingness of the indicator at the time the model is trained may also be
+important. Thus, we could calculate separate leadingness and laggingness scores
+for the trained model and for the evaluation data and examine their combination
+in some way. We do not pursue this combination further and leave this
+investigation for future work.
 
-\section{Examining data in 2021}
+% \section{Examining data in 2021}
 
-In this section, we investigate the sensitivity of the results to the
-period over which we train and evaluate the models.  In the main
-paper, we end all evaluation on December 31, 2020.  Figures
-\ref{fig:fcast-alldates}--\ref{fig:hot-alldates} show how the
-results would differ if we extended this analysis through March
-  31, 2021. Comparing Figure \ref{fig:fcast-alldates} to Figure 3 of
-the main paper, one sees that as ahead increases most methods now improve
-relative to the baseline forecaster. When compared to other methods, 
-\chngcli~appears much better than
-it had previously; however, all forecasters other than \chngcov~and
-\dv~are performing less well relative to the baseline than before.
-These changes are likely due to the differing nature of the pandemic
-in 2021, with flat and downward trends much more common than upward
-trajectories.  Indeed, the nature of the hotspot prediction problem is
-quite different in this period.  With a 21-day training window, it is
-common for there to be many fewer hotspots in training.
+% In this section, we investigate the sensitivity of the results to the
+% period over which we train and evaluate the models.  In the main
+% paper, we end all evaluation on December 31, 2020.  Figures
+% \ref{fig:fcast-alldates}--\ref{fig:hot-alldates} show how the
+% results would differ if we extended this analysis through March
+%   31, 2021. Comparing Figure \ref{fig:fcast-alldates} to Figure 3 of
+% the main paper, one sees that as ahead increases most methods now improve
+% relative to the baseline forecaster. When compared to other methods, 
+% \chngcli~appears much better than
+% it had previously; however, all forecasters other than \chngcov~and
+% \dv~are performing less well relative to the baseline than before.
+% These changes are likely due to the differing nature of the pandemic
+% in 2021, with flat and downward trends much more common than upward
+% trajectories.  Indeed, the nature of the hotspot prediction problem is
+% quite different in this period.  With a 21-day training window, it is
+% common for there to be many fewer hotspots in training.
 
-\section{Examining accumulated error over time and space}
+\section{Aggregation Over Time and Space}
 
 Following the suggestion of an anonymous reviewer, we investigate two
-disaggregated versions of 
-the main forecasting result shown in Figure 3 on the manuscript. The first
-(Figure \ref{fig:cumulative-mean})
-displays the cumulative error (summed over all horizons) of each forecaster
-divided by the cumulative error 
-of the baseline.
-This perspective should illustrate how
-the models perform over time, drawing attention to any nonstationary behavior.
-During the initial increase in cases in July 2020, CTIS-CLIIC and Google-AA gain
-a lot relative to the AR model. All the indicators do much better than the AR
-model during the following downturn (the ebb of the second wave). The AR model
+disaggregated versions of the main forecasting result shown in Figure 3 on the
+manuscript. The first (Figure \ref{fig:cumulative-mean}) displays the cumulative
+error (summed over all horizons) of each forecaster divided by the cumulative
+error of the baseline.  This perspective should illustrate how the models
+perform over time, drawing attention to any nonstationary behavior.  During the
+initial increase in cases in July 2020, CTIS-CLIIC and Google-AA gain a lot
+relative to the AR model. All the indicators do much better than the AR model
+during the following downturn (the ebb of the second wave). The AR model
 actually improves over the indicators in October 2020, before losing a bit in

### 2021-10-05 forecast/paper/supplement.Rmd
Pass over supplement
- Just formatting and copyedits
- Comment out all eval results from 2021 (but I can't seem to
  compile the Rmd because I'm missing the data files)

@@
 <!-- Examining data in 2021 -->
-
+<!--
 ```{r fcast-alldates, fig.cap="Forecast performance over all periods. Performance largely improves for all forecasters with the inclusion of data in 2021."}
@@
 ```
+-->
 

### 2021-10-05 forecast/paper/supplement.tex
recompile supplement without 2021 data

@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{fig/fcast-alldates-1}
-
-}
-
-\caption{Forecast performance over all periods. Performance largely improves for all forecasters with the inclusion of data in 2021.}\label{fig:fcast-alldates}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{fig/fcast-alldates-adjusted-1}
-
-}
-
-\caption{Forcast performance over all periods aggregaged with the geometric mean. Again, the inclusion of data in 2021 leads to improved performance.}\label{fig:fcast-alldates-adjusted}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{fig/hot-alldates-1}
-
-}
-
-\caption{Area under the curve for hotspot predictions including data in 2021.
-Performance degrades relative to the period in 2020.
-However, there are far fewer hotspots during this period as case rates declined in much of the country.}\label{fig:hot-alldates}
-\end{figure}
-
 \clearpage

### 2021-10-06 indicators/paper/supplement.tex
Add lagged case correlation plot to appendix

@@
 COVIDcast Epidata API.}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/case-correlation-lagged-plot-1}
+
+}
+
+\caption{Correlation between confirmed cases and case reports 1, 2, or 3 weeks prior, for all counties in the United States with case reporting.
+Cases are correlated with future cases as one might expect, but note the precipitous drop in correlation in February 2021.
+This matches the correlation drop between other COVIDcast signals and cases during the same time period, supporting the hypothesis that the drop was due to decreased heterogeneity in case rates by county.}\label{fig:case-correlation-lagged-plot}
 \end{figure}
@@
 
-\bibliography{../../common/covidcast.bib,pnas-materials/pnas-sample.bib}
+% \bibliography{../../common/covidcast.bib,pnas-materials/pnas-sample.bib}
 

### 2021-10-06 indicators/paper/paper.tex
Don't depend on prelim.tex
PNAS's system wants to rebuild the LaTeX from scratch, so make it easy.
Include survey consent and IRB information
Requested as part of the PNAS quality check.

@@
 
-\input{../../common/prelim.tex}
+\usepackage{microtype}
+\usepackage{booktabs}
+\def\cor{\mathrm{cor}}
 
@@
 \paragraph{Internet-Based Surveys} Conducted in partnership with Facebook,
-Delphi's COVID-19 Trends and Impact Survey receives an average of 50,000 responses
-daily, and has received over 20 million responses since April 2020
+Delphi's COVID-19 Trends and Impact Survey receives an average of 50,000
+responses daily, and has received over 25 million responses since April 2020
 \cite{Salomon:2021, Kreuter:2020}. From the surveys, we construct indicators on
 symptoms, social distancing, vaccination, and other attitudes and behaviors
-related to COVID.
+related to COVID. The surveys are voluntary and participants are redirected to a
+platform managed by Carnegie Mellon University to give consent and take the
+survey; individual response data is not provided to Facebook and the data
+collection protocol was approved by the Carnegie Mellon Institutional Review
+Board (STUDY2020\_00000162).
 
@@
 Altogether, we produce over 170 signals from 12 distinct sources, and provide
-them in a common format for public access. This unifies both unique (unavailable
+them in a common format foc access. This unifies both unique (unavailable
 anywhere else) and standard COVID data streams into a single common 
@@
 rates by county to decline, making it more difficult for any signal to achieve a
-high geo-wise correlation.
+high geo-wise correlation. Correlations between confirmed cases and cases 1--3
+weeks prior (shown in the Supplementary Information) show a similar correlation
+drop in February 2021, showing that the drop is due to a change in case data and
+not problems with the other signals.
 
@@
 \begin{figure}[t]
-  \includegraphics[width=\columnwidth]{fig/backfill-compare.pdf}
-  \caption{The 95th percentiles of relative changes between early reported
-    values of key signals and final values reported much later. For each date
+  \includegraphics[width=\columnwidth]{backfill-compare.pdf}
+  \caption{The 95th percentiles of relative error of early reported values of
+    key signals compared to final values reported much later. For each date
     between October 15, 2020 and April 15, 2021, the values for each state
@@
 change, showing that reported deaths can incur large relative error in initial
-values, comparable to that in claims-based signals.  However, for deaths, as
-well as cases, the revision pattern is not very systematic, with large
-corrections typically occuring at a sparse subset of locations and times (e.g.,
-due to audits or backlogs being cleared, which can result in thousands of cases
-or deaths can be added or removed all at once).  This backfill is much more
+values, comparable to that in claims-based signals. However, for deaths, as well
+as cases, these large revisions are not very systematic, with large corrections
+typically occurring at a sparse subset of locations and times (e.g., due to 
+audits or backlogs being cleared, which can result in thousands of cases or
+deaths can be added or removed all at once). This backfill is much more
 difficult to predict than that of claims data, therefore the latter (and other

### 2021-10-08 forecast/paper/supplement.Rmd
Revised figures post CHNG situation

--- 2021-10-04 forecast/paper/supplement.Rmd
@@
 actuals <- readRDS(here("data", "confirmed_7dav_incidence_prop.RDS"))
-fcasts_honest <- readRDS(here("data", "results_honest.RDS")) %>%
+fcasts_honest <- readRDS(here("data", "revised_chng", "results_honest.RDS")) %>%
   process_res_cases(actuals) %>%
@@
   filter(forecaster != "gs_inherit")
-hotspots_honest <- readRDS(here("data", "hotspots_honest.RDS")) %>%
+hotspots_honest <- readRDS(here("data", "revised_chng", "hotspots_honest.RDS")) %>%
   process_res_hotspots() %>%
@@
 # Figure 3 in the paper
-fig3 <- plotter(fcasts_honest %>% filter(period != "jm"), 
+fig3 <- plotter(fcasts_honest, 
         "wis", Mean, scaler = "strawman_wis", 
@@
 # Figure 4 in the manuscript
-fig4 <- plotter_hotspots(hotspots_honest %>% filter(period != "jm")) + 
+fig4 <- plotter_hotspots(hotspots_honest) + 
   geom_hline(yintercept = 0.5)
@@
 ```{r fcast-finalized, fig.cap="Forecasting performance using finalized data. Compare to Figure 3 in the manuscript."}
-plotter(fcasts_finalized %>% filter(period != "jm"), 
+plotter(fcasts_finalized, 
         "wis", Mean, scaler = "strawman_wis", 
@@
 ```{r hot-finalized, fig.cap="Hotspot prediction performance using finalized data. Compare to Figure 4 in the manuscript."}
-plotter_hotspots(hotspots_finalized %>% filter(period != "jm")) +
+plotter_hotspots(hotspots_finalized) +
   geom_hline(yintercept = 0.5)
@@
   left_join(
-    fcasts_honest %>% filter(period != "jm") %>% select(forecaster:wis),
-    fcasts_finalized %>% filter(period != "jm") %>% select(forecaster:wis) %>%
+    fcasts_honest %>% select(forecaster:wis),
+    fcasts_finalized %>% select(forecaster:wis) %>%
       rename(finalized_wis = wis)
@@
   hotspots_honest %>%
-    filter(period != "jm") %>%
     group_by(forecaster, ahead) %>%
@@
   hotspots_finalized %>%
-    filter(period != "jm") %>%
     group_by(forecaster, ahead) %>%
@@
 ```{r wis-densities, fig.cap="Weighted interval score appears to more closely resemble a log-Gaussian distribution."}
-ggplot(fcasts_honest %>% filter(period != "jm"), 
+ggplot(fcasts_honest, 
        aes(wis, fill = forecaster)) +
@@
 ```{r fcast-adjusted, fig.cap="Relative forecast performance using vintage data and summarizing with the more robust geometric mean."}
-plotter(fcasts_honest %>% filter(period != "jm"), 
+plotter(fcasts_honest, 
         "wis", GeoMean, scaler = "strawman_wis", 
@@
 st <- fcasts_honest %>%
-  filter(period != "jm") %>%
   mutate(wis = wis / strawman_wis) %>%
@@
 errs <- fcasts_honest %>% 
-  filter(period != "jm") %>%
   group_by(forecaster, forecast_date) %>%
@@
 ```{r bootstrap-loading}
-fcasts_booted <- readRDS(here("data","results_honest_bootstrapped.RDS")) %>%
+fcasts_booted <- readRDS(
+  here("data", "results_honest_bootstrapped.RDS")) %>%
   process_res_cases(actuals) %>%
@@
 }
-hotspots_booted <- readRDS(here("data", "hotspots_honest_bootstrapped.RDS")) %>%
+hotspots_booted <- readRDS(here("data", "revised_chng", "hotspots_honest_bootstrapped.RDS")) %>%
   process_res_hotspots() %>%
@@
 ```{r fcast-booted, fig.cap = "Forecast performance when indicators are replaced with samples from their empirical distribution. Performance is largely similar to the AR model."}
-plotter(fcasts_booted %>% filter(period != "jm"), 
+plotter(fcasts_booted, 
         "wis", Mean, scaler = "strawman_wis", 
@@
 Performance is largely similar to the AR model."}
-plotter(fcasts_booted %>% filter(period != "jm"), 
+plotter(fcasts_booted, 
         "wis", GeoMean, scaler = "strawman_wis", 
@@
 ```{r hot-booted, fig.cap = "Hotspot prediction performance when indicators are replaced with samples from their empirical distribution. Performance is largely similar to the AR model."}
-plotter_hotspots(hotspots_booted %>% filter(period != "jm")) + 
+plotter_hotspots(hotspots_booted) + 
   geom_hline(yintercept = 0.5)
@@
 ```{r up-down-processing}
-preds <- readRDS(here("data", "predictions_honest.RDS"))
+preds <- readRDS(here("data", "revised_chng", "predictions_honest.RDS"))
 preds <- preds %>% 
@@
 up_down_df_summary <- up_down_df %>%
-  filter(period != "jm") %>%
   group_by(forecaster, udf) %>%
@@
   up_down, by = c("geo_value","target_end_date")) %>%
-  filter(period != "jm") %>%
   group_by(forecaster, udf) %>%
@@
 up_down_df %>% 
-  filter(period != "jm") %>%
   group_by(forecaster, udf) %>%
@@
 hot_udf <- inner_join(
-  hotspots_honest %>% filter(period != "jm"),
+  hotspots_honest,
   up_down %>% select(geo_value, target_end_date, udf))
@@
 comparison_df <- left_join(
-  fcasts_honest %>% filter(forecaster != "AR", period != "jm"),
-  fcasts_honest %>% filter(forecaster == "AR", period != "jm") %>%
+  fcasts_honest %>% filter(forecaster != "AR"),
+  fcasts_honest %>% filter(forecaster == "AR") %>%
     select(-c(forecaster,target_end_date)) %>%
@@
 up_down_df %>% 
-  filter(period != "jm") %>% 
   group_by(forecaster, udf) %>%
@@
 df2 %>%
-  filter(period != "jm") %>%
   group_by(forecaster, udf) %>%
@@
 gg
-```
-
-<!-- Examining data in 2021 -->
-
-```{r fcast-alldates, fig.cap="Forecast performance over all periods. Performance largely improves for all forecasters with the inclusion of data in 2021."}
-plotter(fcasts_honest,
-        "wis", Mean, scaler = "strawman_wis", 
-        order_of_operations = c("aggr","scale")) +
-  ylab("Mean WIS (relative to baseline)")
-```
-
-\clearpage
-
-```{r fcast-alldates-adjusted, fig.cap="Forcast performance over all periods aggregaged with the geometric mean. Again, the inclusion of data in 2021 leads to improved performance."}
-plotter(fcasts_honest, 
-        "wis", GeoMean, scaler = "strawman_wis", 
-        order_of_operations = c("scale","aggr")) +
-  ylab("Geometric mean of WIS (relative to baseline)")
-```
-
-\clearpage
-
-```{r hot-alldates, fig.cap="Area under the curve for hotspot predictions including data in 2021.
-Performance degrades relative to the period in 2020.
-However, there are far fewer hotspots during this period as case rates declined in much of the country."}
-plotter_hotspots(hotspots_honest)
 ```
@@
 by_hrr <- fcasts_honest %>% 
-  filter(period != "jm") %>%
   group_by(forecaster, geo_value) %>% 
@@
 # Same as Fig 3 in the paper
-df <- summarizer(fcasts_honest %>% filter(period != "jm") %>%
+df <- summarizer(fcasts_honest %>%
                    group_by(ahead, forecaster), "wis", NULL, 
@@
 df2 <- fcasts_honest %>% 
-  filter(period != "jm") %>%
   group_by(ahead, forecaster) %>%

### 2021-10-08 forecast/code/figures/results_processing_and_plotting.R

@@ -22,7 +22,7 @@ plotter <- function(df, y, aggr,
   facet_layer <- NULL
   if (facet_by_period) {
     period_labels <- c(jj = "June - July", as = "Aug - Sep", o = "Oct",
-                       nd = "Nov - Dec", jm = "Jan - Mar")
+                       nd = "Nov - Dec") #, jm = "Jan - Mar")
     df <- df %>% group_by(period, .add = TRUE)
     facet_layer <- facet_wrap(~period, scales = "free_y",
                               labeller = labeller(period = period_labels))
@@ -45,7 +45,7 @@ plotter_hotspots <- function(df, facet_by_period = FALSE) {
   facet_layer <- NULL
   if (facet_by_period) {
     period_labels <- c(jj = "June - July", as = "Aug - Sep", o = "Oct",
-                       nd = "Nov - Dec", jm = "Jan - Mar")
+                       nd = "Nov - Dec")#, jm = "Jan - Mar")
     df <- df %>% group_by(period, .add = TRUE)
     facet_layer <- facet_wrap(~period, scales = "free_y",
                               labeller = labeller(period = period_labels))
@@ -69,11 +69,11 @@ period_processing <- function(res) {
     month(forecast_date) %in% 6:7 ~ "jj",
     month(forecast_date) %in% 8:9 ~ "as",
     month(forecast_date) %in% 10 ~ "o",
-    month(forecast_date) %in% 11:12 ~ "nd",
-    month(forecast_date) %in% 1:3 ~ "jm"))
+    month(forecast_date) %in% 11:12 ~ "nd"))
+    #month(forecast_date) %in% 1:3 ~ "jm"))
   res <- mutate(res,
                 period = fct_relevel(as.factor(period),
-                                     "jj","as","o","nd", "jm"))
+                                     "jj","as","o","nd"))#, "jm"))
   res
 }
 

### 2021-10-08 forecast/paper/response-to-reviewers.tex

@@
 \begin{resp}
-  During the period discussed in the main paper, vaccinations hadn’t really
-  taken off. On Dec 31 2000, when the main evaluation period ends,  according to
+  During the period discussed in the main paper, vaccinations had essentially no
+  effect. On Dec 31 2000, when the main evaluation period ends,  according to
   CDC data, about 1\% of the population 
@@
   to ignore such factors.
-
-  $\bullet$ We also added some text to the Prediction Period section stating
-  that our end 
-  date (31 December) coincides with the introduction of vaccinations.
 \end{resp}
@@
 \begin{resp}
-  This is an important point that we should have addressed more fully.
-  Conventional methods for assessing statistical significance that we explored
-  were overconfident, to the point of being unbelievable due to the
-  spatiotemporal correlation.
+  As noted in the introduction, there are no options available for fully 
+  rigorous and nonparametric (model-free) statistical significance testing,
+  giving the intricate spatiotemporal dependence in our forecasting problem.
 
-  We also have strong reasons for not using model-based approaches (one being the
-  nonstationarity the Reviewer mentions above, but also the belief that these
-  models are wrong). We felt that the related investigation in 2B was more
-  pertinent.
+  We also have strong reasons for not using model-based approaches (in part due
+  to nonstationarity, but also due to the belief that these models are not
+  applicable here). We felt that our extensive predictive comparisons in the
+  methods section and supplement were more appropriate.
 
-  To respond more directly we have made a number of additions.:
+  That said, to address this more directly we have made a number of additions:
 
@@
   to it around 
-  line 85 in the main text), that performs a Sign Test for differences in
+  line 458 in the main text), that performs a Sign Test for differences in
   accuracy. P-values there tend to cluster near zero and one suggesting some

### 2021-10-08 indicators/paper/paper.tex

@@
 \paragraph{Internet-Based Surveys} Conducted in partnership with Facebook,
-Delphi's COVID-19 Trends and Impact Survey receives an average of 50,000
-responses daily, and has received over 25 million responses since April 2020
+Delphi's COVID-19 Trends and Impact Survey receives an average of 50,000 responses
+daily, and has received over 25 million responses since April 2020
 \cite{Salomon:2021, Kreuter:2020}. From the surveys, we construct indicators on
 symptoms, social distancing, vaccination, and other attitudes and behaviors
-related to COVID. The surveys are voluntary and participants are redirected to a
-platform managed by Carnegie Mellon University to give consent and take the
-survey; individual response data is not provided to Facebook and the data
-collection protocol was approved by the Carnegie Mellon Institutional Review
-Board (STUDY2020\_00000162).
+related to COVID.
 
@@
 as cases, these large revisions are not very systematic, with large corrections
-typically occurring at a sparse subset of locations and times (e.g., due to 
+typically occuring at a sparse subset of locations and times (e.g., due to
 audits or backlogs being cleared, which can result in thousands of cases or

### 2021-10-08 indicators/paper/supplement.Rmd

@@
 \end{figure}
+
+\clearpage
+```{r case-correlation-lagged-data}
+start_day <- "2020-04-15"
+end_day <- "2021-04-15"
+as_of_day <- "2021-06-01"
+
+df_cases <- covidcast_signal("jhu-csse", "confirmed_7dav_incidence_prop",
+                             start_day, end_day, as_of = as_of_day)
+
+names <- c("7 days", "14 days", "21 days")
+lags <- c(7, 14, 21)
+df_cor <- vector("list", length(lags))
+
+for (ii in seq_along(lags)) {
+  df_cor[[ii]] <- covidcast_cor(df_cases, df_cases, dt_y = lags[ii],
+                                by = "time_value")
+  df_cor[[ii]]$signal <- names[ii]
+}
+
+df <- do.call(rbind, df_cor)
+df$signal <- factor(df$signal, levels = names, labels = names)
+```
+
+```{r case-correlation-lagged-plot, fig.cap="Correlation between confirmed cases and case reports 1, 2, or 3 weeks prior, for all counties in the United States with case reporting.
+Cases are correlated with future cases as one might expect, but note the precipitous drop in correlation in February 2021.
+This matches the correlation drop between other COVIDcast signals and cases during the same time period, supporting the hypothesis that the drop was due to decreased heterogeneity in case rates by county."}
+ggplot(df, aes(x = time_value, y = value)) +
+  geom_line(aes(color = signal)) +
+  ylim(0, NA) +
+  labs(x = "Date", y = "Correlation") +
+  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
+               date_minor_breaks = "1 month") +
+  theme_bw() +
+  theme(legend.position = "bottom", legend.title = element_blank())
+```
 

### 2021-10-08 forecast/paper/response-to-reviewers.tex
minor typo in response, recompile

@@
   During the period discussed in the main paper, vaccinations had essentially no
-  effect. On Dec 31 2000, when the main evaluation period ends,  according to
+  effect. On Dec 31 2020, when the main evaluation period ends,  according to
   CDC data, about 1\% of the population 
   had received 1 dose. That said, we should probably be more specific in the
-  paragraph on lines 31-40: our goal is to use very simple models that do not
+  paragraph on lines 31--40: our goal is to use very simple models that do not
   account for extra information (vaccinations but also strains of the virus,
@@
   There is little reason to believe that this data is stationary. The models
-  however are trained and evaluated over reasonably short periods (2-3 months),
-  over which time the behavior may well be stationary. We have added a section
+  however are trained and evaluated over reasonably short periods (2--3 months),
+  over which time the behavior may well be stationary.
+
+  $\bullet$ We have added a section
   to the appendix that uses the cumulative sum normalized by baseline as
@@
   (out-of-sample) comparison of forecasting performance for state-of-the-art
-  models can be found in reference 6 (Cramer et al.). That paper focuses on
+  models can be found in reference 7 (Cramer et al.). That paper focuses on
   forecasting deaths rather than cases and at the state-level rather than
@@
   the HRR level, but once scaled by the baseline, should be roughly comparable.
+  (Note also, these are for cases, not deaths as described in the previous
+  bullet.) 
   Only those forecasters that submitted at least 300 forecasts (roughly 50

### 2021-10-09 forecast/paper/supplement-text.tex
revised forecast paper

@@
 
+It is important to note, that even though some indicators, notably CHNG-COVID 
+and CHNG-CLI underperform relative to the AR model during upswings, all models 
+dramatically outperform the baseline in such periods. Figure 
+\ref{fig:upswing-summary-remake} shows the performance of all forecasters relative
+to the null model. All forecasters suffer relative to the baseline during down 
+periods, but the AR is the worst. In contrast, all models beat the baseline
+during up periods, even CHNG-COVID and CHNG-CLI, though not by quite as much as the
+AR does.
+
 \section{Leadingness and Laggingness}
@@
 Following the suggestion of an anonymous reviewer, we investigate two
-disaggregated versions of the main forecasting result shown in Figure 3 on the
+other disaggregated versions of the main forecasting result shown in Figure 3 on the
 manuscript. The first (Figure \ref{fig:cumulative-mean}) displays the cumulative
@@
 actually improves over the indicators in October 2020, before losing a bit in
-late November.
+late November. The second (Figure \ref{fig:cumulative-geo-mean}) repeats this 
+analysis but aggregating by Geometric mean as described above and displays a
+similar pattern.
 
@@
 appear to be any particular pattern to these locations.
-
-% \section{Comparisons to state-of-the-art forecasters}
-
-% Since mid-July 2020, a number of forecasting teams (48 in total) have submitted
-% 1--4 week ahead state-level case forecasts to the COVID19 Forecast Hub
-% \cite{ForecastHub}. To compare those forecasts to those created in this paper,
-% we selected only those forecasters who submitted at least 300 forecasts for the
-% 2-week-ahead target before
-% January 1, 2021 (approximately 1 forecast for each state for 6 weeks). This left
-% 32 forecasters, one of which, the COVIDhub-4-week-ensemble, was an exact
-% duplicate of another. To make the comparison fair, we scaled by the
-% COVIDhub-baseline as in Figure 3 of the manuscript. Figure
-% \ref{fig:compare-to-hub-mean} shows all submitted forecasts (available 7, 14,
-% and 21 days ahead) overlaid on Figure 3. We have highlighted the
-% COVIDhub-ensemble in light blue for comparison. All the models developed in this
-% paper beat the baseline, and would rank in the top 8 of 31 submitted forecasts
-% over this period.
-
-% A more fair comparison that directly adjusts for spatiotemporal difficulty is to
-% use the geometric mean scaled by the baseline. This is the same idea shown in
-% \ref{fig:fcast-adjusted}. We have also remade this figure overlaying submitted
-% forecasts available on the Forecast Hub. It occasionally happens that a
-% submitted forecast achieved a WIS of 0, so in this case we use the WIS$+ 1$
-% rather than WIS for all forecasters. The result is shown in
-% \ref{fig:compare-to-hub-geomean}. By this metric, the methods employed in this
-% paper would rank in the top 2, beating the COVIDhub-ensemble at all forecast
-% horizons. The improvement of the best indicator assisted model over the AR model
-% is roughly the same as the improvement of the AR model over the COVIDhub-ensemble.

### 2021-10-09 forecast/paper/supplement.Rmd
revised forecast paper

@@
 actuals <- readRDS(here("data", "confirmed_7dav_incidence_prop.RDS"))
-fcasts_honest <- readRDS(here("data", "revised_chng", "results_honest.RDS")) %>%
+fcasts_honest <- readRDS(here("data", "results_honest.RDS")) %>%
   process_res_cases(actuals) %>%
@@
   filter(forecaster != "gs_inherit")
-hotspots_honest <- readRDS(here("data", "revised_chng", "hotspots_honest.RDS")) %>%
+hotspots_honest <- readRDS(here("data", "hotspots_honest.RDS")) %>%
   process_res_hotspots() %>%
@@
 }
-hotspots_booted <- readRDS(here("data", "revised_chng", "hotspots_honest_bootstrapped.RDS")) %>%
+hotspots_booted <- readRDS(here("data", "hotspots_honest_bootstrapped.RDS")) %>%
   process_res_hotspots() %>%
@@
 ```{r up-down-processing}
-preds <- readRDS(here("data", "revised_chng", "predictions_honest.RDS"))
+preds <- readRDS(here("data", "predictions_honest.RDS"))
 preds <- preds %>% 
@@
 
-```{r upswing-summary-remake, fig.cap="REMAKE: Percent change in average WIS of the forecaster (AR or indicator assisted) relative to the baseline.
-All models perform poorly durning down periods, but the indicators help.
-During flat periods, the indicators improve slightly over the AR.
-During up periods, all forecasters do much better than the baseline, but only some do as well as AR."}
-up_down_df_summary <- inner_join(
-  left_join(
-    corr_df, 
-    fcasts_honest %>% select(ahead, geo_value, forecast_date, strawman_wis)),
-  up_down, by = c("geo_value","target_end_date")) %>%
-  group_by(forecaster, udf) %>%
-  summarise(mean = Mean(wis) / Mean(strawman_wis) - 1)
-up_down_df_summary %>%
-  ggplot(aes(udf, mean, fill = forecaster)) +
-  geom_bar(width = 0.6, position = position_dodge(width=0.6),
-           stat = "identity") +
-  scale_fill_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
-  theme_bw() +
-  scale_y_continuous(labels = scales::percent_format()) +
-  geom_hline(yintercept = 0) +
-  ylab("% change in Forecaster WIS\n relative to Baseline") +
-  xlab("") +
-  theme(legend.position = "bottom", legend.title = element_blank())
-```
-
 ```{r upswing-histogram, include=FALSE, fig.width=9, fig.height=4.5}
@@
 ```{r hotspots-upswing-downswing, fig.cap="Classification and loglikelihood separated into periods of upswing, downswing, and flat cases.
-Like the analysis of the forecasting task in the main paper (see Figure 7), performance is better during down and flat periods."}
+Like the analysis of the forecasting task in the main paper (see Figure 7), relative performance is better during down and flat periods."}
 hot_udf <- inner_join(
@@
 In terms of loglikelihood, all forecasters have lower likelihood than the null classifier during up periods.
-In down and flat periods, the indicators generally improve over the AR, while the are worse during up periods."}
+In down and flat periods, the indicators generally improve over the AR, while the are worse during up periods.", eval=FALSE}
 # mycummean <- function(x) {
@@
   filter(!is.na(actual)) %>%
-  mutate(pred = value > cutoff) %>%
+  mutate(pred = value > .5) %>%
   group_by(forecaster, udf) %>%
@@
 #                    "periods then in up periods."))
+```
+
+\clearpage
+
+```{r upswing-summary-remake, fig.cap="Percent change in average WIS of the forecaster (AR or indicator assisted) relative to the baseline.
+All models perform poorly durning down periods, but the indicators help.
+During flat periods, the indicators improve slightly over the AR.
+During up periods, all forecasters do much better than the baseline, but only some do as well as AR."}
+up_down_df_summary <- inner_join(
+  left_join(
+    corr_df, 
+    fcasts_honest %>% select(ahead, geo_value, forecast_date, strawman_wis)),
+  up_down, by = c("geo_value","target_end_date")) %>%
+  group_by(forecaster, udf) %>%
+  summarise(mean = Mean(wis) / Mean(strawman_wis) - 1)
+up_down_df_summary %>%
+  ggplot(aes(udf, mean, fill = forecaster)) +
+  geom_bar(width = 0.6, position = position_dodge(width=0.6),
+           stat = "identity") +
+  scale_fill_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
+  theme_bw() +
+  scale_y_continuous(labels = scales::percent_format()) +
+  geom_hline(yintercept = 0) +
+  ylab("% change in Forecaster WIS\n relative to Baseline") +
+  xlab("") +
+  theme(legend.position = "bottom", legend.title = element_blank())
 ```

### 2021-10-09 forecast/paper/supplement.tex
Revised figures post CHNG situation
revised forecast paper

@@
 \midrule
-Geometric Mean Relative WIS & 0.141 & 0.076 & 0.000 & 0.059 & 0.008\\
-Mean Relative WIS & 0.186 & 0.137 & 0.002 & 0.074 & 0.064\\
+Geometric Mean Relative WIS & 0.116 & 0.067 & 0.000 & 0.059 & 0.008\\
+Mean Relative WIS & 0.159 & 0.111 & 0.002 & 0.074 & 0.064\\
 \bottomrule
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{fig/upswing-summary-remake-1}
-
-}
-
-\caption{REMAKE: Percent change in average WIS of the forecaster (AR or indicator assisted) relative to the baseline.
-All models perform poorly durning down periods, but the indicators help.
-During flat periods, the indicators improve slightly over the AR.
-During up periods, all forecasters do much better than the baseline, but only some do as well as AR.}\label{fig:upswing-summary-remake}
-\end{figure}
-
 \clearpage
@@
 \caption{Classification and loglikelihood separated into periods of upswing, downswing, and flat cases.
-Like the analysis of the forecasting task in the main paper (see Figure 7), performance is better during down and flat periods.}\label{fig:hotspots-upswing-downswing}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{fig/hotspots-upswing-downswing-remake-1}
-
-}
-
-\caption{REMAKE: Classification and loglikelihood separated into periods of upswing, downswing, and flat cases.
-The indicator assisted models always have lower classification error relative to the null classifier, while the AR actually does worse in down and flat periods.
-In terms of loglikelihood, all forecasters have lower likelihood than the null classifier during up periods.
-In down and flat periods, the indicators generally improve over the AR, while the are worse during up periods.}\label{fig:hotspots-upswing-downswing-remake}
+Like the analysis of the forecasting task in the main paper (see Figure 7), relative performance is better during down and flat periods.}\label{fig:hotspots-upswing-downswing}
 \end{figure}
@@
 This conclusion is stronger in down periods then in up periods.}\label{fig:upswing-corr-table}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/upswing-summary-remake-1}
+
+}
+
+\caption{Percent change in average WIS of the forecaster (AR or indicator assisted) relative to the baseline.
+All models perform poorly durning down periods, but the indicators help.
+During flat periods, the indicators improve slightly over the AR.
+During up periods, all forecasters do much better than the baseline, but only some do as well as AR.}\label{fig:upswing-summary-remake}
 \end{figure}

### 2021-10-16 indicators/code/count_signals.R
Better document indicator code

@@ -1,3 +1,6 @@
+# Count the unique signals in the COVIDcast API so the paper can quote an
+# approximate number of signals.
+
 library(covidcast)
 library(dplyr)
 

### 2021-10-16 survey/code/paper-tables.Rmd
Add READMEs, license, and missing data file
The file has to be compressed because GitHub has a 100mb file size
limit. Fortunately R can decompress the file automatically.

@@
 
-# SVI
+# Centers for Disease Control and Prevention/Agency for Toxic Substances and
+# Disease Registry/Geospatial Research, Analysis, and Services Program.
+# CDC/ATSDR Social Vulnerability Index 2018 Database US.
+# https://www.atsdr.cdc.gov/placeandhealth/svi/data_documentation_download.html
 svi <- read.csv("data/SVI2018_US_COUNTY.csv") %>%

### 2021-10-16 indicators/paper/paper.tex
BrowNstein

@@
   Brooks:2020}. Several other examples exist that span a wide variety of data
-platforms and diseases \cite{Browstein:2009, Kass-Hout:2011, Salathe:2012,
+platforms and diseases \cite{Brownstein:2009, Kass-Hout:2011, Salathe:2012, 
   Kass-Hout:2013}. During the COVID-19 pandemic, digital data streams have

### 2021-10-16 indicators/code/indicator-paper-figures.Rmd
Better document indicator code

@@
 author: Alex Reinhart, et al.
-date: June 23, 2021
+date: October 16, 2021
 ---
@@
 [rmarkdown package](https://cran.r-project.org/package=rmarkdown). By default it
-also saves publication-ready PDF versions of figures to the `../paper/fig/`
+saves publication-ready PDF versions of figures to the `../paper/fig/`
 directory.
+
+This file was run with R version 4.0.4 and the versions of all packages
+available in October 2021.
 
@@
 ```{r figure-7, message=FALSE, cache=TRUE}
-cdc <- read.csv("COVID-19_Vaccinations_in_the_United_States_County.csv") %>%
+cdc <- read.csv("COVID-19_Vaccinations_in_the_United_States_County.csv.gz") %>%
   filter(Date == "07/20/2021")

### 2021-10-16 indicators/paper/paper.tex

@@
 \paragraph{Internet-Based Surveys} Conducted in partnership with Facebook,
-Delphi's COVID-19 Trends and Impact Survey receives an average of 50,000 responses
-daily, and has received over 25 million responses since April 2020
+Delphi's COVID-19 Trends and Impact Survey receives an average of 50,000
+responses daily, and has received over 25 million responses since April 2020
 \cite{Salomon:2021, Kreuter:2020}. From the surveys, we construct indicators on
 symptoms, social distancing, vaccination, and other attitudes and behaviors
-related to COVID.
+related to COVID. The surveys are voluntary and participants are redirected to a
+platform managed by Carnegie Mellon University to give consent and take the
+survey; individual response data is not provided to Facebook and the data
+collection protocol was approved by the Carnegie Mellon Institutional Review
+Board (STUDY2020\_00000162).
 
@@
 \begin{figure}[t]
-  \includegraphics[width=\columnwidth]{backfill-compare.pdf}
+  \includegraphics[width=\columnwidth]{fig/backfill-compare.pdf}
   \caption{The 95th percentiles of relative error of early reported values of

### 2021-10-18 perspective/paper.tex
Perspective piece

\documentclass{article}
\input{ryantibs}

\title{Epidemic Tracking and Forecasting: Lessons Learned from a Tumultuous Year}
\author{Roni Rosenfeld \and Ryan J. Tibshirani}

\begin{document}
\maketitle

\section{Introduction}

Epidemic forecasting has garnered increasing interest in the last decade,
nurtured and scaffolded by various forecasting challenges organized by groups
within the U.S. federal government, including the CDC \citep{Biggerstaff:2016,
  Biggerstaff:2018, Reich:2019}, OSTP \citep{Johansson:2019}, DARPA
\citep{DelValle:2018}, and elsewhere \citep{Ajelli:2018, Viboud:2018}.  In 2017,
after several years of experimentation with flu forecasting in academic groups,
the CDC decided to incorporate influenza forecasting into its normal operations,
including weekly public communications citep{FluSight} and briefing to
higher-ups.  To provide more reliable infrastructure and support for its
forecasting needs, the CDC in 2019 designated two national Centers of Excellence
for Influenza Forecasting, one at the University of Massachusetts at
Amherst\footnote{\url{https://reichlab.io/people}} and one at Carnegie Mellon
University\footnote{\url{https://delphi.cmu.edu/about/center-of-excellence/}}.

Not unrelatedly, the last decade has also seen a rise in the importance of
\emph{digital surveillance} streams in public health, with improving epidemic
tracking and forecasting models being a key application of these data.  Digital
streams, such as search and social media trends, have constituted a large part
of the focus \citep{Ginsberg:2009, Brownstein:2009, Salathe:2012,
Kass-Hout:2013, Santillana:2015, Paul:2017}; however, even more broadly, data
from \emph{auxiliary} streams that operate outside of traditional public health
reporting, such as online surveys, medical devices, or electronic medical
records, have received considerable attention as well \citep{Kass-Hout:2011,
Carlson:2013, Viboud:2014, Smolinski:2015, Santillana:2016, Charu:2017,
Koppeschaar:2017, Yang:2019, Leuba:2020, Radin:2020, Ackley:2020}.

The Carnegie Mellon Delphi group, which the two of us co-lead, has worked in
both of these emerging disciplines---epidemic forecasting, and building relevant
auxiliary signals to aid such forecasting models---since 2012.  In 2020, as the
pandemic broke out, we struggled like many other groups to find ways to
contribute to the national efforts to respond to the pandemic.  We ended up
shifting our focus to lie nearly entirely on the data end of the spectrum,
pursuing several different directions in order to build and make available to
the public a variety of new indicators that reflect real-time COVID-19 activity
in the U.S.  Three papers in this collection describe this work from three
different perspectives.  A fourth describes international work by some of our
collaborators that parallels our group's work on online surveys in the U.S.

\section{Papers in This Collection}

Here is a very brief summary of the papers in this collection.

\begin{enumerate}
\item \citet{Reinhart:2021} describes our group's (ongoing) effort in building
  and maintaining COVIDcast: an open repository of real-time, geographically
  detailed COVID-19 indicators in the U.S.  These indicators (a term we use
  interchangeably with signals) are derived from a diverse set of data sources:
  medical testing devices, medical insurance claims, internet search trends,
  app-based mobility data, and online surveys, among others.  Many indicators
  are demonstrated to have meaningful statistical relationships with what have
  become the pandemic's ``topline'' numbers (reported cases, hospitalizations,
  deaths), whereas others uniquely reflect certain activity (not available in
  other publicly-available data sources) that may drive or affect the spread of
  COVID-19.  The paper demonstrates through a number of examples that the
  indicators in the COVIDcast repository can improve on the timeliness,
  robustness, and scope of traditional public health reporting streams.

\item \citet{McDonald:2021} provides a detailed analysis of whether a core set
  of the indicators in the COVIDcast repository can be used to improve the
  accuracy of COVID-19 short-term forecasting and hotspot detection models.
  This speaks to the quantitative utility of the indicators in a way that is
  directly tied to the benefits observed in relevant downstream modeling tasks.
  The paper finds that time series models (that are already competitive with top
  forecasters from the COVID-19 Forecast Hub \citep{ForecastHub}) improve in
  predictive accuracy when they are supplemented with any of the five indicators
  under consideration.  However, the magnitude of these improvements varies with
  the pandemic's local dynamics at prediction time, and only one indicator
  appears to offer a nontrivial improvement during upswings.

\item \citet{Salomon:2021} focuses on the U.S. COVID-19 Trends and Impact Survey
  (CTIS), an (ongoing) online survey operated by our group, in partnership with
  Facebook.  This is a very rich source of data about the pandemic and its
  effect on people, only partially reflected by the indicators (derived from the
  survey) in the COVIDcast repository; the full data set of individual,
  anonymized survey responses is available to researchers under a data use
  agreement.  The paper presents descriptive analyses that reflect the unique
  value of CTIS as an important supplement to public health reporting, in
  particular, as an instrument to measure key information about behaviors,
  attitudes, economic impacts, and other topics not covered in traditional
  public health streams.

\item \emph{Authors' note to the Editor: this is to be filled out later.  The
    Astley et al. paper is still undergoing a significant revision and its main
    messages may change as a result.}
\end{enumerate}

\section{Lessons Learned}

We now take the opportunity to reflect on some ``lessons learned'' from our work
over the past year and a half.  Some of the observations below are described in
more depth in the papers in the collection, and others extend beyond the papers
in the collection (but we give references to relevant articles with more details
in the discussion below).

\paragraph{Deceptively simple data labels often belie the data's true meaning
  and complexity.}

Labels such as ``COVID-19 cases'' or ``COVID-19 hospitalizations'' hide an
enormous amount of complexity and potential ambiguity, especially when applied
to data at fine geographic and temporal resolutions.  We elaborate on this and
several other examples in what follows.

\begin{itemize}
\item Cases may be lab-confirmed only or also suspected (with the definition of
  ``suspected'' varying across jurisdictions and time); they may be listed by
  date reported on the jurisdiction's website, by date reported to the public
  health authority, by date tested, by specimen collection date, or occasionally
  by symptom onset date (most informative but often unavailable or
  inapplicable). A casual review of many websites of local and state health
  departments suggests there is great heterogeneity in what is being reported
  \citep{Simon:2021}.

\item The term ``hospitalizations'' is used ambiguously, sometimes referring to
  incidence (hospital admissions) and sometimes to prevalence (hospital bed
  occupancy).  These two quantities cannot easily be mapped to one another, 
  because COVID-19 hospital discharges are rarely if ever reported.
  Furthermore, people admitted without a COVID-19 diagnosis may acquire the
  infection and/or the diagnosis at any time during their hospital stay.

\item Hospitalizations may be reported by the location (typically county) of the
  patient's residence, but are more often reported by the county of the
  reporting hospital.  This is an important distinction, as many rural COVID-19
  patients in need of advanced care travel to the nearest secondary or tertiary
  hospital, often at a nearby urban county.  For example, the hospitals in
  Pittsburgh, located in Allegheny County, Pennsylvania (population 1.2
  million), treat patients from throughout a 13 county region in Southwestern
  Pennsylvania (population over 4 million).  For tracking and forecasting
  hospitalization burden then, the geographic units of hospital referral regions
  (HRRs, \citep{DartmouthHRR}) may be most appropriate.  Alas, these units do
  not conform to county boundaries, which complicates the projection of cases to
  hospitalizations.

\item Deaths are usually reported by county where they occurred, which for
  hospitalized patients may differ from their county of residence \citep{NCHS}.

\item Hospitalization or deaths with COVID-19 are significantly different from
  hospitalization or deaths due to COVID-19 (as captured by e.g., a
  COVID-related chief complaint or primary ICD-10 code).  The proportion of the
  two varies significantly by age groups and across time \citep{Fillmore:2021}.

\item Test positivity rates are most often reported by lumping together all
  tests performed regardless of the reason for performing them.  Tests taken
  following positive diagnosis, due to symptoms, or due to being a contact of a
  confirmed case, are all likely to have a much higher positivity rate than that
  of the general population.  Screening tests are most likely to reflect the
  true prevalence in the screened population.  Sadly, very few jurisdictions
  report or even retain the breakdown of the test results by reason for testing,
  thereby losing forever valuable information.

\item Medical insurance claims offer rich, detailed information about COVID-19
  and other health conditions, but are not without their weaknesses.  Claims are
  often not filed until weeks and months after the medical encounter.  As such,
  signals derived from claims are usually subject to regular and considerable
  revisions up to 60 days after a given date of service, because signals must be
  updated each time new claims for that day of service are received (specific
  statistics on how this affects signal values are given below).  This tends to
  make projections challenging, especially at finer geographic units such as
  counties, since there tends to be a high degree of heterogeneity across
  locations.

\item Medical claims contain information about lab tests taken, but not their
  results.  More generally, for understandable HIPAA reasons, medical claims
  contain only information necessary for adjudicating and auditing claims.   
\end{itemize}

Data definitions must be disambiguated, clarified, and made consistent to the
greatest extent possible, and remaining inconsistencies must be documented and
saliently communicated.

\paragraph{Understanding the data generation process is critical for downstream
  applications.}

Both traditional public health surveillance data streams and newer digital
surveillance streams are the result of often complex processes, some having to
do with the underlying health status or activities being monitored, others with
the reporting process itself.  Understanding the entire ``data generation
process'' for each data source can be challenging, but is absolutely essential
for proper modeling and effective use of the data.  Some examples are as
follows.

\begin{itemize}
\item In medical claims, relevant diagnoses and comorbidities may not be
  reflected if they are not directly relevant to the charges incurred.  On the
  other hand, because medical claim coding determines reimbursement levels, some
  codes may be over-represented relative to their medical significance.

\item Some populations and some healthcare settings are not reflected in the
  commercial claims stream.  These include the healthcare systems of the
  Department of Defense, Indian Health Services, Veterans Affairs, prison
  systems, and other systems that do not reimburse by procedure or service, as
  well as Medicare fee-for-service and Medicaid.  This can cause significant
  bias in the signals relative to the prevalence in the general population.

\item Public health reporting data are often subject to backlogs and reporting
  delays, and estimates for any particular date can be revised over time as
  errors are found or additional data becomes available.  During the pandemic,
  audits, corrections, and the clearing of backlogs has frequently resulted in
  huge artificial spikes and drops \citep{ArvisaisAnhalt:2021}.  Data
  aggregators like Johns Hopkins CSSE \citep{Dong:2020} have worked tirelessly
  to correct such anomalies after first publication (they attempt to
  back-distribute a spike or dip, by working with a local authority to figure
  out how this should best be done).

\item Data revisioning (also known as ``backfill'') is pervasive not only in
  traditional public health reporting, but also in many (though not all) digital
  surveillance sources.  As already alluded to above, signals based on medical
  claims typically undergo regular revisions because many claims (on which these 
  signals are based) get submitted and processed late; for many COVID-related
  claims-based signals, the median relative error between initial reports and
  final values is over 10\%, and only after 30 days or so do estimates typically
  match finalized values within 5\% \citep{Reinhart:2021}.  However, the
  systematic nature of these revisions suggests that, with suitable historical
  data, statistical models could be fit to estimate the final values from
  preliminary reports.  By comparison, revisions to public health reports during
  the pandemic (the spikes and dips just described) have been much less
  systematic and much less predictable.

\item Traditionally, public health agencies do not publish provisional data
  until it meets a level of stability.  For example, NCHS data on the percentage
  of deaths due to pneumonia, influenza and COVID-19 is not released until at
  least 20\% of the expected deaths in a jurisdiction have been reported
  \citep{PIMortality}, a process that may take several weeks.  However, for
  modeling and forecasting purposes, even highly provisional data can be very 
  informative, as long as sufficient historical provisional data is collected so
  that the statistical relationships between provisional values and finalized
  values can be modeled.

\item Calendar effects permeate not only the reporting process but also
  health-seeking behavior and the epidemic process itself, with the effects on
  these three processes not easy to disentangle.  Major holidays and other
  national or regional events are associated with significant travel, social
  mixing, and other distinct behaviors affecting disease transmission.  However,
  holidays and weekends also affect health-seeking behavior via reduced
  non-emergency healthcare capacity (doctors' offices and labs being partially
  or fully closed).  Perhaps the strongest calendar effects are on reporting,
  including claim filing and hospital reporting.  Using 1- or 2-week trailing
  averages eliminates weekend effects, but at a cost of reduced temporal
  resolution, and it leaves unsolved the effects of holidays.  A better approach
  might be to explicitly model and correct for calendar effects. 
\end{itemize}

\paragraph{Mandated reporting in a time of emergency can be burdensome and
  inflexible.}

COVID-19 reporting by hospitals, as mandated by HHS during the pandemic,
consisted of many dozens of data elements and imposed a significant burden on
the nation's 6,000 or so hospitals at a time that they were already stretched to
their limits.  It also took a huge effort to formulate, communicate,
disambiguate, and monitor for quality assurance and uniformity of
interpretation.  In light of this, it is not surprising that it took a long
time, and pressure, for most hospitals to comply (near universal compliance was
not achieved until December 2020).  When changes needed to be made to the
collected statistics, an arduous and time consuming process of approvals,
re-formulation, re-communication, re-implementation, and re-assessment had to be
followed.

While some aspects of mandated reporting are likely to remain irreplaceable,
effective alternative surveillance sources can be of great use: they can improve
on the timeliness, scope, robustness, and utility of mandated reporting data,
while being less burdensome to collect.  This is a common theme that runs
through all four papers in the collection, but is perhaps most directly
addressed in \citet{Reinhart:2021} (which focuses on the ecosystem of signals
broadly).  That said, we have far from saturated the utility of auxiliary
surveillance.  Much more needs to be developed in this area in order to usher
epidemic tracking into its next phase of reliability, accuracy, and
transparency.  To us, electronic medical records hold a great promise for
surveillance streams, and we elaborate on this in the next section.

\paragraph{Human behavior and its impact on the progression of epidemics is hard
  to measure and hard to model.}

In the nearly 10 years of government-organized epidemic forecasting exercises in
the U.S., efforts were focused on modeling the natural history and likely
evolution of the pathogen, with adaptation of human behavior playing a secondary
role (if any role at all).  The pandemic demonstrated that our forecasting
models must pay closer attention to reactive human behavior, even more so if we
are to consider interventions.  Unfortunately, many highly relevant aspects of
human behavior are not measured by publicly available data streams, like
compliance with policies and recommendations (with perhaps surveys and mobility
reports providing our best glimpse into these hard-to-observe aspects of
behavior \citep{Bilinski:2021}), and even the very process of promulgating these 
policies and recommendations.  Furthermore, even if we had these data in hand,
incorporating their effects will require significant and new cognitive and
behavioral modeling, with uncertain success.  The tragic breakdown and
fragmentation of trust in governments, public health officials, and healthcare
professionals are perhaps the hardest factors to measure and model, yet they
played an undeniable role in the progression of the current pandemic in the
U.S. and other countries.

\section{The Road Ahead}

The focus of the Delphi group during the initial, critical period of the
pandemic (February 2020 -- March 2021) was on short-term goals: trying to
provide signals, analysis, and decision support to federal, state and local
public health officials, as well as to fellow researchers, data journalists, and
the general public.  In spring of 2021, equipped with the hard lessons learned
during this tumultuous year, we turned our attention back to the original vision
of the Delphi group, and asked ourselves: given where we are and what we know
now, what is needed to be able to take a major step forward in epidemic tracking
and forecasting?  In this section, we list some ideas which we hope will elicit
further public discussion and, most importantly, experimentation. Because our
expertise lies in modeling and forecasting, not in public health surveillance,
our perspective and recommendations are necessarily limited to those aspects of
surveillance that are needed to realize our vision.

\paragraph{EMR as a key missing component for epidemic tracking and
  forecasting.}

The success of nowcasting, analytics and forecasting depends crucially on the
availability of rich, real-time data sources.  In light of the limitations of
mandated reporting discussed above, we must consider the complementary value of
other data sources.  Chief among these are electronic medical records (EMR), as
are being created and used daily by inpatient and outpatient healthcare
providers, medical laboratories, and pharmacies.  The advantage of these data
resources: they are rich, real-time, and already being generated (found ``in the
wild'').  The challenges: they are highly fragmented in the U.S., with its
approximately 6,000 hospitals and 100,000 outpatient care facilities.  One
promising avenue for countering this fragmentation are Health Information
Exchanges (HIEs), which were set up in the early 2000s with the support of the
federal government to coordinate the sharing of healthcare information among
healthcare providers in a given region, and eventually nationally.  The primary
goal of the HIEs has been patient continuity of care, but public health
surveillance is recognized as an additional worthy goal.  In the context of
health surveillance, HIEs hold the promise of reducing fragmentation from a
hundred thousand partners to only a few hundred.

Other formidable challenges to using EMR are legal, ethical, commercial, and
operational.  Who owns the data, who has access to it, and who has use rights
are all complex and often open questions.  An overriding concern is of course
patient privacy.  We must find a way to use this highly promising data for the
common good without compromising the privacy of individuals.  Fortunately, a
technological solution appears feasible, in the form of \emph{federated
  surveillance}. An outline is as follows.

\begin{itemize}
\item A common API is developed for querying all participating EMR custodians.

\item EMR databases are queried daily with an agreed upon set of queries, and
  return aggregated counts.

\item These counts are further aggregated across multiple providers and
  localities, and then fused with all other available data sources to provide
  alerts, nowcasts, and forecasts.

\item These model outputs are then shared back to the contributing EMR
  custodians, as well as to CDC and other federal and state agencies.

\item Done in this way, no personal health information (PHI) ever leaves the
  custodians' premises, while aggregated statistics can be combined to increase
  statistical power, thereby shortening alert latency and improving detection
  and prediction capabilities.
\end{itemize}

A successful example of federated querying (albeit designed for research rather
than real-time tracking and forecasting) has been recently demonstrated in the
UK \citep{Williamson:2020}.

One advantage of this approach to health surveillance is that when a new
emerging health crisis is identified, or when specific syndromic signatures are
discovered (e.g., ageusia and anosmia for COVID-19), a new query can be
developed, approved, and deployed literally overnight, allowing us to ``shine a
light'' on it on very short notice.  This can be contrasted with traditional,
legally-mandated public health reporting, which could take weeks and months to
develop, approve, negotiate, disseminate, implement, monitor, and assure quality
of, as has happened during the current pandemic.  In the slightly longer term,
demonstrating the effectiveness and superiority of federated surveillance could
obviate the need for crisis-time mandated reporting, alleviating the reporting
burden on hospitals and other healthcare providers during these difficult
times.

In an interpandemic period, an important use case for federated surveillance is
detection of trends and anomalies.  A set of queries can be designed to
continuously test for unusual recent spikes or trends in any number of
diagnoses, symptoms, lab tests, or prescriptions.  The aggregation of evidence
across many systems, localities, and data streams will make detection both more
sensitive and more robust.  Such a system would likely have detected the opioid
epidemic years before it was actually noticed.

One open technical challenge with the federated surveillance approach is
semantic heterogeneity: the use of emerging Health IT standards like HL7's FHIR
\citep{HL7FHIR} can enable a unified view of EMR data elements, but different 
healthcare systems often have different operational definitions for supposedly
universal concepts like ``high blood pressure'' or ``low blood oxygenation''.
Combining counts of such events across different healthcare systems may be a bit
like adding apples to oranges.  It will take some work to harmonize semantics
across so many diverse data custodians, but this is both doable and well worth
doing.  Note that this problem is less severe for the many surveillance and
anomaly detection tasks where the focus is on changes in a signal (in a given
location) over time, rather than on its absolute meaning.

\paragraph{Different phases of epidemic surveillance call for different analytic
  tools.}

It is important to discuss analytic needs separately for each of three different 
phases of epidemic surveillance and tracking, since each poses different
technical challenges and requires different analytical tools.

In the \emph{interpandemic phase}, the main activity is threat scanning for
threats, namely monitoring data streams and events throughout the world for
disconcerting developments.  Relevant statistical tools include anomaly
detection and scan statistics to help decide when an epidemiological
investigation is warranted.  While it may be possible to rank the risks of
different outbreak triggering events (like species jumping or point mutations)
in different locations, which could in turn be used to inform surveillance
resource allocation, conventional forecasting has a limited role to play in this
phase, as such events have large inherent uncertainty.

In the \emph{containment phase}, a discovered threat must be intensely
monitored, continuously assessed, and ultimately contained.  The analytical,
data-driven tools required in this phase include real-time estimation of
critical epidemiological parameters such as R0, infection fatality rates, the
incubation period, the serial interval, and so on.  These real-time estimates
are necessarily based on provisional data, highlighting the value of modeling
the data generation process discussed above.  In this phase, forecasting still
has a limited role, since the outbreak is still local, its fundamental dynamics
unclear, and point events can have large consequences down the road.

If containment fails, during the \emph{mitigation phase} the goals of analytics
expand significantly to include informing mitigation policies and planning.
Real-time tracking (nowcasting) and short term-forecasting (a few weeks
ahead) can play critical roles in these activities, and indeed have been the
focus of our group's work since its inception.  While there is still important
work to be done and advances to be made in this area, we believe that these
advances are likely to be incremental until we see major progress in either (1)
supporting data streams (e.g., better standardization and cleaning of public
health reporting data, identification of leading indicators from, say, EMR); or
(2) our collective scientific understanding of the real-world geotemporal
dynamics of epidemics (discussed next).

\paragraph{Useful, reliable longer-term forecasting remains an aspiration.}

Influenza forecasting exercises in the last several years demonstrated that it
is often possible to usefully quantify uncertainty over the remainder of an
ongoing flu season \citep{Reich:2019}.  But this success was based mostly on
observing the behavior of seasonable epidemics over several decades.  To
reliably forecast the progression of pandemics, where relevant historical data
is almost nonexistent, we must have a detailed quantitative understanding of how
different, diverse factors affect disease transmissibility.  Such an 
understanding is currently grossly lacking, as evidenced by our collective
failure to predict \citep{Reich:2021} (or even understand, post-hoc
\citep{Hawre:2021}) the high-level temporal and geographic contours of the main
pandemic waves in the U.S.  Yet this very pandemic, the most instrumented in
human history, is also a rare opportunity to attempt this vital scientific and
technological goal.

\subsection*{Acknowledgements}

\emph{Authors' note: to be filled out later.}

\newpage
\bibliographystyle{plainnat}
\bibliography{../common/covidcast}

\end{document}

### 2021-10-22 forecast/paper/paper.tex
Sync up before branch
Make suggested minor changes to fcast

@@
   Reliable, short-term forecasts of traditional public health reporting streams
-  (such as cases, hospitalizations, and deaths) are a key ingredient in
+  (such as cases, hospitalizations, and deaths)
+  ought to be a key
+  % are a key
+  ingredient in
   effective public health decision-making during a pandemic. Since April 2020,
@@
 streams---such as confirmed cases and deaths in the COVID-19 pandemic---is
-crucial for understanding disease spread, formulating public policy responses,
-and anticipating future public health resource needs.  A companion paper
+crucial for understanding disease spread, correctly formulating public policy responses,
+and rationally planning future public health resource needs.  A companion paper
 \cite{Reinhart:2021} describes our research group's (Delphi's) efforts in
@@
 September, 2020, we do not have true vintage data before then.  Furthermore, the
-latency of the signal was always at least one week through 2020.
-% However, this signal is never revised after initial publication (confirmed via
-% personal communication with the Google team that produces this signal)
-However, unlike the claims-based signals, there is no reason for revisions to
-occur after initial publication, 
-and furthermore,
-the latency of the signal is not an unavoidable property of the data type, so we
-simply use finalized signal values, with zero latency, in our analysis.
+latency of the signal was always at least one week through 2020.  However,
+unlike the claims-based signals, there is no reason for revisions to occur after
+initial publication, and furthermore, the latency of the signal is not an
+unavoidable property of the data type, so we simply use finalized signal values,
+with zero latency, in our analysis.
 
@@
 when the survey was first launched). The end date was chosen again with a 
-consideration to align both tasks as best as possible, and because very few
-hotspots exist for the several few months of 2021, due to the general and
-gradual decline of the pandemic during this period.
+consideration to align both tasks as best as possible and to coincide with the
+introduction of widely available vaccines.
 
@@
   given location, its behavior mimics that of future case rates).  On
-  the other hand, if an an indicator model does better in periods of
+  the other hand, if an indicator model does better in periods of
   increase, or worse in periods of increase or decrease, its performance is
@@
 To test this hypothesis, we reran all of the prediction experiments but with
-$X_{\ell,t}$ in the each indicator models replaced with suitable random noise
+$X_{\ell,t}$ in each indicator model replaced with suitable random noise
 (bootstrap samples from a signal's history).  The results, shown and explained
@@
 ``fake'' indicator models offered essentially no improvement over the pure AR
-model, which strongly rejects (informally speaking) the implicit regularization
+model, which---informally speaking---strongly rejects the implicit regularization
 hypothesis.

### 2021-10-23 forecast/paper/paper.tex
Mainly minor changes.
* Add piles of refs to the bib for real-time/leadingness.
* Address a few approved minor wording issues.

@@
 \cite{CovidcastPy} packages for convenient data fetching and processing
-tools. In the current paper, we aim to quantify the utility provided by a core
+tools. In the current paper, we quantify the utility provided by a core
 set of these indicators for two fundamental prediction tasks: probabilistic
@@
 to provide an authoritative take on cutting-edge COVID-19 forecasting methods.
-Instead, our purpose here is to provide a rigorous, quantitative assessment of
+Instead, we provide a rigorous, quantitative assessment of
 the utility that several auxiliary indicators---such as those derived from
@@
 While forecasting carries a rich literature that offers a wide range of
-techniques, see, e.g., \cite{Hyndman:2018}, we purposely constrain ourselves to 
+techniques, see, e.g., \cite{Hyndman:2018},
+% put Ioannidis ref here
+we purposely constrain ourselves to 
 very simple models, avoiding common enhancements such as order selection,
@@
 we use in this paper. See, e.g., \cite{Cramer:2021}, for a detailed study
-of death forecasting. Our internal evaluations suggest that the same is true (at
-least, if not to a greater extent) for case forecasting.
+of death forecasting. 
+% added
+Similarly, the simple autoregressive models used here for case forecasting
+are competitive with top models submitted to the Forecast Hub over the same
+period of time (see the Supplement for this analysis).
+% Our internal evaluations suggest that the same is true (at
+% least, if not to a greater extent) for case forecasting.
 
@@
 and facilitates honest backtesting.
+%
+TODO: Here we cite many refs
+\cite{faust2009a,diebold1991a,patterson1995a,croushore2011a,croushore2006a,harvey1983a,mankiw1986a,trivellato1986a,clark2009a,koenig2003a,howrey1978a}
 
@@
 Why do we define the response variables via 7-day averaging? The short answer 
-is robustness: averaging stabilizes the case time series, and accounts for 
+is robustness: averaging stabilizes the case time series, and moderates
 uninteresting artifacts like day-of-the-week effects in the series.  Note that
@@
 cross correlations to case rates over a short time window around $t$.
-The higher this score, the greater it ``leads'' case activity.  Figure 
+The higher this score, the greater it ``leads'' case activity.
+%
+TODO: Here we cite many refs.
+\cite{yeats1972a,emerson1996a,koch1988a,sargent1977a,engle1987a,shiskin1968a,granger1969a,lahiri1991a,koopmans1947a,StockWatson1989,diebold1989a,mitchell1938a,auerbach1982a,hamilton1996a}
+%
+Figure 
 \ref{fig:leading} displays correlations between the leadingness score of an 
@@
 levels of nuance to such an answer that must be explained.  None of the
-indicators that we have investigated appear to be the ``silver bullet'' that one
-might have hoped for, revolutionizing the tractability of the prediction
+indicators that we have investigated revolutionize the tractability of the prediction
 problem, rendering it easy when it was once hard (in the absence of auxiliary

### 2021-10-23 forecast/code/deprecated/compare-to-hub.R

@@ -1,3 +1,7 @@
+
+# Import from the Hub -----------------------------------------------------
+
+library(tidyverse)
 library(aws.s3)
 Sys.setenv("AWS_DEFAULT_REGION" = "us-east-2")
 s3bucket <- get_bucket("forecast-eval")
@@ -5,7 +9,9 @@ n_keeper_weeks <- 6L
 n_keeper_locs <- 50L
 case_scores <- s3readRDS("score_cards_state_cases.rds", s3bucket)
 case_scores <- case_scores %>% 
-  mutate(forecast_date = target_end_date - ahead * 7) %>% 
+  mutate(
+    ahead = ahead * 7 - 2, # forecast on a Monday
+    forecast_date = target_end_date - ahead) %>% 
   # fix weirdnesses about submission dates
   filter(forecast_date < "2021-01-01") %>%
   select(ahead, geo_value, forecaster, target_end_date, wis, forecast_date)
@@ -19,8 +25,9 @@ case_scores <- left_join(
   )
 case_scores <- case_scores %>% 
   filter(forecaster != "COVIDhub-baseline")
+aheads <- case_scores %>% distinct(ahead) %>% pull()
 n_submitted <- case_scores %>% 
-  filter(ahead == 2) %>% 
+  filter(ahead == aheads[2]) %>% 
   group_by(forecaster) %>% 
   summarise(nfcasts = n())
 keepers <- n_submitted %>% 
@@ -29,7 +36,32 @@ keepers <- n_submitted %>%
   pull(forecaster)
 
 case_scores <- case_scores %>% filter(forecaster %in% keepers)
-all_time_performance <- case_scores %>%
+
+
+# Load our models ---------------------------------------------------------
+
+ours <- readRDS("~/Downloads/results_no_october_honest.RDS")
+pop <- covidcast::state_census %>% select(ABBR, POPESTIMATE2019) %>%
+  mutate(geo_value = tolower(ABBR)) %>% select(-ABBR)
+# scale from prop to num
+ours <- left_join(ours, pop) %>% mutate(wis = wis * POPESTIMATE2019 / 1e5) %>%
+  select(-ae, -POPESTIMATE2019)
+
+common_fd <- as.Date(intersect(
+  case_scores %>% select(forecast_date) %>% distinct() %>% pull(),
+  ours %>% select(forecast_date) %>% distinct() %>% pull()),
+  "1970-01-01")
+
+
[…]

### 2021-10-27 forecast/paper/paper.tex
Finish minor paper mods; all but SS, abstract. Code for hub comparison

@@
 techniques, see, e.g., \cite{Hyndman:2018},
-% put Ioannidis ref here
 we purposely constrain ourselves to 
@@
 may well aid in forecasting, such as age-specific effects, holiday adjustments,
-public health mandates, and so on. That said, analyses of forecasts submitted
+public health mandates, and so on.
+% Paragraph from Rob
+Work in \cite{ioannidis2020a} points out many additional difficulties when it
+comes to
+forecasting COVID-19 case rates and deaths. These include bad data, wrong
+assumptions, lack of incorporation of epidemiological features, etc.  We agree
+with most of these 
+and our group has found that COVID-19 forecasting to be extremely challenging.
+However the point of our  paper is to demonstrate the utility of  auxiliary
+indicators for improving COVID-19 forecasting, and to suggest that these
+indicators might be useful for future pandemics. We are not suggesting that
+these indicators make COVID-19 forecasting  easy. 
+Despite these obstacles, analyses of forecasts submitted
+%
 to the COVID-19 Forecast Hub \cite{ForecastHub} by a large community of modelers 
@@
 %
-TODO: Here we cite many refs
-\cite{faust2009a,diebold1991a,patterson1995a,croushore2011a,croushore2006a,harvey1983a,mankiw1986a,trivellato1986a,clark2009a,koenig2003a,howrey1978a}
+The necessity of real-time data for honest forecast evaluations has a long history in econometrics,
+\cite[see,
+e.g.,][]{faust2009a,diebold1991a,patterson1995a,croushore2011a,croushore2006a,harvey1983a,mankiw1986a,trivellato1986a,clark2009a,koenig2003a,howrey1978a}
+but it is far less acknowledged in public health despite its demonstrable importance.
 
@@
 
-In the supplement, we show how to define a quantitative score to measure the
+In the supplement, we define a quantitative score to measure the
 leadingness of an indicator, at any time $t$ and any location $\ell$, based on
@@
 %
-TODO: Here we cite many refs.
-\cite{yeats1972a,emerson1996a,koch1988a,sargent1977a,engle1987a,shiskin1968a,granger1969a,lahiri1991a,koopmans1947a,StockWatson1989,diebold1989a,mitchell1938a,auerbach1982a,hamilton1996a}
+This analysis is closely related to Granger Causality \cite{engle1987a} and
+draws on a large body of prior work that measures leadingness in economic time-series 
+\cite[e.g.,][]{yeats1972a,emerson1996a,koch1988a,sargent1977a,engle1987a,shiskin1968a,granger1969a,lahiri1991a,koopmans1947a,StockWatson1989,diebold1989a,mitchell1938a,auerbach1982a,hamilton1996a}.
 %

### 2021-10-27 forecast/code/deprecated/compare-baselines.R

@@ -1,0 +1,63 @@
+library(evalcast)
+library(covidcast)
+library(tidyverse)
+library(aws.s3)
+Sys.setenv("AWS_DEFAULT_REGION" = "us-east-2")
+s3bucket <- get_bucket("forecast-eval")
+
+
+# Point to the data Addison sent
+ours <- readRDS("~/Downloads/results_no_october_honest.RDS")
+baseline <- ours %>% filter(forecaster == "Baseline") %>%
+  select(-forecaster) %>% rename(strawman_wis = wis)
+
+# these are epiweek / num
+bpreds <- s3readRDS("predictions_cards.rds", s3bucket) %>%
+  filter(forecaster == "COVIDhub-baseline", 
+         signal == "confirmed_incidence_num",
+         forecast_date < "2021-01-01", 
+         ahead %in% 1:4,
+         !(geo_value %in% c("as", "gu", "pr", "vi", "mp", "us")))
+
+# trying with the prop signal, since that's what we use
+actuals <- covidcast_signal("jhu-csse", "confirmed_7dav_incidence_prop", 
+                            end_day = "2021-02-01", geo_type = "state", 
+                            as_of = "2021-05-18") %>% 
+  select(geo_value, time_value, value) %>%
+  rename(target_end_date = time_value, actual = value)
+
+# Population / 100,000
+pop <- covidcast::state_census %>% select(ABBR, POPESTIMATE2019) %>%
+  mutate(geo_value = tolower(ABBR), POPESTIMATE2019 = POPESTIMATE2019 / 1e5) %>% 
+  select(-ABBR)
+
+# Scale epiweek total to daily prop
+bpreds2 <- left_join(bpreds, pop) %>% 
+  mutate(value = value / (7 * POPESTIMATE2019)) %>%
+  select(-POPESTIMATE2019, -incidence_period)
+
+# Score the Hub Baseline
+bscores <- evaluate_predictions(bpreds2, actuals, 
+                                grp_vars = c("ahead", "forecast_date", "geo_value"),
+                                err_measures = list(wis = weighted_interval_score))
+
+comb <- left_join(bscores %>% 
+                    # submissions made on Monday with Sunday target
+                    mutate(ahead = ahead * 7 - 2, 
+                           forecast_date = target_end_date - ahead) %>%
+                    select(forecast_date, target_end_date, geo_value, wis), 
+                  baseline %>% 
[…]

### 2021-10-27 forecast/code/deprecated/compare-to-hub.R

@@ -44,7 +44,7 @@ ours <- readRDS("~/Downloads/results_no_october_honest.RDS")
 pop <- covidcast::state_census %>% select(ABBR, POPESTIMATE2019) %>%
   mutate(geo_value = tolower(ABBR)) %>% select(-ABBR)
 # scale from prop to num
-ours <- left_join(ours, pop) %>% mutate(wis = wis * POPESTIMATE2019 / 1e5) %>%
+ours <- left_join(ours, pop) %>% mutate(wis = wis * POPESTIMATE2019 / 1e5 * 7) %>%
   select(-ae, -POPESTIMATE2019)
 
 common_fd <- as.Date(intersect(

### 2021-10-27 forecast/code/deprecated/hrr_state_aggr.R

@@ -1,0 +1,42 @@
+library(tidyverse)
+fips_hrr <- read_csv(
+  "https://raw.githubusercontent.com/cmu-delphi/covidcast-indicators/main/_delphi_utils_python/delphi_utils/data/fips_hrr_table.csv", 
+  col_types = "cid")
+
+hf <- pivot_wider(fips_hrr, names_from = fips, values_from = weight)
+hf <- hf %>% arrange(hrr)
+hf_mat <- hf %>% select(-hrr) %>% as.matrix()
+hf_mat[is.na(hf_mat)] <- 0
+fh_mat <- MASS::ginv(hf_mat)
+hrrs <- hf %>% pull(hrr) 
+colnames(fh_mat) <- hrrs
+fh <- as_tibble(fh_mat)
+fh$geo_value <- colnames(hf_mat) %>% 
+  substr(1, 2) %>% 
+  covidcast::fips_to_abbr() %>% 
+  tolower()
+sh <- fh %>% 
+  group_by(geo_value) %>% 
+  summarise(across(everything(), sum))
+sh_mat <- sh %>% select(-geo_value) %>% as.matrix()
+sh_long <- sh %>% 
+  pivot_longer(-geo_value, names_to = "hrr", values_to = "multiplier") %>%
+  mutate(hrr = as.character(hrr)) %>%
+  arrange(hrr)
+
+
+# Process preds -----------------------------------------------------------
+
+preds <- preds %>% 
+  rename(hrr = geo_value) %>%
+  group_by(forecaster, forecast_date, ahead, quantile) %>%
+  arrange(hrr) #%>%
+  group_modify( ~ {
+    left_join(sh_long, .x, by = "hrr") %>%
+      mutate(value = multiplier * value) %>%
+      group_by(geo_value) %>%
+      summarise(value = sum(value))
+  })
+  
+## Need to re-sort and threshold to 0 afterward
+

### 2021-10-29 forecast/code/deprecated/compare-to-hub.R
updated hub comparison with all data

@@ -40,12 +40,22 @@ case_scores <- case_scores %>% filter(forecaster %in% keepers)
 
 # Load our models ---------------------------------------------------------
 
-ours <- readRDS("~/Downloads/results_no_october_honest.RDS")
+ours <- readRDS("~/Downloads/results_honest.RDS")
 pop <- covidcast::state_census %>% select(ABBR, POPESTIMATE2019) %>%
   mutate(geo_value = tolower(ABBR)) %>% select(-ABBR)
 # scale from prop to num
-ours <- left_join(ours, pop) %>% mutate(wis = wis * POPESTIMATE2019 / 1e5 * 7) %>%
-  select(-ae, -POPESTIMATE2019)
+ours <- left_join(ours, pop) %>%
+  mutate(wis = wis * POPESTIMATE2019 / 1e5 * 7,
+         forecaster = recode(forecaster,
+                             AR3 = "AR",
+                             AR3CHCLI3 = "CHNG-CLI",
+                             AR3CHCOV3 = "CHNG-COVID",
+                             AR3DVCLI3 = "DV-CLI",
+                             AR3FBCLI3 = "CTIS-CLIIC",
+                             AR3GSSAA3_Subset = "drop",
+                             AR3GSSAA3_Zero = "Google-AA")) %>%
+  filter(forecaster != "drop") %>%
+  select(-ae, -POPESTIMATE2019) 
 
 common_fd <- as.Date(intersect(
   case_scores %>% select(forecast_date) %>% distinct() %>% pull(),
@@ -57,71 +67,36 @@ our_models <- ours %>% filter(forecaster != "Baseline")
 baseline <- ours %>% filter(forecaster == "Baseline") %>%
   select(-forecaster) %>% rename(strawman_wis = wis)
 our_models <- left_join(our_models, baseline)
-all_models <- case_scores %>% bind_rows(our_models) 
+all_models <- bind_rows(hub = case_scores, ours = our_models, .id = "source") 
 
 
 all_time_performance <- all_models %>%
   filter(forecast_date %in% common_fd) %>%
-  group_by(forecaster, ahead) %>%
+  group_by(forecaster, ahead, source) %>%
   summarise(rel_wis = Mean(wis) / Mean(strawman_wis),
-            geo_wis1 = GeoMean((wis + 1) / (strawman_wis + 1)),
-            geo_wis = GeoMean(wis / strawman_wis))
[…]

### 2021-10-29 forecast/paper/supplement-text.tex
add comparison to hub to the supplement, ready for Ryan

@@
 in which we see the heaviest left tails occuring for \chngcli.
+
+\section{Comparing with the COVID-19 Forecast Hub}
+
+Since July of 2020, a number of teams have been submitting real time forecasts
+to the COVID-19 Forecast Hub \cite{ForecastHub}. While a number of forecast
+targets can be submitted, the most related to that used in this paper---the
+7-day trailing average of COVID-19 case incidence rates at the Hospital Referral
+Region---are state-level forecasts for total case incidence over an epiweek.
+Essentially, for each state, teams are asked to predict the total number of
+cases that will be recorded over the week from Monday to Sunday. Forecasts must
+be submitted by Tuesday for the current Sunday as well as the next three
+Sundays. This corresponds to producing forecasts for 5, 12, 19, and 26 days
+ahead. Some teams choose to submit their forecasts a few days earlier, but
+always for the same targets.
+
+To properly compare the models presented in this paper with those submitted to
+the Hub, we limit ourselves to the same set of forecast dates and states, and
+we predict case incidence over a week by: (1) predicting the 7-day trailing
+average case rate (as in the manuscript);
+(2) multiplying by 7 to give the total over a week; (3) multiplying by the state
+population as reported in the 2019 US Census Population Estimate (available in
+the Covidcast R package \cite{CovidcastR}). Finally, we perform the same
+evaluations as in the manuscript Figure 3 and also using geometric mean as
+described above. The result is shown in Figure \ref{fig:compare-to-hub}. Since
+not all teams submitted forecasts to the Forecast Hub for the entire period, we
+only display those which submitted for at least 6 weeks. By both metrics,
+the models examined in this paper are competitive with the top teams, even
+beating the Ensemble for large target horizons.
 

### 2021-10-29 forecast/paper/supplement.Rmd
add comparison to hub to the supplement, ready for Ryan

@@
 
+<!-- Comparison to the Hub -->
+
+```{r compare-to-hub, fig.cap="Performance of our models re-estimated for state-level, epiweek case incidence.
+Our models are colored the same as before.
+The solid light-blue line is the COVIDhub Ensemble forecast.
+The thin grey lines are the forecasts of other teams."}
+
+knitr::include_graphics("fig/compare-states-to-hub.pdf")
+```
+
+\clearpage
+
 <!-- Statistical significance -->

### 2021-10-29 forecast/paper/supplement.tex
add comparison to hub to the supplement, ready for Ryan

@@
 \caption{Relative forecast performance using vintage data and summarizing with the more robust geometric mean.}\label{fig:fcast-adjusted}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/compare-states-to-hub}
+
+}
+
+\caption{Performance of our models re-estimated for state-level, epiweek case incidence.
+Our models are colored the same as before.
+The solid light-blue line is the COVIDhub Ensemble forecast.
+The thin grey lines are the forecasts of other teams.}\label{fig:compare-to-hub}
 \end{figure}

### 2021-10-29 forecast/code/figures/compare-to-hub.R

@@ -7,10 +7,27 @@ Sys.setenv("AWS_DEFAULT_REGION" = "us-east-2")
 s3bucket <- get_bucket("forecast-eval")
 n_keeper_weeks <- 6L
 n_keeper_locs <- 50L
-case_scores <- s3readRDS("score_cards_state_cases.rds", s3bucket)
+# case_scores <- s3readRDS("score_cards_state_cases.rds", s3bucket) 
+case_preds <- s3readRDS("predictions_cards.rds", s3bucket) %>%
+  filter(signal == "confirmed_incidence_num",
+         forecast_date < "2021-01-01",
+         !(geo_value %in% c("as", "gu", "pr", "vi", "mp", "us"))) %>%
+  select(-data_source, -signal, -incidence_period)
+actuals <- covidcast::covidcast_signal("jhu-csse", "confirmed_7dav_incidence_num",
+                                       geo_type = "state", as_of = "2021-05-18",
+                                       end_day = "2021-03-01") %>%
+  select(geo_value, time_value, value) %>%
+  mutate(value = value * 7) %>%
+  rename(target_end_date = time_value, actual = value)
+case_scores <- evalcast::evaluate_predictions(
+  case_preds, actuals, 
+  err_measures = list(wis = evalcast::weighted_interval_score),
+  grp_vars = c("ahead", "forecaster", "forecast_date", "geo_value")
+)
+rm(case_preds)
 case_scores <- case_scores %>% 
   mutate(
-    ahead = ahead * 7 - 2, # forecast on a Monday
+    ahead = ahead * 7 - 2, # forecast on a Tuesday
     forecast_date = target_end_date - ahead) %>% 
   # fix weirdnesses about submission dates
   filter(forecast_date < "2021-01-01") %>%
@@ -40,7 +57,7 @@ case_scores <- case_scores %>% filter(forecaster %in% keepers)
 
 # Load our models ---------------------------------------------------------
 
-ours <- readRDS("~/Downloads/results_honest.RDS")
+ours <- readRDS(here::here("data", "results_honest_states.RDS"))
 pop <- covidcast::state_census %>% select(ABBR, POPESTIMATE2019) %>%
   mutate(geo_value = tolower(ABBR)) %>% select(-ABBR)
 # scale from prop to num
@@ -69,6 +86,9 @@ baseline <- ours %>% filter(forecaster == "Baseline") %>%
 our_models <- left_join(our_models, baseline)
[…]

### 2021-10-29 survey/code/paper-tables.Rmd

@@
   facet_wrap(vars(signal), nrow = 2, ncol = 3, scales = "free_y") +
-  labs(x = "Date", y = "", color = "Census Region") +
+  labs(x = "Date", y = "", color = "Census region") +
   scale_x_date(date_breaks = "3 months",
@@
             linetype = 1, size = 0.5) +
-  scale_color_manual(values = c("lightblue", "gray50")) +
+  scale_color_manual(values = c("steelblue", "black")) +
   scale_y_continuous(limits = c(0, 20), breaks = c(0, 10, 20),

### 2021-10-29 forecast/code/figures/compare-to-hub-ryan.Rmd
Ryan's post hoc work

# Import from the Hub -----------------------------------------------------
```{r eval = FALSE}
library(tidyverse)
library(aws.s3)
Sys.setenv("AWS_DEFAULT_REGION" = "us-east-2")
s3bucket <- get_bucket("forecast-eval")
n_keeper_weeks <- 6L
n_keeper_locs <- 50L
# case_scores <- s3readRDS("score_cards_state_cases.rds", s3bucket) 
case_preds <- s3readRDS("predictions_cards.rds", s3bucket) %>%
  filter(signal == "confirmed_incidence_num",
         forecast_date < "2021-01-01",
         !(geo_value %in% c("as", "gu", "pr", "vi", "mp", "us"))) %>%
  select(-data_source, -signal, -incidence_period)
actuals <- covidcast::covidcast_signal("jhu-csse", "confirmed_7dav_incidence_num",
                                       geo_type = "state", as_of = "2021-05-18",
                                       end_day = "2021-03-01") %>%
  select(geo_value, time_value, value) %>%
  mutate(value = value * 7) %>%
  rename(target_end_date = time_value, actual = value)
case_scores <- evalcast::evaluate_predictions(
  case_preds, actuals, 
  err_measures = list(wis = evalcast::weighted_interval_score),
  grp_vars = c("ahead", "forecaster", "forecast_date", "geo_value")
)
rm(case_preds)
case_scores <- case_scores %>% 
  mutate(
    ahead = ahead * 7 - 2, # forecast on a Tuesday
    forecast_date = target_end_date - ahead) %>% 
  # fix weirdnesses about submission dates
  filter(forecast_date < "2021-01-01") %>%
  select(ahead, geo_value, forecaster, target_end_date, wis, forecast_date)

strawman <- case_scores %>% filter(forecaster == "COVIDhub-baseline")
case_scores <- left_join(
  case_scores, 
  strawman %>% 
    select(forecast_date, target_end_date, geo_value, wis) %>% 
    rename(strawman_wis = wis)
  )
case_scores <- case_scores %>% 
  filter(forecaster != "COVIDhub-baseline")
aheads <- case_scores %>% distinct(ahead) %>% pull()
n_submitted <- case_scores %>% 
  filter(ahead == aheads[2]) %>% 
  group_by(forecaster) %>% 
  summarise(nfcasts = n())
keepers <- n_submitted %>% 
  filter(nfcasts / n_keeper_locs > n_keeper_weeks - .0001) %>% 
  # submitted at least z weeks for x locations
  pull(forecaster)

case_scores <- case_scores %>% filter(forecaster %in% keepers)
```

# Load our models ---------------------------------------------------------
```{r}
library(tidyverse)
ours <- readRDS(here::here("data", "results_honest_states.RDS")) %>%
  filter(forecaster != "AR3GSSAA3_Subset")
pop <- covidcast::state_census %>% select(ABBR, POPESTIMATE2019) %>%
  mutate(geo_value = tolower(ABBR)) %>% select(-ABBR)
# scale from prop to num
ours2 <- left_join(ours, pop, by = "geo_value") %>%
  mutate(wis = wis * POPESTIMATE2019, #/ 1e5 * 7,
         forecaster = recode(forecaster,
                             AR3 = "AR",
                             AR3CHCLI3 = "CHNG-CLI",
                             AR3CHCOV3 = "CHNG-COVID",
                             AR3DVCLI3 = "DV-CLI",
                             AR3FBCLI3 = "CTIS-CLIIC",
                             AR3GSSAA3_Zero = "Google-AA")) %>%
  select(-ae, -POPESTIMATE2019)

# Join sanity check
nrow(ours) == nrow(ours2)
ours = ours2

our_models <- ours %>% filter(forecaster != "Baseline")
baseline <- ours %>% filter(forecaster == "Baseline") %>%
  select(-forecaster) %>% rename(strawman_wis = wis)
our_models <- left_join(our_models, baseline)

# Join sanity check
nrow(our_models) == nrow(ours) - sum(ours$forecaster == "Baseline")

Mean <- function(x) mean(x, na.rm = TRUE)
GeoMean <- function(x) exp(mean(log(x), na.rm = TRUE))

all_models = our_models
all_time_performance <- all_models %>%
  #filter(forecast_date %in% common_fd) %>%
  group_by(forecaster, ahead) %>% #, source) %>%
  summarise(rel_wis = Mean(wis) / Mean(strawman_wis),
            geo_wis = GeoMean((wis) / (strawman_wis))) %>%
  pivot_longer(contains("wis"))

facet_labs <- c(geo_wis = "Geometric mean of WIS", rel_wis = "Mean of WIS")
fcast_colors <- c(RColorBrewer::brewer.pal(5, "Set1"), "#000000")
names(fcast_colors) <- c("CHNG-CLI", "CHNG-COVID", "CTIS-CLIIC", "DV-CLI",
                         "Google-AA", "AR")

ggplot(all_time_performance %>% filter(ahead <= 21), 
       aes(ahead, value, color = forecaster)) +
  geom_line(aes(ahead, value, color = forecaster)) + 
  geom_point(aes(ahead, value, color = forecaster)) +
  scale_color_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
  theme_bw() +
  ylab("Score relative to baseline") +
  geom_hline(yintercept = 1, size = 1.5) +
  xlab("Days ahead") +
  facet_wrap(~ name, labeller = labeller(name = facet_labs), scales = "free_y") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

# Try with other code
```{r}  
library(tidyverse)
library(lubridate)
source(here::here("code", "figures", "results_processing_and_plotting.R"))
source(here::here("code", "figures", "eval_funs.R"))
ours3 <- readRDS(here::here("data", "results_honest_states.RDS")) %>%
  process_res_cases(actuals) %>%
  filter(forecaster != "gs_subset") %>%
  filter(forecaster != "gs_inherit") %>%
  mutate(forecaster = recode(forecaster, gs = "Google-AA"))

unique(ours3$forecaster)[!(unique(ours3$forecaster) %in% names(fcast_colors))]

p1 <- plotter(ours3,
        "wis", Mean, scaler = "strawman_wis", 
        order_of_operations = c("aggr","scale")) +
  ylab("Mean WIS (relative to baseline)")

p2 <- plotter(ours3,
        "wis", GeoMean, scaler = "strawman_wis", 
        order_of_operations = c("aggr","scale")) +
  ylab("Geometric mean WIS (relative to baseline)")

gridExtra::grid.arrange(p2, p1, nrow = 1)
```

# Load our HRR results ---------------------------------------------------------
```{r}
source(here::here("code", "figures", "results_processing_and_plotting.R"))
source(here::here("code", "figures", "eval_funs.R"))

actuals <- readRDS(here::here("data", "confirmed_7dav_incidence_prop.RDS"))
fcasts_honest <- readRDS(here::here("data", "results_honest.RDS")) %>%
  process_res_cases(actuals) %>%
  filter(forecaster != "gs_inherit")

fig3 <- plotter(fcasts_honest, 
        "wis", Mean, scaler = "strawman_wis", 
        order_of_operations = c("aggr","scale")) +
  ylab("Mean WIS (relative to baseline)")
fig3
```

### 2021-10-29 forecast/code/figures/compare-to-hub-ryan.Rmd
Filter to ahead <= 21

@@
 
-p1 <- plotter(ours3,
+p1 <- plotter(ours3 %>% filter(ahead <= 21), 
         "wis", Mean, scaler = "strawman_wis", 
@@
 
-p2 <- plotter(ours3,
+p2 <- plotter(ours3 %>% filter(ahead <= 21), 
         "wis", GeoMean, scaler = "strawman_wis", 
@@
 
-fig3 <- plotter(fcasts_honest, 
+fig3 <- plotter(fcasts_honest %>% filter(ahead <= 21), 
         "wis", Mean, scaler = "strawman_wis", 

### 2021-10-29 forecast/code/deprecated/compare-baselines.R
Revised comparison figures.

@@ -7,7 +7,7 @@ s3bucket <- get_bucket("forecast-eval")
 
 
 # Point to the data Addison sent
-ours <- readRDS("~/Downloads/results_no_october_honest.RDS")
+ours <- readRDS(here::here("data", "results_honest_states.RDS"))
 baseline <- ours %>% filter(forecaster == "Baseline") %>%
   select(-forecaster) %>% rename(strawman_wis = wis)
 
@@ -51,6 +51,9 @@ comb <- left_join(bscores %>%
                   by = c("geo_value", "forecast_date", "target_end_date")) %>%
   filter(!is.na(strawman_wis), !is.na(wis))
 
+
+
+
 ggplot(comb, aes(wis, strawman_wis, color = geo_value)) + 
   geom_point() + 
   theme_bw() +
@@ -61,3 +64,15 @@ ggplot(comb, aes(wis, strawman_wis, color = geo_value)) +
   xlab("WIS of COVIDhub-baseline") + 
   ylab("WIS or our baseline") + 
   coord_equal()
+
+ncomb <- comb %>%
+  mutate(ahead = target_end_date - forecast_date) %>%
+  pivot_longer(contains("wis"), names_to = "forecaster") %>%
+  group_by(forecaster, ahead) %>%
+  summarise(wis  = Mean(value), geo_wis = GeoMean(value)) %>%
+  mutate(forecaster = recode(forecaster, wis = "Hub", strawman_wis = "Ours")) %>%
+  pivot_longer(contains("wis"))
+ggplot(ncomb, aes(as.character(ahead), value, fill = forecaster)) +
+  geom_col(position = "dodge") +
+  facet_wrap(~name)
+  

### 2021-10-29 forecast/code/figures/compare-to-hub.R

@@ -94,7 +94,13 @@ all_time_performance <- all_models %>%
   filter(forecast_date %in% common_fd) %>%
   group_by(forecaster, ahead, source) %>%
   summarise(rel_wis = Mean(wis) / Mean(strawman_wis),
-            geo_wis = GeoMean((wis + 1) / (strawman_wis + 1))) %>%
+            geo_wis = GeoMean((wis) / (strawman_wis))) %>%
+  pivot_longer(contains("wis"))
+
+our_performance <- our_models %>% 
+  group_by(forecaster, ahead) %>%
+  summarise(rel_wis = Mean(wis) / Mean(strawman_wis),
+            geo_wis = GeoMean((wis) / (strawman_wis))) %>%
   pivot_longer(contains("wis"))
 
 facet_labs <- c(geo_wis = "Geometric mean of WIS", rel_wis = "Mean of WIS")
@@ -102,14 +108,8 @@ fcast_colors <- c(RColorBrewer::brewer.pal(5, "Set1"), "#000000")
 names(fcast_colors) <- c("CHNG-CLI", "CHNG-COVID", "CTIS-CLIIC", "DV-CLI",
                          "Google-AA", "AR")
 
-
-ggplot(all_time_performance %>% filter(source == "ours"),
+ggplot(our_performance %>% filter(ahead > 6, ahead < 22),
        aes(ahead, value, color = forecaster)) +
-  geom_line(data = all_time_performance %>% filter(forecaster != "COVIDhub-ensemble"), 
-            aes(group = forecaster),
-            color = "grey80") +
-  geom_line(data = all_time_performance %>% filter(forecaster == "COVIDhub-ensemble"),
-            color = "lightblue", size = 1.5) +
   geom_line(aes(ahead, value, color = forecaster)) + 
   geom_point(aes(ahead, value, color = forecaster)) +
   scale_color_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
@@ -120,4 +120,47 @@ ggplot(all_time_performance %>% filter(source == "ours"),
   facet_wrap(~ name, labeller = labeller(name = facet_labs)) +
   theme(legend.position = "bottom", legend.title = element_blank())
 
-ggsave("paper/fig/compare-states-to-hub.pdf", width = 6.5, height = 4.5)
+ggsave(here::here("paper", "fig", "state-level-performance.pdf"),
+       width = 6.5, height = 4.5)
+
+
+ggplot(all_time_performance %>% 
+         filter(source == "hub", 
[…]

### 2021-10-29 forecast/code/figures/compare-to-hub-ryan.Rmd

@@
 # Load our HRR results ---------------------------------------------------------
-```{r}
+```{r eval = FALSE}
 source(here::here("code", "figures", "results_processing_and_plotting.R"))
@@
 actuals <- readRDS(here::here("data", "confirmed_7dav_incidence_prop.RDS"))
-fcasts_honest <- readRDS(here::here("data", "results_honest.RDS")) %>%
+fcasts_honest <- readRDS(here::here("data", "results_honest.RDS")) 
   process_res_cases(actuals) %>%
-  filter(forecaster != "gs_inherit")
+  filter(forecaster != "gs_subset") %>%
+  filter(forecaster != "gs_inherit") %>%
+  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
 
-fig3 <- plotter(fcasts_honest %>% filter(ahead <= 21), 
+p1 <- plotter(ours3 %>% filter(ahead <= 21), 
         "wis", Mean, scaler = "strawman_wis", 
@@
   ylab("Mean WIS (relative to baseline)")
-fig3
+
+p2 <- plotter(ours3 %>% filter(ahead <= 21), 
+        "wis", GeoMean, scaler = "strawman_wis", 
+        order_of_operations = c("aggr","scale")) +
+  ylab("Geometric mean WIS (relative to baseline)")
+
+gridExtra::grid.arrange(p2, p1, nrow = 1)
 ```

### 2021-10-29 forecast/paper/paper.tex
Fix small typo

@@
 with most of these 
-and our group has found that COVID-19 forecasting to be extremely challenging.
+and our group has found COVID-19 forecasting to be extremely challenging.
 However the point of our  paper is to demonstrate the utility of  auxiliary
@@
 have shown that simple, robust models have consistently been among the
-best-performing over the pandemic, including time series models similar the ones
+best-performing over the pandemic, including time series models similar to the ones
 we use in this paper. See, e.g., \cite{Cramer:2021}, for a detailed study

### 2021-10-29 forecast/paper/supplement.Rmd
Update supplement with comparison figures

--- 2021-10-09 forecast/paper/supplement.Rmd
@@
 
+<!-- Comparison to the Hub -->
+
+```{r compare-to-hub, fig.cap="Performance of the AR model re-estimated for state-level, epiweek case incidence. The thin grey lines are the forecasts of other teams."}
+knitr::include_graphics("fig/compare-states-to-hub.pdf")
+```
+
+\clearpage
+
+```{r state-level-performance, fig.cap="Performance of our models re-estimated for state-level, epiweek case incidence."}
+knitr::include_graphics("fig/state-level-performance.pdf")
+```
+
+\clearpage
+
 <!-- Statistical significance -->

### 2021-10-29 forecast/paper/supplement.tex
Update supplement with comparison figures

--- 2021-10-09 forecast/paper/supplement.tex
@@
 \caption{Relative forecast performance using vintage data and summarizing with the more robust geometric mean.}\label{fig:fcast-adjusted}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/compare-states-to-hub}
+
+}
+
+\caption{Performance of the AR model re-estimated for state-level, epiweek case incidence. The thin grey lines are the forecasts of other teams.}\label{fig:compare-to-hub}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/state-level-performance}
+
+}
+
+\caption{Performance of our models re-estimated for state-level, epiweek case incidence.}\label{fig:state-level-performance}
 \end{figure}

### 2021-10-29 forecast/code/figures/compare-to-hub-ryan.Rmd
First push done
- Rewrite sig statement
- Rewrite abstract
- Pass over and rewrite muhc of intro
- Generally trying to describe the sience as much as possible,
  while maintaining language that I believe is accurate and not
  over-stating our conclusions

@@
 ours2 <- left_join(ours, pop, by = "geo_value") %>%
-  mutate(wis = wis * POPESTIMATE2019, #/ 1e5 * 7,
+  mutate(wis = wis, #* POPESTIMATE2019, #/ 1e5 * 7,
          forecaster = recode(forecaster,
@@
 ours = ours2
+```
 
+```{r}
 our_models <- ours %>% filter(forecaster != "Baseline")
@@
 
-# Join sanity check
+# Sanity check: left join
 nrow(our_models) == nrow(ours) - sum(ours$forecaster == "Baseline")
+
+# Sanity check: missingness
+our_models %>%
+  group_by(forecaster) %>%
+  summarize(sum(is.na(wis)))
+
+our_models %>% summarize(sum(is.na(strawman_wis)))
 
@@
   theme(legend.position = "bottom", legend.title = element_blank())
+```
+
+```{r}
+p1 <- ours %>% 
+  group_by(forecaster, ahead) %>%
+  summarize(mean_wis = Mean(wis)) %>%
+  ggplot(aes(x = ahead, y = mean_wis, color = forecaster)) + 
+  geom_line() + geom_point() +   
+  xlab("Days ahead") + ylab("Mean WIS") + 
+  scale_color_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
+  theme_bw() +
+  theme(legend.position = "bottom", legend.title = element_blank())
+
+p2 <- ours %>% 
+  group_by(forecaster, ahead) %>%
+  summarize(geomean_wis = GeoMean(wis)) %>%
+  ggplot(aes(x = ahead, y = geomean_wis, color = forecaster)) + 
+  geom_line() + geom_point() +   
+  xlab("Days ahead") + ylab("Geo mean WIS") + 
+  scale_color_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
+  theme_bw() +
+  theme(legend.position = "bottom", legend.title = element_blank())
+
+gridExtra::grid.arrange(p2, p1, nrow = 1)
 ```

### 2021-10-29 forecast/paper/paper.tex
First push done
- Rewrite sig statement
- Rewrite abstract
- Pass over and rewrite muhc of intro
- Generally trying to describe the sience as much as possible,
  while maintaining language that I believe is accurate and not
  over-stating our conclusions

@@
 % Please add a significance statement to explain the relevance of your work
-\significancestatement{Forecasting is a critical element in the public health
-  response to any fast-moving epidemic or pandemic.  Predicting the future
-  spread of some disease process, based on disease activity in the recent
-  past---say, using autoregressive modeling from the time series literature---is 
-  a basic and widely-used method.  In this paper we study whether including
-  auxiliary indicators as additional features in such an autoregressive model
-  can be useful in COVID-19 case forecasting and hotspot prediction.  We find
-  that indicators based on de-identified medical insurance claims, self-reported
-  symptoms through online surveys, and COVID-related internet search activity
-  all provide a nontrivial improvement in forecast accuracy, though their
-  utility depends on the pandemic's local dynamics (upswing, downswing, or flat
-  period) at prediction time. 
+\significancestatement{Validated forecasting methodology should be a key 
+  element in the public health response to any fast-moving epidemic or 
+  pandemic. A widely-used model for predicting the future spread of a temporal
+  process, based on past data, is an autoregressive (AR) model. While basic,
+  such an AR model (properly trained) can already be competitive with top models
+  in operational use for COVID-19 forecasting. In this paper, we 
+  exhibit five auxiliary indicators---based on de-identified medical insurance
+  claims, self-reported symptoms via online surveys, and COVID-related Google
+  searches---that further improve the predictive accuracy of an AR model in
+  COVID-19 forecasting. The most substantial gains appear to be in quiescent  
+  times; but the Google search indicator appears to also offer improvements
+  during upswings in pandemic activity.   
 }
@@
 \begin{abstract}
-  Reliable, short-term forecasts of traditional public health reporting streams
-  (such as cases, hospitalizations, and deaths)
-  ought to be a key
-  % are a key
-  ingredient in
-  effective public health decision-making during a pandemic. Since April 2020,
-  our research group has worked with data partners to collect, curate, and make
-  publicly available numerous real-time COVID-19 indicators, providing multiple
-  views of pandemic activity.  This paper studies the utility of these
-  indicators from a forecasting perspective. We focus on five indicators,
-  derived from medical insurance claims data, web search queries, and online
-  survey responses. For each indicator, we ask whether its inclusion in a simple
-  model leads to improved predictive accuracy relative to a similar model
-  excluding it. We consider both probabilistic forecasting of confirmed COVID-19 
-  case rates and binary prediction of case ``hotspots''.  Since the values of 
-  indicators (and case rates) are commonly revised over time, we take special
-  care to ensure that the data provided to a forecaster is the version that
-  would have been available at the time the forecast was made.   
-  % We demonstrate how less careful backtesting can lead to misleading
-  % retrospective evaluations.   
-  Our  analysis shows that consistent but modest gains in predictive accuracy 
-  are obtained by using these indicators, and furthermore, these gains are
-  related to periods in which the auxiliary indicators behave as ``leading
-  indicators'' of case rates.     
+  Short-term forecasts of traditional public health reporting streams (such as
+  cases, hospitalizations, and deaths) are a key input to public health
+  decision-making during a pandemic. Since early 2020, our research group has
+  worked with data partners to collect, curate, and make publicly available
+  numerous real-time COVID-19 indicators, providing multiple views of pandemic
+  activity in the U.S.  This paper studies the utility of five such
+  indicators---derived from de-identified medical insurance claims,
+  self-reported symptoms from online surveys, and COVID-related Google search
+  activity---from a forecasting perspective. For each indicator, we ask whether
+  its inclusion in an autoregressive (AR) model leads to improved predictive
+  accuracy relative to the same model excluding it. Such an AR model, without
+  external features, is already competitive with many top COVID-19 forecasting
+  models in use today. Our analysis reveals that (a) inclusion of each of these
+  five indicators improves on the overall predictive accuracy of the AR model;
+  (b) predictive gains are in general most pronounced during times in which
+  COVID cases are trending in ``flat'' or ``down'' directions; (c) one
+  indicator, based on Google searches, seems to be particularly helpful
+  during ``up''  trends.      
 \end{abstract}
@@
 streams---such as confirmed cases and deaths in the COVID-19 pandemic---is
-crucial for understanding disease spread, correctly formulating public policy responses,
-and rationally planning future public health resource needs.  A companion paper
-\cite{Reinhart:2021} describes our research group's (Delphi's) efforts in
-curating and maintaining a database of real-time indicators that track COVID-19
-activity and other relevant phenomena. The signals (a term we use synonomously
-with ``indicators'') in this database are accessible through the COVIDcast API
-\cite{CovidcastAPI}, with associated R \cite{CovidcastR} and Python
-\cite{CovidcastPy} packages for convenient data fetching and processing
-tools. In the current paper, we quantify the utility provided by a core
-set of these indicators for two fundamental prediction tasks: probabilistic
-forecasting of COVID-19 case rates and prediction of future COVID-19 case
-hotspots (defined by the event that a relative increase in COVID-19 cases
-exceeds a certain threshold).
+crucial for understanding disease spread, correctly formulating public policy
+responses, and rationally planning future public health resource needs.  A
+companion paper \cite{Reinhart:2021} describes our research group's efforts,
+beginning in April 2020, in curating, and maintaining a database of real-time
+indicators that track COVID-19 activity and other relevant phenomena. The
+signals (a term we use synonomously with ``indicators'') in this database are
+accessible through the COVIDcast API \cite{CovidcastAPI}, as well as associated
+R \cite{CovidcastR} and Python \cite{CovidcastPy} packages, for convenient data  
+fetching and processing. In the current paper, we quantify the utility provided
+by a core set of these indicators for two fundamental prediction tasks:
+probabilistic forecasting of COVID-19 case rates and prediction of future
+COVID-19 case hotspots (defined by the event that a relative increase in
+COVID-19 cases exceeds a certain threshold).
 
@@
 to provide an authoritative take on cutting-edge COVID-19 forecasting methods.
-Instead, we provide a rigorous, quantitative assessment of
-the utility that several auxiliary indicators---such as those derived from
-internet surveys or medical insurance claims---provide in tasks that involve
-predicting future trends in confirmed COVID-19 cases. To assess such utility in
-as simple terms as possible, we center our study in the framework of a basic
-autoregressive model (in which COVID cases in the near future are predicted from  
-a linear combination of COVID cases in the near past), and ask whether the 
-inclusion of an auxiliary indicator as an additional feature in such a model
-improves its predictions.
+Similarly, some authors, e.g., \cite{ioannidis2020a}, have pointed out numerous
+mishaps of forecasting during the pandemic and it is not our general intent to
+fix them here. Instead, we start with a basic and yet reasonably effective
+predictive model for predicting future trends in COVID-19 cases, and present a
+rigorous, quantitative assessment of the added value provided by auxiliary
+indicators, that are derived from data sources that operate outside of 
+traditional public health streams. In particular, we consider five indicators
+derived from de-identified medical insurance claims, self-reported symptoms from
+online surveys, and COVID-related Google searches.
 
-While forecasting carries a rich literature that offers a wide range of
-techniques, see, e.g., \cite{Hyndman:2018},
-we purposely constrain ourselves to 
-very simple models, avoiding common enhancements such as order selection,
-correction of outliers/anomalies in the data, and inclusion of regularization or
-nonlinearities. For the same reason, we do not account for other factors that
-may well aid in forecasting, such as age-specific effects, holiday adjustments,
-public health mandates, and so on.
-% Paragraph from Rob
-Work in \cite{ioannidis2020a} points out many additional difficulties when it
-comes to
-forecasting COVID-19 case rates and deaths. These include bad data, wrong
-assumptions, lack of incorporation of epidemiological features, etc.  We agree
-with most of these 
-and our group has found COVID-19 forecasting to be extremely challenging.
-However the point of our  paper is to demonstrate the utility of  auxiliary
-indicators for improving COVID-19 forecasting, and to suggest that these
-indicators might be useful for future pandemics. We are not suggesting that
-these indicators make COVID-19 forecasting  easy. 
-Despite these obstacles, analyses of forecasts submitted
-%
-to the COVID-19 Forecast Hub \cite{ForecastHub} by a large community of modelers 
-have shown that simple, robust models have consistently been among the
-best-performing over the pandemic, including time series models similar to the ones
-we use in this paper. See, e.g., \cite{Cramer:2021}, for a detailed study
-of death forecasting. 
-% added
-Similarly, the simple autoregressive models used here for case forecasting
-are competitive with top models submitted to the Forecast Hub over the same
-period of time (see the Supplement for this analysis).
-% Our internal evaluations suggest that the same is true (at
-% least, if not to a greater extent) for case forecasting.
+To assess this value in as direct terms as possible, we base our study around a
+very simple basic model: an autoregressive model, in which COVID cases in the
+near future are predicted using a linear combination of COVID cases in the near
+past. Forecasting carries a rich literature, offering a wide range of
+sophisticated techniques, see, e.g., \cite{Hyndman:2018} for a review; however,
+we purposely avoid enhancements such as order selection, correction of
+outliers/anomalies in the data, and inclusion of regularization or
+nonlinearities. Similarly, we do not account for other factors that may well aid
+in forecasting, such as age-specific effects, holiday adjustments, and the
+effects of public health mandates. All that said, despite its simplicity, the
+basic autoregressive model that we consider in this paper exhibits competitive
+performance (see the Supplementary Materials for details) with many of the top
+COVID-19 case forecasters submitted to the U.S.\ COVID-19 Forecast Hub
+\cite{ForecastHub}, which is the official source of forecasts used in public
+communications by the U.S. CDC. The strong performance of the autoregressive
+model here is in line with the fact that simple, robust models have also
+consistently been among the best-performing ones for COVID-19 death
+forecasting \cite{Cramer:2021}.
 
@@
 provide relevant information that is not otherwise contained in past values of
-the case rate series itself. We will assess this in the most direct way
-possible: by inspecting the difference in predictive performance of simple
-autoregressive models trained with and without access to past values of a
-particular indicator.
+the case rate series itself. We assess this by inspecting the difference in
+predictive performance of simple autoregressive models trained with and without
+access to past values of a particular indicator.
 
-There is another, more subtle issue in evaluating predictive utility that 
-deserves explicit mention, as it will play a key role in our analysis.
-Signals computed from surveillance streams will often be subject to  
-latency and/or revision. For example, a signal based on aggregated medical
-insurance claims may be available after just a few days, but it can then be
-substantially revised over the next several weeks as additional claims are
-submitted and/or processed late. Correlations between such a signal and case
-rates calculated ``after the fact'' (i.e., computed retrospectively, using the
-finalized values of this signal) will not deliver an honest answer to the    
-question of whether this signal would have been useful in real time. Instead,
-we build predictive models using only the data that would have been available
-\textit{as of} the prediction date, and compare the ensuing predictions in terms 
-of accuracy. To do so, we leverage Delphi's \texttt{evalcast} R package 
-\cite{EvalcastR}, which plugs into the COVIDcast API's data versioning system,  
-and facilitates honest backtesting.
-%
-The necessity of real-time data for honest forecast evaluations has a long history in econometrics,
-\cite[see,
-e.g.,][]{faust2009a,diebold1991a,patterson1995a,croushore2011a,croushore2006a,harvey1983a,mankiw1986a,trivellato1986a,clark2009a,koenig2003a,howrey1978a}
-but it is far less acknowledged in public health despite its demonstrable importance.
+We find that each of the five indicators we consider--- two based on
+COVID-related outpatient visits from medical insurance claims, one on
+self-reported symptoms from online surveys, and one on Google searches for
+anosmia or ageusia---provide an overall improvement in accuracy when
+incorporated into the autorgressive model. This is true both for COVID-19 case 
+forecasting and hotspot prediction. Further analysis reveals that the gains in
+accuracy depend on the pandemic's dynamics at prediction time: the biggest  
+gains in accuracy appear during times in which cases are ``flat'' or trending
+``down''; but the indicator based on Google searches offers a most notable
+improvement when cases are trending ``up''.
+
+Careful handling of data revisions plays a key role in our analysis. Signals 
+computed from surveillance streams are often subject to latency and/or 
+revision. For example, a signal based on aggregated medical insurance claims may
+be available after just a few days, but it can then be substantially revised
+over the next several weeks as additional claims are submitted and/or processed
+late. Correlations between such a signal and case rates calculated ``after the
+fact'' (i.e., computed retrospectively, using the finalized values of this
+signal) will not deliver an honest answer to the question of whether this signal
+would have been useful in real time. Instead, we build predictive models using
+only the data that would have been available \textit{as of} the prediction date,
+and compare the ensuing predictions in terms of accuracy. The necessity of
+real-time data for honest forecast evaluations has been recognized in
+econometrics for a long time, see
+\cite{howrey1978a, harvey1983a, mankiw1986a, trivellato1986a, diebold1991a,
+  patterson1995a, koenig2003a, croushore2006a, clark2009a, faust2009a,
+  croushore2011a}; but it appears to be far less acknowledged in public health,
+despite its demonstrable importance.
 
@@
   Lei:2018, Rinaldo:2019, Williamsom:2020, Zhang:2020, Dai:2021, Fryer:2021}.
-Our take on this problem is in line with much of this literature; however, in
-order to avoid making any explicit assumptions, we do not attempt to make formal 
-significance statements, and instead, broadly examine the stability of our
-conclusions with respect to numerous modes of analysis.
+Our take in the current work is in line with much of this literature; however,
+in order to avoid making any explicit assumptions, we do not attempt to make
+formal  significance statements, and instead, broadly examine the stability of
+our conclusions with respect to numerous modes of analysis.
 

### 2021-10-29 forecast/paper/supplement-text.tex

@@
 
-Since July of 2020, a number of teams have been submitting real time forecasts
+Since July of 2020, teams have been submitting real time forecasts
 to the COVID-19 Forecast Hub \cite{ForecastHub}. While a number of forecast
@@
 evaluations as in the manuscript Figure 3 and also using geometric mean as
-described above. The result is shown in Figure \ref{fig:compare-to-hub}. Since
-not all teams submitted forecasts to the Forecast Hub for the entire period, we
+described above.
+For the purpose of the comparison, we show only the AR model along with
+submissions collected by the Forecast Hub in Figure \ref{fig:compare-to-hub}.
+Since not all teams submitted forecasts for the entire period, we
 only display those which submitted for at least 6 weeks. By both metrics,
-the models examined in this paper are competitive with the top teams, even
-beating the Ensemble for large target horizons.
+the AR model examined in this paper is competitive with the submitted by top teams, even
+beating the Ensemble for smaller target horizons.
+
+We also are undertook the same analysis shown in Figure 3 in the manuscript and Figure
+\ref{fig:fcast-adjusted} above for all models
+(the AR, as well as the indicator-assisted models). This comparison is shown in Figure
+\ref{fig:state-level-performance}. 
+%
+TODO: some text about why indicators don't help as much.
 

### 2021-10-29 forecast/paper/paper.tex
small changes to intro/discussion

@@
   process, based on past data, is an autoregressive (AR) model. While basic,
-  such an AR model (properly trained) can already be competitive with top models
+  such an AR model (properly trained) is already competitive with top models
   in operational use for COVID-19 forecasting. In this paper, we 
@@
 that motivates the current article: is the information contained in an indicator
-demonstrably useful for the prediction tasks we care about? For such a question,
-lagged correlations (e.g., measuring the correlation between an indicator and
-COVID case rates several days in the future) move us in the right direction, but
-still do not move us all the way there. The question about \textit{utility for
-  prediction} is focused on a much higher standard than simply asking about 
+demonstrably useful for the prediction tasks we care about?
+% For such a question,
+% lagged correlations (e.g., measuring the correlation between an indicator and
+% COVID case rates several days in the future) move us in the right direction, but
+% still do not move us all the way there.
+Demonstrating \textit{utility for
+  prediction} is a much higher standard than simply asking about 
 correlations; to be useful in forecast or hotspot models, an indicator must
 provide relevant information that is not otherwise contained in past values of
-the case rate series itself. We assess this by inspecting the difference in
+the case rate series itself (see the pioneering work on Granger Causality
+\cite{engle1987a,Granger:1969} as well as detailed discussion below). We
+assess this directly by inspecting the difference in  
 predictive performance of simple autoregressive models trained with and without
@@
 \subsection{Effects of Leading or Lagging Behavior}
+\label{sec:leading-lagging}
 
@@
 levels of nuance to such an answer that must be explained.  None of the
-indicators that we have investigated revolutionize the tractability of the prediction
-problem, rendering it easy when it was once hard (in the absence of auxiliary
-information).  Rather, the gains in accuracy from the indicator models (over an
+indicators that we have investigated are transformative, rendering the
+prediction problem  easy when it was once hard (in the absence of auxiliary
+information).
+% This seems like a place to reference Superforecasters (or related) if we're
+% going to do it. ``The expectation should never be
+% that there is a silver bullet. Too much solid work refutes that. Again lack of
+% prior art citations to some very solid prior empirical work, and repetition of
+% a long discredited Hedgehog/Silver bullet strawman.''  
+Rather, the gains in accuracy from the indicator models (over an
 autoregressive model based only on past case rates) appear to be nontrivial, and
@@
 while devoting most of our effort to accounting for as much of the complexity of
-the underlying data and evaluation as possible.  In fact, our paper is as much
-about demonstrating how one might address questions about model comparisons and
-evaluation in forecasting and hotspot prediction in general, as it is about
-providing rigorous answers to such questions in the context of COVID-19 case
-rates in particular.  We hope that others will leverage our framework, and build
-on it, so that it can be used to guide work that advances the frontier of
-predictive modeling for epidemics and pandemics.
+the underlying data and evaluation as possible.
+% The below was held out as particularly anathema to the PNAS ethos.
+%
+% In fact, our paper is as much
+% about demonstrating how one might address questions about model comparisons and
+% evaluation in forecasting and hotspot prediction in general, as it is about
+% providing rigorous answers to such questions in the context of COVID-19 case
+% rates in particular.  We hope that others will leverage our framework, and build
+% on it, so that it can be used to guide work that advances the frontier of
+% predictive modeling for epidemics and pandemics.
+%
+% Added by DJM
+Despite their simplicity, these models are competitive with top forecasting
+models submitted to the COVID-19 Forecast Hub in real time, and they are
+demonstrably robust to the careful scrutiny undertaken here.
 

### 2021-10-30 forecast/paper/paper.tex
Second push done
- Incorporate sentence update in intro from Jacob
- Take a careful pass through rest of the paper, various small
  edits or copyedits
- Restore some of the original discussion, but try to avoid having
  it sound like "use my method"

@@
 % Please add a significance statement to explain the relevance of your work
-\significancestatement{Validated forecasting methodology should be a key 
+\significancestatement{Validated forecasting methodology should be a vital  
   element in the public health response to any fast-moving epidemic or 
@@
   process, based on past data, is an autoregressive (AR) model. While basic,
-  such an AR model (properly trained) is already competitive with top models
-  in operational use for COVID-19 forecasting. In this paper, we 
-  exhibit five auxiliary indicators---based on de-identified medical insurance
-  claims, self-reported symptoms via online surveys, and COVID-related Google
+  such an AR model (properly trained) is already competitive with the top models
+  in operational use for COVID-19 forecasting. In this paper, we exhibit five
+  auxiliary indicators---based on de-identified medical insurance claims,
+  self-reported symptoms via online surveys, and COVID-related Google
   searches---that further improve the predictive accuracy of an AR model in
@@
 \begin{abstract}
-  Short-term forecasts of traditional public health reporting streams (such as
-  cases, hospitalizations, and deaths) are a key input to public health
+  Short-term forecasts of traditional streams from public health reporting (such 
+  as cases, hospitalizations, and deaths) are a key input to public health
   decision-making during a pandemic. Since early 2020, our research group has
@@
 that motivates the current article: is the information contained in an indicator
-demonstrably useful for the prediction tasks we care about?
-% For such a question,
-% lagged correlations (e.g., measuring the correlation between an indicator and
-% COVID case rates several days in the future) move us in the right direction, but
-% still do not move us all the way there.
-Demonstrating \textit{utility for
-  prediction} is a much higher standard than simply asking about 
-correlations; to be useful in forecast or hotspot models, an indicator must
-provide relevant information that is not otherwise contained in past values of
-the case rate series itself (see the pioneering work on Granger Causality
-\cite{engle1987a,Granger:1969} as well as detailed discussion below). We
-assess this directly by inspecting the difference in  
-predictive performance of simple autoregressive models trained with and without
-access to past values of a particular indicator.
+demonstrably useful for the prediction tasks we care about? Note that even
+lagged correlations cannot deliver a complete answer. Demonstrating
+\textit{utility for prediction} is a much higher standard than simply asking
+about correlations; to be useful in forecast or hotspot models, an indicator
+must provide relevant information that is not otherwise contained in past values
+of the case rate series itself (cf.\ the pioneering work on Granger  causality
+\cite{Granger:1969, engle1987a}, as well as the further references given 
+below). We assess this directly by inspecting the difference in predictive  
+performance of simple autoregressive models trained with and without access to  
+past values of a particular indicator.
 
@@
 real-time data for honest forecast evaluations has been recognized in
-econometrics for a long time, see
-\cite{howrey1978a, harvey1983a, mankiw1986a, trivellato1986a, diebold1991a,
-  patterson1995a, koenig2003a, croushore2006a, clark2009a, faust2009a,
-  croushore2011a}; but it appears to be far less acknowledged in public health,
-despite its demonstrable importance.
+econometrics for a long time, see \cite{howrey1978a, harvey1983a, mankiw1986a,
+  trivellato1986a, diebold1991a, patterson1995a, koenig2003a, croushore2006a,
+  clark2009a, faust2009a, croushore2011a}; but it is often overlooked in
+epidemic forecasting despite its critical importance, see, e.g.,
+\cite{chakraborty2018know}.
 
@@
 \end{itemize}
-In short, we choose these indicators because, conceptually speaking, they
-measure aspects of an individual's disease progression that would plausibly
-precede the occurence of (at worst, co-occur  with) the report of a positive
-COVID-19 test, through standard public health reporting streams.
+We choose these indicators because, conceptually speaking, they measure aspects
+of an individual's disease progression that would plausibly precede the  
+occurence of (at worst, co-occur  with) the report of a positive COVID-19 test,
+through standard public health reporting streams.
 
@@
 containing Los Angeles, which has more than 10 million people), generally HRRs  
-are much more homogenous in size than the (approximately) 3200 counties,  
+are much more homogenous in size than the (approximately) 3200 U.S. counties,   
 and they serve as a nice middle ground in between counties and states.
@@
 Because Google's team did not start publishing this signal until early
-September, 2020, we do not have true vintage data before then.  Furthermore, the
-latency of the signal was always at least one week through 2020.  However,
-unlike the claims-based signals, there is no reason for revisions to occur after
-initial publication, and furthermore, the latency of the signal is not an
-unavoidable property of the data type, so we simply use finalized signal values,
-with zero latency, in our analysis.
+September, 2020, we do not have true vintage data before then; the latency of 
+the signal was always at least one week through 2020.  However, unlike the
+claims-based signals, there is no reason for revisions to occur after initial
+publication, and furthermore, the latency of the signal is not an unavoidable
+property of the data type, so we simply use finalized signal values, with zero
+latency, in our analysis.
 
@@
 is robustness: averaging stabilizes the case time series, and moderates
-uninteresting artifacts like day-of-the-week effects in the series.  Note that
-we can also equivalently view this (equivalent up to a constant factor) as
+uninteresting artifacts like day-of-the-week effects in the data.  We note
+that we can also equivalently view this (equivalent up to a constant factor) as 
 predicting the HRR-level case incidence rate \textit{summed} over some 7-day
@@
 the CTIS-CLI-in-community signal, which does not exist before early April 2020,
-when the survey was first launched). The end date was chosen again with a 
-consideration to align both tasks as best as possible and to coincide with the
-introduction of widely available vaccines.
+when the survey was first launched).
 
@@
 levels $\tau' > \tau$, the predicted quantile at level $\tau$ exceeds the
-predicted quantile at level $\tau'$), we apply a simple post-hoc sorting.  See
+predicted quantile at level $\tau'$), we apply a simple post hoc sorting.  See 
 Figure~\ref{fig:trajectory} for an example forecast.
@@
 training and evaluation where, on average, fewer than 30 cases (this refers to a
-count, not a rate) are observed over the preceding 7 days. This avoids having to
+count, not a rate) are observed over the prior 7 days. This avoids having to
 make arbitrary calls for a hotspot (or lack thereof) based on small counts.
@@
 expectation is minimized by the population quantiles of the target variable.
-The use of WIS in COVID-19 forecast scoring is discussed in \cite{Bracher:2021};
+The use of WIS in COVID-19 forecast scoring was proposed by \cite{Bracher:2021}; 
 WIS is also the main evaluation metric used in the COVID-19 Forecast Hub. More
@@
 especially noteworthy in our current study because it reveals WIS to be the same
-as the quantile regression loss that we use to train our forecasting models
-(i.e., we fit by optimizing WIS averaged over the training data).
+as the quantile regression loss that we use to train our forecasting models.
+%(i.e., we fit by optimizing WIS averaged over the training data).
 
@@
 This is the case for the Google-AA signal (due to the way in which the
-underlying search trends time series is self-normalized, see
+Google search trends time series is self-normalized, see
 \cite{GoogleSymptoms}) and the claims-based signals (due to market-share 
@@
 these signals, because we train the models on data pooled over all locations.
-In the current paper, we do not attempt to address this issue (it is again a
-topic of ongoing work in our group), and we simply note that location-specific
-effects (or pre-processing to remove spatial bias) would likely improve the
-performance of the models involving Google-AA and the claims-based indicators.
+In the current paper, we do not attempt to address this issue, and we simply
+note that location-specific effects (or pre-processing to remove spatial bias)
+would likely improve the performance of the models involving Google-AA and the
+claims-based indicators.
 
@@
 \eqref{eq:forecast_ar_x}, or hotspot prediction, \eqref{eq:hotspot_ar_x}, that
-uses this particular indicator as a feature (the meaning should be clear from
+uses this same indicator as a feature (the meaning should be clear from
 the context).  So, for example, the CHNG-CLI model in forecasting is the one in  
@@
 
-We begin with a summary of the high-level conclusions.
+Below is summary of the high-level conclusions.
 
@@
   gains in both forecasting and hotspot prediction.  CHNG-CLI is
-  inconsistent: it provides a big gain in hotspot prediction, but little gain 
+  inconsistent: it provides a large gain in hotspot prediction, but little gain  
   in forecasting (it seems to be hurt by a notable lack of robustness, due
@@
   and DV-CLI). More rarely does an indicator model tend to beat AR when case
-  rates are increasing, but there appears to be some evidence of this for the
+  rates are increasing, but there appears to be evidence of this for the
   Google-AA model.   
@@
 
-\item In this same analysis, when an indicator model performs better than AR in
-  a decreasing period, this tends to co-occur with instances in which the
+\item In this same analysis, when an indicator model performs better than AR 
+  in a decreasing period, this tends to co-occur with instances in which the 
   indicator ``leads'' case rates (meaning, roughly, on a short-time scale in a
@@
 
-Code to reproduce all results (which uses the \texttt{evalcast} R package) can 
-be found at
+Code to reproduce all results can be found at
 \url{https://github.com/cmu-delphi/covidcast-pnas/tree/main/forecast/code}.
@@
 
-One might ask if the benefits we observe in forecasting and hotspot
-prediction have anything to do with the actual indicator themselves. A plausible  
-alternative explanation is that the indicators are simply providing \textit{implicit
-  regularization} on top of the basic AR model, in the same way any noise
-variable might, when we include them as lagged features in
-\eqref{eq:forecast_ar_x} and \eqref{eq:hotspot_ar_x}.
+One might ask if the benefits observed in forecasting and hotspot
+prediction have anything to do with the actual auxiliary indicators
+themselves. A plausible alternative explanation is that the indicators are
+just providing \textit{implicit regularization} on top of the basic AR model,
+in the same way any noise variable might, if we were to use it to create lagged
+features in \eqref{eq:forecast_ar_x} and \eqref{eq:hotspot_ar_x}.
 
 To test this hypothesis, we reran all of the prediction experiments but with
-$X_{\ell,t}$ in each indicator model replaced with suitable random noise
+$X_{\ell,t}$ in each indicator model replaced by suitable random noise
 (bootstrap samples from a signal's history).  The results, shown and explained
-more precisely in the Supplementary Materials, are vastly different (worse) than
-the original set of results.  In both forecasting and hotspot prediction, the
-``fake'' indicator models offered essentially no improvement over the pure AR
-model, which---informally speaking---strongly rejects the implicit regularization
-hypothesis.
+more precisely in the Supplement, are vastly different (i.e., worse) than the
+original set of results.  In both forecasting and hotspot prediction, the
+``fake'' indicator models offer essentially no improvement over the pure AR
+model, which---informally speaking---strongly rejects the implicit
+regularization hypothesis.
 
 On the topic of regularization, it is also worth noting that the use of 
-$\ell_1$ regularization (tuned by cross-validation) in fitting any of the models
-in \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar}, or 
-\eqref{eq:hotspot_ar_x} did not generally improve their performance
-(experiments not shown).  This is likely due to the fact that the number of
-training samples is large compared to the number of features (6,426 training
-samples and only 3--6 features). 
+$\ell_1$ regularization (tuned using cross-validation) in fitting any of the
+models in \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x},
+\eqref{eq:hotspot_ar}, or  \eqref{eq:hotspot_ar_x} did not generally improve
+their performance (experiments not shown).  This is likely due to the fact that
+the number of training samples is large compared to the number of features
+(6,426 training samples and only 3--6 features).  
 % DJM: and that proper CV requires longer training windows, exacerbating the
@@
 increasing period, we use the same definition we used for the hotspot task in
-Table \ref{tab:analysis_tasks}.  Therefore all hotspots are ``up'', while all
-non-hotspots are either ``flat'' or ``down''.  For the ``down'' scenario, we 
-simply use the opposite of the hotspot definition: $Y_{\ell,t}$ decreases by
-more than 20\% relative to the preceding week.
+Table \ref{tab:analysis_tasks}.  Therefore, all hotspots are labeled ``up'',
+while all non-hotspots are either ``flat'' or ``down''.  For the ``down''
+scenario, we simply use the opposite of the hotspot definition: $Y_{\ell,t}$
+decreases by more than 20\% relative to the preceding week.
 
-While the performance of all forecasters, including AR, will generally degrade
-in up periods, different models exhibit different and interesting patterns.
-CHNG-CLI, CHNG-COVID, Google-AA, and especially CTIS-CLI-in-community have
-large right tails (displaying improvements over AR) during the down periods. 
-Google-AA and CTIS-CLI-in-community have large right tails during the flat
-periods. CHNG-CLI and DV-CLI have large left tails (poor forecasts relative to
-AR) in flat and up periods.  Google-AA is the only model that outperforms
-the AR model, on average, over up periods.  Overall, the indicators seem to help
-more during flat or down periods than up periods, with the exception of
-Google-AA.
+While the performance of all models, including AR, generally degrades in up
+periods, different models exhibit different and interesting patterns. CHNG-CLI,
+CHNG-COVID, Google-AA, and especially CTIS-CLI-in-community have large right
+tails (showing improvements over AR) during the down periods.  Google-AA and
+CTIS-CLI-in-community have large right tails during the flat periods. CHNG-CLI
+and DV-CLI have large left tails (poor forecasts relative to AR) in flat and up
+periods.  Google-AA is the only model that outperforms the AR model, on average,
+in up periods.  Overall, the indicators seem to help more during flat or down
+periods than during up periods, with the exception of Google-AA.
 
@@
 cross correlations to case rates over a short time window around $t$.
-The higher this score, the greater it ``leads'' case activity.
-%
-This analysis is closely related to Granger Causality \cite{engle1987a} and
-draws on a large body of prior work that measures leadingness in economic time-series 
-\cite[e.g.,][]{yeats1972a,emerson1996a,koch1988a,sargent1977a,engle1987a,shiskin1968a,granger1969a,lahiri1991a,koopmans1947a,StockWatson1989,diebold1989a,mitchell1938a,auerbach1982a,hamilton1996a}.
-%
-Figure 
-\ref{fig:leading} displays correlations between the leadingness score of an 
-indicator and the WIS difference (AR model minus an indicator model), stratified
-by whether the target is classified as up, down, or flat.  One would naturally
-expect that the WIS difference would be positively correlated with leadingness.
-Somewhat surprisingly, this relationship turns out to be strongest in down
-periods and weakest in up periods.  In fact, it is very nearly the case that for
-each indicator, the strength of correlations only decreases as we move from down
-to flat to up  periods.  In the supplement, we extend this analysis by studying 
+The higher this score, the greater it ``leads'' case activity. This analysis is
+closely related to Granger causality \cite{engle1987a} and draws on a large
+body of prior work that measures leadingness in economic time-series, e.g., 
+\cite{mitchell1938a, koopmans1947a, shiskin1968a, granger1969a, yeats1972a, 
+  sargent1977a, auerbach1982a, koch1988a, StockWatson1989, diebold1989a,
+  lahiri1991a, emerson1996a, hamilton1996a}. Figure \ref{fig:leading} displays
+correlations between the leadingness score of an indicator and the WIS
+difference (AR model minus an indicator model), stratified by whether the target
+is classified as up, down, or flat.  One would naturally expect that the WIS
+difference would be positively correlated with leadingness.  Somewhat
+surprisingly, this relationship turns out to be strongest in down periods and
+weakest in up periods.  In fact, it is very nearly the case that for each
+indicator, the strength of correlations only decreases as we move from down to
+flat to up periods.  In the supplement, we extend this analysis by studying
 analogous ``laggingness'' scores, but we do not find as clear patterns.
@@
 prediction problem easy when it was once hard (in the absence of auxiliary
-information).
-% This seems like a place to reference Superforecasters (or related) if we're
-% going to do it. ``The expectation should never be
-% that there is a silver bullet. Too much solid work refutes that. Again lack of
-% prior art citations to some very solid prior empirical work, and repetition of
-% a long discredited Hedgehog/Silver bullet strawman.''  
-Rather, the gains in accuracy from the indicator models (over an
+information). Rather, the gains in accuracy from the indicator models (over an 
 autoregressive model based only on past case rates) appear to be nontrivial, and
@@
 trending down, rather than periods in which case rates are trending up (as one
-might hope to see is the benefit provided by a hypothetical ``leading
+might hope to see is the benefit being provided by a hypothetical ``leading 
 indicator'').
@@
 sources that are relevant to public health surveillance are often subject to
-revision, sometimes regularly (such as medical claims data) and sometimes
+revision, sometimes regularly (as in medical claims data) and sometimes
 unpredictably (such as COVID-19 case reports).  When analyzing models that are
@@
 greater or lesser extent when they are forced to work with vintage data, so 
-backtesting them with finalized data may lead us to make modeling decisions that
-are suboptimal for true test-time performance.
+backtesting with finalized data may lead us to make modeling decisions that are
+suboptimal for true test-time performance.
 
@@
 while devoting most of our effort to accounting for as much of the complexity of
-the underlying data and evaluation as possible.
-% The below was held out as particularly anathema to the PNAS ethos.
-%
-% In fact, our paper is as much
-% about demonstrating how one might address questions about model comparisons and
-% evaluation in forecasting and hotspot prediction in general, as it is about
-% providing rigorous answers to such questions in the context of COVID-19 case
-% rates in particular.  We hope that others will leverage our framework, and build
-% on it, so that it can be used to guide work that advances the frontier of
-% predictive modeling for epidemics and pandemics.
-%
-% Added by DJM
-Despite their simplicity, these models are competitive with top forecasting
-models submitted to the COVID-19 Forecast Hub in real time, and they are
-demonstrably robust to the careful scrutiny undertaken here.
+the underlying data and evaluation as possible. In fact, our paper is not only
+about providing rigorous answers to questions about model comparisons in
+COVID-19 forecasting and hotspot prediction, but also about demonstrating how 
+one might go about answering such questions in general. We hope that others will
+leverage our work, and build on it, to guide advances on the frontier of
+predictive modeling for epidemics and pandemics.
 

### 2021-10-30 indicators/paper/supplement.Rmd
Rebuild supplement

--- 2021-09-16 indicators/paper/supplement.Rmd
@@
 \includegraphics[width=\textwidth]{fig/severity-pyramid.pdf}
-\caption{The epidemiological ``severity pyramid'' represents the progression of
-cases from the public, through infection, through increasingly severe stages of
-disease. The annotations here represent the data sources collected by Delphi's
-COVIDcast Epidata API.}
+\caption{Epidemiological ``severity pyramid'', representing the progression of disease progression, from relevant public behaviors, through infection, towards increasingly severe stages of disease.
+The annotations here refer the data sources available in Delphi's Epidata API.}
+\label{fig:severity-pyramid}
 \end{figure}
+
+\clearpage
+
+```{r case-correlation-lagged-data}
+start_day <- "2020-04-15"
+end_day <- "2021-04-15"
+as_of_day <- "2021-06-01"
+
+df_cases <- covidcast_signal("jhu-csse", "confirmed_7dav_incidence_prop",
+                             start_day, end_day, as_of = as_of_day)
+
+names <- c("7 days", "14 days", "21 days")
+lags <- c(7, 14, 21)
+df_cor <- vector("list", length(lags))
+
+for (ii in seq_along(lags)) {
+  df_cor[[ii]] <- covidcast_cor(df_cases, df_cases, dt_y = lags[ii],
+                                by = "time_value")
+  df_cor[[ii]]$signal <- names[ii]
+}
+
+df <- do.call(rbind, df_cor)
+df$signal <- factor(df$signal, levels = names, labels = names)
+```
+
+```{r case-correlation-lagged-plot, fig.cap="Geo-wise correlations between cases and lagged cases 1, 2, or 3 weeks prior, for all counties in the U.S.
+Lagged cases are correlated with cases as one might expect, but note the precipitous drop in correlation in February 2021.
+This matches the correlation drop between other COVIDcast signals and cases during the same time period, supporting the hypothesis that the drop was due to decreased heterogeneity in case rates by county."}
+ggplot(df, aes(x = time_value, y = value)) +
+  geom_line(aes(color = signal)) +
+  ylim(0, NA) +
+  labs(x = "Date", y = "Correlation", color = "Lag") +
+  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
+               date_minor_breaks = "1 month") +
+  theme_bw() +
+  theme(legend.position = "bottom")
+```
 
@@
 
-```{r state-trend-grids-1, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States.
-Cases are reported cases per 100,000 population.
+```{r state-trend-grids-1, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for U.S.
+states and territories.
+Cases are displayed on the rate scale: counts per 100,000 people.
 Other signals are scaled to have the same global maximum across all counties and times, so they can be presented in the same range.
-(part 1 of 4)", fig.height=8}
+(Part 1 of 4.)", fig.height=8}
 g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 1)
@@
 
-```{r state-trend-grids-2, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. (part 2 of 4)", fig.height=8}
+```{r state-trend-grids-2, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for U.S. states and territories. (Part 2 of 4.)", fig.height=8}
 g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 2)
@@
 
-```{r state-trend-grids-3, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. (part 3 of 4)", fig.height=8}
+```{r state-trend-grids-3, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for U.S. states and territories. (Part 3 of 4.)", fig.height=8}
 g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 3)
@@
 
-```{r state-trend-grids-4, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. (part 4 of 4)", fig.height=8}
+```{r state-trend-grids-4, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for U.S. states and territories. (Part 4 of 4.)", fig.height=8}
 g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 4)
@@
 
-```{r county-trend-grids-1, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States.
-Cases are reported cases per 100,000 population.
+```{r county-trend-grids-1, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous U.S.
+counties.
+Cases are displayed on the rate scale: counts per 100,000 people.
 Other signals are scaled to have the same global maximum across all counties and times, so they can be presented in the same range.
-(part 1 of 4)", fig.height=8}
+(Part 1 of 4.)", fig.height=8}
 g <- get_trend_grid(
@@
 
-```{r county-trend-grids-2, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 2 of 4)", fig.height=8}
+```{r county-trend-grids-2, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous U.S. counties. (Part 2 of 4.)", fig.height=8}
 g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 2)
@@
 
-```{r county-trend-grids-3, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 3 of 4)", fig.height=8}
+```{r county-trend-grids-3, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous U.S. counties. (Part 3 of 4.)", fig.height=8}
 g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 3)
@@
 
-```{r county-trend-grids-4, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 4 of 4)", fig.height=8}
+```{r county-trend-grids-4, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous U.S. counties. (Part 4 of 4.)", fig.height=8}
 g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 4)
@@
 
-```{r hospitalization_time_trends_national, fig.cap="National trends, from August 2020 to August 2021, of HHS-reported confirmed COVID-19 hospital admissions, plus comparison signals in the COVIDcast API.
-(HHS data was not consistently reported before August 2020.) Here, the HHS data has been processed into a 7-day trailing average rate per 100,000 resident population using Census Bureau estimates for 2019.
-HSP-Hosp is the percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners, smoothed in time and adjusted for systematic day-of-week effects (HSP-Hosp is available from an earlier date when working at finer geographical resolutions).
-HSP-Hosp, CTIS-CLI-in-community, and CHNG-CLI have been scaled to have the same maximum value as the HHS data."}
+```{r hospitalization_time_trends_national, fig.cap="National trends, from August 2020 to August 2021, of HHS-reported confirmed COVID-19 hospital admissions, along with several signals from the COVIDcast API.
+(HHS data was not consistently reported before August 2020; furthermore, as with cases in the previous trend plots, the HHS data has been smoothed using a 7-day trailing average.) Hospitalizations are displayed on the rate scale: counts per 100,000 people.
+Other signals are scaled to have the same maximum value.
+HSP-Hosp is the percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners (smoothed in time and adjusted for systematic day-of-week effects)."}
 start_day = "2020-08-15"

### 2021-10-30 indicators/paper/supplement.tex
Rebuild supplement

@@
 \includegraphics[width=\textwidth]{fig/severity-pyramid.pdf}
-\caption{The epidemiological ``severity pyramid'' represents the progression of
-cases from the public, through infection, through increasingly severe stages of
-disease. The annotations here represent the data sources collected by Delphi's
-COVIDcast Epidata API.}
+\caption{Epidemiological ``severity pyramid'', representing the progression of disease progression, from relevant public behaviors, through infection, towards increasingly severe stages of disease.
+The annotations here refer the data sources available in Delphi's Epidata API.}
+\label{fig:severity-pyramid}
 \end{figure}
@@
 
-\caption{Correlation between confirmed cases and case reports 1, 2, or 3 weeks prior, for all counties in the United States with case reporting.
-Cases are correlated with future cases as one might expect, but note the precipitous drop in correlation in February 2021.
+\caption{Geo-wise correlations between cases and lagged cases 1, 2, or 3 weeks prior, for all counties in the U.S.
+Lagged cases are correlated with cases as one might expect, but note the precipitous drop in correlation in February 2021.
 This matches the correlation drop between other COVIDcast signals and cases during the same time period, supporting the hypothesis that the drop was due to decreased heterogeneity in case rates by county.}\label{fig:case-correlation-lagged-plot}
@@
 
-\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States.
-Cases are reported cases per 100,000 population.
+\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for U.S.
+states and territories.
+Cases are displayed on the rate scale: counts per 100,000 people.
 Other signals are scaled to have the same global maximum across all counties and times, so they can be presented in the same range.
-(part 1 of 4)}\label{fig:state-trend-grids-1}
+(Part 1 of 4.)}\label{fig:state-trend-grids-1}
 \end{figure}
@@
 
-\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. (part 2 of 4)}\label{fig:state-trend-grids-2}
+\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for U.S. states and territories. (Part 2 of 4.)}\label{fig:state-trend-grids-2}
 \end{figure}
@@
 
-\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. (part 3 of 4)}\label{fig:state-trend-grids-3}
+\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for U.S. states and territories. (Part 3 of 4.)}\label{fig:state-trend-grids-3}
 \end{figure}
@@
 
-\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. (part 4 of 4)}\label{fig:state-trend-grids-4}
+\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for U.S. states and territories. (Part 4 of 4.)}\label{fig:state-trend-grids-4}
 \end{figure}
@@
 
-\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States.
-Cases are reported cases per 100,000 population.
+\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous U.S.
+counties.
+Cases are displayed on the rate scale: counts per 100,000 people.
 Other signals are scaled to have the same global maximum across all counties and times, so they can be presented in the same range.
-(part 1 of 4)}\label{fig:county-trend-grids-1}
+(Part 1 of 4.)}\label{fig:county-trend-grids-1}
 \end{figure}
@@
 
-\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 2 of 4)}\label{fig:county-trend-grids-2}
+\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous U.S. counties. (Part 2 of 4.)}\label{fig:county-trend-grids-2}
 \end{figure}
@@
 
-\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 3 of 4)}\label{fig:county-trend-grids-3}
+\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous U.S. counties. (Part 3 of 4.)}\label{fig:county-trend-grids-3}
 \end{figure}
@@
 
-\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 4 of 4)}\label{fig:county-trend-grids-4}
+\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous U.S. counties. (Part 4 of 4.)}\label{fig:county-trend-grids-4}
 \end{figure}
@@
 
-\caption{National trends, from August 2020 to August 2021, of HHS-reported confirmed COVID-19 hospital admissions, plus comparison signals in the COVIDcast API.
-(HHS data was not consistently reported before August 2020.) Here, the HHS data has been processed into a 7-day trailing average rate per 100,000 resident population using Census Bureau estimates for 2019.
-HSP-Hosp is the percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners, smoothed in time and adjusted for systematic day-of-week effects (HSP-Hosp is available from an earlier date when working at finer geographical resolutions).
-HSP-Hosp, CTIS-CLI-in-community, and CHNG-CLI have been scaled to have the same maximum value as the HHS data.}\label{fig:hospitalization_time_trends_national}
+\caption{National trends, from August 2020 to August 2021, of HHS-reported confirmed COVID-19 hospital admissions, along with several signals from the COVIDcast API.
+(HHS data was not consistently reported before August 2020; furthermore, as with cases in the previous trend plots, the HHS data has been smoothed using a 7-day trailing average.) Hospitalizations are displayed on the rate scale: counts per 100,000 people.
+Other signals are scaled to have the same maximum value.
+HSP-Hosp is the percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners (smoothed in time and adjusted for systematic day-of-week effects).}\label{fig:hospitalization_time_trends_national}
 \end{figure}

### 2021-10-31 forecast/paper/paper.tex
New paragraph in disc

@@
 techniques would change the ``topline'' conclusion---that auxiliary indicators
-can provide nontrivial, consistent, but modest gains in forecasting and hotspot
-prediction. Whether a more sophisticated model would be able to leverage the
-indicators in such a way as to change some of the finer conclusions (e.g., by 
-offering clear improvements in periods in which cases are trending up) is less 
-clear to us.
+can provide clear but modest gains in forecasting and hotspot prediction.
+
+However, rigorously vetting the details for more sophisticated models, as well
+as the generalizability of our findings to different geographic resolutions,
+both remain important directions for future study.  For example, in the
+Supplementary Materials we show that for forecasting at the state level, the
+benefits provided by indicators models over AR are generally less clear
+(compared to those observed at the HRR level).  One hypothesis is that at the
+state level, where the signal-to-noise ratio (SNR) is higher, AR performs even
+better and provides a higher standard (when asking whether it can be improved
+upon using indicators).  At the HRR level, where the SNR is lower, including the
+indicators as additional linear features in the AR model likely serves as an
+important variance reduction (just like averaging independent terms would) which
+contributes to improved accuracy.  But at a higher SNR, this variance
+reduction becomes less important, and we may be compelled to use more
+sophisticated modeling techniques to extract comparable value from the
+indicators.
 

### 2021-10-31 cover/cover.Rmd
First attempt at cover fig

---
title: Playground for PNAS cover graphic 
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path = "fig/", cache = TRUE, 
                      autodep = TRUE, cache.comments = TRUE,
                      message = FALSE, warning = FALSE)

library(covidcast)
library(epitools)
library(ggplot2)
library(dplyr)
library(purrr)
```

```{r get-us, eval = FALSE}
sources <- c("jhu-csse", 
             "doctor-visits", 
             "chng", 
             "fb-survey", 
             "safegraph",
             "safegraph")

signals <- c("confirmed_7dav_incidence_prop",
             "smoothed_adj_cli",
             "smoothed_adj_outpatient_covid",
             "smoothed_whh_cmnty_cli",
             "bars_visit_prop",
             "restaurants_visit_prop")

start <- "2020-04-15"
end <- "2021-04-15"

Select <- function(x, ...) { 
  if (nrow(x) == 0) x
  else select(x, ...) 
}

us <- covidcast_signals(
  sources, signals,
  start, end, 
  geo_type = "nation",
) %>% map(function(df) df %>% 
            Select(data_source, signal, geo_value, time_value, value))
```

```{r get-states, eval = FALSE}
state <- covidcast_signals(
  sources, signals,
  start, end, 
  geo_type = "state"
) %>% map(function(df) df %>% 
            Select(data_source, signal, geo_value, time_value, value))
```

```{r get-counties, eval = FALSE}
top_counties <- county_census %>%
  filter(COUNTY != 0) %>% # these are states
  arrange(desc(POPESTIMATE2019)) %>%
  head(n = 50) %>%
  pull(FIPS)

# Work around for an API problem: specifying geo_values in the API call fails on
# the CHNG data. So we filter after the fact instead. DANGER: Very slow.
county_nochng <- covidcast_signals(
  sources[sources != "chng"], signals[sources != "chng"],
  start, end, 
  geo_type = "county", geo_values = top_counties
) %>% map(function(df) df %>% 
            Select(data_source, signal, geo_value, time_value, value))

county_chng <- covidcast_signals(
  sources[sources == "chng"], signals[sources == "chng"],
  start, end, 
  geo_type = "county", 
) %>% map(function(df) df %>% 
            Select(data_source, signal, geo_value, time_value, value) %>%
            filter(geo_value %in% top_counties))

county <- append(county_nochng, county_chng)
```

```{r save-all, eval = FALSE}
save(list = ls(), file = "all.rda")
```

```{r load-all}
load(file = "all.rda")
```

```{r build-sg}
# SafeGraph national signals don't exist for all time, so build them from states
y = state_census %>% 
  mutate(geo_value = tolower(ABBR)) %>% 
  rename(w = POPESTIMATE2019) %>%
  select(geo_value, w)

for (i in which(sources == "safegraph")) {  
  us[[i]] <- state[[i]] %>% 
    left_join(y, by = "geo_value") %>% 
    group_by(data_source, signal, time_value) %>%
    summarize(value = Sum(w * value) / Sum(w)) %>%
    mutate(geo_value = "us") %>%
    ungroup()
} 
```

```{r smooth-sg}
# SafeGraph signals at all levels aren't smoothed, so we use 7-day averaging
for (i in which(sources == "safegraph")) {  
  us[[i]] <- us[[i]] %>% 
    as.epi_tibble() %>% 
    group_by(geo_value) %>%
    epi_slide(~ Mean(.x$value), n = 7, new_col_name = "value") %>%
    ungroup()

  state[[i]] <- state[[i]] %>% 
    as.epi_tibble() %>% 
    group_by(geo_value) %>%
    epi_slide(~ Mean(.x$value), n = 7, new_col_name = "value") %>%
    ungroup()

  county[[i]] <- county[[i]] %>% 
    as.epi_tibble() %>% 
    group_by(geo_value) %>%
    epi_slide(~ Mean(.x$value), n = 7, new_col_name = "value") %>%
    ungroup() 
}
```

```{r aggregate-data}
all <- do.call(rbind, c(us, state, county))
```

```{r plotting-functions}
names <- c("Cases", 
           "DV-CLI", 
           "CHNG-COVID", 
           "CTIS-CLI-in-community",
           "SG-Bars", 
           "SG-Restaurants")

colors <- c("black", scales::hue_pal()(length(signals)-1))

get_trend_plot <- function(df) {
  case_min <- Min(df %>% filter(signal == signals[1]) %>% pull(value))
  case_max <- Max(df %>% filter(signal == signals[1]) %>% pull(value))

  df %>%
    group_by(signal, geo_value) %>%
    mutate(value = (value - Min(value)) / (Max(value) - Min(value)) * 
             case_max + case_min) %>%
    ungroup() %>%
    mutate(signal = factor(signal, levels = signals, labels = names)) %>%
    ggplot(aes(x = time_value, y = value, color = signal)) +
    geom_line() + scale_color_manual(breaks = names, values = colors) +
    facet_wrap(vars(geo_value), scales = "free_y") + 
    labs(x = "Date", y = "Signal value (scaled)") +
    theme_bw() + theme(legend.pos = "bottom", legend.title = element_blank())
}
```

```{r plot-it, fig.width = 10, fig.height = 4}
all %>%
  filter(signal != signals[5]) %>%
  filter(geo_value %in% c("us", "tx", "48113")) %>%
  filter(between(time_value, as.Date("2020-06-01"), as.Date("2020-07-15"))) %>%
  get_trend_plot()
```

### 2021-10-31 forecast/paper/paper.tex
Start supplement copyedits

@@
 upon using indicators).  At the HRR level, where the SNR is lower, including the
-indicators as additional linear features in the AR model likely serves as an
-important variance reduction (just like averaging independent terms would) which
+indicators as additional linear features in the AR model probably delivers a
+kind of variance reduction (just like averaging independent terms would) which 
 contributes to improved accuracy.  But at a higher SNR, this variance

### 2021-10-31 forecast/paper/supplement-text.tex
Start supplement copyedits

@@
 they are, in principle, measuring the same thing (namely, the percentage of
-outpatient visits that are primarily about COVID-related symptoms).  The
-substantial difference in their curves in Figure 3 of the main paper must
-therefore reflect their having very different backfill profiles.
+outpatient visits that are primarily about COVID-related symptoms), but based on
+data from different providers.  The substantial difference in their curves in
+Figure 3 of the main paper must therefore reflect their having very different
+backfill profiles.
 
@@
 This is a reminder that the forecasting and hotspot problems are fundamentally
-distinct.  For example, the hotspot problem does not measure the ability to
-distinguish between flat and downward trends.
+different problems.  For example, the hotspot problem does not measure the
+ability to distinguish between flat and downward trends.
 
@@
 The forecasters based on indicators are thus affected both by revisions to the
-indicators and by revisions to the case rates.  In the case of the \gs~model, in
-which we only used finalized values for the \gs~indicator, the difference in
-performance can be wholly attributed to revisions of case rates.
+indicators and by revisions to the case rates.  And in the case of the
+\gs~model, in which we only used finalized values for the \gs~indicator, the
+difference in performance can be wholly attributed to revisions of case rates.
 

### 2021-10-31 forecast/paper/supplement.Rmd
Start supplement copyedits

@@
   ylab("Mean WIS (vintage / finalized)")
-
 ```
@@
   theme(legend.position = "bottom")
-
 ```

### 2021-10-31 forecast/paper/paper.tex
Rewordings from Jacob and others

@@
 Supplementary Materials we show that for forecasting at the state level, the
-benefits provided by indicators models over AR are generally less clear
-(compared to those observed at the HRR level).  One hypothesis is that at the
-state level, where the signal-to-noise ratio (SNR) is higher, AR performs even
-better and provides a higher standard (when asking whether it can be improved
+benefits of adding indicators to the AR model are generally less clear
+(compared to those observed at the HRR level).  A plausible explanation: at the
+state level, where the signal-to-noise ratio (SNR) is higher, AR performs better
+overall and provides a higher standard (when asking whether it can be improved 
 upon using indicators).  At the HRR level, where the SNR is lower, including the
@@
 will our sense of accuracy be unrealistic, but certain models may degrade by a
-greater or lesser extent when they are forced to work with vintage data, so 
-backtesting with finalized data may lead us to make modeling decisions that are
+greater or lesser extent when they are forced to reckon with vintage data, so  
+backtesting on finalized data may lead us to make modeling decisions that are
 suboptimal for true test-time performance.

### 2021-10-31 forecast/paper/supplement-text.tex
Intermediate progress on supp text

@@
 
-\section{Aggregation with Geometric Mean}
+\section{Aggregation With Geometric Mean}
 
-In this section, we consider using the geometric mean instead of the arithmetic
-mean when aggregating the weighted interval score (WIS) across location-time
-pairs.  There are three reasons why using the geometric mean may be desirable.
+In this section, we consider using the geometric mean instead of the usual
+(arithmetic) mean when aggregating the weighted interval score (WIS) across
+location-time pairs.  There are two reasons why using the geometric mean may be
+desirable.
+
 \begin{enumerate}
@@
   densities appear roughly log-Gaussian.  The geometric mean is a natural choice   
-  in such a context since the relative ordering of forecasters is determined by
-  the arithmetic mean of the {\em logarithm} of their WIS values.
+  in such a context since it can be viewed as a measure of centrality on the log
+  scale (it is the exponential of the arithmetic mean of log values).
 \item In the main paper, we report the ratio of the mean WIS of a forecaster to
@@
   penalize a method less for doing poorly where the baseline forecaster also
-  does poorly.  Using instead the geometric mean makes the order of aggregation
-  and scaling immaterial since the ratio of geometric means is the same as the 
-  geometric mean of ratios.
-\item If one imagines that a forecaster's WIS is composed of multiplicative
-  space-time effects $S_{\ell,t}$ shared across all forecasters,
-  i.e., $\WIS(F_{\ell,t,f},Y_{\ell,t})=S_{\ell,t}E_{f,t}$ with $E_{f,t}$ a
-  forecaster-specific error, then taking the ratio of two forecasters' geometric
-  mean WIS values will effectively cancel these space-time effects. 
+  does poorly.\footnote{In a sense, this is implicitly estimating a
+    nonparametric space-time effect for forecaster error, and assuming that has
+    a shared, multiplicative contribution to forecaster errors.  That is, if one
+    imagines that a forecaster's WIS is composed of multiplicative space-time
+    effects $S_{\ell,t}$ shared across all forecasters,
+    $\WIS(F_{\ell,t,f},Y_{\ell,t}) = S_{\ell,t} E_{f,t}$ with $E_{f,t}$ a 
+    forecaster-specific error, then taking the ratio of individual WIS values
+    cancels these space-time effects.}  Using instead the geometric
+  mean makes the order of aggregation and scaling immaterial since the ratio of
+  geometric means is the same as the geometric mean of ratios.
 \end{enumerate}
@@
 Figure \ref{fig:fcast-adjusted} uses the geometric mean for aggregation.
-Comparing this with Figure 3 of the main paper, we see that the main conclusions
+Comparing this with Figure 3 in the main paper, we see that the main conclusions
 are largely unchanged; however, \chngcli~now appears better than \ar.  This
@@
 moderate errors).  Indeed, Figure 5 of the main paper further corroborates this,
-in which we see the heaviest left tails occuring for \chngcli.
+in which we see the heaviest left tails occur for \chngcli.
 
-\section{Comparing with the COVID-19 Forecast Hub}
+\section{Comparing With the COVID-19 Forecast Hub}
 
-Since July of 2020, teams have been submitting real time forecasts
-to the COVID-19 Forecast Hub \cite{ForecastHub}. While a number of forecast
-targets can be submitted, the most related to that used in this paper---the
-7-day trailing average of COVID-19 case incidence rates at the Hospital Referral
-Region---are state-level forecasts for total case incidence over an epiweek.
-Essentially, for each state, teams are asked to predict the total number of
-cases that will be recorded over the week from Monday to Sunday. Forecasts must
-be submitted by Tuesday for the current Sunday as well as the next three
-Sundays. This corresponds to producing forecasts for 5, 12, 19, and 26 days
-ahead. Some teams choose to submit their forecasts a few days earlier, but
-always for the same targets.
+Since July of 2020, modelers have been submitting real-time forecasts of
+COVID-19 case incidence to the COVID-19 Forecast Hub \cite{ForecastHub}. This 
+(along with forecasts of hospitalizations and deaths collected in the same Hub)
+serves as the source of the CDC's official communications on COVID forecasting.
+
+Our goal in this section is to compare the AR model and indicator models to
+those in the Hub, in terms of forecast errors aggregated over the same forecast
+tasks, to give a clear sense for how robust and effective the models we choose
+to investigate in the paper are relative to those in common operational use.
+This was prompted by a question from an anonymous reviewer of this paper, who
+asked why we chose to build our analysis of indicator utility around the AR
+model in the first place, and why we did not investigate others (say, the SIR
+model or more complex mechanistic models of disease transmission) that have
+occupied more of the spotlight over the course of the pandemic.  The analysis
+presented here corroborates the claim that an AR model is simple, and yet,
+properly trained---using a quantile loss to directly estimate multiple
+conditional quantiles, a trailing training window of 21 days, pooling across all
+locations jointly, and fitting to case rates rather than counts---all as we do
+in the models in the main paper, it can be robust and effective, performing 
+competitively with the top models from COVID-19 Forecast Hub.
+
+The closest forecast target in the Hub to that used in this paper are
+state-level case incidence over an epiweek---defined by the sum of new case
+counts reported between a Sunday and the following Saturday (inclusive). Our
+target is 7-day trailing average of COVID-19 case incidence rates at the HRR
+level, which is different in three regards:
+
+\begin{enumerate}
+\item temporal resolution (daily versus weekly);
+\item geographic resolution (HRRs versus states); and 
+\item scale (rates versus counts).
+\end{enumerate}
+
+While the first and third of these differences can be easily addressed post 
+hoc---meaning, we can take always take our model's outputs and multiply them by
+7 in order to track the incidence over any given trailing week, and rescale them
+per-location by population to put them on the count scale---the second
+difference is not easy to adjust post hoc, because a quantile is not linear
+operator (a quantile of a linear combination of random variables is not simply
+the linear combination of their quantiles, but rather, depends intricately on
+the correlations between random variables).
+
+Therefore, to make a comparison to models in the Hub in as direct a way as
+possible, we retrained our models over the same forecast period as in the main
+paper, and with the same general setup entirely, except at the state rather than 
+HRR level.  We then rescaled them post hoc to account for the different temporal
+resolution and the rate versus count scale (first and third items on the above
+list).  The results are given in Figure \ref{fig:compare-to-hub}.  The
+evaluation was carried out exactly as in the main paper, and the figure displays
+both mean WIS and geometric mean WIS, as a function of ahead.  Furthermore, to
+account for missingness (not all teams submitted forecasts to the Hub for all
+locations and ahead values for the entire period), we first dropped any
+forecaster than submitted for less than 6 weeks, and then restricted the
+aggregation metrics (mean or geometric mean) to errors from commonly available
+forecast tasks.
+
+By either metric, mean and geometric mean WIS, relative to baseline, we can see
+in Figure \ref{fig:compare-to-hub} that the AR model examined in this paper is 
+competitive with top models in the Hub, even outperforming the Forecast Hub
+ensemble model for smaller ahead values.  The same general conclusion can be
+drawn for the indicators models as well.  However, interestingly, a close
+inspection reveals that the AR model here is for the most part in the ``middle
+of the pack'' when compared to the indicator models, and only the Google-AA
+model offers clear improvement over AR for all aheads.  This is likely due to
+the fact that at the state level, the signal-to-noise ratio (SNR) is generally
+higher, and AR model provides a higher standard (on which to expect improvement
+using an auxiliary indicator).  At the HRR level, with lower SNR, using
+indicators as simple additional linear features in the AR model provides an
+important enough variance reduction
+
+% Essentially, for each state, teams are asked to predict the total number of
+% cases that will be recorded over the week from Monday to Sunday. Forecasts must
+% be submitted by Tuesday for the current Sunday as well as the next three
+% Sundays. This corresponds to producing forecasts for 5, 12, 19, and 26 days
+% ahead. Some teams choose to submit their forecasts a few days earlier, but
+% always for the same targets.
 
@@
 Since not all teams submitted forecasts for the entire period, we
-only display those which submitted for at least 6 weeks. By both metrics,
-the AR model examined in this paper is competitive with the submitted by top teams, even
-beating the Ensemble for smaller target horizons.
+only display those which submitted for at least 6 weeks.
 

### 2021-10-31 forecast/paper/supplement-text.tex
Another intermediate commit on supp

@@
 
-\section{Aggregation With Geometric Mean}
+\section{Robust Aggregation}
 
@@
 (arithmetic) mean when aggregating the weighted interval score (WIS) across
-location-time pairs.  There are two reasons why using the geometric mean may be
-desirable.
+location-time pairs.  Aside from the geometric mean being generally more robust
+to large values, there are two reasons why using it may be desirable.
 
@@
     effects $S_{\ell,t}$ shared across all forecasters,
-    $\WIS(F_{\ell,t,f},Y_{\ell,t}) = S_{\ell,t} E_{f,t}$ with $E_{f,t}$ a 
-    forecaster-specific error, then taking the ratio of individual WIS values
-    cancels these space-time effects.}  Using instead the geometric
-  mean makes the order of aggregation and scaling immaterial since the ratio of
-  geometric means is the same as the geometric mean of ratios.
+    \smash{$\WIS(F_{\ell,t,f},Y_{\ell,t}) = S_{\ell,t} E_{f,t}$} with
+    \smash{$E_{f,t}$} a forecaster-specific error, then taking the ratio of
+    individual WIS values cancels these space-time effects.}  Using instead the
+  geometric mean makes the order of aggregation and scaling immaterial since the
+  ratio of geometric means is the same as the geometric mean of ratios.
 \end{enumerate}
@@
 
-\section{Comparing With the COVID-19 Forecast Hub}
+\section{Comparing COVID-19 Forecast Hub Models}
 
@@
 asked why we chose to build our analysis of indicator utility around the AR
-model in the first place, and why we did not investigate others (say, the SIR
-model or more complex mechanistic models of disease transmission) that have
+model in the first place, and why we did not build it around others (say, the
+SIR model or more complex mechanistic models of disease transmission) that have 
 occupied more of the spotlight over the course of the pandemic.  The analysis
-presented here corroborates the claim that an AR model is simple, and yet,
-properly trained---using a quantile loss to directly estimate multiple
-conditional quantiles, a trailing training window of 21 days, pooling across all
-locations jointly, and fitting to case rates rather than counts---all as we do
-in the models in the main paper, it can be robust and effective, performing 
-competitively with the top models from COVID-19 Forecast Hub.
+presented here corroborates the claim that, while simple, the AR model, properly
+trained---using a quantile loss to directly estimate multiple conditional
+quantiles, a trailing training window of 21 days, pooling across all locations
+jointly, and fitting to case rates rather than counts (as we do in all our
+models in the main paper)---can be robust and effective, performing
+competitively many of the top models from COVID-19 Forecast Hub, including the
+Hub ensemble model.
 
-The closest forecast target in the Hub to that used in this paper are
+The closest forecast target in the Hub to that used in the main paper is 
 state-level case incidence over an epiweek---defined by the sum of new case
 counts reported between a Sunday and the following Saturday (inclusive). Our
-target is 7-day trailing average of COVID-19 case incidence rates at the HRR
-level, which is different in three regards:
+forecast target, recall, is a 7-day trailing average of COVID-19 case incidence
+rates at the HRR level, which is different in three regards:
 
@@
 \item temporal resolution (daily versus weekly);
-\item geographic resolution (HRRs versus states); and 
+\item geographic resolution (HRRs versus states); 
 \item scale (rates versus counts).
@@
 
-While the first and third of these differences can be easily addressed post 
-hoc---meaning, we can take always take our model's outputs and multiply them by
-7 in order to track the incidence over any given trailing week, and rescale them
-per-location by population to put them on the count scale---the second
-difference is not easy to adjust post hoc, because a quantile is not linear
-operator (a quantile of a linear combination of random variables is not simply
-the linear combination of their quantiles, but rather, depends intricately on
-the correlations between random variables).
+While the first and third of these differences could be easily addressed post 
+hoc---meaning, we can take always take our model's output and multiply it by 
+7 in order to track the incidence over any given trailing week, and rescale it
+per location by population to bring it to the count scale---the second
+difference is not easy to adjust post hoc due to nonlinearity of the quantiles
+(a quantile of a linear combination of random variables is not simply the linear
+combination of their quantiles, but rather, depends intricately on the
+correlations between the random variables).
 
-Therefore, to make a comparison to models in the Hub in as direct a way as
-possible, we retrained our models over the same forecast period as in the main
-paper, and with the same general setup entirely, except at the state rather than 
-HRR level.  We then rescaled them post hoc to account for the different temporal
-resolution and the rate versus count scale (first and third items on the above
-list).  The results are given in Figure \ref{fig:compare-to-hub}.  The
-evaluation was carried out exactly as in the main paper, and the figure displays
-both mean WIS and geometric mean WIS, as a function of ahead.  Furthermore, to
-account for missingness (not all teams submitted forecasts to the Hub for all
-locations and ahead values for the entire period), we first dropped any
-forecaster than submitted for less than 6 weeks, and then restricted the
-aggregation metrics (mean or geometric mean) to errors from commonly available
-forecast tasks.
+Therefore, to make a comparison to models in the Hub as direct as possible, we
+retrained our models over the same forecast period as in the main paper, and
+with the same general setup entirely, except at the state rather than HRR level.
+We then rescaled them post hoc to account for the different temporal resolution
+and the rate versus count scale (first and third points in the above list).  The
+results are given in Figure \ref{fig:compare-to-hub}.  The evaluation was
+carried out exactly as in the main paper, and the figure displays both mean WIS
+and geometric mean WIS, as a function of ahead, relative to the baseline model.
+Furthermore, to account for missingness (not all teams submitted forecasts to
+the Hub for all locations and ahead values for the entire period), we first
+dropped any forecaster than submitted for less than 6 weeks, and then restricted
+the aggregation metrics (mean or geometric mean) to errors from commonly
+available forecast tasks.
 
-By either metric, mean and geometric mean WIS, relative to baseline, we can see
+By either metric, mean or geometric mean WIS relative to baseline, we can see  
 in Figure \ref{fig:compare-to-hub} that the AR model examined in this paper is 
-competitive with top models in the Hub, even outperforming the Forecast Hub
-ensemble model for smaller ahead values.  The same general conclusion can be
-drawn for the indicators models as well.  However, interestingly, a close
-inspection reveals that the AR model here is for the most part in the ``middle
-of the pack'' when compared to the indicator models, and only the Google-AA
-model offers clear improvement over AR for all aheads.  This is likely due to
-the fact that at the state level, the signal-to-noise ratio (SNR) is generally
-higher, and AR model provides a higher standard (on which to expect improvement
-using an auxiliary indicator).  At the HRR level, with lower SNR, using
-indicators as simple additional linear features in the AR model provides an
-important enough variance reduction
-
-% Essentially, for each state, teams are asked to predict the total number of
-% cases that will be recorded over the week from Monday to Sunday. Forecasts must
-% be submitted by Tuesday for the current Sunday as well as the next three
-% Sundays. This corresponds to producing forecasts for 5, 12, 19, and 26 days
-% ahead. Some teams choose to submit their forecasts a few days earlier, but
-% always for the same targets.
-
-To properly compare the models presented in this paper with those submitted to
-the Hub, we limit ourselves to the same set of forecast dates and states, and
-we predict case incidence over a week by: (1) predicting the 7-day trailing
-average case rate (as in the manuscript);
-(2) multiplying by 7 to give the total over a week; (3) multiplying by the state
-population as reported in the 2019 US Census Population Estimate (available in
-the Covidcast R package \cite{CovidcastR}). Finally, we perform the same
-evaluations as in the manuscript Figure 3 and also using geometric mean as
-described above.
-For the purpose of the comparison, we show only the AR model along with
-submissions collected by the Forecast Hub in Figure \ref{fig:compare-to-hub}.
-Since not all teams submitted forecasts for the entire period, we
-only display those which submitted for at least 6 weeks.
-
-We also are undertook the same analysis shown in Figure 3 in the manuscript and Figure
-\ref{fig:fcast-adjusted} above for all models
-(the AR, as well as the indicator-assisted models). This comparison is shown in Figure
-\ref{fig:state-level-performance}. 
-%
-TODO: some text about why indicators don't help as much.
+competitive with top models in the Hub, even outperforming the Hub ensemble
+model for smaller ahead values.  The same general conclusion can be drawn for
+the indicators models as well.  However, interestingly, a close inspection
+reveals that the AR model here is for the most part in the ``middle of the
+pack'' when compared to the indicator models, and only the Google-AA model
+offers clear improvement over AR for all aheads.  This is likely due to the fact
+that at the state level, the signal-to-noise ratio (SNR) is generally higher,
+and AR model provides a higher standard (on which to expect improvement using an 
+auxiliary indicator), since it is able to extract a clearer signal from past
+lags of case rates.  At the HRR level, with lower SNR, using indicators as
+simple additional linear features in the AR model probably leads to a variance
+reduction that is enough to boost accuracy, but at the state level, perhaps more
+sophisticated modeling techniques are needed to extract comparable value from
+some of the indicators.
 
@@
 
-In the Introduction of the main paper, we give some reasons that we avoid making
-formal statements about statistical significance, preferring instead to examine
-the stability of our results in different contexts. There are strong reasons to
-avoid model-based significance tests because the necessary assumptions about
-stationarity, independence, and the  model being true (or at least approximately
-true) are certainly violated. With those caveats in mind, we undertake two
-relatively assumption-lean investigations in this section. The first is a
-sign-test for whether the difference between the AR model’s relative WIS  and
-each other model’s relative WIS is centered at zero. By ``relative WIS'' we mean
-scaled by the strawman as displayed in Figure 3. To mitigate the dependence
+In the introduction of the main manuscript, we gave some reasons that we avoid
+making formal statements about statistical significance, preferring instead to
+examine the stability of our results in different contexts.  There are strong
+reasons to avoid model-based significance tests because the necessary
+assumptions about stationarity, independence, and the model being true (or at
+least approximately true) are certainly violated.  With those caveats in mind,
+we undertake two relatively assumption-lean investigations in this section.  The
+first is a sign test for whether the difference between the AR model’s relative
+WIS and each other model’s relative WIS is centered at zero.  (Relative WIS here
+means scaled by the WIS of the baseline model.)  To mitigate the dependence 
 across time (which intuitively seems to matter more than that across space), we
 computed these tests in a stratified way, where for each forecast date we run a
-sign test on the scaled errors between two models over all 440 counties. The
-results are plotted as histograms in Figure \ref{fig:sign-test}. In this case,
-we use the total relative WIS over all aheads, but the histograms are largely
-similar for individual target horizons. If there were no difference, we would
-expect to see a uniform distribution. However, for all indicator-assisted
-models, we see many more small p-values than would be expected if the null
-hypothesis (that the AR model is better) were true.
+sign test on the scaled errors between two models over all 306 HRRs.  The
+results are plotted as histograms in Figure \ref{fig:sign-test}.  For this
+figure, we use the total relative WIS over all aheads, but the histograms are
+largely similar for individual target horizons. If there were no difference, we
+would expect to see a uniform distribution. However, for each indicator model, 
+we see many more small p-values than would be expected if the null hypothesis
+(that the AR model is better) were true.
 
 Another relatively assumption-lean method of testing for differences in forecast
-accuracy is the Diebold-Mariano test \cite{Diebold:2002, Diebold:2015,
-Harvey:1997}. Essentially, the differences between forecast losses are assumed
+accuracy is the Diebold-Mariano (DM) test \cite{Diebold:2002, Diebold:2015,
+Harvey:1997}. Essentially, the differences between forecast errors are assumed
 to have a constant mean and a covariance that depends on time.  Under these
@@
 differences is limiting normal provided that a heteroskedasticity and
-autocorrelation robust estimate of the variance is used. Using the loss as
-weighted interval score across all HRRs and horizons (7 to 21 days ahead), we
-perform the DM test using both the mean relative to the strawman (as reported in
-the manuscript) and the geometric mean relative to the strawman as described
-above. The first two rows of Table \ref{tab:dm-test} displays p-values for the
-test that the indicator-assisted model is no-better than the AR model. Only the
-CHNG-CLI model's p-value exceeds conventional statistical significance
-thresholds.
+autocorrelation robust estimate of the variance is used.  Using the error as WIS
+across all HRRs and horizons (7 to 21 days ahead), we perform the DM test using
+both the mean relative to the baseline (as reported in the manuscript) and the
+geometric mean relative to the baseline as described above.  The first two rows
+of Table \ref{tab:dm-test} displays p-values for the test that each indicator
+model is no better than the AR model.  In only a few instances---geometric mean
+and mean for the CHNG-CLI model, and mean for the CHNG-COVID model---do the
+p-values exceed conventional statistical significance thresholds.
 
@@
 
-As explained in Section 2.B. of the main paper, a (somewhat cynical) hypothesis
+As explained in Section 2.B of the main paper, a (somewhat cynical) hypothesis
 for why we see benefits in forecasting and hotspot prediction is that the
@@
 investigate this hypothesis, we consider fitting ``noise features'' that in
-truth should have zero coefficients.  Recall (from the main paper) that at each
-forecast date, we train a model on 6,426 location-time pairs.  Indicator models
-are based on six features, corresponding to the three autoregressive terms and
-the three lagged indicator values.  To form noise indicator features, we replace
-their values with those from a randomly chosen time-space pair (while keeping
-the autoregressive features fixed).  In particular, at each location $\ell$ and
-time $t$, for the forecasting task we replace the triplet $(X_{\ell,t},
-X_{\ell,t-7}, X_{\ell,t-14})$ in Eq. (3) of the main paper with the triplet
-$(X_{\ell^*,t^*}, X_{\ell^*,t^*-7}, X_{\ell^*,t^*-14})$, where $(\ell^*,t^*)$ is
-a location-time pair sampled with replacement from the 6,426 location-time
-pairs.  Likewise in the hotspot prediction task, we replace the triplet
-$(X_{\ell,t}^\Delta, X_{\ell,t-7}^\Delta, X_{\ell,t-14}^\Delta)$ in Eq. (5) of
-the main paper with $(X_{\ell^*,t^*}^\Delta, X_{\ell^*,t^*-7}^\Delta,
-X_{\ell^*,t^*-14}^\Delta)$.  Figures
+truth have zero relationship to the response.  Recall (from the main paper) that
+at each forecast date, we train a model on 6,426 location-time pairs.  Indicator
+models are based on six features, corresponding to the three autoregressive
+terms and the three lagged indicator values.  To form noise indicator features,
+we replace their values with those from a randomly chosen time-space pair (while
+keeping the autoregressive features fixed).  In particular, at each location
+$\ell$ and time $t$, for the forecasting task we replace the triplet
+\smash{$(X_{\ell,t}, X_{\ell,t-7}, X_{\ell,t-14})$} in Eq.\ (3) of the main
+paper with the triplet \smash{$(X_{\ell^*,t^*}, X_{\ell^*,t^*-7},
+  X_{\ell^*,t^*-14})$}, where $(\ell^*,t^*)$ is a location-time pair sampled with
+replacement from the 6,426 total location-time pairs.  Likewise in the hotspot 
+prediction task, we replace the triplet \smash{$(X_{\ell,t}^\Delta,
+  X_{\ell,t-7}^\Delta, X_{\ell,t-14}^\Delta)$} in Eq.\ (5) of the main paper with
+\smash{$(X_{\ell^*,t^*}^\Delta, X_{\ell^*,t^*-7}^\Delta,
+  X_{\ell^*,t^*-14}^\Delta)$}.  Figures
 \ref{fig:fcast-booted}--\ref{fig:hot-booted} show the results.  No method
@@
 
-In this section we provide extra details about the upswing / flat / downswing
-analysis described in the main text. Figure \ref{fig:upswing-summary} shows the
-overall results, examining the average difference $\WIS(AR) - \WIS(F)$ in
-period. Figure \ref{fig:hotspots-upswing-downswing} shows the same information
-for the hotspot task. On average, during downswings and flat periods, the
-indicator-assisted models have lower classification error and higher log
-likelihood than the AR model. For hotspots, both Google-AA and CTIS-CLIIC
-perform better than the AR model during upswings, in contrast to the forecasting
-task, where only Google-AA improves. For a related analysis, Figure
-\ref{fig:cor-wis-ratio} shows histograms of the Spearman correlation (Spearman's
-$\rho$, a rank-based measure of association) between the $\WIS(F)/\WIS(AR)$ and
-the magnitude of the swing. Again we see that case rate increases are positively
-related to diminished performance of the indicator models.
+In this section, we provide extra details about the up/down/flat analysis
+described in the main text. Figure \ref{fig:upswing-summary} shows the overall
+results, examining the average difference $\WIS(AR) - \WIS(F)$ in period. Figure
+\ref{fig:hotspots-upswing-downswing} shows the same information for the hotspot
+task. On average, during downswings and flat periods, the indicator models have
+lower classification error and higher log likelihood than the AR model. For
+hotspots, both Google-AA and CTIS-CLIIC perform better than the AR model during
+upswings, in contrast to the forecasting task, where only Google-AA
+improves. For a related analysis, Figure \ref{fig:cor-wis-ratio} shows
+histograms of the Spearman correlation (Spearman's $\rho$, a rank-based measure
+of association) between the $\WIS(F)/\WIS(AR)$ and the magnitude of the
+swing. Again we see that case rate increases are positively related to
+diminished performance of the indicator models.
 
@@
 Figure \ref{fig:errs-in-space} examines the spatial behavior over the entire
-period of the indicator-assisted models relative to the AR. For ease of
+period of the indicator models relative to the AR. For ease of
 comparison, we show the percent improvement in each HRR. Negative numbers (blue)

### 2021-10-31 forecast/paper/supplement.Rmd
Another intermediate commit on supp

--- 2021-10-29 forecast/paper/supplement.Rmd
@@
 fig4 <- plotter_hotspots(hotspots_honest) + 
-  geom_hline(yintercept = 0.5)
+  geom_hline(yintercept = 0.5, linetype = "dashed")
 fig4
@@
 plotter_hotspots(hotspots_finalized) +
-  geom_hline(yintercept = 0.5)
+  geom_hline(yintercept = 0.5, linetype = "dashed")
 ```
@@
 
-```{r fcast-honest-v-finalized, fig.cap="Relative forecast WIS with vintage compared to finalized data. Using finalized data leads to overly optimistic performance."}
+```{r fcast-honest-v-finalized, fig.cap="Forecast performance with vintage compared to finalized data. Using finalized data leads to overly optimistic performance."}
 plotter(
@@
   order_of_operations = c("aggr","scale")) +
-  geom_hline(yintercept = 1) +
+  geom_hline(yintercept = 1, linetype = "dashed") +
   ylab("Mean WIS (vintage / finalized)")
-
 ```
@@
 
-```{r hot-honest-v-finalized, fig.cap="Relative AUC with vintage compared to finalized data. Using finalized data leads to overly optimistic hotspot performance."}
+```{r hot-honest-v-finalized, fig.cap="Hotspot prediction performance with vintage compared to finalized data. Using finalized data leads to overly optimistic performance."}
 left_join(
@@
   theme(legend.position = "bottom", legend.title = element_blank()) +
-  geom_hline(yintercept = 1) +
+  geom_hline(yintercept = 1, linetype = "dashed") +
   ylab("AUC (vintage / finalized)")
@@
 
-```{r wis-densities, fig.cap="Weighted interval score appears to more closely resemble a log-Gaussian distribution."}
+```{r wis-densities, fig.cap="WIS values from forecast models, which appear to be roughly log-Gaussian."}
 ggplot(fcasts_honest, 
@@
   theme_bw() +
-  xlab("WIS") +
+  xlab("WIS") + ylab("Density") +
   scale_fill_manual(values = fcast_colors) +
@@
 
-```{r fcast-adjusted, fig.cap="Relative forecast performance using vintage data and summarizing with the more robust geometric mean."}
+```{r fcast-adjusted, fig.cap="Forecast performance (using vintage data), summarized by geometric mean."}
 plotter(fcasts_honest, 
@@
         order_of_operations = c("scale","aggr")) +
-  ylab("Geometric mean of WIS (relative to baseline)")
+  ylab("Geometric mean WIS (relative to baseline)")
 ```
@@
 
-```{r compare-to-hub, fig.cap="Performance of our models re-estimated for state-level, epiweek case incidence.
-Our models are colored the same as before.
-The solid light-blue line is the COVIDhub Ensemble forecast.
-The thin grey lines are the forecasts of other teams."}
-
+```{r compare-to-hub, fig.cap="Forecast performance for AR and indicators models, each retrained at the state level, compared to models submitted to the COVID-19 Forecast Hub over the same period.
+The thin grey lines individual models from the Hub; the blue line is the Hub ensemble model.
+(Note that, to align prediction dates as best as possible, we look at the AR and indicator model forecasts for 5, 12, 19, and 26 days ahead; this roughly corresponds to 1, 2, 3, and 4 weeks ahead, respectively, since in the Hub, models typically submit forecasts on a Tuesday for the epiweeks aligned to end on each of the following 4 Saturdays.)"}
 knitr::include_graphics("fig/compare-states-to-hub.pdf")
@@
 
-```{r sign-test, fig.cap="P-values for a sign test that the WIS of the AR forecaster is smaller than that of the indicator-assisted model. Each P-value corresponds a particular forecast date."}
+```{r sign-test, fig.cap="P-values from a one-sided sign test for equality of forecast error distributions.
+Each P-value corresponds a forecast date.
+The alternative hypothesis is that the the AR model is better (median difference between the relative WIS of the AR model and an indicator model is negative)."}
 fcast_colors2 <- fcast_colors[names(fcast_colors) != "AR"]
@@
   theme_bw() +
-  theme(legend.position = "bottom")
-
+  theme(legend.position = "bottom", legend.title = element_blank())
 ```
@@
   summarise(
-    `Mean Relative WIS` = mean(wis) / mean(strawman_wis),
-    # mwis = mean(wis), 
-    `Geometric Mean Relative WIS` = GeoMean(wis / strawman_wis) #,
-    # gmwis = GeoMean(wis)
+    `Mean` = mean(wis) / mean(strawman_wis),
+    `Geometric mean` = GeoMean(wis) / GeoMean(strawman_wis),
     ) %>%
@@
   filter(forecaster != "AR") %>% 
-  pivot_longer(-c("forecaster", "forecast_date"), names_to = "metric", values_to = "err")
+  pivot_longer(-c("forecaster", "forecast_date"), 
+               names_to = "Aggregation metric",
+               values_to = "err")
 
@@
     pivot_wider(names_from = forecaster, values_from = dm), 
-  digits = 3, booktabs = TRUE, caption = "$P$-values for a one-sided Diebold-Mariano test for improvement in forecast error.
-The test is for the null hypothesis of equal performance against the alternative that the indicator assisted model is better.")
+  digits = 3, booktabs = TRUE, caption = "P-values from a one-sided Diebold-Mariano test for equality of forecast error distributions. The alternative hypothesis is that the AR model is better.")
 ```
@@
 plotter_hotspots(hotspots_booted) + 
-  geom_hline(yintercept = 0.5)
+  geom_hline(yintercept = 0.5, linetype = "dashed")
 ```
@@
 
-```{r upswing-summary, fig.cap="Average difference between the WIS of the AR model and the WIS of the other forecasters. The indicator-assisted forecasters do best during down and flat periods."}
+```{r upswing-summary, fig.cap="Average difference between the WIS of the AR model and of the indicator models. The indicator models generally do best during down and flat periods."}
 up_down_df_summary <- up_down_df %>%
@@
            stat = "identity") +
-  scale_fill_manual(values = fcast_colors[1:5], guide = guide_legend(nrow = 1)) +
+  scale_fill_manual(values = fcast_colors2, guide = guide_legend(nrow = 1)) +
   scale_y_continuous() +
   theme_bw() +
-  geom_hline(yintercept = 0) +
+  geom_hline(yintercept = 0, linetype = "dashed") +
   ylab("Mean of WIS(AR) - WIS(F)") +
@@
   geom_vline(xintercept = 0)
-```
-
-```{r upswing-histogram-logged, eval = FALSE, fig.cap="Not sure if we want this here. Similar to Figure 5 in the manuscript but taking logs. "}
-up_down_df %>% 
-  group_by(forecaster, udf) %>%
-  ggplot(aes((AR_wis + 1) / (wis + 1), fill = forecaster)) +
-  geom_histogram(bins = 100) +
-  facet_grid(udf ~ forecaster) +
-  theme_bw() +
-  ylab("Count") +
-  theme(legend.position = "none") +
-  scale_fill_manual(values = fcast_colors) +
-  scale_x_log10() +
-  scale_y_log10(breaks = c(10,1000,100000), 
-                labels = trans_format("log10", math_format(10^.x))) +
-  xlab("AR WIS - forecaster WIS") +
-  geom_vline(xintercept = 1)
 ```
@@
            stat = "identity") +
-  scale_fill_manual(values = fcast_colors[1:5], guide = guide_legend(nrow = 1)) +
+  scale_fill_manual(values = fcast_colors2, guide = guide_legend(nrow = 1)) +
   scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
   theme_bw() +
-  geom_hline(yintercept = 0) +
+  geom_hline(yintercept = 0, linetype = "dashed") +
   facet_wrap(~err) +
@@
 ```
-
-```{r hotspots-upswing-downswing-remake, fig.cap="REMAKE: Classification and loglikelihood separated into periods of upswing, downswing, and flat cases.
-The indicator assisted models always have lower classification error relative to the null classifier, while the AR actually does worse in down and flat periods.
-In terms of loglikelihood, all forecasters have lower likelihood than the null classifier during up periods.
-In down and flat periods, the indicators generally improve over the AR, while the are worse during up periods.", eval=FALSE}
-# mycummean <- function(x) {
-#   isgood <- ! is.na(x)
-#   denom <- cumsum(isgood)
-#   x[!isgood] <- 0
-#   num <- cumsum(x)
-#   y <- num / denom
-#   y[is.na(y)] <- .5 # deal with empties at the beginning
-#   y
-# }
-# 
-# null_classifier <- readRDS(here("data", 
-#                                 "all_signals_wide_as_of_2020-05-18.RDS")) %>%
-#   rename(jhu = `value+0:jhu-csse_confirmed_7dav_incidence_prop`) %>%
-#   select(geo_value, time_value, jhu) %>%
-#   group_by(geo_value) %>%
-#   arrange(time_value) %>%
-#   mutate(lag_value = lag(jhu, 7),
-#          apc = (jhu - lag_value) / lag_value,
-#          hot = case_when(
-#            apc > .25 & lag_value > 30 ~ 1,
-#            apc <= .25 & lag_value > 30 ~ 0),
-#          null_class = mycummean(hot))
-
-hot_udf <- inner_join(
-  hotspots_honest,
-  up_down %>% select(geo_value, target_end_date, udf))
-
-cutoff <- mean(hot_udf %>% filter(forecaster == "AR") %>% pull(actual))
-
-con_tab <- hot_udf %>% 
-  filter(!is.na(actual)) %>%
-  mutate(pred = value > .5) %>%
-  group_by(forecaster, udf) %>%
-  summarise(m = (mean(pred != actual) - cutoff) / cutoff,
-            err = mean(pred != actual)) %>%
-  ungroup()
-
-# con_tab <- left_join(
-#   con_tab %>% filter(forecaster != "AR"),
-#   con_tab %>% filter(forecaster == "AR") %>%
-#     select(-forecaster) %>%
-#     rename(mar = m))
-
-llike_tab <- hot_udf %>%
-  filter(!is.na(actual)) %>%
-  # left_join(null_classifier %>% select(geo_value, time_value, null_class) %>%
-  #             rename(target_end_date = time_value)) %>%
-  mutate( # kill prob 0-1 predictions
-    value = case_when(
-      value < 1e-8 ~ 1e-8,
-      value > 1-1e-8 ~ 1-1e-8,
-      TRUE ~ value),
-    # null_class = case_when(
-    #   null_class < 1e-8 ~ 1e-8,
-    #   null_class > 1-1e-8 ~ 1-1e-8,
-    #   TRUE ~ null_class),
-    llike = (actual == 1) * log(value) + (actual == 0) * log(1 - value),
-    nulldev = (actual == 1) * log(cutoff) + 
-      (actual == 0) * log(1 - cutoff)) %>%
-  group_by(forecaster, udf) %>%
-  summarise(m = (mean(llike) - mean(nulldev)) / abs(mean(nulldev)),
-            mm = mean(llike),
-            mmm = mean(nulldev)) %>%
-  ungroup()
-
-# llike_tab <- left_join(
-#   llike_tab %>% filter(forecaster != "AR"),
-#   llike_tab %>% filter(forecaster == "AR") %>%
-#     select(-forecaster) %>%
-#     rename(mar = m))
-
-bind_rows(con_tab %>% mutate(err = "Classification error"), 
-          llike_tab %>% mutate(err = "Log likelihood")) %>%
-  ggplot(aes(udf, m , fill = forecaster)) +
-  geom_bar(width = 0.6, position = position_dodge(width=0.6),
-           stat = "identity") +
-  scale_fill_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
-  #scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
-  theme_bw() +
-  geom_hline(yintercept = 0) +
-  scale_y_continuous(labels = scales::percent_format()) +
-  facet_wrap(~err, scales = "free_y") +
-  ylab("Improvement relative to the null classifier") +
-  xlab("") +
-  theme(legend.position = "bottom", legend.title = element_blank())
-```
-
-\clearpage
 
@@
 
-```{r cor-wis-ratio-m1, eval = FALSE, fig.cap="This is Alden's second set of histograms.
-Here we have the correlation of the absolute value of WIS ratio - 1 with the percent change in 7dav cases relative to 7 days earlier"}
-ggplot(pct_chng_df %>% 
-         group_by(forecaster) %>% 
-         mutate(median = median(cor_abs_wisratio_minus_1_pctchange)), 
-       aes(x = cor_abs_wisratio_minus_1_pctchange)) +
-  geom_histogram(aes(y = ..count.. / sum(..count..), fill = forecaster), bins = 50) +
-  scale_fill_manual(values = fcast_colors) +
-  theme_bw() +
-  theme(legend.position = "none") +
-  geom_vline(aes(xintercept = median), linetype = "dotted") + 
-  geom_vline(xintercept = 0) +
-  facet_wrap(~forecaster) + 
-  xlab("Spearman correlation") + ylab("Relative frequency") +
-  scale_y_continuous(labels = scales::label_percent(accuracy = 1))
-```
-
-\clearpage
-
 ```{r upswing-corr-table, fig.cap="Correlation of the difference in WIS with the difference in median predictions for the AR model relative to the indicator-assisted forecaster.
@@
            stat = "identity") +
-  scale_fill_manual(values = fcast_colors[1:5], guide = guide_legend(nrow = 1)) +
+  scale_fill_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
   theme_bw() +
-  geom_hline(yintercept = 0) +
+  geom_hline(yintercept = 0, linetype = "dashed") +
   theme(legend.position = "bottom", legend.title = element_blank())
@@
   scale_y_continuous(labels = scales::percent_format()) +
-  geom_hline(yintercept = 0) +
+  geom_hline(yintercept = 0, linetype = "dashed") +
   ylab("% change in Forecaster WIS\n relative to Baseline") +
@@
            stat = "identity") +
-  geom_hline(yintercept = 0) +
-  scale_fill_manual(values = fcast_colors[1:5]) +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  scale_fill_manual(values = fcast_colors2) +
   guides(fill = guide_legend(nrow = 1)) +
@@
         legend.background = element_rect(fill = "transparent"))
-```
-
-```{r leading-and-lagging, eval=FALSE}
-# Deprecated
-df2 %>%
-  group_by(forecaster, udf) %>%
-  summarise(
-    leadingness = cor(AR_wis - wis, leading, use = "pairwise.complete.obs"),
-    laggingness = cor(AR_wis - wis, lagging, use = "pairwise.complete.obs")) %>%
-  pivot_longer(leadingness:laggingness) %>%
-  ggplot(aes(udf, value, fill = forecaster)) +
-  geom_bar(width = 0.6, position = position_dodge(width=0.6),
-           stat = "identity") +
-  geom_hline(yintercept = 0) +
-  scale_fill_manual(values = fcast_colors) +
-  facet_wrap(~name) +
-  theme_bw() +
-  ylab("Correlation") +
-  xlab("") +
-  theme(legend.position = "bottom", legend.title = element_blank())
 ```
@@
            stat = "identity") +
-  geom_hline(yintercept = 0) +
-  scale_fill_manual(values = fcast_colors[1:5], guide = guide_legend(nrow = 1)) +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  scale_fill_manual(values = fcast_colors2, guide = guide_legend(nrow = 1)) +
   theme_bw() +
@@
            stat = "identity") +
-  geom_hline(yintercept = 0) +
-  scale_fill_manual(values = fcast_colors[1:5]) +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  scale_fill_manual(values = fcast_colors2) +
   theme_bw() +
@@
 source(here("code", "figures", "trajectory.R"))
-gg
 ```
@@
                color = "grey70") +
-  geom_hline(yintercept = 1) +
+  geom_hline(yintercept = 1, linetype = "dashed") +
   xlab("forecast date") +
@@
                color = "grey70") +
-  geom_hline(yintercept = 1) +
+  geom_hline(yintercept = 1, linetype = "dashed") +
   xlab("forecast date") + ylab("Cumulative Geometric Mean") +
@@
 ```
-
-\clearpage
-
-<!-- Comparison to submitted forecasts -->
-
-```{r grab-submitted-case-fcasts}
-library(aws.s3)
-Sys.setenv("AWS_DEFAULT_REGION" = "us-east-2")
-s3bucket <- get_bucket("forecast-eval")
-# ensure that forecasters have submitted a reasonable amount
-n_keeper_weeks <- 6L
-n_keeper_locs <- 50L
-case_scores <- s3readRDS("score_cards_state_cases.rds", s3bucket) %>% 
-  mutate(forecast_date = target_end_date - ahead * 7) %>% 
-  # fix weirdnesses about submission dates, 
-  # this is generous to the forecaster
-  filter(forecast_date < "2021-01-01", ahead < 4) %>% 
-  select(ahead, geo_value, forecaster, target_end_date, wis, forecast_date)
-
-strawman <- case_scores %>% filter(forecaster == "COVIDhub-baseline")
-case_scores <- left_join(
-  case_scores, 
-  strawman %>% 
-    select(forecast_date, target_end_date, geo_value, wis) %>% 
-    rename(strawman_wis = wis)
-) %>% filter(forecaster != "COVIDhub-baseline")
-
-# Subset to those forecasters that submitted enough 14-day ahead forecasts
-n_submitted <- case_scores %>% 
-  filter(ahead == 2) %>% 
-  group_by(forecaster) %>% 
-  summarise(nfcasts = n())
-keepers <- n_submitted %>% 
-  filter(nfcasts / n_keeper_locs > n_keeper_weeks - .0001, 
-         forecaster != "COVIDhub-4_week_ensemble") %>% # same as CH-Ensemble 
-  pull(forecaster)
-
-case_scores <- case_scores %>% 
-  filter(forecaster %in% keepers) %>%
-  group_by(forecaster, ahead) %>%
-  summarise(rel_wis = Mean(wis) / Mean(strawman_wis),
-            geo_wis1 = GeoMean((wis + 1) / (strawman_wis + 1)),
-            geo_wis = GeoMean(wis / strawman_wis))
-```
-
-```{r compare-to-hub-mean, fig.height = 8.5, fig.cap="This figure reproduces Figure 3 in the main paper but overlays scores for the forecasts submitted to the COVID-19 Forecast Hub.
-Grey lines correspond to the various teams that submitted during period our evaluation period.
-We have highlighted the COVIDhub-ensemble, which is the official forecast of the CDC.", include=FALSE}
-# Same as Fig 3 in the paper
-df <- summarizer(fcasts_honest %>%
-                   group_by(ahead, forecaster), "wis", NULL, 
-                 "strawman_wis", Mean, c("aggr","scale"))
-ggplot(df) +
-  geom_line(aes(ahead, wis, color = forecaster)) + 
-  geom_point(aes(ahead, wis, color = forecaster)) +
-  theme_bw() +
-  geom_hline(yintercept = 1, size = 1.5) +
-  xlab("Days ahead") +
-  ylab("Mean WIS (relative to baseline)") +
-  scale_color_manual(values = c(fcast_colors, "CH-ensemble" = "lightblue"),
-                     guide = guide_legend(nrow = 2)) +
-  theme(legend.position = "bottom", legend.title = element_blank()) +
-  geom_line(data = case_scores, 
-            aes(ahead * 7, rel_wis, group = forecaster),
-            color = "grey70") +
-  geom_line(data = case_scores %>% filter(forecaster == "COVIDhub-ensemble"),
-            aes(ahead * 7, rel_wis), color = "lightblue", size = 1.5) +
-  scale_y_log10()
-```
-
-\clearpage
-
-```{r compare-to-hub-geomean, fig.height = 8.5, fig.cap="This figure is similar to Figure \\ref{fig:fcast-adjusted}.
-In this case, we add 1 to both the forecaster WIS and the baseline WIS before scaling (to allow forecasters that achieve 0 error to appear), and we overlay scores for the forecasts submitted to the COVID-19 Forecast Hub.
-Grey lines correspond to the various teams that submitted during period our evaluation period.
-We have highlighted the COVIDhub-ensemble, which is the official forecast of the CDC.", include=FALSE}
-# Like supplement GeoMean figure, but adding 1
-df2 <- fcasts_honest %>% 
-  group_by(ahead, forecaster) %>%
-  summarise(wis = GeoMean((wis + 1) / (strawman_wis + 1)))
-
-ggplot(df2) +
-  geom_line(aes(ahead, wis, color = forecaster)) + 
-  geom_point(aes(ahead, wis, color = forecaster)) +
-  theme_bw() +
-  geom_hline(yintercept = 1, size = 1.5) +
-  xlab("Days ahead") +
-  ylab("Geometric mean of WIS (relative to baseline)") +
-  scale_color_manual(values = c(fcast_colors, "CH-ensemble" = "lightblue"),
-                     guide = guide_legend(nrow = 2)) +
-  theme(legend.position = "bottom", legend.title = element_blank()) +
-  geom_line(data = case_scores, 
-            aes(ahead * 7, geo_wis1, group = forecaster),
-            color = "grey70") +
-  geom_line(data = case_scores %>% filter(forecaster == "COVIDhub-ensemble"),
-            aes(ahead * 7, geo_wis1), color = "lightblue", size = 1.5) +
-  scale_y_log10()
-```
-
-<!-- All code below is deprecated -->
-
-```{r fcast-gs-locations, eval=FALSE}
-plotter(fcasts_gs %>% filter(period != "jm"),
-        "wis", Mean, scaler = "strawman_wis", 
-        order_of_operations = c("aggr","scale")) +
-  ylab("Mean WIS (relative to baseline)")
-```
-
-```{r fcast-gs-locations-adjusted, eval=FALSE}
-plotter(fcasts_gs %>% filter(period != "jm"), 
-        "wis", GeoMean, scaler = "strawman_wis", 
-        order_of_operations = c("scale","aggr")) +
-  ylab("Geometric mean of WIS (relative to baseline)")
-```
-
-```{r hot-gs-locations, eval=FALSE}
-plotter_hotspots(hotspots_gs %>% filter(period != "jm")) +
-  geom_hline(yintercept = 0.5)
-```
-
-```{r traj-data, eval = FALSE}
-source(here("code", "deprecated", "trajectory_plot_funs.R"))
-preds <- readRDS(here("data", "predictions_honest.RDS"))
-traj_best <- get_trajectory_plots(fcasts_honest, preds, actuals, hrr_tab, 
-                                  "only2020", "best")
-traj_worst <- get_trajectory_plots(fcasts_honest, preds, actuals, hrr_tab,
-                                   "only2020", "worst")
-rm(preds)
-```
-
-```{r trajectory-plots, eval = FALSE}
-for (nm in names(traj_best)) print(trajectory_panel(nm, traj_best, traj_worst))
-```

### 2021-10-31 forecast/code/figures/compare-to-hub.R

@@ -7,24 +7,31 @@ Sys.setenv("AWS_DEFAULT_REGION" = "us-east-2")
 s3bucket <- get_bucket("forecast-eval")
 n_keeper_weeks <- 6L
 n_keeper_locs <- 50L
-# case_scores <- s3readRDS("score_cards_state_cases.rds", s3bucket) 
-case_preds <- s3readRDS("predictions_cards.rds", s3bucket) %>%
-  filter(signal == "confirmed_incidence_num",
-         forecast_date < "2021-01-01",
-         !(geo_value %in% c("as", "gu", "pr", "vi", "mp", "us"))) %>%
-  select(-data_source, -signal, -incidence_period)
-actuals <- covidcast::covidcast_signal("jhu-csse", "confirmed_7dav_incidence_num",
-                                       geo_type = "state", as_of = "2021-05-18",
-                                       end_day = "2021-03-01") %>%
-  select(geo_value, time_value, value) %>%
-  mutate(value = value * 7) %>%
-  rename(target_end_date = time_value, actual = value)
-case_scores <- evalcast::evaluate_predictions(
-  case_preds, actuals, 
-  err_measures = list(wis = evalcast::weighted_interval_score),
-  grp_vars = c("ahead", "forecaster", "forecast_date", "geo_value")
-)
-rm(case_preds)
+case_scores <- s3readRDS("score_cards_state_cases.rds", s3bucket) 
+# case_preds <- s3readRDS("predictions_cards.rds", s3bucket) %>%
+#   filter(signal == "confirmed_incidence_num",
+#          forecast_date < "2021-01-01",
+#          !(geo_value %in% c("as", "gu", "pr", "vi", "mp", "us"))) 
+# actuals <- covidcast::covidcast_signal("jhu-csse", "confirmed_7dav_incidence_num",
+#                                        geo_type = "state", as_of = "2021-05-18",
+#                                        end_day = "2021-03-01") %>%
+#   mutate(value = value * 7) %>%
+#   rename(target_end_date = time_value, actual = value)
+# case_scores <- evalcast::evaluate_predictions(
+#   case_preds, actuals, 
+#   err_measures = list(wis = evalcast::weighted_interval_score),
+#   grp_vars = c("ahead", "forecaster", "forecast_date", "geo_value")
+# )
+# rm(case_preds)
+
+### RJT: using evalcast above leads to the following error for me. I even made 
[…]

### 2021-10-31 forecast/code/figures/eval_funs.R

@@ -1,7 +1,6 @@
-fcast_colors <- c(RColorBrewer::brewer.pal(5, "Set1"), "#000000")
-names(fcast_colors) <- c("CHNG-CLI", "CHNG-COVID", "CTIS-CLIIC", "DV-CLI",
-                         "Google-AA", "AR")
-
+fcast_colors <- c("#000000", RColorBrewer::brewer.pal(5, "Set1"))
+names(fcast_colors) <- c("AR", "CHNG-CLI", "CHNG-COVID", "CTIS-CLIIC", "DV-CLI",
+                         "Google-AA")
 
 Mean <- function(x) mean(x, na.rm = TRUE)
 Median <- function(x) median(x, na.rm = TRUE)
@@ -10,7 +9,6 @@ TrimMean025 <- function(x) mean(x, trim = .025, na.rm = TRUE)
 TrimMean05 <- function(x) mean(x, trim = .05, na.rm = TRUE)
 TrimMean1 <- function(x) mean(x, trim = .1, na.rm = TRUE)
 GeoMean <- function(x) exp(mean(log(x), na.rm = TRUE))
-
 
 pct_change <- function(x, n = 14, col_name = "pct_change"){
   # courtesy of Alden

### 2021-10-31 forecast/paper/paper.tex
Copyedits to discussion

@@
 Supplementary Materials we show that for forecasting at the state level, the
-benefits of adding indicators to the AR model are generally less clear
+benefits of including indicators in the AR model are generally less clear
 (compared to those observed at the HRR level).  A plausible explanation: at the
 state level, where the signal-to-noise ratio (SNR) is higher, AR performs better
-overall and provides a higher standard (when asking whether it can be improved 
-upon using indicators).  At the HRR level, where the SNR is lower, including the
-indicators as additional linear features in the AR model probably delivers a
+overall and represents a higher standard (when asking whether it can be improved
+upon using the indicators).  At the HRR level, where the SNR is lower, including
+the indicators as additional linear features in the AR model probably delivers a
 kind of variance reduction (just like averaging independent terms would) which
-contributes to improved accuracy.  But at a higher SNR, this variance
-reduction becomes less important, and we may be compelled to use more
-sophisticated modeling techniques to extract comparable value from the
-indicators.
+contributes to improved accuracy.  But as the SNR increases, this variance
+reduction becomes less important, and perhaps we must use more sophisticated
+modeling techniques to extract comparable value from the indicators.
 

### 2021-10-31 forecast/paper/supplement-text.tex
Copyedits to discussion

@@
 truth have zero relationship to the response.  Recall (from the main paper) that
-at each forecast date, we train a model on 6,426 location-time pairs.  Indicator
-models are based on six features, corresponding to the three autoregressive
-terms and the three lagged indicator values.  To form noise indicator features,
-we replace their values with those from a randomly chosen time-space pair (while
-keeping the autoregressive features fixed).  In particular, at each location
-$\ell$ and time $t$, for the forecasting task we replace the triplet
-\smash{$(X_{\ell,t}, X_{\ell,t-7}, X_{\ell,t-14})$} in Eq.\ (3) of the main
+at each forecast date, we train a model on 6,426 location-time pairs.  Each
+indicator model uses 6 features, corresponding to the 3 autoregressive terms and
+the 3 lagged indicator values.  To form noise indicator features, we replace
+their values with those from a randomly chosen time-space pair (while keeping
+the autoregressive features fixed).  In particular, at each location $\ell$ and
+time $t$, for the forecasting task we replace the triplet \smash{$(X_{\ell,t},
+X_{\ell,t-7}, X_{\ell,t-14})$} in Eq.\ (3) of the main
 paper with the triplet \smash{$(X_{\ell^*,t^*}, X_{\ell^*,t^*-7},
@@
 In this section, we provide extra details about the up/down/flat analysis
-described in the main text. Figure \ref{fig:upswing-summary} shows the overall
-results, examining the average difference $\WIS(AR) - \WIS(F)$ in period. Figure
-\ref{fig:hotspots-upswing-downswing} shows the same information for the hotspot
-task. On average, during downswings and flat periods, the indicator models have
-lower classification error and higher log likelihood than the AR model. For
-hotspots, both Google-AA and CTIS-CLIIC perform better than the AR model during
-upswings, in contrast to the forecasting task, where only Google-AA
-improves. For a related analysis, Figure \ref{fig:cor-wis-ratio} shows
+described in the main manuscript. Figure \ref{fig:upswing-summary} shows the
+overall results, examining the average difference $\WIS(AR) - \WIS(F)$ in
+period. Figure \ref{fig:hotspots-upswing-downswing} shows the same information
+for the hotspot task. On average, during downswings and flat periods, the
+indicator models have lower classification error and higher log likelihood than
+the AR model. For hotspots, both Google-AA and CTIS-CLIIC perform better than
+the AR model during upswings, in contrast to the forecasting task, where only
+Google-AA improves. For a related analysis, Figure \ref{fig:cor-wis-ratio} shows
 histograms of the Spearman correlation (Spearman's $\rho$, a rank-based measure

### 2021-10-31 forecast/paper/supplement.Rmd
Copyedits to discussion

@@
 Compare to Figure 5 in the manuscript."}
-# Deprecated
 df2 %>%
@@
 The relationship is essentially the same as described in the manuscript and shown in Figure 5."}
-# Deprecated
 df2 %>%

### 2021-10-31 forecast/paper/supplement-text.tex
DONE with pass over supp

@@
 the autoregressive features fixed).  In particular, at each location $\ell$ and
-time $t$, for the forecasting task we replace the triplet \smash{$(X_{\ell,t},
-X_{\ell,t-7}, X_{\ell,t-14})$} in Eq.\ (3) of the main
+time $t$, for the forecasting task we replace the triplet $(X_{\ell,t},
+X_{\ell,t-7}, X_{\ell,t-14})$ in Eq.\ (3) of the main
 paper with the triplet \smash{$(X_{\ell^*,t^*}, X_{\ell^*,t^*-7},
@@
 
-In this section, we provide extra details about the up/down/flat analysis
-described in the main manuscript. Figure \ref{fig:upswing-summary} shows the
-overall results, examining the average difference $\WIS(AR) - \WIS(F)$ in
-period. Figure \ref{fig:hotspots-upswing-downswing} shows the same information
-for the hotspot task. On average, during downswings and flat periods, the
-indicator models have lower classification error and higher log likelihood than
-the AR model. For hotspots, both Google-AA and CTIS-CLIIC perform better than
-the AR model during upswings, in contrast to the forecasting task, where only
-Google-AA improves. For a related analysis, Figure \ref{fig:cor-wis-ratio} shows
-histograms of the Spearman correlation (Spearman's $\rho$, a rank-based measure
-of association) between the $\WIS(F)/\WIS(AR)$ and the magnitude of the
-swing. Again we see that case rate increases are positively related to
-diminished performance of the indicator models.
+In this section, we provide extra details about the upswing/downswing analysis
+described in the main manuscript, Section 2.C. Figure \ref{fig:upswing-summary}
+shows the overall results, examining the average difference $\WIS(\mathrm{AR}) - 
+\WIS(F)$ for each forecaster $F$, in each in each period.  Figure
+\ref{fig:hotspots-upswing-downswing} shows the same information for the hotspot 
+task.  On average, during downswings and flat periods, the indicator models have
+lower classification error and higher log likelihood than the AR model.  For
+hotspots, both Google-AA and CTIS-CLIIC perform better than the AR model during
+upswings, in contrast to the forecasting task, where only Google-AA improves.
+In a related analysis, Figure \ref{fig:cor-wis-ratio} shows histograms of the
+Spearman correlation (Spearman's $\rho$, a rank-based measure of association)
+between the ratio $\WIS(F)/\WIS(\mathrm{AR})$ and the magnitude of the swing.
+Again we see that case rate increases are positively related to diminished
+performance of the indicator models.
 
@@
 on upswings.  Figure \ref{fig:upswing-corr-table} shows the correlation between
-between $\WIS(AR) - \WIS(F)$ and the difference of their median forecasts.
-During downswings, this correlation is large, implying that improved relative
-performance of $F$ is related to making lower forecasts than the AR model. The
-opposite is true during upswings. This is largely to be expected. However, the
-relationship attenuates in flat periods and during upswings. That is, when
-performance is better in those cases, it may be due to other factors than simply
-making predictions in the correct direction, for example, narrower confidence
-intervals.
+between $\WIS(\mathrm{AR}) - \WIS(F)$ and the difference of their median
+forecasts.  During downswings, this correlation is large, implying that improved
+relative performance of $F$ is related to making lower forecasts than the AR
+model.  The opposite is true during upswings.  This is largely to be expected.
+However, the relationship attenuates in flat periods and during upswings. That
+is, when performance is better in those cases, it may be due to other factors
+than simply making predictions in the correct direction, for example, narrower
+confidence intervals.
 
-It is important to note, that even though some indicators, notably CHNG-COVID 
-and CHNG-CLI underperform relative to the AR model during upswings, all models 
-dramatically outperform the baseline in such periods. Figure 
-\ref{fig:upswing-summary-remake} shows the performance of all forecasters relative
-to the null model. All forecasters suffer relative to the baseline during down 
-periods, but the AR is the worst. In contrast, all models beat the baseline
-during up periods, even CHNG-COVID and CHNG-CLI, though not by quite as much as the
-AR does.
+It is important to note that, even though some indicators---notably CHNG-COVID
+and CHNG-CLI---underperform relative to the AR model during upswings, all models
+dramatically outperform the baseline in such periods.  Furthermore, Figure
+\ref{fig:upswing-summary-remake} shows the performance of all forecasters
+relative to the baseline model.  All forecasters suffer relative to the baseline
+during down periods, but the AR is the worst. In contrast, all models beat the
+baseline during up periods, even CHNG-COVID and CHNG-CLI, though not by quite as
+much as the AR does.
 
@@
 function (CCF) between the two time series. The $\CCF_\ell(a)$ of an indicator
-$X_{\ell}$ and case rates $Y_{\ell}$ is defined as
-% \[
-% \CCF_{\ell,t}(a) = \frac{1}{n_a} \sum_{i=1}^{n_a} \left( ( X_{\ell,t,i+a} -
-%   \overline{X}_{\ell,t}) / s^{X}_{\ell,t}\right) \left( ( Y_{\ell,t,i} -
-%   \overline{Y}_{\ell,t}) / s^{Y}_{\ell,t}\right), 
-% \]
-% where $n_a$ is the number of available time points when $X_{\ell,t}$ has been
-% shifted in time by $a$ days, and $\overline{X}_{\ell,t}$, $s^X_{\ell,t}$
-% are the sample mean and
-% standard deviation of $X_{\ell,t}$ (respectively $Y_{\ell,t}$). The result is a
-their Pearson correlation where $X_\ell$ has been aligned with the values of
-$Y_{\ell}$ that occurred $a$ days earlier.  Thus, for any $a>0$,
-$\CCF_{\ell}(a)>0$ indicates that $Y_{\ell,t}$ is moving together with
-$X_{\ell,t+a}$.  In this case we say that $X_{\ell}$ is lagging $Y_{\ell}$. For
-$a<0$, $\CCF_{\ell}(a)>0$ means that $Y_{\ell,t}$ is positively correlated with
-$X_{\ell,t-a}$, so we say that $X_{\ell}$ leads $Y_{\ell}$.
+series $X_{\ell}$ and case rate series $Y_{\ell}$ is defined as their Pearson 
+correlation where $X_\ell$ has been aligned with the values of $Y_{\ell}$ that
+occurred $a$ days earlier.  Thus, for any $a>0$, $\CCF_{\ell}(a)>0$ indicates
+that $Y_{\ell,t}$ is moving together with $X_{\ell,t+a}$.  In this case we say
+that $X_{\ell}$ is lagging $Y_{\ell}$. For $a<0$, $\CCF_{\ell}(a)>0$ means that
+$Y_{\ell,t}$ is positively correlated with $X_{\ell,t-a}$, so we say that
+$X_{\ell}$ leads $Y_{\ell}$.
 
@@
 containing Charlotte, North Carolina, from August 1, 2020 until the end of
-September. These are the same signals shown in Figure 1 in the manuscript but
+September. These are the same signals shown in Figure 1 in the manuscript, but 
 using finalized data. To define ``leadingness'' we compute $\CCF_{\ell}(a)$ (as
@@
 Figure \ref{fig:lagging-only} shows the correlation between laggingness and the
-difference in indicator WIS and AR WIS. Unlike leadingness (Figure 5 in the
+difference in forecaster WIS and AR WIS. Unlike leadingness (Figure 5 in the
 manuscript) there is no obvious relationship that holds consistently across
-indicators. This is heartening as laggingness should not aid forecasting
+indicators. This is encouraging as laggingness should not aid forecasting 
 performance. On the other hand, if an indicator is more lagging than it is
 leading, this may suggest diminished performance.
-Figure~\ref{fig:diff-in-lead-lag} shows the correlation of the difference in
+Figure \ref{fig:diff-in-lead-lag} shows the correlation of the difference in
 leadingness and laggingness with the difference in WIS. The pattern here is
@@
 for the trained model and for the evaluation data and examine their combination
-in some way. We do not pursue this combination further and leave this
-investigation for future work.
+in some way.  However, we do not pursue this further.
 
-% \section{Examining data in 2021}
-
-% In this section, we investigate the sensitivity of the results to the
-% period over which we train and evaluate the models.  In the main
-% paper, we end all evaluation on December 31, 2020.  Figures
-% \ref{fig:fcast-alldates}--\ref{fig:hot-alldates} show how the
-% results would differ if we extended this analysis through March
-%   31, 2021. Comparing Figure \ref{fig:fcast-alldates} to Figure 3 of
-% the main paper, one sees that as ahead increases most methods now improve
-% relative to the baseline forecaster. When compared to other methods, 
-% \chngcli~appears much better than
-% it had previously; however, all forecasters other than \chngcov~and
-% \dv~are performing less well relative to the baseline than before.
-% These changes are likely due to the differing nature of the pandemic
-% in 2021, with flat and downward trends much more common than upward
-% trajectories.  Indeed, the nature of the hotspot prediction problem is
-% quite different in this period.  With a 21-day training window, it is
-% common for there to be many fewer hotspots in training.
-
-\section{Aggregation Over Time and Space}
+\section{Disaggregation Over Time and Space}
 
@@
 other disaggregated versions of the main forecasting result shown in Figure 3 on the
-manuscript. The first (Figure \ref{fig:cumulative-mean}) displays the cumulative
-error (summed over all horizons) of each forecaster divided by the cumulative
-error of the baseline.  This perspective should illustrate how the models
-perform over time, drawing attention to any nonstationary behavior.  During the
-initial increase in cases in July 2020, CTIS-CLIIC and Google-AA gain a lot
-relative to the AR model. All the indicators do much better than the AR model
-during the following downturn (the ebb of the second wave). The AR model
-actually improves over the indicators in October 2020, before losing a bit in
-late November. The second (Figure \ref{fig:cumulative-geo-mean}) repeats this 
-analysis but aggregating by Geometric mean as described above and displays a
-similar pattern.
+manuscript. Below we use the term ``error'' to refer to the WIS summed over all
+ahead values.  The first (Figure \ref{fig:cumulative-mean}) displays the
+cumulative error, up through any point in time, of each forecaster divided by
+the cumulative error of the baseline.  This perspective should illustrate how
+the models perform over time, drawing attention to any nonstationary behavior.
+During the initial increase in cases in July 2020, CTIS-CLIIC and Google-AA gain
+quite a bit acurracy compared to the AR model.  All the indicators do much
+better than the AR model during the following downturn (the ebb of the second 
+wave). The AR model actually improves over the indicators in October 2020,
+before losing a bit in late November. % The second (Figure
+% \ref{fig:cumulative-geo-mean}) repeats this analysis but aggregating by 
+% geometric mean, and presents a similar message.
 
-Figure \ref{fig:errs-in-space} examines the spatial behavior over the entire
-period of the indicator models relative to the AR. For ease of
-comparison, we show the percent improvement in each HRR. Negative numbers (blue)
-mean that the indicator helped while positives (red) mean that the indicator
-hurt forecasting performance. The clear pattern is that in most HRRs, the
-indicators improved performance, though usually by small amounts (2.5\%--10\%).
-In some isolated HRRs, performance was markedly worse, though there does not
-appear to be any particular pattern to these locations.
+Figure \ref{fig:errs-in-space} examines the spatial behavior of errors over the
+entire period of the indicator models relative to the AR.  For ease of
+comparison, we show the percent improvement in each HRR.  Negative numbers
+(blue) mean that the indicator helped while positives (red) mean that the
+indicator hurt forecasting performance.  The clear pattern is that in most HRRs,
+the indicators improved performance, though usually by small amounts
+(2.5\%--10\%).  In some isolated HRRs, performance was markedly worse, though
+there does not appear to be any particular pattern to these locations.
+Interestingly, the geographic patterns of improvement differ quite a lot in
+between the indicators.  This suggests that a forecasting model that carefully
+combines all of indicators could be a considerable improvement.

### 2021-10-31 forecast/paper/supplement.Rmd
DONE with pass over supp

@@
 
-```{r upswing-summary, fig.cap="Average difference between the WIS of the AR model and of the indicator models. The indicator models generally do best during down and flat periods."}
+```{r upswing-summary, fig.cap="Average difference between the WIS of the AR model and of the indicator models, separated into up, down, and flat periods.
+The indicator models generally do best during down and flat periods."}
 up_down_df_summary <- up_down_df %>%
@@
   geom_hline(yintercept = 0, linetype = "dashed") +
-  ylab("Mean of WIS(AR) - WIS(F)") +
+  ylab("Mean of AR WIS - forecaster WIS") +
   xlab("") +
@@
 
-```{r hotspots-upswing-downswing, fig.cap="Classification and loglikelihood separated into periods of upswing, downswing, and flat cases.
-Like the analysis of the forecasting task in the main paper (see Figure 7), relative performance is better during down and flat periods."}
+```{r hotspots-upswing-downswing, fig.cap="Percentage change in classification error and log likelihood, relative that of the AR model, separated into up, down, and flat periods.
+Like the analogous forecasting analysis, the indicator models generally do better during down and flat periods."}
 hot_udf <- inner_join(
@@
   facet_wrap(~err) +
-  ylab("Change relative to AR") +
+  ylab("% change relative to AR") +
   xlab("") +
@@
 
-```{r cor-wis-ratio, fig.cap="Histograms of the Spearman correlation between the ratio of AR to AR WIS with the percent change in smoothed case rates relative to 7 days earlier."}
+```{r cor-wis-ratio, fig.cap="Histograms of the Spearman correlation between the ratio of forecaster WIS to AR WIS with the % change in case rates, relative to case rates 7 days earlier."}
 ggplot(pct_chng_df %>% 
@@
   facet_wrap(~forecaster) + 
-  xlab("Spearman correlation") + ylab("Relative frequency") +
+  xlab("Spearman correlation") + ylab("Frequency") +
   scale_y_continuous(labels = scales::label_percent(accuracy = .1))
@@
 
-```{r upswing-corr-table, fig.cap="Correlation of the difference in WIS with the difference in median predictions for the AR model relative to the indicator-assisted forecaster.
-In down periods, improvements in forecast risk are highlycorrelated with lower median predictions.
-The opposite is true in up periods.
-This suggests, as one might expect that improved performance of the indicator-assisted model is attributable to being closer to the truth then the AR model.
-This conclusion is stronger in down periods then in up periods."}
+```{r upswing-corr-table, fig.cap="Correlation of the difference in WIS with the difference in median predictions (each difference being between the AR model and an indicator model), separated into up, down, and flat periods.
+In down periods, improvements in forecast error are highly correlated with lower median predictions.
+The opposite is true in up periods, but the conclusion here appears to be weaker."}
 up_down_df %>% 
@@
   scale_fill_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
-  theme_bw() +
+  theme_bw() + ylab("Correlation") + xlab("") +
   geom_hline(yintercept = 0, linetype = "dashed") +
   theme(legend.position = "bottom", legend.title = element_blank())
-
-# up_down_df %>% 
-#   filter(period != "jm") %>% 
-#   group_by(forecaster, udf) %>%
-#   summarise(cor = cor(AR_wis - wis, AR_med - med)) %>%
-#   pivot_wider(names_from = forecaster, values_from = cor) %>%
-#   kableExtra::kbl(
-#     booktabs = TRUE, digits = 2, centering = TRUE,
-#     caption = paste("Correlation of the difference in WIS between the AR",
-#                    "model with the difference in median predictions. In down",
- #                   "periods, improvements in forecast risk are highly",
-#                    "correlated with lower median predictions. The opposite",
-#                    "is true in up periods. This suggests, as one might expect",
-#                    "that improved performance of the indicator-assisted",
-#                    "model is attributable to being closer to the truth then",
-#                    "the AR model. This conclusion is stronger in down",
-#                    "periods then in up periods."))
 ```
@@
 
-```{r upswing-summary-remake, fig.cap="Percent change in average WIS of the forecaster (AR or indicator assisted) relative to the baseline.
+```{r upswing-summary-remake, fig.cap="Percentage change in average WIS of the forecaster (AR or indicator assisted), relative to the baseline.
 All models perform poorly durning down periods, but the indicators help.
@@
   geom_hline(yintercept = 0, linetype = "dashed") +
-  ylab("% change in Forecaster WIS\n relative to Baseline") +
+  ylab("% change in WIS relative to baseline") +
   xlab("") +
@@
 
-```{r ccf-dv-finalized, eval=TRUE, fig.cap="Illustration of the cross-correlation function between DV-CLI and cases.
+```{r ccf-dv-finalized, fig.cap="Illustration of the cross-correlation function between DV-CLI and cases.
 The left panel shows the standardized signals over the period from August 1 to September 28 (as of May 15, 2021).
@@
 The orange dashed lines indicate the 95\\% significance threshold.
-By our leadingness/laggingness metric, DV-CLI is leading (but notlagging) cases over this period."}
+By our leadingness/laggingness metric, DV-CLI is leading (but not lagging) cases over this period."}
 source(here("code", "figures", "ccf-dv-finalized.R"))
@@
 
-```{r cumulative-mean, fig.cap="This figure shows the cumulative sum of WIS for each forecaster divided by the cumulative sum of WIS for the baseline model.
-The shaded background shows total U.S.
-cases as reported by JHU-CSSE for the 14-day ahead target.
-Hashes along the $x$-axis denote weeks." }
+```{r cumulative-mean, fig.cap="Cumulative sum of WIS for each forecaster divided by the cumulative sum of WIS for the baseline model.
+The shaded background shows national case incidence for the 14-day ahead target.
+Hashes along the x-axis denote weeks." }
 cumulatives <- fcasts_honest %>%
@@
 ggplot(cumulatives %>% filter(forecast_date < "2021-01-01")) + 
-  geom_line(aes(forecast_date, y = `Cumulative Sum`, color = forecaster)) +
+  geom_line(aes(forecast_date, y = `Cumulative sum`, color = forecaster)) +
   annotate(x = us_cases$time_value - 14, ymax = us_cases$value, ymin = -Inf,
@@
   geom_hline(yintercept = 1, linetype = "dashed") +
-  xlab("forecast date") +
+  xlab("Forecast date") + ylab("Cumulative sum") +
   coord_cartesian(#ylim = c(.76, 1), 
@@
 
-```{r cumulative-geo-mean, fig.cap="This figure shows the cumulative geometric mean of WIS for each forecaster divided by the WIS for the baseline model.
-The shaded background shows total U.S.
-cases as reported by JHU-CSSE for the 14-day ahead target.
-Hashes along the $x$-axis denote weeks." }
+```{r cumulative-geo-mean, fig.cap="Cumulative geometric mean of WIS for each forecaster divided by WIS for the baseline model.
+The shaded background shows national case incidence for the 14-day ahead target.
+Hashes along the x-axis denote weeks." , eval=FALSE}
 ggplot(cumulatives %>% filter(forecast_date < "2021-01-01")) + 
@@
   geom_hline(yintercept = 1, linetype = "dashed") +
-  xlab("forecast date") + ylab("Cumulative Geometric Mean") +
+  xlab("Forecast date") + ylab("Cumulative geo mean") +
   coord_cartesian(#ylim = c(.76, 1), 
@@
 
-```{r errs-in-space, fig.cap="Percent improvement in WIS relative to the AR forecaster by region (negative numbers indicate improved performance, positives indicate worsening)."}
+```{r errs-in-space, fig.cap="Percentage improvement in WIS, relative to the AR forecaster, by HRR (negative numbers indicate improved performance, positives indicate worsening)."}
 # fig.height = 9.6, fig.width = 6, out.width = "5in", out.height="8in", 

### 2021-10-31 forecast/code/deprecated/supplement.Rmd
Deprecated supplement code

\clearpage

<!-- Comparison to submitted forecasts -->

```{r grab-submitted-case-fcasts}
library(aws.s3)
Sys.setenv("AWS_DEFAULT_REGION" = "us-east-2")
s3bucket <- get_bucket("forecast-eval")
# ensure that forecasters have submitted a reasonable amount
n_keeper_weeks <- 6L
n_keeper_locs <- 50L
case_scores <- s3readRDS("score_cards_state_cases.rds", s3bucket) %>% 
  mutate(forecast_date = target_end_date - ahead * 7) %>% 
  # fix weirdnesses about submission dates, 
  # this is generous to the forecaster
  filter(forecast_date < "2021-01-01", ahead < 4) %>% 
  select(ahead, geo_value, forecaster, target_end_date, wis, forecast_date)

strawman <- case_scores %>% filter(forecaster == "COVIDhub-baseline")
case_scores <- left_join(
  case_scores, 
  strawman %>% 
    select(forecast_date, target_end_date, geo_value, wis) %>% 
    rename(strawman_wis = wis)
) %>% filter(forecaster != "COVIDhub-baseline")

# Subset to those forecasters that submitted enough 14-day ahead forecasts
n_submitted <- case_scores %>% 
  filter(ahead == 2) %>% 
  group_by(forecaster) %>% 
  summarise(nfcasts = n())
keepers <- n_submitted %>% 
  filter(nfcasts / n_keeper_locs > n_keeper_weeks - .0001, 
         forecaster != "COVIDhub-4_week_ensemble") %>% # same as CH-Ensemble 
  pull(forecaster)

case_scores <- case_scores %>% 
  filter(forecaster %in% keepers) %>%
  group_by(forecaster, ahead) %>%
  summarise(rel_wis = Mean(wis) / Mean(strawman_wis),
            geo_wis1 = GeoMean((wis + 1) / (strawman_wis + 1)),
            geo_wis = GeoMean(wis / strawman_wis))
```

```{r compare-to-hub-mean, fig.height = 8.5, fig.cap="This figure reproduces Figure 3 in the main paper but overlays scores for the forecasts submitted to the COVID-19 Forecast Hub. Grey lines correspond to the various teams that submitted during period our evaluation period. We have highlighted the COVIDhub-ensemble, which is the official forecast of the CDC.", include=FALSE}
# Same as Fig 3 in the paper
df <- summarizer(fcasts_honest %>%
                   group_by(ahead, forecaster), "wis", NULL, 
                 "strawman_wis", Mean, c("aggr","scale"))
ggplot(df) +
  geom_line(aes(ahead, wis, color = forecaster)) + 
  geom_point(aes(ahead, wis, color = forecaster)) +
  theme_bw() +
  geom_hline(yintercept = 1, size = 1.5) +
  xlab("Days ahead") +
  ylab("Mean WIS (relative to baseline)") +
  scale_color_manual(values = c(fcast_colors, "CH-ensemble" = "lightblue"),
                     guide = guide_legend(nrow = 2)) +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  geom_line(data = case_scores, 
            aes(ahead * 7, rel_wis, group = forecaster),
            color = "grey70") +
  geom_line(data = case_scores %>% filter(forecaster == "COVIDhub-ensemble"),
            aes(ahead * 7, rel_wis), color = "lightblue", size = 1.5) +
  scale_y_log10()
```

\clearpage

```{r compare-to-hub-geomean, fig.height = 8.5, fig.cap="This figure is similar to Figure \\ref{fig:fcast-adjusted}. In this case, we add 1 to both the forecaster WIS and the baseline WIS before scaling (to allow forecasters that achieve 0 error to appear), and we overlay scores for the forecasts submitted to the COVID-19 Forecast Hub. Grey lines correspond to the various teams that submitted during period our evaluation period. We have highlighted the COVIDhub-ensemble, which is the official forecast of the CDC.", include=FALSE}
# Like supplement GeoMean figure, but adding 1
df2 <- fcasts_honest %>% 
  group_by(ahead, forecaster) %>%
  summarise(wis = GeoMean((wis + 1) / (strawman_wis + 1)))

ggplot(df2) +
  geom_line(aes(ahead, wis, color = forecaster)) + 
  geom_point(aes(ahead, wis, color = forecaster)) +
  theme_bw() +
  geom_hline(yintercept = 1, size = 1.5) +
  xlab("Days ahead") +
  ylab("Geometric mean WIS (relative to baseline)") +
  scale_color_manual(values = c(fcast_colors, "CH-ensemble" = "lightblue"),
                     guide = guide_legend(nrow = 2)) +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  geom_line(data = case_scores, 
            aes(ahead * 7, geo_wis1, group = forecaster),
            color = "grey70") +
  geom_line(data = case_scores %>% filter(forecaster == "COVIDhub-ensemble"),
            aes(ahead * 7, geo_wis1), color = "lightblue", size = 1.5) +
  scale_y_log10()
```

```{r hotspots-upswing-downswing-remake, fig.cap="REMAKE: Classification and loglikelihood separated into periods of upswing, downswing, and flat cases. The indicator assisted models always have lower classification error relative to the null classifier, while the AR actually does worse in down and flat periods. In terms of loglikelihood, all forecasters have lower likelihood than the null classifier during up periods. In down and flat periods, the indicators generally improve over the AR, while the are worse during up periods.", eval=FALSE}
# mycummean <- function(x) {
#   isgood <- ! is.na(x)
#   denom <- cumsum(isgood)
#   x[!isgood] <- 0
#   num <- cumsum(x)
#   y <- num / denom
#   y[is.na(y)] <- .5 # deal with empties at the beginning
#   y
# }
# 
# null_classifier <- readRDS(here("data", 
#                                 "all_signals_wide_as_of_2020-05-18.RDS")) %>%
#   rename(jhu = `value+0:jhu-csse_confirmed_7dav_incidence_prop`) %>%
#   select(geo_value, time_value, jhu) %>%
#   group_by(geo_value) %>%
#   arrange(time_value) %>%
#   mutate(lag_value = lag(jhu, 7),
#          apc = (jhu - lag_value) / lag_value,
#          hot = case_when(
#            apc > .25 & lag_value > 30 ~ 1,
#            apc <= .25 & lag_value > 30 ~ 0),
#          null_class = mycummean(hot))

hot_udf <- inner_join(
  hotspots_honest,
  up_down %>% select(geo_value, target_end_date, udf))

cutoff <- mean(hot_udf %>% filter(forecaster == "AR") %>% pull(actual))

con_tab <- hot_udf %>% 
  filter(!is.na(actual)) %>%
  mutate(pred = value > .5) %>%
  group_by(forecaster, udf) %>%
  summarise(m = (mean(pred != actual) - cutoff) / cutoff,
            err = mean(pred != actual)) %>%
  ungroup()

# con_tab <- left_join(
#   con_tab %>% filter(forecaster != "AR"),
#   con_tab %>% filter(forecaster == "AR") %>%
#     select(-forecaster) %>%
#     rename(mar = m))

llike_tab <- hot_udf %>%
  filter(!is.na(actual)) %>%
  # left_join(null_classifier %>% select(geo_value, time_value, null_class) %>%
  #             rename(target_end_date = time_value)) %>%
  mutate( # kill prob 0-1 predictions
    value = case_when(
      value < 1e-8 ~ 1e-8,
      value > 1-1e-8 ~ 1-1e-8,
      TRUE ~ value),
    # null_class = case_when(
    #   null_class < 1e-8 ~ 1e-8,
    #   null_class > 1-1e-8 ~ 1-1e-8,
    #   TRUE ~ null_class),
    llike = (actual == 1) * log(value) + (actual == 0) * log(1 - value),
    nulldev = (actual == 1) * log(cutoff) + 
      (actual == 0) * log(1 - cutoff)) %>%
  group_by(forecaster, udf) %>%
  summarise(m = (mean(llike) - mean(nulldev)) / abs(mean(nulldev)),
            mm = mean(llike),
            mmm = mean(nulldev)) %>%
  ungroup()

# llike_tab <- left_join(
#   llike_tab %>% filter(forecaster != "AR"),
#   llike_tab %>% filter(forecaster == "AR") %>%
#     select(-forecaster) %>%
#     rename(mar = m))

bind_rows(con_tab %>% mutate(err = "Classification error"), 
          llike_tab %>% mutate(err = "Log likelihood")) %>%
  ggplot(aes(udf, m , fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  scale_fill_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
  #scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_bw() +
  geom_hline(yintercept = 0) +
  scale_y_continuous(labels = scales::percent_format()) +
  facet_wrap(~err, scales = "free_y") +
  ylab("Improvement relative to the null classifier") +
  xlab("") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

```{r cor-wis-ratio-m1, eval = FALSE, fig.cap="This is Alden's second set of histograms. Here we have the correlation of the absolute value of WIS ratio - 1 with the percent change in 7dav cases relative to 7 days earlier"}
ggplot(pct_chng_df %>% 
         group_by(forecaster) %>% 
         mutate(median = median(cor_abs_wisratio_minus_1_pctchange)), 
       aes(x = cor_abs_wisratio_minus_1_pctchange)) +
  geom_histogram(aes(y = ..count.. / sum(..count..), fill = forecaster), bins = 50) +
  scale_fill_manual(values = fcast_colors) +
  theme_bw() +
  theme(legend.position = "none") +
  geom_vline(aes(xintercept = median), linetype = "dotted") + 
  geom_vline(xintercept = 0) +
  facet_wrap(~forecaster) + 
  xlab("Spearman correlation") + ylab("Relative frequency") +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1))
```

```{r upswing-histogram-logged, eval = FALSE, fig.cap="Not sure if we want this here. Similar to Figure 5 in the manuscript but taking logs. "}
up_down_df %>% 
  group_by(forecaster, udf) %>%
  ggplot(aes((AR_wis + 1) / (wis + 1), fill = forecaster)) +
  geom_histogram(bins = 100) +
  facet_grid(udf ~ forecaster) +
  theme_bw() +
  ylab("Count") +
  theme(legend.position = "none") +
  scale_fill_manual(values = fcast_colors) +
  scale_x_log10() +
  scale_y_log10(breaks = c(10,1000,100000), 
                labels = trans_format("log10", math_format(10^.x))) +
  xlab("AR WIS - forecaster WIS") +
  geom_vline(xintercept = 1)
```

```{r leading-and-lagging, eval=FALSE}
# Deprecated
df2 %>%
  group_by(forecaster, udf) %>%
  summarise(
    leadingness = cor(AR_wis - wis, leading, use = "pairwise.complete.obs"),
    laggingness = cor(AR_wis - wis, lagging, use = "pairwise.complete.obs")) %>%
  pivot_longer(leadingness:laggingness) %>%
  ggplot(aes(udf, value, fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  geom_hline(yintercept = 0) +
  scale_fill_manual(values = fcast_colors) +
  facet_wrap(~name) +
  theme_bw() +
  ylab("Correlation") +
  xlab("") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

```{r fcast-gs-locations, eval=FALSE}
plotter(fcasts_gs %>% filter(period != "jm"),
        "wis", Mean, scaler = "strawman_wis", 
        order_of_operations = c("aggr","scale")) +
  ylab("Mean WIS (relative to baseline)")
```

```{r fcast-gs-locations-adjusted, eval=FALSE}
plotter(fcasts_gs %>% filter(period != "jm"), 
        "wis", GeoMean, scaler = "strawman_wis", 
        order_of_operations = c("scale","aggr")) +
  ylab("Geometric mean WIS (relative to baseline)")
```

```{r hot-gs-locations, eval=FALSE}
plotter_hotspots(hotspots_gs %>% filter(period != "jm")) +
  geom_hline(yintercept = 0.5)
```

```{r traj-data, eval = FALSE}
source(here("code", "deprecated", "trajectory_plot_funs.R"))
preds <- readRDS(here("data", "predictions_honest.RDS"))
traj_best <- get_trajectory_plots(fcasts_honest, preds, actuals, hrr_tab, 
                                  "only2020", "best")
traj_worst <- get_trajectory_plots(fcasts_honest, preds, actuals, hrr_tab,
                                   "only2020", "worst")
rm(preds)
```

```{r trajectory-plots, eval = FALSE}
for (nm in names(traj_best)) print(trajectory_panel(nm, traj_best, traj_worst))
```

### 2021-10-31 forecast/paper/supplement.Rmd
recompile supplement with Ryan's revisions.

@@
   arrange(forecast_date) %>%
-  mutate(`Cumulative Mean` = cummean(mw) / cummean(smw),
-         `Cumulative Sum` = cumsum(mw) / cumsum(smw),
+  mutate(`Cumulative mean` = cummean(mw) / cummean(smw),
+         `Cumulative sum` = cumsum(mw) / cumsum(smw),
          `Cumulative Geo Mean` = exp(cummean(log(gmw))),

### 2021-10-31 forecast/paper/supplement.tex
recompile supplement with Ryan's revisions.

--- 2021-10-09 forecast/paper/supplement.tex
@@
 
-\caption{Relative forecast WIS with vintage compared to finalized data. Using finalized data leads to overly optimistic performance.}\label{fig:fcast-honest-v-finalized}
+\caption{Forecast performance with vintage compared to finalized data. Using finalized data leads to overly optimistic performance.}\label{fig:fcast-honest-v-finalized}
 \end{figure}
@@
 
-\caption{Relative AUC with vintage compared to finalized data. Using finalized data leads to overly optimistic hotspot performance.}\label{fig:hot-honest-v-finalized}
+\caption{Hotspot prediction performance with vintage compared to finalized data. Using finalized data leads to overly optimistic performance.}\label{fig:hot-honest-v-finalized}
 \end{figure}
@@
 
-\caption{Weighted interval score appears to more closely resemble a log-Gaussian distribution.}\label{fig:wis-densities}
+\caption{WIS values from forecast models, which appear to be roughly log-Gaussian.}\label{fig:wis-densities}
 \end{figure}
@@
 
-\caption{Relative forecast performance using vintage data and summarizing with the more robust geometric mean.}\label{fig:fcast-adjusted}
+\caption{Forecast performance (using vintage data), summarized by geometric mean.}\label{fig:fcast-adjusted}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/compare-states-to-hub}
+
+}
+
+\caption{Forecast performance for AR and indicators models, each retrained at the state level, compared to models submitted to the COVID-19 Forecast Hub over the same period.
+The thin grey lines individual models from the Hub; the blue line is the Hub ensemble model.
+(Note that, to align prediction dates as best as possible, we look at the AR and indicator model forecasts for 5, 12, 19, and 26 days ahead; this roughly corresponds to 1, 2, 3, and 4 weeks ahead, respectively, since in the Hub, models typically submit forecasts on a Tuesday for the epiweeks aligned to end on each of the following 4 Saturdays.)}\label{fig:compare-to-hub}
 \end{figure}
@@
 
-\caption{P-values for a sign test that the WIS of the AR forecaster is smaller than that of the indicator-assisted model. Each P-value corresponds a particular forecast date.}\label{fig:sign-test}
+\caption{P-values from a one-sided sign test for equality of forecast error distributions.
+Each P-value corresponds a forecast date.
+The alternative hypothesis is that the the AR model is better (median difference between the relative WIS of the AR model and an indicator model is negative).}\label{fig:sign-test}
 \end{figure}
@@
 
-\caption{\label{tab:dm-test}$P$-values for a one-sided Diebold-Mariano test for improvement in forecast error.
-The test is for the null hypothesis of equal performance against the alternative that the indicator assisted model is better.}
+\caption{\label{tab:dm-test}P-values from a one-sided Diebold-Mariano test for equality of forecast error distributions. The alternative hypothesis is that the AR model is better.}
 \centering
@@
 \midrule
-Geometric Mean Relative WIS & 0.116 & 0.067 & 0.000 & 0.059 & 0.008\\
-Mean Relative WIS & 0.159 & 0.111 & 0.002 & 0.074 & 0.064\\
+Geometric mean & 0.072 & 0.036 & 0.005 & 0.032 & 0.026\\
+Mean & 0.177 & 0.132 & 0.006 & 0.092 & 0.103\\
 \bottomrule
@@
 
-\caption{Average difference between the WIS of the AR model and the WIS of the other forecasters. The indicator-assisted forecasters do best during down and flat periods.}\label{fig:upswing-summary}
+\caption{Average difference between the WIS of the AR model and of the indicator models, separated into up, down, and flat periods.
+The indicator models generally do best during down and flat periods.}\label{fig:upswing-summary}
 \end{figure}
@@
 
-\caption{Classification and loglikelihood separated into periods of upswing, downswing, and flat cases.
-Like the analysis of the forecasting task in the main paper (see Figure 7), relative performance is better during down and flat periods.}\label{fig:hotspots-upswing-downswing}
+\caption{Percentage change in classification error and log likelihood, relative that of the AR model, separated into up, down, and flat periods.
+Like the analogous forecasting analysis, the indicator models generally do better during down and flat periods.}\label{fig:hotspots-upswing-downswing}
 \end{figure}
 
-\clearpage
+\textbackslash begin\{figure\}
 
-\begin{figure}
+\{\centering \includegraphics[width=\textwidth]{fig/cor-wis-ratio-1}
 
-{\centering \includegraphics[width=\textwidth]{fig/cor-wis-ratio-1}
+\}
 
-}
-
-\caption{Histograms of the Spearman correlation between the ratio of AR to AR WIS with the percent change in smoothed case rates relative to 7 days earlier.}\label{fig:cor-wis-ratio}
-\end{figure}
-
-\clearpage
+\textbackslash caption\{Histograms of the Spearman correlation between
+the ratio of forecaster WIS to AR WIS with the \% change in case rates,
+relative to case rates 7 days earlier.\}\label{fig:cor-wis-ratio}
+\textbackslash end\{figure\}
 
@@
 
-\caption{Correlation of the difference in WIS with the difference in median predictions for the AR model relative to the indicator-assisted forecaster.
-In down periods, improvements in forecast risk are highlycorrelated with lower median predictions.
-The opposite is true in up periods.
-This suggests, as one might expect that improved performance of the indicator-assisted model is attributable to being closer to the truth then the AR model.
-This conclusion is stronger in down periods then in up periods.}\label{fig:upswing-corr-table}
+\caption{Correlation of the difference in WIS with the difference in median predictions (each difference being between the AR model and an indicator model), separated into up, down, and flat periods.
+In down periods, improvements in forecast error are highly correlated with lower median predictions.
+The opposite is true in up periods, but the conclusion here appears to be weaker.}\label{fig:upswing-corr-table}
 \end{figure}
@@
 
-\caption{Percent change in average WIS of the forecaster (AR or indicator assisted) relative to the baseline.
+\caption{Percentage change in average WIS of the forecaster (AR or indicator assisted), relative to the baseline.
 All models perform poorly durning down periods, but the indicators help.
@@
 The orange dashed lines indicate the 95\% significance threshold.
-By our leadingness/laggingness metric, DV-CLI is leading (but notlagging) cases over this period.}\label{fig:ccf-dv-finalized}
+By our leadingness/laggingness metric, DV-CLI is leading (but not lagging) cases over this period.}\label{fig:ccf-dv-finalized}
 \end{figure}
@@
 
-\caption{This figure shows the cumulative sum of WIS for each forecaster divided by the cumulative sum of WIS for the baseline model.
-The shaded background shows total U.S.
-cases as reported by JHU-CSSE for the 14-day ahead target.
-Hashes along the $x$-axis denote weeks.}\label{fig:cumulative-mean}
+\caption{Cumulative sum of WIS for each forecaster divided by the cumulative sum of WIS for the baseline model.
+The shaded background shows national case incidence for the 14-day ahead target.
+Hashes along the x-axis denote weeks.}\label{fig:cumulative-mean}
 \end{figure}
@@
 \clearpage
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{fig/cumulative-geo-mean-1}
-
-}
-
-\caption{This figure shows the cumulative geometric mean of WIS for each forecaster divided by the WIS for the baseline model.
-The shaded background shows total U.S.
-cases as reported by JHU-CSSE for the 14-day ahead target.
-Hashes along the $x$-axis denote weeks.}\label{fig:cumulative-geo-mean}
-\end{figure}
 
@@
 
-\caption{Percent improvement in WIS relative to the AR forecaster by region (negative numbers indicate improved performance, positives indicate worsening).}\label{fig:errs-in-space}
+\caption{Percentage improvement in WIS, relative to the AR forecaster, by HRR (negative numbers indicate improved performance, positives indicate worsening).}\label{fig:errs-in-space}
 \end{figure}
-
-\clearpage
-
-\clearpage
 

### 2021-10-31 forecast/code/figures/compare-to-hub.R

@@ -7,30 +7,24 @@ Sys.setenv("AWS_DEFAULT_REGION" = "us-east-2")
 s3bucket <- get_bucket("forecast-eval")
 n_keeper_weeks <- 6L
 n_keeper_locs <- 50L
-case_scores <- s3readRDS("score_cards_state_cases.rds", s3bucket) 
-# case_preds <- s3readRDS("predictions_cards.rds", s3bucket) %>%
-#   filter(signal == "confirmed_incidence_num",
-#          forecast_date < "2021-01-01",
-#          !(geo_value %in% c("as", "gu", "pr", "vi", "mp", "us"))) 
-# actuals <- covidcast::covidcast_signal("jhu-csse", "confirmed_7dav_incidence_num",
-#                                        geo_type = "state", as_of = "2021-05-18",
-#                                        end_day = "2021-03-01") %>%
-#   mutate(value = value * 7) %>%
-#   rename(target_end_date = time_value, actual = value)
-# case_scores <- evalcast::evaluate_predictions(
-#   case_preds, actuals, 
-#   err_measures = list(wis = evalcast::weighted_interval_score),
-#   grp_vars = c("ahead", "forecaster", "forecast_date", "geo_value")
-# )
-# rm(case_preds)
 
-### RJT: using evalcast above leads to the following error for me. I even made 
-### sure that I'm running the latest versions of evalcast and covidcast ...
-### Error in max(target_response$end) + backfill_buffer : 
-### non-numeric argument to binary operator
-### In addition: Warning message:
-###  In evaluate_predictions_single_ahead(filter(predictions_cards, .
-###    data$ahead == : Incompatible methods ("+.Date", "Ops.data.frame") for "+"
+case_preds <- s3readRDS("predictions_cards.rds", s3bucket) %>%
+  filter(signal == "confirmed_incidence_num",
+         forecast_date < "2021-01-01",
+         !(geo_value %in% c("as", "gu", "pr", "vi", "mp", "us")))
+actuals <- covidcast::covidcast_signal("jhu-csse", "confirmed_7dav_incidence_num",
+                                       geo_type = "state", as_of = "2021-05-18",
+                                       end_day = "2021-03-01") %>%
+  mutate(value = value * 7) %>%
+  rename(target_end_date = time_value, actual = value) %>%
+  select(geo_value, target_end_date, actual)
+case_scores <- evalcast::evaluate_predictions(
[…]

### 2021-10-31 forecast/paper/supplement-text.tex
Rebuilt with edits from me.

@@
 affected by this distinction, reflecting the latency in claims-based reporting.
-This underscores the importance of efforts to provide ``nowcasts'' for claims
+This highlights the importance of efforts to provide ``nowcasts'' for claims
 signals (which corresponds to a 0-ahead forecast of what the claims signal's
@@
 data from different providers.  The substantial difference in their curves in
-Figure 3 of the main paper must therefore reflect their having very different
+Figure 3 of the main paper must, therefore, reflect their having very different
 backfill profiles.
@@
 models in the main paper)---can be robust and effective, performing
-competitively many of the top models from COVID-19 Forecast Hub, including the
-Hub ensemble model.
+competitively to the top models submitted to the COVID-19 Forecast Hub,
+including the Hub's ensemble model.
 
@@
 
-Therefore, to make a comparison to models in the Hub as direct as possible, we
+Therefore, to make the comparison to models in the Hub as direct as possible, we
 retrained our models over the same forecast period as in the main paper, and
@@
 We then rescaled them post hoc to account for the different temporal resolution
-and the rate versus count scale (first and third points in the above list).  The
+and the rate-versus-count scaling (first and third points in the above list).
+The 
 results are given in Figure \ref{fig:compare-to-hub}.  The evaluation was
@@
 model for smaller ahead values.  The same general conclusion can be drawn for
-the indicators models as well.  However, interestingly, a close inspection
+the indicator-assisted models as well.  However, interestingly, a close inspection
 reveals that the AR model here is for the most part in the ``middle of the

### 2021-10-31 forecast/paper/supplement.Rmd
Rebuilt with edits from me.

@@
 
-```{r compare-to-hub, fig.cap="Forecast performance for AR and indicators models, each retrained at the state level, compared to models submitted to the COVID-19 Forecast Hub over the same period.
-The thin grey lines individual models from the Hub; the blue line is the Hub ensemble model.
-(Note that, to align prediction dates as best as possible, we look at the AR and indicator model forecasts for 5, 12, 19, and 26 days ahead; this roughly corresponds to 1, 2, 3, and 4 weeks ahead, respectively, since in the Hub, models typically submit forecasts on a Tuesday for the epiweeks aligned to end on each of the following 4 Saturdays.)"}
+```{r compare-to-hub, fig.cap="Forecast performance for AR and indicator models, each retrained at the state level, compared to models submitted to the COVID-19 Forecast Hub over the same period.
+The thin grey lines are individual models from the Hub; the blue line is the Hub ensemble model.
+(To align prediction dates as best as possible, we look at the AR and indicator model forecasts for 5, 12, 19, and 26 days ahead; this roughly corresponds to 1, 2, 3, and 4 weeks ahead, respectively, since in the Hub, models typically submit forecasts on a Tuesday for the epiweeks aligned to end on each of the following 4 Saturdays.)"}
 knitr::include_graphics("fig/compare-states-to-hub.pdf")

### 2021-10-31 forecast/paper/supplement.tex
Rebuilt with edits from me.

@@
 
-\caption{Forecast performance for AR and indicators models, each retrained at the state level, compared to models submitted to the COVID-19 Forecast Hub over the same period.
-The thin grey lines individual models from the Hub; the blue line is the Hub ensemble model.
-(Note that, to align prediction dates as best as possible, we look at the AR and indicator model forecasts for 5, 12, 19, and 26 days ahead; this roughly corresponds to 1, 2, 3, and 4 weeks ahead, respectively, since in the Hub, models typically submit forecasts on a Tuesday for the epiweeks aligned to end on each of the following 4 Saturdays.)}\label{fig:compare-to-hub}
+\caption{Forecast performance for AR and indicator models, each retrained at the state level, compared to models submitted to the COVID-19 Forecast Hub over the same period.
+The thin grey lines are individual models from the Hub; the blue line is the Hub ensemble model.
+(To align prediction dates as best as possible, we look at the AR and indicator model forecasts for 5, 12, 19, and 26 days ahead; this roughly corresponds to 1, 2, 3, and 4 weeks ahead, respectively, since in the Hub, models typically submit forecasts on a Tuesday for the epiweeks aligned to end on each of the following 4 Saturdays.)}\label{fig:compare-to-hub}
 \end{figure}

### 2021-10-31 forecast/code/figures/compare-to-hub.R
Proofreading done. Revision 2 complete.

@@ -123,7 +123,7 @@ ggplot(all_time_performance %>% filter(source == "ours"),
   geom_line(aes(ahead, value, color = forecaster)) +
   geom_point(aes(ahead, value, color = forecaster)) +
   scale_color_manual(values = c(fcast_colors, "Ensemble" = "lightblue"),
-                     guide = guide_legend(nrow = 1)) +
+                     guide = guide_legend(nrow = 2)) +
   ylab("WIS (relative to baseline)") +
   geom_hline(yintercept = 1, linetype = "dashed") +
   xlab("Days ahead") +

### 2021-10-31 forecast/paper/supplement.Rmd

@@
 ```{r compare-to-hub, fig.cap="Forecast performance for AR and indicator models, each retrained at the state level, compared to models submitted to the COVID-19 Forecast Hub over the same period.
-The thin grey lines are individual models from the Hub; the blue line is the Hub ensemble model.
-(To align prediction dates as best as possible, we look at the AR and indicator model forecasts for 5, 12, 19, and 26 days ahead; this roughly corresponds to 1, 2, 3, and 4 weeks ahead, respectively, since in the Hub, models typically submit forecasts on a Tuesday for the epiweeks aligned to end on each of the following 4 Saturdays.)"}
+The thin grey lines are individual models from the Hub; the light blue line is the Hub ensemble model.
+To align prediction dates as best as possible, we look at the AR and indicator model forecasts for 5, 12, 19, and 26 days ahead; this roughly corresponds to 1, 2, 3, and 4 weeks ahead, respectively, since in the Hub, models typically submit forecasts on a Tuesday for the epiweeks aligned to end on each of the following 4 Saturdays."}
 knitr::include_graphics("fig/compare-states-to-hub.pdf")
@@
 
-```{r sign-test, fig.cap="P-values from a one-sided sign test for equality of forecast error distributions.
-Each P-value corresponds a forecast date.
-The alternative hypothesis is that the the AR model is better (median difference between the relative WIS of the AR model and an indicator model is negative)."}
+```{r sign-test, fig.cap="P-values from a one-sided sign test for improved forecast performance of the indicator-assisted models.
+Each p-value corresponds to a forecast date.
+The alternative hypothesis is that the AR model is better (median difference between the relative WIS of the AR model and an indicator model is negative)."}
 fcast_colors2 <- fcast_colors[names(fcast_colors) != "AR"]
@@
     pivot_wider(names_from = forecaster, values_from = dm), 
-  digits = 3, booktabs = TRUE, caption = "P-values from a one-sided Diebold-Mariano test for equality of forecast error distributions. The alternative hypothesis is that the AR model is better.")
+  digits = 3, booktabs = TRUE, caption = "P-values from a one-sided Diebold-Mariano test for improvemed forecast performance when adding the indicators.
+The alternative hypothesis is that the AR model is better.")
 ```
@@
 
-```{r cor-wis-ratio, fig.cap="Histograms of the Spearman correlation between the ratio of forecaster WIS to AR WIS with the % change in case rates, relative to case rates 7 days earlier."}
+```{r cor-wis-ratio, fig.cap="Histograms of the Spearman correlation between the ratio of forecaster WIS to AR WIS with the \\% change in case rates, relative to case rates 7 days earlier."}
 ggplot(pct_chng_df %>% 
@@
 ```{r upswing-summary-remake, fig.cap="Percentage change in average WIS of the forecaster (AR or indicator assisted), relative to the baseline.
-All models perform poorly durning down periods, but the indicators help.
+All models perform poorly during down periods, but the indicators help.
 During flat periods, the indicators improve slightly over the AR.
@@
 The right panel shows $\\CCF_{\\ell}(a)$ for different values of $a$ as vertical blue bars.
-The orange dashed lines indicate the 95\\% significance threshold.
+The orange dashed line indicates the 95\\% significance threshold.
 By our leadingness/laggingness metric, DV-CLI is leading (but not lagging) cases over this period."}
@@
 The shaded background shows national case incidence for the 14-day ahead target.
-Hashes along the x-axis denote weeks." }
+Hash marks along the x-axis denote weeks." }
 cumulatives <- fcasts_honest %>%

### 2021-10-31 forecast/paper/supplement.tex

@@
 \caption{Forecast performance for AR and indicator models, each retrained at the state level, compared to models submitted to the COVID-19 Forecast Hub over the same period.
-The thin grey lines are individual models from the Hub; the blue line is the Hub ensemble model.
-(To align prediction dates as best as possible, we look at the AR and indicator model forecasts for 5, 12, 19, and 26 days ahead; this roughly corresponds to 1, 2, 3, and 4 weeks ahead, respectively, since in the Hub, models typically submit forecasts on a Tuesday for the epiweeks aligned to end on each of the following 4 Saturdays.)}\label{fig:compare-to-hub}
+The thin grey lines are individual models from the Hub; the light blue line is the Hub ensemble model.
+To align prediction dates as best as possible, we look at the AR and indicator model forecasts for 5, 12, 19, and 26 days ahead; this roughly corresponds to 1, 2, 3, and 4 weeks ahead, respectively, since in the Hub, models typically submit forecasts on a Tuesday for the epiweeks aligned to end on each of the following 4 Saturdays.}\label{fig:compare-to-hub}
 \end{figure}
@@
 
-\caption{P-values from a one-sided sign test for equality of forecast error distributions.
-Each P-value corresponds a forecast date.
-The alternative hypothesis is that the the AR model is better (median difference between the relative WIS of the AR model and an indicator model is negative).}\label{fig:sign-test}
+\caption{P-values from a one-sided sign test for improved forecast performance of the indicator-assisted models.
+Each p-value corresponds to a forecast date.
+The alternative hypothesis is that the AR model is better (median difference between the relative WIS of the AR model and an indicator model is negative).}\label{fig:sign-test}
 \end{figure}
@@
 
-\caption{\label{tab:dm-test}P-values from a one-sided Diebold-Mariano test for equality of forecast error distributions. The alternative hypothesis is that the AR model is better.}
+\caption{\label{tab:dm-test}P-values from a one-sided Diebold-Mariano test for improvemed forecast performance when adding the indicators. The alternative hypothesis is that the AR model is better.}
 \centering
@@
 
-\textbackslash begin\{figure\}
+\begin{figure}
 
-\{\centering \includegraphics[width=\textwidth]{fig/cor-wis-ratio-1}
+{\centering \includegraphics[width=\textwidth]{fig/cor-wis-ratio-1}
 
-\}
+}
 
-\textbackslash caption\{Histograms of the Spearman correlation between
-the ratio of forecaster WIS to AR WIS with the \% change in case rates,
-relative to case rates 7 days earlier.\}\label{fig:cor-wis-ratio}
-\textbackslash end\{figure\}
+\caption{Histograms of the Spearman correlation between the ratio of forecaster WIS to AR WIS with the \% change in case rates, relative to case rates 7 days earlier.}\label{fig:cor-wis-ratio}
+\end{figure}
 
@@
 \caption{Percentage change in average WIS of the forecaster (AR or indicator assisted), relative to the baseline.
-All models perform poorly durning down periods, but the indicators help.
+All models perform poorly during down periods, but the indicators help.
 During flat periods, the indicators improve slightly over the AR.
@@
 The right panel shows $\CCF_{\ell}(a)$ for different values of $a$ as vertical blue bars.
-The orange dashed lines indicate the 95\% significance threshold.
+The orange dashed line indicates the 95\% significance threshold.
 By our leadingness/laggingness metric, DV-CLI is leading (but not lagging) cases over this period.}\label{fig:ccf-dv-finalized}
@@
 The shaded background shows national case incidence for the 14-day ahead target.
-Hashes along the x-axis denote weeks.}\label{fig:cumulative-mean}
+Hash marks along the x-axis denote weeks.}\label{fig:cumulative-mean}
 \end{figure}

### 2021-11-03 indicators/code/indicator-paper-figures.Rmd

@@
 
-case_max <- max(cs[[1]]$value)
+Min <- function(x) min(x, na.rm = TRUE)
+Max <- function(x) max(x, na.rm = TRUE)
+case_min <- Min(cs[[1]]$value)
+case_max <- Max(cs[[1]]$value)
 
@@
   group_by(signal) %>%
-  mutate(value = value / max(value, na.rm = TRUE) * case_max) %>%
+    mutate(value = (value - Min(value)) / (Max(value) - Min(value)) * 
+             case_max + case_min) %>%
   ungroup()

### 2021-11-03 indicators/paper/supplement.Rmd
Rebuild again

@@
 
+Min <- function(x) min(x, na.rm = TRUE)
+Max <- function(x) max(x, na.rm = TRUE)
+
 # geo_labeler is a function from geo_value to human-readable name
 get_trend_grid <- function(trends, geo_labeler) {
-  case_max <- max(trends[[1]]$value)
+  case_min <- Min(trends[[1]]$value)
+  case_max <- Max(trends[[1]]$value)
 
@@
     group_by(signal) %>%
-    mutate(value = value / max(value, na.rm = TRUE) * case_max) %>%
+    mutate(value = (value - Min(value)) / (Max(value) - Min(value)) * 
+             case_max + case_min) %>%
     ungroup() %>%
@@
 Cases are displayed on the rate scale: counts per 100,000 people.
-Other signals are scaled to have the same global maximum across all counties and times, so they can be presented in the same range.
+Other signals are scaled to have the same global range across all counties and times.
 (Part 1 of 4.)", fig.height=8}
@@
 (HHS data was not consistently reported before August 2020; furthermore, as with cases in the previous trend plots, the HHS data has been smoothed using a 7-day trailing average.) Hospitalizations are displayed on the rate scale: counts per 100,000 people.
-Other signals are scaled to have the same maximum value.
+Other signals are scaled to have the same range.
 HSP-Hosp is the percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners (smoothed in time and adjusted for systematic day-of-week effects)."}
@@
 
-hosp_main_signal_max <- max(hosp_signal_data[[1]]$value, na.rm=TRUE)
+hosp_main_signal_min <- Min(hosp_signal_data[[1]]$value)
+hosp_main_signal_max <- Max(hosp_signal_data[[1]]$value)
 
@@
   group_by(signal) %>%
-  mutate(value = value / max(value, na.rm=TRUE) * hosp_main_signal_max) %>%
+    mutate(value = (value - Min(value)) / (Max(value) - Min(value)) * 
+             hosp_main_signal_max + hosp_main_signal_min) %>%
   ungroup()

### 2021-11-03 indicators/paper/supplement.tex

@@
 Cases are displayed on the rate scale: counts per 100,000 people.
-Other signals are scaled to have the same global maximum across all counties and times, so they can be presented in the same range.
+Other signals are scaled to have the same global range across all counties and times.
 (Part 1 of 4.)}\label{fig:state-trend-grids-1}
@@
 Cases are displayed on the rate scale: counts per 100,000 people.
-Other signals are scaled to have the same global maximum across all counties and times, so they can be presented in the same range.
+Other signals are scaled to have the same global range across all counties and times.
 (Part 1 of 4.)}\label{fig:county-trend-grids-1}
@@
 (HHS data was not consistently reported before August 2020; furthermore, as with cases in the previous trend plots, the HHS data has been smoothed using a 7-day trailing average.) Hospitalizations are displayed on the rate scale: counts per 100,000 people.
-Other signals are scaled to have the same maximum value.
+Other signals are scaled to have the same range.
 HSP-Hosp is the percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners (smoothed in time and adjusted for systematic day-of-week effects).}\label{fig:hospitalization_time_trends_national}

### 2021-11-04 cover/cover.Rmd
New idea

@@
 
-```{r get-us, eval = FALSE}
+```{r get-data, eval = FALSE}
 sources <- c("jhu-csse", 
              "doctor-visits", 
-             "chng", 
              "fb-survey", 
-             "safegraph",
-             "safegraph")
+             "google-symptoms")
 
@@
              "smoothed_adj_cli",
-             "smoothed_adj_outpatient_covid",
              "smoothed_whh_cmnty_cli",
-             "bars_visit_prop",
-             "restaurants_visit_prop")
+             "sum_anosmia_ageusia_smoothed_search")
 
 start <- "2020-04-15"
-end <- "2021-04-15"
+end <- "2021-10-25"
 
@@
 
-us <- covidcast_signals(
-  sources, signals,
-  start, end, 
-  geo_type = "nation",
-) %>% map(function(df) df %>% 
-            Select(data_source, signal, geo_value, time_value, value))
-```
-
-```{r get-states, eval = FALSE}
-state <- covidcast_signals(
+x <- covidcast_signals(
   sources, signals,
@@
             Select(data_source, signal, geo_value, time_value, value))
-```
-
-```{r get-counties, eval = FALSE}
-top_counties <- county_census %>%
-  filter(COUNTY != 0) %>% # these are states
-  arrange(desc(POPESTIMATE2019)) %>%
-  head(n = 50) %>%
-  pull(FIPS)
-
-# Work around for an API problem: specifying geo_values in the API call fails on
-# the CHNG data. So we filter after the fact instead. DANGER: Very slow.
-county_nochng <- covidcast_signals(
-  sources[sources != "chng"], signals[sources != "chng"],
-  start, end, 
-  geo_type = "county", geo_values = top_counties
-) %>% map(function(df) df %>% 
-            Select(data_source, signal, geo_value, time_value, value))
-
-county_chng <- covidcast_signals(
-  sources[sources == "chng"], signals[sources == "chng"],
-  start, end, 
-  geo_type = "county", 
-) %>% map(function(df) df %>% 
-            Select(data_source, signal, geo_value, time_value, value) %>%
-            filter(geo_value %in% top_counties))
-
-county <- append(county_nochng, county_chng)
 ```
@@
 
-```{r build-sg}
-# SafeGraph national signals don't exist for all time, so build them from states
-y = state_census %>% 
-  mutate(geo_value = tolower(ABBR)) %>% 
-  rename(w = POPESTIMATE2019) %>%
-  select(geo_value, w)
+```{r trend-plots}
+names <- c("Reported COVID cases", 
+           "COVID symptoms in medical claims", 
+           "Self-reported COVID symptoms in surveys",
+           "Searches for COVID symptoms on Google")
+colors <- c("#FFFFFF", "#E69F00", "#56B4E9", "#CC79A7")
 
-for (i in which(sources == "safegraph")) {  
-  us[[i]] <- state[[i]] %>% 
-    left_join(y, by = "geo_value") %>% 
-    group_by(data_source, signal, time_value) %>%
-    summarize(value = Sum(w * value) / Sum(w)) %>%
-    mutate(geo_value = "us") %>%
-    ungroup()
+g <- vector(mode = "list", length = length(x))
+delta <- 1/500
+alpha <- 0.35
+ylim <- c(-0.05, 1.05 + 50 * delta)
+
+for (i in 1:length(x)) {
+  g[[i]] <- x[[i]] %>%
+    filter(geo_value %in% tolower(state.abb)) %>%
+    mutate(geo_num = as.integer(as.factor(geo_value))) %>%
+    group_by(geo_value) %>%
+    mutate(value = (value - Min(value)) / (Max(value) - Min(value)) + 
+             delta * (geo_num - 1)) %>%
+    ungroup() %>%
+    ggplot(aes(x = time_value, y = value, group = geo_value)) +
+    labs(y = names[i]) + coord_cartesian(ylim = ylim) +
+    geom_line(alpha = 0.5, color = colors[i]) + theme_void() +
+    theme(panel.background = element_rect(fill = "black"),
+          axis.title.y = element_text(color = colors[i], angle = 90,
+                                      margin = margin(l = 10, r = -15)))
 }
+
+pdf(file = "cover-graphic.pdf", width = 15, height = 15)
+do.call(gridExtra::marrangeGrob, list(g, nrow = 5, ncol = 1))
+graphics.off()
 ```
-
-```{r smooth-sg}
-# SafeGraph signals at all levels aren't smoothed, so we use 7-day averaging
-for (i in which(sources == "safegraph")) {  
-  us[[i]] <- us[[i]] %>% 
-    as.epi_tibble() %>% 
-    group_by(geo_value) %>%
-    epi_slide(~ Mean(.x$value), n = 7, new_col_name = "value") %>%
-    ungroup()
-
-  state[[i]] <- state[[i]] %>% 
-    as.epi_tibble() %>% 
-    group_by(geo_value) %>%
-    epi_slide(~ Mean(.x$value), n = 7, new_col_name = "value") %>%
-    ungroup()
-
-  county[[i]] <- county[[i]] %>% 
-    as.epi_tibble() %>% 
-    group_by(geo_value) %>%
-    epi_slide(~ Mean(.x$value), n = 7, new_col_name = "value") %>%
-    ungroup() 
-}
-```
-
-```{r aggregate-data}
-all <- do.call(rbind, c(us, state, county))
-```
-
-```{r plotting-functions}
-names <- c("Cases", 
-           "DV-CLI", 
-           "CHNG-COVID", 
-           "CTIS-CLI-in-community",
-           "SG-Bars", 
-           "SG-Restaurants")
-
-colors <- c("black", scales::hue_pal()(length(signals)-1))
-
-get_trend_plot <- function(df) {
-  case_min <- Min(df %>% filter(signal == signals[1]) %>% pull(value))
-  case_max <- Max(df %>% filter(signal == signals[1]) %>% pull(value))
-
-  df %>%
-    group_by(signal, geo_value) %>%
-    mutate(value = (value - Min(value)) / (Max(value) - Min(value)) * 
-             case_max + case_min) %>%
-    ungroup() %>%
-    mutate(signal = factor(signal, levels = signals, labels = names)) %>%
-    ggplot(aes(x = time_value, y = value, color = signal)) +
-    geom_line() + scale_color_manual(breaks = names, values = colors) +
-    facet_wrap(vars(geo_value), scales = "free_y") + 
-    labs(x = "Date", y = "Signal value (scaled)") +
-    theme_bw() + theme(legend.pos = "bottom", legend.title = element_blank())
-}
-```
-
-```{r plot-it, fig.width = 10, fig.height = 4}
-all %>%
-  filter(signal != signals[5]) %>%
-  filter(geo_value %in% c("us", "tx", "48113")) %>%
-  filter(between(time_value, as.Date("2020-06-01"), as.Date("2020-07-15"))) %>%
-  get_trend_plot()
-```

### 2021-11-04 forecast/paper/paper.tex
minor data deposit revision issues

@@
 companion paper \cite{Reinhart:2021} describes our research group's efforts,
-beginning in April 2020, in curating, and maintaining a database of real-time
+beginning in April 2020, in curating and maintaining a database of real-time
 indicators that track COVID-19 activity and other relevant phenomena. The
@@
 Similarly, some authors, e.g., \cite{ioannidis2020a}, have pointed out numerous
-mishaps of forecasting during the pandemic and it is not our general intent to
+mishaps of forecasting during the pandemic, and it is not our general intent to
 fix them here. Instead, we start with a basic and yet reasonably effective
@@
 containing Los Angeles, which has more than 10 million people), generally HRRs  
-are much more homogenous in size than the (approximately) 3200 U.S. counties,   
+are much more homogenous in size than the (approximately) 3200 U.S.\ counties,   
 and they serve as a nice middle ground in between counties and states.
@@
 
-Below is summary of the high-level conclusions.
+Below is a summary of the high-level conclusions.
 

### 2021-11-04 cover/cover.Rmd
Some tweaks

@@
 
-start <- "2020-04-15"
-end <- "2021-10-25"
+start <- as.Date("2020-04-15")
+end <- as.Date("2021-10-22")
 
@@
 alpha <- 0.35
+xlim <- c(start, end)
 ylim <- c(-0.05, 1.05 + 50 * delta)
+xp <- xlim[1] - 15
+yp <- mean(ylim)
 
@@
     ggplot(aes(x = time_value, y = value, group = geo_value)) +
-    labs(y = names[i]) + coord_cartesian(ylim = ylim) +
+    labs(y = names[i]) + coord_cartesian(xlim = xlim, ylim = ylim) +
     geom_line(alpha = 0.5, color = colors[i]) + theme_void() +
-    theme(panel.background = element_rect(fill = "black"),
-          axis.title.y = element_text(color = colors[i], angle = 90,
-                                      margin = margin(l = 10, r = -15)))
+    annotate("text", x = xp, y = yp, label = names[i], 
+             color = colors[i], angle = 90) +
+    theme(panel.background = element_rect(fill = "black"))
+          # axis.title.y = element_text(color = colors[i], angle = 90,
+          #                             margin = margin(l = 10, r = -15)))
 }
 
-pdf(file = "cover-graphic.pdf", width = 15, height = 15)
-do.call(gridExtra::marrangeGrob, list(g, nrow = 5, ncol = 1))
-graphics.off()
+gg <- do.call(gridExtra::arrangeGrob, list(grobs = g, ncol = 1))
+ggsave(file = "cover-graphic.pdf", gg, height = 12, width = 15)
 ```

### 2021-11-04 cover/cover.Rmd

--- 2021-11-04 cover/cover.Rmd
@@
 names <- c("Reported COVID cases",
-           "COVID symptoms in medical claims", 
-           "Self-reported COVID symptoms in surveys",
-           "Searches for COVID symptoms on Google")
+           "COVID symptoms\nin medical claims",
+           "Self-reported COVID\nsymptoms in surveys",
+           "Searches for COVID\nsymptoms on Google")
 colors <- c("#FFFFFF", "#E69F00", "#56B4E9", "#CC79A7")
+
+dates <- seq.Date(start, end, by = "1 day")
+n_days <- length(dates)
 
@@
 alpha <- 0.35
+xlim <- c(start - 10, end)
 ylim <- c(-0.05, 1.05 + 50 * delta)
+xp <- xlim[1] - 8
+yp <- mean(ylim)
 
 for (i in 1:length(x)) {
+  y <- tibble(geo_value = rep(unique(x[[i]]$geo_value), each = n_days),
+              time_value = rep(seq.Date(start, end, by = "day"),
+                               length(unique(x[[i]]$geo_value))))
+  x[[i]] <- full_join(x[[i]], y, by = c("geo_value", "time_value"))
+
   g[[i]] <- x[[i]] %>%
@@
     ggplot(aes(x = time_value, y = value, group = geo_value)) +
-    labs(y = names[i]) + coord_cartesian(ylim = ylim) +
-    geom_line(alpha = 0.5, color = colors[i]) + theme_void() +
-    theme(panel.background = element_rect(fill = "black"),
-          axis.title.y = element_text(color = colors[i], angle = 90,
-                                      margin = margin(l = 10, r = -15)))
+    labs(y = names[i]) + coord_cartesian(xlim = xlim, ylim = ylim) +
+    geom_line(alpha = 0.5, color = colors[i], size = 0.3) +
+    theme_void() +
+    annotate("text", x = xp, y = yp, label = names[i],
+             color = colors[i], angle = 90, size = 3) +
+    theme(panel.background = element_rect(fill = "black"))
 }
 
-pdf(file = "cover-graphic.pdf", width = 15, height = 15)
-do.call(gridExtra::marrangeGrob, list(g, nrow = 5, ncol = 1))
-graphics.off()
+gg <- do.call(gridExtra::arrangeGrob, list(grobs = g, ncol = 1))
+# PNAS wants 21 cm wide by 22.5 cm high
+ggsave(file = "cover-graphic.pdf", gg, height = 22.5 / 2.54, width = 21 / 2.54)
 ```

### 2021-11-04 cover/cover.Rmd
Add Astley paragraph

@@
 library(covidcast)
-library(epitools)
 library(ggplot2)
@@
 ```{r get-data, eval = FALSE}
-sources <- c("jhu-csse",
+sources <- c("usa-facts",
              "doctor-visits",
@@
 end <- as.Date("2021-10-22")
-
-Select <- function(x, ...) {
-  if (nrow(x) == 0) x
-  else select(x, ...)
-}
+dates <- seq.Date(start, end, by = "1 day")
+geos <- tolower(state.abb)
 
@@
   start, end,
-  geo_type = "state"
-) %>% map(function(df) df %>%
-            Select(data_source, signal, geo_value, time_value, value))
+  geo_type = "state", 
+  geo_values = geos) %>% 
+  map(function(df) df %>%
+        select(data_source, signal, geo_value, time_value, value))
 ```
@@
 
+```{r load-all}
+# Inspect missingness: google has lots (privacy threshold)
+do.call(rbind, x) %>%
+  filter(geo_value %in% tolower(state.abb)) %>%
+  group_by(signal, geo_value) %>%
+  summarize(nas = length(dates) - n()) %>%
+  ungroup() %>% group_by(signal) %>%
+  filter(nas > 0)
+
+y <- tibble(geo_value = rep(geos, each = length(dates)),
+            time_value = rep(dates, length(geos)))
+
+x[[4]] %>%
+  full_join(y, by = c("geo_value", "time_value")) %>%
+  group_by(time_value) %>%
+  summarize(nas = sum(is.na(value))) %>% 
+  filter(nas > 40)
+```
+
 ```{r trend-plots}
@@
 
-dates <- seq.Date(start, end, by = "1 day")
-n_days <- length(dates)
-
 g <- vector(mode = "list", length = length(x))
-delta <- 1/500
+delta <- 1 / length(geos) * 0.1
 alpha <- 0.35
+line_width <- 0.3
+label_size <- 3
 xlim <- c(start - 10, end)
-ylim <- c(-0.05, 1.05 + 50 * delta)
+ylim <- c(-0.05, 1.05 + length(geos) * delta)
 xp <- xlim[1] - 8
@@
 for (i in 1:length(x)) {
-  y <- tibble(geo_value = rep(unique(x[[i]]$geo_value), each = n_days),
-              time_value = rep(seq.Date(start, end, by = "day"),
-                               length(unique(x[[i]]$geo_value))))
-  x[[i]] <- full_join(x[[i]], y, by = c("geo_value", "time_value"))
-
   g[[i]] <- x[[i]] %>%
-    filter(geo_value %in% tolower(state.abb)) %>%
+    full_join(y, by = c("geo_value", "time_value")) %>%
     mutate(geo_num = as.integer(as.factor(geo_value))) %>%
@@
     labs(y = names[i]) + coord_cartesian(xlim = xlim, ylim = ylim) +
-    geom_line(alpha = 0.5, color = colors[i], size = 0.3) +
+    geom_line(alpha = 0.5, color = colors[i], size = line_width) +
     theme_void() +
     annotate("text", x = xp, y = yp, label = names[i],
-             color = colors[i], angle = 90, size = 3) +
+             color = colors[i], angle = 90, size = label_size) +
     theme(panel.background = element_rect(fill = "black"))
@@
 gg <- do.call(gridExtra::arrangeGrob, list(grobs = g, ncol = 1))
-# PNAS wants 21 cm wide by 22.5 cm high
-ggsave(file = "cover-graphic.pdf", gg, height = 22.5 / 2.54, width = 21 / 2.54)
+ggsave(file = "cover-graphic.pdf", gg, height = 22.5, width = 21, units = "cm")
 ```

### 2021-11-04 perspective/paper.tex
Order citations by appearance; use numbered style
Affiliations and reordering
Add Astley paragraph

@@
 \documentclass{article}
-\input{ryantibs}
+
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}     % use 8-bit T1 fonts
+\usepackage{booktabs}        % professional-quality tables
+\usepackage{nicefrac}          % compact symbols for 1/2, etc.
+\usepackage{microtype}       % microtypography
+\usepackage{newtxtext}       % times font
+
+\usepackage[numbers]{natbib}
+\usepackage{amssymb,amsmath,amsthm,bbm}
+
+\usepackage[margin=1in]{geometry}
+\usepackage{verbatim,float,url,dsfont}
+\usepackage{graphicx,subfigure,psfrag}
+\usepackage{algorithm,algorithmic}
+\usepackage{mathtools,enumitem}
+\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue,linkcolor=blue]{hyperref}
+\usepackage{multirow}
 
 \title{Epidemic Tracking and Forecasting: Lessons Learned from a Tumultuous Year}
-\author{Roni Rosenfeld \and Ryan J. Tibshirani}
+\author{Roni Rosenfeld$^a$ \and Ryan J. Tibshirani$^{a,b}$}
+\date{$^a$Machine Learning Department, Carnegie Mellon University \\ 
+$^b$Department of Statistics \& Data Science, Carnegie Mellon University}
 
@@
 
-\item \emph{Authors' note to the Editor: this is to be filled out later.  The
-    Astley et al. paper is still undergoing a significant revision and its main
-    messages may change as a result.}
+\item \citet{Astley:2021} focuses on the international version of CTIS, which is
+  an (ongoing) online survey operated by the University of Maryland, again in
+  partnership with Facebook.  This international survey covers over 100
+  countries, and is run in coordination with the U.S. one, so that they bear a
+  similar structure and undergo similar updates; the full data set of
+  individual, anonymized survey responses is again available to individual
+  researchers under a data use agreement.  The paper presents analyses that
+  reflect some basic and important characteristics of the international survey,
+  reflecting its value abroad, where public health reporting efforts may be more
+  limited than those in the U.S.
 \end{enumerate}
@@
 \newpage
-\bibliographystyle{plainnat}
+\bibliographystyle{unsrt}
 \bibliography{../common/covidcast}

### 2021-11-04 perspective/paper.tex
Add acks paragraph

@@
 \begin{enumerate}
-\item \citet{Reinhart:2021} describes our group's (ongoing) effort in building
-  and maintaining COVIDcast: an open repository of real-time, geographically
-  detailed COVID-19 indicators in the U.S.  These indicators (a term we use
-  interchangeably with signals) are derived from a diverse set of data sources:
-  medical testing devices, medical insurance claims, internet search trends,
-  app-based mobility data, and online surveys, among others.  Many indicators
-  are demonstrated to have meaningful statistical relationships with what have
-  become the pandemic's ``topline'' numbers (reported cases, hospitalizations,
-  deaths), whereas others uniquely reflect certain activity (not available in
-  other publicly-available data sources) that may drive or affect the spread of
-  COVID-19.  The paper demonstrates through a number of examples that the
-  indicators in the COVIDcast repository can improve on the timeliness,
-  robustness, and scope of traditional public health reporting streams.
+\item Reinhart et al.\ \citep{Reinhart:2021} describes our group's (ongoing)
+  effort in building and maintaining COVIDcast: an open repository of real-time,
+  geographically detailed COVID-19 indicators in the U.S.  These indicators (a
+  term we use interchangeably with signals) are derived from a diverse set of
+  data sources: medical testing devices, medical insurance claims, internet
+  search trends, app-based mobility data, and online surveys, among others.
+  Many indicators are demonstrated to have meaningful statistical relationships
+  with what have become the pandemic's ``topline'' numbers (reported cases,
+  hospitalizations, deaths), whereas others uniquely reflect certain activity
+  (not available in other publicly-available data sources) that may drive or
+  affect the spread of COVID-19.  The paper demonstrates through a number of
+  examples that the indicators in the COVIDcast repository can improve on the
+  timeliness, robustness, and scope of traditional public health reporting streams.
 
-\item \citet{McDonald:2021} provides a detailed analysis of whether a core set
-  of the indicators in the COVIDcast repository can be used to improve the
-  accuracy of COVID-19 short-term forecasting and hotspot detection models.
-  This speaks to the quantitative utility of the indicators in a way that is
-  directly tied to the benefits observed in relevant downstream modeling tasks.
-  The paper finds that time series models (that are already competitive with top
-  forecasters from the COVID-19 Forecast Hub \citep{ForecastHub}) improve in
-  predictive accuracy when they are supplemented with any of the five indicators
-  under consideration.  However, the magnitude of these improvements varies with
-  the pandemic's local dynamics at prediction time, and only one indicator
-  appears to offer a nontrivial improvement during upswings.
+\item McDonald et al.\ \citep{McDonald:2021} provides a detailed analysis of
+  whether a core set of the indicators in the COVIDcast repository can be used
+  to improve the accuracy of COVID-19 short-term forecasting and hotspot
+  detection models. This speaks to the quantitative utility of the indicators in
+  a way that is directly tied to the benefits observed in relevant downstream
+  modeling tasks. The paper finds that time series models (that are already
+  competitive with top forecasters from the COVID-19 Forecast Hub
+  \citep{ForecastHub}) improve in predictive accuracy when they are supplemented
+  with any of the five indicators under consideration, based on COVID-related
+  medical insurance claims, self-reported symptoms from surveys (in fact, from
+  CTIS, described next), and COVID-related Google searches.
 
-\item \citet{Salomon:2021} focuses on the U.S. COVID-19 Trends and Impact Survey
-  (CTIS), an (ongoing) online survey operated by our group, in partnership with
-  Facebook.  This is a very rich source of data about the pandemic and its
-  effect on people, only partially reflected by the indicators (derived from the
-  survey) in the COVIDcast repository; the full data set of individual,
-  anonymized survey responses is available to researchers under a data use
-  agreement.  The paper presents descriptive analyses that reflect the unique
-  value of CTIS as an important supplement to public health reporting, in
+\item Salomon et al.\ \citep{Salomon:2021} focuses on the U.S. COVID-19 Trends 
+  and Impact Survey (CTIS), an (ongoing) online survey operated by our group, in
+  partnership with Facebook.  This is a very rich source of data about the
+  pandemic and its effect on people, only partially reflected by the indicators
+  (derived from the survey) in the COVIDcast repository; the full data set of
+  individual, anonymized survey responses is available to researchers under a
+  data use agreement.  The paper presents descriptive analyses that reflect the
+  unique value of CTIS as an important supplement to public health reporting, in
   particular, as an instrument to measure key information about behaviors, 
@@
 
-\item \citet{Astley:2021} focuses on the international version of CTIS, which is
-  an (ongoing) online survey operated by the University of Maryland, again in
-  partnership with Facebook.  This international survey covers over 100
-  countries, and is run in coordination with the U.S. one, so that they bear a
-  similar structure and undergo similar updates; the full data set of
-  individual, anonymized survey responses is again available to individual
-  researchers under a data use agreement.  The paper presents analyses that
-  reflect some basic and important characteristics of the international survey,
-  reflecting its value abroad, where public health reporting efforts may be more
-  limited than those in the U.S.
+\item Astley et al.\ \citep{Astley:2021} focuses on the international version of
+  CTIS, which is an (ongoing) online survey operated by the University of
+  Maryland, again in partnership with Facebook.  This international survey
+  covers over 100 countries and territories, and is run in coordination with the
+  U.S.\ one, so that the two bear similar structures and undergo similar
+  updates; the full data set of individual, anonymized international survey
+  responses is again available to individual researchers under a data use
+  agreement.  The paper presents analyses that reflect some basic and important
+  characteristics of the international survey, reflecting its value abroad,
+  where public health reporting efforts may be more limited than those in the
+  U.S. 
 \end{enumerate}
@@
 through all four papers in the collection, but is perhaps most directly
-addressed in \citet{Reinhart:2021} (which focuses on the ecosystem of signals
+addressed in \citep{Reinhart:2021} (which focuses on the ecosystem of signals
 broadly).  That said, we have far from saturated the utility of auxiliary
@@
 
-\emph{Authors' note: to be filled out later.}
+We thank Michael Finke, Dawn Rucker, and Nigam Shah for sustained counsel over 
+the past year; Donald Rucker for educating us on Health Information Exchanges;
+and Logan Brooks and Maria Jahja for their help assembling references for this
+perspective. More broadly, we are grateful to the entire Delphi team whose work
+is described in this special issue and is the basis for the lessons we discussed
+here. Delphi's work from its inception was enabled, encouraged, and supported by 
+the CDC's Influenza Division, and we are especially indebted to Matt
+Biggerstaff, Michael Johansson, and Carrie Reed for many years of productive
+collaboration and friendship. Lastly, Delphi's work would not be possible
+without the invaluable data and support provided by our industry partners and
+friends; a detailed list is given in the acknowledgements section of
+\citep{Reinhart:2021}.
+
+This material is based on work supported by gifts from Google.or and the McCune 
+Foundation; and by Centers for Disease Control and Prevention (CDC) grant
+U01IP001121.
 

### 2021-11-05 cover/cover.Rmd
Fix jhu and gs

@@
 library(covidcast)
+library(epitools)
 library(ggplot2)
@@
 ```{r get-data, eval = FALSE}
-sources <- c("usa-facts",
+sources <- c("jhu-csse",
              "doctor-visits",
@@
 
-```{r load-all}
+```{r missingness-checks}
 # Inspect missingness: google has lots (privacy threshold)
@@
 
+# Are there any days where google is missing almost everything?
 y <- tibble(geo_value = rep(geos, each = length(dates)),

### 2021-11-05 forecast/paper/pnas-suppl-template.tex
Add medRXiv submission doc; fix Addison's middle initial
add proper DOIs to forecast supplement

--- 2021-06-18 forecast/paper/pnas-suppl-template.tex
@@
 
-\author{Daniel J.\ McDonald, Jacob Bien, Alden Green, Addison Hu, Nat DeFries, Sangwon Hyun, Natalia L.\ Oliveira, James Sharpnack, Jingjing Tang, Robert Tibshirani, Valerie Ventura, Larry Wasserman, and Ryan J.\ Tibshirani}
+\author{Daniel J. McDonald, Jacob Bien, Alden Green, Addison J. Hu, Nat DeFries,
+  Sangwon Hyun, Natalia L. Oliveira, James Sharpnack, Jingjing Tang, Robert
+  Tibshirani, Valerie Ventura, Larry Wasserman, and Ryan J. Tibshirani} 
 \correspondingauthor{Daniel J. McDonald.\\E-mail: daniel@stat.ubc.ca}
@@
 
+\FloatBarrier
+
+\dataset{Code}{Code for reproducibility. {\tt R} script files and small
+  datasets. 
+  Persistent DOI: \url{https://doi.org/10.5281/zenodo.5639567}}
+
+\dataset{Data}{Large intermediate data sets. Archived {\tt .RDS} (R objects)
+  files 
+  containing all evaluations for forecasting and hotspots using vintage and
+  finalized data. Can be used to produce the graphics and conclusions in the
+  manuscript without rerunning the entire pipeline. 
+  Persistent DOI: \url{https://doi.org/10.5683/SP3/UW4VTC}}
+
 \bibliography{$for(bibliography)$$bibliography$$sep$,$endfor$}

### 2021-11-05 forecast/paper/supplement.tex
add proper DOIs to forecast supplement

@@
 
-\dataset{predictions.zip}{Archived {\tt .RDS} (R objects) files containing
-  all predictions for forecasting and hotspots using vintage and finalized data.
-Persistent DOI to be added at publication.}
+\dataset{Code}{Code for reproducibility. {\tt R} script files and small
+  datasets. 
+  Persistent DOI: \url{https://doi.org/10.5281/zenodo.5639567}}
 
-\dataset{evaluations.zip}{Archived {\tt .RDS} (R objects) files containing
-  all evaluations for forecasting and hotspots using vintage and finalized data.
-  Persistent DOI to be added at publication.}
-
-\dataset{analysis.zip}{Archived {\tt .RDS} (R objects) files containing
-  additional data used to produce graphics and conclusions in the manuscript.
-  Persistent DOI to be added at publication.}
-
-\dataset{code.zip}{R script files containing
-  all code used to reproduce the analysis described in the manuscript.
-  Persistent DOI to be added at publication.}
+\dataset{Data}{Large intermediate data sets. Archived {\tt .RDS} (R objects)
+  files 
+  containing all evaluations for forecasting and hotspots using vintage and
+  finalized data. Can be used to produce the graphics and conclusions in the
+  manuscript without rerunning the entire pipeline. 
+  Persistent DOI: \url{https://doi.org/10.5683/SP3/UW4VTC}}
 

### 2021-11-08 forecast/code/deprecated/build_map.Rmd
update forecast readme

---
title: Build the mapping from HRRs to states
---

```{r, message = FALSE}
library(covidcast)
library(dplyr)
library(tidyr)

start_day <- "2020-03-01"
end_day <- "2021-10-15"

y <- covidcast_signal(data_source = "jhu-csse", 
                      signal = "confirmed_incidence_prop",
                      start_day = start_day, 
                      end_day = end_day,
                      geo_type = "state") %>%
  select(geo_value, time_value, value) %>%
  pivot_wider(names_from = geo_value)

dim(y)
head(y)

x <- covidcast_signal(data_source = "jhu-csse", 
                      signal = "confirmed_incidence_prop",
                      start_day = start_day, 
                      end_day = end_day,
                      geo_type = "hrr") %>%
  select(geo_value, time_value, value) %>%
  pivot_wider(names_from = geo_value)

dim(x)
head(x)

# Check we have the same time values
all(y$time_value == x$time_value)

# Convert to matrices, compute regression coefficients
library(MASS) # Using ginv() because I got singularity errors with solve()
x_mat = as.matrix(x[,-1])
y_mat = as.matrix(y[,-1])
b = t(cov(y_mat, x_mat) %*% ginv(cov(x_mat)))

# Sanity check: check against first 4 individual regressions
par(mfrow = c(2, 2))
for (i in 1:4) plot(b[,i], coef(lm(y_mat[,i] ~ x_mat + 0)))

# Sanity check: look at reconstruction error by state
err = colMeans((y_mat - x_mat %*% b)^2)
barplot(err)
```

### 2021-11-08 forecast/paper/paper-medrxiv.tex
update forecast medrXiv

@@
 \begin{abstract}
-  Reliable, short-term forecasts of traditional public health reporting streams
-  (such as cases, hospitalizations, and deaths) are a key ingredient in
-  effective public health decision-making during a pandemic. Since April 2020,
-  our research group has worked with data partners to collect, curate, and make
-  publicly available numerous real-time COVID-19 indicators, providing multiple
-  views of pandemic activity.  This paper studies the utility of these
-  indicators from a forecasting perspective. We focus on five indicators,
-  derived from medical insurance claims data, web search queries, and online
-  survey responses. For each indicator, we ask whether its inclusion in a simple
-  model leads to improved predictive accuracy relative to a similar model
-  excluding it. We consider both probabilistic forecasting of confirmed COVID-19
-  case rates and binary prediction of case ``hotspots''.  Since the values of 
-  indicators (and case rates) are commonly revised over time, we take special
-  care to ensure that the data provided to a forecaster is the version that
-  would have been available at the time the forecast was made.   
-  % We demonstrate how less careful backtesting can lead to misleading
-  % retrospective evaluations.   
-  Our  analysis shows that consistent but modest gains in predictive accuracy 
-  are obtained by using these indicators, and furthermore, these gains are
-  related to periods in which the auxiliary indicators behave as ``leading
-  indicators'' of case rates.
+ Short-term forecasts of traditional streams from public health reporting (such 
+  as cases, hospitalizations, and deaths) are a key input to public health
+  decision-making during a pandemic. Since early 2020, our research group has
+  worked with data partners to collect, curate, and make publicly available
+  numerous real-time COVID-19 indicators, providing multiple views of pandemic
+  activity in the U.S.  This paper studies the utility of five such
+  indicators---derived from de-identified medical insurance claims,
+  self-reported symptoms from online surveys, and COVID-related Google search
+  activity---from a forecasting perspective. For each indicator, we ask whether
+  its inclusion in an autoregressive (AR) model leads to improved predictive
+  accuracy relative to the same model excluding it. Such an AR model, without
+  external features, is already competitive with many top COVID-19 forecasting
+  models in use today. Our analysis reveals that (a) inclusion of each of these
+  five indicators improves on the overall predictive accuracy of the AR model;
+  (b) predictive gains are in general most pronounced during times in which
+  COVID cases are trending in ``flat'' or ``down'' directions; (c) one
+  indicator, based on Google searches, seems to be particularly helpful
+  during ``up''  trends.
 
@@
 streams---such as confirmed cases and deaths in the COVID-19 pandemic---is
-crucial for understanding disease spread, formulating public policy responses,
-and anticipating future public health resource needs.  In a companion paper, we
-describe our research group's (Delphi's) efforts in curating and maintaining a 
-database of real-time indicators that track COVID-19 activity and other relevant
-phenomena. The signals (a term we use synonomously with ``indicators'') in this
-database are accessible through the COVIDcast API \cite{CovidcastAPI}, with
-associated R \cite{CovidcastR} and Python \cite{CovidcastPy} packages for
-convenient data fetching and processing tools. In the current paper, we aim to
-quantify the utility provided by a core set of these indicators for two
-fundamental prediction tasks: probabilistic forecasting of COVID-19 case 
-rates and prediction of future COVID-19 case hotspots (defined by the event that
-a relative increase in COVID-19 cases exceeds a certain threshold).
+crucial for understanding disease spread, correctly formulating public policy
+responses, and rationally planning future public health resource needs.  A
+companion paper \cite{Reinhart:2021} describes our research group's efforts,
+beginning in April 2020, in curating and maintaining a database of real-time
+indicators that track COVID-19 activity and other relevant phenomena. The
+signals (a term we use synonomously with ``indicators'') in this database are
+accessible through the COVIDcast API \cite{CovidcastAPI}, as well as associated
+R \cite{CovidcastR} and Python \cite{CovidcastPy} packages, for convenient data  
+fetching and processing. In the current paper, we quantify the utility provided
+by a core set of these indicators for two fundamental prediction tasks:
+probabilistic forecasting of COVID-19 case rates and prediction of future
+COVID-19 case hotspots (defined by the event that a relative increase in
+COVID-19 cases exceeds a certain threshold).
 
@@
 to provide an authoritative take on cutting-edge COVID-19 forecasting methods.
-Instead, our purpose here is to provide a rigorous, quantitative assessment of
-the utility that several auxiliary indicators---such as those derived from
-internet surveys or medical insurance claims---provide in tasks that involve
-predicting future trends in confirmed COVID-19 cases. To assess such utility in
-as simple terms as possible, we center our study in the framework of a basic
-autoregressive model (in which COVID cases in the near future are predicted from  
-a linear combination of COVID cases in the near past), and ask whether the 
-inclusion of an auxiliary indicator as an additional feature in such a model
-improves its predictions.
+Similarly, some authors, e.g., \cite{ioannidis2020a}, have pointed out numerous
+mishaps of forecasting during the pandemic, and it is not our general intent to
+fix them here. Instead, we start with a basic and yet reasonably effective
+predictive model for predicting future trends in COVID-19 cases, and present a
+rigorous, quantitative assessment of the added value provided by auxiliary
+indicators, that are derived from data sources that operate outside of 
+traditional public health streams. In particular, we consider five indicators
+derived from de-identified medical insurance claims, self-reported symptoms from
+online surveys, and COVID-related Google searches.
 
-While forecasting carries a rich literature that offers a wide range of
-techniques, see e.g., \cite{Hyndman:2018}, we purposely constrain ourselves to
-very simple models, avoiding common enhancements such as order selection,
-correction of outliers/anomalies in the data, and inclusion of regularization or
-nonlinearities. That said, analyses of forecasts submitted to the COVID-19
-Forecast Hub \cite{ForecastHub} by a large community of modelers have shown that
-simple, robust models have consistently been among the best-performing over the
-pandemic \cite{Cramer:2021}, including time series models similar to those we
-consider in what follows.
+To assess this value in as direct terms as possible, we base our study around a
+very simple basic model: an autoregressive model, in which COVID cases in the
+near future are predicted using a linear combination of COVID cases in the near
+past. Forecasting carries a rich literature, offering a wide range of
+sophisticated techniques, see, e.g., \cite{Hyndman:2018} for a review; however,
+we purposely avoid enhancements such as order selection, correction of
+outliers/anomalies in the data, and inclusion of regularization or
+nonlinearities. Similarly, we do not account for other factors that may well aid
+in forecasting, such as age-specific effects, holiday adjustments, and the
+effects of public health mandates. All that said, despite its simplicity, the
+basic autoregressive model that we consider in this paper exhibits competitive
+performance (see the Supplementary Materials for details) with many of the top
+COVID-19 case forecasters submitted to the U.S.\ COVID-19 Forecast Hub
+\cite{ForecastHub}, which is the official source of forecasts used in public
+communications by the U.S. CDC. The strong performance of the autoregressive
+model here is in line with the fact that simple, robust models have also
+consistently been among the best-performing ones for COVID-19 death
+forecasting \cite{Cramer:2021}.
 
-In our companion paper, we analyze correlations between various indicators and   
-COVID case rates. These correlations are natural summaries of the
-contemporaneous association between an indicator and COVID cases, but they fall 
-short of delivering a satisfactory answer to the question that motivates the
-current article: is the information contained in an indicator demonstrably
-useful for the prediction tasks we care about? For such a question, lagged
-correlations (e.g., measuring the correlation between an indicator and COVID
-case rates several days in the future) move us in the right direction, but still
-do not move us all the way there. The question about \textit{utility for 
-  prediction} is focused on a much higher standard than simply asking 
+In the companion paper \cite{Reinhart:2021}, we analyze correlations between
+various indicators and COVID case rates. These correlations are natural
+summaries of the contemporaneous association between an indicator and COVID
+cases, but they fall short of delivering a satisfactory answer to the question
+that motivates the current article: is the information contained in an indicator
+demonstrably useful for the prediction tasks we care about? Note that even
+lagged correlations cannot deliver a complete answer. Demonstrating
+\textit{utility for prediction} is a much higher standard than simply asking
 about correlations; to be useful in forecast or hotspot models, an indicator
 must provide relevant information that is not otherwise contained in past values
-of the case rate series itself. We will assess this in the most direct way
-possible: by inspecting the difference in predictive performance of simple
-autoregressive models trained with and without access to past values of a
-particular indicator.
+of the case rate series itself (cf.\ the pioneering work on Granger  causality
+\cite{Granger:1969, engle1987a}, as well as the further references given 
+below). We assess this directly by inspecting the difference in predictive  
+performance of simple autoregressive models trained with and without access to  
+past values of a particular indicator.
 
-There is another, more subtle issue in evaluating predictive utility that 
-deserves explicit mention, as it will play a key role in our analysis.
-Signals computed from surveillance streams will often be subject to  
-latency and/or revision. For example, a signal based on aggregated medical
-insurance claims may be available after just a few days, but it can then be
-substantially revised over the next several weeks as additional claims are
-submitted and/or processed late. Correlations between such a signal and case
-rates calculated ``after the fact'' (i.e., computed retrospectively, using the
-finalized values of this signal) will not deliver an honest answer to the    
-question of whether this signal would have been useful in real time. Instead,
-we build predictive models using only the data that would have been available
-\textit{as of} the prediction date, and compare the ensuing predictions in terms 
-of accuracy. To do so, we leverage Delphi's \texttt{evalcast} R package 
-\cite{EvalcastR}, which plugs into the COVIDcast API's data versioning system,  
-and facilitates honest backtesting.
+We find that each of the five indicators we consider--- two based on
+COVID-related outpatient visits from medical insurance claims, one on
+self-reported symptoms from online surveys, and one on Google searches for
+anosmia or ageusia---provide an overall improvement in accuracy when
+incorporated into the autorgressive model. This is true both for COVID-19 case 
+forecasting and hotspot prediction. Further analysis reveals that the gains in
+accuracy depend on the pandemic's dynamics at prediction time: the biggest  
+gains in accuracy appear during times in which cases are ``flat'' or trending
+``down''; but the indicator based on Google searches offers a most notable
+improvement when cases are trending ``up''.
+
+Careful handling of data revisions plays a key role in our analysis. Signals 
+computed from surveillance streams are often subject to latency and/or 
+revision. For example, a signal based on aggregated medical insurance claims may
+be available after just a few days, but it can then be substantially revised
+over the next several weeks as additional claims are submitted and/or processed
+late. Correlations between such a signal and case rates calculated ``after the
+fact'' (i.e., computed retrospectively, using the finalized values of this
+signal) will not deliver an honest answer to the question of whether this signal
+would have been useful in real time. Instead, we build predictive models using
+only the data that would have been available \textit{as of} the prediction date,
+and compare the ensuing predictions in terms of accuracy. The necessity of
+real-time data for honest forecast evaluations has been recognized in
+econometrics for a long time, see \cite{howrey1978a, harvey1983a, mankiw1986a,
+  trivellato1986a, diebold1991a, patterson1995a, koenig2003a, croushore2006a,
+  clark2009a, faust2009a, croushore2011a}; but it is often overlooked in
+epidemic forecasting despite its critical importance, see, e.g.,
+\cite{chakraborty2018know}.
 
@@
 containing Los Angeles, which has more than 10 million people), generally HRRs  
-are much more homogenous in size than the (approximately) 3200 counties,  
+are much more homogenous in size than the (approximately) 3200 U.S.\ counties,   
 and they serve as a nice middle ground in between counties and states.
@@
 point in time.)  This is possible because of the COVIDcast API's comprehensive
-data versioning system (described in more detail in our companion paper).  We
-also use the \texttt{evalcast} R package \cite{EvalcastR}, which streamlines the 
-process of training arbitrary prediction models over a sequence of prediction
-dates, by constructing the proper sequence of vintage training data sets.
+data versioning system (described in more detail in the companion paper
+\cite{Reinhart:2021}).  We also use the \texttt{evalcast} R package
+\cite{EvalcastR}, which streamlines the process of training arbitrary
+prediction models over a sequence of prediction dates, by constructing the
+proper sequence of vintage training data sets.
 
 \begin{figure}[tb!]
-  \centering
-  \includegraphics[width=.5\linewidth]{fig/revisions.pdf}
+  \includegraphics[width=\columnwidth]{fig/revisions.pdf}
   % RJT: Ensuring we start with a "declarative title" :) 
@@
 Because Google's team did not start publishing this signal until early
-September, 2020, we do not have true vintage data before then.  Furthermore, the
-latency of the signal was always at least one week through 2020.  However, this
-signal is never revised after initial publication (confirmed via personal
-communication with the Google team that produces this signal) and furthermore
-the latency of the signal is not an unavoidable property of the data type, so we
-simply use finalized signal values, with zero latency, in our analysis.
+September, 2020, we do not have true vintage data before then; the latency of 
+the signal was always at least one week through 2020.  However, unlike the
+claims-based signals, there is no reason for revisions to occur after initial
+publication, and furthermore, the latency of the signal is not an unavoidable
+property of the data type, so we simply use finalized signal values, with zero
+latency, in our analysis.
 
@@
 Why do we define the response variables via 7-day averaging? The short answer 
-is robustness: averaging stabilizes the case time series, and accounts for 
-uninteresting artifacts like day-of-the-week effects in the series.  Note that
-we can also equivalently view this (equivalent up to a constant factor) as
+is robustness: averaging stabilizes the case time series, and moderates
+uninteresting artifacts like day-of-the-week effects in the data.  We note
+that we can also equivalently view this (equivalent up to a constant factor) as 
 predicting the HRR-level case incidence rate \textit{summed} over some 7-day
@@
 the CTIS-CLI-in-community signal, which does not exist before early April 2020,
-when the survey was first launched). The end date was chosen again with a 
-consideration to align both tasks as best as possible, and because few hotspots
-exist post December 31, 2020, due to the general and gradual decline of the
-pandemic in 2021.
+when the survey was first launched).
 
@@
 \begin{figure}[t]
-  \centering
-\includegraphics[width=.5\columnwidth]{fig/trajectory.pdf}
+\includegraphics[width=\columnwidth]{fig/trajectory.pdf}
 \caption{Forecast for the HRR containing New York City from an autoregressive
@@
 quantile models in \eqref{eq:forecast_ar} and \eqref{eq:forecast_ar_x} using
-\textit{weighted interval score} (WIS), a quantile-based scoring rule; see e.g.,
-\cite{Gneiting:2007}.  WIS is a proper score, which means that its expectation 
-is minimized by the population quantiles of the target variable.  The use of WIS
-in COVID-19 forecast scoring is discussed in \cite{Bracher:2021}; WIS is also
-the main evaluation metric used in the COVID-19 Forecast Hub.
+\textit{weighted interval score} (WIS), a quantile-based scoring rule; see,
+e.g., \cite{Gneiting:2007}.  WIS is a proper score, which means that its
+expectation is minimized by the population quantiles of the target variable.
+The use of WIS in COVID-19 forecast scoring was proposed by \cite{Bracher:2021}; 
+WIS is also the main evaluation metric used in the COVID-19 Forecast Hub. More
+broadly, the specific form of WIS used here is a standard metric in the
+forecasting community for evaluating quantile-based probabilistic forecasts,
+just as mean squared forecast error is standard for point forecasts.
 
@@
 \eqref{eq:forecast_ar_x}, or hotspot prediction, \eqref{eq:hotspot_ar_x}, that
-uses this particular indicator as a feature (the meaning should be clear from
+uses this same indicator as a feature (the meaning should be clear from
 the context).  So, for example, the CHNG-CLI model in forecasting is the one in  
 \eqref{eq:forecast_ar_x} that sets $X_{\ell,t}$ to be the value of the CHNG-CLI
-indicator at location $\ell$ and time $t$.
-% fits the coefficients based on the training data, and makes its predictions by
-% setting $X_{\ell,t}$ to the appropriate CHNG-CLI values.  
-Finally, we use the term ``indicator model'' to refer to any one of the ten
-models of the form \eqref{eq:forecast_ar_x} or \eqref{eq:hotspot_ar_x} (five
-from each of the forecasting and hotspot prediction tasks).
+indicator at location $\ell$ and time $t$.  Finally, we use the term ``indicator
+model'' to refer to any one of the ten models of the form
+\eqref{eq:forecast_ar_x} or \eqref{eq:hotspot_ar_x} (five from each of the
+forecasting and hotspot prediction tasks).
 
-We begin with a summary of the high-level conclusions.
+Below is a summary of the high-level conclusions.
 
@@
   indicator models tend to be better than AR when case rates are flat or
-  decreasing (most notable in CHNG-COVID and CTIS-CLI-in-community), but they
-  are worse than AR when case rates are increasing (this is most notable in
-  CHNG-CLI and DV-CLI). More rarely does an indicator model tend to beat AR when
-  case rates are increasing, but there appears to be some evidence of this for
-  the Google-AA model. 
+  decreasing (most notable in CHNG-COVID and CTIS-CLI-in-community), but can be
+  worse than AR when case rates are increasing (this is most notable in CHNG-CLI
+  and DV-CLI). More rarely does an indicator model tend to beat AR when case
+  rates are increasing, but there appears to be evidence of this for the
+  Google-AA model.   
   % In an increasing period, when an indicator model is worse than AR, it tends
@@
 
-\item In this same analysis, when an indicator model performs better than AR in
-  a decreasing period, this tends to co-occur with instances in which the
+\item In this same analysis, when an indicator model performs better than AR 
+  in a decreasing period, this tends to co-occur with instances in which the 
   indicator ``leads'' case rates (meaning, roughly, on a short-time scale in a
   given location, its behavior mimics that of future case rates).  On
-  the other hand, if an an indicator model does better in periods of
+  the other hand, if an indicator model does better in periods of
   increase, or worse in periods of increase or decrease, its performance is
@@
 proper vintage data, we ran a parallel set of forecasting and hotspot prediction
-experiments using finalized data. The results, given in the supplement, show
-that training and making predictions on finalized data can result in overly
-optimistic estimates of true test-time performance (up to 10\% better in terms
-of average WIS or AUC). Furthermore, since indicators can have greatly different
-backfill profiles, the use of finalized data in retrospective evaluations
-changes the relative ranking of models.  For example, CHNG-CLI and DV-CLI,
-when trained on finalized data, perform very similarly in forecasting.
+experiments using finalized data. The results, given in the Supplementary
+Materials, show that training and making predictions on finalized data can
+result in overly optimistic estimates of true test-time performance (up to 10\%
+better in terms of average WIS or AUC). Furthermore, since indicators can have
+greatly different backfill profiles, the use of finalized data in retrospective
+evaluations changes the relative ranking of models.  For example, CHNG-CLI and
+DV-CLI, when trained on finalized data, perform very similarly in forecasting.
 This makes sense since they are both claims-based indicators that are supposedly
-measuring the same thing.  However, DV-CLI outperforms CHNG-CLI on vintage
-data, reflecting its has a less severe backfill profile.
+measuring the same thing.  However, DV-CLI outperforms CHNG-CLI on vintage data,
+reflecting its has a less severe backfill profile.
 
-Code to reproduce all results (which uses the \texttt{evalcast} R package) can 
-be found at
+The Supplement provides a number of other additional analyses; for example, we
+examine two assumption-lean methods for assessing the statistical significance
+of our main results.
+
+Code to reproduce all results can be found at
 \url{https://github.com/cmu-delphi/covidcast-pnas/tree/main/forecast/code}.
@@
 dates.  Shown is the average WIS for each forecast model divided by that from a
-baseline model, which is basically a flat-line forecaster (its median forecast
-for $Y_{\ell,t+a}$ is always $Y_{\ell,t}$, with predicted quantiles defined
-around this based on historical variation). This is the same baseline model as
-in the COVID-19 Forecast Hub. Here, we use the baseline model in order to scale
-mean WIS so that it gets put on an interpretable, unitless scale.  In the
-figure, we can see that all curves are below 1, which means (smaller WIS is
-better) that all of the models, including AR, outperform the baseline on average
-over the forecasting period.  On the other hand, the models deliver at best an
-improvement of about 20\% in average WIS over the baseline model, with this gap
-narrowing to about 10\% at the largest ahead values, illustrating the difficulty 
-of the forecasting problem.
+baseline model. The baseline model is a flat-line forecaster which forms its
+median forecast by using the most recent value $Y_{\ell,t}$ for all aheads
+$Y_{\ell,t+a}$, with predicted quantiles defined by the empirical distribution
+of the residuals from this median forecast over recent history. This is the same
+baseline model as in the COVID-19 Forecast Hub. Here, we use the baseline model
+in order to scale mean WIS to put it on an interpretable, unitless
+scale.  In the figure, we can see that all curves are below 1, which means
+(smaller WIS is better) that all of the models, including AR, outperform the
+baseline on average over the forecasting period.  On the other hand, the models
+deliver at best an improvement of about 20\% in average WIS over the baseline
+model, with this gap narrowing to about 10\% at the largest ahead values,
+illustrating the difficulty of the forecasting problem.
 
@@
 
-One might ask if the benefits we observe in forecasting and hotspot
-prediction have anything to do with the actual indicator themselves. A plausible  
-alternative explanation is that the indicators are simply providing \textit{implicit
-  regularization} on top of the basic AR model, in the same way any noise
-variable might, when we include them as lagged features in
-\eqref{eq:forecast_ar_x} and \eqref{eq:hotspot_ar_x}.
+One might ask if the benefits observed in forecasting and hotspot
+prediction have anything to do with the actual auxiliary indicators
+themselves. A plausible alternative explanation is that the indicators are
+just providing \textit{implicit regularization} on top of the basic AR model,
+in the same way any noise variable might, if we were to use it to create lagged
+features in \eqref{eq:forecast_ar_x} and \eqref{eq:hotspot_ar_x}.
 
 To test this hypothesis, we reran all of the prediction experiments but with
-$X_{\ell,t}$ in the each indicator models replaced with suitable random noise
+$X_{\ell,t}$ in each indicator model replaced by suitable random noise
 (bootstrap samples from a signal's history).  The results, shown and explained
-more precisely in the supplement, are vastly different (worse) than the original
-set of results.  In both forecasting and hotspot prediction, the ``fake''
-indicator models offered essentially no improvement over the pure AR model,
-which strongly rejects (informally speaking) the implicit regularization
-hypothesis.
+more precisely in the Supplement, are vastly different (i.e., worse) than the
+original set of results.  In both forecasting and hotspot prediction, the
+``fake'' indicator models offer essentially no improvement over the pure AR
+model, which---informally speaking---strongly rejects the implicit
+regularization hypothesis.
 
 On the topic of regularization, it is also worth noting that the use of 
-$\ell_1$ regularization (tuned by cross-validation) in fitting any of the models
-in \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar}, or 
-\eqref{eq:hotspot_ar_x} did not generally improve their performance
-(experiments not shown).  This is likely due to the fact that the number of
-training samples is large compared to the number of features (6,426 training
-samples and only 3--6 features). 
+$\ell_1$ regularization (tuned using cross-validation) in fitting any of the
+models in \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x},
+\eqref{eq:hotspot_ar}, or  \eqref{eq:hotspot_ar_x} did not generally improve
+their performance (experiments not shown).  This is likely due to the fact that
+the number of training samples is large compared to the number of features
+(6,426 training samples and only 3--6 features).  
 % DJM: and that proper CV requires longer training windows, exacerbating the
@@
 increasing period, we use the same definition we used for the hotspot task in
-Table \ref{tab:analysis_tasks}.  Therefore all hotspots are ``up'', while all
-non-hotspots are either ``flat'' or ``down''.  For the ``down'' scenario, we 
-simply use the opposite of the hotspot definition: $Y_{\ell,t}$ decreases by
-more than 20\% relative to the preceding week.
+Table \ref{tab:analysis_tasks}.  Therefore, all hotspots are labeled ``up'',
+while all non-hotspots are either ``flat'' or ``down''.  For the ``down''
+scenario, we simply use the opposite of the hotspot definition: $Y_{\ell,t}$
+decreases by more than 20\% relative to the preceding week.
 
-While the performance of all forecasters, including AR, will generally degrade
-in up periods, different models exhibit different and interesting patterns.
-CHNG-CLI, CHNG-COVID, Google-AA, and especially CTIS-CLI-in-community have
-large right tails (displaying improvements over AR) during the down periods. 
-Google-AA and CTIS-CLI-in-community have large right tails during the flat
-periods. CHNG-CLI and DV-CLI have large left tails (poor forecasts relative to
-AR) in flat and up periods.  Google-AA is the only model that outperforms
-the AR model, on average, over up periods.  Overall, the indicators seem to help
-more during flat or down periods than up periods, with the exception of
-Google-AA.
+While the performance of all models, including AR, generally degrades in up
+periods, different models exhibit different and interesting patterns. CHNG-CLI,
+CHNG-COVID, Google-AA, and especially CTIS-CLI-in-community have large right
+tails (showing improvements over AR) during the down periods.  Google-AA and
+CTIS-CLI-in-community have large right tails during the flat periods. CHNG-CLI
+and DV-CLI have large left tails (poor forecasts relative to AR) in flat and up
+periods.  Google-AA is the only model that outperforms the AR model, on average,
+in up periods.  Overall, the indicators seem to help more during flat or down
+periods than during up periods, with the exception of Google-AA.
 
-The supplement pursues this analysis further.  For example, we examine
+The Supplement pursues this analysis further.  For example, we examine
 classification accuracy and log-likelihood for the hotspot task and find a
@@
 \subsection{Effects of Leading or Lagging Behavior}
+\label{sec:leading-lagging}
 
 \begin{figure}[t]
-  \centering
-  \includegraphics[width=.5\columnwidth]{fig/leading-only-1.pdf}
+  \includegraphics[width=\columnwidth]{fig/leading-only-1.pdf}
   \caption{Correlation of the difference in WIS with the ``leadingness'' of the
@@
 
-In the supplement, we show how to define a quantitative score to measure the
+In the supplement, we define a quantitative score to measure the 
 leadingness of an indicator, at any time $t$ and any location $\ell$, based on 
 cross correlations to case rates over a short time window around $t$.
-The higher this score, the greater it ``leads'' case activity.  Figure 
-\ref{fig:leading} displays correlations between the leadingness score of an 
-indicator and the WIS difference (AR model minus an indicator model), stratified
-by whether the target is classified as up, down, or flat.  One would naturally
-expect that the WIS difference would be positively correlated with leadingness.
-Somewhat surprisingly, this relationship turns out to be strongest in down
-periods and weakest in up periods.  In fact, it is very nearly the case that for
-each indicator, the strength of correlations only decreases as we move from down
-to flat to up  periods.  In the supplement, we extend this analysis by studying 
+The higher this score, the greater it ``leads'' case activity. This analysis is
+closely related to Granger causality \cite{engle1987a} and draws on a large
+body of prior work that measures leadingness in economic time-series, e.g., 
+\cite{mitchell1938a, koopmans1947a, shiskin1968a, granger1969a, yeats1972a, 
+  sargent1977a, auerbach1982a, koch1988a, StockWatson1989, diebold1989a,
+  lahiri1991a, emerson1996a, hamilton1996a}. Figure \ref{fig:leading} displays
+correlations between the leadingness score of an indicator and the WIS
+difference (AR model minus an indicator model), stratified by whether the target
+is classified as up, down, or flat.  One would naturally expect that the WIS
+difference would be positively correlated with leadingness.  Somewhat
+surprisingly, this relationship turns out to be strongest in down periods and
+weakest in up periods.  In fact, it is very nearly the case that for each
+indicator, the strength of correlations only decreases as we move from down to
+flat to up periods.  In the supplement, we extend this analysis by studying
 analogous ``laggingness'' scores, but we do not find as clear patterns.
@@
 levels of nuance to such an answer that must be explained.  None of the
-indicators that we have investigated appear to be the ``silver bullet'' that one
-might have hoped for, revolutionizing the tractability of the prediction
-problem, rendering it easy when it was once hard (in the absence of auxiliary
+indicators that we have investigated are transformative, rendering the
+prediction problem easy when it was once hard (in the absence of auxiliary
 information). Rather, the gains in accuracy from the indicator models (over an 
@@
 techniques would change the ``topline'' conclusion---that auxiliary indicators
-can provide nontrivial, consistent, but modest gains in forecasting and hotspot
-prediction. Whether a more sophisticated model would be able to leverage the
-indicators in such a way as to change some of the finer conclusions (e.g., by 
-offering clear improvements in periods in which cases are trending up) is less 
-clear to us.
+can provide clear but modest gains in forecasting and hotspot prediction.
+
+However, rigorously vetting the details for more sophisticated models, as well
+as the generalizability of our findings to different geographic resolutions,
+both remain important directions for future study.  For example, in the
+Supplementary Materials we show that for forecasting at the state level, the
+benefits of including indicators in the AR model are generally less clear
+(compared to those observed at the HRR level).  A plausible explanation: at the
+state level, where the signal-to-noise ratio (SNR) is higher, AR performs better
+overall and represents a higher standard (when asking whether it can be improved
+upon using the indicators).  At the HRR level, where the SNR is lower, including
+the indicators as additional linear features in the AR model probably delivers a
+kind of variance reduction (just like averaging independent terms would) which
+contributes to improved accuracy.  But as the SNR increases, this variance
+reduction becomes less important, and perhaps we must use more sophisticated
+modeling techniques to extract comparable value from the indicators.
 
@@
 will our sense of accuracy be unrealistic, but certain models may degrade by a
-greater or lesser extent when they are forced to work with vintage data, so 
-backtesting them with finalized data may lead us to make modeling decisions that
-are suboptimal for true test-time performance.
+greater or lesser extent when they are forced to reckon with vintage data, so  
+backtesting on finalized data may lead us to make modeling decisions that are
+suboptimal for true test-time performance.
 
@@
 while devoting most of our effort to accounting for as much of the complexity of
-the underlying data and evaluation as possible.  In fact, our paper is as much
-about demonstrating how one might address questions about model comparisons and
-evaluation in forecasting and hotspot prediction in general, as it is about
-providing rigorous answers to such questions in the context of COVID-19 case
-rates in particular.  We hope that others will leverage our framework, and build
-on it, so that it can be used to guide work that advances the frontier of
+the underlying data and evaluation as possible. In fact, our paper is not only
+about providing rigorous answers to questions about model comparisons in
+COVID-19 forecasting and hotspot prediction, but also about demonstrating how 
+one might go about answering such questions in general. We hope that others will
+leverage our work, and build on it, to guide advances on the frontier of
 predictive modeling for epidemics and pandemics.

### 2021-11-09 forecast/paper/supplement-medrxiv.tex

@@
 
-\caption{Relative forecast performance using vintage data and summarizing with the more robust geometric mean.}\label{fig:fcast-adjusted}
+\caption{Forecast performance (using vintage data), summarized by geometric mean.}\label{fig:fcast-adjusted}
 \end{figure}
@@
 \clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/compare-states-to-hub}
+
+}
+
+\caption{Forecast performance for AR and indicator models, each retrained at the state level, compared to models submitted to the COVID-19 Forecast Hub over the same period.
+The thin grey lines are individual models from the Hub; the light blue line is the Hub ensemble model.
+To align prediction dates as best as possible, we look at the AR and indicator model forecasts for 5, 12, 19, and 26 days ahead; this roughly corresponds to 1, 2, 3, and 4 weeks ahead, respectively, since in the Hub, models typically submit forecasts on a Tuesday for the epiweeks aligned to end on each of the following 4 Saturdays.}\label{fig:compare-to-hub}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/sign-test-1}
+
+}
+
+\caption{P-values from a one-sided sign test for improved forecast performance of the indicator-assisted models.
+Each p-value corresponds to a forecast date.
+The alternative hypothesis is that the AR model is better (median difference between the relative WIS of the AR model and an indicator model is negative).}\label{fig:sign-test}
+\end{figure}
+
+\begin{table}
+
+\caption{\label{tab:dm-test}P-values from a one-sided Diebold-Mariano test for improvemed forecast performance when adding the indicators. The alternative hypothesis is that the AR model is better.}
+\centering
+\begin{tabular}[t]{lrrrrr}
+\toprule
+metric & CHNG-CLI & CHNG-COVID & CTIS-CLIIC & DV-CLI & Google-AA\\
+\midrule
+Geometric mean & 0.072 & 0.036 & 0.005 & 0.032 & 0.026\\
+Mean & 0.177 & 0.132 & 0.006 & 0.092 & 0.103\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
-\caption{Classification and loglikelihood separated into periods of upswing, downswing, and flat cases.
-Like the analysis of the forecasting task in the main paper (see Figure 7), performance is better during down and flat periods.}\label{fig:hotspots-upswing-downswing}
+\caption{Percentage change in classification error and log likelihood, relative that of the AR model, separated into up, down, and flat periods.
+Like the analogous forecasting analysis, the indicator models generally do better during down and flat periods.}\label{fig:hotspots-upswing-downswing}
 \end{figure}
-
-\clearpage
 
@@
 
-\caption{Histograms of the Spearman correlation between the ratio of AR to AR WIS with the percent change in smoothed case rates relative to 7 days earlier.}\label{fig:cor-wis-ratio}
+\caption{Histograms of the Spearman correlation between the ratio of forecaster WIS to AR WIS with the \% change in case rates, relative to case rates 7 days earlier.}\label{fig:cor-wis-ratio}
 \end{figure}
-
-\clearpage
 
@@
 
-\caption{Correlation of the difference in WIS with the difference in median predictions for the AR model relative to the indicator-assisted forecaster.
-In down periods, improvements in forecast risk are highlycorrelated with lower median predictions.
-The opposite is true in up periods.
-This suggests, as one might expect that improved performance of the indicator-assisted model is attributable to being closer to the truth then the AR model.
-This conclusion is stronger in down periods then in up periods.}\label{fig:upswing-corr-table}
+\caption{Correlation of the difference in WIS with the difference in median predictions (each difference being between the AR model and an indicator model), separated into up, down, and flat periods.
+In down periods, improvements in forecast error are highly correlated with lower median predictions.
+The opposite is true in up periods, but the conclusion here appears to be weaker.}\label{fig:upswing-corr-table}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/upswing-summary-remake-1}
+
+}
+
+\caption{Percentage change in average WIS of the forecaster (AR or indicator assisted), relative to the baseline.
+All models perform poorly during down periods, but the indicators help.
+During flat periods, the indicators improve slightly over the AR.
+During up periods, all forecasters do much better than the baseline, but only some do as well as AR.}\label{fig:upswing-summary-remake}
 \end{figure}
@@
 The right panel shows $\CCF_{\ell}(a)$ for different values of $a$ as vertical blue bars.
-The orange dashed lines indicate the 95\% significance threshold.
-By our leadingness/laggingness metric, DV-CLI is leading (but notlagging) cases over this period.}\label{fig:ccf-dv-finalized}
+The orange dashed line indicates the 95\% significance threshold.
+By our leadingness/laggingness metric, DV-CLI is leading (but not lagging) cases over this period.}\label{fig:ccf-dv-finalized}
 \end{figure}
@@
 
+\clearpage
+
 \begin{figure}
 
-{\centering \includegraphics[width=\textwidth]{fig/fcast-alldates-1}
+{\centering \includegraphics[width=\textwidth]{fig/cumulative-mean-1}
 
@@
 
-\caption{Forecast performance over all periods. Performance largely improves for all forecasters with the inclusion of data in 2021.}\label{fig:fcast-alldates}
+\caption{Cumulative sum of WIS for each forecaster divided by the cumulative sum of WIS for the baseline model.
+The shaded background shows national case incidence for the 14-day ahead target.
+Hash marks along the x-axis denote weeks.}\label{fig:cumulative-mean}
 \end{figure}
+
+\clearpage
 
@@
 
-{\centering \includegraphics[width=\textwidth]{fig/fcast-alldates-adjusted-1}
+{\centering \includegraphics[width=\textwidth]{fig/errs-in-space-1}
 
@@
 
-\caption{Forcast performance over all periods aggregaged with the geometric mean. Again, the inclusion of data in 2021 leads to improved performance.}\label{fig:fcast-alldates-adjusted}
+\caption{Percentage improvement in WIS, relative to the AR forecaster, by HRR (negative numbers indicate improved performance, positives indicate worsening).}\label{fig:errs-in-space}
 \end{figure}
-
-\clearpage
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{fig/hot-alldates-1}
-
-}
-
-\caption{Area under the curve for hotspot predictions including data in 2021.
-Performance degrades relative to the period in 2020.
-However, there are far fewer hotspots during this period as case rates declined in much of the country.}\label{fig:hot-alldates}
-\end{figure}
-
-%\bibliography{../../common/covidcast.bib,pnas-materials/pnas-sample.bib}

### 2021-11-29 cover/cover.Rmd
Make eval=TRUE
Fix latex typo

@@
 
-```{r get-data, eval = FALSE}
+```{r get-data}
 sources <- c("jhu-csse",
@@
 
-```{r load-all}
+```{r load-all, eval = FALSE}
 load(file = "all.rda")
@@
 ```
+
+```{r trend-plots-no-y-labels}
+for (i in 1:length(x)) {
+  g[[i]] <- x[[i]] %>%
+    full_join(y, by = c("geo_value", "time_value")) %>%
+    mutate(geo_num = as.integer(as.factor(geo_value))) %>%
+    group_by(geo_value) %>%
+    mutate(value = (value - Min(value)) / (Max(value) - Min(value)) +
+             delta * (geo_num - 1)) %>%
+    ungroup() %>%
+    ggplot(aes(x = time_value, y = value, group = geo_value)) +
+    labs(y = names[i]) + coord_cartesian(xlim = xlim, ylim = ylim) +
+    geom_line(alpha = 0.5, color = colors[i], size = line_width) +
+    theme_void() +
+    theme(panel.background = element_rect(fill = "black"))
+}
+
+gg <- do.call(gridExtra::arrangeGrob, list(grobs = g, ncol = 1))
+ggsave(file = "cover-no-labs.pdf", gg, height = 22.5, width = 21, units = "cm")
+ggsave(file = "cover-no-labs.png", gg, height = 15.9, width = 15, units = "cm")
+```

### 2021-11-29 perspective/paper.tex
Fix latex typo

@@
 the CDC decided to incorporate influenza forecasting into its normal operations,
-including weekly public communications citep{FluSight} and briefing to
+including weekly public communications \citep{FluSight} and briefing to
 higher-ups.  To provide more reliable infrastructure and support for its
@@
 are to consider interventions.  Unfortunately, many highly relevant aspects of
-human behavior are not measured by publicly available data streams, like
-compliance with policies and recommendations (with perhaps surveys and mobility
+human behavior, like compliance with policies and recommendations, are not
+measured by publicly available data streams (with perhaps surveys and mobility  
 reports providing our best glimpse into these hard-to-observe aspects of
-behavior \citep{Bilinski:2021}), and even the very process of promulgating these 
-policies and recommendations.  Furthermore, even if we had these data in hand,
-incorporating their effects will require significant and new cognitive and
+behavior \citep{Bilinski:2021}).  Furthermore, even if we had these data in
+hand, incorporating their effects will require significant and new cognitive and 
 behavioral modeling, with uncertain success.  The tragic breakdown and

### 2021-11-30 perspective/paper.tex
Compress refs

@@
 
-\usepackage[numbers]{natbib}
+\usepackage[numbers, sort&compress]{natbib}
 \usepackage{amssymb,amsmath,amsthm,bbm}
@@
 
-Not unrelatedly, the last decade has also seen a rise in the importance of
-\emph{digital surveillance} streams in public health, with improving epidemic
+Not unrelatedly, the last decade or so also seen a rise in the importance of 
+\emph{digital surveillance} in public health, with improving epidemic
 tracking and forecasting models being a key application of these data.  Digital
@@
 of the focus \citep{Ginsberg:2009, Brownstein:2009, Salathe:2012,
-Kass-Hout:2013, Santillana:2015, Paul:2017}; however, even more broadly, data
-from \emph{auxiliary} streams that operate outside of traditional public health
-reporting, such as online surveys, medical devices, or electronic medical
-records, have received considerable attention as well \citep{Kass-Hout:2011,
-Carlson:2013, Viboud:2014, Smolinski:2015, Santillana:2016, Charu:2017,
-Koppeschaar:2017, Yang:2019, Leuba:2020, Radin:2020, Ackley:2020}.
+Kass-Hout:2013, Santillana:2015, Paul:2017}; however, even more broadly,
+\emph{auxiliary surveillance} streams, which operate outside of traditional
+public health reporting, such as online surveys, medical devices, or electronic
+medical records, have received considerable attention as well  
+\citep{Kass-Hout:2011, Carlson:2013, Viboud:2014, Smolinski:2015,
+  Santillana:2016, Charu:2017, Koppeschaar:2017, Yang:2019, Leuba:2020,
+  Radin:2020, Ackley:2020}.
 
@@
 different perspectives.  A fourth describes international work by some of our
-collaborators that parallels our group's work on online surveys in the U.S.
+collaborators that parallels our group's work on surveys in the U.S.
 
@@
 
-\newpage
 \bibliographystyle{unsrt}

### 2021-12-01 perspective/paper.tex
Fix typo

@@
 
-In the \emph{interpandemic phase}, the main activity is threat scanning for
+In the \emph{interpandemic phase}, the main activity is scanning for
 threats, namely monitoring data streams and events throughout the world for

### 2021-12-07 perspective/paper.tex
Typo in acks

@@
 
-This material is based on work supported by gifts from Google.or and the McCune 
+This material is based on work supported by gifts from Google.org and the McCune  
 Foundation; and by Centers for Disease Control and Prevention (CDC) grant

### 2021-12-15 indicators/paper/paper.tex
Fix typo
Fix indicator paper typos

@@
 Altogether, we produce over 170 signals from 12 distinct sources, and provide
-them in a common format foc access. This unifies both unique (unavailable
+them in a common format for access. This unifies both unique (unavailable
 anywhere else) and standard COVID data streams into a single common 
@@
 results in January 2021, based on an item asking survey respondents whether they
-would accept a COVID-19 vaccine if one were vaccinated today, predicted actual
+would accept a COVID-19 vaccine if one were offered today, predicted actual
 uptake of COVID-19 vaccines by July 2021 (as reported by the Centers for Disease
@@
 Claims data tends also to be subject to biases during major national holidays
-and other events that change health-seeking behavior. Characterizing these
+and other events that change healthcare-seeking behavior. Characterizing these
 biases will be important for future research and operational systems that use

### 2021-12-15 indicators/paper/paper-medrxiv.tex
Make indicator medRxiv version match final version

@@
 
+% Print author/affiliation in smaller font, because it takes too much space
+\renewcommand\Authfont{\small}
+\renewcommand\Affilfont{\small}
+
 % Use letters for affiliations, numbers to show equal authorship (if applicable)
@@
 % Middle authors
-% \author[d]{Sumit Agrawal}
+\author[d]{Sumit Agrawal}
 \author[e]{Wael Al Saeed}
@@
 \author[b]{Brian Clark}
-% \author[d]{Sarah Colquhoun}
+\author[d]{Sarah Colquhoun}
 \author[b]{Nat DeFries}
-% \author[d]{David C. Farrow}
+\author[d]{David C. Farrow}
 \author[i]{Jodi Forlizzi}
-% \author[d]{Jed Grabman}
+\author[d]{Jed Grabman}
 \author[b]{Samuel Gratzl}
@@
 \author[i]{Robin Han}
-% \author[d]{Kate Harwood}
+\author[d]{Kate Harwood}
 \author[a,b]{Addison J. Hu}
-% \author[d]{Raphael Hyde}
+\author[d]{Raphael Hyde}
 \author[h]{Sangwon Hyun}
@@
 \author[n]{Daniel J. McDonald}
-% \author[d]{Phillip McGuinness}
+\author[d]{Phillip McGuinness}
 \author[o,p]{Balasubramanian Narasimhan}
-% \author[d]{Michael P. O'Brien}
+\author[d]{Michael P. O'Brien}
 \author[a,b]{Natalia L. Oliveira}
@@
 \author[e]{Dawn Rucker}
-% \author[d]{Chris Scott}
+\author[d]{Chris Scott}
 \author[q]{Nigam H. Shah}
@@
 \author[s]{Noah Simon}
-% \author[d]{Benjamin Y. Smith}
+\author[d]{Benjamin Y. Smith}
 \author[b]{Vishakha Srivastava}
@@
 \author[a,b]{Larry Wasserman}
-% \author[d]{Benjamin Weaver}
+\author[d]{Benjamin Weaver}
 \author[t]{Jeremy C. Weiss}
-% \author[d]{Spencer Whitman}
+\author[d]{Spencer Whitman}
 \author[i]{Kristin Williams}
@@
 \affil[c]{Computational Biology Department, Carnegie Mellon University}
-% \affil[d]{Google.org}
+\affil[d]{Google.org Fellows, Google LLC}
 \affil[e]{Computer Science Department, Carnegie Mellon University}
@@
   Mellon University}
+% Middle authors
 
@@
   areinhar@stat.cmu.edu}
-\footnotetext[2]{L.B., M.J., A.R., and J.T. contributed equally to this work.}
+\footnotetext[2]{L.B., M.J., A. Rumack, and J.T. contributed equally to this work.}
 
@@
   Brooks:2020}. Several other examples exist that span a wide variety of data
-platforms and diseases \cite{Browstein:2009, Kass-Hout:2011, Salathe:2012,
+platforms and diseases \cite{Brownstein:2009, Kass-Hout:2011, Salathe:2012,
   Kass-Hout:2013}. During the COVID-19 pandemic, digital data streams have
@@
 
-The Delphi Group works with partner organizations and public data sets to build
-a massive database of indicators tracking COVID-19 activity and other relevant
-phenomena in the United States, which has been publicly available and
+The Delphi group worked with partner organizations and public data sets to build
+a large-scale database of indicators tracking COVID-19 activity and other
+relevant phenomena in the United States, which has been publicly available and
 continuously updated since April 2020. Alongside public data on reported cases
@@
 previous (historical) versions of the signals, a key feature that exposes the
-effects of data revisions.  Lastly, we provide R \cite{CovidcastR} and Python
+effects of data revisions. Moreover, we provide R \cite{CovidcastR} and Python
 \cite{CovidcastPy} packages to facilitate interaction with the API, and an
@@
 In a companion paper, we analyze the utility provided by a core set of the
-indicators in COVID-19 forecasting and hotspot prediction models.  In another
-companion paper, we elaborate on our research group's (Delphi's) large-scale
-public surveys, run in partnership with Facebook and available in aggregate form
-in the COVIDcast API.
+indicators in short-term COVID-19 forecasting and hotspot prediction models
+\cite{McDonald:2021}. In another companion paper, we elaborate on our research
+group's (Delphi's) large-scale public surveys, run in partnership with Facebook
+and available in aggregate form in the COVIDcast API \cite{Salomon:2021}. This
+paper focuses on the COVID-19 indicators themselves, describing the
+data streams, how they are processed and made publicly available, and insights
+that can be gained by combining novel data sources with standard public health
+surveillance data.
 
@@
 \paragraph{Internet-Based Surveys} Conducted in partnership with Facebook,
-Delphi's COVID-19 Trends and Impact Survey receives an average of 50,000 responses
-daily, and has received over 20 million responses since April 2020
-\cite{DelphiSurvey, Kreuter:2020}. From the surveys, we construct indicators on
+Delphi's COVID-19 Trends and Impact Survey receives an average of 50,000
+responses daily, and has received over 25 million responses since April 2020
+\cite{Salomon:2021, Kreuter:2020}. From the surveys, we construct indicators on
 symptoms, social distancing, vaccination, and other attitudes and behaviors
-related to COVID.
+related to COVID. The surveys are voluntary and participants are redirected to a
+platform managed by Carnegie Mellon University to give consent and take the
+survey; individual response data is not provided to Facebook and the data
+collection protocol was approved by the Carnegie Mellon Institutional Review
+Board (STUDY2020\_00000162).
 
@@
 
-Altogether, we collect 110 signals from 12 distinct sources, aggregate and
-process them, and provide them in a common format for public access. This
-unifies both unique (not available anywhere else) and standard COVID data
-streams into a single common format, enabling efficient comparison and modeling.
-A summary of the data sources and signals in the API is in
-Table~\ref{tab:sources_signals}, and detailed documentation is available at
+Altogether, we produce over 170 signals from 12 distinct sources, and provide
+them in a common format for access. This unifies both unique (unavailable
+anywhere else) and standard COVID data streams into a single common
+format, enabling efficient comparison and modeling. A summary of the data
+sources and signals in the API is in Table~\ref{tab:sources_signals}, and
+detailed documentation is available online at
 \url{https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html}.
@@
 \centering
-\caption{Data sources available in Delphi's COVIDcast API \cite{CovidcastAPI},
-  as of date of publication. The first group of data sources are produced by
-  Delphi from data not otherwise available publicly (or only available in
-  limited form); the second group is mirrored from public sources.}
-\begin{tabular}{>{\raggedright}p{1.2in} p{5.0in} l >{\raggedright\arraybackslash}p{0.5in}}
+\caption{Data sources available in the COVIDcast API \cite{CovidcastAPI}, as of
+  date of publication. The first group of data sources are produced from data
+  not otherwise available publicly (or only available in limited form); the
+  second group is mirrored from public sources.}
+\small
+\begin{tabular}{>{\raggedright}p{1.2in} p{4.0in} l >{\raggedright\arraybackslash}p{0.5in}}
   \toprule
@@
 the \textit{time value}, the date the underlying events (such as tests or
-doctor's visits) occurred, and the \textit{issue date} when Delphi aggregated
-and reported the data for that time value. Importantly, there can be multiple
+doctor's visits) occurred, and the \textit{issue date} when we aggregated and
+reported the data for that time value. Importantly, there can be multiple
 observations for a single time value with different issue dates, for example if
-data is revised or claims records arrive late. Delphi tracks revisions to all
-data sources we ingest, including external data sources (such as sources
+data is revised or claims records arrive late. We track revisions to all data
+sources included in the API, including external data sources (such as sources
 tracking cases and deaths). Many external sources do not keep a public or
@@
 improved as new sources of data arrive, and in response to interviews with users
-and health experts, usage analytics from the site, and user surveys. Key design
-goals for visualizations displaying COVIDcast data have included representing
-uncertainty and missing data, making features consistent and discoverable
-despite a wide variety of signal types, and smoothly switching between different
-geographic aggregations and time scales.
+and health experts, usage analytics from the site, and user surveys.% Key design
+% goals for visualizations displaying COVIDcast data have included representing
+% uncertainty and missing data, making features consistent and discoverable
+% despite a wide variety of signal types, and smoothly switching between
+% different geographic aggregations and time scales.
 
@@
 
-Since July 2020, Delphi has been regularly submitting short-term forecasts of
-COVID-19 case and death incidence, at the state and county levels, to the
+Since July 2020, we have been regularly submitting short-term forecasts of
+COVID-19 case and death incidence, at both the state and county levels, to the
 COVID-19 Forecast Hub \cite{ForecastHub}, with ``CMU-TimeSeries'' as the
@@
 R package and fetches data from the API) can be found at
-\url{https://github.com/cmu-delphi/covidcast-pnas/tree/main/indicators/code/}.
+\url{https://doi.org/10.5281/zenodo.5639567}.
 
@@
 Besides tracking contemporaneous COVID activity, these and other indicators can
-be used to improve forecasts of future COVID case trends, as investigated in a
-companion paper.
+be used to improve forecasts of future COVID case trends, as investigated in our
+companion paper \cite{McDonald:2021}.
 
@@
 performed over all counties with at least 500 cumulative cases by the end of
-this period, and at which all indicators are available (956 counties in
-total). The large positive correlations suggest that these signals could be
-useful in hotspot detection (identifying counties that have relatively high
-COVID activity, at a given time). Somewhat surprisingly, the survey-based
+this period, and at which all indicators are available (956 counties in total).
+The large positive correlations suggest that these signals could be useful in
+hotspot detection (identifying counties that have relatively high COVID
+activity, at a given time). Somewhat surprisingly, the survey-based
 CLI-in-community signal shows the strongest correlations for much of the time
 period. This clearly demonstrates the value of a large-scale survey such as CTIS
-for tracking symptoms and case trends, especially when other data is unavailable.
+for tracking symptoms and case trends, especially when other data is
+unavailable.
+
+Also notable is the fact that the correlations fluctuate over time in complex
+ways; while some of this variation is likely due to changes in public behavior,
+reporting and testing practices, and so on, some is also related to changes in
+overall COVID-19 case trends. For example, comparing against
+Figure~\ref{fig:time_trends_national}, we see that correlations drop in February
+2021 for many of the signals, roughly matching the point in time when COVID
+cases also sharply decline. This decline likely caused the heterogeneity in case
+rates by county to decline, making it more difficult for any signal to achieve a
+high geo-wise correlation. Correlations between confirmed cases and cases 1--3
+weeks prior (shown in the Supplementary Information) show a similar correlation
+drop in February 2021, showing that the drop is due to a change in case data and
+not problems with the other signals.
 
@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.6\columnwidth]{fig/time_wise_correlation.pdf}
+  \caption{Time-wise correlations with case rates, from April 15, 2020 to April
+    15, 2021, calculated over all counties for which all signals were available
+    and which had at least 500 cumulative cases by the end of this period.}
+  \label{fig:time_wise_correlation}
+\end{figure}
+
 \begin{itemize}
@@
 
-As an example, looking back at Figure \ref{fig:time_trends_national}, we can see
+As an example, looking back at Figure~\ref{fig:time_trends_national}, we can see
 clear dips in the confirmed COVID case curve that occur around the Thanksgiving
@@
 death reporting (for this reason, the artificial dips persist at the state- and
-county-level as well). The CLI signal from the survey, on the other hand,
-displays no such dips. The claims signals actually display holiday effects going
-in the other direction: they exhibit \textit{spikes} around Thanksgiving and New
-Year's. This is because they measure the fraction of all outpatient visits with
-a certain condition, and the denominator here (total outpatient visits) drops
-disproportionately during holiday periods, as people are likely less willing to
-go to the doctor for more routine issues. Fortunately, in principle, the holiday
-effects in claims signals should be correctable: they are mainly due to
-\textit{overall} changes in medical seeking behavior during holiday, periods,
-and we can estimate such effects using historical claims data.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=0.6\columnwidth]{fig/time_wise_correlation.pdf}
-  \caption{Time-wise correlations with case rates, from April 15, 2020 to April
-    15, 2021, calculated over all counties for which all signals were available
-    and which had at least 500 cumulative cases by the end of this period.}
-  \label{fig:time_wise_correlation}
-\end{figure}
+county-level as well). This delay denies public health officials timely signals
+of current trends for the duration of the holidays. The CLI signal from the
+survey, on the other hand, displays no such dips. The claims signals actually
+display holiday effects going in the other direction: they exhibit
+\textit{spikes} around Thanksgiving and New Year's. This is because they measure
+the fraction of all outpatient visits with a certain condition, and the
+denominator (total outpatient visits) drops disproportionately during holiday
+periods, as people are likely less willing to go to the doctor for more routine
+issues. Fortunately, in principle, the holiday effects in claims signals should
+be correctable: they are mainly due to \textit{overall} changes in medical
+seeking behavior during holiday, periods, and we can estimate such effects using
+historical claims data.
 
@@
 population who know someone in their local community with COVID-Like Illness
-(CLI). As we can see in the right-hand panel of the figure, this signal was not
-affected by Bexar County's reporting problems and, as shown in the last
+(CLI). As we can see in the right-hand panel of the figure, this indicator was
+not affected by Bexar County's reporting problems and, as shown in the last
 subsection, it is (in general) highly correlated with case rates, providing an
-alternate stream of data about COVID activity unaffected by backlogs. Similar
-reporting problems have occurred in many jurisdictions across the United States,
-making it valuable to cross-check against external sources not part of the same
-reporting systems.
+alternate stream of data about COVID activity unaffected by backlogs. In
+general, reporting problems have occurred in many jurisdictions across the
+United States, and audits have regularly discovered misclassified or unreported
+cases and deaths, making it valuable to cross-check against external sources not
+part of the same reporting systems.
 
@@
   \centering
-  \includegraphics[width=0.6\columnwidth]{fig/bexar_compare.pdf}
+  \includegraphics[width=0.8\columnwidth]{fig/bexar_compare.pdf}
   \caption{Reported cases per day in Bexar County, Texas during the summer of
@@
   \centering
-  \includegraphics[width=0.6\columnwidth]{fig/dv_as_of.pdf}
-  \caption{Estimated percentage of doctor's visits due to COVID-like illness
-    displayed across multiple issue dates, with later issue dates adding
-    additional data and revising past data from prior issue dates.}
+  \includegraphics[width=0.8\columnwidth]{fig/dv_as_of.pdf}
+  \caption{Estimated percentage of outpatient visits due to COVID-like illness
+    (DV-CLI) displayed across multiple issue dates, with later issue dates
+    adding additional data and revising past data from prior issue dates.}
   \label{fig:dv_as_of}
@@
 suggests that statistical models could be fit (potentially separately for each
-location) to estimate the final values from preliminary reports. On the other
-hand, official public health reporting of COVID cases and deaths can be subject
-to revision as death certificates are audited and backlogs cleared, resulting in
-thousands of cases and deaths being added or removed. This process is much more
-difficult to predict, and thus claims data and other sources may be a useful
-stand-in while public health reports are aggregated and corrected.
+location) to estimate the final values from preliminary reports.
 
-To reiterate a previous point, when training and validating forecast models (on
-historical data), users will want to use data that was known \textit{as of} the
-forecast date, not revised versions that only became available much later. The
-COVIDcast API makes all historical versions available and easily accessible for
-this purpose; and this feature plays a prominent role in our own analysis of
-forecasting and hotspot prediction models appearing in a companion paper.
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.6\columnwidth]{fig/backfill-compare.pdf}
+  \caption{The 95th percentiles of relative error of early reported values of
+    key signals compared to final values reported much later. For each date
+    between October 15, 2020 and April 15, 2021, the values for each state
+    reported between 10 and 90 days later are compared to ``final'' versions
+    recorded as of August 13, 2021. Even officially reported case and death data
+    can have large revisions 30--60 days or more after initial reporting; much
+    of this is driven by individual large revisions affecting specific states
+    and dates, rather than systematic changes affecting all states and dates.}
+  \label{fig:backfill-compare}
+\end{figure}
+
+Figure~\ref{fig:backfill-compare} shows relative differences between early
+indicator values, reported 10--90 days after the underlying events, and later
+versions at least four months later. As the distribution of these revision
+amounts is highly skewed, the figure plots the 95th percentile of relative
+change, showing that reported deaths can incur large relative error in initial
+values, comparable to that in claims-based signals. However, for deaths, as well
+as cases, these large revisions are not very systematic, with large corrections
+typically occurring at a sparse subset of locations and times (e.g., due to
+audits or backlogs being cleared, which can result in thousands of cases or
+deaths can be added or removed all at once). This backfill is much more
+difficult to predict than that of claims data, therefore the latter (and other
+sources) may be useful for nowcasting cases and deaths while public health
+reports are being aggregated and corrected.
+
+To reiterate a previous point, when building forecast models (on historical
+data) for retrospective evaluation purposes, users will want to use data that
+was known \textit{as of} the forecast date, not revised versions that only
+became available at a later time. Note that not only model training, but also
+model assessment, can be affected by the revision process (comparisons of
+forecasts to the ground truth may shift when the ground truth is revised weeks
+later). Only by systematically tracking revisions can all these effects be
+monitored and properly accounted for.  The COVIDcast API makes all historical
+versions available and easily accessible for this purpose; and this feature
+plays a prominent role in our own analysis of forecasting and hotspot prediction
+models appearing in a companion paper \cite{McDonald:2021}.
 
@@
   \centering
-  \includegraphics[width=0.6\columnwidth]{fig/mobility-drop.pdf}
-  \caption{Percentage point increase in SafeGraph estimates of the percentage of
-  people staying completely at home, based on aggregated mobile device data,
-  between March 1, 2020 and April 15, 2020.}
-  \label{fig:mobility-drop}
+  \includegraphics[width=0.6\columnwidth]{fig/vaccine-intent-compare.pdf}
+  \caption{CTIS estimates of the percentage of people willing to get
+    vaccinated, back on January 20, 2021, compared to CDC reporting of the
+    percentage of people vaccinated, on July 20, 2021. Each point is a county
+    (with at least 250 survey responses between January 14--20, 2021), colored
+    by its parent United States Census region.}
+  \label{fig:vaccine-intent-compare}
 \end{figure}
 
-As an illustration, Figure~\ref{fig:mobility-drop} maps the sharp increase in
-rates of people staying at home between March 1 and April 15, 2020. Using
-SafeGraph data on the fraction of mobile devices included in SafeGraph's panel
-that did not leave the immediate area of their home, it illustrates the sharp
-drop in travel and work outside the home that occurred in the early stages of
-the pandemic. It also shows that this drop was much more pronounced in some
-states than others, enabling analysis of policy impacts and disease spread.
-Similar maps can be quickly constructed for any signal in the COVIDcast API.
+As an illustration, Figure~\ref{fig:vaccine-intent-compare} illustrates how CTIS
+results in January 2021, based on an item asking survey respondents whether they
+would accept a COVID-19 vaccine if one were offered today, predicted actual
+uptake of COVID-19 vaccines by July 2021 (as reported by the Centers for Disease
+Control and Prevention). It also reveals a geographical disparity: in the
+Northeast, actual vaccination rates more closely match vaccine willingness rates
+in January than they do in the South, where vaccination rates lag overall. While
+these results must be interpreted carefully due to potential sampling biases in
+the survey, they illustrate the potential of data in the COVIDcast API to inform
+public health decision-making.
 
@@
 
-\input{supplement-medrxiv.tex}
+\include{supplement-medrxiv.tex}
 

### 2021-12-15 indicators/paper/supplement-medrxiv.tex
Make indicator medRxiv version match final version

@@
 \includegraphics[width=\textwidth]{fig/severity-pyramid.pdf}
-\caption{The epidemiological ``severity pyramid'' represents the progression of
-cases from the public, through infection, through increasingly severe stages of
-disease. The annotations here represent the data sources collected by Delphi's
-COVIDcast Epidata API.}
+\caption{Epidemiological ``severity pyramid'', representing the progression of disease progression, from relevant public behaviors, through infection, towards increasingly severe stages of disease.
+The annotations here refer to the data sources available in Delphi's Epidata API.}
+\label{fig:severity-pyramid}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{fig/case-correlation-lagged-plot-1}
+
+}
+
+\caption{Geo-wise correlations between cases and lagged cases 1, 2, or 3 weeks prior, for all counties in the U.S.
+Lagged cases are correlated with cases as one might expect, but note the precipitous drop in correlation in February 2021.
+This matches the correlation drop between other COVIDcast signals and cases during the same time period, supporting the hypothesis that the drop was due to decreased heterogeneity in case rates by county.}\label{fig:case-correlation-lagged-plot}
 \end{figure}
@@
 
-\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States.
-Cases are reported cases per 100,000 population.
-Other signals are scaled to have the same global maximum across all counties and times, so they can be presented in the same range.
-(part 1 of 4)}\label{fig:state-trend-grids-1}
+\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for U.S.
+states and territories.
+Cases are displayed on the rate scale: counts per 100,000 people.
+Other signals are scaled to have the same global range across all counties and times.
+(Part 1 of 4.)}\label{fig:state-trend-grids-1}
 \end{figure}
@@
 
-\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States.
-Cases are reported cases per 100,000 population.
-Other signals are scaled to have the same global maximum across all counties and times, so they can be presented in the same range.
-(part 1 of 4)}\label{fig:county-trend-grids-1}
+\caption{Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous U.S.
+counties.
+Cases are displayed on the rate scale: counts per 100,000 people.
+Other signals are scaled to have the same global range across all counties and times.
+(Part 1 of 4.)}\label{fig:county-trend-grids-1}
 \end{figure}
@@
 
-\caption{National trends, from April 2020 to April 2021, of HHS-reported confirmed COVID-19 hospital admissions, plus another hospitalization signal in the COVIDcast API.
-Here, the HHS data has been limited to values from 2020-08-01 onward due to changes in reporting behavior, and processed into a 7-day trailing average rate per 100,000 resident population using Census Bureau estimates for 2019.
-HSP-Hosp is the percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners, smoothed in time and adjusted for systematic day-of-week effects (HSP-Hosp is available from an earlier date when working at finer geographical resolutions).
-HSP-Hosp has been scaled to have the same maximum value as the HHS data.}\label{fig:hospitalization_time_trends_national}
+\caption{National trends, from August 2020 to August 2021, of HHS-reported confirmed COVID-19 hospital admissions, along with several signals from the COVIDcast API.
+(HHS data was not consistently reported before August 2020; furthermore, as with cases in the previous trend plots, the HHS data has been smoothed using a 7-day trailing average.) Hospitalizations are displayed on the rate scale: counts per 100,000 people.
+Other signals are scaled to have the same range.
+HSP-Hosp is the percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners (smoothed in time and adjusted for systematic day-of-week effects).}\label{fig:hospitalization_time_trends_national}
 \end{figure}
@@
 
-\caption{Time-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which both signals are available.}\label{fig:hosp-correlations-by-state}
+\caption{Time-wise correlations with hospitalization rates derived from HHS data, from August 15, 2020 to August 15, 2021, calculated over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which all signals are available.}\label{fig:hosp-correlations-by-state}
 \end{figure}
